### 2020-08-31 HW1/HW1.Rmd
homework 1 template

---
title: 'DATA 621 – Business Analytics and Data Mining: Homework #1'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, Bryan Persaud,
  Jeff Shamp
date: "9/27/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. DATA EXPLORATION (25 Points)
Describe the size and the variables in the moneyball training data set. Consider that too much detail will cause a manager to lose interest while too little detail will make the manager consider that you aren’t doing your job. Some suggestions are given below. Please do NOT treat this as a check list of things to do to complete the assignment. You should have your own thoughts on what to tell the boss. These are just ideas.

### a. Mean / Standard Deviation / Median
### b. Bar Chart or Box Plot of the data
### c. Is the data correlated to the target variable (or to other variables?)
### d. Are any of the variables missing and need to be imputed “fixed”?

## 2. DATA PREPARATION (25 Points)
Describe how you have transformed the data by changing the original variables or creating new variables. If you did transform the data or create new variables, discuss why you did this. Here are some possible transformations.
### a. Fix missing values (maybe with a Mean or Median value)
### b. Create flags to suggest if a variable was missing
### c. Transform data by putting it into buckets
### d. Mathematical transforms such as log or square root (or use Box-Cox)
### e. Combine variables (such as ratios or adding or multiplying) to create new variables

## 3. BUILD MODELS (25 Points)
Using the training data set, build at least three different multiple linear regression models, using different variables (or the same variables with different transformations). Since we have not yet covered automated variable selection methods, you should select the variables manually (unless you previously learned Forward or Stepwise selection, etc.). Since you manually selected a variable for inclusion into the model or exclusion into the model, indicate why this was done.
Discuss the coefficients in the models, do they make sense? For example, if a team hits a lot of Home Runs, it would be reasonably expected that such a team would win more games. However, if the coefficient is negative (suggesting that the team would lose more games), then that needs to be discussed. Are you keeping the model even though it is counter intuitive? Why? The boss needs to know.

## 4. SELECT MODELS (25 Points)
Decide on the criteria for selecting the best multiple linear regression model. Will you select a model with slightly worse performance if it makes more sense or is more parsimonious? Discuss why you selected your model.
For the multiple linear regression model, will you use a metric such as Adjusted R2, RMSE, etc.? Be sure to explain how you can make inferences from the model, discuss multi-collinearity issues (if any), and discuss other relevant model output. Using the training data set, evaluate the multiple linear regression model based on (a) mean squared error, (b) R2, (c) F-statistic, and (d) residual plots. Make predictions using the evaluation data set.

## APPENDIX
### Assignment Overview
In this homework assignment, you will explore, analyze and model a data set containing approximately 2200
records. Each record represents a professional baseball team from the years 1871 to 2006 inclusive. Each record
has the performance of the team for the given year, with all of the statistics adjusted to match the performance of
a 162 game season.

Your objective is to build a multiple linear regression model on the training data to predict the number of wins
for the team. You can only use the variables given to you (or variables that you derive from the variables
provided). Below is a short description of the variables of interest in the data set:

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{VARIABLE NAME} & \textbf{DEFINITION} & \textbf{THEORETICAL EFFECT} \\ \hline
INDEX & Identification Variable (do not use) & None \\ \hline
TARGET\_WINS & Number of wins & \\ \hline
TEAM\_BATTING\_H & Base Hits by batters (1B,2B,3B,HR) & Positive Impact on Wins  \\ \hline
TEAM\_BATTING\_2B & Doubles by batters (2B) & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_3B & Triples by batters (3B) & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_HR & Homeruns by batters (4B) & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_BB & Walks by batters & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_HBP & Batters hit by pitch (get a free base) & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_SO & Strikeouts by batters & Negative Impact on Wins \\ \hline
TEAM\_BASERUN\_SB & Stolen bases & Positive Impact on Wins \\ \hline
TEAM\_BASERUN\_CS & Caught stealing & Negative Impact on Wins \\ \hline
TEAM\_FIELDING\_E & Errors & Negative Impact on Wins \\ \hline
TEAM\_FIELDING\_DP & Double Plays & Positive Impact on Wins \\ \hline
TEAM\_PITCHING\_BB & Walks allowed & Negative Impact on Wins \\ \hline
TEAM\_PITCHING\_H & Hits allowed & Negative Impact on Wins \\ \hline
TEAM\_PITCHING\_HR & Homeruns allowed & Negative Impact on Wins \\ \hline
TEAM\_PITCHING\_SO & Strikeouts by pitchers & Positive Impact on Wins \\ \hline

\end{tabular}
\end{center}

### R Programming code

```{r}

train_set <- read.csv('data/moneyball-training-data.csv')
eval_set <- read.csv('data/moneyball-evaluation-data.csv')

colnames(train_set)

```

### 2020-08-31 HW1/HW1.Rmd
homework 1 desciption updates

@@
 
-## 1. DATA EXPLORATION (25 Points)
+## 1. DATA EXPLORATION (Amber)
 Describe the size and the variables in the moneyball training data set.
@@
 
-### a. Mean / Standard Deviation / Median
-### b. Bar Chart or Box Plot of the data
-### c. Is the data correlated to the target variable (or to other variables?)
-### d. Are any of the variables missing and need to be imputed “fixed”?
+a. Mean / Standard Deviation / Median
+b. Bar Chart or Box Plot of the data
+c. Is the data correlated to the target variable (or to other variables?)
+d. Are any of the variables missing and need to be imputed “fixed”?
 
@@
 Here are some possible transformations.
-### a. Fix missing values (maybe with a Mean or Median value)
-### b. Create flags to suggest if a variable was missing
-### c. Transform data by putting it into buckets
-### d. Mathematical transforms such as log or square root (or use Box-Cox)
-### e. Combine variables (such as ratios or adding or multiplying) to create new variables
+
+a. Fix missing values (maybe with a Mean or Median value)
+b. Create flags to suggest if a variable was missing
+c. Transform data by putting it into buckets
+d. Mathematical transforms such as log or square root (or use Box-Cox)
+e. Combine variables (such as ratios or adding or multiplying) to create new variables
 
@@
 Since you manually selected a variable for inclusion into the model or exclusion into the model, indicate why this was done.
+
 Discuss the coefficients in the models, do they make sense?
@@
 Discuss why you selected your model.
+
 For the multiple linear regression model, will you use a metric such as Adjusted R2, RMSE, etc.?

### 2020-09-02 HW1/missing_value_eda.Rmd
Initial EDA on missing values, imputation

---
title: "621 - HW1 -EDA"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Moneyball EDA

Initial exploration into the moneyball data sets.

```{r}
sessionInfo()
```

```{r}
library(tidyverse)
train_df<- 
  read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv?token=AJP53VW2N3ER7LBQPZRTCN27J6ZTQ")
```

```{r}
glimpse(train_df)
```

```{r}
train_df %>%
  filter(complete.cases(train_df) == TRUE) %>%
  count() %>%
  rename("complete cases"= n) %>%
  bind_cols(
    train_df %>%
    select(TARGET_WINS) %>%
    count() %>%
    rename("target count" = n)
  )
```

Lots of not complete cases. Let's look at percent `NA`.

```{r}
glimpse(
  train_df %>%
  summarize_all(funs(sum(is.na(.))/length(.)))
) 
```

This is good news, it seems. The most missing values come from what we would expect; hit-by-pitch, caught stealing, double plays. We will need to see about columns with zeros and `NA` values.

```{r}
missing_df<-
  train_df %>%
  select(TEAM_BATTING_SO, TEAM_BASERUN_SB, TEAM_BASERUN_CS,
         TEAM_BATTING_HBP, TEAM_PITCHING_SO, TEAM_FIELDING_DP)

# Are there zeros and NAs in the same columns with missing values
missing_df %>%
  summarize_all(funs(length(which(. ==0))))
```

For the three most missing columns (HBP, CS, DP) we have neither one or zero, zeros. Seems reasonable that for these columns `NA` values can be zeros as these are rare events and they don't seem to be tallied as a zero. Alternatively, these stats just might not have been counted back in the day so those team season stats are entered as `NA`.

For `TEAM_BATTING_SO` and `..PITCHING_SO` there are 20 zero values and 99 `NA`. **It seems completely unreasonable** to have zero strike outs in a season or to **not** record them.

### Impute Strategy

Given some issues outlined above, it will likely be helpful to impute some of the values. For that, distribution of the known values are displayed below. Red vertical line is the mean value (normality assumed, some aren't).

There seem to be good candidates for imputation by mean/median given the variables distribution (normal, binomial, geometric, etc).

```{r}
missing_df %>%
  ggplot(aes(TEAM_BATTING_HBP)) +
  geom_histogram(na.rm = TRUE, binwidth = 1) +
   geom_vline(xintercept = mean(
              missing_df$TEAM_BATTING_HBP,
              na.rm = TRUE), color="red")

missing_df %>%
  ggplot(aes(TEAM_BASERUN_CS)) +
  geom_histogram(na.rm = TRUE, binwidth = 1) +
  geom_vline(xintercept = mean(
              missing_df$TEAM_BASERUN_CS,
              na.rm = TRUE), color="red")

missing_df %>%
  ggplot(aes(TEAM_FIELDING_DP)) +
  geom_histogram(na.rm=TRUE, binwidth = 1) +
  geom_vline(xintercept = mean(
              missing_df$TEAM_FIELDING_DP,
              na.rm = TRUE), color="red")

missing_df %>%
  ggplot(aes(TEAM_BATTING_SO)) +
  geom_histogram(na.rm=TRUE, binwidth = 1) +
  geom_vline(xintercept = mean(
              missing_df$TEAM_BATTING_SO,
              na.rm = TRUE), color="red")

missing_df %>%
  ggplot(aes(TEAM_BASERUN_SB)) +
  geom_histogram(na.rm = TRUE, binwidth = 1) +
   geom_vline(xintercept = mean(
              missing_df$TEAM_BASERUN_SB,
              na.rm = TRUE), color="red")

missing_df %>%
  ggplot(aes(TEAM_PITCHING_SO)) +
  geom_histogram(na.rm = TRUE, binwidth = 1) +
   geom_vline(xintercept = mean(
              missing_df$TEAM_PITCHING_SO,
              na.rm = TRUE), color="red")
```

May also be easy to regress missing values by association.

```{r}
missing_df %>%
  na.omit() %>%
  ggplot(aes(TEAM_BASERUN_SB, TEAM_BASERUN_CS)) +
  geom_point()
```

## Outliers

`TEAM_PITCHING_SO` seemed to have a compressed distribution. Checking that out.

```{r}
missing_df %>%
  select(TEAM_PITCHING_SO) %>%
  summary()

missing_df %>%
  ggplot(aes(TEAM_PITCHING_SO)) +
  geom_boxplot(na.rm = TRUE)
```

No way. The MLB team pitching strikeout record is 1450 and set in 2014 (CLE). Also 20k SO in a season is 123 SO per game for all 162 games. There are a minimum of 27 at-bats per game and most teams average ~35 at-bats.

```{r}
missing_df %>%
  select(TEAM_PITCHING_SO) %>%
  na.omit() %>%
  summarize(iqr = IQR(TEAM_PITCHING_SO)) %>%
  bind_cols(
            missing_df %>% 
            select(TEAM_PITCHING_SO)
            )  %>%
  filter(
    TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO,
                                probs=c(0.25), 
                                na.rm = TRUE) - 1.5*iqr, 
    TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, 
                                probs=c(0.75), 
                                na.rm=TRUE) + 1.5*iqr
        ) %>%
  ggplot(aes(TEAM_PITCHING_SO)) +
  geom_boxplot(fill="lightblue3")

```

That's better. The record values are still outliers on the plot, but not totally out of reason.

### 2020-09-03 HW1/missing_value_eda.Rmd
updates with some analysis.

@@
 train_df<- 
-  read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv?token=AJP53VW2N3ER7LBQPZRTCN27J6ZTQ")
+  read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv?token=AJP53VTYKQ4BKNZQERSMBYK7KBD52")
 ```
@@
 ```
+
+### Missing Values
 
@@
 
-For the three most missing columns (HBP, CS, DP) we have neither one or zero, zeros.
-Seems reasonable that for these columns `NA` values can be zeros as these are rare events and they don't seem to be tallied as a zero.
-Alternatively, these stats just might not have been counted back in the day so those team season stats are entered as `NA`.
+For the three most missing columns (HBP, CS, DP) we have neither one or zero, zeros. These stats just might not have been counted back-in-the-day so those team season stats are entered as `NA`.
 
-For `TEAM_BATTING_SO` and `..PITCHING_SO` there are 20 zero values and 99 `NA`. **It seems completely unreasonable** to have zero strike outs in a season or to **not** record them.
+For `TEAM_BATTING_SO` and `..PITCHING_SO` there are 20 zero values and 99 `NA`. **It seems completely unreasonable** to have zero strike outs in a season. It seems wild to **not** record them.
 
@@
 There seem to be good candidates for imputation by mean/median given the variables distribution (normal, binomial, geometric, etc).
+For all of the variables with missing values a simple imputation strategy will cause a large "pile up" of values where imputation was done.
+For example, `TEAM_BASERUN_SB` is only 5.75% missing with 131 missing observations, but the largest count total for any given value is 20.
+So if we impute via median value, we will have an order magnitude more values where we filled compared to the values present in the data.
 
@@
 
-May also be easy to regress missing values by association.
+May also be easy (or better) to regress missing values by association.
 
@@
 Also 20k SO in a season is 123 SO per game for all 162 games.
-There are a minimum of 27 at-bats per game and most teams average ~35 at-bats.
+There are a minimum of 27 at-bats per game and most teams average ~35 at-bats, making 123 SOs impossible.
 
 ```{r}
+so_pitch_outliers<-
   missing_df %>%
@@
   filter(
+    # Used median and IQR for outlier removal. 
+    # mean and std could also work.
     TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO,
@@
                                 na.rm=TRUE) + 1.5*iqr
-        ) %>%
+        )
+
+so_pitch_outliers %>%
   ggplot(aes(TEAM_PITCHING_SO)) +
@@
 
+so_pitch_outliers %>%
+  ggplot(aes(TEAM_PITCHING_SO)) +
+  geom_histogram(binwidth=5)
 ```

### 2020-09-06 HW1/Regression Model - Deokinanan.rmd
Added regression model #1

---
title: "DATA 621 - HW1 - Regression Model"
author: "Samantha Deokinanan"
date: "September 3, 2020"
output: html_document
---

***
#### Task
***

*Using the training data set, build at least three different multiple linear regression models, using different variables(or the same variables with different transformations). Since we have not yet covered automated variable selection methods, you should select the variables manually (unless you previously learned Forward or Stepwise selection, etc.). Since you manually selected a variable for inclusion into the model or exclusion into the model, indicate why this was done.*

*Discuss the coefficients in the models, do they make sense? For example, if a team hits a lot of Home Runs, it would be reasonably expected that such a team would win more games. However, if the coefficient is negative (suggesting that the team would lose more games), then that needs to be discussed. Are you keeping the model even though it is counter intuitive? Why? The boss needs to know.*

***
#### The dataset
***

The data set contains approximately 2200 records. Each record represents a professional baseball team from the years 1871 to 2006 inclusive. Each record has the performance of the team for the given year, with all of the statistics adjusted to match the performance of a 162 game season. Aim is to predict the number of wins for the team.

```{r message=FALSE, warning=FALSE}
# The required R packages
library(tidyverse)
library(psych)
library(olsrr)
library(rcompanion)
library(mice)
library(VIM)
library(mctest)
library(caret)

# Load training data set
train_df = read.csv("C:/Users/Deokinanan/Desktop/2019 -2021 CUNY SPS/FALL 2020/DATA 621/WEEK 3/moneyball-training-data.csv")
train_df = subset(train_df, select = -c(INDEX))
```

***
#### Data Transformation, Outliers and Missing Data
***

A quick exploration of the data, and it is already apparent that the `TEAM_BATTING_HBP` variable has nearly 92% of the its data missing. Since it would be very difficult to accurately impute such a large proportion, this variable will be excluded from further analysis.

```{r}
# Quick Exploration
sapply(train_df, describe)

# Data Transformation
train_df = subset(train_df, select = -c(TEAM_BATTING_HBP))
```

Next, the outlier plots revealed that there are a few extreme values that can influence the analysis. Because the objective is so create a multivariate regression model, declaring an observation as an outlier based on a just one feature could lead to unrealistic inferences. There Cook's distance is use to decide if an individual entity is an extreme value or not.

Cook’s distance is a measure computed with respect to a given regression model and therefore is impacted only by the predictors included in the model. It computes the influence exerted by each data point on the response variable. Fitting the full model during this exploratory stage, the process remove the most influential points. Note from the plots that all influential observations are not necessarily outliers. Least than 3% of the data was removed.

```{r}
# Outlier Plots
model = lm(TARGET_WINS ~ ., data = train_df)
ols_plot_diagnostics(model)

# Remove Outlier
cooksd = cooks.distance(model)
influential = as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
train_df = train_df[-c(influential), ]

# New diagnostic plots
model_new = lm(TARGET_WINS ~ ., data = train_df)
ols_plot_diagnostics(model_new)
```

Now, an investigation is perform to calculate the amount of missing/imputed values in each variable. The plot helps to understand that almost 85% of the samples are not missing any information, whereas below states the percentage of missing values per variables. The method of choice to handle these missing values is imputation using predictive mean matching, to replace missing data with a randomly chosen value from several similar cases.

```{r}
aggr(train_df, col = c('darkseagreen2','darkslateblue'), numbers = TRUE, sortVars = TRUE, oma = c(10,5,5,3), labels = names(train_df), cex.axis = 0.8, gap = 3, axes = TRUE, Prop = TRUE, ylab = c("Proportion of Missing Data", "Combination"))

train_df.clean = complete(mice(data = train_df, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
sapply(train_df.clean, describe)
```

***
#### Initial Tests 
***

Because the data consist of variable called `TEAM_BATTING_H`, base hits by batters, i.e. it is linear combination of H = 1B + 2B + 3B + HR, there are concerns of possible multicollinearity. Therefore, the single hits by batter found and base hits were removed. Next, a collinearity diagnostic test is done to examining the diagnostic output for variance inflation factor, tolerance, and Farrar-Glauber F-test. The F-statistic for the variable `TEAM_BATTING_HR` is quite high (42.1158) followed by the variable `TEAM_PITCHING_HR` (F-value of 33.8885). So, the test shows that there are multiple variables that will be the root cause of multicollinearity. Moreover, as expected, there are high partial correlations found to be statistically significant. As a solution to deal with multicollinearity, there are several remedial measures will be used as a result of this diagnostic test. Some included removal of highly correlated variables and stepwise regression analysis were done.

```{r}
# Data Transformation
train_df.clean$TEAM_BATTING_1B = train_df.clean$TEAM_BATTING_H - (train_df.clean$TEAM_BATTING_2B + train_df.clean$TEAM_BATTING_3B + train_df.clean$TEAM_BATTING_HR)
train_df.clean = subset(train_df.clean, select = -c(TEAM_BATTING_H))

# Multicollinearity Diagnostic Measures
model_new = lm(TARGET_WINS ~ ., data = train_df.clean)
imcdiag(model_new, method = "VIF", all = TRUE)

# Data Transformation & Multicollinearity Diagnostic Measures
train_df.clean = subset(train_df.clean, select = -c(TEAM_PITCHING_HR))
model_new = lm(TARGET_WINS ~ ., data = train_df.clean)
imcdiag(model_new, method = "VIF", all = TRUE)
```

```{r, fig.width=7, fig.height=9}
# Correlation
corr = round(cor(train_df.clean), 2)
corr = cbind(names = rownames(corr), data.frame(corr, row.names=NULL))
g.corr = gather(as.data.frame(corr), key, value, 2:length(corr), na.rm = TRUE)

ggplot(data = g.corr, aes(key, names, fill = value)) + geom_tile(color = "white") +
 scale_fill_gradient2(low = "darkseagreen2", high = "darkslateblue", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) +
 coord_fixed() + geom_text(aes(key, names, label = value), color = "black", size = 3) 
```

Lastly, because a robust regression process will be performed, higher order polynomials variables were introduced into the full model.

```{r}
stepdata = cbind(train_df.clean,sapply(train_df.clean[2:length(train_df.clean)], function(x) x^2))
names(stepdata) = make.unique(names(stepdata), sep = "_")
```

***
#### Regression Model #1
***

A stepwise variable selection model is conducted to determine what are the variables that can help predict the number of wins for the team. The stepwise variable selection allows variables to be added one at a time to the model, as long as the F-statistic is below the specified $\alpha$, in this case $\alpha = 0.05$. However, variables already in the model do not necessarily stay in. The steps evaluate all of the variables already included in the model and remove any variable that has an insignificant F-statistic. Only after this test ends, is the best model found, that is when none of the variables cstepan be excluded and every variable included in the model is significant.

Here the dependent variable is the continuous variable, `TARGET_WINS`, and the independent variables are the full model to identify the most contributing predictors. In addition, a robust method for estimating the accuracy of a model, the k-fold cross-validation method, was performed evaluate the model performance on different subset of the training data and then calculate the average prediction error rate.

After the steps, the final model resulted below, with $adj. R^2 = 0.42$, suggesting that this model accounts for nearly 42% of the variation in the dependent variable with the independent variables, which is acceptable as a good model. Within the method of `lmStepAIC`, AIC (Akaike Information Criteria) quantifies the amount of information loss due to simplification. That is, based on the AIC, the final model outputted is the simplest model without impacting much on the performance.

```{r  message=FALSE, warning=FALSE, paged.print=FALSE}
# Set up repeated k-fold cross-validation
set.seed(525)
train.control = trainControl(method = "cv", number = 10)

# Train the model
step.model = train(TARGET_WINS ~ ., data = stepdata, method = "lmStepAIC", trControl = train.control, trace = FALSE)

# Model accuracy
step.model$results

# Final model coefficients
step.model$finalModel

# Summary of the model
summary(step.model$finalModel)
```

***
#### Prediction
***

Studying the coefficients of the model suggest that winning is in favor if the team batting hits more doubles, triples and home runs. Moreover, increase in the number of stolen bases, and a decrease in caught steals, double plays, error, and walks allowed would all lead to a win for the batting team. It is noteworthy that the model suggests that a decrease in single hits by batter and an increase in strikeouts by batters which seems counter intuitive. But these variables were kept because when a batter steps to the plate, the player is more likely to strike out than to get a hit. Trying to hit the ball out of the park will come with strikeouts but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry out.

```{r message=FALSE, warning=FALSE}
# load the test data
test_df = read.csv("C:/Users/Deokinanan/Desktop/2019 -2021 CUNY SPS/FALL 2020/DATA 621/WEEK 3/moneyball-evaluation-data.csv")

# Data Transformation to match train data
test_df = complete(mice(data = test_df, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
test_df$TEAM_BATTING_1B = test_df$TEAM_BATTING_H - (test_df$TEAM_BATTING_2B + test_df$TEAM_BATTING_3B + test_df$TEAM_BATTING_HR)
test_df = cbind(test_df,sapply(test_df[1:length(test_df)], function(x) x^2))
names(test_df) = make.unique(names(test_df), sep = "_")

predictions = predict(step.model, newdata = test_df)

results = cbind(dataset = c('Training Data', 'Test Prediction'), bind_rows(describe(train_df$TARGET_WINS), describe(predictions))) 
rownames(results) = NULL

results
```

### 2020-09-08 HW1/RegModel_Adler.Rmd
Start at data exploration

---
title: "Data 621---HW1---Regression Model Pt 2"
author: "Group 1"
date: "9/8/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(data.table)
library(ggplot2)
```

# Data Exploration
## Task
Describe the size and the variables in the moneyballtraining data set. Consider
that too much detail will cause a manager to lose interest while too little
detail will make the manager consider that you aren’t doing your job. Some
suggestions are given below. Please do NOT treat this as a check list of things
to do to complete the assignment. You should have your own thoughts on what to
tell the boss. These are just ideas.

  a. Mean / Standard Deviation / Median
  b. Bar Chart or Box Plot of the data
  c. Is the data correlated to the target variable (or to other variables?)
  d. Are any of the variables missing and need to be imputed “fixed”?

## Analysis
```{r readData}
MB <- fread('./data/moneyball-training-data.csv')
n <- dim(MB)[[1]]
k <- dim(MB)[[2]]

# Meltng data from wide to narrow for munging & ggplot; simultaneously removing
# INDEX.
MBmelt <- melt(MB[, -1L], measure.vars = names(MB[, -1L]),
               variable.name = 'Metric', value.name = 'Value',
               variable.factor = FALSE)
```

### Description
The data is composed of `r n` observations of `r k` fields, all of which are
discrete numerical variables.

### Missing Observations
Most variables are present in their entirety. Some have a few observations
missing that may be addressed through imputation. Unfortunately,
`TEAM_BATTING_HBP`---the metric representing batters getting first base awarded
due to being hit by a pitch---is missing almost all observations, it is most
prudent to **exclude** from the analysis. The table below shows all variables
with missing values

```{r missingTable}
missingObs <- MBmelt[, .(Missing = sum(is.na(Value)),
                         Percentage = sum(is.na(Value)) / length(Value)),
                     keyby = Metric][Missing > 0][order(-Missing)]
knitr::kable(missingObs, digits = 3L)
```

\clearpage
### Data Distribution
The table and plots below give an overview as to the distribution of the
variables.

```{r summaryTable}
MBmelt <- MBmelt[Metric != 'TEAM_BATTING_HBP']
# Casting to double since medians and quartiles may return non-integer values.
MBmelt[, Value := as.double(Value)]

statTable <- MBmelt[, .(Min = min(Value, na.rm = TRUE),
                        Q1 = quantile(Value, 0.25, na.rm = TRUE),
                        Median = median(Value, na.rm = TRUE),
                        Q3 = round(quantile(Value, 0.75, na.rm = TRUE)),
                        Max = max(Value, na.rm = TRUE),
                        Mean = mean(Value, na.rm = TRUE),
                        StDev = sd(Value, na.rm = TRUE),
                        IQR = IQR(Value, na.rm = TRUE)), keyby = Metric]
knitr::kable(statTable, digits = 2L)
```

```{r boxPlot1}
ggplot(MBmelt, aes(x = Value, y = Metric)) + geom_boxplot()
```

One immediate observation from both the table and the boxplots, is that the
range for `TEAM_PITCHING_H` is one to two orders of magnitude greater than that
of the other variables. Below are boxplots of the remaining variables for a
better view, followed by a histogram of `TEAM_PITCHING_H`.

```{r boxPlot2}
ggplot(MBmelt[Metric != 'TEAM_PITCHING_H'], aes(x = Value, y = Metric)) +
  geom_boxplot()

# Use the Freedman-Diaconis rule for historgram bin widths instead of ggplot
# defaults

FreedmanDiaconis <- function(x) {
  2 * IQR(x, na.rm = TRUE) / (length(x) ^ (1 / 3))
}

ggplot(MBmelt[Metric == 'TEAM_PITCHING_H'], aes(x = Value)) +
  geom_histogram(binwidth = FreedmanDiaconis) +
  scale_x_continuous(labels = scales::label_comma())
```

It is very clear that `TEAM_PITCHING_H` is extremely right tailed and may be a
prime subject for a transformation.

# Appendix
The code chunks below represent the code called in order during the analysis.
They are reproduced in the appendix for review and comment.

```{r readData, echo=TRUE, eval=FALSE}
```
```{r missingTable, echo=TRUE, eval=FALSE}
```
```{r summaryTable, echo=TRUE, eval=FALSE}
```
```{r boxPlot1, echo=TRUE, eval=FALSE}
```
```{r boxPlot2, echo=TRUE, eval=FALSE}
```

### 2020-09-08 HW1/RegModel_Adler.Rmd
Stopping for now

@@
 ---
-title: "Data 621---HW1---Regression Model Pt 2"
-author: "Group 1"
+title: "Data 621---HW1---Regression Model---Version 2"
+author: "Group 2"
 date: "9/8/2020"
@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+```
+```{r loadPackages}
+library(corrplot)
+library(RColorBrewer)
+library(ggplot2)
+library(scales)
 library(data.table)
-library(ggplot2)
 ```
@@
 ### Missing Observations
-Most variables are present in their entirety. Some have a few observations
-missing that may be addressed through imputation. Unfortunately,
-`TEAM_BATTING_HBP`---the metric representing batters getting first base awarded
-due to being hit by a pitch---is missing almost all observations, it is most
-prudent to **exclude** from the analysis. The table below shows all variables
-with missing values
+The table below shows all variables with missing values:
 
@@
 
-\clearpage
+Most variables are present in their entirety. Some have a few observations
+missing that may be addressed through imputation. Unfortunately,
+`TEAM_BATTING_HBP`---the metric representing batters getting first base awarded
+due to being hit by a pitch---is missing almost all observations. It is prudent
+to remove that variable from the analysis. It is also probably prudent to
+remove `TEAM_BASERUN_CS` as it is missing more than one-third of the time. The
+variables missing less than 10% of the time may be imputed using median or
+k-means clustering. Whether `TEAM_FIELDING_DP` will be removed or imputed will
+be a decision made later on in the analysis.
+
 ### Data Distribution
@@
 ```{r summaryTable}
+# Removing 'TEAM_BATTING_HBP' and index from wide data. Easier to calculate
+# correlations later on using wide data.
+MB[, `:=`(INDEX = NULL, TEAM_BATTING_HBP = NULL)]
+
+# Removing 'TEAM_BATTING_HBP' from tall data
 MBmelt <- MBmelt[Metric != 'TEAM_BATTING_HBP']
+
 # Casting to double since medians and quartiles may return non-integer values.
@@
 
-```{r boxPlot1}
-ggplot(MBmelt, aes(x = Value, y = Metric)) + geom_boxplot()
+```{r boxPlot1, fig.width = 7}
+ggplot(MBmelt, aes(x = Value, y = Metric)) + geom_boxplot() +
+  scale_x_continuous(labels = comma)
 ```
@@
 
-```{r boxPlot2}
+```{r boxPlot2, fig.width = 7}
 ggplot(MBmelt[Metric != 'TEAM_PITCHING_H'], aes(x = Value, y = Metric)) +
-  geom_boxplot()
+  geom_boxplot() + scale_x_continuous(labels = comma)
 
@@
   geom_histogram(binwidth = FreedmanDiaconis) +
-  scale_x_continuous(labels = scales::label_comma())
+  scale_x_continuous(labels = comma)
 ```
@@
 
+### Correlations
+A quick way to identify correlations between variables is through a
+corrgram.[^1] In the corrgram below, blue represents positive correlation and
+red represents negative correlation, as shown in the color bar. The shape of the
+images in each cell reflect the scatterplots of the features with each other.
+The main diagonal, therefore, are all upward sloping blue lines, as each feature
+is perfectly correlated with itself.
+
+[^1]:This is the proper name for a multicolored matrix representation of a
+correlation matrix (Friendly 2002). Correlelograms refer to the time-series
+plots used to check for auto-correlation. The terms have been conflated, though.
+
+```{r corrgram, fig.height=6}
+MBcor <- cor(MB, method = 'pearson', use = 'pairwise.complete.obs')
+corrplot(MBcor, 'ellipse', type = 'lower', order = 'hclust',
+         col=brewer.pal(n=8, name="RdYlBu"))
+```
+
+Most of the ellipses are pretty wide, so there is not *much* correlation, but
+there are some clear clusters. For example, `TEAM_FIELDING_E` and 
+`TEAM_BATTING_3B` are both negatively correlated with `TEAM_BATTING_SO` and
+`TEAM_BATTING_HR`. As regards `TEAM_FIELDING_E`, a hypothesis is that if batters
+are either hitting home runs or striking out, *they are***not***getting on base*.
+Therefore there are fewer opportunity for fielding errors. Why triples are 
+negatively correlated with HRs or SOs is not as logical. There is a positive
+correlation between triples and fielding errors of around 51%. Perhaps triples
+are more a function of fielding than batting?
+
+Regardless, there are some correlations in the data, but nothing that screams
+for immediate action.
+
+# References
+
+ * Friendly, Michael (2002) Corrgrams: Exploratory displays for correlation
+ matrices, *The American Statistician*, 56:4, 316-324, DOI: 10.1198/000313002533,
+ http://euclid.psych.yorku.ca/datavis/papers/corrgram.pdf
+
 # Appendix
@@
 
+```{r loadPackages, echo=TRUE, eval=FALSE}
+```
 ```{r readData, echo=TRUE, eval=FALSE}
@@
 ```
+```{r corrgram, echo=TRUE, eval=FALSE}
+```

### 2020-09-08 HW1/RegModel_Adler.Rmd
Imputation, center/scaling, start commentary on results of c/s

@@
 ```{r loadPackages}
+library(caret)
 library(corrplot)
@@
                      keyby = Metric][Missing > 0][order(-Missing)]
-knitr::kable(missingObs, digits = 3L)
+knitr::kable(missingObs, digits = 3L,
+caption = 'Training Data with Missing Observations')
 ```
@@
                         IQR = IQR(Value, na.rm = TRUE)), keyby = Metric]
-knitr::kable(statTable, digits = 2L)
+knitr::kable(statTable, digits = 2L,
+             caption = 'Training Data Summary Statistics')
 ```
@@
 
+# Data Preperation
+## Task
+Describe how you have transformed the data by changing the original variables or
+creating new variables. If you did transform the data or create new variables,
+discuss why you did this. Here are some possible transformations:
+
+  a. Fix missing values (maybe with a Mean or Median value)
+  b. Create flags to suggest if a variable was missing
+  c. Transform data by putting it into buckets
+  d. Mathematical transforms such as log or square root(or use Box-Cox)
+  e. Combine variables (such as ratios or adding or multiplying) to create new
+  variables
+
+## Analysis
+The `caret` package for R is the gold-standard catch-all interface for machine
+learning techniques for those not using the tidyverse. Otherwise, the same
+author's `tidymodels` package may be used as well. When using `caret`, almost
+all of the following imputation and transformation is done as pre-processing
+to the model. For the purposes of this homework, they are being performed
+manually to demonstrate understanding.
+
+### Imputation
+Since we are restricting ourselves to the world of the simple linear regression
+model, it makes sense to use a Z-transformation to standardize the models.
+Prior to this, the missing variables should be imputed. Looking at the summary
+statistics table, the median is not that far off of the mean in most cases with
+the exceptions of `TEAM_BASERUN_SB` and `TEAM_BATTING_SO`. Therefore, for an
+initial imputation, we will use the mean so that the overall mean is unaffected
+for the Z-transform. This **will** artificially reduce the SD, however. We could
+address that by calculating the mean and SD, *prior* to imputation. We will not
+do so now for simplicity.
+
+```{r imputeMean}
+MBmelt[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
+       keyby = Metric]
+```
+
+### Transformations
+The first transform applied will be centering and scaling, as discussed above.
+
+```{r centerScale}
+MBmelt[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+```
+
+At this point we would hope to see Gaussian-looking densities.
+```{r plotScaled}
+ggplot(MBmelt, aes(x = Value)) + geom_density() +
+  facet_wrap(~ Metric, scales = 'free')
+```
+
+Some of the variables, like `TEAM_BATTING_2B` or `TEAM_BATTING_SO` show
+Gaussian-like behavior, with some allowances for the discrete nature of the
+input data (those wiggles at the crest of `TEAM_BATTING_SO` for example). Others
+are classic right tailed (exponential/gamma-esque) such as `TEAM_FIELDING_E` or
+`TEAM_BATTING_3B`. Still others are clearly multimodal. Then there are the
+weird spike-like ones as well.
+
 # References
@@
 ```
+```{r imputeMean, echo=TRUE, eval=FALSE}
+```
+```{r centerScale, echo=TRUE, eval=FALSE}
+```

### 2020-09-08 HW1/RegModel_Adler.Rmd
Center & Scale not enough. Moving to `caret` package

@@
 ### Transformations
+#### Centering and Scaling
 The first transform applied will be centering and scaling, as discussed above.
+We will store the means and standard deviations so that we can transform the
+predictions back to their respective distributions.
 
 ```{r centerScale}
-MBmelt[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+MBmeltSC <- copy(MBmelt)
+MBmeltUnscale <- MBmeltSC[, .(Mean = mean(Value), SD = sd(Value)),
+                          keyby = Metric]
+MBmeltSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+knitr::kable(MBmeltUnscale, digits = 2L, caption = 'Data Means and SDs')
 ```
@@
 ```{r plotScaled}
-ggplot(MBmelt, aes(x = Value)) + geom_density() +
+ggplot(MBmeltSC, aes(x = Value)) + geom_density() +
   facet_wrap(~ Metric, scales = 'free')
@@
 weird spike-like ones as well.
+
+All of this indicates that there will need to be further processing.
+
+#### Box-Cox transformations
+One of classic data transformations to induce normality is the Box-Cox
+transform. It is defined as:
+\[
+y(\lambda) =
+\begin{cases}
+\frac{y^\lambda - 1}{\lambda}, &\quad \lambda \neq 0\\
+\ln{y}, &\quad \lambda = 0
+\end{cases}
+\]
+
+Unfortunately, this is only defined on positive data. Therefore, the unscaled
+data will need to be used. While in theory it would be nice to transform each
+column as best possible, that makes for very difficult reconstruction. At this
+point, we will pivot to continuing the analysis using the `caret` package which
+takes care of a lot of the computational issues.
 

### 2020-09-08 HW1/RegModel_Adler.Rmd
Hand-picking Box-Cox parameters

@@
 
-Most of the ellipses are pretty wide, so there is not *much* correlation, but
-there are some clear clusters. For example, `TEAM_FIELDING_E` and 
+Most of the ellipses are pretty wide, so there is not *much* correlation. There
+is one notable exception. `TEAM_BATTING_HR` is about **97%** correlated with
+`TEAM_PITCHING_HR`, as the individual HR directly affects the single determinant
+of both wins and losses---the score! There is no other correlation greater than
+67%. That one is `TEAM_FIELDING_E` with `TEAM_PITCHING_H` which stands to reason,
+as one can only be charged with an error if the ball is in play, which is also
+the only way to get a hit.
+
+There are some interesting clusters. For example, `TEAM_FIELDING_E` and 
 `TEAM_BATTING_3B` are both negatively correlated with `TEAM_BATTING_SO` and
@@
 
-Regardless, there are some correlations in the data, but nothing that screams
-for immediate action.
+Our suggestion would be to use only `TEAM_PITCHING_HR` as that has a slightly
+higher correlation with target wins, and leave all the other variables in the
+analysis.
 
@@
 ```
+```{r plotScaled, fig.width=7}
+ggplot(MBmelt, aes(x = Value)) + geom_density() +
+  facet_wrap(~ Metric, scales = 'free') +
+  ggtitle(label = 'Kernel-Smoothed Densities of Raw Data')
+ggplot(MBmeltSC, aes(x = Value)) + geom_density() +
+  facet_wrap(~ Metric, scales = 'free') +
+  ggtitle(label = 'Kernel-Smoothed Densities of Centered and Scaled Data')
+```
 
-At this point we would hope to see Gaussian-looking densities.
-```{r plotScaled}
-ggplot(MBmeltSC, aes(x = Value)) + geom_density() +
-  facet_wrap(~ Metric, scales = 'free')
-```
+As expected, centering and scaling did not change the shape of the
+distributions. As an aside, the relationship we saw between `TEAM_PITCHING_HR`
+and `TEAM_BATTING_HR` above remains rather clear in the plots.
 
@@
 input data (those wiggles at the crest of `TEAM_BATTING_SO` for example). Others
-are classic right tailed (exponential/gamma-esque) such as `TEAM_FIELDING_E` or
-`TEAM_BATTING_3B`. Still others are clearly multimodal. Then there are the
-weird spike-like ones as well.
+are classic right tailed (exponential/gamma-esque) such as `TEAM_FIELDING_E`
+or `TEAM_BATTING_3B`.
+
+Still others are clearly multimodal. Variables like `TEAM_BATTING_HR` and
+`TEAM_PITCHING_HR` indicate the existence of some clustering of better and worse
+players. With these two, As said above, the "wiggles" in `TEAM_BATTING_SO` and
+`TEAM_BASERUN_SB` may be more artifacts of discrete data than indications of
+multimodality.
+
+Then there is `TEAM_BASERUN_CS` which has an extreme spike at one point and then
+some outliers on either side.
 
@@
 
-Unfortunately, this is only defined on positive data. Therefore, the unscaled
-data will need to be used. While in theory it would be nice to transform each
-column as best possible, that makes for very difficult reconstruction. At this
-point, we will pivot to continuing the analysis using the `caret` package which
-takes care of a lot of the computational issues.
+with \(\lambda\) usually allowed to range between -5 and 5. Unfortunately, this
+is only defined on positive data. Therefore, the unscaled data will need to be
+used and then subsequently rescaled.
+
+As mentioned in the introduction, most of the actual computation of transforms
+is handled by the software, such as the `caret` package. For the purposes of
+pedagogy, we will make some hand-selected transforms.
+
+```{r boxCox}
+boxCox <- function(x, l) {
+  if (l == 0) {
+    log(x)
+  } else {
+    (x ^ l - 1) / l
+  }
+}
+
+# All of these values determined by visual inspection of resulting density
+# functions. A better approach would be some kind of optimization to normality,
+# perhaps minimizing the square of the skewness and *excess* kurtosis, both of
+# which are 0 in the standard normal distribution.
+
+TEAM_PITCHING_H <- -2.5
+TEAM_FIELDING_E <- -0.5
+TEAM_BASERUN_SB <- 0
+TEAM_PITCHING_HR <- 0.6
+TEAM_PITCHING_BB <- -0.15
+TEAM_PITCHING_SO <- -0.35
+
+```
+
+# Build Models
+## Task
+Using the training data set, build at least three different multiple linear
+regression models, using different variables (or the same variables with
+different transformations). Since we have not yet covered automated variable
+selection methods, you should select the variables manually (unless you
+previously learned Forward or Stepwise selection, etc.). Since you manually
+selected a variable for inclusion into the model or exclusion into [*sic*] the
+model, indicate why this was done. Discuss the coefficients in the models, do
+they make sense? For example, if a team hits a lot of Home Runs, it would be
+reasonably expected that such a team would win more games. However, if the
+coefficient is negative (suggesting that the team would lose more games), then
+that needs to be discussed. Are you keeping the model even though it is counter
+intuitive? Why? The boss needs to know.
 
@@
 ```
+```{r plotScaled, echo=TRUE, eval=FALSE}
+```

### 2020-09-08 HW1/RegModel_Adler.Rmd
Complete the Box-Cox saga

@@
 ```{r boxCox}
-boxCox <- function(x, l) {
-  if (l == 0) {
-    log(x)
-  } else {
-    (x ^ l - 1) / l
-  }
+BxCox <- function(x, l) {
+  ifelse(l == 0, log(x), (x ^ l - 1) / l)
 }
@@
 
-TEAM_PITCHING_H <- -2.5
-TEAM_FIELDING_E <- -0.5
-TEAM_BASERUN_SB <- 0
-TEAM_PITCHING_HR <- 0.6
-TEAM_PITCHING_BB <- -0.15
-TEAM_PITCHING_SO <- -0.35
+BCL <- data.table(Metric = c('TEAM_PITCHING_H', 'TEAM_FIELDING_E',
+                             'TEAM_BASERUN_SB', 'TEAM_PITCHING_HR',
+                             'TEAM_PITCHING_BB', 'TEAM_PITCHING_SO',
+                             'TEAM_BATTING_3B'),
+                  Lambda = c(-2.5, -0.5, 0, 0.6, -0.15, -0.35, 0))
+setkey(BCL, Metric)
+knitr::kable(BCL, digits = 2L, caption = 'Hand-selected Box-Cox Lambdas')
 
+# Left join MBmelt to BCL. 
+MBmelt <- BCL[MBmelt]
+
+# Transform data where there are selected lambdas
+MBmelt[!is.na(Lambda), Value := BxCox(Value, Lambda)]
+
+# Remove TEAM_BATTING_HR
+
+MBmelt <- MBmelt[Metric != 'TEAM_BATTING_HR']
+
+# Plot results
+ggplot(MBmelt, aes(x = Value)) + geom_density() +
+  facet_wrap(~ Metric, scales = 'free') +
+  ggtitle(label = 'Kernel-Smoothed Densities of Box-Cox Transformed Data')
 ```
 
+Most of the variables are now closer to a Gaussian distribution. Also note we
+have removed `TEAM_BATTING_HR` from the data. Now the data needs to be rescaled
+and recentered.
+
+```{r reScale}
+MBmeltSC <- copy(MBmelt)
+MBmeltSC[, Lambda := NULL]
+MBmeltUnscale <- MBmeltSC[, .(Mean = mean(Value), SD = sd(Value)),
+                          keyby = Metric]
+
+MBmeltSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+knitr::kable(MBmeltUnscale, digits = 2L,
+             caption = 'Post-Transform Data Means and SDs')
+```
+
+Unfortunately, the scaling introduces infinities, as most of the Box-Cox lamdas
+are non-positive, which transforms 0 into infinity. These were converted to NAs
+for the purposes of imputation, after which scaling and centering will be done
+again. This is the danger of too many transforms. Each one induces a "stress" on
+the relationship between the independent and dependent variables. At what point
+does that relationship fracture?
+
+```{r reImputeThenScale}
+MBmeltSC <- copy(MBmelt)
+MBmeltSC[, Lambda := NULL]
+# Remove infinities
+MBmeltSC[, Value := ifelse(is.infinite(Value), NA, Value)]
+# Center and Scale
+MBmeltSC[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
+       keyby = Metric]
+MBmeltUnscale <- MBmeltSC[, .(Mean = mean(Value), SD = sd(Value)),
+                          keyby = Metric]
+MBmeltSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+knitr::kable(MBmeltUnscale, digits = 2L,
+             caption = 'Final Post-Transform Data Means and SDs')
+ggplot(MBmeltSC, aes(x = Value)) + geom_density() +
+  facet_wrap(~ Metric, scales = 'free') +
+  ggtitle(label = 'Kernel-Smoothed Densities of Scaled Box-Cox Transformed Data')
+```
 # Build Models
@@
 intuitive? Why? The boss needs to know.
+
+## Analysis
 
@@
 ```
+```{r boxCox, echo=TRUE, eval=FALSE}
+```
+```{r reScale, echo=TRUE, eval=FALSE}
+```
+```{r reImputeThenScale, echo=TRUE, eval=FALSE}
+```

### 2020-09-09 HW1/RegModel_Adler.Rmd
Needed INDEX for casting to work properly!!

@@
 
-# Meltng data from wide to narrow for munging & ggplot; simultaneously removing
-# INDEX.
-MBmelt <- melt(MB[, -1L], measure.vars = names(MB[, -1L]),
-               variable.name = 'Metric', value.name = 'Value',
-               variable.factor = FALSE)
+# Meltng data from wide to narrow for munging & ggplot
+MBmelt <- melt(MB, id.vars = 'INDEX', variable.name = 'Metric',
+               value.name = 'Value', variable.factor = FALSE)
 ```
@@
 # correlations later on using wide data.
-MB[, `:=`(INDEX = NULL, TEAM_BATTING_HBP = NULL)]
+MB[, `:=`(TEAM_BATTING_HBP = NULL)]
 
@@
 ```{r corrgram, fig.height=6}
-MBcor <- cor(MB, method = 'pearson', use = 'pairwise.complete.obs')
+MBcor <- cor(MB[, -1L], method = 'pearson', use = 'pairwise.complete.obs')
 corrplot(MBcor, 'ellipse', type = 'lower', order = 'hclust',
@@
 `TEAM_BATTING_HR`. As regards `TEAM_FIELDING_E`, a hypothesis is that if batters
-are either hitting home runs or striking out, *they are***not***getting on base*.
-Therefore there are fewer opportunity for fielding errors. Why triples are 
-negatively correlated with HRs or SOs is not as logical. There is a positive
-correlation between triples and fielding errors of around 51%. Perhaps triples
-are more a function of fielding than batting?
+are either hitting home runs or striking out, *they are* **not**
+*getting on base*. Therefore there are fewer opportunity for fielding errors.
+Why triples are negatively correlated with home runs or strikeouts is not as
+logical. There is a positive correlation between triples and fielding errors of
+around 51%. Perhaps triples are more a function of fielding than batting?
 
@@
 
+# Remove Lambda
+MBmelt[, Lambda := NULL]
+
 # Plot results
@@
 MBmeltSC <- copy(MBmelt)
-MBmeltSC[, Lambda := NULL]
 MBmeltUnscale <- MBmeltSC[, .(Mean = mean(Value), SD = sd(Value)),
@@
 ```
+
 # Build Models
@@
 ## Analysis
+At this point we have a set of transformed independent variables on which we
+can build a model. The target variable, `TARGET_WINS` does not need to be
+scaled. Normally, we would have thrown everything into `caret`, but for the
+purposes of this homework, we will use the `stepAIC` function from the `MASS`
+package to perform a backwards-stepwise linear regression using AIC as our
+goodness of fit metric. For ease, we will cast the dataset we were using from
+tall back to wide first.
+
+```{r lm1}
+MBsc <- dcast(MBmeltSC, INDEX ~ Metric, value.var = 'Value')
+```
 

### 2020-09-09 HW1/RegModel_Adler.Rmd
In this example will not use caret. Just an example.

@@
 ```{r loadPackages}
-library(caret)
 library(corrplot)

### 2020-09-09 HW1/reg_model_shamp.Rmd
update eda, add model

---
title: "HW! - Regression Model - 621"
author: "Jeff Shamp"
date: "9/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(mice)
library(psych)
```

## Data

We have the mlb training data set. See `missing_value_eda.Rmd` for details on data distribution.

We saw from the above file that at least one variable should be omitted for low reporting of data. There are also outlier concerns for several of the variables, namely `TEAM_PITCHING_SO`.

First, we will look into obvious outliers in the data set and remove them. There appears to be enough data in the training set to naive yet cautious approach to outlier removal.

```{r}
set.seed(9450)
train_df<- 
  read.csv("/Users/jeffshamp/Documents/Github/DT621_Fall2020_Group2/HW1/data/moneyball-training-data.csv")
train_df<- 
  train_df %>% 
  select(-c(TEAM_BATTING_HBP, INDEX)) %>%
  mutate_if(is.integer, as.numeric)
```

### Isolate base hits.

The variable `TEAM_BATTING_H` is all hits. We will separate out singles from the rest.

```{r}
train_df<- 
    train_df %>%
    mutate(TEAM_BATTING_1B =
           TEAM_BATTING_H -
           (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR))

```

Looking at some distributions.

```{r}
train_df %>%
  ggplot(aes(TEAM_BATTING_1B)) +
  geom_histogram(na.rm = TRUE, binwidth = 1) +
   geom_vline(xintercept = mean(
              train_df$TEAM_BATTING_1B,
              na.rm = TRUE), color="red") 
```

## Outlier Removal

Again, we will take a cautious approach to removing outliers. Removing values that appear to be so far out from the normal that is it seems likely that there is some kind of data input issue.

```{r}
train_df %>%
  pivot_longer(everything(), 
               names_to ="Variable", 
               values_to="Value") %>%
  ggplot(aes(x=Variable, y=Value)) +
  geom_boxplot(na.rm = TRUE) + coord_flip()
```

There are some clear issues with hitting singles and some that we know about with strikeouts. We will remove the SO outliers first and see what needs to happen with singles hitting. These appear to right skewed so we will use median and IQR to remove outliers.

```{r}
train_df<-
train_df %>%
  na.omit() %>%
  summarize(iqr = IQR(TEAM_PITCHING_SO)) %>%
  bind_cols(
            train_df 
            )  %>%
  filter(
    TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO,
                                probs=c(0.25), 
                                na.rm = TRUE) - 1.5*iqr, 
    TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, 
                                probs=c(0.75), 
                                na.rm=TRUE) + 1.5*iqr
        ) %>%
  select(-iqr)

train_df %>%
  pivot_longer(everything(), 
               names_to ="Variable", 
               values_to="Value") %>%
  ggplot(aes(x=Variable, y=Value)) +
  geom_boxplot(na.rm = TRUE) + coord_flip()
```

We will take a similar approach to removing outliers for pitching hitting allowed. At least one seems unrealisticly extreme.

```{r}
train_df<-
train_df %>%
  na.omit() %>%
  summarize(iqr = IQR(TEAM_PITCHING_H)) %>%
  bind_cols(
            train_df 
            )  %>%
  filter(
    TEAM_PITCHING_H > quantile(TEAM_PITCHING_H,
                                probs=c(0.25), 
                                na.rm = TRUE) - 1.5*iqr, 
    TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, 
                                probs=c(0.75), 
                                na.rm=TRUE) + 1.5*iqr
        ) %>%
  select(-iqr)

```

```{r}
train_df %>%
  pivot_longer(everything(), 
               names_to ="Variable", 
               values_to="Value") %>%
  ggplot(aes(x=Variable, y=Value)) +
  geom_boxplot(na.rm = TRUE) + coord_flip()
```

This has eliminated what appears to be the most extreme outliers. There are still large outlier sets for Errors and stolen bases, but none that seem to dwarf the other variables. Basically, things appear to be on a scale that seems logical given the data and source of information that we have.

## Imputation of Values

From the `missing_value_eda.Rmd` the vast majority of the data is complete with only a few variables with missing values. We will use the "mice" imputation method [described here](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3074241/#:~:text=Missing%20data%20are%20a%20common,method%20of%20addressing%20missing%20data.).

We will use the predictive mean matching method.

```{r, message=FALSE}
train_df <- complete(mice(data = train_df, 
                         method = "norm", 
                         seed = 9450, 
                         print=FALSE), 3)
```

## Feature Engineering

With outliers removed and missing values imputated via MICE, we can add a few features. We will take the obvious and approximate the following features; on-base percent, slugging, and on-base plu slugging. To reduce colinearity, we may only keep on-base percent and slugging.

### On Base Percentage

On-base percentage is calculated using this formula:

${\displaystyle OBP={\frac {H+BB+HBP}{AB+BB+HBP+SF}}}$

We are missing several of the needed variables but we can approximate this percentage in the following way.

${\displaystyle \sim OBP={\frac {H+BB}{SO_{batting}+BB+H}}}$

We do not have the total At-bats for each observation so we will approximate at-bats by summing batting strikeouts and total hits.

### Slugging Percentage

Slugging is calculated in the following way

${\displaystyle \mathrm {SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{AB}}}$

We will take the same approximation for at-bats.

${\displaystyle \mathrm {\sim SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{SO_{batting} + H + BB}}}$

```{r}
train_df<-
train_df %>%
  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
         OBP = (TEAM_BATTING_H + TEAM_BATTING_BB)/AB,
         SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) %>%
  # Removing AB and total hits
  select(-c(AB, TEAM_BATTING_H))
```

## Modeling

We will start with a simple linear model with resampling and see what kind of results we get.

```{r}
lm_reg = lm(data=train_df, TARGET_WINS ~ .)
summary(lm_reg)
```

Triples and singles seem to be candidates to remove from the analysis.

```{r}
lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
          TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
          TEAM_PITCHING_HR+ TEAM_PITCHING_BB+TEAM_PITCHING_SO+TEAM_PITCHING_H+
          TEAM_FIELDING_E+ TEAM_FIELDING_DP+OBP+ SLG, data=train_df)
summary(lm_1)
```

Let's take this backward selection process and use it in a resampling scheme and see how our R values hold up.

```{r}
folds<- vfold_cv(train_df,
                 v = 10)

lm_model<- 
  linear_reg() %>%
  set_engine(engine="lm")

lm_wf<- 
  workflow() %>%
  add_formula(
    TARGET_WINS ~ 
          TEAM_BATTING_2B +TEAM_BATTING_HR+
          TEAM_BATTING_BB +TEAM_BATTING_SO+ 
          TEAM_BASERUN_SB +TEAM_BASERUN_CS+ 
          TEAM_PITCHING_HR +TEAM_PITCHING_BB+
          TEAM_PITCHING_SO +TEAM_PITCHING_H+
          TEAM_FIELDING_E +TEAM_FIELDING_DP+
          OBP+ SLG) %>%
  add_model(lm_model)

lm_resample<- 
  fit_resamples(
    lm_wf,
    folds, 
    control = control_resamples(save_pred = TRUE)
  )

lm_resample %>%
  collect_metrics()
```
So the metrics didn't tank with cross validation, that's good. Let's fit the entire training data one last time and look at $R^{2}$ and RMSE.

```{r}
results_train<-
  lm_wf %>%
  fit(train_df) %>%
  predict(new_data = train_df) %>%
  mutate(
    truth = train_df$TARGET_WINS,
    model = "lm"
  )

multi_metric<- metric_set(rmse, rsq)

results_train %>%
  multi_metric(truth=truth, estimate = .pred)

```

## Test Data

Let's prep, fit, and evaluate the testing data.

```{r}
test_df<- 
  read.csv("/Users/jeffshamp/Documents/Github/DT621_Fall2020_Group2/HW1/data/moneyball-evaluation-data.csv")
test_df<- 
  test_df %>% 
  select(-c(TEAM_BATTING_HBP, INDEX)) %>%
  mutate(TEAM_BATTING_1B =
         TEAM_BATTING_H -
         (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR)) %>%
  mutate_if(is.integer, as.numeric)

test_df <- complete(mice(data = test_df,
                         method = "norm",
                         seed = 9450,
                         print=FALSE), 3)

test_df<-
  test_df %>%
  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
         OBP = (TEAM_BATTING_H + TEAM_BATTING_BB)/AB,
         SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) %>%
  select(-c(AB, TEAM_BATTING_H))

```

```{r}
test_pred<-
lm_wf %>%
  fit(train_df) %>%
  predict(new_data = test_df) %>%
  rename(predictions = .pred)

summary(test_pred)
```

Wow. max of 944, not good. How many terrible predictions are there? Let's filter by predictions beyond the currentl MLB season wins record of 116.

```{r}
test_pred %>%
  filter(predictions >= 116)
```

Ten beyond the record, two of which are somewhat reasonable. The rest are not reasonable.

### 2020-09-09 HW1/RegModel_Adler.Rmd
Model discussion

@@
 ```{r loadPackages}
+library(MASS)
 library(corrplot)
@@
 ## Analysis
+Normally, we would have thrown everything into `caret`, but for the purposes of
+this homework, we will use calls to base R for the null and saturated models and
+the `stepAIC` function from the `MASS` package to perform a stepwise linear
+regression using AIC as our goodness of fit metric.
+
 At this point we have a set of transformed independent variables on which we
-can build a model. The target variable, `TARGET_WINS` does not need to be
-scaled. Normally, we would have thrown everything into `caret`, but for the
-purposes of this homework, we will use the `stepAIC` function from the `MASS`
-package to perform a backwards-stepwise linear regression using AIC as our
-goodness of fit metric. For ease, we will cast the dataset we were using from
-tall back to wide first.
+can build a model. The reasoning for the variable selection and transformations
+are in previous sections. For ease, we will cast the dataset we were using from
+tall back to wide first. Furthermore, the target variable, `TARGET_WINS` does
+not need to be scaled.
 
-```{r lm1}
+We will show three models: the null model, the saturated model, and the model
+resulting from the stepwise regression. For the regression, we will consider at
+most pairwise interactions within the classes of metrics.
+
+```{r lm, include=FALSE}
 MBsc <- dcast(MBmeltSC, INDEX ~ Metric, value.var = 'Value')
+# Replace TARGET_WINS with original. Could have restored it from MBmeltUnscale.
+MBsc[, TARGET_WINS := MB[, TARGET_WINS]]
+# Removing INDEX for convenience. Could have used "-INDEX" in the formula call.
+MBsc[, INDEX := NULL]
+
+nullModel <- lm(TARGET_WINS ~ 1, data = MBsc)
+snm <- summary(nullModel)
+
+saturatedModel <- lm(TARGET_WINS ~ ., data = MBsc)
+ssm <- summary(saturatedModel)
+
+stepModel <- stepAIC(saturatedModel, scope = list(upper = ~
+  (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
+   TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_SO) ^ 2 + (TEAM_FIELDING_DP +
+   TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB + TEAM_PITCHING_H +
+   TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, lower = ~ 1), direction = 'both')
+
+sstm <- summary(stepModel)
 ```
+```{r nullOutput}
+knitr::kable(snm$coefficients, digits = 3L,
+             caption = 'Null Model Output')
+```
+The null model has an adjusted \(R^2\) of
+`r prettyNum(snm$adj.r.squared, digits = 3L)` and an AIC of
+`r prettyNum(AIC(nullModel), digits = 2L, big.mark = ',')`.
+
+```{r satOutput}
+knitr::kable(ssm$coefficients, digits = 3L,
+             caption = 'Saturated Model Output')
+```
+The saturated model has an adjusted \(R^2\) of
+`r prettyNum(ssm$adj.r.squared, digits = 3L)` and an AIC of
+`r prettyNum(AIC(saturatedModel), digits = 2L, big.mark = ',')`.
+
+```{r stepOutput}
+knitr::kable(sstm$coefficients, digits = 3L,
+             caption = 'Step Model with Pairwise Class Interactions Output')
+```
+
+The step model allowing pairwise interactions between the four classes of
+metric: base running, batting, fielding, and pitching, has an adjusted \(R^2\)
+of `r prettyNum(sstm$adj.r.squared, digits = 3L)` and an AIC of
+`r prettyNum(AIC(stepModel), digits = 2L, big.mark = ',')`.
 
@@
 ```
+```{r lm, echo=TRUE, eval=FALSE}
+```
+```{r nullOutput, echo=TRUE, eval=FALSE}
+```
+```{r satOutput, echo=TRUE, eval=FALSE}
+```
+```{r stepOutput, echo=TRUE, eval=FALSE}
+```

### 2020-09-09 HW1/RegModel_Adler.Rmd
Finished part 3 on part 4

@@
 ## Task
-Describe the size and the variables in the moneyballtraining data set. Consider
+Describe the size and the variables in the moneyball training data set. Consider
 that too much detail will cause a manager to lose interest while too little
@@
 
-Our suggestion would be to use only `TEAM_PITCHING_HR` as that has a slightly
-higher correlation with target wins, and leave all the other variables in the
-analysis.
+Initially, our thought was to only use `TEAM_PITCHING_HR` as that has a slightly
+higher correlation with target wins, but after running the model, it was clear
+that there was valid information captured in both variables so both were left in.
 
@@
 
-# Remove TEAM_BATTING_HR
-
-MBmelt <- MBmelt[Metric != 'TEAM_BATTING_HR']
-
 # Remove Lambda
@@
 
-Unfortunately, the scaling introduces infinities, as most of the Box-Cox lamdas
+Unfortunately, the scaling introduces infinities, as most of the Box-Cox lambdas
 are non-positive, which transforms 0 into infinity. These were converted to NAs
@@
 ## Analysis
-Normally, we would have thrown everything into `caret`, but for the purposes of
-this homework, we will use calls to base R for the null and saturated models and
-the `stepAIC` function from the `MASS` package to perform a stepwise linear
-regression using AIC as our goodness of fit metric.
+Normally, we would use a statistical packahe like `caret`, which not only
+simplifies model construction but also takes care of cross-validation. For the
+purposes of this homework, and to demonstrate understanding of the underlying
+decisions involved in predictive modeling, we will use calls to base R for the
+null and saturated models and the `stepAIC` function from the `MASS` package to
+perform a stepwise linear regression using AIC as our goodness of fit metric.
+In the final section, we may pivot to using `caret` for final model selection.
 
@@
 
+### Linear Models
 We will show three models: the null model, the saturated model, and the model
@@
 snm <- summary(nullModel)
+nmr2 <- prettyNum(snm$r.squared, digits = 3L)
+nmar2 <- prettyNum(snm$adj.r.squared, digits = 3L)
+nmAIC <- prettyNum(AIC(nullModel), digits = 2L, big.mark = ',')
 
@@
 ssm <- summary(saturatedModel)
+smr2 <- prettyNum(ssm$r.squared, digits = 3L)
+smar2 <- prettyNum(ssm$adj.r.squared, digits = 3L)
+smAIC <- prettyNum(AIC(saturatedModel), digits = 2L, big.mark = ',')
 
@@
   (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
-   TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_SO) ^ 2 + (TEAM_FIELDING_DP +
-   TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB + TEAM_PITCHING_H +
-   TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, lower = ~ 1), direction = 'both')
+  TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_SO) ^ 2 +
+  (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
+  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, lower = ~ 1),
+  direction = 'both')
 
 sstm <- summary(stepModel)
+stmr2 <- prettyNum(sstm$r.squared, digits = 3L)
+stmar2 <- prettyNum(sstm$adj.r.squared, digits = 3L)
+stmAIC <- prettyNum(AIC(stepModel), digits = 2L, big.mark = ',')
+
 ```
@@
 ```
-The null model has an adjusted \(R^2\) of
-`r prettyNum(snm$adj.r.squared, digits = 3L)` and an AIC of
-`r prettyNum(AIC(nullModel), digits = 2L, big.mark = ',')`.
+The null model has an \(R^2\) of `r nmr2`, an adjusted \(R^2\) of `r nmar2` and
+an AIC of `r nmAIC`.
 
@@
 ```
-The saturated model has an adjusted \(R^2\) of
-`r prettyNum(ssm$adj.r.squared, digits = 3L)` and an AIC of
-`r prettyNum(AIC(saturatedModel), digits = 2L, big.mark = ',')`.
+The saturated model has an \(R^2\) of `r smr2`, an adjusted \(R^2\) of `r smar2`
+and an AIC of `r smAIC`.
 
@@
 
-The step model allowing pairwise interactions between the four classes of
-metric: base running, batting, fielding, and pitching, has an adjusted \(R^2\)
-of `r prettyNum(sstm$adj.r.squared, digits = 3L)` and an AIC of
-`r prettyNum(AIC(stepModel), digits = 2L, big.mark = ',')`.
+The step model allowing pairwise interactions within the four classes of
+metric: base running, batting, fielding, and pitching, has an \(R^2\) of
+`r stmr2`, an adjusted \(R^2\) of `r stmar2` and an AIC of `r stmAIC`.
+
+### Discussion
+The first item for discussion is the observation that the intercept does **not**
+change between the null (intercept-only) model and the saturated (all variables
+included) model. There are 162 games in a baseball season. According to both the
+null and saturated models, given no other information, a team will *lose*
+slightly more than half its games. Whereas in the stepwise model allowing for
+pairwise interactions within classes, given no other information, a team will
+*win* slightly more than half of its games. Apparently, inteactions unearth
+somewhat more information about why teams **lose** rather than why they **win**.
+
+The stepwise model has the best \(R^2\) of the three and the best AIC. Using
+accepted rules of thumb, a difference of more than 10 indicates no support for
+the model with the higher AIC. The magnitude is irrelevant; it is a function of
+the number of observations. The difference between the AIC of the saturated
+model and the step model is
+`r prettyNum(AIC(saturatedModel) - AIC(stepModel), digits = 2L)`. This shows
+that the saturated model, and obviously the null model, should be completely
+ignored in favor of the stepwise model with pairwise within class interactions.
+
+Most of the coefficients are reasonable within the context of the game of
+baseball. These two statements are axiomatic:
+
+  * The only way to win is to have the higher score at the end of the game.
+  * The only way to score is for a baserunner to cross home plate.
+
+With those in mind, we can make the following observations about the linear
+non-interactive coefficients.
+
+ * Reasonable
+   * Caught stealing removes baserunners; negative coefficient makes sense.
+   * Stolen bases get a runner closer to home plate; positive coefficient makes
+   sense.
+   * Triples get a runner very close to home plate; positive coefficient makes
+   sense.
+   * Walks are free baserunners; positive coefficient makes sense.
+   * Hits increase the number of baserunners; positive coefficient makes sense.
+   * Hitting home runs directly increase the score; positive coefficient makes
+   sense.
+   * Striking out reduces the number of baserunners; negative coefficient makes
+   sense.
+   * Errors allow the other team free baserunners; negative coefficient makes
+   sense.
+   * Giving up home runs gives the opponent scores; negative coefficient makes
+   sense.
+   * Getting strikeouts reduces the opponents baserunners; positive coefficient
+   makes sense.
+ * Curious
+   * Hitting doubles get runners on base; why negative?
+   * Turning double plays reduce baserunners; why negative?
+   * Giving up walks allows the other team free baserunners; its coefficient is
+   both insignificant and near 0.
+   * Giving up hits allows the other team baserunners; its coefficient is
+   both insignificant and near 0.
+
+An absolutely fascinating observation. In the first draft of this exercise,
+`TEAM_BATTING_HR` was removed due to its high correlation with
+`TEAM_PITCHING_HR`. Note the coefficients, batting is +10 and pitching is -7.
+However, this is such as strong indicator, that in the first run, pitching was
+given a *postive* coefficient of around +3. In hindsight this is because
+*it was being used as an indicator for* ***batting!!*** The correlation allowed
+the use of pitching HRs as an indicator for the hidden batting HRs! Once both
+were restored to the model, the logical coefficients surfaced. Another reason
+why models should not be trusted out of the box, but all model results should
+be reviewed for sanity and sense!!
+
+The curiosities above can *possibly* be resolved by looking at the interaction
+terms. When giving up walks is interacted with strikeouts,
+`TEAM_PITCHING_BB:TEAM_PITCHING_SO`, the coefficent is positive and significant.
+Perhaps pitchers with many strikeouts also walk a lot of batters so giving up
+walks on its own is less relevant. Similarly, the interaction between giving
+up hits and strikeouts, `TEAM_PITCHING_H:TEAM_PITCHING_SO `, is highly negative
+and significant. Once again, it probably captures most of the giving up hits
+information making the singleton less relevant.
+
+A possible explanation for the negative coefficent for getting double plays, is
+that double plays require at least two people on base. That means that the
+oppenent has a lot of base runners, which is very highly correlated with
+scoring.
+
+The behavior of doubles remains confusing. Unless its hiding something like
+a team's propensity to strand runners on base. It would be interesting to see
+a breakdown between the American and National leagues, as the latter tends to be
+somewhat better at "small ball" and moving the runners along.
+
+# Build Models
+## Task
+Decide on the criteria for selecting the best multiple linear regression model.
+Will you select a model with slightly worse performance if it makes more sense
+or is more parsimonious? Discuss why you selected your model. For the multiple
+linear regression model, will you use a metric such as Adjusted \(R^2\), RMSE,
+etc.? Be sure to explain how you can make inferences from the model,
+discuss multi-collinearity issues (if any), and discuss other relevant model
+output.
+
+Using the training data set, evaluate the multiple linear regression model based
+on (a) mean squared error, (b) \(R^2\), (c) F-statistic, and (d) residual plots.
+Make predictions using the evaluation data set.
+
+## Analysis
+At this point, to properly fulfill this portion, cross-validation should be used
+in model selection. That is not possible here given the model selection above.
+That being said, given that the purpose of this exercise is **not** to
+demonstrate facility with tuning, the three models developed above will be used
+as the representatives.
+
+### Criteria
+As a devout follower of Burnham & Anderson (2002), I prefer to use AIC whenever
+possible. This includes inherent recognition of parsimony. As discussed in the
+previous section, for an AIC difference of 10 or greater, the model with the
+higher AIC has effectively no support. The stepwise model with within-class
+pairwise interactions had the lowest AIC by hundreds.
 
@@
 
+ * Burnham, K.P. and Anderson, D.R. (2002) "Model Selection and Inference: A
+ Practical Information-Theoretic Approach". 2nd Edition, Springer-Verlag,
+ New York. http://dx.doi.org/10.1007/b97636
  * Friendly, Michael (2002) Corrgrams: Exploratory displays for correlation

### 2020-09-09 HW1/RegModel_Adler.Rmd
Refactor to build a 75/25 train/test split and proceed

@@
 
+### Training and Test Sets
+As we are performing this analysis manually, we will break up the supplied data
+into two sets, one for training and one for testing. This needs to be done
+before any normalisation or imputation, otherwise the testing set is polluted
+with information from the training set. Please note that obervations prior to
+this point were on the entirety of the data set.
+
+```{r testTrain}
+set.seed(811247)
+trainObs <- sample(x = n, size = 0.75 * n, replace = FALSE)
+MBtrn <- MB[trainObs]
+MBtst <- MB[!trainObs]
+MBtrnlng <- melt(MBtrn, id.vars = 'INDEX', variable.name = 'Metric',
+                 value.name = 'Value', variable.factor = FALSE)
+MBtstlng <- melt(MBtst, id.vars = 'INDEX', variable.name = 'Metric',
+                 value.name = 'Value', variable.factor = FALSE)
+
+# Casting to double since medians and quartiles may return non-integer values.
+MBtrnlng[, Value := as.double(Value)]
+MBtstlng[, Value := as.double(Value)]
+```
+
 ### Imputation
@@
 ```{r imputeMean}
-MBmelt[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
+MBtrnlng[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
        keyby = Metric]
@@
 ```{r centerScale}
-MBmeltSC <- copy(MBmelt)
-MBmeltUnscale <- MBmeltSC[, .(Mean = mean(Value), SD = sd(Value)),
+MBtrnlngSC <- copy(MBtrnlng)
+MBtrnlngUnscale <- MBtrnlngSC[, .(Mean = mean(Value), SD = sd(Value)),
                               keyby = Metric]
-MBmeltSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
-knitr::kable(MBmeltUnscale, digits = 2L, caption = 'Data Means and SDs')
+MBtrnlngSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+knitr::kable(MBtrnlngUnscale, digits = 2L, caption = 'Data Means and SDs')
 ```
 ```{r plotScaled, fig.width=7}
-ggplot(MBmelt, aes(x = Value)) + geom_density() +
+ggplot(MBtrnlng, aes(x = Value)) + geom_density() +
   facet_wrap(~ Metric, scales = 'free') +
   ggtitle(label = 'Kernel-Smoothed Densities of Raw Data')
-ggplot(MBmeltSC, aes(x = Value)) + geom_density() +
+ggplot(MBtrnlngSC, aes(x = Value)) + geom_density() +
   facet_wrap(~ Metric, scales = 'free') +
@@
                              'TEAM_BASERUN_SB', 'TEAM_PITCHING_HR',
-                             'TEAM_PITCHING_BB', 'TEAM_PITCHING_SO',
-                             'TEAM_BATTING_3B'),
-                  Lambda = c(-2.5, -0.5, 0, 0.6, -0.15, -0.35, 0))
+                             'TEAM_BATTING_HR', 'TEAM_PITCHING_BB',
+                             'TEAM_PITCHING_SO', 'TEAM_BATTING_3B',
+                             'TEAM_BASERUN_CS'),
+                  Lambda = c(-2.5, -0.7, 0, 0.6, 0.6, -0.1, -0.35, 0, 0.5))
 setkey(BCL, Metric)
@@
 
-# Left join MBmelt to BCL. 
-MBmelt <- BCL[MBmelt]
+# Left join MBtrnlng to BCL. 
+MBtrnlng <- BCL[MBtrnlng]
 
 # Transform data where there are selected lambdas
-MBmelt[!is.na(Lambda), Value := BxCox(Value, Lambda)]
+MBtrnlng[!is.na(Lambda), Value := BxCox(Value, Lambda)]
 
 # Remove Lambda
-MBmelt[, Lambda := NULL]
+MBtrnlng[, Lambda := NULL]
 
 # Plot results
-ggplot(MBmelt, aes(x = Value)) + geom_density() +
+ggplot(MBtrnlng, aes(x = Value)) + geom_density() +
   facet_wrap(~ Metric, scales = 'free') +
@@
 
-Most of the variables are now closer to a Gaussian distribution. Also note we
-have removed `TEAM_BATTING_HR` from the data. Now the data needs to be rescaled
-and recentered.
+Most of the variables are now closer to a Gaussian distribution. Now the data
+needs to be rescaled and recentered.
 
 ```{r reScale}
-MBmeltSC <- copy(MBmelt)
-MBmeltUnscale <- MBmeltSC[, .(Mean = mean(Value), SD = sd(Value)),
+MBtrnlngSC <- copy(MBtrnlng)
+MBtrnlngUnscale <- MBtrnlngSC[, .(Mean = mean(Value), SD = sd(Value)),
                           keyby = Metric]
 
-MBmeltSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
-knitr::kable(MBmeltUnscale, digits = 2L,
+MBtrnlngSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+knitr::kable(MBtrnlngUnscale, digits = 2L,
              caption = 'Post-Transform Data Means and SDs')
@@
 ```{r reImputeThenScale}
-MBmeltSC <- copy(MBmelt)
-MBmeltSC[, Lambda := NULL]
+MBtrnlngSC <- copy(MBtrnlng)
+
 # Remove infinities
-MBmeltSC[, Value := ifelse(is.infinite(Value), NA, Value)]
+MBtrnlngSC[, Value := ifelse(is.infinite(Value), NA, Value)]
 # Center and Scale
-MBmeltSC[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
+MBtrnlngSC[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
        keyby = Metric]
-MBmeltUnscale <- MBmeltSC[, .(Mean = mean(Value), SD = sd(Value)),
+MBtrnlngUnscale <- MBtrnlngSC[, .(Mean = mean(Value), SD = sd(Value)),
                           keyby = Metric]
-MBmeltSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
-knitr::kable(MBmeltUnscale, digits = 2L,
+MBtrnlngSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+knitr::kable(MBtrnlngUnscale, digits = 2L,
              caption = 'Final Post-Transform Data Means and SDs')
-ggplot(MBmeltSC, aes(x = Value)) + geom_density() +
+ggplot(MBtrnlngSC, aes(x = Value)) + geom_density() +
   facet_wrap(~ Metric, scales = 'free') +
@@
 ## Analysis
-Normally, we would use a statistical packahe like `caret`, which not only
-simplifies model construction but also takes care of cross-validation. For the
-purposes of this homework, and to demonstrate understanding of the underlying
-decisions involved in predictive modeling, we will use calls to base R for the
-null and saturated models and the `stepAIC` function from the `MASS` package to
-perform a stepwise linear regression using AIC as our goodness of fit metric.
-In the final section, we may pivot to using `caret` for final model selection.
+Normally, we would use a statistical package like `caret`, which not only
+simplifies model construction but also takes of splitting into training and
+test sets and takes care of cross-validation. For the purposes of this homework,
+and to demonstrate understanding of the underlying decisions involved in
+predictive modeling, we will use calls to base R and the `MASS` package.
+
+As the transforms were imposed on the entirety of the training data, we ar
+
+A null model is once where there are no explanatory variables other than the
+grand mean. A saturated model is one where every variable appears.
 
@@
 ### Linear Models
-We will show three models: the null model, the saturated model, and the model
-resulting from the stepwise regression. For the regression, we will consider at
+We will show three models:
+
+ 1. The null model: This is the model with only the intercept, the grand mean.
+ 2. The saturated model: this is the model with every term included
+ 3. A model allowing pairwise interactions within the data types of baserunning,
+ batting, pitching, and fielding, fit using a forward and backwards stepwise
+ regression, and the model
+resulting from the stepwise regression. A nuFor the regression, we will consider at
 most pairwise interactions within the classes of metrics.
@@
 ```{r lm, include=FALSE}
-MBsc <- dcast(MBmeltSC, INDEX ~ Metric, value.var = 'Value')
-# Replace TARGET_WINS with original. Could have restored it from MBmeltUnscale.
-MBsc[, TARGET_WINS := MB[, TARGET_WINS]]
+MBtrnSC <- dcast(MBtrnlngSC, INDEX ~ Metric, value.var = 'Value')
+
+# Rscale TARGET_WINS to its standard value.
+MBtrnSC[, TARGET_WINS := TARGET_WINS * 
+          MBtrnlngUnscale[Metric == 'TARGET_WINS', SD] +
+          MBtrnlngUnscale[Metric == 'TARGET_WINS', Mean]]
+
 # Removing INDEX for convenience. Could have used "-INDEX" in the formula call.
-MBsc[, INDEX := NULL]
+MBtrnSC[, INDEX := NULL]
 
-nullModel <- lm(TARGET_WINS ~ 1, data = MBsc)
+nullModel <- lm(TARGET_WINS ~ 1, data = MBtrnSC)
 snm <- summary(nullModel)
@@
 nmAIC <- prettyNum(AIC(nullModel), digits = 2L, big.mark = ',')
+nmRMSE <- prettyNum(sqrt(mean(nullModel$residuals ^ 2)), digits = 5L)
 
-saturatedModel <- lm(TARGET_WINS ~ ., data = MBsc)
+saturatedModel <- lm(TARGET_WINS ~ ., data = MBtrnSC)
 ssm <- summary(saturatedModel)
@@
 smAIC <- prettyNum(AIC(saturatedModel), digits = 2L, big.mark = ',')
+smRMSE <- prettyNum(sqrt(mean(saturatedModel$residuals ^ 2)), digits = 5L)
+smF <- prettyNum(ssm$fstatistic, digits = 5L)
+smFp <- pf(ssm$fstatistic[[1]], ssm$fstatistic[[2]], ssm$fstatistic[[3]],
+            lower.tail = FALSE)
 
-stepModel <- stepAIC(saturatedModel, scope = list(upper = ~
+crazyModel <- lm(TARGET_WINS ~
+  (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
+  TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_SO) ^ 2 +
+  (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
+  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = MBtrnSC)
+
+stepModel <- stepAIC(crazyModel, scope = list(upper = ~
   (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
@@
 stmAIC <- prettyNum(AIC(stepModel), digits = 2L, big.mark = ',')
-
+stmRMSE <- prettyNum(sqrt(mean(stepModel$residuals ^ 2)), digits = 5L)
+stmF <- prettyNum(sstm$fstatistic, digits = 5L)
+stmFp <- pf(sstm$fstatistic[[1]], sstm$fstatistic[[2]], sstm$fstatistic[[3]],
+            lower.tail = FALSE)
 ```
@@
 ```
-The null model has an \(R^2\) of `r nmr2`, an adjusted \(R^2\) of `r nmar2` and
-an AIC of `r nmAIC`.
+The null model has an \(R^2\) of `r nmr2`, an adjusted \(R^2\) of `r nmar2`, an
+AIC of `r nmAIC`, and an RMSE on the training set of `r nmRMSE`. There is no
+F-statistic for the entirety of the null model as there is only one axis for
+degrees of freedom.
 
@@
 ```
-The saturated model has an \(R^2\) of `r smr2`, an adjusted \(R^2\) of `r smar2`
-and an AIC of `r smAIC`.
+The saturated model has an \(R^2\) of `r smr2`, an adjusted \(R^2\) of
+`r smar2`, an AIC of `r smAIC`, and an RMSE on the training set of `r smRMSE`.
+The p-value for the F-test with a critical value of `r smF[[1]]` with
+`r smF[[2]]` and `r smF[[3]]` degrees of freedom is below machine error
+(`r smFp`).
 
@@
 metric: base running, batting, fielding, and pitching, has an \(R^2\) of
-`r stmr2`, an adjusted \(R^2\) of `r stmar2` and an AIC of `r stmAIC`.
+`r stmr2`, an adjusted \(R^2\) of `r stmar2`, an AIC of `r stmAIC`, and an RMSE
+on the training set of `r stmRMSE`. The p-value for the F-test with a critical
+value of `r stmF[[1]]` with `r stmF[[2]]` and `r stmF[[3]]` degrees of freedom
+is below machine error (`r stmFp`).
 
 ### Discussion
-The first item for discussion is the observation that the intercept does **not**
-change between the null (intercept-only) model and the saturated (all variables
-included) model. There are 162 games in a baseball season. According to both the
-null and saturated models, given no other information, a team will *lose*
-slightly more than half its games. Whereas in the stepwise model allowing for
-pairwise interactions within classes, given no other information, a team will
-*win* slightly more than half of its games. Apparently, inteactions unearth
-somewhat more information about why teams **lose** rather than why they **win**.
+The first item for discussion is the observation that while the intercept does
+not change between the null model and the saturated model, it does decrease for
+the model allowing within class pairwise interactions. There are 162 games in a
+baseball season. According to all the models, given no other information, a team
+will lose slightly more than half its games. Apparently, interctions in the
+stepwise model unearth somewhat more information about why teams win.
 
@@
    * Turning double plays reduce baserunners; why negative?
-   * Giving up walks allows the other team free baserunners; its coefficient is
-   both insignificant and near 0.
-   * Giving up hits allows the other team baserunners; its coefficient is
-   both insignificant and near 0.
+   * Giving up walks allows the other team free baserunners; negative
+   coefficient makes sense but why insignificant?
+   * Giving up hits allows the other team baserunners; why positive athough
+   insignificant?
 
@@
 `TEAM_BATTING_HR` was removed due to its high correlation with
-`TEAM_PITCHING_HR`. Note the coefficients, batting is +10 and pitching is -7.
-However, this is such as strong indicator, that in the first run, pitching was
-given a *postive* coefficient of around +3. In hindsight this is because
+`TEAM_PITCHING_HR`. Note the coefficients, batting is almost +12 and pitching is
+-9. However, this is such as strong indicator, that in the first run, pitching
+was given a *postive* coefficient of around +3. In hindsight this is because
 *it was being used as an indicator for* ***batting!!*** The correlation allowed
@@
 The curiosities above can *possibly* be resolved by looking at the interaction
-terms. When giving up walks is interacted with strikeouts,
-`TEAM_PITCHING_BB:TEAM_PITCHING_SO`, the coefficent is positive and significant.
-Perhaps pitchers with many strikeouts also walk a lot of batters so giving up
-walks on its own is less relevant. Similarly, the interaction between giving
-up hits and strikeouts, `TEAM_PITCHING_H:TEAM_PITCHING_SO `, is highly negative
-and significant. Once again, it probably captures most of the giving up hits
-information making the singleton less relevant.
+terms. The interaction between giving up hits and strikeouts,
+`TEAM_PITCHING_H:TEAM_PITCHING_SO `, is highly negative and significant. It
+probably captures most of the giving up hits information making the singleton
+less relevant.
 
@@
 that double plays require at least two people on base. That means that the
-oppenent has a lot of base runners, which is very highly correlated with
+opponent has a lot of base runners, which is very highly correlated with
 scoring.
@@
 pairwise interactions had the lowest AIC by hundreds.
+
+General inferences about the nature of the coefficients is in the previous
+section. The fascinating example of how multi-collinearity can override logic is
+also discussed in the previous secton.
+
+In order to make inferences from the model, the incoming data needs to be
+processed in the exact same way:
+
+ 1. Mean imputation
+ 2. Box-Cox with the selected lambdas
+ 3. Mean imputation again to remove `Inf` issues
+ 4. Centering and scaling
+
+### Model Evaluations
+In the table below are the RMSE, \(R^2\), and significance of the F-statistic
+for each of the three models as applied to the **testing** set.
+
+```{r buildTests}
+MBtstlng <- BCL[MBtstlng]
+MBtstlng[!is.na(Lambda), Value := BxCox(Value, Lambda)]
+MBtstlng[, Lambda := NULL]
+MBtstlngSC <- copy(MBtstlng)
+
+# Remove infinities
+MBtstlngSC[, Value := ifelse(is.infinite(Value), NA, Value)]
+
+# Center and Scale
+MBtstlngSC[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
+       keyby = Metric]
+MBtstlngUnscale <- MBtstlngSC[, .(Mean = mean(Value), SD = sd(Value)),
+                          keyby = Metric]
+MBtstlngSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+MBtstSC <- dcast(MBtstlngSC, INDEX ~ Metric, value.var = 'Value')
+MBtstSC[, TARGET_WINS := TARGET_WINS * 
+          MBtstlngUnscale[Metric == 'TARGET_WINS', SD] +
+          MBtstlngUnscale[Metric == 'TARGET_WINS', Mean]]
+```
 

### 2020-09-09 HW1/RegModel_Adler.Rmd
Finish test set. Up to predictions.

@@
 # Left join MBtrnlng to BCL. 
-MBtrnlng <- BCL[MBtrnlng]
+MBtrnlng <- BCL[MBtrnlng, on = 'Metric']
 
@@
 
-The stepwise model has the best \(R^2\) of the three and the best AIC. Using
-accepted rules of thumb, a difference of more than 10 indicates no support for
-the model with the higher AIC. The magnitude is irrelevant; it is a function of
-the number of observations. The difference between the AIC of the saturated
-model and the step model is
+#### Statistics
+The stepwise model has the best \(R^2\), adjusted  \(R^2\), and AIC of the three
+models. Using accepted rules of thumb, a difference of more than 10 indicates no
+support for the model with the higher AIC. The magnitude is irrelevant; it is a
+function of the number of observations. The difference between the AIC of the
+saturated model and the step model is
 `r prettyNum(AIC(saturatedModel) - AIC(stepModel), digits = 2L)`. This shows
@@
 
+#### Residual Plots
+```{r residPlots, fig.width=7}
+par(mfrow = c(2, 2))
+plot(saturatedModel$fitted.values, studres(saturatedModel),
+     main = 'Saturated Model', xlab = 'Predicted Values',
+     ylab = 'Studentized Residuals', xlim = c(0, 160), ylim = c(-6, 6))
+plot(stepModel$fitted.values, studres(stepModel),
+     main = 'Stepwise Pairwise Model', xlab = 'Predicted Values',
+     ylab = 'Studentized Residuals', xlim = c(0, 160), ylim = c(-6, 6))
+plot(nullModel$fitted.values, studres(nullModel),
+     main = 'Null Model', xlab = 'Predicted Values',
+     ylab = 'Studentized Residuals', xlim = c(0, 160), ylim = c(-6, 6))
+par(mfrow = c(1, 1))
+```
+
+Both the saturated and the stepwise models look to have a basically normal
+distribution of their residuals. The null model returns a single value,
+explaining the bizarness of the graph.
+
+The **actual** model to be used for the prediction will be the model which
+returns the best RMSE and MAPE on the test set.
+
+#### Explanation of Selected Model
 Most of the coefficients are reasonable within the context of the game of
@@
 higher AIC has effectively no support. The stepwise model with within-class
-pairwise interactions had the lowest AIC by hundreds.
+pairwise interactions had the lowest AIC by hundreds. The various statistics and
+plots requested for the training data are also in the previous section.
 
@@
 ### Model Evaluations
-In the table below are the RMSE, \(R^2\), and significance of the F-statistic
-for each of the three models as applied to the **testing** set.
+In the table below are the RMSE and MAPE of each of the three models developed
+above when applied to the **testing** set.
 
-```{r buildTests}
-MBtstlng <- BCL[MBtstlng]
+```{r buildTestSet}
+MBtstlng <- BCL[MBtstlng, on = 'Metric']
 MBtstlng[!is.na(Lambda), Value := BxCox(Value, Lambda)]
@@
           MBtstlngUnscale[Metric == 'TARGET_WINS', Mean]]
+nmP <- predict(nullModel, MBtstSC)
+smP <- predict(saturatedModel, MBtstSC)
+stmP <- predict(stepModel, MBtstSC)
+nmRMSE <- sqrt(mean((MBtstSC$TARGET_WINS - nmP) ^ 2))
+nmMAPE <- mean(abs((MBtstSC$TARGET_WINS - nmP) / MBtstSC$TARGET_WINS))
+smRMSE <- sqrt(mean((MBtstSC$TARGET_WINS - smP) ^ 2))
+smMAPE <- mean(abs((MBtstSC$TARGET_WINS - smP) / MBtstSC$TARGET_WINS))
+stmRMSE <- sqrt(mean((MBtstSC$TARGET_WINS - stmP) ^ 2))
+stmMAPE <- mean(abs((MBtstSC$TARGET_WINS - stmP) / MBtstSC$TARGET_WINS))
+knitr::kable(data.frame(Model = c("Null", "Saturated", "Stepwise Pairwise"),
+                        RMSE = c(nmRMSE, smRMSE, stmRMSE),
+                        MAPE = c(nmMAPE, smMAPE, stmMAPE)),
+             digits = 3L, caption = 'Test Set Results')
 ```
+
+The stepwise model allowing for pairwise interactions is the winner in both
+cases, as was expected.
+
+### Predictions
+We will use the stepwise-pairwise model to predict `TARGET_WINS` for the
+separately provided evaluation set.
 
@@
 ```
+```{r testTrain, echo=TRUE, eval=FALSE}
+```
 ```{r imputeMean, echo=TRUE, eval=FALSE}
@@
 ```
+```{r residPlots, echo=TRUE, eval=FALSE}
+```
+```{r buildTestSet, echo=TRUE, eval=FALSE}
+```

### 2020-09-09 HW1/RegModel_Adler.Rmd
Looks complete for now

@@
 We will use the stepwise-pairwise model to predict `TARGET_WINS` for the
-separately provided evaluation set.
+separately provided evaluation set by following the same preprocessing used for
+the training and testing sets.
+
+```{r predictions}
+MBT <- fread('./data/moneyball-evaluation-data.csv')
+MBT[, `:=`(TEAM_BATTING_HBP = NULL)]
+MBTlng <- melt(MBT, id.vars = 'INDEX', variable.name = 'Metric',
+               value.name = 'Value', variable.factor = FALSE)
+MBTlng[, Value := as.double(Value)]
+MBTlng <- BCL[MBTlng, on = 'Metric']
+MBTlng[!is.na(Lambda), Value := BxCox(Value, Lambda)]
+MBTlng[, Lambda := NULL]
+MBTlngSC <- copy(MBTlng)
+
+# Remove infinities
+MBTlngSC[, Value := ifelse(is.infinite(Value), NA, Value)]
+
+# Impute, Center, and Scale
+MBTlngSC[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
+       keyby = Metric]
+MBTlngSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+MBTSC <- dcast(MBTlngSC, INDEX ~ Metric, value.var = 'Value')
+predictions <- predict(stepModel, MBTSC)
+predictions
+```
 
@@
 ```
+```{r predictions, echo=TRUE, eval=FALSE}
+```

### 2020-09-09 HW1/missing_value_eda.Rmd

@@
 train_df<- 
-  read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv?token=AJP53VTYKQ4BKNZQERSMBYK7KBD52")
+  read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv?token=AJP53VUTHKQNVGSBWE3U7KC7KDWP4")
 ```

### 2020-09-09 HW1/RegModel_Adler.Rmd
Copyedit and add R2/aR2 to test set and table to training

@@
 
-# Meltng data from wide to narrow for munging & ggplot
+# Melting data from wide to narrow for munging & ggplot
 MBmelt <- melt(MB, id.vars = 'INDEX', variable.name = 'Metric',
@@
 
-# Use the Freedman-Diaconis rule for historgram bin widths instead of ggplot
+# Use the Freedman-Diaconis rule for histogram bin widths instead of ggplot
 # defaults
@@
 into two sets, one for training and one for testing. This needs to be done
-before any normalisation or imputation, otherwise the testing set is polluted
-with information from the training set. Please note that obervations prior to
+before any normalization or imputation, otherwise the testing set is polluted
+with information from the training set. Please note that observations prior to
 this point were on the entirety of the data set.
@@
 MBtst <- MB[!trainObs]
+ntrn <- dim(MBtrn)[[1]]
+ntst <- dim(MBtst)[[1]]
 MBtrnlng <- melt(MBtrn, id.vars = 'INDEX', variable.name = 'Metric',
@@
 Normally, we would use a statistical package like `caret`, which not only
-simplifies model construction but also takes of splitting into training and
-test sets and takes care of cross-validation. For the purposes of this homework,
-and to demonstrate understanding of the underlying decisions involved in
-predictive modeling, we will use calls to base R and the `MASS` package.
-
-As the transforms were imposed on the entirety of the training data, we ar
+simplifies model construction but also takes care of cross-validation. For the
+purposes of this homework, and to demonstrate understanding of the underlying
+decisions involved in predictive modeling, we will use calls to base R and the
+`MASS` package.
 
@@
  1. The null model: This is the model with only the intercept, the grand mean.
- 2. The saturated model: this is the model with every term included
+ 2. The saturated model: this is the model with every term included.
  3. A model allowing pairwise interactions within the data types of baserunning,
  batting, pitching, and fielding, fit using a forward and backwards stepwise
- regression, and the model
-resulting from the stepwise regression. A nuFor the regression, we will consider at
-most pairwise interactions within the classes of metrics.
+ regression.
 
@@
 
-# Rscale TARGET_WINS to its standard value.
+# Rescale TARGET_WINS to its standard value.
 MBtrnSC[, TARGET_WINS := TARGET_WINS * 
@@
 baseball season. According to all the models, given no other information, a team
-will lose slightly more than half its games. Apparently, interctions in the
+will lose slightly more than half its games. Apparently, interactions in the
 stepwise model unearth somewhat more information about why teams win.
@@
 
+```{r trainStatTable}
+knitr::kable(data.frame(Model = c("Null", "Saturated", "Stepwise Pairwise"),
+                        Rsq = c(snm$r.squared, ssm$r.squared, sstm$r.squared),
+                        AdjRsq = c(snm$adj.r.squared, ssm$adj.r.squared,
+                                   ssm$adj.r.squared),
+                        RMSE = c(sqrt(mean(nullModel$residuals ^ 2)),
+                                 sqrt(mean(saturatedModel$residuals ^ 2)),
+                                 sqrt(mean(stepModel$residuals ^ 2)))),
+             digits = 3L, caption = 'Train Set Results')
+```
+
 #### Residual Plots
-```{r residPlots, fig.width=7}
+```{r residPlots, fig.height=7}
 par(mfrow = c(2, 2))
@@
 distribution of their residuals. The null model returns a single value,
-explaining the bizarness of the graph.
+explaining the bizarreness of the graph.
 
 The **actual** model to be used for the prediction will be the model which
-returns the best RMSE and MAPE on the test set.
+returns the best RMSE on the test set. Unfortunately, MAPE cannot be used as
+there are observations which are 0.
 
@@
    coefficient makes sense but why insignificant?
-   * Giving up hits allows the other team baserunners; why positive athough
+   * Giving up hits allows the other team baserunners; why positive although
    insignificant?
@@
 -9. However, this is such as strong indicator, that in the first run, pitching
-was given a *postive* coefficient of around +3. In hindsight this is because
+was given a *positive* coefficient of around +3. In hindsight this is because
 *it was being used as an indicator for* ***batting!!*** The correlation allowed
@@
 
-A possible explanation for the negative coefficent for getting double plays, is
+A possible explanation for the negative coefficient for getting double plays, is
 that double plays require at least two people on base. That means that the
@@
 section. The fascinating example of how multi-collinearity can override logic is
-also discussed in the previous secton.
+also discussed in the previous section.
 
@@
 nmP <- predict(nullModel, MBtstSC)
+nmPR <- (MBtstSC$TARGET_WINS - nmP)
 smP <- predict(saturatedModel, MBtstSC)
+smPR <- (MBtstSC$TARGET_WINS - smP)
 stmP <- predict(stepModel, MBtstSC)
-nmRMSE <- sqrt(mean((MBtstSC$TARGET_WINS - nmP) ^ 2))
-nmMAPE <- mean(abs((MBtstSC$TARGET_WINS - nmP) / MBtstSC$TARGET_WINS))
-smRMSE <- sqrt(mean((MBtstSC$TARGET_WINS - smP) ^ 2))
-smMAPE <- mean(abs((MBtstSC$TARGET_WINS - smP) / MBtstSC$TARGET_WINS))
-stmRMSE <- sqrt(mean((MBtstSC$TARGET_WINS - stmP) ^ 2))
-stmMAPE <- mean(abs((MBtstSC$TARGET_WINS - stmP) / MBtstSC$TARGET_WINS))
+stmPR <- (MBtstSC$TARGET_WINS - stmP)
+SST <- sum((MBtstSC$TARGET_WINS - mean(MBtstSC$TARGET_WINS)) ^ 2)
+nmPR2 <- 1 - sum(nmPR ^ 2) / SST
+nmPaR2 <- 1 - (1 - nmPR2) * (snm$df[[1]] + snm$df[[2]] - 1) / (snm$df[[2]])
+nmPRMSE <- sqrt(mean(nmPR ^ 2))
+
+smPR2 <- 1 - sum(smPR ^ 2) / SST
+smPaR2 <- 1 - (1 - smPR2) * (ssm$df[[1]] + ssm$df[[2]] - 1) / (ssm$df[[2]])
+smPRMSE <- sqrt(mean(smPR ^ 2))
+
+stmPR2 <- 1 - sum(stmPR ^ 2) / SST
+stmPaR2 <- 1 - (1 - stmPR2) * (sstm$df[[1]] + sstm$df[[2]] - 1) / (sstm$df[[2]])
+stmPRMSE <- sqrt(mean(stmPR ^ 2))
 knitr::kable(data.frame(Model = c("Null", "Saturated", "Stepwise Pairwise"),
-                        RMSE = c(nmRMSE, smRMSE, stmRMSE),
-                        MAPE = c(nmMAPE, smMAPE, stmMAPE)),
+                        Rsq = c(nmPR2, smPR2, stmPR2),
+                        AdjRsq = c(nmPaR2, smPaR2, stmPaR2),
+                        RMSE = c(nmPRMSE, smPRMSE, stmPRMSE)),
              digits = 3L, caption = 'Test Set Results')
@@
 
-The stepwise model allowing for pairwise interactions is the winner in both
-cases, as was expected.
+The stepwise model allowing for pairwise interactions is the winner in all
+cases, as was expected. The metrics are all worse than those on the training
+set. This is a classic example of overfitting, which may have been helped by
+k-fold cross-validation.
 
-### Predictions
+Also, this is one of those rare times where the actual \(R^2\) can be negative.
+The null model prediction is a constant which is not equal to the grand mean of
+the testing set, so it actually **adds** variance.
+
+### Predictions on Evaluation Set
 We will use the stepwise-pairwise model to predict `TARGET_WINS` for the
@@
 
-# Appendix
+# Code Appendix
 The code chunks below represent the code called in order during the analysis.
@@
 ```
+```{r trainStatTable, echo=TRUE, eval=FALSE}
+```
 ```{r residPlots, echo=TRUE, eval=FALSE}

### 2020-09-09 HW1/RegModel_Adler.Rmd
Minor corrections

@@
                         AdjRsq = c(snm$adj.r.squared, ssm$adj.r.squared,
-                                   ssm$adj.r.squared),
+                                   sstm$adj.r.squared),
                         RMSE = c(sqrt(mean(nullModel$residuals ^ 2)),
@@
 #### Residual Plots
+
 ```{r residPlots, fig.height=7}
@@
 ### Model Evaluations
-In the table below are the RMSE and MAPE of each of the three models developed
-above when applied to the **testing** set.
+In the table below are the \(R^2\), adjusted \(R^2\), and RMSEfor each of the
+three models developed above when applied to the **testing** set.
 

### 2020-09-09 HW1/HW1_AF.Rmd
data exploration - initial

--- 2020-08-31 HW1/HW1.Rmd
@@
 ---
-title: 'DATA 621 – Business Analytics and Data Mining: Homework #1'
-author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, Bryan Persaud,
-  Jeff Shamp
+title: 'DATA 621---Business Analytics and Data Mining'
+subtitle: 'Fall 2020---Group 2---Homework #1'
+author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
+    Bryan Persaud, Jeff Shamp
 date: "9/27/2020"
@@
 
+```{r loadData, include=FALSE}
+library(tidyverse)
+library(ggplot2)
+library(corrplot)
+train_set <- read.csv('data/moneyball-training-data.csv')
+eval_set <- read.csv('data/moneyball-evaluation-data.csv')
+```
+
 ## 1. DATA EXPLORATION (Amber)
-Describe the size and the variables in the moneyball training data set.
-Consider that too much detail will cause a manager to lose interest while too little detail will make the manager consider that you aren’t doing your job.
-Some suggestions are given below.
-Please do NOT treat this as a check list of things to do to complete the assignment.
-You should have your own thoughts on what to tell the boss.
-These are just ideas.
+Describe the size and the variables in the moneyball training data set. Consider
+that too much detail will cause a manager to lose interest while too little
+detail will make the manager consider that you aren't doing your job. Some
+suggestions are given below. Please do NOT treat this as a check list of things
+to do to complete the assignment. You should have your own thoughts on what to
+tell the boss. These are just ideas.
 
@@
 
+We are given a data set containing 2,276 records, each of which includes the 
+number of wins the team had that year and 15 additional statistics about the 
+season. A first look at the data shows that only about 8% of the records have 
+a full set of data. The good news is that most of the missing values come from 
+what we would expect; hit-by-pitch, caught stealing, double plays.
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
+
+pct_null <- data.frame(do.call("rbind", map(train_set %>% select(-INDEX), ~mean(is.na(.)))))
+colnames(pct_null) <- c('PCT_NULL')
+
+pct_null %>%
+  arrange(desc(PCT_NULL)) %>%
+  filter(PCT_NULL > 0)
+
+```
+
+  * Total cases: 2276, Complete cases: 191 (only ~8% of the data has fully filled in values)
+  * 15 stats (variables), 1 target (wins)
+  * Best team won 146 out of 162 games, worst team won none, Average number of wins is 80 (show win dist)
+  * Discuss missing values
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
+
+# histogram of wins
+p <- ggplot(train_set, aes(x=TARGET_WINS)) + 
+  geom_histogram() + 
+  geom_vline(aes(xintercept=mean(TARGET_WINS)),
+            color="blue", linetype="dashed", size=1)+
+  labs(title="Distribution of Wins",x="Number of wins", y = "Number of teams")
+
+p
+
+# Correlation matrix with significance levels (p-value)
+library("Hmisc")
+res2 <- rcorr(as.matrix(train_set %>% select(-INDEX)))
+
+# Insignificant correlation are crossed
+corrplot(res2$r, type="upper", order="hclust", 
+          tl.col = "black", p.mat = res2$P, sig.level = 0.01, insig = "blank")
+
+train_set %>%
+  select(-INDEX) %>%
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE") %>%
+  ggplot(aes(x = VALUE, y = TARGET_WINS)) +
+    geom_point() +
+  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
+    facet_wrap(~ STAT, scales = "free")
+
+```
+
 ## 2. DATA PREPARATION (25 Points)
-Describe how you have transformed the data by changing the original variables or creating new variables.
-If you did transform the data or create new variables, discuss why you did this.
-Here are some possible transformations.
+Describe how you have transformed the data by changing the original variables or
+creating new variables. If you did transform the data or create new variables,
+discuss why you did this. Here are some possible transformations.
 
@@
   d. Mathematical transforms such as log or square root (or use Box-Cox)
-e. Combine variables (such as ratios or adding or multiplying) to create new variables
+  e. Combine variables (such as ratios or adding or multiplying) to create new
+  variables
 
 ## 3. BUILD MODELS (25 Points)
-Using the training data set, build at least three different multiple linear regression models, using different variables (or the same variables with different transformations).
-Since we have not yet covered automated variable selection methods, you should select the variables manually (unless you previously learned Forward or Stepwise selection, etc.).
-Since you manually selected a variable for inclusion into the model or exclusion into the model, indicate why this was done.
+Using the training data set, build at least three different multiple linear
+regression models, using different variables (or the same variables with
+different transformations). Since we have not yet covered automated variable
+selection methods, you should select the variables manually (unless you
+previously learned Forward or Stepwise selection, etc.). Since you manually
+selected a variable for inclusion into the model or exclusion into the model,
+indicate why this was done.
 
-Discuss the coefficients in the models, do they make sense?
-For example, if a team hits a lot of Home Runs, it would be reasonably expected that such a team would win more games.
-However, if the coefficient is negative (suggesting that the team would lose more games), then that needs to be discussed.
-Are you keeping the model even though it is counter intuitive?
-Why?
-The boss needs to know.
+Discuss the coefficients in the models, do they make sense? For example, if a
+team hits a lot of Home Runs, it would be reasonably expected that such a team
+would win more games. However, if the coefficient is negative (suggesting that
+the team would lose more games), then that needs to be discussed. Are you
+keeping the model even though it is counter intuitive? Why? The boss needs to
+know.
 
@@
 Decide on the criteria for selecting the best multiple linear regression model.
-Will you select a model with slightly worse performance if it makes more sense or is more parsimonious?
-Discuss why you selected your model.
+Will you select a model with slightly worse performance if it makes more sense
+or is more parsimonious? Discuss why you selected your model.
 
-For the multiple linear regression model, will you use a metric such as Adjusted R2, RMSE, etc.?
-Be sure to explain how you can make inferences from the model, discuss multi-collinearity issues (if any), and discuss other relevant model output.
-Using the training data set, evaluate the multiple linear regression model based on (a) mean squared error, (b) R2, (c) F-statistic, and (d) residual plots.
-Make predictions using the evaluation data set.
+For the multiple linear regression model, will you use a metric such as Adjusted
+\(R^2\), RMSE, etc.? Be sure to explain how you can make inferences from the
+model, discuss multi-collinearity issues (if any), and discuss other relevant
+model output. Using the training data set, evaluate the multiple linear
+regression model based on (a) mean squared error, (b) \(R^2\), (c) F-statistic,
+and (d) residual plots. Make predictions using the evaluation data set.
 
@@
 ### Assignment Overview
-In this homework assignment, you will explore, analyze and model a data set containing approximately 2200
-records. Each record represents a professional baseball team from the years 1871 to 2006 inclusive. Each record
-has the performance of the team for the given year, with all of the statistics adjusted to match the performance of
-a 162 game season.
+In this homework assignment, you will explore, analyze and model a data set
+containing approximately 2200 records. Each record represents a professional
+baseball team from the years 1871 to 2006 inclusive. Each record has the
+performance of the team for the given year, with all of the statistics adjusted
+to match the performance of a 162 game season.
 
-Your objective is to build a multiple linear regression model on the training data to predict the number of wins
-for the team. You can only use the variables given to you (or variables that you derive from the variables
-provided). Below is a short description of the variables of interest in the data set:
+Your objective is to build a multiple linear regression model on the training
+data to predict the number of wins for the team. You can only use the variables
+given to you (or variables that you derive from the variables provided). Below
+is a short description of the variables of interest in the data set:
 
@@
 
-```{r}
+```{r initial, warning = FALSE}
+########### libraries 
+library(tidyverse)
 
+########### data
 train_set <- read.csv('data/moneyball-training-data.csv')
@@
 
-colnames(train_set)
+```
 
+```{r dataExploration, results = 'hide'}
+# initial summary of all values 
+summary(train_set)
+p <- ggplot(train_set, aes(x=TARGET_WINS)) + 
+  geom_histogram() + 
+  geom_vline(aes(xintercept=mean(TARGET_WINS)),
+            color="blue", linetype="dashed", size=1)+
+  labs(title="Distribution of Wins",x="Number of wins", y = "Number of teams")
+
+# complete vs total cases
+train_set %>%
+  filter(complete.cases(.) == TRUE) %>%
+  count() %>%
+  rename("complete cases"= n) %>%
+  bind_cols(
+    train_set %>%
+    select(TARGET_WINS) %>%
+    count() %>%
+    rename("target count" = n)
+  )
 ```

### 2020-09-09 HW1/HW1_BP.rmd
Add files via upload

--- 2020-08-31 HW1/HW1.Rmd
@@
 ```{r}
+library(dplyr)
+library(reshape2)
+library(ggplot2)
+```
 
-train_set <- read.csv('data/moneyball-training-data.csv')
-eval_set <- read.csv('data/moneyball-evaluation-data.csv')
+```{r}
+
+train_set <- read.csv('/Users/Bryan/Documents/Data 621/HW 1/moneyball-training-data.csv')
+eval_set <- read.csv('/Users/Bryan/Documents/Data 621/HW 1/moneyball-evaluation-data.csv')
 
@@
 ```
+
+# Data Transformation
+
+```{r}
+train_set <- train_set %>%
+  filter(TARGET_WINS >= 20 & TARGET_WINS <= 125 & TEAM_BATTING_H <= 1880 & TEAM_BATTING_2B >= 115 & TEAM_BATTING_2B <= 380 & TEAM_BATTING_BB >= 300 & TEAM_BATTING_BB <= 870 & TEAM_BATTING_SO <= 1540 & TEAM_PITCHING_HR <= 255 & TEAM_PITCHING_SO <= 1450)
+train_set[train_set == 0] <- NA
+```
+
+```{r}
+sub = melt(train_set)
+ggplot(sub, aes(x = value)) + geom_density() + facet_wrap(~ variable, scales = 'free')
+```
+
+# Build Model
+
+```{r}
+model1 <- lm(TARGET_WINS ~ ., train_set)
+summary(model1)
+```
+
+```{r}
+model2 <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_SO + TEAM_PITCHING_HR, train_set)
+summary(model2)
+```

### 2020-09-09 HW1/HW1_AF.Rmd
data exploration updates

--- 2020-08-31 HW1/HW1.Rmd
@@
 
-## 1. DATA EXPLORATION (Amber)
-Describe the size and the variables in the moneyball training data set.
-Consider that too much detail will cause a manager to lose interest while too little detail will make the manager consider that you aren’t doing your job.
-Some suggestions are given below.
-Please do NOT treat this as a check list of things to do to complete the assignment.
-You should have your own thoughts on what to tell the boss.
-These are just ideas.
-
-a. Mean / Standard Deviation / Median
-b. Bar Chart or Box Plot of the data
-c. Is the data correlated to the target variable (or to other variables?)
-d. Are any of the variables missing and need to be imputed “fixed”?
-
-## 2. DATA PREPARATION (25 Points)
-Describe how you have transformed the data by changing the original variables or creating new variables.
-If you did transform the data or create new variables, discuss why you did this.
-Here are some possible transformations.
-
-a. Fix missing values (maybe with a Mean or Median value)
-b. Create flags to suggest if a variable was missing
-c. Transform data by putting it into buckets
-d. Mathematical transforms such as log or square root (or use Box-Cox)
-e. Combine variables (such as ratios or adding or multiplying) to create new variables
-
-## 3. BUILD MODELS (25 Points)
-Using the training data set, build at least three different multiple linear regression models, using different variables (or the same variables with different transformations).
-Since we have not yet covered automated variable selection methods, you should select the variables manually (unless you previously learned Forward or Stepwise selection, etc.).
-Since you manually selected a variable for inclusion into the model or exclusion into the model, indicate why this was done.
-
-Discuss the coefficients in the models, do they make sense?
-For example, if a team hits a lot of Home Runs, it would be reasonably expected that such a team would win more games.
-However, if the coefficient is negative (suggesting that the team would lose more games), then that needs to be discussed.
-Are you keeping the model even though it is counter intuitive?
-Why?
-The boss needs to know.
-
-## 4. SELECT MODELS (25 Points)
-Decide on the criteria for selecting the best multiple linear regression model.
-Will you select a model with slightly worse performance if it makes more sense or is more parsimonious?
-Discuss why you selected your model.
-
-For the multiple linear regression model, will you use a metric such as Adjusted R2, RMSE, etc.?
-Be sure to explain how you can make inferences from the model, discuss multi-collinearity issues (if any), and discuss other relevant model output.
-Using the training data set, evaluate the multiple linear regression model based on (a) mean squared error, (b) R2, (c) F-statistic, and (d) residual plots.
-Make predictions using the evaluation data set.
-
-## APPENDIX
-### Assignment Overview
-In this homework assignment, you will explore, analyze and model a data set containing approximately 2200
-records. Each record represents a professional baseball team from the years 1871 to 2006 inclusive. Each record
-has the performance of the team for the given year, with all of the statistics adjusted to match the performance of
-a 162 game season.
-
-Your objective is to build a multiple linear regression model on the training data to predict the number of wins
-for the team. You can only use the variables given to you (or variables that you derive from the variables
-provided). Below is a short description of the variables of interest in the data set:
-
-\begin{center}
-\begin{tabular}{|l|l|l|}
-\hline
-\textbf{VARIABLE NAME} & \textbf{DEFINITION} & \textbf{THEORETICAL EFFECT} \\ \hline
-INDEX & Identification Variable (do not use) & None \\ \hline
-TARGET\_WINS & Number of wins & \\ \hline
-TEAM\_BATTING\_H & Base Hits by batters (1B,2B,3B,HR) & Positive Impact on Wins  \\ \hline
-TEAM\_BATTING\_2B & Doubles by batters (2B) & Positive Impact on Wins \\ \hline
-TEAM\_BATTING\_3B & Triples by batters (3B) & Positive Impact on Wins \\ \hline
-TEAM\_BATTING\_HR & Homeruns by batters (4B) & Positive Impact on Wins \\ \hline
-TEAM\_BATTING\_BB & Walks by batters & Positive Impact on Wins \\ \hline
-TEAM\_BATTING\_HBP & Batters hit by pitch (get a free base) & Positive Impact on Wins \\ \hline
-TEAM\_BATTING\_SO & Strikeouts by batters & Negative Impact on Wins \\ \hline
-TEAM\_BASERUN\_SB & Stolen bases & Positive Impact on Wins \\ \hline
-TEAM\_BASERUN\_CS & Caught stealing & Negative Impact on Wins \\ \hline
-TEAM\_FIELDING\_E & Errors & Negative Impact on Wins \\ \hline
-TEAM\_FIELDING\_DP & Double Plays & Positive Impact on Wins \\ \hline
-TEAM\_PITCHING\_BB & Walks allowed & Negative Impact on Wins \\ \hline
-TEAM\_PITCHING\_H & Hits allowed & Negative Impact on Wins \\ \hline
-TEAM\_PITCHING\_HR & Homeruns allowed & Negative Impact on Wins \\ \hline
-TEAM\_PITCHING\_SO & Strikeouts by pitchers & Positive Impact on Wins \\ \hline
-
-\end{tabular}
-\end{center}
-
-### R Programming code
-
-```{r}
-
+```{r loadData, include=FALSE}
+library(tidyverse)
+library(ggplot2)
+library(corrplot)
 train_set <- read.csv('data/moneyball-training-data.csv')
 eval_set <- read.csv('data/moneyball-evaluation-data.csv')
+```
 
-colnames(train_set)
+## 1. DATA EXPLORATION
+
+### How often does the team win?
+
+We are given a data set containing 2,276 records, each of which includes the 
+number of wins the team had that year and 15 additional statistics about the 
+season. The team wins on average about 50% of the games they play (81 games 
+out of 162), with the best season having 146 wins and the worst season having 
+0 wins. The data is pretty normally distributed, with most seasons having between
+49 and 112 wins (blue lines below). This means that there aren't too many extreme 
+seasons where the team won significantly more or significantly less games than 
+usual. This serves as a good gut-check for our final predictions; if the 
+predicted wins are too high or too low, we know something is probably off.
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}
+
+avgWins <- mean(train_set$TARGET_WINS)
+sdev <- sd(train_set$TARGET_WINS)
+l1 <- mean(train_set$TARGET_WINS) - (2*sdev)
+l2 <- mean(train_set$TARGET_WINS) + (2*sdev)
+
+# histogram of wins
+p <- ggplot(train_set, aes(x=TARGET_WINS)) + 
+  geom_histogram() + 
+  geom_vline(aes(xintercept= avgWins),
+            color="red", linetype="dashed", size=1)+
+  geom_vline(aes(xintercept=l1),
+            color="blue", linetype="dashed", size=1) +
+    geom_vline(aes(xintercept=l2),
+            color="blue", linetype="dashed", size=1) +
+  labs(title="Distribution of Wins",x="Number of wins", y = "Number of seasons")
+
+p
+```
+
+### What's missing? 
+A first look at the data shows that only about 8% of the records have 
+a full set of information. The good news is that most of the missing values come 
+from what we would expect because they don't happen too often; hit-by-pitch 
+(92% missing!), caught stealing (34% missing), and double plays (13% missing). 
+Since we have so little hit-by-pitch data, we expect that it doesn't contribute 
+to overall wins and will eliminate it from a few of the models we propose. The other
+two stats have less than half of the data missing, so we'll need to think of a 
+clever way to impute these values. The remaining missing information is from a 
+combination of stolen bases and strikeouts (by batters and pitchers). **It seems** 
+**completely unreasonable** to have zero strike outs in a season, so this is something 
+we'll most certainly have to impute.
+
+```{r, include = FALSE}
+
+pct_null <- data.frame(do.call("rbind", map(train_set %>% select(-INDEX), ~mean(is.na(.)))))
+colnames(pct_null) <- c('PCT_NULL')
+
+pct_null %>%
+  arrange(desc(PCT_NULL)) %>%
+  filter(PCT_NULL > 0)
 
 ```
+
+### Do the individual stats affect winning?
+Intuitively, we expect that Caught stealing, Errors, Hits allowed, Homeruns allowed, 
+Strikeouts by batters, and Walks allowed would all have a **negative** impact on
+the total wins. In other words, as these stats increase, we would expect that the 
+team is less likely to win.
+
+However, when we look at the data for each of these statistics with respect to the 
+total number of wins, we don't see a clear relationship across the board. In fact, 
+only Errors and Hits allowed seem to have a negative impact on the wins. Caught stealing
+and strikeouts by batters appear to be random -- whether the stat for a particular season
+is high or low doesn't affect the overall number of wins.
+
+Even more interestingly, Homeruns and Walks allowed have the *opposite* effect; as the number
+of homeruns and walks that are allowed from the opposing team increase, the number of wins 
+appears to increase as well!
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
+
+negSet <- train_set %>%
+  select(TARGET_WINS, 
+         'Errors' = TEAM_FIELDING_E, 
+         'Hits allowed' = TEAM_PITCHING_H, 
+         'Strikeouts by batters' = TEAM_BATTING_SO, 
+         'Caught stealing' = TEAM_BASERUN_CS, 
+         'Walks allowed' = TEAM_PITCHING_BB, 
+         'Homeruns allowed' = TEAM_PITCHING_HR)
+
+ggplot(data = negSet %>%
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
+  facet_wrap(~ STAT, scales = "free")
+
+```
+
+We can look at the same information for the stats that we expect to have a **positive**
+effect on wins: Base hits, Doubles, Triples, Homeruns, and Walks by batters, Batters
+getting hit by pitches, Stolen bases, Double Plays, and Strikeouts by pitchers.
+
+Many of these stats *do* seem to have an effect on the number of wins. Most noteably,
+as the number of base hits and walks increase, the team is more likely to win. We see
+a slighter, positive relationship between homeruns, doubles, triples, and stolen bases.
+This makes sense when we think about it -- these things tend to happen less often in
+games than pure base hits and walks. Double plays and batters hit don't appear to have
+an effect on the number of wins either. Once again, this intuitively makes sense because
+they are less likely to happen in a game.
+
+One thing to note is the number of strikeouts compared to the number of wins. We can see
+that there are a few outliers -- abnormally high numbers of strikeouts in a season. This
+should be taken with caution, as they don't represent a typical season's stats.
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
+
+posSet <- train_set %>%
+  select(TARGET_WINS, 
+         'Base hits' = TEAM_BATTING_H, 
+         'Doubles' = TEAM_BATTING_2B, 
+         'Triples' = TEAM_BATTING_3B, 
+         'Homeruns' = TEAM_BATTING_HR, 
+         'Walks' = TEAM_BATTING_BB, 
+         'Batters hit' = TEAM_BATTING_HBP,
+         'Stolen bases' = TEAM_BASERUN_SB,
+         'Double plays' = TEAM_FIELDING_DP,
+         'Strikeouts' = TEAM_PITCHING_SO)
+
+ggplot(data = posSet %>%
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
+  facet_wrap(~ STAT, scales = "free")
+
+```
+
+Some of the provided statistics are well-balanced in the sense that there are
+very few (or no) extreme values. **Homeruns by batters**, **Strikeouts by** 
+**batters**, and **Homeruns allowed** are examples of this. Aside from
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
+
+set1 <-  train_set %>%
+  select('Homeruns by batters' = TEAM_BATTING_HR, 
+         'Strikeouts by batters' = TEAM_BATTING_SO,  
+         'Homeruns allowed' = TEAM_PITCHING_HR) %>%
+  gather("stat", "value") %>%
+  filter(complete.cases(.) == TRUE)
+
+vals <- ggplot(set1 , aes(x=stat, y=value)) + 
+    geom_boxplot() +
+  labs(title="",x="", y = "Value") +
+  facet_wrap(~stat, scale="free")
+
+vals
+
+# sapply(train_set %>% select(-INDEX, -TARGET_WINS), function(x){length(boxplot.stats(x)$out)})
+
+```
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
+
+# Correlation matrix with significance levels (p-value)
+library("Hmisc")
+res2 <- rcorr(as.matrix(train_set %>% select(-INDEX)))
+
+# Insignificant correlation are crossed
+corrplot(res2$r, type="upper", order="hclust", 
+          tl.col = "black", p.mat = res2$P, sig.level = 0.01, insig = "blank")
+
+```

### 2020-09-10 HW1/HW1_AF.Rmd
updating text

@@
 We are given a data set containing 2,276 records, each of which includes the 
-number of wins the team had that year and 15 additional statistics about the 
-season. The team wins on average about 50% of the games they play (81 games 
+number of wins the team had during a particular year and 15 additional statistics 
+about the season. The team wins on average about 50% of the games they play (81 games 
 out of 162), with the best season having 146 wins and the worst season having 
-0 wins. The data is pretty normally distributed, with most seasons having between
+0 wins. The data is pretty normally distributed and most seasons have between
 49 and 112 wins (blue lines below). This means that there aren't too many extreme 
@@
 ### What's missing? 
-A first look at the data shows that only about 8% of the records have 
-a full set of information. The good news is that most of the missing values come 
-from what we would expect because they don't happen too often; hit-by-pitch 
-(92% missing!), caught stealing (34% missing), and double plays (13% missing). 
-Since we have so little hit-by-pitch data, we expect that it doesn't contribute 
-to overall wins and will eliminate it from a few of the models we propose. The other
-two stats have less than half of the data missing, so we'll need to think of a 
-clever way to impute these values. The remaining missing information is from a 
-combination of stolen bases and strikeouts (by batters and pitchers). **It seems** 
-**completely unreasonable** to have zero strike outs in a season, so this is something 
-we'll most certainly have to impute.
+A first look at the data shows that only about 8% of the records have a full set
+of information. The good news is that most of the missing values come from 
+statistics that don't happen too often: hit-by-pitch (92% missing!), caught 
+stealing (34% missing), and double plays (13% missing). Since we have so little 
+hit-by-pitch data, we expect that it doesn't contribute much to overall wins and will 
+eliminate it from a few of the models we propose. The other two stats have less 
+than half of the data missing, so we'll need to think of a clever way to fill in 
+these values. The remaining missing information is from a combination of stolen 
+bases and strikeouts (by batters and pitchers). **It seems completely unreasonable** 
+to have zero strike outs in a season, so this is something we'll most certainly have 
+to impute.
 
@@
 Strikeouts by batters, and Walks allowed would all have a **negative** impact on
-the total wins. In other words, as these stats increase, we would expect that the 
-team is less likely to win.
+the total wins. In other words, as these values increase, we expect that the team 
+is less likely to win.
 
-However, when we look at the data for each of these statistics with respect to the 
-total number of wins, we don't see a clear relationship across the board. In fact, 
-only Errors and Hits allowed seem to have a negative impact on the wins. Caught stealing
-and strikeouts by batters appear to be random -- whether the stat for a particular season
-is high or low doesn't affect the overall number of wins.
+However, when we take a closer look at the data, these relationships aren't obvious. 
+In fact, only Errors and Hits allowed seem to have a negative impact on wins. Caught 
+stealing and Strikeouts by batters appear to be random; this means that whether the 
+stat for a particular season is high or low doesn't affect the overall number of wins.
 
-Even more interestingly, Homeruns and Walks allowed have the *opposite* effect; as the number
-of homeruns and walks that are allowed from the opposing team increase, the number of wins 
-appears to increase as well!
+Even more interestingly, Homeruns allowed and Walks allowed have the *opposite* 
+effect; as these stats increase, so do the number of wins!
 
@@
 We can look at the same information for the stats that we expect to have a **positive**
-effect on wins: Base hits, Doubles, Triples, Homeruns, and Walks by batters, Batters
-getting hit by pitches, Stolen bases, Double Plays, and Strikeouts by pitchers.
+effect on wins: Base hits, Doubles, Triples, Homeruns, Walks, Batters getting hit by 
+pitches, Stolen bases, Double Plays, and Strikeouts by pitchers.
 
-Many of these stats *do* seem to have an effect on the number of wins. Most noteably,
-as the number of base hits and walks increase, the team is more likely to win. We see
-a slighter, positive relationship between homeruns, doubles, triples, and stolen bases.
+Many of these stats *do* seem to have an effect on the number of wins. Most notably,
+as the number of base hits and walks increase, so do the number of wins. We see
+weaker positive relationships for homeruns, doubles, triples, and stolen bases.
 This makes sense when we think about it -- these things tend to happen less often in
-games than pure base hits and walks. Double plays and batters hit don't appear to have
-an effect on the number of wins either. Once again, this intuitively makes sense because
-they are less likely to happen in a game.
+games than pure base hits and walks, so they don't have as much of an effect on winning. 
+Double plays and batters hit don't appear to have an effect on the number of wins either. 
+Once again, this intuitively makes sense because they are less likely to happen in a game.
 
@@
 
+### Are some stats more skewed than others?
+
 Some of the provided statistics are well-balanced in the sense that there are
@@
 
+### Are stats correlated?
 ```{r, echo = FALSE, message = FALSE, warning = FALSE}

### 2020-09-10 HW1/HW1_AF.Rmd
updating grammar

@@
 
-We are given a data set containing 2,276 records, each of which includes the 
-number of wins the team had during a particular year and 15 additional statistics 
-about the season. The team wins on average about 50% of the games they play (81 games 
-out of 162), with the best season having 146 wins and the worst season having 
-0 wins. The data is pretty normally distributed and most seasons have between
-49 and 112 wins (blue lines below). This means that there aren't too many extreme 
-seasons where the team won significantly more or significantly less games than 
-usual. This serves as a good gut-check for our final predictions; if the 
-predicted wins are too high or too low, we know something is probably off.
+We are given a data set containing 2,276 records, each of which includes 15 statistics
+and the total number of wins a team had in a given year. On average, the team wins about 
+50% of the games they play (81 games out of 162), with the best season having 146 wins 
+and the worst season having 0 wins. The data is normally distributed and most seasons have 
+between 49 and 112 wins (blue lines below). This means that there aren't too many extreme 
+seasons where the team won significantly more or significantly less games than usual. This 
+serves as a good gut-check for our final predictions; if the predicted wins are too high or 
+too low, we know something is probably off.
 

### 2020-09-10 HW1/HW1_AF.Rmd
more grammar updates

@@
 We are given a data set containing 2,276 records, each of which includes 15 statistics
-and the total number of wins a team had in a given year. On average, the team wins about 
-50% of the games they play (81 games out of 162), with the best season having 146 wins 
-and the worst season having 0 wins. The data is normally distributed and most seasons have 
-between 49 and 112 wins (blue lines below). This means that there aren't too many extreme 
-seasons where the team won significantly more or significantly less games than usual. This 
-serves as a good gut-check for our final predictions; if the predicted wins are too high or 
-too low, we know something is probably off.
+and the total number of wins a team had in a given year. On average, about 50% of games 
+played are won (81 games out of 162), with the best season having 146 wins and the worst 
+season having 0 wins. The data is normally distributed and most years have between 49 
+and 112 wins (blue lines below). The nature of the distribution means there aren't too 
+many extreme seasons where wins are significantly higher or lower than usual. This serves 
+as a good gut-check for our final predictions; if the predicted wins are too high or too 
+low, we know something in our model is probably off.
 

### 2020-09-10 HW1/reg_model_shamp.Rmd
updated modeling

@@
 train_df <- complete(mice(data = train_df, 
-                         method = "norm", 
+                         method = "pmm", 
                          seed = 9450, 
@@
 
-Triples and singles seem to be candidates to remove from the analysis.
+Triples, singles, and pitching hits seem to be candidates to remove from the analysis.
 
@@
           TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
-          TEAM_PITCHING_HR+ TEAM_PITCHING_BB+TEAM_PITCHING_SO+TEAM_PITCHING_H+
+          TEAM_PITCHING_HR+ TEAM_PITCHING_BB+TEAM_PITCHING_H+
           TEAM_FIELDING_E+ TEAM_FIELDING_DP+OBP+ SLG, data=train_df)
 summary(lm_1)
+```
+
+## Remove Outliers that Influence Fit
+
+We need further outlier treatment. We will use Cook's distance to remove outliers that are influencing the fit of the model above.
+
+```{r}
+cooks_dis<- cooks.distance(lm_1)
+influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4/nrow(train_df)))])
+train_df<- train_df[-influential, ]
 ```
@@
 test_df <- complete(mice(data = test_df,
-                         method = "norm",
+                         method = "pmm",
                          seed = 9450,
@@
 
-Wow. max of 944, not good. How many terrible predictions are there? Let's filter by predictions beyond the currentl MLB season wins record of 116.
+Wow. max of 1231, not good. How many terrible predictions are there? Let's filter by predictions beyond the currentl MLB season wins record of 116.
 
@@
 test_pred %>%
-  filter(predictions >= 116)
+  filter(predictions >= 116) %>%
+  arrange(predictions)
 ```
 
-Ten beyond the record, two of which are somewhat reasonable. The rest are not reasonable.
+Nine beyond the record, three of which are somewhat reasonable. The rest are not reasonable.

### 2020-09-10 HW1/reg_model_shamp.Rmd
fix data load files

@@
 train_df<- 
-  read.csv("/Users/jeffshamp/Documents/Github/DT621_Fall2020_Group2/HW1/data/moneyball-training-data.csv")
-train_df<- 
+  read.csv("./data/moneyball-training-data.csv")
+train_df <-
   train_df %>%
-  select(-c(TEAM_BATTING_HBP, INDEX)) %>%
+  select(-c(TEAM_BATTING_HBP, -INDEX)) %>%
   mutate_if(is.integer, as.numeric)
@@
 ```{r}
+# Base R, disgusting
 cooks_dis<- cooks.distance(lm_1)
@@
 test_df<- 
-  read.csv("/Users/jeffshamp/Documents/Github/DT621_Fall2020_Group2/HW1/data/moneyball-evaluation-data.csv")
+  read.csv("./data/moneyball-evaluation-data.csv")
 test_df<- 

### 2020-09-10 HW1/Regression Model - Deokinanan.rmd
Code added to get data from Git

@@
 # Load training data set
-train_df = read.csv("C:/Users/Deokinanan/Desktop/2019 -2021 CUNY SPS/FALL 2020/DATA 621/WEEK 3/moneyball-training-data.csv")
+theURL = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv"
+train_df = data.frame(read.csv(file = theURL, header = TRUE, sep = ","))
 train_df = subset(train_df, select = -c(INDEX))
@@
 After the steps, the final model resulted below, with $adj.
-R^2 = 0.42$, suggesting that this model accounts for nearly 42% of the variation in the dependent variable with the independent variables, which is acceptable as a good model.
+R^2 = 0.43$, suggesting that this model accounts for nearly 42% of the variation in the dependent variable with the independent variables, which is acceptable as a good model.
 Within the method of `lmStepAIC`, AIC (Akaike Information Criteria) quantifies the amount of information loss due to simplification.
@@
 # load the test data
-test_df = read.csv("C:/Users/Deokinanan/Desktop/2019 -2021 CUNY SPS/FALL 2020/DATA 621/WEEK 3/moneyball-evaluation-data.csv")
+theURL = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-evaluation-data.csv"
+test_df = data.frame(read.csv(file = theURL, header = TRUE, sep = ","))
 

### 2020-09-10 HW1/HW1_BP.rmd
Update HW1_BP.rmd

@@
 
-train_set <- read.csv('/Users/Bryan/Documents/Data 621/HW 1/moneyball-training-data.csv')
-eval_set <- read.csv('/Users/Bryan/Documents/Data 621/HW 1/moneyball-evaluation-data.csv')
+train_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
+eval_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-evaluation-data.csv')
 

### 2020-09-10 HW1/HW1_AF.Rmd
more updates

@@
 
-```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}
+```{r wins, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}
 
@@
 
-```{r, include = FALSE}
+```{r pctNull, include = FALSE}
 
@@
 
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
+```{r negFactors, echo = FALSE, message = FALSE, warning = FALSE}
 
@@
   facet_wrap(~ STAT, scales = "free")
-
 ```
@@
 
-Many of these stats *do* seem to have an effect on the number of wins. Most notably,
-as the number of base hits and walks increase, so do the number of wins. We see
-weaker positive relationships for homeruns, doubles, triples, and stolen bases.
-This makes sense when we think about it -- these things tend to happen less often in
-games than pure base hits and walks, so they don't have as much of an effect on winning. 
-Double plays and batters hit don't appear to have an effect on the number of wins either. 
-Once again, this intuitively makes sense because they are less likely to happen in a game.
+Many of these stats *do* seem to have an effect on the number of wins, most notably,
+base hits and walks. We see weaker positive relationships for homeruns, doubles, triples, 
+and stolen bases. This makes sense when we think about it; these things tend to happen 
+less often in games than pure base hits and walks, so they don't have as much of an 
+effect on winning. Finally, double plays and batters hit don't appear to have any 
+correlation with the number of wins. Once again, this intuitively makes sense because 
+they are less likely to happen in a game.
 
 One thing to note is the number of strikeouts compared to the number of wins. We can see
-that there are a few outliers -- abnormally high numbers of strikeouts in a season. This
+that there are a few outliers (abnormally high numbers of strikeouts in a season). This
 should be taken with caution, as they don't represent a typical season's stats.
 
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
+```{r posFactors, echo = FALSE, message = FALSE, warning = FALSE}
 
@@
 very few (or no) extreme values. **Homeruns by batters**, **Strikeouts by** 
-**batters**, and **Homeruns allowed** are examples of this. Aside from
+**batters**, and **Homeruns allowed** are examples of this.
 
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
+```{r summaryInfo, include = FALSE}
+
+summary(train_set)
+```
+
+```{r fewOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}
 
@@
 
-# sapply(train_set %>% select(-INDEX, -TARGET_WINS), function(x){length(boxplot.stats(x)$out)})
+```
+Some things to note about each of these statistics:
+
+* Homeruns allowed (average = ~100/year) and Homeruns by batters (average = ~106/year) 
+have a very similar mid-range distribution (50% of the data lies between ~50 and 150). 
+The slight difference in average stats means that teams tend to have a higher number
+of homeruns than the opposition team. 
+* The only thing that stands out about strikeouts by batters (average = ~736/year) is 
+how nearly perfectly normal it is. 50% of the data is between about 500 and 1000 and 
+there are absolutely no outliers in the dataset! This means that there were no surprisingly
+high or low seasons.
+
+Conversely, some of the stats have a very high number of outliers, indicating that there
+are some seasons with some abnormally high or low values. **Walks by batters**, **Stolen**
+**bases**, **Hits allowed**, and **Errors** are examples of this.
+
+```{r manyOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5, fig.align = 'center'}
+
+set2 <-  train_set %>%
+  select('Walks by batters' = TEAM_BATTING_BB, 
+         'Stolen bases' = TEAM_BASERUN_SB,  
+         'Hits allowed' = TEAM_PITCHING_H,
+         'Errors' = TEAM_FIELDING_E) %>%
+  gather("stat", "value") %>%
+  filter(complete.cases(.) == TRUE)
+
+vals2 <- ggplot(set2 , aes(x=stat, y=value)) + 
+    geom_boxplot() +
+  labs(title="",x="", y = "Value") +
+  facet_wrap(~ stat, scale = 'free')
+
+vals2
+
+#sapply(train_set %>% select(-INDEX, -TARGET_WINS), function(x){length(boxplot.stats(x)$out)})
 
 ```
+Some things to note about each of these statistics:
 

### 2020-09-10 HW1/reg_model_shamp.Rmd

@@
 date: "9/8/2020"
-output: html_document
+output: pdf_document
 ---
@@
 
-## Data
+## Data Preparation
 
-We have the mlb training data set. See `missing_value_eda.Rmd` for details on data distribution.
-
-We saw from the above file that at least one variable should be omitted for low reporting of data. There are also outlier concerns for several of the variables, namely `TEAM_PITCHING_SO`.
+We saw from the above that at least one variable should be omitted for low reporting of data. There are also outlier concerns for several of the variables, namely `TEAM_PITCHING_SO`.
 
@@
 
-```{r}
+```{r load, echo=FALSE}
 set.seed(9450)
@@
 
-```{r}
+```{r singles, echo=FALSE}
 train_df<- 
@@
 
-Looking at some distributions.
+### Outlier Removal
 
-```{r}
-train_df %>%
-  ggplot(aes(TEAM_BATTING_1B)) +
-  geom_histogram(na.rm = TRUE, binwidth = 1) +
-   geom_vline(xintercept = mean(
-              train_df$TEAM_BATTING_1B,
-              na.rm = TRUE), color="red") 
-```
+Again, we will take a cautious approach to removing outliers at first. Removing values that appear to be so far out from the normal that is it seems likely that there is some kind of data input issue.
 
-## Outlier Removal
-
-Again, we will take a cautious approach to removing outliers. Removing values that appear to be so far out from the normal that is it seems likely that there is some kind of data input issue.
-
-```{r}
+```{r outliers_1, echo=FALSE}
 train_df %>%
@@
 There are some clear issues with hitting singles and some that we know about with strikeouts.
-We will remove the SO outliers first and see what needs to happen with singles hitting.
+We will remove the `SO` outliers first and see what needs to happen with singles hitting.
 These appear to right skewed so we will use median and IQR to remove outliers.
 
-```{r}
+```{r iqr_removal_SO, echo=FALSE}
 train_df<-
@@
 
-```{r}
+```{r iqr_removal_H, echo=FALSE}
 train_df<-
@@
 
-```{r}
+```{r outlier_boxplot, echo=FALSE}
 train_df %>%
@@
 
-```{r, message=FALSE}
+```{r impute_train,echo=FALSE, message=FALSE, warning=FALSE}
 train_df <- complete(mice(data = train_df, 
@@
 
-With outliers removed and missing values imputated via MICE, we can add a few features.
-We will take the obvious and approximate the following features; on-base percent, slugging, and on-base plu slugging.
-To reduce colinearity, we may only keep on-base percent and slugging.
-
-### On Base Percentage
-
-On-base percentage is calculated using this formula:
-
-${\displaystyle OBP={\frac {H+BB+HBP}{AB+BB+HBP+SF}}}$
-
-We are missing several of the needed variables but we can approximate this percentage in the following way.
-
-${\displaystyle \sim OBP={\frac {H+BB}{SO_{batting}+BB+H}}}$
-
-We do not have the total At-bats for each observation so we will approximate at-bats by summing batting strikeouts and total hits.
+With outliers removed and missing values imputed via MICE, we can add a few features. We will take the obvious and approximate the slugging feature.
 
@@
 
-```{r}
+```{r make_slg_obp, echo=FALSE}
 train_df<-
@@
   mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
-         OBP = (TEAM_BATTING_H + TEAM_BATTING_BB)/AB,
          SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) %>%
@@
 
-We will start with a simple linear model with resampling and see what kind of results we get.
+We will start with a simple linear model and select out features based on null hypothesis testing for non-zero slope.
 
-```{r}
+Below is the output of coefficients for backward selection model.
+
+```{r lm_first, echo=FALSE, eval=FALSE}
 lm_reg = lm(data=train_df, TARGET_WINS ~ .)
@@
 
-Triples, singles, and pitching hits seem to be candidates to remove from the analysis.
-
-```{r}
+```{r lm_selected, echo=FALSE}
 lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
           TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
-          TEAM_PITCHING_HR+ TEAM_PITCHING_BB+TEAM_PITCHING_H+
-          TEAM_FIELDING_E+ TEAM_FIELDING_DP+OBP+ SLG, data=train_df)
-summary(lm_1)
+          TEAM_PITCHING_HR+ TEAM_PITCHING_H+
+          TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_df)
+knitr::kable(lm_1$coefficients, digits = 3L)
 ```
 
-## Remove Outliers that Influence Fit
+### Model Coefficient Discussion
 
-We need further outlier treatment. We will use Cook's distance to remove outliers that are influencing the fit of the model above.
+First, we are keeping the coefficient for `TEAM_BATTING_HR` since it's p-value is marginally above the general threshold and knowledge of the game suggests it is important.
+There are some counter intuitive results, which is expected given that baseball is a messy, imprecise game that has evolved over time.
+To that end, we should expect some seemingly strange results from algorithmic regression.
+We used a combination of domain knowledge and data analysis to justify retaining features.
 
-```{r}
-# Base R, disgusting
+Fielding double plays and batting doubles both appear to have negative impacts on wins even though they _should_ be postive impacts.
+Turning double plays, while a good for the defensive team, may suggest a larger, negative issue.
+Namely, weak pitching that leads to runners on base.
+Similarly, batting doubles, leaves runners open to double plays.
+Allowed hits by pitching, caught stealing, and batting strike-outs are all counter intuitive results as well, but they seem to be small contributors and these are events that happen regularly in every game.
+
+Slugging as an approximation is the major predictor in this regression, by far.
+The other predictors other than the intercept are orders of magnitude less in predictive value.
+It should be noted that slugging alone is not a good predictor for wins overall.
+
+### Remove Outliers that Influence Fit
+
+We need further outlier treatment. We will use Cook's distance to remove outliers that are influencing the fit of the model above. We will use a cutoff of $\frac{4}{N}$.
+
+```{r cooks_distance, echo=FALSE}
 cooks_dis<- cooks.distance(lm_1)
@@
 
+### Model Evaluation
+
 Let's take this backward selection process and use it in a resampling scheme and see how our R values hold up.
 
-```{r}
+```{r resample, echo=FALSE}
 folds<- vfold_cv(train_df,
@@
   add_formula(
-    TARGET_WINS ~ 
-          TEAM_BATTING_2B +TEAM_BATTING_HR+
-          TEAM_BATTING_BB +TEAM_BATTING_SO+ 
-          TEAM_BASERUN_SB +TEAM_BASERUN_CS+ 
-          TEAM_PITCHING_HR +TEAM_PITCHING_BB+
-          TEAM_PITCHING_SO +TEAM_PITCHING_H+
-          TEAM_FIELDING_E +TEAM_FIELDING_DP+
-          OBP+ SLG) %>%
+    TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
+          TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
+          TEAM_PITCHING_HR+ TEAM_PITCHING_H+
+          TEAM_FIELDING_E+ TEAM_FIELDING_DP+ SLG
+    ) %>%
   add_model(lm_model)
@@
 lm_resample %>%
-  collect_metrics()
+  collect_metrics() %>%
+  knitr::kable(digits = 3L)
 ```
-So the metrics didn't tank with cross validation, that's good. Let's fit the entire training data one last time and look at $R^{2}$ and RMSE.
 
-```{r}
+The metrics hold up with cross validation, that's good. Let's fit the entire training dataset one last time and look at $R^{2}$ and RMSE.
+
+```{r last_fit, echo=FALSE}
 results_train<-
@@
 results_train %>%
-  multi_metric(truth=truth, estimate = .pred)
+  multi_metric(truth=truth, estimate = .pred) %>%
+  knitr::kable(digits = 3L)
 
@@
 
-```{r}
+```{r test_prep, echo=FALSE, warning=FALSE, message=FALSE}
 test_df<- 
@@
   mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
-         OBP = (TEAM_BATTING_H + TEAM_BATTING_BB)/AB,
          SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) %>%
   select(-c(AB, TEAM_BATTING_H))
-
 ```
 
-```{r}
+```{r make_predictions, echo=FALSE}
 test_pred<-
@@
   fit(train_df) %>%
-  predict(new_data = test_df) %>%
+  predict(new_data=test_df) %>%
   rename(predictions = .pred)
 
-summary(test_pred)
+knitr::kable(summary(test_pred))
 ```
 
-Wow. max of 1231, not good. How many terrible predictions are there? Let's filter by predictions beyond the currentl MLB season wins record of 116.
-
-```{r}
+```{r prediction, echo=FALSE}
 test_pred %>%
-  filter(predictions >= 116) %>%
-  arrange(predictions)
+  filter(predictions >= 116 | predictions <=0) %>%
+  arrange(predictions) %>%
+  knitr::kable()
 ```
 
-Nine beyond the record, three of which are somewhat reasonable. The rest are not reasonable.
+Three predictions are unreasonably large, but not bad given the metrics.
+Two predictions are above the MLB reccord, but only by two or three wins.
+With an RMSE of 9.53 and $R^{2}=0.446$, and $R^{2}_{ADJ} =0.43$ we have a model that explains 44% of the variance in the data and can predict wins within $\pm 10$ wins for a given season.
+That's pretty good at predicting wins relative to seasoned baseball veterans.

### 2020-09-10 HW1/Regression Model - Deokinanan.rmd
edited for pdf format

--- 2020-09-06 HW1/Regression Model - Deokinanan.rmd
@@
 date: "September 3, 2020"
-output: html_document
+output:
+  pdf_document: default
 ---
 
-***
-#### Task
-***
+## Task
 
@@
 
-***
-#### The dataset
-***
+## The Dataset
 
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE, include=FALSE}
 # The required R packages
@@
 # Load training data set
-train_df = read.csv("C:/Users/Deokinanan/Desktop/2019 -2021 CUNY SPS/FALL 2020/DATA 621/WEEK 3/moneyball-training-data.csv")
+theURL = "https://raw.githubusercontent.com/aadler/
+DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv"
+train_df = data.frame(read.csv(file = theURL, header = TRUE, sep = ","))
 train_df = subset(train_df, select = -c(INDEX))
@@
 
-***
-#### Data Transformation, Outliers and Missing Data
-***
+## Data Transformation, Outliers and Missing Data
 
@@
 
-```{r}
+```{r include=FALSE}
 # Quick Exploration
-sapply(train_df, describe)
+# sapply(train_df, describe)
 
@@
 Because the objective is so create a multivariate regression model, declaring an observation as an outlier based on a just one feature could lead to unrealistic inferences.
-There Cook's distance is use to decide if an individual entity is an extreme value or not.
+Therefore Cook's distance is use to decide if an individual entity is an extreme value or not.
 
@@
 Note from the plots that all influential observations are not necessarily outliers.
-Least than 3% of the data was removed.
+Less than 3% of the data was removed.
 
-```{r}
+```{r include=FALSE}
 # Outlier Plots
 model = lm(TARGET_WINS ~ ., data = train_df)
-ols_plot_diagnostics(model)
+# ols_plot_diagnostics(model)
+p1 = ols_plot_cooksd_chart(model)
+p2 = ols_plot_resid_fit(model)
+p3 = ols_plot_resid_hist(model)
 
@@
 model_new = lm(TARGET_WINS ~ ., data = train_df)
-ols_plot_diagnostics(model_new)
+# ols_plot_diagnostics(model_new)
+p4 = ols_plot_cooksd_chart(model_new)
+p5 = ols_plot_resid_fit(model_new)
+p6 = ols_plot_resid_hist(model_new)
+```
+```{r echo=FALSE, fig.height=7}
+gridExtra::grid.arrange(p1,p4,p2,p5,p3,p6, nrow=3, top = "Before and After Outliers were Removed")
 ```
@@
 
-```{r}
-aggr(train_df, col = c('darkseagreen2','darkslateblue'), numbers = TRUE, sortVars = TRUE, oma = c(10,5,5,3), labels = names(train_df), cex.axis = 0.8, gap = 3, axes = TRUE, Prop = TRUE, ylab = c("Proportion of Missing Data", "Combination"))
+```{r echo=FALSE, fig.height=5}
+aggr(train_df, col = c('darkseagreen2','darkslateblue'), numbers = TRUE, sortVars = FALSE, 
+     oma = c(10,5,5,3), labels = names(train_df), cex.axis = 0.8, gap = 3, axes = TRUE, 
+     Prop = TRUE, ylab = c("Proportion of Missing Data", "Combination"))
 
-train_df.clean = complete(mice(data = train_df, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
-sapply(train_df.clean, describe)
+train_df.clean = complete(mice(data = train_df, method = "pmm", maxit = 5, 
+                               seed = 525, print = FALSE), 3)
+# sapply(train_df.clean, describe)
 ```
 
-***
-#### Initial Tests 
-***
+## Initial Tests
 
@@
 
-```{r}
+```{r include=FALSE}
 # Data Transformation
-train_df.clean$TEAM_BATTING_1B = train_df.clean$TEAM_BATTING_H - (train_df.clean$TEAM_BATTING_2B + train_df.clean$TEAM_BATTING_3B + train_df.clean$TEAM_BATTING_HR)
+train_df.clean$TEAM_BATTING_1B = train_df.clean$TEAM_BATTING_H - 
+  (train_df.clean$TEAM_BATTING_2B + train_df.clean$TEAM_BATTING_3B + 
+     train_df.clean$TEAM_BATTING_HR)
 train_df.clean = subset(train_df.clean, select = -c(TEAM_BATTING_H))
@@
 
-```{r, fig.width=7, fig.height=9}
+```{r, fig.width=7, fig.height=7, echo=FALSE}
 # Correlation
@@
 ```
-
 Lastly, because a robust regression process will be performed, higher order polynomials variables were introduced into the full model.
 
-```{r}
-stepdata = cbind(train_df.clean,sapply(train_df.clean[2:length(train_df.clean)], function(x) x^2))
+```{r echo=FALSE}
+stepdata = cbind(train_df.clean,sapply(train_df.clean[2:length(train_df.clean)], 
+                                       function(x) x^2))
 names(stepdata) = make.unique(names(stepdata), sep = "_")
@@
 
-***
-#### Regression Model #1
-***
+## Regression Model: Stepwise Regression with Repeated k-fold Cross-Validation
 
@@
 The steps evaluate all of the variables already included in the model and remove any variable that has an insignificant F-statistic.
-Only after this test ends, is the best model found, that is when none of the variables cstepan be excluded and every variable included in the model is significant.
+Only after this test ends, is the best model found, that is when none of the variables can be excluded and every variable included in the model is significant.
 
@@
 After the steps, the final model resulted below, with $adj.
-R^2 = 0.42$, suggesting that this model accounts for nearly 42% of the variation in the dependent variable with the independent variables, which is acceptable as a good model.
-Within the method of `lmStepAIC`, AIC (Akaike Information Criteria) quantifies the amount of information loss due to simplification.
+R^2 = 0.43$, suggesting that this model accounts for nearly 43% of the variation in the dependent variable with the independent variables, which is acceptable as a good model.
+With this method of stepwise regression, AIC (Akaike Information Criteria) quantifies the amount of information loss due to simplification.
 That is, based on the AIC, the final model outputted is the simplest model without impacting much on the performance.
 
-```{r  message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
 # Set up repeated k-fold cross-validation
@@
 # Train the model
-step.model = train(TARGET_WINS ~ ., data = stepdata, method = "lmStepAIC", trControl = train.control, trace = FALSE)
-
-# Model accuracy
-step.model$results
+step.model = train(TARGET_WINS ~ ., data = stepdata, method = "lmStepAIC", 
+                   trControl = train.control, trace = FALSE)
 
@@
 # Summary of the model
-summary(step.model$finalModel)
+mod_sum = summary(step.model$finalModel)
+
+# Prediction on training data
+predictions = predict(step.model, newdata = stepdata)
 ```
 
-***
-#### Prediction
-***
+```{r echo=FALSE}
+# Model accuracy
+data.frame(r.squared = c(mod_sum[["r.squared"]]), adj.r.squared = c(mod_sum[["adj.r.squared"]]), 
+           rsme = c(sqrt(mean((stepdata$TARGET_WINS - predictions)^2)))) %>% knitr::kable(digits = 2L)
+```
+
+## Prediction
 
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r echo=FALSE}
+step.model$finalModel[["coefficients"]] %>% knitr::kable()
+```
+
+Using the test data and the final step model, a comparison in the prediction statistic was conducted.
+It is apparent that the model's prediction are not too off from the training data.
+However, the predictions resulted in a higher kurtosis, thus there are expectation of greater extremity of the deviations, and not centered near the mean.
+
+```{r message=FALSE, warning=FALSE, echo=FALSE}
 # load the test data
-test_df = read.csv("C:/Users/Deokinanan/Desktop/2019 -2021 CUNY SPS/FALL 2020/DATA 621/WEEK 3/moneyball-evaluation-data.csv")
+theURL = "https://raw.githubusercontent.com/aadler/
+DT621_Fall2020_Group2/master/HW1/data/moneyball-evaluation-data.csv"
+test_df = data.frame(read.csv(file = theURL, header = TRUE, sep = ","))
 
@@
 test_df = complete(mice(data = test_df, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
-test_df$TEAM_BATTING_1B = test_df$TEAM_BATTING_H - (test_df$TEAM_BATTING_2B + test_df$TEAM_BATTING_3B + test_df$TEAM_BATTING_HR)
+test_df$TEAM_BATTING_1B = test_df$TEAM_BATTING_H - 
+  (test_df$TEAM_BATTING_2B + test_df$TEAM_BATTING_3B + test_df$TEAM_BATTING_HR)
 test_df = cbind(test_df,sapply(test_df[1:length(test_df)], function(x) x^2))
@@
 
-results = cbind(dataset = c('Training Data', 'Test Prediction'), bind_rows(describe(train_df$TARGET_WINS), describe(predictions))) 
+results = cbind(dataset = c('Training Data', 'Test Prediction'), 
+                bind_rows(describe(train_df$TARGET_WINS), describe(predictions))) 
 rownames(results) = NULL
 
-results
+results[,-c(2,8,11)] %>% knitr::kable(digits = 2L)
 ```

### 2020-09-10 HW1/reg_model_shamp.Rmd
fix typo, remove index

@@
   train_df %>%
-  select(-c(TEAM_BATTING_HBP, -INDEX)) %>%
+  select(-c(TEAM_BATTING_HBP, INDEX)) %>%
   mutate_if(is.integer, as.numeric)
@@
 Three predictions are unreasonably large, but not bad given the metrics.
-Two predictions are above the MLB reccord, but only by two or three wins.
-With an RMSE of 9.53 and $R^{2}=0.446$, and $R^{2}_{ADJ} =0.43$ we have a model that explains 44% of the variance in the data and can predict wins within $\pm 10$ wins for a given season.
+One prediction is above the MLB reccord, but only by one win.
+With an RMSE of 9.53 and $R^{2}=0.443$, and $R^{2}_{ADJ} =0.43$ we have a model that explains 44% of the variance in the data and can predict wins within $\pm 10$ wins for a given season.
 That's pretty good at predicting wins relative to seasoned baseball veterans.

### 2020-09-13 HW1/HW1_AF.Rmd
another round of editing

@@
 
-```{r loadData, include=FALSE}
+```{r loadData, include=FALSE, warning = FALSE, message = FALSE}
 library(tidyverse)
@@
 library(corrplot)
+library(data.table)
 train_set <- read.csv('data/moneyball-training-data.csv')
@@
 
-We are given a data set containing 2,276 records, each of which includes 15 statistics
-and the total number of wins a team had in a given year. On average, about 50% of games 
+We are given a data set of 2,276 records containing 15 seasonal statistics and the
+total number of wins a team had in a given year. On average, about 50% of games 
 played are won (81 games out of 162), with the best season having 146 wins and the worst 
@@
 
-```{r wins, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}
+```{r wins, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
 
@@
 
-```{r pctNull, include = FALSE}
+```{r pctNull, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2, fig.width = 5, fig.align = 'center'}
 
@@
 
-pct_null %>%
+totalNulls <- pct_null %>%
+  mutate(VARIABLE = rownames(.)) %>%
   arrange(desc(PCT_NULL)) %>%
-  filter(PCT_NULL > 0)
+  filter(PCT_NULL > 0) %>%
+  select(VARIABLE, PCT_NULL)
+
+ggplot(totalNulls, aes(x=reorder(VARIABLE, PCT_NULL), y=PCT_NULL, label = round(PCT_NULL, 2))) + 
+  geom_text(vjust = 0.5, hjust = -0.05)+
+  geom_bar(stat = "identity") +
+  ggtitle("Variables with Missing Information") +
+  xlab("Statistic") + ylab("Percent Missing") + 
+  coord_flip() +
+  expand_limits(y = 1)
 
@@
 ### Do the individual stats affect winning?
-Intuitively, we expect that Caught stealing, Errors, Hits allowed, Homeruns allowed, 
-Strikeouts by batters, and Walks allowed would all have a **negative** impact on
-the total wins. In other words, as these values increase, we expect that the team 
-is less likely to win.
 
-However, when we take a closer look at the data, these relationships aren't obvious. 
-In fact, only Errors and Hits allowed seem to have a negative impact on wins. Caught 
-stealing and Strikeouts by batters appear to be random; this means that whether the 
-stat for a particular season is high or low doesn't affect the overall number of wins.
-
-Even more interestingly, Homeruns allowed and Walks allowed have the *opposite* 
-effect; as these stats increase, so do the number of wins!
+**Stats with an expected negative impact: ** Intuitively, we expect that Caught 
+stealing, Errors, Hits allowed, Homeruns allowed, Strikeouts by batters, and Walks 
+allowed would all have a **negative** impact on the total wins. In other words, as 
+these values increase, we expect that the team is less likely to win.
 
@@
 
-We can look at the same information for the stats that we expect to have a **positive**
-effect on wins: Base hits, Doubles, Triples, Homeruns, Walks, Batters getting hit by 
-pitches, Stolen bases, Double Plays, and Strikeouts by pitchers.
+When we take a closer look at the data, these negative relationships aren't obvious. 
+In fact, only Errors and Hits allowed seem to have a negative impact on wins. Caught 
+stealing and Strikeouts by batters appear to be random; this means that whether the 
+stat for a particular season is high or low doesn't affect the overall number of wins.
 
-Many of these stats *do* seem to have an effect on the number of wins, most notably,
-base hits and walks. We see weaker positive relationships for homeruns, doubles, triples, 
-and stolen bases. This makes sense when we think about it; these things tend to happen 
-less often in games than pure base hits and walks, so they don't have as much of an 
-effect on winning. Finally, double plays and batters hit don't appear to have any 
-correlation with the number of wins. Once again, this intuitively makes sense because 
-they are less likely to happen in a game.
+Even more interestingly, Homeruns allowed and Walks allowed have the *opposite* 
+effect; as these stats increase, so do the number of wins!
 
-One thing to note is the number of strikeouts compared to the number of wins. We can see
-that there are a few outliers (abnormally high numbers of strikeouts in a season). This
-should be taken with caution, as they don't represent a typical season's stats.
+**Stats with an expected positive impact: ** We can look at the same information for the 
+stats that we expect to have a **positive** effect on wins: Base hits, Doubles, Triples, 
+Homeruns, Walks, Batters getting hit by pitches, Stolen bases, Double Plays, and Strikeouts 
+by pitchers.
 
@@
 
+Many of these stats *do* seem to have an effect on the number of wins, most notably,
+base hits and walks. We see weaker positive relationships for homeruns, doubles, triples, 
+and stolen bases. This makes sense when we think about it; these things tend to happen 
+less often in games than pure base hits and walks, so they don't have as much of an 
+effect on winning. Finally, double plays and batters hit don't appear to have any 
+correlation with the number of wins. Once again, this intuitively makes sense because 
+they are less likely to happen in a game.
+
+One thing to note is the number of strikeouts compared to the number of wins. We can see
+that there are a few outliers (abnormally high numbers of strikeouts in a season). This
+should be taken with caution, as they don't represent a typical season's stats.
+
 ### Are some stats more skewed than others?
 
-Some of the provided statistics are well-balanced in the sense that there are
-very few (or no) extreme values. **Homeruns by batters**, **Strikeouts by** 
-**batters**, and **Homeruns allowed** are examples of this.
+Before using any of the statistics in a model, we need to take a closer look at the 
+variation in the data. We call out of the ordinary values (exceptionally high or
+low values) **outliers**. We need to take these into account in our modeling because
+we want to make sure our predictions aren't skewed because of them.
 
@@
 
-```{r fewOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}
+Some of the provided statistics are well-balanced in the sense that there are
+very few (or no) extreme values. **Homeruns by batters**, **Strikeouts by** 
+**batters**, and **Homeruns allowed** are examples of this.
+
+```{r fewOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
 
@@
 
-* Homeruns allowed (average = ~100/year) and Homeruns by batters (average = ~106/year) 
+* **Homeruns allowed** (average = ~100/year) and **Homeruns by batters** (average = ~106/year) 
 have a very similar mid-range distribution (50% of the data lies between ~50 and 150). 
@@
 of homeruns than the opposition team. 
-* The only thing that stands out about strikeouts by batters (average = ~736/year) is 
+* The only thing that stands out about **Strikeouts by batters** (average = ~736/year) is 
 how nearly perfectly normal it is. 50% of the data is between about 500 and 1000 and 
@@
 Conversely, some of the stats have a very high number of outliers, indicating that there
-are some seasons with some abnormally high or low values. **Walks by batters**, **Stolen**
-**bases**, **Hits allowed**, and **Errors** are examples of this.
+are some seasons with some abnormally high or low values. **Errors**, **Hits allowed**,
+**Stolen bases**, and **Walks by batters** are examples of this.
 
-```{r manyOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5, fig.align = 'center'}
+```{r manyOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 6, fig.align = 'center'}
 
@@
   labs(title="",x="", y = "Value") +
-  facet_wrap(~ stat, scale = 'free')
+  facet_wrap(~ stat, scale = 'free', ncol = 4)
 
@@
 Some things to note about each of these statistics:
+
+* All of the outliers for **Errors** and **Hits allowed** are above the upper 
+tail of the data set. This is further illustrated by the mean and median values
+for both of these stats; in both instances, the mean per year (Errors = ~246/year
+and Hits allowed = ~1779/year) are higher than the median per year (Errors = 
+~159/year and Hits allowed = ~1518/year). This means that some seasons with
+exceptionally high values for both of these statistics skew the dataset. 
+* The maxiumum value
 

### 2020-09-14 HW1/HW1_draft_kellogg.Rmd
Rough Draft of Final submission
Combined most of the RMD files available here.
Exception HW1_BP.rmd - If we want to add the model here or the ggplot subplot, let me know

--- 2020-09-13 HW1/HW1_AF.Rmd
@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 library(data.table)
+library(tidymodels)
+library(mice)
+library(psych)
 train_set <- read.csv('data/moneyball-training-data.csv')
 eval_set <- read.csv('data/moneyball-evaluation-data.csv')
+
+set.seed(9450)
 ```
@@
 total number of wins a team had in a given year. On average, about 50% of games 
-played are won (81 games out of 162), with the best season having 146 wins and the worst 
+played are won (81 games out of 162), with the best individual season having 146 wins and the worst 
 season having 0 wins. The data is normally distributed and most years have between 49 
@@
 **Stats with an expected negative impact: ** Intuitively, we expect that Caught 
-stealing, Errors, Hits allowed, Homeruns allowed, Strikeouts by batters, and Walks 
+stealing, Errors, Hits allowed, Home Runs allowed, Strikeouts by batters, and Walks 
 allowed would all have a **negative** impact on the total wins. In other words, as 
@@
          'Walks allowed' = TEAM_PITCHING_BB, 
-         'Homeruns allowed' = TEAM_PITCHING_HR)
+         'Home Runs allowed' = TEAM_PITCHING_HR)
 
@@
 
-Even more interestingly, Homeruns allowed and Walks allowed have the *opposite* 
+Even more interestingly, Home Runs allowed and Walks allowed have the *opposite* 
 effect; as these stats increase, so do the number of wins!
@@
 stats that we expect to have a **positive** effect on wins: Base hits, Doubles, Triples, 
-Homeruns, Walks, Batters getting hit by pitches, Stolen bases, Double Plays, and Strikeouts 
+Home Runs, Walks, Batters getting hit by pitches, Stolen bases, Double Plays, and Strikeouts 
 by pitchers.
@@
          'Triples' = TEAM_BATTING_3B, 
-         'Homeruns' = TEAM_BATTING_HR, 
+         'Home Runs' = TEAM_BATTING_HR, 
          'Walks' = TEAM_BATTING_BB, 
@@
 Many of these stats *do* seem to have an effect on the number of wins, most notably,
-base hits and walks. We see weaker positive relationships for homeruns, doubles, triples, 
+base hits and walks. We see weaker positive relationships for home runs, doubles, triples, 
 and stolen bases. This makes sense when we think about it; these things tend to happen 
@@
 Some of the provided statistics are well-balanced in the sense that there are
-very few (or no) extreme values. **Homeruns by batters**, **Strikeouts by** 
-**batters**, and **Homeruns allowed** are examples of this.
+very few (or no) extreme values. **Home Runs by batters**, **Strikeouts by** 
+**batters**, and **Home Runs allowed** are examples of this.
 
@@
 set1 <-  train_set %>%
-  select('Homeruns by batters' = TEAM_BATTING_HR, 
+  select('Home Runs by batters' = TEAM_BATTING_HR, 
          'Strikeouts by batters' = TEAM_BATTING_SO,  
-         'Homeruns allowed' = TEAM_PITCHING_HR) %>%
+         'Home Runs allowed' = TEAM_PITCHING_HR) %>%
   gather("stat", "value") %>%
@@
 
-* **Homeruns allowed** (average = ~100/year) and **Homeruns by batters** (average = ~106/year) 
+* **Home Runs allowed** (average = ~100/year) and **Home Runs by batters** (average = ~106/year) 
 have a very similar mid-range distribution (50% of the data lies between ~50 and 150). 
 The slight difference in average stats means that teams tend to have a higher number
-of homeruns than the opposition team. 
+of Home Runs than the opposition team. 
 * The only thing that stands out about **Strikeouts by batters** (average = ~736/year) is 
@@
 exceptionally high values for both of these statistics skew the dataset. 
-* The maxiumum value
+* The maximum value
 
@@
 ```
+
+## Data Preparation
+
+Now we have a good idea of the data we are looking at we can take the next steps to 
+prepare it for building a solid model.  We saw from the above that at least one 
+variable should be omitted for low reporting of data. There are also outlier concerns 
+for several of the variables, namely Team Pitching Strikeouts (`TEAM_PITCHING_SO`), 
+Team Batting Hits (`TEAM_BATTING_H`) and Team Pitching Hits ('TEAM_PITCHING_H').
+
+First, we will look into obvious outliers in the data set and remove them. There appears 
+to be enough data in the training set to take a naive yet cautious approach to outlier removal.
+
+```{r load, echo=FALSE}
+train_set <-
+  train_set %>%
+  select(-c(TEAM_BATTING_HBP, INDEX)) %>%
+  mutate_if(is.integer, as.numeric)
+```
+
+### Isolate base hits.
+
+The variable Team Batting Hits (`TEAM_BATTING_H`) is a combination of all hits for 
+the season. So we can cleanly use the Second base ('TEAM_BATTING_2B'), Third Base 
+('TEAM_BATTING_3B') and Home Run ('TEAM_BATTING_HR') variables we create a First base variable
+('TEAM_BATTING_1B') by subtracting the 2B, 3B, and HR respectively from the Team Batting Hits total.
+
+~Do we need a graph here?
+
+```{r singles, echo=FALSE}
+train_set<- 
+    train_set %>%
+    mutate(TEAM_BATTING_1B =
+           TEAM_BATTING_H -
+           (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR))
+
+```
+
+### Outlier Removal
+
+When performing the processes of outlier removal a cautious approach is always best.  Each 
+outlier is evaluated to ensure its clearly from incorrectly or misentered data, the removal 
+does not affect our later assumptions, and the outlier creates a significant association or relation 
+at without it the relation is no longer present.
+
+There are some clear issues with two of the variables.  Since both appear with a heavy right skew, we 
+will use median and IQR to remove the outliers.
+
+* TEAM_PITCHING_SO
+  +We discussed earlier Team Pitching Strikeouts.
+* TEAM_PITCHING_H 
+  + Team Pitching Hits has a value over 30,000 and two over 20,000
+  + These values seem highly unlikely and will be adjusted
+
+```{r outliers_1, echo=FALSE}
+train_set %>%
+  pivot_longer(everything(), 
+               names_to ="Variable", 
+               values_to="Value") %>%
+  ggplot(aes(x=Variable, y=Value)) +
+  geom_boxplot(na.rm = TRUE) + coord_flip()
+```
+
+```{r iqr_removal_SO, echo=FALSE}
+train_set<-
+train_set %>%
+  na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
+  bind_cols(
+            train_set 
+            )  %>%
+  filter(
+    TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO,
+                                probs=c(0.25), 
+                                na.rm = TRUE) - 1.5*iqr, 
+    TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, 
+                                probs=c(0.75), 
+                                na.rm=TRUE) + 1.5*iqr
+        ) %>%
+  select(-iqr)
+```
+
+```{r iqr_removal_H, echo=FALSE}
+train_set<-
+train_set %>%
+  na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
+  bind_cols(
+            train_set 
+            )  %>%
+  filter(
+    TEAM_PITCHING_H > quantile(TEAM_PITCHING_H,
+                                probs=c(0.25), 
+                                na.rm = TRUE) - 1.5*iqr, 
+    TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, 
+                                probs=c(0.75), 
+                                na.rm=TRUE) + 1.5*iqr
+        ) %>%
+  select(-iqr)
+
+```
+
+```{r outlier_boxplot, echo=FALSE}
+train_set %>%
+  pivot_longer(everything(), 
+               names_to ="Variable", 
+               values_to="Value") %>%
+  ggplot(aes(x=Variable, y=Value)) +
+  geom_boxplot(na.rm = TRUE) + coord_flip()
+```
+
+This has eliminated what appears to be the most extreme outliers. There are 
+still large outlier sets for Errors and stolen bases, but none that seem to 
+dwarf the other variables. According to the box plot, things appear to be on a 
+scale that seems logical given the data and source of information that we have.
+
+## "MICE" imputation method and Feature Engineering
+
+From our early exploration of the data, the vast majority of the data is complete 
+with only a few variables with missing values. We will use the "mice" imputation method, 
+specifically predictive mean matching method.
+
+MICE (Multivariate imputation by chained equations) is a principled method of dealing 
+with missing data.  It creates multiple imputations, as opposed to single imputations 
+and accounts for the statistical uncertainty in the imputations.  It creates predictive 
+values for the mean instead of imputing the IQR values.
+
+[described here](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3074241/#:~:text=Missing%20data%20are%20a%20common,method%20of%20addressing%20missing%20data.).  
+```{r impute_train,echo=FALSE, message=FALSE, warning=FALSE}
+train_set <- complete(mice(data = train_set, 
+                         method = "pmm", 
+                         seed = 9450, 
+                         print=FALSE), 3)
+```
+
+With outliers removed and missing values imputed via MICE, we can build a feature, 
+specifically one for slugging.
+
+### Slugging Percentage
+
+~Why are we building a feature?
+
+Slugging is calculated in the following way
+
+${\displaystyle \mathrm {SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{AB}}}$
+
+We will take the same approximation for at-bats.
+
+${\displaystyle \mathrm {\sim SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{SO_{batting} + H + BB}}}$
+
+```{r make_slg_obp, echo=FALSE}
+train_set<-
+train_set %>%
+  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
+         SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) %>%
+  # Removing AB and total hits
+  select(-c(AB, TEAM_BATTING_H))
+```
+
+## Modeling
+
+We will start with a simple linear model and select out features based on null hypothesis \
+testing for non-zero slope.
+
+Below is the output of coefficients for backward selection model.
+
+```{r lm_first, echo=FALSE, eval=FALSE}
+lm_reg = lm(data=train_set, TARGET_WINS ~ .)
+summary(lm_reg)
+```
+
+```{r lm_selected, echo=FALSE}
+lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
+          TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
+          TEAM_PITCHING_HR+ TEAM_PITCHING_H+
+          TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_set)
+knitr::kable(lm_1$coefficients, digits = 3L)
+```

### 2020-09-14 HW1/HW1_draft_kellogg.Rmd
Update HW1_draft_kellogg.Rmd
Changed to read csv files from github.

@@
 library(psych)
-train_set <- read.csv('data/moneyball-training-data.csv')
-eval_set <- read.csv('data/moneyball-evaluation-data.csv')
+train_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
+eval_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-evaluation-data.csv')
 

### 2020-09-14 HW1/HW1_draft_kellogg.Rmd
part 1 complete

--- 2020-09-14 HW1/HW1_draft_kellogg.Rmd
@@
 output:
+  pdf_document: default
   html_document: default
-  pdf_document: default
 ---
@@
 library(psych)
-train_set <- read.csv('data/moneyball-training-data.csv')
-eval_set <- read.csv('data/moneyball-evaluation-data.csv')
+library(Hmisc)
+train_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
+eval_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-evaluation-data.csv')
 
@@
 
-## 1. DATA EXPLORATION
+## DATA EXPLORATION
 
@@
 We are given a data set of 2,276 records containing 15 seasonal statistics and the
-total number of wins a team had in a given year. On average, about 50% of games 
-played are won (81 games out of 162), with the best individual season having 146 wins and the worst 
+total number of wins a team had in a given year. On average, about 50% of games played 
+are won (81 games out of 162), with the best individual season having 146 wins and the worst 
 season having 0 wins. The data is normally distributed and most years have between 49 
@@
 ### What's missing? 
-A first look at the data shows that only about 8% of the records have a full set
-of information. The good news is that most of the missing values come from 
-statistics that don't happen too often: hit-by-pitch (92% missing!), caught 
-stealing (34% missing), and double plays (13% missing). Since we have so little 
-hit-by-pitch data, we expect that it doesn't contribute much to overall wins and will 
-eliminate it from a few of the models we propose. The other two stats have less 
-than half of the data missing, so we'll need to think of a clever way to fill in 
-these values. The remaining missing information is from a combination of stolen 
-bases and strikeouts (by batters and pitchers). **It seems completely unreasonable** 
-to have zero strike outs in a season, so this is something we'll most certainly have 
-to impute.
+A first look at the data shows that only about 8% of the records have a full set of 
+information. The good news is that most of the missing values come from statistics 
+that don't happen too often: hit-by-pitch (`TEAM_BATTING_HBP`, 92% missing!), caught 
+stealing (`TEAM_BASERUN_CS`, 34% missing), and double plays (`TEAM_FIELDING_DP`, 13% 
+missing). Since we have so little hit-by-pitch data, we expect that it doesn't contribute 
+much to overall wins and will eliminate it from a few of the models we propose. The 
+other two stats have less than half of the data missing, so we'll need to think of a 
+clever way to fill in these values. The remaining missing information is from a 
+combination of stolen bases and strikeouts by pitchers and batters (`TEAM_BASERUN_SB`,
+`TEAM_PITCHING_SO`, `TEAM_BATTING_SO`). **It seems completely unreasonable** to have zero 
+strike outs in a season, so this is something we'll most certainly have to impute.
 
@@
 
-**Stats with an expected negative impact: ** Intuitively, we expect that Caught 
-stealing, Errors, Hits allowed, Home Runs allowed, Strikeouts by batters, and Walks 
-allowed would all have a **negative** impact on the total wins. In other words, as 
-these values increase, we expect that the team is less likely to win.
+**Stats with an expected negative impact: ** Intuitively, we expect that Caught stealing, 
+Errors, Hits allowed, Home Runs allowed, Strikeouts by batters, and Walks allowed would 
+all have a **negative** impact on the total wins. In other words, as these values increase, 
+we expect that the team is less likely to win.
 
@@
 
-When we take a closer look at the data, these negative relationships aren't obvious. 
-In fact, only Errors and Hits allowed seem to have a negative impact on wins. Caught 
-stealing and Strikeouts by batters appear to be random; this means that whether the 
+When we take a closer look at the data, these negative relationships aren't obvious. In 
+fact, only **Errors** and **Hits allowed** seem to have a negative impact on wins. **Caught**
+**stealing** and **Strikeouts by batters** appear to be random; this means that whether the 
 stat for a particular season is high or low doesn't affect the overall number of wins.
 
-Even more interestingly, Home Runs allowed and Walks allowed have the *opposite* 
+Even more interestingly, **Home Runs allowed** and **Walks allowed** have the *opposite* 
 effect; as these stats increase, so do the number of wins!
@@
 **Stats with an expected positive impact: ** We can look at the same information for the 
-stats that we expect to have a **positive** effect on wins: Base hits, Doubles, Triples, 
-Home Runs, Walks, Batters getting hit by pitches, Stolen bases, Double Plays, and Strikeouts 
-by pitchers.
+stats that we expect to have a **positive** effect on wins: Base hits, Doubles, Triples, Home 
+Runs, Walks, Batters getting hit by pitches, Stolen bases, Double Plays, and Strikeouts by 
+pitchers.
 
@@
 Many of these stats *do* seem to have an effect on the number of wins, most notably,
-base hits and walks. We see weaker positive relationships for home runs, doubles, triples, 
-and stolen bases. This makes sense when we think about it; these things tend to happen 
-less often in games than pure base hits and walks, so they don't have as much of an 
-effect on winning. Finally, double plays and batters hit don't appear to have any 
-correlation with the number of wins. Once again, this intuitively makes sense because 
-they are less likely to happen in a game.
+**Base hits** and **Walks**. We see weaker positive relationships for **Home runs**, 
+**Doubles**, **Triples**, and **Stolen bases**. This makes sense when we think about it; 
+these things tend to happen less often in games than pure base hits and walks, so they don't 
+have as much of an effect on winning. Finally, **Double plays** and **Batters hit** don't 
+appear to have any correlation with the number of wins. Once again, this intuitively makes 
+sense because they are less likely to happen in a game.
 
@@
 
-Before using any of the statistics in a model, we need to take a closer look at the 
-variation in the data. We call out of the ordinary values (exceptionally high or
-low values) **outliers**. We need to take these into account in our modeling because
-we want to make sure our predictions aren't skewed because of them.
+Before using any of the statistics in a model, we need to take a closer look at the variation 
+in the data. We call out-of-the-ordinary values (exceptionally high or low values) **outliers**. 
+We need to take these into account in our modeling because we want to make sure our predictions 
+aren't skewed because of them.
 
-```{r summaryInfo, include = FALSE}
+```{r summaryInfo, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}
 
-summary(train_set)
+#summary(train_set)
+
+totalOutliers <- data.frame(sapply(train_set %>% select(-INDEX, -TARGET_WINS, - TEAM_BATTING_HBP), 
+       function(x){length(boxplot.stats(x)$out)/nrow(train_set)}))
+
+totalOutliers$VARIABLE_NM <- rownames(totalOutliers)
+colnames(totalOutliers) <- c('PCT_OUTLIERS', 'VARIABLE_NM')
+
+ggplot(totalOutliers , aes(x=reorder(VARIABLE_NM, PCT_OUTLIERS), y=PCT_OUTLIERS, label = round(PCT_OUTLIERS, 3))) + 
+  geom_text(vjust = 0.5, hjust = -0.05)+
+  geom_bar(stat = "identity") +
+  ggtitle("Percent Outliers") +
+  xlab("Statistic") + ylab("Percent of Data that is an Outlier") + 
+  coord_flip() +
+  expand_limits(y = 0.15)
+
 ```
 
-Some of the provided statistics are well-balanced in the sense that there are
-very few (or no) extreme values. **Home Runs by batters**, **Strikeouts by** 
-**batters**, and **Home Runs allowed** are examples of this.
+As we can see, some of the provided statistics are well-balanced in the sense that there are very 
+*few* (or no) extreme values. **Home Runs allowed** (`TEAM_PITCHING_HR`), **Strikeouts by batters** 
+(`TEAM_BATTING_SO`), and **Home Runs by batters** (`TEAM_BATTING_HR`) are examples of this.
 
-```{r fewOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
+```{r fewOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2, fig.width = 6, fig.align = 'center'}
 
@@
 
-Conversely, some of the stats have a very high number of outliers, indicating that there
-are some seasons with some abnormally high or low values. **Errors**, **Hits allowed**,
-**Stolen bases**, and **Walks by batters** are examples of this.
+Conversely, some of the stats have a very *high* number of outliers, indicating that there
+are some seasons with some abnormally high or low values. **Errors** (`TEAM_FIELDING_E`), 
+**Hits allowed** (`TEAM_PITCHING_H`), **Walks by batters** (`TEAM_BATTING_BB`), and 
+**Stolen bases** (`TEAM_BASERUN_SB`) are examples of this.
 
-```{r manyOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 6, fig.align = 'center'}
+```{r manyOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2, fig.width = 6, fig.align = 'center'}
 
@@
 
-#sapply(train_set %>% select(-INDEX, -TARGET_WINS), function(x){length(boxplot.stats(x)$out)})
-
 ```
@@
 
-* All of the outliers for **Errors** and **Hits allowed** are above the upper 
-tail of the data set. This is further illustrated by the mean and median values
-for both of these stats; in both instances, the mean per year (Errors = ~246/year
-and Hits allowed = ~1779/year) are higher than the median per year (Errors = 
-~159/year and Hits allowed = ~1518/year). This means that some seasons with
-exceptionally high values for both of these statistics skew the dataset. 
-* The maximum value
+* All of the outliers for **Errors**, **Hits allowed**, and **Stolen bases** are above the upper 
+tail of the data set. This is further illustrated by the mean and median values for these stats; in 
+all instances, the mean per year (Errors = ~246/year, Hits allowed = ~1779/year, Stolen bases = ~125/year) 
+are higher than the median per year (Errors = ~159/year, Hits allowed = ~1518/year, Stolen bases = 
+~101/year). This means that some seasons with exceptionally high values skew the dataset. 
+* There are a few *very* extreme outliers for **Hits allowed**. The maximum value is 30,132, which
+is over 16 times the average number of hits allowed per season!
+* There are outliers both above *and* below the tails of the data for the **Walks by batters** stat. 
+This means that we have exceptionally low (min = 0!) and exceptionally high (max = 878) seasons.
+
+The remaining stats, **Walks allowed** (`TEAM_PITCHING_BB`), **Base Hits by batters** (`TEAM_BATTING_H`), 
+**Caught stealing** (`TEAM_BASERUN_CS`), **Strikeouts by pitchers** (`TEAM_PITCHING_SO`), **Double plays** 
+(`TEAM_FIELDING_DP`), **Triples by batters** (`TEAM_BATTING_3B`), and **Doubles by batters** 
+(`TEAM_BATTING_2B`) have between 29 and 99 outliers.
+
+```{r otherOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7, fig.align = 'center'}
+
+set3 <-  train_set %>%
+  select('Base Hits by batters' = TEAM_BATTING_H, 
+         'Doubles by batters' = TEAM_BATTING_2B,  
+         'Triples by batters' = TEAM_BATTING_3B,
+         'Caught stealing' = TEAM_BASERUN_CS,
+         'Walks allowed' = TEAM_PITCHING_BB,
+         'Strikeouts by pitchers' = TEAM_PITCHING_SO,
+         'Double plays' = TEAM_FIELDING_DP) %>%
+  gather("stat", "value") %>%
+  filter(complete.cases(.) == TRUE)
+
+vals3 <- ggplot(set3 , aes(x=stat, y=value)) + 
+    geom_boxplot() +
+  labs(title="",x="", y = "Value") +
+  facet_wrap(~ stat, scale = 'free', ncol = 4)
+
+vals3
+
+```
+
+Some things to note:
+
+* All variables are very narrowly distributed, meaning that most of the data falls within
+a small range. 
+* **Strikeouts by pitchers** and **Walks allowed** have a few very extreme outliers; these 
+represent seasons that have abnormally high values for the statistics. 
+* The average number of pure **Base Hits** (1469/season) is greater than the average number
+of Doubles (~241/season) and Triples (~55/season). This isn't at all surprising, but serves 
+as a good gut check on the validity of the data.
 
 ### Are stats correlated?
+
+We would expect that a few things in the dataset might be correlated: perhaps number of errors and
+hits/homeruns allowed or the number of base hits by batters and doubles/triples/homeruns. We can
+visualize the correlations between the statistics to determine if there is a significant relationship
+between the them: blue dots represent positively correlated variables (as one increases,
+so does the other) and red dots represent negatively correlated variables (as one increases,
+the other decreases).
+
 ```{r, echo = FALSE, message = FALSE, warning = FALSE}
@@
 # Correlation matrix with significance levels (p-value)
-library("Hmisc")
 res2 <- rcorr(as.matrix(train_set %>% select(-INDEX)))
@@
 
-## Data Preparation
+Some noteworthy relationships (coincidental or not):
+
+* **Errors** are highly, negatively correlated with walks by batters, strikeouts by batters, and
+homeruns (both by batters and allowed). 
+* **Triples by batters** are highly, negatively correlated with strikeouts by batters and homeruns
+(both by batters and allowed).
+* **Strikeouts by batters** are highly, positively correlated with homeruns (both by batters and
+allowed). 
+* **Homeruns by batters** and **Homeruns allowed** are both positively correlated with walks by 
+batters.
+* As expected, **basehits** are positively correlated with doubles and triples.
+
+We can keep these correlations in mind when developing our models: if we have correlated statistics, 
+there could be in-built redundancy in the features, and we may be able to create a simpler, more
+accurate model by eliminating some.
+
+## DATA PREPARATION
 
@@
 for several of the variables, namely Team Pitching Strikeouts (`TEAM_PITCHING_SO`), 
-Team Batting Hits (`TEAM_BATTING_H`) and Team Pitching Hits ('TEAM_PITCHING_H').
+Team Batting Hits (`TEAM_BATTING_H`) and Team Pitching Hits (`TEAM_PITCHING_H`).
 
@@
 
-## "MICE" imputation method and Feature Engineering
+### "MICE" imputation method and Feature Engineering
 
@@
 
-## Modeling
+## BUILD MODELS
 
+### Base Model
 We will start with a simple linear model and select out features based on null hypothesis \
@@
 ```
+
+## SELECT MODELS

### 2020-09-15 HW1/HW1_draft_kellogg.Rmd
ambers updates

@@
 library(Hmisc)
+library(MASS)
+library(dplyr)
+library(caret)
+
 train_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
@@
 
-pct_null <- data.frame(do.call("rbind", map(train_set %>% select(-INDEX), ~mean(is.na(.)))))
+pct_null <- data.frame(do.call("rbind", map(train_set %>% dplyr::select(-INDEX), ~mean(is.na(.)))))
 colnames(pct_null) <- c('PCT_NULL')
@@
   filter(PCT_NULL > 0) %>%
-  select(VARIABLE, PCT_NULL)
+  dplyr::select(VARIABLE, PCT_NULL)
 
@@
 negSet <- train_set %>%
-  select(TARGET_WINS, 
+  dplyr::select(TARGET_WINS, 
          'Errors' = TEAM_FIELDING_E, 
@@
 posSet <- train_set %>%
-  select(TARGET_WINS, 
+  dplyr::select(TARGET_WINS, 
          'Base hits' = TEAM_BATTING_H, 
@@
 
-totalOutliers <- data.frame(sapply(train_set %>% select(-INDEX, -TARGET_WINS, - TEAM_BATTING_HBP), 
+totalOutliers <- data.frame(sapply(train_set %>% dplyr::select(-INDEX, -TARGET_WINS, - TEAM_BATTING_HBP), 
        function(x){length(boxplot.stats(x)$out)/nrow(train_set)}))
@@
 set1 <-  train_set %>%
-  select('Home Runs by batters' = TEAM_BATTING_HR, 
+  dplyr::select('Home Runs by batters' = TEAM_BATTING_HR, 
          'Strikeouts by batters' = TEAM_BATTING_SO,  
@@
 set2 <-  train_set %>%
-  select('Walks by batters' = TEAM_BATTING_BB, 
+  dplyr::select('Walks by batters' = TEAM_BATTING_BB, 
          'Stolen bases' = TEAM_BASERUN_SB,  
@@
 set3 <-  train_set %>%
-  select('Base Hits by batters' = TEAM_BATTING_H, 
+  dplyr::select('Base Hits by batters' = TEAM_BATTING_H, 
          'Doubles by batters' = TEAM_BATTING_2B,  
@@
 
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
+```{r corrPlot, echo = FALSE, message = FALSE, warning = FALSE}
 
 # Correlation matrix with significance levels (p-value)
-res2 <- rcorr(as.matrix(train_set %>% select(-INDEX)))
+res2 <- rcorr(as.matrix(train_set %>% dplyr::select(-INDEX)))
 
@@
 Now we have a good idea of the data we are looking at we can take the next steps to 
-prepare it for building a solid model.  We saw from the above that at least one 
-variable should be omitted for low reporting of data. There are also outlier concerns 
-for several of the variables, namely Team Pitching Strikeouts (`TEAM_PITCHING_SO`), 
-Team Batting Hits (`TEAM_BATTING_H`) and Team Pitching Hits (`TEAM_PITCHING_H`).
-
-First, we will look into obvious outliers in the data set and remove them. There appears 
-to be enough data in the training set to take a naive yet cautious approach to outlier removal.
-
-```{r load, echo=FALSE}
-train_set <-
-  train_set %>%
-  select(-c(TEAM_BATTING_HBP, INDEX)) %>%
-  mutate_if(is.integer, as.numeric)
-```
-
-### Isolate base hits.
-
-The variable Team Batting Hits (`TEAM_BATTING_H`) is a combination of all hits for 
-the season. So we can cleanly use the Second base ('TEAM_BATTING_2B'), Third Base 
-('TEAM_BATTING_3B') and Home Run ('TEAM_BATTING_HR') variables we create a First base variable
-('TEAM_BATTING_1B') by subtracting the 2B, 3B, and HR respectively from the Team Batting Hits total.
-
-~Do we need a graph here?
-
-```{r singles, echo=FALSE}
-train_set<- 
-    train_set %>%
-    mutate(TEAM_BATTING_1B =
-           TEAM_BATTING_H -
-           (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR))
-
-```
+prepare it for building a solid model.
 
@@
 
-When performing the processes of outlier removal a cautious approach is always best.  Each 
-outlier is evaluated to ensure its clearly from incorrectly or misentered data, the removal 
-does not affect our later assumptions, and the outlier creates a significant association or relation 
-at without it the relation is no longer present.
+As we saw in the data analysis, there are some outlier concerns for some of the variables,
+so we will need to account for this in our modeling. When performing the processes of outlier 
+removal, a cautious approach is always best. Each outlier is evaluated to ensure:
 
-There are some clear issues with two of the variables.  Since both appear with a heavy right skew, we 
-will use median and IQR to remove the outliers.
+* It is clearly from incorrectly or mis-centered data.
+* Its removal does not affect later assumptions.
+* It creates a significant association/relation that is eliminated with its removal.
 
-* TEAM_PITCHING_SO
-  +We discussed earlier Team Pitching Strikeouts.
-* TEAM_PITCHING_H 
-  + Team Pitching Hits has a value over 30,000 and two over 20,000
-  + These values seem highly unlikely and will be adjusted
+We can take another look at the outliers for each variable:
 
-```{r outliers_1, echo=FALSE}
+```{r outliers_1, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 8, fig.align = 'center'}
 train_set %>%
+  dplyr::select(-TARGET_WINS, -INDEX) %>%
   pivot_longer(everything(), 
@@
   ggplot(aes(x=Variable, y=Value)) +
-  geom_boxplot(na.rm = TRUE) + coord_flip()
+  geom_boxplot(na.rm = TRUE) + 
+  labs(title="Outliers for each Statistic",x="Variable", y = "Value") +
+  coord_flip()
 ```
 
-```{r iqr_removal_SO, echo=FALSE}
-train_set<-
-train_set %>%
+Off the bat, there are some clear issues with two of the variables: `TEAM_PITCHING_SO` and
+`TEAM_PITCHING_H.`
+
+* **TEAM_PITCHING_SO**: The largest outlier is close to 20,000, which averages ~123 strikeouts
+per game. The MLB team pitching strikeout record is 1,450 and set in 2014 (CLE). There are a 
+minimum of 27 at-bats per game and most teams average ~35 at-bats, thus making 123 strikeouts
+impossible. 
+* **TEAM_PITCHING_H**: The largest outlier for this vaiable is over 30,000 which is highly 
+unlikely.
+
+Since both appear with a heavy right skew, we will use median and IQR to remove the outliers.
+
+```{r iqr_removal_SO, include=FALSE}
+train_set<- train_set %>%
   na.omit() %>%
@@
         ) %>%
-  select(-iqr)
+  dplyr::select(-iqr)
 ```
 
-```{r iqr_removal_H, echo=FALSE}
-train_set<-
-train_set %>%
+```{r iqr_removal_H, include=FALSE}
+train_set<- train_set %>%
   na.omit() %>%
@@
         ) %>%
-  select(-iqr)
+  dplyr::select(-iqr)
 
@@
 
-```{r outlier_boxplot, echo=FALSE}
+We can take a look at the distributions of these variables after outlier removal:
+
+```{r outlier_boxplot, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 8, fig.align = 'center'}
 train_set %>%
+  dplyr::select(-TARGET_WINS, -INDEX) %>%
   pivot_longer(everything(), 
@@
   ggplot(aes(x=Variable, y=Value)) +
-  geom_boxplot(na.rm = TRUE) + coord_flip()
+  geom_boxplot(na.rm = TRUE) + 
+  labs(title="Outliers for each Statistic",x="Variable", y = "Value") +
+  coord_flip()
 ```
@@
 
-### "MICE" imputation method and Feature Engineering
+### "MICE" imputation method
 
@@
 
-[described here](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3074241/#:~:text=Missing%20data%20are%20a%20common,method%20of%20addressing%20missing%20data.).  
 ```{r impute_train,echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-With outliers removed and missing values imputed via MICE, we can build a feature, 
-specifically one for slugging.
+### Feature Engineering: Single base hits
 
-### Slugging Percentage
+With outliers removed and missing values imputed via MICE, we can create a few new 
+variables. The first will be single base hits, and it will be derived from some of the
+variables we do have. We know that Team Batting Hits (`TEAM_BATTING_H`) is a combination 
+of *all* hits for the season, so we can create Single Base Hits as follows:
 
-~Why are we building a feature?
+$$Singles = Total\ Hits - (Doubles + Triples + Homeruns)$$
 
-Slugging is calculated in the following way
+```{r singles, include=FALSE}
 
-${\displaystyle \mathrm {SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{AB}}}$
+train_set<- train_set %>%
+  dplyr::select(-INDEX) %>%
+  mutate_if(is.integer, as.numeric) %>%
+  mutate(TEAM_BATTING_1B =
+           TEAM_BATTING_H -
+           (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR))
 
-We will take the same approximation for at-bats.
+```
 
-${\displaystyle \mathrm {\sim SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{SO_{batting} + H + BB}}}$
+### Feature Engineering: Slugging Percentage
 
-```{r make_slg_obp, echo=FALSE}
-train_set<-
-train_set %>%
+The second variable we will create is **Slugging Percentage**. Slugging is an offensive 
+statistic that is a good predictor of winning and it tends to have better variance and
+correlation than most other variables. It is composed of singles, doubles, triples, 
+home runs and at-bats.
+
+Slugging is calculated in the following way:
+
+$${\displaystyle \mathrm {SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{AB}}}$$
+
+Because we don't have a statistic for at-bats, we will approximate it as follows:
+
+$${\displaystyle \mathrm {\sim SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{SO_{batting} + H + BB}}}$$
+
+```{r make_slg_obp, include=FALSE}
+train_set<- train_set %>%
   mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
-         SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) %>%
-  # Removing AB and total hits
-  select(-c(AB, TEAM_BATTING_H))
+         SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) 
+```
+
+As a gut-check, we can take a look at Slugging in comparison to Triples (`TEAM_BATTING_3B`) 
+and Homeruns (`TEAM_BATTING_HR`). All of these variables should have a positive correlation 
+with the total number of wins.
+
+```{r slugging_compare, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 8, fig.align = 'center'}
+
+modified_set <- train_set %>%
+  dplyr::select(TARGET_WINS, SLG, TEAM_BATTING_3B, TEAM_BATTING_HR)
+
+ggplot(data = modified_set %>%
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
+  facet_wrap(~ STAT, scales = "free")
 ```
@@
 ### Base Model
-We will start with a simple linear model and select out features based on null hypothesis \
-testing for non-zero slope.
+We will start with a simple linear model to serve as a baseline. This includes all variables 
+in the dataset.
 
-Below is the output of coefficients for backward selection model.
-
-```{r lm_first, echo=FALSE, eval=FALSE}
+```{r baseModel, include = FALSE}
 lm_reg = lm(data=train_set, TARGET_WINS ~ .)
@@
 
-```{r lm_selected, echo=FALSE}
+### Model 1
+Model 1 includes all variables except Singles, Triples, Base Hits, Strikeouts by
+pitchers, Walks allowed, and Batters Hit by Pitch. The final variables were chosen as 
+a result of backwards dplyr::selection based on null hypothesis testing for non-zero slope.
+
+We can take a look at the final coefficients for each variable.
+
+**Coefficient Discussion: **
+```{r model1, echo=FALSE}
 lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
@@
 
+First, we are keeping the coefficient for `TEAM_BATTING_HR` since it's p-value is marginally above the general threshold and knowledge of the game suggests it is important.
+There are some counter intuitive results, which is expected given that baseball is a messy, imprecise game that has evolved over time.
+To that end, we should expect some seemingly strange results from algorithmic regression.
+We used a combination of domain knowledge and data analysis to justify retaining features.
+
+Fielding double plays and batting doubles both appear to have negative impacts on wins even though they _should_ be postive impacts.
+Turning double plays, while a good for the defensive team, may suggest a larger, negative issue.
+Namely, weak pitching that leads to runners on base.
+Similarly, batting doubles, leaves runners open to double plays.
+Allowed hits by pitching, caught stealing, and batting strike-outs are all counter intuitive results as well, but they seem to be small contributors and these are events that happen regularly in every game.
+
+Slugging as an approximation is the major predictor in this regression, by far.
+The other predictors other than the intercept are orders of magnitude less in predictive value.
+It should be noted that slugging alone is not a good predictor for wins overall.
+
+**Further data transformation: **  
+We will use Cook's distance to remove outliers that are influencing the fit of the model above. We will use a cutoff of $\frac{4}{N}$.
+
+```{r cooks_distance, include=FALSE}
+cooks_dis<- cooks.distance(lm_1)
+influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4/nrow(train_set)))])
+train_df2<- train_set[-influential, ]
+```
+
+### Model 2
+This is a model allowing pairwise interactions within the data types of baserunning,
+batting, pitching, and fielding, fit using a forward and backwards stepwise regression.
+
+```{r model2, include = FALSE}
+
+crazyModel <- lm(TARGET_WINS ~
+  (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
+  TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_SO) ^ 2 +
+  (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
+  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = train_set)
+
+stepModel <- stepAIC(crazyModel, scope = list(upper = ~
+  (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
+  TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_SO) ^ 2 +
+  (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
+  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, lower = ~ 1),
+  direction = 'both')
+
+sstm <- summary(stepModel)
+stmr2 <- prettyNum(sstm$r.squared, digits = 3L)
+stmar2 <- prettyNum(sstm$adj.r.squared, digits = 3L)
+stmAIC <- prettyNum(AIC(stepModel), digits = 2L, big.mark = ',')
+stmRMSE <- prettyNum(sqrt(mean(stepModel$residuals ^ 2)), digits = 5L)
+stmF <- prettyNum(sstm$fstatistic, digits = 5L)
+stmFp <- pf(sstm$fstatistic[[1]], sstm$fstatistic[[2]], sstm$fstatistic[[3]],
+            lower.tail = FALSE)
+```
+
+**Coefficient Discussion: **
+```{r model2_discuss, echo = FALSE, message = FALSE, warning = FALSE,}
+knitr::kable(sstm$coefficients, digits = 3L,
+             caption = 'Step Model with Pairwise Class Interactions Output')
+
+```
+
+Most of the coefficients are reasonable within the context of the game of
+baseball. These two statements are axiomatic:
+
+  * The only way to win is to have the higher score at the end of the game.
+  * The only way to score is for a baserunner to cross home plate.
+
+With those in mind, we can make the following observations about the linear
+non-interactive coefficients.
+
+ * Reasonable
+   * Caught stealing removes baserunners; negative coefficient makes sense.
+   * Stolen bases get a runner closer to home plate; positive coefficient makes
+   sense.
+   * Triples get a runner very close to home plate; positive coefficient makes
+   sense.
+   * Walks are free baserunners; positive coefficient makes sense.
+   * Hits increase the number of baserunners; positive coefficient makes sense.
+   * Hitting home runs directly increase the score; positive coefficient makes
+   sense.
+   * Striking out reduces the number of baserunners; negative coefficient makes
+   sense.
+   * Errors allow the other team free baserunners; negative coefficient makes
+   sense.
+   * Giving up home runs gives the opponent scores; negative coefficient makes
+   sense.
+   * Getting strikeouts reduces the opponents baserunners; positive coefficient
+   makes sense.
+ * Curious
+   * Hitting doubles get runners on base; why negative?
+   * Turning double plays reduce baserunners; why negative?
+   * Giving up walks allows the other team free baserunners; negative
+   coefficient makes sense but why insignificant?
+   * Giving up hits allows the other team baserunners; why positive although
+   insignificant?
+
+An absolutely fascinating observation. In the first draft of this exercise,
+`TEAM_BATTING_HR` was removed due to its high correlation with
+`TEAM_PITCHING_HR`. Note the coefficients, batting is almost +12 and pitching is
+-9. However, this is such as strong indicator, that in the first run, pitching
+was given a *positive* coefficient of around +3. In hindsight this is because
+*it was being used as an indicator for* ***batting!!*** The correlation allowed
+the use of pitching HRs as an indicator for the hidden batting HRs! Once both
+were restored to the model, the logical coefficients surfaced. Another reason
+why models should not be trusted out of the box, but all model results should
+be reviewed for sanity and sense!!
+
+The curiosities above can *possibly* be resolved by looking at the interaction
+terms. The interaction between giving up hits and strikeouts,
+`TEAM_PITCHING_H:TEAM_PITCHING_SO `, is highly negative and significant. It
+probably captures most of the giving up hits information making the singleton
+less relevant.
+
+A possible explanation for the negative coefficient for getting double plays, is
+that double plays require at least two people on base. That means that the
+opponent has a lot of base runners, which is very highly correlated with
+scoring.
+
+The behavior of doubles remains confusing. Unless its hiding something like
+a team's propensity to strand runners on base. It would be interesting to see
+a breakdown between the American and National leagues, as the latter tends to be
+somewhat better at "small ball" and moving the runners along.
+
+### Model 3
+The final model is a Stepwise Regression with Repeated k-fold Cross-Validation.
+
+A stepwise variable selection model is conducted to determine what are the variables that can help predict the number of wins for the team.
+The stepwise variable selection allows variables to be added one at a time to the model, as long as the F-statistic is below the specified $\alpha$, in this case $\alpha = 0.05$.
+However, variables already in the model do not necessarily stay in.
+The steps evaluate all of the variables already included in the model and remove any variable that has an insignificant F-statistic.
+Only after this test ends, is the best model found, that is when none of the variables can be excluded and every variable included in the model is significant.
+
+Here the dependent variable is the continuous variable, `TARGET_WINS`, and the independent variables are the full model to identify the most contributing predictors.
+In addition, a robust method for estimating the accuracy of a model, the k-fold cross-validation method, was performed evaluate the model performance on different subset of the training data and then calculate the average prediction error rate.
+
+```{r model3, include = FALSE}
+
+# Data Transformation & Multicollinearity Diagnostic Measures
+train_df.clean = subset(train_set, select = -c(TEAM_PITCHING_HR, TEAM_BATTING_H, TEAM_BATTING_HBP))
+
+stepdata = cbind(train_df.clean,sapply(train_df.clean[2:length(train_df.clean)], 
+                                       function(x) x^2))
+names(stepdata) = make.unique(names(stepdata), sep = "_")
+train.control = trainControl(method = "cv", number = 10)
+
+# Train the model
+step.model = train(TARGET_WINS ~ ., data = stepdata, method = "lmStepAIC", 
+                   trControl = train.control, trace = FALSE)
+
+# Final model coefficients
+step.model$finalModel
+
+```
+
+**Coefficient Discussion: **
+```{r model3_discuss, echo=FALSE}
+step.model$finalModel[["coefficients"]] %>% knitr::kable()
+```
+
+Studying the coefficients of the model suggest that winning is in favor if the team batting hits more doubles, triples and home runs.
+Moreover, increase in the number of stolen bases, and a decrease in caught steals, double plays, error, and walks allowed would all lead to a win for the batting team.
+It is noteworthy that the model suggests that a decrease in single hits by batter and an increase in strikeouts by batters which seems counter intuitive.
+But these variables were kept because when a batter steps to the plate, the player is more likely to strike out than to get a hit.
+Trying to hit the ball out of the park will come with strikeouts but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry out.
+
 ## SELECT MODELS

### 2020-09-15 HW1/HW1_draft_kellogg.Rmd
Minor updates by Samantha

@@
 library(caret)
-
 train_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
 eval_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-evaluation-data.csv')
-
 set.seed(9450)
@@
 ```{r wins, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
-
 avgWins <- mean(train_set$TARGET_WINS)
@@
 l2 <- mean(train_set$TARGET_WINS) + (2*sdev)
-
 # histogram of wins
@@
   labs(title="Distribution of Wins",x="Number of wins", y = "Number of seasons")
-
 p
@@
 ```{r pctNull, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2, fig.width = 5, fig.align = 'center'}
-
 pct_null <- data.frame(do.call("rbind", map(train_set %>% dplyr::select(-INDEX), ~mean(is.na(.)))))
 colnames(pct_null) <- c('PCT_NULL')
-
 totalNulls <- pct_null %>%
@@
   dplyr::select(VARIABLE, PCT_NULL)
-
 ggplot(totalNulls, aes(x=reorder(VARIABLE, PCT_NULL), y=PCT_NULL, label = round(PCT_NULL, 2))) + 
@@
   expand_limits(y = 1)
-
 ```
@@
 ```{r negFactors, echo = FALSE, message = FALSE, warning = FALSE}
-
 negSet <- train_set %>%
@@
          'Home Runs allowed' = TEAM_PITCHING_HR)
-
 ggplot(data = negSet %>%
@@
 ```{r posFactors, echo = FALSE, message = FALSE, warning = FALSE}
-
 posSet <- train_set %>%
@@
          'Strikeouts' = TEAM_PITCHING_SO)
-
 ggplot(data = posSet %>%
@@
   geom_smooth(method = "lm", se = FALSE) +
-  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
+  labs(title="Number of Wins with Respect to Recorded Stats",x="", y = "Number of Wins") +
   facet_wrap(~ STAT, scales = "free")
-
 ```
@@
 ```{r summaryInfo, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}
-
 #summary(train_set)
-
 totalOutliers <- data.frame(sapply(train_set %>% dplyr::select(-INDEX, -TARGET_WINS, - TEAM_BATTING_HBP), 
        function(x){length(boxplot.stats(x)$out)/nrow(train_set)}))
-
 totalOutliers$VARIABLE_NM <- rownames(totalOutliers)
 colnames(totalOutliers) <- c('PCT_OUTLIERS', 'VARIABLE_NM')
-
 ggplot(totalOutliers , aes(x=reorder(VARIABLE_NM, PCT_OUTLIERS), y=PCT_OUTLIERS, label = round(PCT_OUTLIERS, 3))) + 
@@
   geom_bar(stat = "identity") +
-  ggtitle("Percent Outliers") +
+  ggtitle("Percentage of Outliers") +
   xlab("Statistic") + ylab("Percent of Data that is an Outlier") + 
@@
   expand_limits(y = 0.15)
-
 ```
@@
 ```{r fewOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2, fig.width = 6, fig.align = 'center'}
-
 set1 <-  train_set %>%
@@
   filter(complete.cases(.) == TRUE)
-
 vals <- ggplot(set1 , aes(x=stat, y=value)) + 
@@
   facet_wrap(~stat, scale="free")
-
 vals
-
 ```
@@
 ```{r manyOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2, fig.width = 6, fig.align = 'center'}
-
 set2 <-  train_set %>%
@@
   filter(complete.cases(.) == TRUE)
-
 vals2 <- ggplot(set2 , aes(x=stat, y=value)) + 
@@
   facet_wrap(~ stat, scale = 'free', ncol = 4)
-
 vals2
-
 ```
@@
 ```{r otherOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7, fig.align = 'center'}
-
 set3 <-  train_set %>%
@@
   filter(complete.cases(.) == TRUE)
-
 vals3 <- ggplot(set3 , aes(x=stat, y=value)) + 
@@
   facet_wrap(~ stat, scale = 'free', ncol = 4)
-
 vals3
-
 ```
@@
 ```{r corrPlot, echo = FALSE, message = FALSE, warning = FALSE}
-
 # Correlation matrix with significance levels (p-value)
 res2 <- rcorr(as.matrix(train_set %>% dplyr::select(-INDEX)))
-
 # Insignificant correlation are crossed
@@
           tl.col = "black", p.mat = res2$P, sig.level = 0.01, insig = "blank")
-
 ```
@@
   dplyr::select(-iqr)
-
 ```
@@
 ```{r singles, include=FALSE}
-
 train_set<- train_set %>%
@@
            (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR))
-
 ```
@@
 ```{r slugging_compare, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 8, fig.align = 'center'}
-
 modified_set <- train_set %>%
   dplyr::select(TARGET_WINS, SLG, TEAM_BATTING_3B, TEAM_BATTING_HR)
-
 ggplot(data = modified_set %>%
@@
           TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_set)
-knitr::kable(lm_1$coefficients, digits = 3L)
+knitr::kable(summary(lm_1)$coefficients, digits = 3L, caption = 'Backwards Model Regression Output')
 ```
@@
 
-Fielding double plays and batting doubles both appear to have negative impacts on wins even though they _should_ be postive impacts.
+Fielding double plays and batting doubles both appear to have negative impacts on wins even though they _should_ be positive impacts.
 Turning double plays, while a good for the defensive team, may suggest a larger, negative issue.
@@
 ```{r model2, include = FALSE}
-
 crazyModel <- lm(TARGET_WINS ~
@@
   TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = train_set)
-
 stepModel <- stepAIC(crazyModel, scope = list(upper = ~
@@
   direction = 'both')
-
 sstm <- summary(stepModel)
-stmr2 <- prettyNum(sstm$r.squared, digits = 3L)
-stmar2 <- prettyNum(sstm$adj.r.squared, digits = 3L)
+stmr2 <- sstm$r.squared
+stmar2 <- sstm$adj.r.squared
 stmAIC <- prettyNum(AIC(stepModel), digits = 2L, big.mark = ',')
-stmRMSE <- prettyNum(sqrt(mean(stepModel$residuals ^ 2)), digits = 5L)
+stmRMSE <- sqrt(mean(stepModel$residuals ^ 2))
 stmF <- prettyNum(sstm$fstatistic, digits = 5L)
@@
              caption = 'Step Model with Pairwise Class Interactions Output')
-
 ```
@@
 ### Model 3
-The final model is a Stepwise Regression with Repeated k-fold Cross-Validation.
+The final model is a Stepwise Regression with Repeated k-fold Cross-Validation, and higher order polynomials variables were introduced into the full model.
 
 A stepwise variable selection model is conducted to determine what are the variables that can help predict the number of wins for the team.
-The stepwise variable selection allows variables to be added one at a time to the model, as long as the F-statistic is below the specified $\alpha$, in this case $\alpha = 0.05$.
+The method allows variables to be added one at a time to the model, as long as the F-statistic is below the specified $\alpha$, in this case $\alpha = 0.05$.
 However, variables already in the model do not necessarily stay in.
@@
 
-Here the dependent variable is the continuous variable, `TARGET_WINS`, and the independent variables are the full model to identify the most contributing predictors.
+Here, the dependent variable is the continuous variable, `TARGET_WINS`, and the independent variables are the full model to identify the most contributing predictors.
 In addition, a robust method for estimating the accuracy of a model, the k-fold cross-validation method, was performed evaluate the model performance on different subset of the training data and then calculate the average prediction error rate.
@@
 ```{r model3, include = FALSE}
-
 # Data Transformation & Multicollinearity Diagnostic Measures
 train_df.clean = subset(train_set, select = -c(TEAM_PITCHING_HR, TEAM_BATTING_H, TEAM_BATTING_HBP))
-
 stepdata = cbind(train_df.clean,sapply(train_df.clean[2:length(train_df.clean)], 
@@
 train.control = trainControl(method = "cv", number = 10)
-
 # Train the model
@@
                    trControl = train.control, trace = FALSE)
-
 # Final model coefficients
 step.model$finalModel 
-
+# Summary of the model
+mod_sum = summary(step.model$finalModel)
 ```
@@
 ```{r model3_discuss, echo=FALSE}
-step.model$finalModel[["coefficients"]] %>% knitr::kable()
+mod_sum[["coefficients"]] %>% knitr::kable(digits = 3L, caption = 'K-fold Step Model with Higher Order Polynomials Output')
 ```
@@
 
-## SELECT MODELS
+## SELECTING A MODEL
+
+In order to select the best model to make predictions, we looked at some measurements that tells us how well each model fits.
+These include the 1) $R^2$, which represents the proportion of the variance explained by a model; 2) $adj R^2$, which is a modified version of $R^2$; 3) *Root Mean Squared Error* (RMSE), which is the square root of the mean squared error; and 4) *Akaike Information Criterion* (AIC), which is an estimator of out-of-sample prediction error.
+
+It is apparent that Model #2: Pairwise accounts for nearly 43% of the variation in the dependent variable with the independent variables, which is acceptable as a good model ($R^2 = 0.427$, $adjR^2 = 0.417$.
+Because the RMSE and adjusted $R^2$ statistics already include a minor adjustment for the number of coefficients estimated, to evaluate the model complexity, we compared the AIC.
+While a smaller AIC is deemed less complex, the difference between Model #2 and Model #3 is quite insignificant.
+
+```{r echo=FALSE}
+# Predictions on training data
+mod_sum1 = summary(lm_1)  
+pred_lm1 = predict(lm_1, newdata = train_set)
+pred_step.model = predict(step.model, newdata = stepdata)
+
+# Model accuracy
+results = data.frame(Models = c("Model #1: Backwards" , "Model #2: Pairwise", "Model #3: K-fold"),
+                R.squared = c(mod_sum1[["r.squared"]], stmr2, mod_sum[["r.squared"]]), 
+                adj.R.squared = c(mod_sum[["adj.r.squared"]], stmar2, mod_sum[["adj.r.squared"]]),
+                AIC = c(AIC(lm_1), AIC(stepModel), AIC(step.model$finalModel)),
+                RSME = c(sqrt(mean((train_set$TARGET_WINS - pred_lm1)^2)), stmRMSE,
+                         sqrt(mean((stepdata$TARGET_WINS - pred_step.model)^2))))
+results %>% knitr::kable(digits = 3L, caption = 'Performance Statistics & Error Measure of Models' )
+```
+
+## PREDICTIONS
+
+Using the test data and the selected final model, a comparison in the prediction statistic was conducted.
+
+```{r include=FALSE, eval=FALSE}
+# Data Transformation to match train data
+eval_set = complete(mice(data = eval_set, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
+
+# Add final model
+predictions = predict(ADD-MODEL-HERE, newdata = eval_set)
+
+results = cbind(dataset = c('Training Data', 'Test Prediction'), 
+                bind_rows(describe(train_set$TARGET_WINS), describe(predictions))) 
+rownames(results) = NULL
+
+results[,-c(2,8,11)] %>% knitr::kable(digits = 3L, caption = 'Prediction Comparison with Final Model')
+```
+
+## CONCLUSIONS

### 2020-09-16 Adler_personal/Blog1.Rmd
First Blog and stopped one which is going nowhere

---
title: "Blog Post 1: Orthogonal Regression"
subtitle: "DT 621---Fall 2020"
author: "Avraham Adler"
date: "9/15/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# What is Orthogonal Regression
Orthogonal regression is when the error between the observations, \(y_i\) and
the regression line is not measured solely along the y-axis (vertically) but 
perpendicularly---orthogonally---to the regreression line. The "error" is split
evenly between the x-axis and the y-axis. This is a special case of
[Deming regression](https://en.wikipedia.org/wiki/Deming_regression), where the
relationship between the x and y error is a constant reflecting the ratio of the
variance of their errors, not necessarily equal to one as it is in orthogonal
regression.

# Why Use Orthogonal Regression?
Orthogonal and Deming regressions are special cases of
[errors-in-variables models](https://en.wikipedia.org/wiki/Errors-in-variables_models).
These are models where there is explicit recognition that the *independant*
variables are subject to error as well. Perhaps a tool was faulty, perhaps we
ran into Heisenberg's uncertainty principle and we **cannot** measure precisely.
Regardless, these kinds of models help correct for the attenuation bias that
results when independent variables have measurement error.

# Example
Let's assume that the true relationship between \(X\) and \(Y\) is:
\[
y_i = 3.1 + 6x_i + \epsilon_i\\
\epsilon \sim \mathcal{N}(0, 2)
\]

In case one, we will generate data assuming we know the \(x\)'s exactly. In case
two, we will generate white noise to add to each \(x\), representing the
measurement error. We will then compare the results of OLS with orthogonal least
squares.

```{r createData, fig.width=8}
set.seed(33)
n <- 15
x <- rnorm(n, 0, 1)
y <- 3.1 + 6 * x + rnorm(n, 0, 2)
x2 <- x + rnorm(n, 0, 2)
par(mfrow = c(1, 2))
plot(x2, y, col = 'red', xlim = c(min(x, x2), max(x, x2)),
                                  ylim = c(min(y), max(y)),
     main = "Observations With Measurment Error")
plot(x, y, xlim = c(min(x, x2), max(x, x2)), ylim = c(min(y), max(y)),
     main = "Observations w/o Measurment Error")
```

Clearly, the measurement error will inducing attenuation error in the model. We
will now regress the observed \(y\)'s on the **observed** \(x\)'s.

```{r model, fig.width=8}
set.seed(33)
library(deming)
fit1 <- lm(y ~ x2)
fit2 <- deming(y ~ x2)
Tfit <- lm(y ~ x)
summary(fit1)
fit2
plot(x2, y, col = 'red', xlim = c(min(x, x2), max(x, x2)),
                                  ylim = c(min(y), max(y)),
     main = "Observations With Measurment Error")
abline(fit1$coefficients[[1]], fit1$coefficients[[2]], col = 'red', lty = 2)
abline(fit2$coefficients[[1]], fit2$coefficients[[2]], col = 'darkgreen', lty = 2)
abline(Tfit$coefficients[[1]], Tfit$coefficients[[2]], col = 'black', lty = 2)
```

The red line is the linear regression assuming the measured \(x\)'s have no
error. The black line is the regression through the "true" value of the \(x\)'s.
The green line is the orthogonal regression line. While we were unable to fully
recover the true regression line, using orthogonal regression did address the
attenuation error, and shifted the slope more towards the actual than to 0.

### 2020-09-16 Adler_personal/BlogZ.Rmd
First Blog and stopped one which is going nowhere

---
title: "Blog Post 1: The Use of the Offset"
subtitle: "DT 621---Fall 2020"
author: "Avraham Adler"
date: "9/13/2020"
output: html_document
bibliography: 'D:\\Users\\Parents\\Documents\\SpiderOak\\LaTeX\\LaTeX\\BibTeX\\AviRefsBiblatex.bib'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
In my area of expertise, non-life reinsurance, there is not that much call for
linear regression. However, in primary insurance, generalized linear models
(GLM) are used very often to set rates and class plans. A common example is in
personal auto insurance, where there is a lot of driver data, and models are fit
to features such as marital status, sex, age, and location.

Focusing on auto physical damage, the loss an insurer pays for first party
property damage to the car (when the driver hits another car), the expected
losses are a function of both frequency and severity. Severity is a function of
the value of car being driven. Frequency, however, depends on the driver. Often,
a GLM is used to estimate frequency.

As accidents, at least at first glance, can be considered a Poisson process, a
GLM with a Poisson link function is very common. As frequency of events is
clearly a function of the number of drivers, the models are often constructed to
measure the rate of accident per unit of exposure, where the exposure could be
drivers, cars, miles driven, or car-years insured. It is in this context I first
came across offsets in 
@Yan.etal-2009-ApplicationsOfTheOffsetInPropertyCasualtyPredictiveModeling. The
authors describe using an offset in a Poisson model to measure the rate instead
of the amount.

In OLS, which were are studying, the use of the offset term is less common, for
one can simply subtract the offset from the observations and proceed. But in
non-additive models,---the Poisson is multiplicative---it has a larger role.
Thus, I decided to take this opportunity compare running a Poisson model with
and without an offset and compare them.

# Create Synthetic Data
## True Regression Model
Let's create data where the frequency of accidents is a Poisson process with the
following rates per unit of exposure:

|Variable|Value|
|--------|-----|
|Grand Mean|0.01|
|Male|.10|
|Married| -0.08|
|Region: NE|0.05|
|Region: S|0.1|
|Region: NC (North Central)|0|
|Region: W| -0.03|

So the

Let's generate 5 years worth of exposures as a Poisson random variable with mean
of 100, and calculate

# References

### 2020-09-16 HW1/HW1_draft_kellogg.Rmd
fix model 1 training set

@@
 
-We can take a look at the final coefficients for each variable.
+**Further data transformation: **  
+We will use Cook's distance to remove outliers that are influencing the fit of the model above. We will use a cutoff of $\frac{4}{N}$.
 
-**Coefficient Discussion: **
-```{r model1, echo=FALSE}
+```{r cooks_distance, include=TRUE}
 lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
@@
           TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_set)
+
+cooks_dis<- cooks.distance(lm_1)
+influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4/nrow(train_set)))])
+train_df2<- train_set[-influential, ]
+
+lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
+          TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
+          TEAM_PITCHING_HR+ TEAM_PITCHING_H+
+          TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_df2)
 knitr::kable(summary(lm_1)$coefficients, digits = 3L, caption = 'Backwards Model Regression Output')
@@
 
+**Coefficient Discussion: **
 First, we are keeping the coefficient for `TEAM_BATTING_HR` since it's p-value is marginally above the general threshold and knowledge of the game suggests it is important.
@@
 It should be noted that slugging alone is not a good predictor for wins overall.
-
-**Further data transformation: **  
-We will use Cook's distance to remove outliers that are influencing the fit of the model above. We will use a cutoff of $\frac{4}{N}$.
-
-```{r cooks_distance, include=FALSE}
-cooks_dis<- cooks.distance(lm_1)
-influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4/nrow(train_set)))])
-train_df2<- train_set[-influential, ]
-```
 

### 2020-09-16 HW1/HW1_draft_kellogg.Rmd
updated selection model

@@
           TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_set)
-
 cooks_dis<- cooks.distance(lm_1)
@@
 train_df2<- train_set[-influential, ]
-
 lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
@@
           TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_df2)
+
+mod_sum1 = summary(lm_1)  
+pred_lm1 = predict(lm_1, newdata = train_df2)
 knitr::kable(summary(lm_1)$coefficients, digits = 3L, caption = 'Backwards Model Regression Output')
@@
 ```{r model3, include = FALSE}
+train_df = read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
+train_df = subset(train_df, select = -c(INDEX, TEAM_BATTING_HBP))
+model = lm(TARGET_WINS ~ ., data = train_df)
+# Remove Outlier and more cleaning
+cooksd = cooks.distance(model)
+influential = as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
+train_df = train_df[-c(influential), ]
+train_df.clean = complete(mice(data = train_df, method = "pmm", maxit = 5, 
+                               seed = 525, print = FALSE), 3)
+train_df.clean$TEAM_BATTING_1B = train_df.clean$TEAM_BATTING_H - 
+  (train_df.clean$TEAM_BATTING_2B + train_df.clean$TEAM_BATTING_3B + 
+     train_df.clean$TEAM_BATTING_HR)
+train_df.clean = subset(train_df.clean, select = -c(TEAM_BATTING_H, TEAM_PITCHING_HR))
 # Data Transformation & Multicollinearity Diagnostic Measures
-train_df.clean = subset(train_set, select = -c(TEAM_PITCHING_HR, TEAM_BATTING_H, TEAM_BATTING_HBP))
 stepdata = cbind(train_df.clean,sapply(train_df.clean[2:length(train_df.clean)], 
@@
 names(stepdata) = make.unique(names(stepdata), sep = "_")
+# Set up repeated k-fold cross-validation
+set.seed(525)
 train.control = trainControl(method = "cv", number = 10)
@@
 mod_sum = summary(step.model$finalModel)
+# Predictions on train data
+pred_step.model = predict(step.model, newdata = stepdata)
 ```
@@
 
-In order to select the best model to make predictions, we looked at some measurements that tells us how well each model fits.
+In order to select the best model to make predictions, we looked at some measurements that tells us how well each model fits the training.
 These include the 1) $R^2$, which represents the proportion of the variance explained by a model; 2) $adj R^2$, which is a modified version of $R^2$; 3) *Root Mean Squared Error* (RMSE), which is the square root of the mean squared error; and 4) *Akaike Information Criterion* (AIC), which is an estimator of out-of-sample prediction error.
 
-It is apparent that Model #2: Pairwise accounts for nearly 43% of the variation in the dependent variable with the independent variables, which is acceptable as a good model ($R^2 = 0.427$, $adjR^2 = 0.417$.
+When comparing the three models, it was interesting to select which would be our best model given that their performance statistics and error measurements were not significantly different from each other.
+It is apparent that Model #1: Backwards accounts for nearly 44% of the variation in the dependent variable with the independent variables, which is acceptable as a good model.
+But it's explanatory power based on an $adj R^2 = 0.414$ suggests it is no different from Model #3: K-fold.
+In addition, Model #1: Backwards has a highest RSME of all the models, which ranks as one of the major criteria we are using to decide on a model.
+Now, Model #2: Pairwise resulted in the $R^2 = 0.427$, $adjR^2 = 0.417$, and the lowest $RSME = 10.193$.
 Because the RMSE and adjusted $R^2$ statistics already include a minor adjustment for the number of coefficients estimated, to evaluate the model complexity, we compared the AIC.
-While a smaller AIC is deemed less complex, the difference between Model #2 and Model #3 is quite insignificant.
+While a smaller AIC is deemed less complex, Model #1: Backwards would have been the way to go, however, it was deemed that the AIC differences among the models are quite insignificant.
+Thus, it is unanimous that Model #2: Pairwise is our final model.
 
-```{r echo=FALSE}
-# Predictions on training data
-mod_sum1 = summary(lm_1)  
-pred_lm1 = predict(lm_1, newdata = train_set)
-pred_step.model = predict(step.model, newdata = stepdata)
-
+```{r performance, echo=FALSE, message=FALSE, warning=FALSE}
 # Model accuracy
@@
 
-Using the test data and the selected final model, a comparison in the prediction statistic was conducted.
+Using the test data and the selected final model, Model #2: Pairwise, a comparison in the prediction statistic was conducted.
 
-```{r include=FALSE, eval=FALSE}
-# Data Transformation to match train data
-eval_set = complete(mice(data = eval_set, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
+```{r predictions, echo=FALSE, message=FALSE, warning=FALSE}
+# Cleaning test set for Model#1
+test_df1<-  eval_set %>% dplyr::select(-c(TEAM_BATTING_HBP, INDEX)) %>%
+  mutate(TEAM_BATTING_1B =
+         TEAM_BATTING_H -
+         (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR)) %>%
+  mutate_if(is.integer, as.numeric)
+test_df1 <- complete(mice(data = test_df1,
+                         method = "pmm",
+                         seed = 9450,
+                         print=FALSE), 3)
 
-# Add final model
-predictions = predict(ADD-MODEL-HERE, newdata = eval_set)
+test_df1<-
+  test_df1 %>%  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
+         SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) %>% dplyr::select(-c(AB, TEAM_BATTING_H))
 
+# Cleaning test set for Model #2
+BxCox <- function(x, l) {
+  ifelse(l == 0, log(x), (x ^ l - 1) / l)
+}
+BCL <- data.table(Metric = c('TEAM_PITCHING_H', 'TEAM_FIELDING_E',
+                             'TEAM_BASERUN_SB', 'TEAM_PITCHING_HR',
+                             'TEAM_BATTING_HR', 'TEAM_PITCHING_BB',
+                             'TEAM_PITCHING_SO', 'TEAM_BATTING_3B',
+                             'TEAM_BASERUN_CS'),
+                  Lambda = c(-2.5, -0.7, 0, 0.6, 0.6, -0.1, -0.35, 0, 0.5))
+setkey(BCL, Metric)
+MBT <- setDT(eval_set)
+#MBT[, `:=`(TEAM_BATTING_HBP = NULL)]
+MBTlng <- melt(MBT, id.vars = 'INDEX', variable.name = 'Metric',
+               value.name = 'Value', variable.factor = FALSE)
+MBTlng[, Value := as.double(Value)]
+MBTlng <- BCL[MBTlng, on = 'Metric']
+MBTlng[!is.na(Lambda), Value := BxCox(Value, Lambda)]
+MBTlng[, Lambda := NULL]
+MBTlngSC <- copy(MBTlng)
+# Remove infinities
+MBTlngSC[, Value := ifelse(is.infinite(Value), NA, Value)]
+# Impute, Center, and Scale
+MBTlngSC[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
+       keyby = Metric]
+MBTlngSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
+MBTSC <- dcast(MBTlngSC, INDEX ~ Metric, value.var = 'Value')
+
+# Cleaning test set for Model #3
+test_df = complete(mice(data = eval_set, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
+test_df$TEAM_BATTING_1B = test_df$TEAM_BATTING_H - 
+  (test_df$TEAM_BATTING_2B + test_df$TEAM_BATTING_3B + test_df$TEAM_BATTING_HR)
+test_df = cbind(test_df, sapply(test_df[1:length(test_df)], function(x) x^2))
+names(test_df) = make.unique(names(test_df), sep = "_")
+```
+
+```{r echo=FALSE}
+# Model #1
+predictions = predict(lm_1, newdata = test_df1)
 results = cbind(dataset = c('Training Data', 'Test Prediction'), 
-                bind_rows(describe(train_set$TARGET_WINS), describe(predictions))) 
+                rbind(psych::describe(train_set$TARGET_WINS), psych::describe(predictions))) 
 rownames(results) = NULL
+results[,-c(2,8,11)] %>% knitr::kable(digits = 2L, caption = 'Prediction Comparison with Model #1: Backwards')
 
-results[,-c(2,8,11)] %>% knitr::kable(digits = 3L, caption = 'Prediction Comparison with Final Model')
+# Model #2
+predictions <- predict(stepModel, MBTSC)
+results = cbind(dataset = c('Training Data', 'Test Prediction'), 
+                rbind(psych::describe(train_set$TARGET_WINS), psych::describe(predictions))) 
+rownames(results) = NULL
+results[,-c(2,8,11)] %>% knitr::kable(digits = 2L, caption = 'Prediction Comparison with Model #2: Pairwise')
+
+# Model #3
+predictions = predict(step.model$finalModel, newdata = test_df)
+results = cbind(dataset = c('Training Data', 'Test Prediction'), 
+                rbind(psych::describe(train_set$TARGET_WINS), psych::describe(predictions))) 
+rownames(results) = NULL
+results[,-c(2,8,11)] %>% knitr::kable(digits = 2L, caption = 'Prediction Comparison with Model #3: K-fold')
 ```
@@
 ## CONCLUSIONS
+
+## CODE APPENDIX
+
+The code chunks below represent the R code called in order during the analysis.
+They are reproduced in the appendix for review and comment.

### 2020-09-16 HW1/HW1_draft_kellogg.Rmd
minor fix in output

@@
 
-```{r cooks_distance, include=TRUE}
+```{r cooks_distance, include=FALSE}
 lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+

### 2020-09-16 HW1/HW1_draft_kellogg.Rmd
fixed training set backward selection

@@
                 AIC = c(AIC(lm_1), AIC(stepModel), AIC(step.model$finalModel)),
-                RSME = c(sqrt(mean((train_set$TARGET_WINS - pred_lm1)^2)), stmRMSE,
+                RSME = c(sqrt(mean((train_df2$TARGET_WINS - pred_lm1)^2)), stmRMSE,
                          sqrt(mean((stepdata$TARGET_WINS - pred_step.model)^2))))

### 2020-09-17 HW1/HW1_draft_kellogg_AA.Rmd
Copyedit Model 1 & Model 2 and redo Model 2 so that it is consistent with the changed data

--- 2020-09-16 HW1/HW1_draft_kellogg.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
 ```
 
-```{r loadData, include=FALSE, warning = FALSE, message = FALSE}
+```{r loadData, include=FALSE}
 library(tidyverse)
@@
 library(corrplot)
-library(data.table)
 library(tidymodels)
@@
 library(caret)
+library(data.table)
 train_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
 eval_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-evaluation-data.csv')
+ts2 <- train_set
+es2 <- eval_set
 set.seed(9450)
@@
 
-## DATA EXPLORATION
-
-### How often does the team win?
-
-We are given a data set of 2,276 records containing 15 seasonal statistics and the
-total number of wins a team had in a given year. On average, about 50% of games played 
-are won (81 games out of 162), with the best individual season having 146 wins and the worst 
-season having 0 wins. The data is normally distributed and most years have between 49 
-and 112 wins (blue lines below). The nature of the distribution means there aren't too 
-many extreme seasons where wins are significantly higher or lower than usual. This serves 
-as a good gut-check for our final predictions; if the predicted wins are too high or too 
+# DATA EXPLORATION
+## How Often Does The Team Win?
+We are given a data set of 2,276 records containing 15 seasonal statistics and
+the total number of wins a team had in a given year. On average, about 50% of
+games played are won (81 games out of 162), with the best individual season
+having 146 wins and the worst season having 0 wins. The data is normally
+distributed and most years have between 49 and 112 wins (blue lines below). The
+nature of the distribution means there aren't too many extreme seasons where
+wins are significantly higher or lower than usual. This serves as a good
+gut-check for our final predictions; if the predicted wins are too high or too 
 low, we know something in our model is probably off.
 
-```{r wins, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
+```{r wins, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
 avgWins <- mean(train_set$TARGET_WINS)
 sdev <- sd(train_set$TARGET_WINS)
-l1 <- mean(train_set$TARGET_WINS) - (2*sdev)
-l2 <- mean(train_set$TARGET_WINS) + (2*sdev)
+l1 <- mean(train_set$TARGET_WINS) - (2 * sdev)
+l2 <- mean(train_set$TARGET_WINS) + (2 * sdev)
+
 # histogram of wins
+
 p <- ggplot(train_set, aes(x=TARGET_WINS)) +
   geom_histogram() + 
-  geom_vline(aes(xintercept= avgWins),
-            color="red", linetype="dashed", size=1)+
-  geom_vline(aes(xintercept=l1),
-            color="blue", linetype="dashed", size=1) +
-    geom_vline(aes(xintercept=l2),
-            color="blue", linetype="dashed", size=1) +
-  labs(title="Distribution of Wins",x="Number of wins", y = "Number of seasons")
+  geom_vline(aes(xintercept = avgWins),
+            color = "red", linetype = "dashed", size = 1)+
+  geom_vline(aes(xintercept = l1),
+            color="blue", linetype = "dashed", size = 1) +
+    geom_vline(aes(xintercept = l2),
+            color = "blue", linetype = "dashed", size = 1) +
+  labs(title = "Distribution of Wins", x = "Number of wins",
+       y = "Number of seasons")
 p
@@
 
-### What's missing? 
-A first look at the data shows that only about 8% of the records have a full set of 
-information. The good news is that most of the missing values come from statistics 
-that don't happen too often: hit-by-pitch (`TEAM_BATTING_HBP`, 92% missing!), caught 
-stealing (`TEAM_BASERUN_CS`, 34% missing), and double plays (`TEAM_FIELDING_DP`, 13% 
-missing). Since we have so little hit-by-pitch data, we expect that it doesn't contribute 
-much to overall wins and will eliminate it from a few of the models we propose. The 
-other two stats have less than half of the data missing, so we'll need to think of a 
-clever way to fill in these values. The remaining missing information is from a 
-combination of stolen bases and strikeouts by pitchers and batters (`TEAM_BASERUN_SB`,
-`TEAM_PITCHING_SO`, `TEAM_BATTING_SO`). **It seems completely unreasonable** to have zero 
-strike outs in a season, so this is something we'll most certainly have to impute.
+## What's Missing? 
+A first look at the data shows that only about 8% of the records have a full set
+of information. The good news is that most of the missing values come from
+statistics that don't happen too often: hit-by-pitch (`TEAM_BATTING_HBP`, 92%
+missing!), caught stealing (`TEAM_BASERUN_CS`, 34% missing), and double plays
+(`TEAM_FIELDING_DP`, 13% missing). Since we have so little hit-by-pitch data, we
+expect that it doesn't contribute much to overall wins and will eliminate it
+from a few of the models we propose. The other two stats have less than half of
+the data missing, so we'll need to think of a clever way to fill in these
+values. The remaining missing information is from a combination of stolen bases
+and strikeouts by pitchers and batters (`TEAM_BASERUN_SB`, `TEAM_PITCHING_SO`,
+`TEAM_BATTING_SO`). **It seems completely unreasonable** to have zero strike
+outs in a season, so this is something we'll most certainly have to impute.
 
-```{r pctNull, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2, fig.width = 5, fig.align = 'center'}
-pct_null <- data.frame(do.call("rbind", map(train_set %>% dplyr::select(-INDEX), ~mean(is.na(.)))))
+```{r pctNull, fig.height = 2, fig.width = 5, fig.align = 'center'}
+pct_null <- data.frame(do.call("rbind", map(train_set %>% dplyr::select(-INDEX),
+                                            ~ mean(is.na(.)))))
 colnames(pct_null) <- c('PCT_NULL')
@@
   dplyr::select(VARIABLE, PCT_NULL)
-ggplot(totalNulls, aes(x=reorder(VARIABLE, PCT_NULL), y=PCT_NULL, label = round(PCT_NULL, 2))) + 
+ggplot(totalNulls, aes(x = reorder(VARIABLE, PCT_NULL), y = PCT_NULL,
+                       label = round(PCT_NULL, 2))) + 
   geom_text(vjust = 0.5, hjust = -0.05)+
@@
   xlab("Statistic") + ylab("Percent Missing") + 
-  coord_flip() +
-  expand_limits(y = 1)
+  coord_flip() + expand_limits(y = 1)
 ```
 
-### Do the individual stats affect winning?
+## Do the Individual Stats Affect Winning?
+**Stats with an expected negative impact:** Intuitively, we expect that Caught
+Stealing, Errors, Hits Allowed, Home Runs Allowed, Strikeouts by Batters, and
+Walks Allowed would all have a **negative** impact on the total wins. In other
+words, as these values increase, we expect that the team is less likely to win.
 
-**Stats with an expected negative impact: ** Intuitively, we expect that Caught stealing, 
-Errors, Hits allowed, Home Runs allowed, Strikeouts by batters, and Walks allowed would 
-all have a **negative** impact on the total wins. In other words, as these values increase, 
-we expect that the team is less likely to win.
-
-```{r negFactors, echo = FALSE, message = FALSE, warning = FALSE}
+```{r negFactors}
 negSet <- train_set %>%
@@
 ggplot(data = negSet %>%
-  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
-  facet_wrap(~ STAT, scales = "free")
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
+  aes(x = VALUE, y = TARGET_WINS)) + 
+  geom_point() + geom_smooth(method = "lm", se = FALSE) +
+  labs(title = "Number of wins with respect to recorded stats",
+       x = "", y = "Number of Wins") + facet_wrap(~ STAT, scales = "free")
 ```
 
-When we take a closer look at the data, these negative relationships aren't obvious. In 
-fact, only **Errors** and **Hits allowed** seem to have a negative impact on wins. **Caught**
-**stealing** and **Strikeouts by batters** appear to be random; this means that whether the 
-stat for a particular season is high or low doesn't affect the overall number of wins.
+When we take a closer look at the data, these negative relationships aren't
+obvious. In fact, only **Errors** and **Hits Allowed** seem to have a negative
+impact on wins. **Caught Stealing** and **Strikeouts by Batters** appear to be
+random; this means that whether the stat for a particular season is high or low
+doesn't affect the overall number of wins.
 
-Even more interestingly, **Home Runs allowed** and **Walks allowed** have the *opposite* 
-effect; as these stats increase, so do the number of wins!
+Even more interestingly, **Home Runs Allowed** and **Walks Allowed** have the
+*opposite* effect---as these stats increase, so do the number of wins!
 
-**Stats with an expected positive impact: ** We can look at the same information for the 
-stats that we expect to have a **positive** effect on wins: Base hits, Doubles, Triples, Home 
-Runs, Walks, Batters getting hit by pitches, Stolen bases, Double Plays, and Strikeouts by 
-pitchers.
+**Stats with an expected positive impact:** We can look at the same information
+for the stats that we expect to have a **positive** effect on wins: Base Hits,
+Doubles, Triples, Home Runs, Walks, Batters getting hit by pitches, Stolen
+Bases, Double Plays, and Strikeouts by Pitchers.
 
-```{r posFactors, echo = FALSE, message = FALSE, warning = FALSE}
+```{r posFactors}
 posSet <- train_set %>%
@@
 ggplot(data = posSet %>%
-  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
-  geom_point() +
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
+  aes(x = VALUE, y = TARGET_WINS)) + geom_point() +
   geom_smooth(method = "lm", se = FALSE) +
-  labs(title="Number of Wins with Respect to Recorded Stats",x="", y = "Number of Wins") +
-  facet_wrap(~ STAT, scales = "free")
+  labs(title = "Number of Wins with Respect to Recorded Stats", x = "",
+       y = "Number of Wins") + facet_wrap(~ STAT, scales = "free")
 ```
 
-Many of these stats *do* seem to have an effect on the number of wins, most notably,
-**Base hits** and **Walks**. We see weaker positive relationships for **Home runs**, 
-**Doubles**, **Triples**, and **Stolen bases**. This makes sense when we think about it; 
-these things tend to happen less often in games than pure base hits and walks, so they don't 
-have as much of an effect on winning. Finally, **Double plays** and **Batters hit** don't 
-appear to have any correlation with the number of wins. Once again, this intuitively makes 
-sense because they are less likely to happen in a game.
+Many of these stats *do* seem to have an effect on the number of wins, most
+notably, **Base Hits** and **Walks**. We see weaker positive relationships for
+**Home Runs**, **Doubles**, **Triples**, and **Stolen Bases**. This makes sense
+when we think about it; these things tend to happen less often in games than
+pure base hits and walks, so they don't have as much of an effect on winning.
+Finally, **Double Plays** and **Batters Hit** don't appear to have any
+correlation with the number of wins. Once again, this intuitively makes sense
+because they are less likely to happen in a game.
 
-One thing to note is the number of strikeouts compared to the number of wins. We can see
-that there are a few outliers (abnormally high numbers of strikeouts in a season). This
-should be taken with caution, as they don't represent a typical season's stats.
+One thing to note is the number of strikeouts compared to the number of wins. We
+can see that there are a few outliers (abnormally high numbers of strikeouts in
+a season). This should be taken with caution, as they don't represent a typical
+season's stats.
 
-### Are some stats more skewed than others?
+## Are Some Stats More Skewed Than Others?
+Before using any of the statistics in a model, we need to take a closer look at
+the variation in the data. We call out-of-the-ordinary values (exceptionally
+high or low values) **outliers**. We need to take these into account in our
+modeling because we want to make sure our predictions aren't skewed because of
+them.
 
-Before using any of the statistics in a model, we need to take a closer look at the variation 
-in the data. We call out-of-the-ordinary values (exceptionally high or low values) **outliers**. 
-We need to take these into account in our modeling because we want to make sure our predictions 
-aren't skewed because of them.
-
-```{r summaryInfo, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}
+```{r summaryInfo, fig.height = 3, fig.width = 5, fig.align = 'center'}
 #summary(train_set)
-totalOutliers <- data.frame(sapply(train_set %>% dplyr::select(-INDEX, -TARGET_WINS, - TEAM_BATTING_HBP), 
+totalOutliers <- data.frame(
+  sapply(train_set %>% dplyr::select(-INDEX, -TARGET_WINS, - TEAM_BATTING_HBP), 
        function(x){length(boxplot.stats(x)$out)/nrow(train_set)}))
@@
 colnames(totalOutliers) <- c('PCT_OUTLIERS', 'VARIABLE_NM')
-ggplot(totalOutliers , aes(x=reorder(VARIABLE_NM, PCT_OUTLIERS), y=PCT_OUTLIERS, label = round(PCT_OUTLIERS, 3))) + 
-  geom_text(vjust = 0.5, hjust = -0.05)+
-  geom_bar(stat = "identity") +
-  ggtitle("Percentage of Outliers") +
-  xlab("Statistic") + ylab("Percent of Data that is an Outlier") + 
-  coord_flip() +
+ggplot(totalOutliers,
+       aes(x = reorder(VARIABLE_NM, PCT_OUTLIERS), y=PCT_OUTLIERS,
+           label = round(PCT_OUTLIERS, 3))) + 
+  geom_text(vjust = 0.5, hjust = -0.05)+ geom_bar(stat = "identity") +
+  ggtitle("Percentage of Outliers") + xlab("Statistic") +
+  ylab("Percent of Data that is an Outlier") + coord_flip() +
   expand_limits(y = 0.15)
@@
 
-As we can see, some of the provided statistics are well-balanced in the sense that there are very 
-*few* (or no) extreme values. **Home Runs allowed** (`TEAM_PITCHING_HR`), **Strikeouts by batters** 
-(`TEAM_BATTING_SO`), and **Home Runs by batters** (`TEAM_BATTING_HR`) are examples of this.
+As we can see, some of the provided statistics are well-balanced in the sense
+that there are very *few* (or no) extreme values. **Home Runs allowed**
+(`TEAM_PITCHING_HR`), **Strikeouts by batters** (`TEAM_BATTING_SO`), and
+**Home Runs by batters** (`TEAM_BATTING_HR`) are some examples.
 
-```{r fewOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2, fig.width = 6, fig.align = 'center'}
+```{r fewOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
 set1 <- train_set %>%
@@
   filter(complete.cases(.) == TRUE)
-vals <- ggplot(set1 , aes(x=stat, y=value)) + 
-    geom_boxplot() +
-  labs(title="",x="", y = "Value") +
-  facet_wrap(~stat, scale="free")
+vals <- ggplot(set1 , aes(x = stat, y = value)) + 
+    geom_boxplot() + labs(title = "", x = "", y = "Value") +
+  facet_wrap( ~ stat, scales = "free")
 vals
 ```
+
 Some things to note about each of these statistics:
 
-* **Home Runs allowed** (average = ~100/year) and **Home Runs by batters** (average = ~106/year) 
-have a very similar mid-range distribution (50% of the data lies between ~50 and 150). 
-The slight difference in average stats means that teams tend to have a higher number
-of Home Runs than the opposition team. 
-* The only thing that stands out about **Strikeouts by batters** (average = ~736/year) is 
-how nearly perfectly normal it is. 50% of the data is between about 500 and 1000 and 
-there are absolutely no outliers in the dataset! This means that there were no surprisingly
-high or low seasons.
+  * **Home Runs Allowed** (average ~100/year) and **Home Runs by Batters**
+  (average ~106/year) have a very similar mid-range distribution (50% of the
+  data lies between ~50 and 150). The slight difference in average stats means
+  that teams tend to have a higher number of Home Runs than the opposition team.
+  * The only thing that stands out about **Strikeouts by Batters** (average
+  ~736/year) is how nearly perfectly normal it is. 50% of the data is between
+  about 500 and 1000 and there are absolutely no outliers in the dataset! This
+  means that there were no surprisingly high or low seasons.
 
-Conversely, some of the stats have a very *high* number of outliers, indicating that there
-are some seasons with some abnormally high or low values. **Errors** (`TEAM_FIELDING_E`), 
-**Hits allowed** (`TEAM_PITCHING_H`), **Walks by batters** (`TEAM_BATTING_BB`), and 
-**Stolen bases** (`TEAM_BASERUN_SB`) are examples of this.
+Conversely, some of the stats have a very *high* number of outliers, indicating
+that there are some seasons with some abnormally high or low values. **Errors**
+(`TEAM_FIELDING_E`), **Hits allowed** (`TEAM_PITCHING_H`), **Walks by batters**
+(`TEAM_BATTING_BB`), and **Stolen bases** (`TEAM_BASERUN_SB`) are some examples.
 
-```{r manyOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2, fig.width = 6, fig.align = 'center'}
+```{r manyOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
 set2 <-  train_set %>%
@@
     geom_boxplot() +
-  labs(title="",x="", y = "Value") +
-  facet_wrap(~ stat, scale = 'free', ncol = 4)
+  labs(title = "",x = "", y = "Value") +
+  facet_wrap(~ stat, scales = 'free', ncol = 4)
 vals2
 ```
+
 Some things to note about each of these statistics:
 
-* All of the outliers for **Errors**, **Hits allowed**, and **Stolen bases** are above the upper 
-tail of the data set. This is further illustrated by the mean and median values for these stats; in 
-all instances, the mean per year (Errors = ~246/year, Hits allowed = ~1779/year, Stolen bases = ~125/year) 
-are higher than the median per year (Errors = ~159/year, Hits allowed = ~1518/year, Stolen bases = 
-~101/year). This means that some seasons with exceptionally high values skew the dataset. 
-* There are a few *very* extreme outliers for **Hits allowed**. The maximum value is 30,132, which
-is over 16 times the average number of hits allowed per season!
-* There are outliers both above *and* below the tails of the data for the **Walks by batters** stat. 
-This means that we have exceptionally low (min = 0!) and exceptionally high (max = 878) seasons.
+  * All of the outliers for **Errors**, **Hits Allowed**, and **Stolen Bases**
+  are above the upper tail of the data set. This is further illustrated by the
+  mean and median values for these stats; in all instances, the means per year
+  (Errors ~246/year, Hits Allowed ~1779/year, Stolen Bases ~125/year) are higher
+  than the medians per year (Errors ~159/year, Hits Allowed ~1518/year, Stolen
+  Bases ~101/year). This means that some seasons with exceptionally high values
+  skew the dataset.
+  * There are a few *very* extreme outliers for **Hits Allowed**. The maximum
+  value is 30,132, which is over 16 times the average number of hits allowed per
+  season!
+  * There are outliers both above *and* below the tails of the data for the
+  **Walks by Batters** stat. This means that we have exceptionally low
+  (min = 0!) and exceptionally high (max = 878) seasons.
 
-The remaining stats, **Walks allowed** (`TEAM_PITCHING_BB`), **Base Hits by batters** (`TEAM_BATTING_H`), 
-**Caught stealing** (`TEAM_BASERUN_CS`), **Strikeouts by pitchers** (`TEAM_PITCHING_SO`), **Double plays** 
-(`TEAM_FIELDING_DP`), **Triples by batters** (`TEAM_BATTING_3B`), and **Doubles by batters** 
-(`TEAM_BATTING_2B`) have between 29 and 99 outliers.
+The remaining stats, **Walks Allowed** (`TEAM_PITCHING_BB`),
+**Base Hits by Batters** (`TEAM_BATTING_H`), **Caught Stealing**
+(`TEAM_BASERUN_CS`), **Strikeouts by Pitchers** (`TEAM_PITCHING_SO`),
+**Double Plays** (`TEAM_FIELDING_DP`), **Triples by Batters**
+(`TEAM_BATTING_3B`), and **Doubles by Batters** (`TEAM_BATTING_2B`) have between
+29 and 99 outliers.
 
-```{r otherOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3.5, fig.width = 7, fig.align = 'center'}
+```{r otherOutliers, fig.height = 3.5, fig.width = 7, fig.align = 'center'}
 set3 <-  train_set %>%
@@
   filter(complete.cases(.) == TRUE)
-vals3 <- ggplot(set3 , aes(x=stat, y=value)) + 
+vals3 <- ggplot(set3, aes(x=stat, y=value)) + 
     geom_boxplot() +
-  labs(title="",x="", y = "Value") +
-  facet_wrap(~ stat, scale = 'free', ncol = 4)
+  labs(title = "", x = "",y = "Value") +
+  facet_wrap(~ stat, scales = 'free', ncol = 4)
 vals3
@@
 
-* All variables are very narrowly distributed, meaning that most of the data falls within
-a small range. 
-* **Strikeouts by pitchers** and **Walks allowed** have a few very extreme outliers; these 
-represent seasons that have abnormally high values for the statistics. 
-* The average number of pure **Base Hits** (1469/season) is greater than the average number
-of Doubles (~241/season) and Triples (~55/season). This isn't at all surprising, but serves 
-as a good gut check on the validity of the data.
+  * All variables are very narrowly distributed, meaning that most of the data
+  falls within a small range. 
+  * **Strikeouts by Pitchers** and **Walks Allowed** have a few very extreme
+  outliers; these   represent seasons that have abnormally high values for the
+  statistics. 
+  * The average number of pure **Base Hits** (1469/season) is greater than the
+  average number of Doubles (~241/season) and Triples (~55/season). This isn't
+  at all surprising, but serves as a good gut check on the validity of the data.
 
-### Are stats correlated?
-
-We would expect that a few things in the dataset might be correlated: perhaps number of errors and
-hits/homeruns allowed or the number of base hits by batters and doubles/triples/homeruns. We can
-visualize the correlations between the statistics to determine if there is a significant relationship
-between the them: blue dots represent positively correlated variables (as one increases,
-so does the other) and red dots represent negatively correlated variables (as one increases,
-the other decreases).
+## Are Stats Correlated?
+We would expect that a few things in the dataset might be correlated: perhaps
+number of errors and hits/homeruns allowed or the number of base hits by batters
+and doubles/triples/homeruns. We can visualize the correlations between the
+statistics to determine if there is a significant relationship between the them:
+blue dots represent positively correlated variables (as one increases, so does
+the other) and red dots represent negatively correlated variables (as one
+increases, the other decreases).
 
@@
 
-* **Errors** are highly, negatively correlated with walks by batters, strikeouts by batters, and
-homeruns (both by batters and allowed). 
-* **Triples by batters** are highly, negatively correlated with strikeouts by batters and homeruns
+  * **Errors** are highly, negatively correlated with walks by batters,
+  strikeouts by batters, and homeruns (both by batters and allowed).
+  * **Triples by Batters** are highly, negatively correlated with strikeouts by
+  batters and homeruns (both by batters and allowed).
+  * **Strikeouts by Batters** are highly, positively correlated with homeruns
   (both by batters and allowed). 
-* **Strikeouts by batters** are highly, positively correlated with homeruns (both by batters and
-allowed). 
-* **Homeruns by batters** and **Homeruns allowed** are both positively correlated with walks by 
-batters.
-* As expected, **basehits** are positively correlated with doubles and triples.
+  * **Homeruns by Batters** and **Homeruns Allowed** are both positively
+  correlated with walks by batters.
+  * As expected, **Base Hits** are positively correlated with doubles and
+  triples.
 
-We can keep these correlations in mind when developing our models: if we have correlated statistics, 
-there could be in-built redundancy in the features, and we may be able to create a simpler, more
-accurate model by eliminating some.
+We can keep these correlations in mind when developing our models: if we have
+correlated statistics, there could be in-built redundancy in the features, and
+we may be able to create a simpler, more accurate model by eliminating some.
 
-## DATA PREPARATION
+# DATA PREPARATION
+Now we have a good idea of the data we are looking at we can take the next steps
+to prepare it for building a solid model.
 
-Now we have a good idea of the data we are looking at we can take the next steps to 
-prepare it for building a solid model.
-
-### Outlier Removal
-
-As we saw in the data analysis, there are some outlier concerns for some of the variables,
-so we will need to account for this in our modeling. When performing the processes of outlier 
-removal, a cautious approach is always best. Each outlier is evaluated to ensure:
+## Outlier Removal
+As we saw in the data analysis, there are some outlier concerns for some of the
+variables, so we will need to account for this in our modeling. When performing
+the processes of outlier removal, a cautious approach is always best. Each
+outlier is evaluated to ensure:
 
@@
 
-```{r outliers_1, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 8, fig.align = 'center'}
+```{r outliers_1, fig.height = 4, fig.width = 8, fig.align = 'center'}
 train_set %>%
@@
   pivot_longer(everything(), 
-               names_to ="Variable", 
-               values_to="Value") %>%
-  ggplot(aes(x=Variable, y=Value)) +
+               names_to = "Variable", 
+               values_to = "Value") %>%
+  ggplot(aes(x = Variable, y = Value)) +
   geom_boxplot(na.rm = TRUE) + 
@@
 
-Off the bat, there are some clear issues with two of the variables: `TEAM_PITCHING_SO` and
-`TEAM_PITCHING_H.`
+Off the bat (pun intended), there are some clear issues with two of the
+variables: `TEAM_PITCHING_SO` and `TEAM_PITCHING_H.`
 
-* **TEAM_PITCHING_SO**: The largest outlier is close to 20,000, which averages ~123 strikeouts
-per game. The MLB team pitching strikeout record is 1,450 and set in 2014 (CLE). There are a 
-minimum of 27 at-bats per game and most teams average ~35 at-bats, thus making 123 strikeouts
-impossible. 
-* **TEAM_PITCHING_H**: The largest outlier for this vaiable is over 30,000 which is highly 
-unlikely.
+  * **TEAM_PITCHING_SO**: The largest outlier is close to 20,000, which averages
+  ~123 strikeouts per game. This is only possible if every one of the 162 games
+  went approximately 40 innings, and each out was a strikeout---which clearly has
+  never happened.
+  * **TEAM_PITCHING_H**: The largest outlier for this vaiable is over 30,000
+  which is simlarly highly unlikely.
 
-Since both appear with a heavy right skew, we will use median and IQR to remove the outliers.
+Since both appear with a heavy right skew, we will use median and IQR to remove
+the outliers.
 
-```{r iqr_removal_SO, include=FALSE}
+```{r iqr_removal_SO}
 train_set<- train_set %>%
@@
     TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO,
-                                probs=c(0.25), 
-                                na.rm = TRUE) - 1.5*iqr, 
+                                probs = c(0.25), 
+                                na.rm = TRUE) - 1.5 * iqr, 
     TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, 
-                                probs=c(0.75), 
-                                na.rm=TRUE) + 1.5*iqr
+                                probs = c(0.75), 
+                                na.rm = TRUE) + 1.5 * iqr
         ) %>%
@@
 ```
-
-```{r iqr_removal_H, include=FALSE}
+```{r iqr_removal_H}
 train_set<- train_set %>%
@@
     TEAM_PITCHING_H > quantile(TEAM_PITCHING_H,
-                                probs=c(0.25), 
-                                na.rm = TRUE) - 1.5*iqr, 
+                                probs = c(0.25), 
+                                na.rm = TRUE) - 1.5 * iqr, 
     TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, 
-                                probs=c(0.75), 
-                                na.rm=TRUE) + 1.5*iqr
+                                probs = c(0.75), 
+                                na.rm = TRUE) + 1.5 * iqr
         ) %>%
@@
 
-We can take a look at the distributions of these variables after outlier removal:
+We can take a look at the distributions of these variables after removing the
+outliers:
 
-```{r outlier_boxplot, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 8, fig.align = 'center'}
+```{r outlier_boxplot, fig.height = 4, fig.width = 8, fig.align = 'center'}
 train_set %>%
@@
 This has eliminated what appears to be the most extreme outliers. There are
-still large outlier sets for Errors and stolen bases, but none that seem to 
+still large outlier sets for Errors and Stolen Bases, but none that seem to
 dwarf the other variables. According to the box plot, things appear to be on a 
@@
 
-### "MICE" imputation method
+## Imputation
+From our early exploration of the data, the vast majority of the data is
+complete with only a few variables with missing values. We will use the **MICE**
+(Multivariate Imputation by Chained Equations) method. MICE is a principled
+method for dealing with missing data. It creates multiple imputations, as
+opposed to single imputations, and accounts for the statistical uncertainty in
+these imputations. It creates predictive values for the mean instead of imputing
+the IQR values.
 
-From our early exploration of the data, the vast majority of the data is complete 
-with only a few variables with missing values. We will use the "mice" imputation method, 
-specifically predictive mean matching method.
-
-MICE (Multivariate imputation by chained equations) is a principled method of dealing 
-with missing data. It creates multiple imputations, as opposed to single imputations 
-and accounts for the statistical uncertainty in the imputations. It creates predictive 
-values for the mean instead of imputing the IQR values.
-
-```{r impute_train,echo=FALSE, message=FALSE, warning=FALSE}
+```{r impute_train}
 train_set <- complete(mice(data = train_set, 
@@
 
-### Feature Engineering: Single base hits
+## Feature Engineering
+### Single base hits
+With outliers removed and missing values imputed via MICE, we can create a few
+new variables. The first will be single base hits, and it will be derived from
+some of the variables we do have. We know that Team Batting Hits
+(`TEAM_BATTING_H`) is a combination of *all* hits for the season, so we can
+create Single Base Hits as follows: 
+\[
+Singles = Total\ Hits - (Doubles + Triples + Homeruns)
+\]
 
-With outliers removed and missing values imputed via MICE, we can create a few new 
-variables. The first will be single base hits, and it will be derived from some of the
-variables we do have. We know that Team Batting Hits (`TEAM_BATTING_H`) is a combination 
-of *all* hits for the season, so we can create Single Base Hits as follows:
-
-$$Singles = Total\ Hits - (Doubles + Triples + Homeruns)$$
-
-```{r singles, include=FALSE}
+```{r singles}
 train_set<- train_set %>%
@@
 
-### Feature Engineering: Slugging Percentage
+### Slugging Percentage
+The second variable we will create is **Slugging Percentage**. Slugging is an
+offensive statistic that is a good predictor of winning and it tends to have
+better variance and correlation behavior than most other variables. It is
+composed of singles, doubles, triples, home runs and at-bats:
 
-The second variable we will create is **Slugging Percentage**. Slugging is an offensive 
-statistic that is a good predictor of winning and it tends to have better variance and
-correlation than most other variables. It is composed of singles, doubles, triples, 
-home runs and at-bats.
-
-Slugging is calculated in the following way:
-
-$${\displaystyle \mathrm {SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{AB}}}$$
-
+\[
+\mathrm{SLG} ={\frac{{\mathit{1B}} + 2\times{\mathit{2B}} +
+3\times{\mathit{3B}} + 4\times {\mathit{HR}}}{\mathit{AtBats}}}
+\]
 Because we don't have a statistic for at-bats, we will approximate it as follows:
 
-$${\displaystyle \mathrm {\sim SLG} ={\frac {({\mathit {1B}})+(2\times {\mathit {2B}})+(3\times {\mathit {3B}})+(4\times {\mathit {HR}})}{SO_{batting} + H + BB}}}$$
+\[
+\widehat{\mathrm{SLG}} = {\frac{\mathit{1B} + 2\times\mathit{2B} +
+3\times\mathit{3B} + 4\times\mathit{HR}}{SO_{batting} + H + BB}}
+\]
 
-```{r make_slg_obp, include=FALSE}
+```{r make_slg_obp}
 train_set<- train_set %>%
   mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
-         SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) 
+         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
+                  4 * TEAM_BATTING_HR) / AB)
 ```
 
-As a gut-check, we can take a look at Slugging in comparison to Triples (`TEAM_BATTING_3B`) 
-and Homeruns (`TEAM_BATTING_HR`). All of these variables should have a positive correlation 
-with the total number of wins.
+As a gut-check, we can take a look at Slugging in comparison to Triples
+(`TEAM_BATTING_3B`) and Home Runs (`TEAM_BATTING_HR`). All of these variables
+should have a positive correlation with the total number of wins.
 
-```{r slugging_compare, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 8, fig.align = 'center'}
+```{r slugging_compare, fig.height = 2.5, fig.width = 8, fig.align = 'center'}
 modified_set <- train_set %>%
@@
 ggplot(data = modified_set %>%
-  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
+  aes(x = VALUE, y = TARGET_WINS)) + 
   geom_point() +
   geom_smooth(method = "lm", se = FALSE) +
-  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
+  labs(title = "Number of wins with respect to recorded stats", x = "",
+       y = "Number of Wins") +
   facet_wrap(~ STAT, scales = "free")
@@
 
-## BUILD MODELS
+# BUILD MODELS
+## Base Model
+We will start with a simple linear model to serve as a baseline. This includes
+all variables in the dataset.
 
-### Base Model
-We will start with a simple linear model to serve as a baseline. This includes all variables 
-in the dataset.
-
-```{r baseModel, include = FALSE}
+```{r baseModel}
 lm_reg = lm(data=train_set, TARGET_WINS ~ .)
@@
 
-### Model 1
+## Model 1
 Model 1 includes all variables except Singles, Triples, Base Hits, Strikeouts by
-pitchers, Walks allowed, and Batters Hit by Pitch. The final variables were chosen as 
-a result of backwards dplyr::selection based on null hypothesis testing for non-zero slope.
+pitchers, Walks allowed, and Batters Hit by Pitch. The final variables were
+chosen as a result of backwards `dplyr::selection` based on null hypothesis
+testing for non-zero slope.
 
-**Further data transformation: **  
-We will use Cook's distance to remove outliers that are influencing the fit of the model above. We will use a cutoff of $\frac{4}{N}$.
+### Further data transformation
+We will use Cook's distance to remove outliers that are influencing the fit of
+the model above. We will use a cutoff of \(\frac{4}{N}\).
 
-```{r cooks_distance, include=FALSE}
-lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
-          TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
-          TEAM_PITCHING_HR+ TEAM_PITCHING_H+
-          TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_set)
+```{r cooks_distance}
+lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
+          TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB +
+            TEAM_BASERUN_CS + TEAM_PITCHING_HR + TEAM_PITCHING_H +
+            TEAM_FIELDING_E + TEAM_FIELDING_DP + SLG, data = train_set)
 cooks_dis<- cooks.distance(lm_1)
-influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4/nrow(train_set)))])
+influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4 / nrow(train_set)))])
 train_df2<- train_set[-influential, ]
-lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
-          TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
-          TEAM_PITCHING_HR+ TEAM_PITCHING_H+
-          TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_df2)
-
+lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
+            TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB +
+            TEAM_BASERUN_CS + TEAM_PITCHING_HR + TEAM_PITCHING_H +
+            TEAM_FIELDING_E + TEAM_FIELDING_DP + SLG, data = train_df2)
 mod_sum1 = summary(lm_1)  
 pred_lm1 = predict(lm_1, newdata = train_df2)
-knitr::kable(summary(lm_1)$coefficients, digits = 3L, caption = 'Backwards Model Regression Output')
+knitr::kable(summary(lm_1)$coefficients, digits = 3L,
+             caption = 'Backwards Model Regression Output')
 ```
 
-**Coefficient Discussion: **
-First, we are keeping the coefficient for `TEAM_BATTING_HR` since it's p-value is marginally above the general threshold and knowledge of the game suggests it is important.
-There are some counter intuitive results, which is expected given that baseball is a messy, imprecise game that has evolved over time.
-To that end, we should expect some seemingly strange results from algorithmic regression.
-We used a combination of domain knowledge and data analysis to justify retaining features.
+### Coefficient Discussion
+First, we are keeping the coefficient for `TEAM_BATTING_HR` since it's p-value
+is marginally above the general threshold and knowledge of the game suggests it
+is important. There are some counter intuitive results, which is expected given
+that baseball is a messy, imprecise game that has evolved over time. To that
+end, we should expect some seemingly strange results from algorithmic
+regression. We used a combination of domain knowledge and data analysis to
+justify retaining features.
 
-Fielding double plays and batting doubles both appear to have negative impacts on wins even though they _should_ be positive impacts.
-Turning double plays, while a good for the defensive team, may suggest a larger, negative issue.
-Namely, weak pitching that leads to runners on base.
-Similarly, batting doubles, leaves runners open to double plays.
-Allowed hits by pitching, caught stealing, and batting strike-outs are all counter intuitive results as well, but they seem to be small contributors and these are events that happen regularly in every game.
+Fielding double plays and batting doubles both appear to have negative impacts
+on wins even though they *should* be positive impacts. Turning double plays,
+while a good for the defensive team, may suggest a larger, negative issue.
+Namely, weak pitching that leads to runners on base. Similarly, batting doubles,
+leaves runners open to double plays. Allowed hits by pitching, caught stealing,
+and batting strike-outs are all counter intuitive results as well, but they seem
+to be small contributors and these are events that happen regularly in every
+game.
 
 Slugging as an approximation is the major predictor in this regression, by far.
-The other predictors other than the intercept are orders of magnitude less in predictive value.
-It should be noted that slugging alone is not a good predictor for wins overall.
+The other predictors other than the intercept are orders of magnitude less in
+predictive value. It should be noted that slugging alone is not a good predictor
+for wins overall.
 
-### Model 2
-This is a model allowing pairwise interactions within the data types of baserunning,
-batting, pitching, and fielding, fit using a forward and backwards stepwise regression.
+## Model 2
+This is a model allowing pairwise interactions within the data types of
+baserunning, batting, pitching, and fielding. It is fit using a forward and
+backwards stepwise regression based on AIC.
 
-```{r model2, include = FALSE}
+### Centering and Scaling
+All the features will be centered and scaled. The target variable will be left
+in normal space. This will make prediction easier.
+
+```{r model2C+S}
+ts2 <- train_set[, -(17:19)]
+ts2 <- data.frame(apply(ts2, 2, scale))
+ts2$TARGET_WINS <- train_set$TARGET_WINS
+
+```
+
+A fully specified model with all the interactions will be the starting point for
+the stepwise regression, which uses AIC as its target metric.
+
+```{r model2}
 crazyModel <- lm(TARGET_WINS ~
@@
   (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
-  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = train_set)
+  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = ts2)
+
 stepModel <- stepAIC(crazyModel, scope = list(upper = ~
@@
   TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, lower = ~ 1),
-  direction = 'both')
+  direction = 'both', trace = 0)
 sstm <- summary(stepModel)
@@
 
-**Coefficient Discussion: **
-```{r model2_discuss, echo = FALSE, message = FALSE, warning = FALSE,}
+### Coefficient Discussion:
+```{r model2_discuss}
 knitr::kable(sstm$coefficients, digits = 3L,
@@
  * Reasonable
-   * Caught stealing removes baserunners; negative coefficient makes sense.
+   * Caught stealing removes baserunners but it isn't significant and can be
+   ignored on its own.
    * Stolen bases get a runner closer to home plate; positive coefficient makes
@@
    sense.
-   * Walks are free baserunners; positive coefficient makes sense.
-   * Hits increase the number of baserunners; positive coefficient makes sense.
+   * Walks are free baserunners; positive coefficient makes sense but it isn't
+   significant and can be ignored on its own.
    * Hitting home runs directly increase the score; positive coefficient makes
@@
    sense.
+   * Giving up walks allows the other team free baserunners; negative
+   coefficient makes sense but it isn't significant and can be ignored on its
+   own.
    * Giving up home runs gives the opponent scores; negative coefficient makes
@@
  * Curious
+   * Hits increase the number of baserunners; why negative and insignificant?
    * Hitting doubles get runners on base; why negative?
    * Turning double plays reduce baserunners; why negative?
-   * Giving up walks allows the other team free baserunners; negative
-   coefficient makes sense but why insignificant?
    * Giving up hits allows the other team baserunners; why positive although
@@
 
-An absolutely fascinating observation. In the first draft of this exercise,
-`TEAM_BATTING_HR` was removed due to its high correlation with
-`TEAM_PITCHING_HR`. Note the coefficients, batting is almost +12 and pitching is
--9. However, this is such as strong indicator, that in the first run, pitching
-was given a *positive* coefficient of around +3. In hindsight this is because
-*it was being used as an indicator for* ***batting!!*** The correlation allowed
-the use of pitching HRs as an indicator for the hidden batting HRs! Once both
-were restored to the model, the logical coefficients surfaced. Another reason
-why models should not be trusted out of the box, but all model results should
-be reviewed for sanity and sense!!
+First, an absolutely fascinating observation from an earlier state of the model.
+In the very first draft of this exercise, prior to some of the data adjustments
+performed here, `TEAM_BATTING_HR` was removed due to its high correlation with
+`TEAM_PITCHING_HR`. In that first draft, pitching giving up home runs was given
+a *positive* coefficient. This made no sense. Allowing batting home runs to
+re-enter the model, In hindsight this becomes obvious. The pitching
+*was being used as an indicator for* ***batting!!*** They are almost 100%
+correlated! This correlation allowed the use of pitching HRs as an indicator for
+the hidden batting HRs which has a larger magnitude! Once both variables were
+restored to the model, the logical coefficients surfaced. This is yet another
+reason why models should not be trusted out of the box, but all model results
+should be reviewed for sanity and sense!!
 
 The curiosities above can *possibly* be resolved by looking at the interaction
-terms. The interaction between giving up hits and strikeouts,
-`TEAM_PITCHING_H:TEAM_PITCHING_SO `, is highly negative and significant. It
-probably captures most of the giving up hits information making the singleton
-less relevant.
+terms. The pitching and batting home run and strikeout coefficients are an order
+of magnitude greater than all the others. It is likely that the negative
+coefficients in some of the batting stats and positive coefficents in some of
+the pitching stats are being used to "temper" the effect of SOs and HRs.
 
@@
 
-The behavior of doubles remains confusing. Unless its hiding something like
-a team's propensity to strand runners on base. It would be interesting to see
-a breakdown between the American and National leagues, as the latter tends to be
-somewhat better at "small ball" and moving the runners along.
+The behavior of doubles remains confusing. Perhaps its presence in a number of
+the interaction terms means that the singleton variable needs to act as a
+balance. Another theory may be that it's hiding something like a team's
+propensity to strand runners on base. It would be interesting to see a breakdown
+between the American and National leagues, as the latter tends to be somewhat
+better at "small ball" and moving the runners along.
 
 ### Model 3
-The final model is a Stepwise Regression with Repeated k-fold Cross-Validation, and higher order polynomials variables were introduced into the full model.
+The final model is a Stepwise Regression with Repeated k-fold Cross-Validation,
+and higher order polynomials variables were introduced into the full model.
 
-A stepwise variable selection model is conducted to determine what are the variables that can help predict the number of wins for the team.
-The method allows variables to be added one at a time to the model, as long as the F-statistic is below the specified $\alpha$, in this case $\alpha = 0.05$.
-However, variables already in the model do not necessarily stay in.
-The steps evaluate all of the variables already included in the model and remove any variable that has an insignificant F-statistic.
-Only after this test ends, is the best model found, that is when none of the variables can be excluded and every variable included in the model is significant.
+A stepwise variable selection model is conducted to determine what are the
+variables that can help predict the number of wins for the team. The method
+allows variables to be added one at a time to the model, as long as the
+F-statistic is below the specified \(\alpha\), in this case \(\alpha = 0.05\).
+However, variables already in the model do not necessarily stay in. The steps
+evaluate all of the variables already included in the model and remove any
+variable that has an insignificant F-statistic. Only after this test ends, is
+the best model found, that is when none of the variables can be excluded
+and every variable included in the model is significant.
 
-Here, the dependent variable is the continuous variable, `TARGET_WINS`, and the independent variables are the full model to identify the most contributing predictors.
-In addition, a robust method for estimating the accuracy of a model, the k-fold cross-validation method, was performed evaluate the model performance on different subset of the training data and then calculate the average prediction error rate.
+Here, the dependent variable is the continuous variable, `TARGET_WINS`, and the
+independent variables are the full model to identify the most contributing
+predictors. In addition, a robust method for estimating the accuracy of a model,
+the k-fold cross-validation method, was performed evaluate the model performance
+on different subset of the training data and then calculate the average
+prediction error rate.
 
-```{r model3, include = FALSE}
+```{r model3}
 train_df = read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
@@
 **Coefficient Discussion: **
-```{r model3_discuss, echo=FALSE}
+```{r model3_discuss}
 mod_sum[["coefficients"]] %>% knitr::kable(digits = 3L, caption = 'K-fold Step Model with Higher Order Polynomials Output')
@@
 
-Studying the coefficients of the model suggest that winning is in favor if the team batting hits more doubles, triples and home runs.
-Moreover, increase in the number of stolen bases, and a decrease in caught steals, double plays, error, and walks allowed would all lead to a win for the batting team.
-It is noteworthy that the model suggests that a decrease in single hits by batter and an increase in strikeouts by batters which seems counter intuitive.
-But these variables were kept because when a batter steps to the plate, the player is more likely to strike out than to get a hit.
-Trying to hit the ball out of the park will come with strikeouts but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry out.
+Studying the coefficients of the model suggest that winning is in favor if the
+team batting hits more doubles, triples and home runs. Moreover, increase in the
+number of stolen bases, and a decrease in caught steals, double plays, error,
+and walks allowed would all lead to a win for the batting team. It is noteworthy
+that the model suggests that a decrease in single hits by batter and an increase
+in strikeouts by batters which seems counter intuitive. But these variables were
+kept because when a batter steps to the plate, the player is more likely to
+strike out than to get a hit. Trying to hit the ball out of the park will come
+with strikeouts but it will also increase the chances of hitting home runs (even
+1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry
+out.
 
-## SELECTING A MODEL
+## SELECTING A MODEL **NEEDS TO BE REDONE**
 
@@
 ## CODE APPENDIX
-
 The code chunks below represent the R code called in order during the analysis.
 They are reproduced in the appendix for review and comment.
+```{r appendix, include=FALSE}
+knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
+```
+```{r setup}
+```
+```{r loadData}
+```
+```{r wins}
+```
+```{r pctNull}
+```
+```{r negFactors}
+```
+```{r posFactors}
+```
+```{r summaryInfo}
+```
+```{r fewOutliers}
+```
+```{r manyOutliers}
+```
+```{r otherOutliers}
+```
+```{r corrPlot}
+```
+```{r outliers_1}
+```
+```{r iqr_removal_SO}
+```
+```{r iqr_removal_H}
+```
+```{r outlier_boxplot}
+```
+```{r impute_train}
+```
+```{r singles}
+```
+```{r make_slg_obp}
+```
+```{r slugging_compare}
+```
+```{r baseModel}
+```
+```{r cooks_distance}
+```
+```{r model2C+S}
+```
+```{r model2}
+```
+```{r model2_discuss}
+```
+```{r model3}
+```
+```{r model3_discuss}
+```

### 2020-09-17 HW1/HW1_draft_kellogg.Rmd
Model - Prediction Conformity updated - Deokinanan

--- 2020-09-16 HW1/HW1_draft_kellogg.Rmd
@@
 library(caret)
-train_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
-eval_set <- read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-evaluation-data.csv')
+library(data.table)
+
+urlRemote  = "https://raw.githubusercontent.com/"
+pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
+fileTrain   = "moneyball-training-data.csv"
+train_set = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
+
+fileEval   = "moneyball-evaluation-data.csv"
+eval_set = paste0(urlRemote, pathGithub, fileEval) %>% read.csv()
+
 set.seed(9450)
@@
 
-## DATA EXPLORATION
-
-### How often does the team win?
-
-We are given a data set of 2,276 records containing 15 seasonal statistics and the
-total number of wins a team had in a given year. On average, about 50% of games played 
-are won (81 games out of 162), with the best individual season having 146 wins and the worst 
-season having 0 wins. The data is normally distributed and most years have between 49 
-and 112 wins (blue lines below). The nature of the distribution means there aren't too 
-many extreme seasons where wins are significantly higher or lower than usual. This serves 
-as a good gut-check for our final predictions; if the predicted wins are too high or too 
+# DATA EXPLORATION
+## How Often Does The Team Win?
+We are given a data set of 2,276 records containing 15 seasonal statistics and
+the total number of wins a team had in a given year. On average, about 50% of
+games played are won (81 games out of 162), with the best individual season
+having 146 wins and the worst season having 0 wins. The data is normally
+distributed and most years have between 49 and 112 wins (blue lines below). The
+nature of the distribution means there aren't too many extreme seasons where
+wins are significantly higher or lower than usual. This serves as a good
+gut-check for our final predictions; if the predicted wins are too high or too 
 low, we know something in our model is probably off.
 
-```{r wins, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
+```{r wins, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
 avgWins <- mean(train_set$TARGET_WINS)
 sdev <- sd(train_set$TARGET_WINS)
-l1 <- mean(train_set$TARGET_WINS) - (2*sdev)
-l2 <- mean(train_set$TARGET_WINS) + (2*sdev)
+l1 <- mean(train_set$TARGET_WINS) - (2 * sdev)
+l2 <- mean(train_set$TARGET_WINS) + (2 * sdev)
 # histogram of wins
@@
   geom_histogram() + 
-  geom_vline(aes(xintercept= avgWins),
-            color="red", linetype="dashed", size=1)+
-  geom_vline(aes(xintercept=l1),
-            color="blue", linetype="dashed", size=1) +
-    geom_vline(aes(xintercept=l2),
-            color="blue", linetype="dashed", size=1) +
-  labs(title="Distribution of Wins",x="Number of wins", y = "Number of seasons")
+  geom_vline(aes(xintercept = avgWins),
+            color = "red", linetype = "dashed", size = 1)+
+  geom_vline(aes(xintercept = l1),
+            color="blue", linetype = "dashed", size = 1) +
+    geom_vline(aes(xintercept = l2),
+            color = "blue", linetype = "dashed", size = 1) +
+  labs(title = "Distribution of Wins", x = "Number of wins",
+       y = "Number of seasons")
 p
@@
   xlab("Statistic") + ylab("Percent Missing") + 
-  coord_flip() +
-  expand_limits(y = 1)
+  coord_flip() + expand_limits(y = 1)
 ```
 
-### Do the individual stats affect winning?
+## Do the Individual Stats Affect Winning?
+**Stats with an expected negative impact:** Intuitively, we expect that Caught
+Stealing, Errors, Hits Allowed, Home Runs Allowed, Strikeouts by Batters, and
+Walks Allowed would all have a **negative** impact on the total wins. In other
+words, as these values increase, we expect that the team is less likely to win.
 
-**Stats with an expected negative impact: ** Intuitively, we expect that Caught stealing, 
-Errors, Hits allowed, Home Runs allowed, Strikeouts by batters, and Walks allowed would 
-all have a **negative** impact on the total wins. In other words, as these values increase, 
-we expect that the team is less likely to win.
-
-```{r negFactors, echo = FALSE, message = FALSE, warning = FALSE}
+```{r negFactors, fig.align = 'center'}
 negSet <- train_set %>%
@@
 ggplot(data = negSet %>%
-  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
-  facet_wrap(~ STAT, scales = "free")
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
+  aes(x = VALUE, y = TARGET_WINS)) + 
+  geom_point() + geom_smooth(method = "lm", se = FALSE) +
+  labs(title = "Number of wins with respect to recorded stats",
+       x = "", y = "Number of Wins") + facet_wrap(~ STAT, scales = "free")
 ```
 
-When we take a closer look at the data, these negative relationships aren't obvious. In 
-fact, only **Errors** and **Hits allowed** seem to have a negative impact on wins. **Caught**
-**stealing** and **Strikeouts by batters** appear to be random; this means that whether the 
-stat for a particular season is high or low doesn't affect the overall number of wins.
+When we take a closer look at the data, these negative relationships aren't
+obvious. In fact, only **Errors** and **Hits Allowed** seem to have a negative
+impact on wins. **Caught Stealing** and **Strikeouts by Batters** appear to be
+random; this means that whether the stat for a particular season is high or low
+doesn't affect the overall number of wins.
 
-Even more interestingly, **Home Runs allowed** and **Walks allowed** have the *opposite* 
-effect; as these stats increase, so do the number of wins!
+Even more interestingly, **Home Runs Allowed** and **Walks Allowed** have the
+*opposite* effect---as these stats increase, so do the number of wins!
 
-**Stats with an expected positive impact: ** We can look at the same information for the 
-stats that we expect to have a **positive** effect on wins: Base hits, Doubles, Triples, Home 
-Runs, Walks, Batters getting hit by pitches, Stolen bases, Double Plays, and Strikeouts by 
-pitchers.
+**Stats with an expected positive impact:** We can look at the same information
+for the stats that we expect to have a **positive** effect on wins: Base Hits,
+Doubles, Triples, Home Runs, Walks, Batters getting hit by pitches, Stolen
+Bases, Double Plays, and Strikeouts by Pitchers.
 
-```{r posFactors, echo = FALSE, message = FALSE, warning = FALSE}
+```{r posFactors, fig.align = 'center'}
 posSet <- train_set %>%
@@
 
-* All variables are very narrowly distributed, meaning that most of the data falls within
-a small range. 
-* **Strikeouts by pitchers** and **Walks allowed** have a few very extreme outliers; these 
-represent seasons that have abnormally high values for the statistics. 
-* The average number of pure **Base Hits** (1469/season) is greater than the average number
-of Doubles (~241/season) and Triples (~55/season). This isn't at all surprising, but serves 
-as a good gut check on the validity of the data.
+  * All variables are very narrowly distributed, meaning that most of the data
+  falls within a small range. 
+  * **Strikeouts by Pitchers** and **Walks Allowed** have a few very extreme
+  outliers; these   represent seasons that have abnormally high values for the
+  statistics. 
+  * The average number of pure **Base Hits** (1469/season) is greater than the
+  average number of Doubles (~241/season) and Triples (~55/season). This isn't
+  at all surprising, but serves as a good gut check on the validity of the data.
 
-### Are stats correlated?
+## Are Stats Correlated?
+We would expect that a few things in the dataset might be correlated: perhaps
+number of errors and hits/homeruns allowed or the number of base hits by batters
+and doubles/triples/homeruns. We can visualize the correlations between the
+statistics to determine if there is a significant relationship between the them:
+blue dots represent positively correlated variables (as one increases, so does
+the other) and red dots represent negatively correlated variables (as one
+increases, the other decreases).
 
-We would expect that a few things in the dataset might be correlated: perhaps number of errors and
-hits/homeruns allowed or the number of base hits by batters and doubles/triples/homeruns. We can
-visualize the correlations between the statistics to determine if there is a significant relationship
-between the them: blue dots represent positively correlated variables (as one increases,
-so does the other) and red dots represent negatively correlated variables (as one increases,
-the other decreases).
-
-```{r corrPlot, echo = FALSE, message = FALSE, warning = FALSE}
+```{r corrPlot, fig.align = 'center'}
 # Correlation matrix with significance levels (p-value)
@@
 
-## BUILD MODELS
+# BUILD MODELS
+## Base Model
+We will start with a simple linear model to serve as a baseline. This includes
+all variables in the dataset.
 
-### Base Model
-We will start with a simple linear model to serve as a baseline. This includes all variables 
-in the dataset.
-
-```{r baseModel, include = FALSE}
+```{r baseModel}
 lm_reg = lm(data=train_set, TARGET_WINS ~ .)
-summary(lm_reg)
 ```
 
-### Model 1
+## Model 1
 Model 1 includes all variables except Singles, Triples, Base Hits, Strikeouts by
-pitchers, Walks allowed, and Batters Hit by Pitch. The final variables were chosen as 
-a result of backwards dplyr::selection based on null hypothesis testing for non-zero slope.
+pitchers, Walks allowed, and Batters Hit by Pitch. The final variables were
+chosen as a result of backwards `dplyr::selection` based on null hypothesis
+testing for non-zero slope.
 
-**Further data transformation: **  
-We will use Cook's distance to remove outliers that are influencing the fit of the model above. We will use a cutoff of $\frac{4}{N}$.
+### Further data transformation
+We will use Cook's distance to remove outliers that are influencing the fit of
+the model above. We will use a cutoff of \(\frac{4}{N}\).
 
-```{r cooks_distance, include=TRUE}
-lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
-          TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
-          TEAM_PITCHING_HR+ TEAM_PITCHING_H+
-          TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_set)
+```{r cooks_distance}
+lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
+          TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB +
+            TEAM_BASERUN_CS + TEAM_PITCHING_HR + TEAM_PITCHING_H +
+            TEAM_FIELDING_E + TEAM_FIELDING_DP + SLG, data = train_set)
 cooks_dis<- cooks.distance(lm_1)
-influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4/nrow(train_set)))])
+influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4 / nrow(train_set)))])
 train_df2<- train_set[-influential, ]
-lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B +TEAM_BATTING_HR+
-          TEAM_BATTING_BB+TEAM_BATTING_SO+ TEAM_BASERUN_SB+TEAM_BASERUN_CS+
-          TEAM_PITCHING_HR+ TEAM_PITCHING_H+
-          TEAM_FIELDING_E+ TEAM_FIELDING_DP+  SLG, data=train_df2)
-
+lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
+            TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB +
+            TEAM_BASERUN_CS + TEAM_PITCHING_HR + TEAM_PITCHING_H +
+            TEAM_FIELDING_E + TEAM_FIELDING_DP + SLG, data = train_df2)
 mod_sum1 = summary(lm_1)  
-pred_lm1 = predict(lm_1, newdata = train_df2)
-knitr::kable(summary(lm_1)$coefficients, digits = 3L, caption = 'Backwards Model Regression Output')
+```
+```{r model1_coef}
+knitr::kable(summary(lm_1)$coefficients, digits = 3L,
+             caption = 'Backwards Model Regression Output')
 ```
 
-**Coefficient Discussion: **
-First, we are keeping the coefficient for `TEAM_BATTING_HR` since it's p-value is marginally above the general threshold and knowledge of the game suggests it is important.
-There are some counter intuitive results, which is expected given that baseball is a messy, imprecise game that has evolved over time.
-To that end, we should expect some seemingly strange results from algorithmic regression.
-We used a combination of domain knowledge and data analysis to justify retaining features.
+### Coefficient Discussion
+First, we are keeping the coefficient for `TEAM_BATTING_HR` since it's p-value
+is marginally above the general threshold and knowledge of the game suggests it
+is important. There are some counter intuitive results, which is expected given
+that baseball is a messy, imprecise game that has evolved over time. To that
+end, we should expect some seemingly strange results from algorithmic
+regression. We used a combination of domain knowledge and data analysis to
+justify retaining features.
 
-Fielding double plays and batting doubles both appear to have negative impacts on wins even though they _should_ be positive impacts.
-Turning double plays, while a good for the defensive team, may suggest a larger, negative issue.
-Namely, weak pitching that leads to runners on base.
-Similarly, batting doubles, leaves runners open to double plays.
-Allowed hits by pitching, caught stealing, and batting strike-outs are all counter intuitive results as well, but they seem to be small contributors and these are events that happen regularly in every game.
+Fielding double plays and batting doubles both appear to have negative impacts
+on wins even though they *should* be positive impacts. Turning double plays,
+while a good for the defensive team, may suggest a larger, negative issue.
+Namely, weak pitching that leads to runners on base. Similarly, batting doubles,
+leaves runners open to double plays. Allowed hits by pitching, caught stealing,
+and batting strike-outs are all counter intuitive results as well, but they seem
+to be small contributors and these are events that happen regularly in every
+game.
 
 Slugging as an approximation is the major predictor in this regression, by far.
-The other predictors other than the intercept are orders of magnitude less in predictive value.
-It should be noted that slugging alone is not a good predictor for wins overall.
+The other predictors other than the intercept are orders of magnitude less in
+predictive value. It should be noted that slugging alone is not a good predictor
+for wins overall.
 
-### Model 2
-This is a model allowing pairwise interactions within the data types of baserunning,
-batting, pitching, and fielding, fit using a forward and backwards stepwise regression.
+## Model 2
+This is a model allowing pairwise interactions within the data types of
+baserunning, batting, pitching, and fielding. It is fit using a forward and
+backwards stepwise regression based on AIC.
 
-```{r model2, include = FALSE}
+### Further data transformation
+All the features will be centered and scaled. The target variable will be left
+in normal space. This will make prediction easier.
+
+```{r model2C+S}
+ts2 <- train_set[, -(17:19)]
+ts2 <- data.frame(apply(ts2, 2, scale))
+ts2$TARGET_WINS <- train_set$TARGET_WINS
+```
+
+A fully specified model with all the interactions will be the starting point for
+the stepwise regression, which uses AIC as its target metric.
+
+```{r model2}
 crazyModel <- lm(TARGET_WINS ~
@@
   (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
-  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = train_set)
+  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = ts2)
 stepModel <- stepAIC(crazyModel, scope = list(upper = ~
@@
 
-The behavior of doubles remains confusing. Unless its hiding something like
-a team's propensity to strand runners on base. It would be interesting to see
-a breakdown between the American and National leagues, as the latter tends to be
-somewhat better at "small ball" and moving the runners along.
+The behavior of doubles remains confusing. Perhaps its presence in a number of
+the interaction terms means that the singleton variable needs to act as a
+balance. Another theory may be that it's hiding something like a team's
+propensity to strand runners on base. It would be interesting to see a breakdown
+between the American and National leagues, as the latter tends to be somewhat
+better at "small ball" and moving the runners along.
 
 ### Model 3
-The final model is a Stepwise Regression with Repeated k-fold Cross-Validation, and higher order polynomials variables were introduced into the full model.
+The final model is a Stepwise Regression with Repeated k-fold Cross-Validation,
+and higher order polynomials variables were introduced into the full model.
 
-A stepwise variable selection model is conducted to determine what are the variables that can help predict the number of wins for the team.
-The method allows variables to be added one at a time to the model, as long as the F-statistic is below the specified $\alpha$, in this case $\alpha = 0.05$.
-However, variables already in the model do not necessarily stay in.
-The steps evaluate all of the variables already included in the model and remove any variable that has an insignificant F-statistic.
-Only after this test ends, is the best model found, that is when none of the variables can be excluded and every variable included in the model is significant.
+A stepwise variable selection model is conducted to determine what are the
+variables that can help predict the number of wins for the team. The method
+allows variables to be added one at a time to the model, as long as the
+F-statistic is below the specified \(\alpha\), in this case \(\alpha = 0.05\).
+However, variables already in the model do not necessarily stay in. The steps
+evaluate all of the variables already included in the model and remove any
+variable that has an insignificant F-statistic. Only after this test ends, is
+the best model found, that is when none of the variables can be excluded
+and every variable included in the model is significant.
 
-Here, the dependent variable is the continuous variable, `TARGET_WINS`, and the independent variables are the full model to identify the most contributing predictors.
-In addition, a robust method for estimating the accuracy of a model, the k-fold cross-validation method, was performed evaluate the model performance on different subset of the training data and then calculate the average prediction error rate.
+Here, the dependent variable is the continuous variable, `TARGET_WINS`, and the
+independent variables are the full model to identify the most contributing
+predictors. In addition, a robust method for estimating the accuracy of a model,
+the k-fold cross-validation method, was performed evaluate the model performance
+on different subset of the training data and then calculate the average
+prediction error rate.
 
-```{r model3, include = FALSE}
-train_df = read.csv('https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW1/data/moneyball-training-data.csv')
-train_df = subset(train_df, select = -c(INDEX, TEAM_BATTING_HBP))
-model = lm(TARGET_WINS ~ ., data = train_df)
-# Remove Outlier and more cleaning
+### Further data transformation
+Once again, Cook's distance is use to decide if an individual entity is an extreme 
+value or not.
+
+```{r model3}
+stepdata = subset(train_set, select = -c(AB, SLG, TEAM_BATTING_HBP))
+stepdata = cbind(stepdata, sapply(stepdata[2:length(stepdata)], function(x) x^2))
+names(stepdata) = make.unique(names(stepdata), sep = "_")
+model = lm(TARGET_WINS ~ ., data = stepdata)
 cooksd = cooks.distance(model)
 influential = as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
-train_df = train_df[-c(influential), ]
-train_df.clean = complete(mice(data = train_df, method = "pmm", maxit = 5, 
-                               seed = 525, print = FALSE), 3)
-train_df.clean$TEAM_BATTING_1B = train_df.clean$TEAM_BATTING_H - 
-  (train_df.clean$TEAM_BATTING_2B + train_df.clean$TEAM_BATTING_3B + 
-     train_df.clean$TEAM_BATTING_HR)
-train_df.clean = subset(train_df.clean, select = -c(TEAM_BATTING_H, TEAM_PITCHING_HR))
-# Data Transformation & Multicollinearity Diagnostic Measures
-stepdata = cbind(train_df.clean,sapply(train_df.clean[2:length(train_df.clean)], 
-                                       function(x) x^2))
-names(stepdata) = make.unique(names(stepdata), sep = "_")
+stepdata.2 = stepdata[-c(influential), ]
 # Set up repeated k-fold cross-validation
@@
 # Train the model
-step.model = train(TARGET_WINS ~ ., data = stepdata, method = "lmStepAIC", 
+step.model = train(TARGET_WINS ~ ., data = stepdata.2, method = "lmStepAIC", 
                    trControl = train.control, trace = FALSE)
-# Final model coefficients
-step.model$finalModel 
+# Final model 
+step.model = lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BASERUN_SB + 
+                  TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_FIELDING_E + 
+                  TEAM_FIELDING_DP + TEAM_BATTING_BB_1 + TEAM_BATTING_SO_1 + 
+                  TEAM_BASERUN_CS_1 + TEAM_PITCHING_H_1 + TEAM_PITCHING_BB_1 + 
+                  TEAM_FIELDING_E_1 + TEAM_FIELDING_DP_1 + TEAM_BATTING_1B_1, 
+                data = stepdata.2)
 # Summary of the model
-mod_sum = summary(step.model$finalModel)
-# Predictions on train data
-pred_step.model = predict(step.model, newdata = stepdata)
+mod_sum3 = summary(step.model)
 ```
@@
 **Coefficient Discussion: **
-```{r model3_discuss, echo=FALSE}
-mod_sum[["coefficients"]] %>% knitr::kable(digits = 3L, caption = 'K-fold Step Model with Higher Order Polynomials Output')
+```{r model3_discuss}
+mod_sum3[["coefficients"]] %>% 
+  knitr::kable(digits = 3L, 
+               caption = 'K-fold Step Model with Higher Order Polynomials Output')
 ```
 
-Studying the coefficients of the model suggest that winning is in favor if the team batting hits more doubles, triples and home runs.
-Moreover, increase in the number of stolen bases, and a decrease in caught steals, double plays, error, and walks allowed would all lead to a win for the batting team.
-It is noteworthy that the model suggests that a decrease in single hits by batter and an increase in strikeouts by batters which seems counter intuitive.
-But these variables were kept because when a batter steps to the plate, the player is more likely to strike out than to get a hit.
-Trying to hit the ball out of the park will come with strikeouts but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry out.
+Studying the coefficients of the model suggest that winning is in favor if the
+team batting hits more doubles, triples and home runs. Moreover, increase in the
+number of stolen bases, and a decrease in caught steals, double plays, error,
+and walks allowed would all lead to a win for the batting team. It is noteworthy
+that the model suggests that a decrease in single hits by batter and an increase
+in strikeouts by batters which seems counter intuitive. But these variables were
+kept because when a batter steps to the plate, the player is more likely to
+strike out than to get a hit. Trying to hit the ball out of the park will come
+with strikeouts but it will also increase the chances of hitting home runs (even
+1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry
+out.
 
-## SELECTING A MODEL
+## SELECTING A MODEL & PREDICTIONS
 
@@
 
-When comparing the three models, it was interesting to select which would be our best model given that their performance statistics and error measurements were not significantly different from each other.
-It is apparent that Model #1: Backwards accounts for nearly 44% of the variation in the dependent variable with the independent variables, which is acceptable as a good model.
-But it's explanatory power based on an $adj R^2 = 0.414$ suggests it is no different from Model #3: K-fold.
-In addition, Model #1: Backwards has a highest RSME of all the models, which ranks as one of the major criteria we are using to decide on a model.
-Now, Model #2: Pairwise resulted in the $R^2 = 0.427$, $adjR^2 = 0.417$, and the lowest $RSME = 10.193$.
-Because the RMSE and adjusted $R^2$ statistics already include a minor adjustment for the number of coefficients estimated, to evaluate the model complexity, we compared the AIC.
-While a smaller AIC is deemed less complex, Model #1: Backwards would have been the way to go, however, it was deemed that the AIC differences among the models are quite insignificant.
-Thus, it is unanimous that Model #2: Pairwise is our final model.
-
-```{r performance, echo=FALSE, message=FALSE, warning=FALSE}
+```{r performance}
 # Model accuracy
-results = data.frame(Models = c("Model #1: Backwards" , "Model #2: Pairwise", "Model #3: K-fold"),
-                R.squared = c(mod_sum1[["r.squared"]], stmr2, mod_sum[["r.squared"]]), 
-                adj.R.squared = c(mod_sum[["adj.r.squared"]], stmar2, mod_sum[["adj.r.squared"]]),
-                AIC = c(AIC(lm_1), AIC(stepModel), AIC(step.model$finalModel)),
-                RSME = c(sqrt(mean((train_set$TARGET_WINS - pred_lm1)^2)), stmRMSE,
-                         sqrt(mean((stepdata$TARGET_WINS - pred_step.model)^2))))
-results %>% knitr::kable(digits = 3L, caption = 'Performance Statistics & Error Measure of Models' )
+results = data.frame(
+  Models = c("Model #1: Backwards" , "Model #2: Pairwise", "Model #3: K-fold"),
+  R.squared = c(mod_sum1[["r.squared"]], stmr2, mod_sum3[["r.squared"]]), 
+  adj.R.squared = c(mod_sum1[["adj.r.squared"]], stmar2, mod_sum3[["adj.r.squared"]]),
+  RSME = c(sqrt(mean(lm_1$residuals ^ 2)), stmRMSE, sqrt(mean(step.model$residuals ^ 2))),
+  AIC = c(AIC(lm_1), AIC(stepModel), AIC(step.model)))
+results %>% knitr::kable(digits = 3L, 
+                         caption = 'Performance Statistics & Error Measure of Models')
 ```
 
-## PREDICTIONS
+It was an interesting process when it came comparing the three models and selecting which would be our best model given that their performance statistics and error measurements were not significantly different from each other.
+From the table above, it is apparent that Model #3: K-fold accounts for nearly 45% of the variation in the dependent variable with the independent variables because $R^2 = 0.456$, $adjR^2 = 0.450$ which is acceptable as a good model.
+However, Model #1: Backwards has a smallest RSME of all the models, which ranks as number one as the criteria we are using to decide on a model.
+This measure suggest how far off an actual value is from the model's prediction for that value, and it indicates that there is a greater reduction in the randomness for Model #1: Backwards than the other model.
+But, how different is the RSME of Model #1: Backwards versus the RSME of Model #3: K-fold?
+We agreed that it is insignificant.
+Therefore, because the RMSE and adjusted $R^2$ statistics already include a minor adjustment for the number of coefficients estimated, to evaluate the model complexity, we compared the AIC.
+As a result, Model #1: Backwards would have been the way to go, however, it was deemed that the AIC difference with Model #3 is also insignificant.
 
-Using the test data and the selected final model, Model #2: Pairwise, a comparison in the prediction statistic was conducted.
+### PREDICTIONS
 
-```{r predictions, echo=FALSE, message=FALSE, warning=FALSE}
-# Cleaning test set for Model#1
-test_df1<-  eval_set %>% dplyr::select(-c(TEAM_BATTING_HBP, INDEX)) %>%
-  mutate(TEAM_BATTING_1B =
-         TEAM_BATTING_H -
-         (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR)) %>%
-  mutate_if(is.integer, as.numeric)
-test_df1 <- complete(mice(data = test_df1,
-                         method = "pmm",
-                         seed = 9450,
-                         print=FALSE), 3)
+Using the evaluation data set and the three models, a comparison in the predictions statistic with the training data was conducted.
+Interestingly, while Model #1: Backwards and Model #3: K-fold resulted in better performance statistics and error measurements, it is evident that Model #2: Pairwise yields closer predictions when the other two slightly under-predicts.
+With all in agreement, under-predicting wins and leaving teams and managers with potential perspective of costs is worse than over-predicting and causing inconvenience.
 
-test_df1<-
-  test_df1 %>%  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
-         SLG = (TEAM_BATTING_1B + (2*TEAM_BATTING_2B)+ (3*TEAM_BATTING_3B)+ (4*TEAM_BATTING_HR))/AB) %>% dplyr::select(-c(AB, TEAM_BATTING_H))
+```{r eval_set}
+eval_set = eval_set %>% na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
+  bind_cols(eval_set) %>%
+  filter(TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO, probs = c(0.25), na.rm = TRUE) 
+         - 1.5 * iqr, 
+         TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, probs = c(0.75), na.rm = TRUE) 
+         + 1.5 * iqr) %>%
+  dplyr::select(-iqr)
 
-# Cleaning test set for Model #2
-BxCox <- function(x, l) {
-  ifelse(l == 0, log(x), (x ^ l - 1) / l)
-}
-BCL <- data.table(Metric = c('TEAM_PITCHING_H', 'TEAM_FIELDING_E',
-                             'TEAM_BASERUN_SB', 'TEAM_PITCHING_HR',
-                             'TEAM_BATTING_HR', 'TEAM_PITCHING_BB',
-                             'TEAM_PITCHING_SO', 'TEAM_BATTING_3B',
-                             'TEAM_BASERUN_CS'),
-                  Lambda = c(-2.5, -0.7, 0, 0.6, 0.6, -0.1, -0.35, 0, 0.5))
-setkey(BCL, Metric)
-MBT <- setDT(eval_set)
-#MBT[, `:=`(TEAM_BATTING_HBP = NULL)]
-MBTlng <- melt(MBT, id.vars = 'INDEX', variable.name = 'Metric',
-               value.name = 'Value', variable.factor = FALSE)
-MBTlng[, Value := as.double(Value)]
-MBTlng <- BCL[MBTlng, on = 'Metric']
-MBTlng[!is.na(Lambda), Value := BxCox(Value, Lambda)]
-MBTlng[, Lambda := NULL]
-MBTlngSC <- copy(MBTlng)
-# Remove infinities
-MBTlngSC[, Value := ifelse(is.infinite(Value), NA, Value)]
-# Impute, Center, and Scale
-MBTlngSC[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
-       keyby = Metric]
-MBTlngSC[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
-MBTSC <- dcast(MBTlngSC, INDEX ~ Metric, value.var = 'Value')
+eval_set = eval_set %>% na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
+  bind_cols(eval_set) %>%
+  filter(TEAM_PITCHING_H > quantile(TEAM_PITCHING_H, probs = c(0.25), na.rm = TRUE) 
+         - 1.5 * iqr,
+         TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, probs = c(0.75), na.rm = TRUE) 
+         + 1.5 * iqr) %>%
+  dplyr::select(-iqr)
 
-# Cleaning test set for Model #3
-test_df = complete(mice(data = eval_set, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
-test_df$TEAM_BATTING_1B = test_df$TEAM_BATTING_H - 
-  (test_df$TEAM_BATTING_2B + test_df$TEAM_BATTING_3B + test_df$TEAM_BATTING_HR)
-test_df = cbind(test_df, sapply(test_df[1:length(test_df)], function(x) x^2))
-names(test_df) = make.unique(names(test_df), sep = "_")
+eval_set = complete(mice(data = eval_set, method = "pmm", seed = 9450,  print=FALSE), 3)
+
+eval_set = eval_set %>% dplyr::select(-INDEX) %>%
+  mutate_if(is.integer, as.numeric) %>%
+  mutate(TEAM_BATTING_1B = TEAM_BATTING_H - (TEAM_BATTING_2B + TEAM_BATTING_3B +
+                                               TEAM_BATTING_HR))
+
+# Cleaning Test Set for Model #1
+test_df1 = eval_set %>%
+  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
+         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
+                  4 * TEAM_BATTING_HR) / AB)
+# Cleaning Test Set for Model #2
+test_df2 = test_df1[, -(17:19)]
+test_df2 = data.frame(apply(test_df2, 2, scale))
+# Cleaning Test Set for Model #3
+test_df3 = cbind(test_df1, sapply(test_df1[1:length(test_df1)], function(x) x^2))
+names(test_df3) = make.unique(names(test_df3), sep = "_")
 ```
 
-```{r echo=FALSE}
-# Model #1
-predictions = predict(lm_1, newdata = test_df1)
-results = cbind(dataset = c('Training Data', 'Test Prediction'), 
+```{r predictions}
+predictions = data.frame(predict(lm_1, newdata = test_df1), 
+                         predict(stepModel, test_df2), 
+                         predict(step.model, newdata = test_df3))
+results = cbind(
+  Datasets = c('Training Data', 'Model #1: Backwards', 'Model #2: Pairwise',
+               'Model #3: K-fold'), 
   rbind(psych::describe(train_set$TARGET_WINS), psych::describe(predictions))) 
 rownames(results) = NULL
-results[,-c(2,8,11)] %>% knitr::kable(digits = 2L, caption = 'Prediction Comparison with Model #1: Backwards')
+results[,-c(2,8,11)] %>% 
+  knitr::kable(digits = 2L, caption = 'Prediction Comparison on Evaluation Set')
+```
 
-# Model #2
-predictions <- predict(stepModel, MBTSC)
-results = cbind(dataset = c('Training Data', 'Test Prediction'), 
-                rbind(psych::describe(train_set$TARGET_WINS), psych::describe(predictions))) 
-rownames(results) = NULL
-results[,-c(2,8,11)] %>% knitr::kable(digits = 2L, caption = 'Prediction Comparison with Model #2: Pairwise')
-
-# Model #3
-predictions = predict(step.model$finalModel, newdata = test_df)
-results = cbind(dataset = c('Training Data', 'Test Prediction'), 
-                rbind(psych::describe(train_set$TARGET_WINS), psych::describe(predictions))) 
-rownames(results) = NULL
-results[,-c(2,8,11)] %>% knitr::kable(digits = 2L, caption = 'Prediction Comparison with Model #3: K-fold')
-```
+Therefore, based on these reasoning, it is unanimous that **Model #2: Pairwise** is our final model because the coefficients of the model with it's the interacting variables makes for better, realistic predictions.
 
@@
 ## CODE APPENDIX
-
 The code chunks below represent the R code called in order during the analysis.
 They are reproduced in the appendix for review and comment.
+```{r appendix, include=FALSE}
+knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
+```
+```{r setup}
+```
+```{r loadData}
+```
+```{r wins}
+```
+```{r pctNull}
+```
+```{r negFactors}
+```
+```{r posFactors}
+```
+```{r summaryInfo}
+```
+```{r fewOutliers}
+```
+```{r manyOutliers}
+```
+```{r otherOutliers}
+```
+```{r corrPlot}
+```
+```{r outliers_1}
+```
+```{r iqr_removal_SO}
+```
+```{r iqr_removal_H}
+```
+```{r outlier_boxplot}
+```
+```{r impute_train}
+```
+```{r singles}
+```
+```{r make_slg_obp}
+```
+```{r slugging_compare}
+```
+```{r baseModel}
+```
+```{r cooks_distance}
+```
+```{r model2C+S}
+```
+```{r model2}
+```
+```{r model3}
+```
+```{r performance}
+```
+```{r eval_set}
+```
+```{r predictions}
+```

### 2020-09-17 Adler_personal/Blog2.Rmd
Blog2

---
title: "Blog Post 2: Training & Testing: Linear Regression"
subtitle: "DT 621---Fall 2020"
author: "Avraham Adler"
date: "9/17/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, error=FALSE)
```

# Training & Testing Sets
When originally taught, most students perform their regression analysis on the
totality of their data. If one is solely interested in past behavior, this is
proper. However, if one is interested in predicting future behavior, it is
better practice to reserve some of the data you have for testing purposes. This
reflects the "unknown" future data which is the true intent of the model.

# Experiment
We will create synthetic data representing both past and future observations.
A quarter of the data will be the "true" future data. This will never be seen
during the modeling process. The remaining 75% reflects the "observed"
data---the full complement of what is available to the data scientist.

Two models will be built using this data. One using all of it and the second
using A 75/25 split into a training and testing set. The selected models from
each group will then be compared against the "future" data.

# Raw Data
We will use a simple linear model as our example:
\[
y_i = 3.2 + 1.6x_{1i} + 3.4x_{2i} - 0.8x_{3i} + \epsilon_i
\]

We will call the model using glm with a Gaussian link.

```{r loadPackages}
library(caret)
```
```{r createData}
set.seed(245)
x1 <- runif(400, 0, 100)
x2 <- rnorm(400, 0, 10)
x3 <- rexp(400, 1/5)
y <- 3.2 + 1.6 * x1 + 3.4 * x2 - 0.8 * x3 + rnorm(400, 0, 10)
futIDX <- sample(400, 100)
x1f <- x1[futIDX]
x1p <- x1[-futIDX]
x2f <- x2[futIDX]
x2p <- x2[-futIDX]
x3f <- x3[futIDX]
x3p <- x3[-futIDX]
yf <- y[futIDX]
yp <- y[-futIDX]
testIDX <- sample(300, 300 * 0.25)
x1trn <- x1p[-testIDX]
x1tst <- x1p[testIDX]
x2trn <- x2p[-testIDX]
x2tst <- x2p[testIDX]
x3trn <- x3p[-testIDX]
x3tst <- x3p[testIDX]
ytrn <- yp[-testIDX]
ytst <- yp[testIDX]
```

# Linear Regression Model
This is almost unfair with regards to the training/testing split. Normally, the
validation set is used to fine tune the hyperparameters. In simple linear
regression there are none. If the second model returns better results, it is
very strong evidence that there is value in simply fitting to less data to
prevent overfitting!

## Training
```{r linearFit}
trc <- trainControl(method = 'cv', number = 10)
m1 <- train(data.frame(x1 = x1p, x2 = x2p, x3 = x3p), yp, method = 'glm',
            trControl = trc, family = gaussian)
m2 <- train(data.frame(x1 = x1trn, x2 = x2trn, x3 = x3trn), ytrn, method = 'glm',
            trControl = trc, family = gaussian)
summary(m1)
summary(m2)
```

Both models recover the parameters decently, although the intercept is somewhat
off. The last is almost always due to process risk, the natural vagaries of
stichastic simulation.

## Testing
```{r linearTest}
m1p <- predict(m1, data.frame(x1 = x1f, x2 = x2f, x3 = x3f))
m2p <- predict(m2, data.frame(x1 = x1f, x2 = x2f, x3 = x3f))
defaultSummary(data.frame(pred = m1p, obs = yf))
defaultSummary(data.frame(pred = m2p, obs = yf))
```

The second model, fit on less data, has better accuracy on the "future" data.
This can only be due to overfitting. In a future blog post, we may investigate
results when there **are** hyperparameters which can be tuned.

### 2020-09-18 HW1/HW1_draft_kellogg.Rmd
Significant rewrite, especially of the conclusion, in light of model performance. Added TOC. Spelling and Grammar reviewed.

@@
 output:
-  pdf_document: default
+  pdf_document:
+    toc: TRUE
+    toc_depth: 4
   html_document: default
@@
   never happened.
-  * **TEAM_PITCHING_H**: The largest outlier for this vaiable is over 30,000
-  which is simlarly highly unlikely.
+  * **TEAM_PITCHING_H**: The largest outlier for this variable is over 30,000
+  which is similarly highly unlikely.
 
@@
 
-# BUILD MODELS
-## Base Model
+# MODEL TRAINING
+## Model 1
+### Base Model
 We will start with a simple linear model to serve as a baseline. This includes
@@
 
-## Model 1
+### Enhanced Model
 Model 1 includes all variables except Singles, Triples, Base Hits, Strikeouts by
@@
 
+### Stepwise Model
 A fully specified model with all the interactions will be the starting point for
-the stepwise regression, which uses AIC as its target metric.
+the stepwise regression, which uses AIC as its target metric. For interest, this
+model will be compared with both the null model and the fully-specified model
+from where the stepwise process begins
 
 ```{r model2}
-crazyModel <- lm(TARGET_WINS ~
+nullModel <- lm(TARGET_WINS ~ 1, data = ts2)
+fullModel <- lm(TARGET_WINS ~
   (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
@@
   TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = ts2)
-stepModel <- stepAIC(crazyModel, scope = list(upper = ~
+stepModel <- stepAIC(fullModel, scope = list(upper = ~
   (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
@@
   direction = 'both', trace = 0)
+snm <- summary(nullModel)
+sfm <- summary(fullModel)
 sstm <- summary(stepModel)
@@
 stmar2 <- sstm$adj.r.squared
-stmAIC <- prettyNum(AIC(stepModel), digits = 2L, big.mark = ',')
+stmAIC <- AIC(stepModel)
 stmRMSE <- sqrt(mean(stepModel$residuals ^ 2))
-stmF <- prettyNum(sstm$fstatistic, digits = 5L)
-stmFp <- pf(sstm$fstatistic[[1]], sstm$fstatistic[[2]], sstm$fstatistic[[3]],
-            lower.tail = FALSE)
 ```
 
-### Coefficient Discussion:
+### Discussion
+#### Models
 ```{r model2_discuss}
+knitr::kable(snm$coefficients, digits = 3L,
+             caption = 'Null Model')
+knitr::kable(sfm$coefficients, digits = 3L,
+             caption = 'Full Model with All Features & Pairwise Class Interactions')
 knitr::kable(sstm$coefficients, digits = 3L,
-             caption = 'Step Model with Pairwise Class Interactions Output')
+             caption = 'Stepwise Model with Pairwise Class Interactions')
 ```
 
+The null model expects teams to lose slightly more than half their games. The
+full model has a lower intercept, which indicates that the impact of all the
+features explains why teams *win* slightly more than just the grand mean does.
+The stepwise model is similar, having a lower intercept than the null but a
+higher intercept that the full model.
+
+Also, the full model has the higher \(R^2\) and the stepwise model has the
+higher *adjusted* \(R^2\). This is expected, as there is no penalty for adding
+more parameters in regular \(R^2\) measures.
+
+The **stepwise** model will be the one discussed below.
+
+#### Coefficients
 Most of the coefficients are reasonable within the context of the game of
@@
 of magnitude greater than all the others. It is likely that the negative
-coefficients in some of the batting stats and positive coefficents in some of
+coefficients in some of the batting stats and positive coefficients in some of
 the pitching stats are being used to "temper" the effect of SOs and HRs.
@@
 
-### Model 3
+## Model 3
 The final model is a Stepwise Regression with Repeated k-fold Cross-Validation,
@@
 ```{r model3}
-stepdata = subset(train_set, select = -c(AB, SLG, TEAM_BATTING_HBP))
-stepdata = cbind(stepdata, sapply(stepdata[2:length(stepdata)], function(x) x^2))
-names(stepdata) = make.unique(names(stepdata), sep = "_")
-model = lm(TARGET_WINS ~ ., data = stepdata)
-cooksd = cooks.distance(model)
-influential = as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
-stepdata.2 = stepdata[-c(influential), ]
+stepdata <- subset(train_set, select = -c(AB, SLG, TEAM_BATTING_HBP))
+stepdata <- cbind(stepdata, sapply(stepdata[2:length(stepdata)], function(x) x^2))
+names(stepdata) <- make.unique(names(stepdata), sep = "_")
+model <- lm(TARGET_WINS ~ ., data = stepdata)
+cooksd <- cooks.distance(model)
+influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
+stepdata.2 <- stepdata[-c(influential), ]
 # Set up repeated k-fold cross-validation
 set.seed(525)
-train.control = trainControl(method = "cv", number = 10)
+train.control <- trainControl(method = "cv", number = 10)
 # Train the model
-step.model = train(TARGET_WINS ~ ., data = stepdata.2, method = "lmStepAIC", 
+step.model <- train(TARGET_WINS ~ ., data = stepdata.2, method = "lmStepAIC", 
                    trControl = train.control, trace = FALSE)
 # Final model 
-step.model = lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BASERUN_SB + 
+step.model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BASERUN_SB + 
                   TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_FIELDING_E + 
@@
 # Summary of the model
-mod_sum3 = summary(step.model)
+mod_sum3 <- summary(step.model)
 ```
 
-**Coefficient Discussion: **
+### Coefficient Discussion
 ```{r model3_discuss}
@@
 
-## SELECTING A MODEL & PREDICTIONS
+# MODEL SELECTION
+## Criteria
+In order to select the best model to make predictions, we looked at the
+following goodness of fit metrics:
 
-In order to select the best model to make predictions, we looked at some measurements that tells us how well each model fits the training.
-These include the 1) $R^2$, which represents the proportion of the variance explained by a model; 2) $adj R^2$, which is a modified version of $R^2$; 3) *Root Mean Squared Error* (RMSE), which is the square root of the mean squared error; and 4) *Akaike Information Criterion* (AIC), which is an estimator of out-of-sample prediction error.
+ 1. \(R^2\), which represents the proportion of the variance explained by the
+ model
+ 2. Adjusted \(R^2\), which is a modified version of \(R^2\) taking parsimony
+ into consideration
+ 3. *Root Mean Squared Error* (RMSE), which is the square root of the mean
+ squared difference between the observation and the fitted value
+ 4. *Akaike Information Criterion* (AIC), which is an information-criteria based
+ estimate of the Kullback-Leibler divergence from the current model to the
+ "true" model. One of the nice results of the theory of information-criteria
+ based measures is that there is included recognition of parsimony.
 
+## Performance
 ```{r performance}
 # Model accuracy
-results = data.frame(
+results <- data.frame(
   Models = c("Model #1: Backwards" , "Model #2: Pairwise", "Model #3: K-fold"),
   R.squared = c(mod_sum1[["r.squared"]], stmr2, mod_sum3[["r.squared"]]), 
-  adj.R.squared = c(mod_sum1[["adj.r.squared"]], stmar2, mod_sum3[["adj.r.squared"]]),
-  RSME = c(sqrt(mean(lm_1$residuals ^ 2)), stmRMSE, sqrt(mean(step.model$residuals ^ 2))),
+  adj.R.squared = c(mod_sum1[["adj.r.squared"]], stmar2,
+                    mod_sum3[["adj.r.squared"]]),
+  RSME = c(sqrt(mean(lm_1$residuals ^ 2)), stmRMSE,
+           sqrt(mean(step.model$residuals ^ 2))),
   AIC = c(AIC(lm_1), AIC(stepModel), AIC(step.model)))
@@
 
-It was an interesting process when it came comparing the three models and selecting which would be our best model given that their performance statistics and error measurements were not significantly different from each other.
-From the table above, it is apparent that Model #3: K-fold accounts for nearly 45% of the variation in the dependent variable with the independent variables because $R^2 = 0.456$, $adjR^2 = 0.450$ which is acceptable as a good model.
-However, Model #1: Backwards has a smallest RSME of all the models, which ranks as number one as the criteria we are using to decide on a model.
-This measure suggest how far off an actual value is from the model's prediction for that value, and it indicates that there is a greater reduction in the randomness for Model #1: Backwards than the other model.
-But, how different is the RSME of Model #1: Backwards versus the RSME of Model #3: K-fold?
-We agreed that it is insignificant.
-Therefore, because the RMSE and adjusted $R^2$ statistics already include a minor adjustment for the number of coefficients estimated, to evaluate the model complexity, we compared the AIC.
-As a result, Model #1: Backwards would have been the way to go, however, it was deemed that the AIC difference with Model #3 is also insignificant.
+It was interesting to compare the three models given that their performance
+statistics and error measurements were not significantly different from each other.
+**Model #2: Pairwise** was the worst performing model for all of our metrics, and
+was eliminated immediately. **Model #3: K-fold** accounts for nearly 45% of the
+variation in the dependent variable with the independent variables because its
+\(R^2\) is 0.456 and its adjusted \(R^2\) is = 0.450, both of which are
+acceptable for a good model. However, **Model #1: Backwards** has the smallest
+RSME of all the models. RMSE is a measure of how far off predictions are from
+the actual, giving extra weight to outliers through summing the squared
+differences. This demonstrates a greater reduction in the randomness by
+**Model #1: Backwards**. But its difference from **Model #3: K-fold** is only
+0.022. It is hard to judge if that is a significant difference. We thus also
+considered AIC. With AIC, it is not the magnitude of the value which is
+important, but the difference between the various model AIC's. The magnitude is
+a function of the number of observations. Accepted rules of thumb in the
+statistical literature are that if the difference is greater than 10, there is
+effectively no support for the model with the larger AIC
+(Section 2.6, Burnham & Anderson 2002). As a result, **Model #1: Backwards**
+is our selected model.
 
-### PREDICTIONS
+## Discussion
+This is an interesting model in that given no other information, a team would be
+expected to win negative games, which is clearly impossible. That being said, it
+is an indication of the predictive power of the engineereed **slugging** feature
+that allows small increments in slugging to result in significant shifts in
+target wins. This engineered feature contains within it numerous interactions.
+It combines scaled versions of some features and normalizes them for the
+combination of other features. Perhaps that is why it is so significant.
 
-Using the evaluation data set and the three models, a comparison in the predictions statistic with the training data was conducted.
-Interestingly, while Model #1: Backwards and Model #3: K-fold resulted in better performance statistics and error measurements, it is evident that Model #2: Pairwise yields closer predictions when the other two slightly under-predicts.
-With all in agreement, under-predicting wins and leaving teams and managers with potential perspective of costs is worse than over-predicting and causing inconvenience.
+# PREDICTIONS
+**Model #1: Backwards** was used to predict the target wins base on the
+evaluation set, once processed similarly to the training data. As the number of
+games won must be integral, the values were rounded to the closest integer.
 
-```{r eval_set}
+```{r eval_set, eval=FALSE}
 eval_set = eval_set %>% na.omit() %>%
@@
 ```{r predictions}
-predictions = data.frame(predict(lm_1, newdata = test_df1), 
-                         predict(stepModel, test_df2), 
-                         predict(step.model, newdata = test_df3))
-results = cbind(
-  Datasets = c('Training Data', 'Model #1: Backwards', 'Model #2: Pairwise',
-               'Model #3: K-fold'), 
-  rbind(psych::describe(train_set$TARGET_WINS), psych::describe(predictions))) 
-rownames(results) = NULL
-results[,-c(2,8,11)] %>% 
-  knitr::kable(digits = 2L, caption = 'Prediction Comparison on Evaluation Set')
+eval_set <- paste0(urlRemote, pathGithub, fileEval) %>% read.csv()
+eval_set <- eval_set %>% na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
+  bind_cols(eval_set) %>%
+  filter(TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO, probs = c(0.25), na.rm = TRUE) 
+         - 1.5 * iqr, 
+         TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, probs = c(0.75), na.rm = TRUE) 
+         + 1.5 * iqr) %>%
+  dplyr::select(-iqr)
+
+eval_set <- eval_set %>% na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
+  bind_cols(eval_set) %>%
+  filter(TEAM_PITCHING_H > quantile(TEAM_PITCHING_H, probs = c(0.25), na.rm = TRUE) 
+         - 1.5 * iqr,
+         TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, probs = c(0.75), na.rm = TRUE) 
+         + 1.5 * iqr) %>%
+  dplyr::select(-iqr)
+
+eval_set <- complete(mice(data = eval_set, method = "pmm", seed = 9450, print=FALSE), 3)
+
+eval_set <- eval_set %>% dplyr::select(-INDEX) %>%
+  mutate_if(is.integer, as.numeric) %>%
+  mutate(TEAM_BATTING_1B = TEAM_BATTING_H - (TEAM_BATTING_2B + TEAM_BATTING_3B +
+                                               TEAM_BATTING_HR))
+eval_set <- eval_set %>%
+  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
+         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
+                  4 * TEAM_BATTING_HR) / AB)
+predictions <- round(predict(lm_1, newdata = eval_set))
+summary(predictions)
+predictions
 ```
 
-Therefore, based on these reasoning, it is unanimous that **Model #2: Pairwise** is our final model because the coefficients of the model with it's the interacting variables makes for better, realistic predictions.
+# REFERENCES
 
-## CONCLUSIONS
+ * Burnham, Kenneth P., and David R. Anderson. 2002. Model Selection and
+ Multimodel Inference: A Practical Information-Theoretic Approach. Second.
+ New York: Springer Science+Business Media, Inc.
 
-## CODE APPENDIX
+# CODE APPENDIX
 The code chunks below represent the R code called in order during the analysis.
@@
 ```
-```{r eval_set}
-```
+<!--```{r eval_set}
+```-->
 ```{r predictions}

### 2020-09-18 HW1/HW1_draft_kellogg.Rmd
amber's edits to final draft

@@
     toc: TRUE
-    toc_depth: 4
+    toc_depth: 2
   html_document: default
@@
 ```{r loadData, include=FALSE}
+##################### libraries and variables 
 library(tidyverse)
@@
 ```{r wins, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
+##################### DATA EXPLORATION
+
 avgWins <- mean(train_set$TARGET_WINS)
@@
 ```{r pctNull, fig.height = 2, fig.width = 5, fig.align = 'center'}
+
+# percent null for each variable
 pct_null <- data.frame(do.call("rbind", map(train_set %>% dplyr::select(-INDEX),
@@
 ```{r negFactors, fig.align = 'center'}
+# variables expected to have a negative impact on wins
 negSet <- train_set %>%
@@
 ```{r posFactors, fig.align = 'center'}
+# variables expected to have a positive impact on wins
 posSet <- train_set %>%
@@
 ```{r summaryInfo, fig.height = 3, fig.width = 5, fig.align = 'center'}
-#summary(train_set)
+# understanding outliers by variable
 totalOutliers <- data.frame(
@@
 ```{r fewOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
+# distributions of variables with few outliers
 set1 <- train_set %>%
@@
 ```{r manyOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
+# distributions of variables with many outliers
 set2 <-  train_set %>%
@@
 ```{r otherOutliers, fig.height = 3.5, fig.width = 7, fig.align = 'center'}
+# distribution of all other variables 
 set3 <-  train_set %>%
@@
 ```{r outliers_1, fig.height = 4, fig.width = 8, fig.align = 'center'}
+
+##################### DATA PREPARATION
+
+# visualizing variables pre outlier removal
 train_set %>%
@@
 ```{r iqr_removal_SO}
+# remove outliers
 train_set<- train_set %>%
@@
 ```{r outlier_boxplot, fig.height = 4, fig.width = 8, fig.align = 'center'}
+# visualizing variables post outlier removal
 train_set %>%
@@
 ```{r impute_train}
+# imputing values
 train_set <- complete(mice(data = train_set, 
@@
 ```{r singles}
+# defining singles variable 
 train_set<- train_set %>%
@@
 ```{r make_slg_obp}
+# defining at bats and slugging variables
 train_set<- train_set %>%
@@
 ```{r slugging_compare, fig.height = 2.5, fig.width = 8, fig.align = 'center'}
+# visualizing slugging compared to wins
 modified_set <- train_set %>%
@@
 
-# MODEL TRAINING
+# BUILD MODELS
 ## Model 1
@@
 ```{r baseModel}
+##################### BUILD MODELS
+##### Model 1: Backwards Selection
 lm_reg = lm(data=train_set, TARGET_WINS ~ .)
-```
+
+knitr::kable(summary(lm_reg)$coefficients, digits = 3L,
+             caption = 'Base Model Regression Output')
+```
+
+We can see off the bat that a few variables *exceed* the 0.05 p-value threshold for 
+significance: `TEAM_BATTING_H`, `TEAM_PITCHING_H` and `TEAM_PITCHING_BB`. Additionally, 
+we see that both `TEAM_BATTING_1B` and `AB` have no coefficients at all! This is likely 
+due to the fact that they are colinear with other variables in the model.
 
 ### Enhanced Model
-Model 1 includes all variables except Singles, Triples, Base Hits, Strikeouts by
-pitchers, Walks allowed, and Batters Hit by Pitch. The final variables were
-chosen as a result of backwards `dplyr::selection` based on null hypothesis
-testing for non-zero slope.
+
+We used a combination of domain knowledge and data analysis to justify removing/retaining 
+features. First, we are keeping the coefficient for `TEAM_BATTING_HR` since its p-value
+is marginally above the general threshold and knowledge of the game suggests it is important. 
+We also chose to remove the `TEAM_BATTING_HBP` variable because most of the values are
+missing from the dataset.
+
+The remaining variables were chosen as a result of **backwards selection** based on null 
+hypothesis testing for non-zero slope. This is done by removing one variable at a
+time to determine the minimum number of significant variables needed for prediction.
+
+Altogether, we end up **removing** 7 total variables: `TEAM_BATTING_1B`, `AB`, 
+`TEAM_BATTING_3B`, `TEAM_BATTING_H`, `TEAM_PITCHING_BB`, `TEAM_PITCHING_SO`, and
+`TEAM_BATTING_1B`.
 
 ### Further data transformation
-We will use Cook's distance to remove outliers that are influencing the fit of
+We will use Cook's distance to remove additional outliers that are influencing the fit of
 the model above. We will use a cutoff of \(\frac{4}{N}\).
@@
 ```{r cooks_distance}
+# removing outliers via cooks distance
 lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
@@
 ```
+
 ```{r model1_coef}
@@
 ### Coefficient Discussion
-First, we are keeping the coefficient for `TEAM_BATTING_HR` since it's p-value
-is marginally above the general threshold and knowledge of the game suggests it
-is important. There are some counter intuitive results, which is expected given
-that baseball is a messy, imprecise game that has evolved over time. To that
-end, we should expect some seemingly strange results from algorithmic
-regression. We used a combination of domain knowledge and data analysis to
-justify retaining features.
+There are some counter intuitive results, which is expected given that baseball is a 
+messy, imprecise game that has evolved over time. To that end, we should expect some 
+seemingly strange results from algorithmic regression.
 
@@
 ```{r model2C+S}
+##### Model 2: Pairwise Model
+# cleaning data for model 2
 ts2 <- train_set[, -(17:19)]
@@
 ```{r model3}
+
+##### Model 3: K-fold Model
 stepdata <- subset(train_set, select = -c(AB, SLG, TEAM_BATTING_HBP))
@@
 
-# MODEL SELECTION
+# SELECT MODELS
 ## Criteria
@@
 ```{r performance}
+##################### SELECT MODELS
 # Model accuracy
@@
 
+```{r errors}
+
+# calculate predictions for each model 
+mod1_preds <- data.frame(cbind(train_df2$TARGET_WINS, round(mod_sum1$residuals,0), train_df2$TARGET_WINS + round(mod_sum1$residuals,0) ))
+colnames(mod1_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')
+
+mod2_preds <- data.frame(cbind(ts2$TARGET_WINS, round(sstm$residuals,0), ts2$TARGET_WINS + round(sstm$residuals,0) ))
+colnames(mod2_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')
+
+mod3_preds <- data.frame(cbind(stepdata.2$TARGET_WINS, round(mod_sum3$residuals,0), stepdata.2$TARGET_WINS + round(mod_sum3$residuals,0) ))
+colnames(mod3_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')
+
+# summary statistics for residuals
+resid_stats <- data.frame(rbind(summary(mod1_preds$RESIDUAL), summary(mod2_preds$RESIDUAL), summary(mod3_preds$RESIDUAL)))
+
+resid_stats$Models <- c('Model #1: Backwards' , 'Model #2: Pairwise', 'Model #3: K-fold')
+
+```
+
 It was interesting to compare the three models given that their performance
-statistics and error measurements were not significantly different from each other.
-**Model #2: Pairwise** was the worst performing model for all of our metrics, and
-was eliminated immediately. **Model #3: K-fold** accounts for nearly 45% of the
-variation in the dependent variable with the independent variables because its
-\(R^2\) is 0.456 and its adjusted \(R^2\) is = 0.450, both of which are
-acceptable for a good model. However, **Model #1: Backwards** has the smallest
-RSME of all the models. RMSE is a measure of how far off predictions are from
-the actual, giving extra weight to outliers through summing the squared
+statistics, error measurements, and residual distributions were not drastically 
+different from each other. **Model #2: Pairwise** was the worst performing model 
+for all of our metrics, and was eliminated immediately. **Model #3: K-fold** accounts 
+for nearly 45% of the variation in the dependent variable with the independent 
+variables because its \(R^2\) is 0.456 and its adjusted \(R^2\) is = 0.450, both of 
+which are acceptable for a good model. However, **Model #1: Backwards** has the 
+smallest RSME of all the models. RMSE is a measure of how far off predictions are 
+from the actual, giving extra weight to outliers through summing the squared
 differences. This demonstrates a greater reduction in the randomness by
@@
 
-# PREDICTIONS
+## Predictions
 **Model #1: Backwards** was used to predict the target wins base on the
@@
 ```{r eval_set, eval=FALSE}
+# predict target wins
 eval_set = eval_set %>% na.omit() %>%
@@
 ```{r predictions}
-eval_set <- paste0(urlRemote, pathGithub, fileEval) %>% read.csv()
+
+############# predict target wins
+
+# outlier removal - TEAM_PITCHING_SO
 eval_set <- eval_set %>% na.omit() %>%
@@
 
+# outlier removal - TEAM_PITCHING_H
 eval_set <- eval_set %>% na.omit() %>%
@@
 
+# mice imputation
 eval_set <- complete(mice(data = eval_set, method = "pmm", seed = 9450, print=FALSE), 3)
 
+# feature engineering
 eval_set <- eval_set %>% dplyr::select(-INDEX) %>%
@@
                   4 * TEAM_BATTING_HR) / AB)
+
+# final predictions
 predictions <- round(predict(lm_1, newdata = eval_set))

### 2020-09-18 HW1/HW1_draft_kellogg.Rmd
ambers final edits

@@
 We can see off the bat that a few variables *exceed* the 0.05 p-value threshold for 
-significance: `TEAM_BATTING_H`, `TEAM_PITCHING_H` and `TEAM_PITCHING_BB`. Additionally, 
-we see that both `TEAM_BATTING_1B` and `AB` have no coefficients at all! This is likely 
-due to the fact that they are colinear with other variables in the model.
+significance. Additionally, we see that both `TEAM_BATTING_1B` and `AB` have no 
+coefficients at all! This is likely due to the fact that they are colinear with other 
+variables in the model.
 
@@
 
-Fielding double plays and batting doubles both appear to have negative impacts
+First, we see that the intercept itself is negative. This means that, given no information
+about any of the statistics, we would predict that a team would lose (by a lot!).
+
+Next, fielding double plays and batting doubles both appear to have negative impacts
 on wins even though they *should* be positive impacts. Turning double plays,

### 2020-09-19 HW1/HW1_draft_kellogg.Rmd
Updated Model 3 coef discussion

@@
 
-Studying the coefficients of the model suggest that winning is in favor if the
-team batting hits more doubles, triples and home runs. Moreover, increase in the
-number of stolen bases, and a decrease in caught steals, double plays, error,
-and walks allowed would all lead to a win for the batting team. It is noteworthy
-that the model suggests that a decrease in single hits by batter and an increase
-in strikeouts by batters which seems counter intuitive. But these variables were
-kept because when a batter steps to the plate, the player is more likely to
-strike out than to get a hit. Trying to hit the ball out of the park will come
-with strikeouts but it will also increase the chances of hitting home runs (even
-1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry
-out.
+From studying each coefficient, this discussion begins with the variables that follow 
+the logic of winning and are self-evident. Firstly, the model suggests that winning is 
+in favor of the team batting hits more bases, (1B, 2B, 3B, HR). It further reveals that 
+an increase in singles would lead to better chances of winning. A hit is a hit, and more 
+hits than the opponent will inevitably lead to a win. Likewise, an increase in walks by 
+the batting team is awarded first base, further increasing base counts and, thus the 
+chances of winning. In terms of the base run, an increase in the number of stolen bases, 
+and a decrease in caught steals would lead to a win for the batting team. These variables 
+have a positive effect on the number of wins, and it is understandable as to why.
+
+When interpreting the model based on the pitching team's plays, it is perceptible that 
+walks allowed should be reduced so that the batting team hits does not increase. Moreover, 
+preventing errors and increasing double plays in which two players are put out will also 
+increase the chances of winning.
+
+On the other hand, it is noteworthy that the model suggests some counter-intuitive play 
+strategies. Firstly, when all variables are null, the intercept predicts that a team would 
+win, with a target win of 129. Moreover, the model suggests that a decrease in double hits 
+and an increase in strikeouts by the batting team will improve winning. These variables 
+were kept because when a batter steps to the plate, the player is more likely to strike 
+out than to get a hit. Trying to hit the ball out of the park will come with strikeouts 
+but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is 
+a pretty good exchange that most teams are willing to perform.
 

### 2020-09-21 HW1/HW1_Group2.Rmd
Copyediting

--- 2020-09-19 HW1/HW1_draft_kellogg.Rmd
@@
 ```{r loadData, include=FALSE}
-##################### libraries and variables 
+# Libraries and variables 
 library(tidyverse)
@@
 library(caret)
-library(data.table)
 
@@
 fileTrain   = "moneyball-training-data.csv"
-train_set = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
+train_set = read.csv(paste0(urlRemote, pathGithub, fileTrain))
 
@@
 ## How Often Does The Team Win?
-We are given a data set of 2,276 records containing 15 seasonal statistics and
-the total number of wins a team had in a given year. On average, about 50% of
-games played are won (81 games out of 162), with the best individual season
-having 146 wins and the worst season having 0 wins. The data is normally
-distributed and most years have between 49 and 112 wins (blue lines below). The
-nature of the distribution means there aren't too many extreme seasons where
-wins are significantly higher or lower than usual. This serves as a good
-gut-check for our final predictions; if the predicted wins are too high or too 
-low, we know something in our model is probably off.
+The data set provided comprises 2,276 records of baseball teams seasons. There
+are 15 features, statistics corresponding to a team's performance in a season,
+and a target variable of the team's number of wins in that season. A baseball
+season contains 162 games. On average, about 50% of games played are won---81
+games out of 162---with the best individual season having 146 wins and the worst
+season having 0 wins. The data is near normally distributed and most years have
+between 49 and 112 wins, represented by the blue lines in the histogram below.
+The nature of the distribution indicates that there aren't too many extreme
+seasons where wins are significantly higher or lower than usual. This serves as
+a good gut-check for our final predictions; if the predicted wins are too high
+or too low, we know something in our model is probably off.
 
 ```{r wins, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
-##################### DATA EXPLORATION
-
+# DATA EXPLORATION
 avgWins <- mean(train_set$TARGET_WINS)
@@
 and strikeouts by pitchers and batters (`TEAM_BASERUN_SB`, `TEAM_PITCHING_SO`,
-`TEAM_BATTING_SO`). **It seems completely unreasonable** to have zero strike
+`TEAM_BATTING_SO`). It seems **completely unreasonable** to have zero strike
 outs in a season, so this is something we'll most certainly have to impute.
@@
 ```{r pctNull, fig.height = 2, fig.width = 5, fig.align = 'center'}
-
 # percent null for each variable
@@
 ```{r negFactors, fig.align = 'center'}
-# variables expected to have a negative impact on wins
+# Variables expected to have a negative impact on wins
 negSet <- train_set %>%
@@
 ```{r posFactors, fig.align = 'center'}
-# variables expected to have a positive impact on wins
+# Variables expected to have a positive impact on wins
 posSet <- train_set %>%
@@
 
-Many of these stats *do* seem to have an effect on the number of wins, most
+Many of these stats *do* seem to have an effect on the number of wins; most
 notably, **Base Hits** and **Walks**. We see weaker positive relationships for
-**Home Runs**, **Doubles**, **Triples**, and **Stolen Bases**. This makes sense
-when we think about it; these things tend to happen less often in games than
-pure base hits and walks, so they don't have as much of an effect on winning.
-Finally, **Double Plays** and **Batters Hit** don't appear to have any
-correlation with the number of wins. Once again, this intuitively makes sense
-because they are less likely to happen in a game.
+**Home Runs**, **Doubles**, **Triples**, and **Stolen Bases**. On the surface,
+this makes sense as these events tend to happen less often in games than base
+hits and walks, so they don't have as much of an effect on winning. Finally,
+**Double Plays** and **Batters Hit** don't appear to have any correlation with
+the number of wins. Once again, this intuitively makes sense because they are
+less likely to happen in a game.
 
 One thing to note is the number of strikeouts compared to the number of wins. We
-can see that there are a few outliers (abnormally high numbers of strikeouts in
-a season). This should be taken with caution, as they don't represent a typical
+can see that there are a few outliers---abnormally high numbers of strikeouts in
+a season. This should be taken with caution, as they don't represent a typical
 season's stats.
@@
 Before using any of the statistics in a model, we need to take a closer look at
-the variation in the data. We call out-of-the-ordinary values (exceptionally
-high or low values) **outliers**. We need to take these into account in our
-modeling because we want to make sure our predictions aren't skewed because of
-them.
+the variation in the data. We call out-of-the-ordinary values, exceptionally
+high or low values, **outliers**. We need to take these into account in our
+modeling to prevent them from skewing our predictions.
 
 ```{r summaryInfo, fig.height = 3, fig.width = 5, fig.align = 'center'}
-# understanding outliers by variable
+# Understanding outliers by variable
 totalOutliers <- data.frame(
@@
 As we can see, some of the provided statistics are well-balanced in the sense
-that there are very *few* (or no) extreme values. **Home Runs allowed**
+that there are very *few*, or no, extreme values. **Home Runs allowed**
 (`TEAM_PITCHING_HR`), **Strikeouts by batters** (`TEAM_BATTING_SO`), and
@@
 ```{r fewOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
-# distributions of variables with few outliers
+# Distributions of variables with few outliers
 set1 <- train_set %>%
@@
   * **Home Runs Allowed** (average ~100/year) and **Home Runs by Batters**
-  (average ~106/year) have a very similar mid-range distribution (50% of the
-  data lies between ~50 and 150). The slight difference in average stats means
+  (average ~106/year) have a very similar mid-range distribution---50% of the
+  data lies between ~50 and 150. The slight difference in average stats means
   that teams tend to have a higher number of Home Runs than the opposition team.
@@
 ```{r manyOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
-# distributions of variables with many outliers
+# Distributions of variables with many outliers
 set2 <-  train_set %>%
@@
 ```{r otherOutliers, fig.height = 3.5, fig.width = 7, fig.align = 'center'}
-# distribution of all other variables 
+# Distribution of all other variables 
 set3 <-  train_set %>%
@@
 ## Are Stats Correlated?
-We would expect that a few things in the dataset might be correlated: perhaps
+We would expect that a few things in the dataset might be correlated. Perhaps
 number of errors and hits/homeruns allowed or the number of base hits by batters
 and doubles/triples/homeruns. We can visualize the correlations between the
-statistics to determine if there is a significant relationship between the them:
-blue dots represent positively correlated variables (as one increases, so does
-the other) and red dots represent negatively correlated variables (as one
-increases, the other decreases).
+statistics to determine if there is a significant relationship between the them.
+In the graph below, blue dots represent positively correlated variables and red
+dots represent negatively correlated variables.
 
@@
 ```{r outliers_1, fig.height = 4, fig.width = 8, fig.align = 'center'}
-
-##################### DATA PREPARATION
-
-# visualizing variables pre outlier removal
+# Visualizing variables prior to outlier removal
 train_set %>%
@@
   ~123 strikeouts per game. This is only possible if every one of the 162 games
-  went approximately 40 innings, and each out was a strikeout---which clearly has
-  never happened.
+  went approximately 40 innings, and each out was a strikeout---which clearly
+  has never happened.
   * **TEAM_PITCHING_H**: The largest outlier for this variable is over 30,000
@@
 ```{r iqr_removal_SO}
-# remove outliers
+# Remove outliers
 train_set<- train_set %>%
@@
 still large outlier sets for Errors and Stolen Bases, but none that seem to
-dwarf the other variables. According to the box plot, things appear to be on a 
-scale that seems logical given the data and source of information that we have.
+dwarf the other variables. According to the box plot, observations appear to be
+on a logical scale given the data.
 
 ## Imputation
-From our early exploration of the data, the vast majority of the data is
+Our early exploration of the data showed that the majority of the data is
 complete with only a few variables with missing values. We will use the **MICE**
 (Multivariate Imputation by Chained Equations) method. MICE is a principled
-method for dealing with missing data. It creates multiple imputations, as
-opposed to single imputations, and accounts for the statistical uncertainty in
-these imputations. It creates predictive values for the mean instead of imputing
-the IQR values.
+method for dealing with missing data. It creates multiple imputations for the
+mean, as opposed to single imputations, and accounts for the statistical
+uncertainty in these imputations.
 
 ```{r impute_train}
-# imputing values
+# Imputing values
 train_set <- complete(mice(data = train_set, 
@@
 ### Single base hits
-With outliers removed and missing values imputed via MICE, we can create a few
-new variables. The first will be single base hits, and it will be derived from
-some of the variables we do have. We know that Team Batting Hits
-(`TEAM_BATTING_H`) is a combination of *all* hits for the season, so we can
-create Single Base Hits as follows: 
+With outliers removed and missing values imputed via MICE, we can create new
+variables. The first will be single base hits, and it will be derived from some
+of the variables we do have. We know that Team Batting Hits (`TEAM_BATTING_H`)
+is a combination of *all* hits for the season, so we can create Single Base Hits
+as follows:
+
 \[
@@
 ```{r singles}
-# defining singles variable 
+# Defining singles variable 
 train_set<- train_set %>%
@@
 \]
-Because we don't have a statistic for at-bats, we will approximate it as follows:
+
+Because we don't have a statistic for at-bats, we will approximate it as:
 
@@
 ```{r make_slg_obp}
-# defining at bats and slugging variables
+# Defining at bats and slugging variables
 train_set<- train_set %>%
@@
 
-As a gut-check, we can take a look at Slugging in comparison to Triples
+As a sanity check, we can take a look at Slugging in comparison to Triples
 (`TEAM_BATTING_3B`) and Home Runs (`TEAM_BATTING_HR`). All of these variables
@@
 ```{r slugging_compare, fig.height = 2.5, fig.width = 8, fig.align = 'center'}
-# visualizing slugging compared to wins
+# Visualizing slugging compared to wins
 modified_set <- train_set %>%
@@
 ```{r baseModel}
-##################### BUILD MODELS
-##### Model 1: Backwards Selection
+# BUILD MODELS
+# Model 1: Backwards Selection
 lm_reg = lm(data=train_set, TARGET_WINS ~ .)
@@
 
-We can see off the bat that a few variables *exceed* the 0.05 p-value threshold for 
-significance. Additionally, we see that both `TEAM_BATTING_1B` and `AB` have no 
-coefficients at all! This is likely due to the fact that they are colinear with other 
-variables in the model.
+We can immediately see that a few variables *exceed* the 0.05 p-value threshold
+for significance. Additionally, we see that both `TEAM_BATTING_1B` and `AB` have
+no coefficients at all! This is likely due to the fact that they are colinear
+with other variables in the model.
 
 ### Enhanced Model
-
-We used a combination of domain knowledge and data analysis to justify removing/retaining 
-features. First, we are keeping the coefficient for `TEAM_BATTING_HR` since its p-value
-is marginally above the general threshold and knowledge of the game suggests it is important. 
-We also chose to remove the `TEAM_BATTING_HBP` variable because most of the values are
-missing from the dataset.
-
-The remaining variables were chosen as a result of **backwards selection** based on null 
-hypothesis testing for non-zero slope. This is done by removing one variable at a
-time to determine the minimum number of significant variables needed for prediction.
+We used a combination of domain knowledge and data analysis to justify removing
+some features. First, we are keeping the coefficient for `TEAM_BATTING_HR` since
+its p-value is marginally above the general threshold and knowledge of the game
+suggests it is important. We also chose to remove the `TEAM_BATTING_HBP`
+variable because most of the values are missing from the dataset.
+
+The remaining variables were chosen as a result of **backwards selection** based
+on null hypothesis testing for non-zero slope. This is done by removing one
+variable at a time to determine the minimum number of significant variables
+needed for prediction.
 
@@
 ### Further data transformation
-We will use Cook's distance to remove additional outliers that are influencing the fit of
-the model above. We will use a cutoff of \(\frac{4}{N}\).
+We used Cook's distance to remove additional outliers that are influencing the
+fit of the model above. We used a cutoff of \(\frac{4}{N}\).
 

### 2020-09-21 HW1/HW1_Group2.Rmd
Copyedit and switch prose. We are predicting target WINS not the score of a game, so the team will not lose BY a lot. Rather, they will lose ALL there games (and then some).

@@
 ```{r cooks_distance}
-# removing outliers via cooks distance
+# Removing outliers via cooks distance
 lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
@@
 ### Coefficient Discussion
-There are some counter intuitive results, which is expected given that baseball is a 
-messy, imprecise game that has evolved over time. To that end, we should expect some 
-seemingly strange results from algorithmic regression.
-
-First, we see that the intercept itself is negative. This means that, given no information
-about any of the statistics, we would predict that a team would lose (by a lot!).
-
-Next, fielding double plays and batting doubles both appear to have negative impacts
-on wins even though they *should* be positive impacts. Turning double plays,
+There are some counter intuitive results, which is expected given that baseball
+is a messy, imprecise game that has evolved over time. To that end, we should
+expect some seemingly strange results from algorithmic regression.
+
+First, we see that the intercept itself is negative. This means that, given no
+information about any of the statistics, we would predict that a team would lose
+all of its games, and then some!
+
+Next, fielding double plays and batting doubles both appear to have negative
+impacts on wins even though they *should* be positive impacts. Turning double plays,
 while a good for the defensive team, may suggest a larger, negative issue.

### 2020-09-21 HW1/HW1_Group2.Rmd
More copyediting

@@
 Next, fielding double plays and batting doubles both appear to have negative
-impacts on wins even though they *should* be positive impacts. Turning double plays,
-while a good for the defensive team, may suggest a larger, negative issue.
-Namely, weak pitching that leads to runners on base. Similarly, batting doubles,
-leaves runners open to double plays. Allowed hits by pitching, caught stealing,
-and batting strike-outs are all counter intuitive results as well, but they seem
-to be small contributors and these are events that happen regularly in every
-game.
+impacts on wins even though they *should* be positive impacts. Turning double
+plays, while a good for the defensive team, may suggest a larger, negative
+issue. Namely, weak pitching that leads to runners on base. Similarly, batting
+doubles, leaves runners open to double plays. Allowed hits by pitching, caught
+stealing, and batting strike-outs are all counter intuitive results as well, but
+they seem to be small contributors and these are events that happen regularly in
+every game.
 
@@
 ```{r model2C+S}
-##### Model 2: Pairwise Model
-# cleaning data for model 2
+# Model 2: Pairwise Model
+# Cleaning data for model 2
 ts2 <- train_set[, -(17:19)]
@@
 ### Further data transformation
-Once again, Cook's distance is use to decide if an individual entity is an extreme 
-value or not.
+Once again, Cook's distance is use to decide if an individual entity is an
+extreme value or not.
 
 ```{r model3}
-
-##### Model 3: K-fold Model
+# Model 3: K-fold Model
 stepdata <- subset(train_set, select = -c(AB, SLG, TEAM_BATTING_HBP))
-stepdata <- cbind(stepdata, sapply(stepdata[2:length(stepdata)], function(x) x^2))
+stepdata <- cbind(stepdata,
+                  sapply(stepdata[2:length(stepdata)], function(x) x^2))
 names(stepdata) <- make.unique(names(stepdata), sep = "_")
@@
 cooksd <- cooks.distance(model)
-influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
+influential <- as.numeric(
+  names(cooksd)[(cooksd > 4 * mean(cooksd, na.rm = TRUE))])
 stepdata.2 <- stepdata[-c(influential), ]
+
 # Set up repeated k-fold cross-validation
@@
 # Final model 
-step.model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BASERUN_SB + 
-                  TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_FIELDING_E + 
-                  TEAM_FIELDING_DP + TEAM_BATTING_BB_1 + TEAM_BATTING_SO_1 + 
-                  TEAM_BASERUN_CS_1 + TEAM_PITCHING_H_1 + TEAM_PITCHING_BB_1 + 
-                  TEAM_FIELDING_E_1 + TEAM_FIELDING_DP_1 + TEAM_BATTING_1B_1, 
-                data = stepdata.2)
+step.model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B +
+                   TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H +
+                   TEAM_PITCHING_HR + TEAM_FIELDING_E +  TEAM_FIELDING_DP +
+                   TEAM_BATTING_BB_1 + TEAM_BATTING_SO_1 +  TEAM_BASERUN_CS_1 +
+                   TEAM_PITCHING_H_1 + TEAM_PITCHING_BB_1 +  TEAM_FIELDING_E_1 +
+                   TEAM_FIELDING_DP_1 + TEAM_BATTING_1B_1, data = stepdata.2)
+
 # Summary of the model
@@
 
-From studying each coefficient, this discussion begins with the variables that follow 
-the logic of winning and are self-evident. Firstly, the model suggests that winning is 
-in favor of the team batting hits more bases, (1B, 2B, 3B, HR). It further reveals that 
-an increase in singles would lead to better chances of winning. A hit is a hit, and more 
-hits than the opponent will inevitably lead to a win. Likewise, an increase in walks by 
-the batting team is awarded first base, further increasing base counts and, thus the 
-chances of winning. In terms of the base run, an increase in the number of stolen bases, 
-and a decrease in caught steals would lead to a win for the batting team. These variables 
-have a positive effect on the number of wins, and it is understandable as to why.
-
-When interpreting the model based on the pitching team's plays, it is perceptible that 
-walks allowed should be reduced so that the batting team hits does not increase. Moreover, 
-preventing errors and increasing double plays in which two players are put out will also 
-increase the chances of winning.
-
-On the other hand, it is noteworthy that the model suggests some counter-intuitive play 
-strategies. Firstly, when all variables are null, the intercept predicts that a team would 
-win, with a target win of 129. Moreover, the model suggests that a decrease in double hits 
-and an increase in strikeouts by the batting team will improve winning. These variables 
-were kept because when a batter steps to the plate, the player is more likely to strike 
-out than to get a hit. Trying to hit the ball out of the park will come with strikeouts 
-but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is 
-a pretty good exchange that most teams are willing to perform.
+This discussion begins with the variables that follow the logic of winning and
+are self-evident. Firstly, the model suggests that winning is in favor of the
+team which when batting has more total bases, (1B, 2B, 3B, HR). It further
+reveals that an increase in singles would lead to better chances of winning. A
+hit is a hit, and more hits than the opponent will inevitably lead to a win.
+Likewise, an increase in walks by the batting team is awarded first base,
+further increasing base counts and, thus the chances of winning. In terms of the
+base run, an increase in the number of stolen bases, and a decrease in caught
+steals would lead to a win for the batting team. These variables have a positive
+effect on the number of wins, and it is understandable as to why.
+
+When interpreting the model based on the pitching team's plays, it is
+perceptible that walks allowed should be reduced so that the batting team hits
+does not increase. Moreover, preventing errors and increasing double plays in
+which two players are put out will also increase the chances of winning.
+
+On the other hand, it is noteworthy that the model suggests some
+counter-intuitive play strategies. Firstly, when all variables are null, the
+intercept predicts that a team would win, with a target win of 129. Moreover,
+the model suggests that a decrease in double hits and an increase in strikeouts
+by the batting team will improve winning. These variables were kept because when
+a batter steps to the plate, the player is more likely to strike out than to get
+a hit. Trying to hit the ball out of the park will come with strikeouts but it
+will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that
+is a pretty good exchange that most teams are willing to perform.
 

### 2020-09-21 HW1/HW1_Group2.Rmd
Unused

@@
 ```{r performance}
-##################### SELECT MODELS
+# SELECT MODELS
 # Model accuracy
@@
                          caption = 'Performance Statistics & Error Measure of Models')
-```
-
-```{r errors}
-
-# calculate predictions for each model 
-mod1_preds <- data.frame(cbind(train_df2$TARGET_WINS, round(mod_sum1$residuals,0), train_df2$TARGET_WINS + round(mod_sum1$residuals,0) ))
-colnames(mod1_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')
-
-mod2_preds <- data.frame(cbind(ts2$TARGET_WINS, round(sstm$residuals,0), ts2$TARGET_WINS + round(sstm$residuals,0) ))
-colnames(mod2_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')
-
-mod3_preds <- data.frame(cbind(stepdata.2$TARGET_WINS, round(mod_sum3$residuals,0), stepdata.2$TARGET_WINS + round(mod_sum3$residuals,0) ))
-colnames(mod3_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')
-
-# summary statistics for residuals
-resid_stats <- data.frame(rbind(summary(mod1_preds$RESIDUAL), summary(mod2_preds$RESIDUAL), summary(mod3_preds$RESIDUAL)))
-
-resid_stats$Models <- c('Model #1: Backwards' , 'Model #2: Pairwise', 'Model #3: K-fold')
-
 ```

### 2020-09-21 HW1/HW1_Group2.Rmd
Copyedit

@@
 different from each other. **Model #2: Pairwise** was the worst performing model 
-for all of our metrics, and was eliminated immediately. **Model #3: K-fold** accounts 
-for nearly 45% of the variation in the dependent variable with the independent 
-variables because its \(R^2\) is 0.456 and its adjusted \(R^2\) is = 0.450, both of 
-which are acceptable for a good model. However, **Model #1: Backwards** has the 
-smallest RSME of all the models. RMSE is a measure of how far off predictions are 
-from the actual, giving extra weight to outliers through summing the squared
-differences. This demonstrates a greater reduction in the randomness by
-**Model #1: Backwards**. But its difference from **Model #3: K-fold** is only
-0.022. It is hard to judge if that is a significant difference. We thus also
-considered AIC. With AIC, it is not the magnitude of the value which is
-important, but the difference between the various model AIC's. The magnitude is
-a function of the number of observations. Accepted rules of thumb in the
-statistical literature are that if the difference is greater than 10, there is
-effectively no support for the model with the larger AIC
-(Section 2.6, Burnham & Anderson 2002). As a result, **Model #1: Backwards**
-is our selected model.
+for all of our metrics, and was eliminated immediately. **Model #3: K-fold**
+accounts for nearly 45% of the variation in the dependent variable with the
+independent variables because its \(R^2\) is 0.456 and its adjusted \(R^2\) is
+0.450, both of which are acceptable for a good model. However,
+**Model #1: Backwards** has the smallest RSME of all the models. RMSE is a
+measure of how far off predictions are from the actual, giving extra weight to
+outliers through summing the squared differences. This demonstrates a greater
+reduction in the randomness by **Model #1: Backwards**. But its difference from
+**Model #3: K-fold** is only 0.022. It is hard to judge if that is a significant
+difference. We thus also considered AIC. With AIC, it is not the magnitude of
+the value which is important, but the difference between the various model
+AIC's. The magnitude is a function of the number of observations. Accepted rules
+of thumb in the statistical literature are that if the difference is greater
+than 10, there is effectively no support for the model with the larger AIC
+(Section 2.6, Burnham & Anderson 2002). As a result, **Model #1: Backwards** is
+our selected model.
 
@@
 expected to win negative games, which is clearly impossible. That being said, it
-is an indication of the predictive power of the engineereed **slugging** feature
+is an indication of the predictive power of the engineered **slugging** feature
 that allows small increments in slugging to result in significant shifts in
@@
 ```{r eval_set, eval=FALSE}
-# predict target wins
+# Predict target wins
 eval_set = eval_set %>% na.omit() %>%
@@
 ```{r predictions}
-
-############# predict target wins
-
+# Predict target wins
 # outlier removal - TEAM_PITCHING_SO

### 2020-09-21 HW1/HW1_Group2.Rmd
Unused

@@
 
-```{r eval_set, eval=FALSE}
-# Predict target wins
-eval_set = eval_set %>% na.omit() %>%
-  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
-  bind_cols(eval_set) %>%
-  filter(TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO, probs = c(0.25), na.rm = TRUE) 
-         - 1.5 * iqr, 
-         TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, probs = c(0.75), na.rm = TRUE) 
-         + 1.5 * iqr) %>%
-  dplyr::select(-iqr)
-
-eval_set = eval_set %>% na.omit() %>%
-  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
-  bind_cols(eval_set) %>%
-  filter(TEAM_PITCHING_H > quantile(TEAM_PITCHING_H, probs = c(0.25), na.rm = TRUE) 
-         - 1.5 * iqr,
-         TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, probs = c(0.75), na.rm = TRUE) 
-         + 1.5 * iqr) %>%
-  dplyr::select(-iqr)
-
-eval_set = complete(mice(data = eval_set, method = "pmm", seed = 9450,  print=FALSE), 3)
-
-eval_set = eval_set %>% dplyr::select(-INDEX) %>%
-  mutate_if(is.integer, as.numeric) %>%
-  mutate(TEAM_BATTING_1B = TEAM_BATTING_H - (TEAM_BATTING_2B + TEAM_BATTING_3B +
-                                               TEAM_BATTING_HR))
-
-# Cleaning Test Set for Model #1
-test_df1 = eval_set %>%
-  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
-         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
-                  4 * TEAM_BATTING_HR) / AB)
-# Cleaning Test Set for Model #2
-test_df2 = test_df1[, -(17:19)]
-test_df2 = data.frame(apply(test_df2, 2, scale))
-# Cleaning Test Set for Model #3
-test_df3 = cbind(test_df1, sapply(test_df1[1:length(test_df1)], function(x) x^2))
-names(test_df3) = make.unique(names(test_df3), sep = "_")
-```
-
 ```{r predictions}
@@
 ```
-<!--```{r eval_set}
-```-->
 ```{r predictions}

### 2020-09-21 Adler_personal/Blog3.Rmd
Blog Post 3

---
title: "Blog Post 3: Training & Testing: Other Models"
subtitle: "DT 621---Fall 2020"
author: "Avraham Adler"
date: "9/21/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, error=FALSE)
```

# Training and Testing with Hyperparameters
We saw in the [previous blog post](https://rpubs.com/Avraham/DT621Blog2) that
splitting data into training and testing sets in and of itself can be valuable
to prevent overfitting. In this post, we will demonstrate how it can help with
hyperparameter tuning and model selection.

We will train two sets of models. The first will train on the entire available
data set and the best performing model will be selected. The second set will
train on the "train" portion of the dataset, and the model which performs best
on the "test" portion of the dataset will be selected. The two selected models
will go head-to-head on the unobserved "future" data. The decision metric will
be RMSE.

This is not the best example as it is so simple, but it should suffice.

```{r loadPackages}
library(caret)
trc <- trainControl(method = 'cv', number = 10)
```
```{r createData}
set.seed(245)
x1 <- runif(400, 0, 100)
x2 <- rnorm(400, 0, 10)
x3 <- rexp(400, 1/5)
y <- 3.2 + 1.6 * x1 + 3.4 * x2 - 0.8 * x3 + rnorm(400, 0, 10)
futIDX <- sample(400, 100)
x1f <- x1[futIDX]
x1p <- x1[-futIDX]
x2f <- x2[futIDX]
x2p <- x2[-futIDX]
x3f <- x3[futIDX]
x3p <- x3[-futIDX]
yf <- y[futIDX]
yp <- y[-futIDX]
testIDX <- sample(300, 300 * 0.25)
x1trn <- x1p[-testIDX]
x1tst <- x1p[testIDX]
x2trn <- x2p[-testIDX]
x2tst <- x2p[testIDX]
x3trn <- x3p[-testIDX]
x3tst <- x3p[testIDX]
ytrn <- yp[-testIDX]
ytst <- yp[testIDX]
```

# Model Training
For a serious project, one-pass tuning would be insufficient. The results of
each run would be used to fine-tune the parameters.

## Random Forest
```{r rfFit}
set.seed(245)
rf1 <- train(data.frame(x1 = x1p, x2 = x2p, x3 = x3p), yp, method = 'ranger',
            trControl = trc, tuneLength = 2L)
rf2 <- train(data.frame(x1 = x1trn, x2 = x2trn, x3 = x3trn), ytrn,
            method = 'ranger', tuneLength = 2L)
plot(rf1)
plot(rf2)
```

## Support Vector Machines
```{r svmFit}
tgrid <- expand.grid(list(C = seq(0.01, 1, 0.01)))
set.seed(245)
sv1 <- train(data.frame(x1 = x1p, x2 = x2p, x3 = x3p), yp, method = 'svmLinear',
            trControl = trc, tuneGrid = tgrid)
sv2 <- train(data.frame(x1 = x1trn, x2 = x2trn, x3 = x3trn), ytrn,
             method = 'svmLinear', trControl = trc, tuneGrid = tgrid)
plot(sv1)
plot(sv2)
```

# Model Selection
## Models Trained on All the Data
```{r MS1}
min(rf1$results$RMSE)
min(sv1$results$RMSE)
```

## Using Testing Set
```{r MS2}
rf2p <- predict(rf2, data.frame(x1 = x1tst, x2 = x2tst, x3 = x3tst))
sv2p <- predict(sv2, data.frame(x1 = x1tst, x2 = x2tst, x3 = x3tst))
RMSE(rf2p, ytst)
RMSE(sv2p, ytst)
```

## Model Testing
In both cases the support vector machine outperformed the random forest. This is
almost assuredly due to the truth being a linear model. In a more complicated
case, one may see something different.

```{r Test}
sv1p <- predict(sv1, data.frame(x1 = x1f, x2 = x2f, x3 = x3f))
sv2p <- predict(sv2, data.frame(x1 = x1f, x2 = x2f, x3 = x3f))
defaultSummary(data.frame(pred = sv1p, obs = yf))
defaultSummary(data.frame(pred = sv2p, obs = yf))
```

The second model, fit on less data, has better accuracy on the "future" data.
This is also a demonstration of robsutness against overfitting.

Once again, being that there was a simple linear model underlying truth, the SVM
dominated. A benfit of using train/testing sets allows models of very different
structure to be compared across more difficult problems. Splitting the data into
two sets allows for the holdout set to act as the unknown future.

### 2020-09-22 HW2/HW2-Group2.rmd
Create HW2-Group2.rmd

---
title: 'DATA 621 - Business Analytics and Data Mining'
subtitle: 'Fall 2020 - Group 2 - Homework #2'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "10/11/2020"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)

# urlRemote  = "https://raw.githubusercontent.com/"
# pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
# fileTrain   = "classification-output-data.csv"
# data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()

data = read.csv("classification-output-data.csv")
```

## The Dataset

The data set contains n = 181 data points, and has three key columns that will be used:

* class: the actual class for the observation  
* scored.class: the predicted class for the observation (based on a threshold of 0.5)  
* scored.probability: the predicted probability of success for the observation

```{r data_table}
class_data = data %>% select(class, pred_class = scored.class, pred_prob = scored.probability) 
class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
table(class_data[,1:2]) %>% knitr::kable(caption = "Counts of the Classification Data")
```

## Confusion Matrix & Functions

A [confusion matrix](http://www.saedsayad.com/model_evaluation_c.htm) is a technique for summarizing the performance of a classification algorithm. It shows the number of correct and incorrect predictions made by the classification model compared to the actual outcomes (target value) in the data. The matrix is $N \times N$, where N is the number of target values (classes). The following table displays a 2x2 confusion matrix for two classes (Positive and Negative).

```{r}
cat("----------------------------------------------------------------------------
                 |           Target          |
Confusion Matrix |---------------------------|
                 |   Positive  |   Negative  |
-----------------|-------------|-------------|------------------------------
      | Positive |      a      |      b      | Positive Predictive | a/(a+b)
Model |----------|-------------|-------------|------------------------------
      | Negative |      c      |      d      | Negative Predictive | d/(c+d)
-----------------|-------------|-------------|------------------------------
                 | Sensitivity | Specificity |                    
                 |-------------|-------------|  Accuracy = (a+d)/(a+b+c+d)
                 |   a/(a+c)   |   d/(b+d)   | 
----------------------------------------------------------------------------")
```

* Accuracy: the proportion of the total number of predictions that were correct.  
* Positive Predictive Value or Precision: the proportion of positive cases that were correctly identified.  
* Negative Predictive Value: the proportion of negative cases that were correctly identified.  
* Sensitivity or Recall: the proportion of actual positive cases which are correctly identified.   
* Specificity: the proportion of actual negative cases which are correctly identified.

### Accuracy of the Predictions

Function `accuracy` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the accuracy of the predictions. The formula is:

\[Accuracy = \frac{TP + TN}{TP + FP + TN + FN}\]

```{r accuracy, echo = TRUE}
accuracy = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    TN = dt %>% filter(class == 0 & pred_class == 0) %>% select(Freq)
    TP = dt %>% filter(class == 1 & pred_class == 1) %>% select(Freq)
    return((TP + TN) / nrow(df))
  }
}
```

```{r accuracy_check}
sprintf("The accuracy of the predictions for the classification data set is %1.3f%%.", accuracy(class_data)*100)
```

### Error Rate of the Predictions

Function `error` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the classification error rate of the predictions. The formula is:

\[\text{Classification Error Rate} = \frac{FP + FN}{TP + FP + TN + FN}\]

```{r error, echo = TRUE}
error = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    FP = dt %>% filter(class == 0 & pred_class == 1) %>% select(Freq)
    FN = dt %>% filter(class == 1 & pred_class == 0) %>% select(Freq)
    return((FP + FN) / nrow(df))
  }
}
```

```{r error_check}
sprintf("The error rate of the predictions for the classification data set is %0.3f.", error(class_data))

sprintf("The accuracy and error rate of the predictions must sum to 1. For the classification data set, there summation equals to %.3f.", accuracy(class_data) + error(class_data))
```

### Precision of the Predictions

Function `precision` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the precision of the predictions. The formula is:

\[precision = \frac{TP}{TP + FP}\]

```{r precision, echo = TRUE}
precision = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    FP = dt %>% filter(class == 0 & pred_class == 1) %>% select(Freq)
    TP = dt %>% filter(class == 1 & pred_class == 1) %>% select(Freq)
    return(TP / (TP + FP))
  }
}
```

```{r precision_check}
sprintf("The precision of the predictions for the classification data set is %0.3f.", 
precision(class_data))
```

### Sensitivity of the Predictions

Function `sensitivity` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the sensitivity of the predictions. Sensitivity is also known as recall. The formula is:

\[sensitivity = \frac{TP}{TP + FN}\]

```{r sensitivity, echo = TRUE}
sensitivity = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    FN = dt %>% filter(class == 1 & pred_class == 0) %>% select(Freq)
    TP = dt %>% filter(class == 1 & pred_class == 1) %>% select(Freq)
    return(TP / (TP + FN))
  }
}
```

```{r sensitivity_check}
sprintf("The sensitivity of the predictions for the classification data set is %0.3f.", 
sensitivity(class_data))
```

### Specificity of the Predictions

Function `specificity` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the specificity of the predictions. The formula is:

\[specificity = \frac{TN}{TN + FP}\]

```{r specificity, echo = TRUE}
specificity = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    TN = dt %>% filter(class == 0 & pred_class == 0) %>% select(Freq)
    FP = dt %>% filter(class == 0 & pred_class == 1) %>% select(Freq)
    return(TN / (TN + FP))
  }
}
```

```{r specificity_check}
sprintf("The specificity of the predictions for the classification data set is %0.3f.", 
specificity(class_data))
```

### F1 Score of the Predictions

Function `f1_score` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the F1 score of the predictions. The formula is:

\[\text{F1 Score} = \frac{2 \times Precision \times Sensitivity}{Precision + Sensitivity}\]

```{r f1_score, echo = TRUE}
f1_score = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    return((2*precision(df)*sensitivity(df)) / (precision(df) + sensitivity(df)))
  }
}
```

```{r f1_score_check}
sprintf("The F1 score of the predictions for the classification data set is %0.3f.", 
f1_score(class_data))
```

### 2020-09-22 HW2/HW2-Group2.rmd
updated data path to git

@@
 
-# urlRemote  = "https://raw.githubusercontent.com/"
-# pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
-# fileTrain   = "classification-output-data.csv"
-# data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
-
-data = read.csv("classification-output-data.csv")
+urlRemote  = "https://raw.githubusercontent.com/"
+pathGithub = "aadler/DT621_Fall2020_Group2/master/HW2/"
+fileTrain   = "classification-output-data.csv"
+data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
 ```

### 2020-09-22 HW1/Drafts/HW1.Rmd

--- 2020-08-31 HW1/HW1.Rmd
@@
 ## 1. DATA EXPLORATION (Amber)
-Describe the size and the variables in the moneyball training data set.
-Consider that too much detail will cause a manager to lose interest while too little detail will make the manager consider that you aren’t doing your job.
-Some suggestions are given below.
-Please do NOT treat this as a check list of things to do to complete the assignment.
-You should have your own thoughts on what to tell the boss.
-These are just ideas.
+Describe the size and the variables in the moneyball training data set. Consider
+that too much detail will cause a manager to lose interest while too little
+detail will make the manager consider that you aren't doing your job. Some
+suggestions are given below. Please do NOT treat this as a check list of things
+to do to complete the assignment. You should have your own thoughts on what to
+tell the boss. These are just ideas.
 
@@
 ## 2. DATA PREPARATION (25 Points)
-Describe how you have transformed the data by changing the original variables or creating new variables.
-If you did transform the data or create new variables, discuss why you did this.
-Here are some possible transformations.
+Describe how you have transformed the data by changing the original variables or
+creating new variables. If you did transform the data or create new variables,
+discuss why you did this. Here are some possible transformations.
 

### 2020-09-22 HW1/Drafts/HW1_draft_kellogg.Rmd
Cleanup folder by moving to draft

--- 2020-09-19 HW1/HW1_draft_kellogg.Rmd
@@
+<<<<<<< HEAD:HW1/HW1_draft_kellogg.Rmd
 ---
@@
 
-From studying each coefficient, this discussion begins with the variables that follow 
-the logic of winning and are self-evident. Firstly, the model suggests that winning is 
-in favor of the team batting hits more bases, (1B, 2B, 3B, HR). It further reveals that 
-an increase in singles would lead to better chances of winning. A hit is a hit, and more 
-hits than the opponent will inevitably lead to a win. Likewise, an increase in walks by 
-the batting team is awarded first base, further increasing base counts and, thus the 
-chances of winning. In terms of the base run, an increase in the number of stolen bases, 
-and a decrease in caught steals would lead to a win for the batting team. These variables 
-have a positive effect on the number of wins, and it is understandable as to why.
-
-When interpreting the model based on the pitching team's plays, it is perceptible that 
-walks allowed should be reduced so that the batting team hits does not increase. Moreover, 
-preventing errors and increasing double plays in which two players are put out will also 
-increase the chances of winning.
-
-On the other hand, it is noteworthy that the model suggests some counter-intuitive play 
-strategies. Firstly, when all variables are null, the intercept predicts that a team would 
-win, with a target win of 129. Moreover, the model suggests that a decrease in double hits 
-and an increase in strikeouts by the batting team will improve winning. These variables 
-were kept because when a batter steps to the plate, the player is more likely to strike 
-out than to get a hit. Trying to hit the ball out of the park will come with strikeouts 
-but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is 
-a pretty good exchange that most teams are willing to perform.
+Studying the coefficients of the model suggest that winning is in favor if the
+team batting hits more doubles, triples and home runs. Moreover, increase in the
+number of stolen bases, and a decrease in caught steals, double plays, error,
+and walks allowed would all lead to a win for the batting team. It is noteworthy
+that the model suggests that a decrease in single hits by batter and an increase
+in strikeouts by batters which seems counter intuitive. But these variables were
+kept because when a batter steps to the plate, the player is more likely to
+strike out than to get a hit. Trying to hit the ball out of the park will come
+with strikeouts but it will also increase the chances of hitting home runs (even
+1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry
+out.
 
@@
 ```
+=======
+---
+title: 'DATA 621---Business Analytics and Data Mining'
+subtitle: 'Fall 2020---Group 2---Homework #1'
+author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
+    Bryan Persaud, Jeff Shamp
+date: "9/27/2020"
+output:
+  pdf_document:
+    toc: TRUE
+    toc_depth: 2
+  html_document: default
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
+```
+
+```{r loadData, include=FALSE}
+##################### libraries and variables 
+library(tidyverse)
+library(ggplot2)
+library(corrplot)
+library(tidymodels)
+library(mice)
+library(psych)
+library(Hmisc)
+library(MASS)
+library(dplyr)
+library(caret)
+library(data.table)
+
+urlRemote  = "https://raw.githubusercontent.com/"
+pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
+fileTrain   = "moneyball-training-data.csv"
+train_set = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
+
+fileEval   = "moneyball-evaluation-data.csv"
+eval_set = paste0(urlRemote, pathGithub, fileEval) %>% read.csv()
+
+set.seed(9450)
+```
+
+# DATA EXPLORATION
+## How Often Does The Team Win?
+We are given a data set of 2,276 records containing 15 seasonal statistics and
+the total number of wins a team had in a given year. On average, about 50% of
+games played are won (81 games out of 162), with the best individual season
+having 146 wins and the worst season having 0 wins. The data is normally
+distributed and most years have between 49 and 112 wins (blue lines below). The
+nature of the distribution means there aren't too many extreme seasons where
+wins are significantly higher or lower than usual. This serves as a good
+gut-check for our final predictions; if the predicted wins are too high or too 
+low, we know something in our model is probably off.
+
+```{r wins, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
+##################### DATA EXPLORATION
+
+avgWins <- mean(train_set$TARGET_WINS)
+sdev <- sd(train_set$TARGET_WINS)
+l1 <- mean(train_set$TARGET_WINS) - (2 * sdev)
+l2 <- mean(train_set$TARGET_WINS) + (2 * sdev)
+# histogram of wins
+p <- ggplot(train_set, aes(x=TARGET_WINS)) +
+  geom_histogram() + 
+  geom_vline(aes(xintercept = avgWins),
+            color = "red", linetype = "dashed", size = 1)+
+  geom_vline(aes(xintercept = l1),
+            color="blue", linetype = "dashed", size = 1) +
+    geom_vline(aes(xintercept = l2),
+            color = "blue", linetype = "dashed", size = 1) +
+  labs(title = "Distribution of Wins", x = "Number of wins",
+       y = "Number of seasons")
+p
+```
+
+## What's Missing? 
+A first look at the data shows that only about 8% of the records have a full set
+of information. The good news is that most of the missing values come from
+statistics that don't happen too often: hit-by-pitch (`TEAM_BATTING_HBP`, 92%
+missing!), caught stealing (`TEAM_BASERUN_CS`, 34% missing), and double plays
+(`TEAM_FIELDING_DP`, 13% missing). Since we have so little hit-by-pitch data, we
+expect that it doesn't contribute much to overall wins and will eliminate it
+from a few of the models we propose. The other two stats have less than half of
+the data missing, so we'll need to think of a clever way to fill in these
+values. The remaining missing information is from a combination of stolen bases
+and strikeouts by pitchers and batters (`TEAM_BASERUN_SB`, `TEAM_PITCHING_SO`,
+`TEAM_BATTING_SO`). **It seems completely unreasonable** to have zero strike
+outs in a season, so this is something we'll most certainly have to impute.
+
+```{r pctNull, fig.height = 2, fig.width = 5, fig.align = 'center'}
+
+# percent null for each variable
+pct_null <- data.frame(do.call("rbind", map(train_set %>% dplyr::select(-INDEX),
+                                            ~ mean(is.na(.)))))
+colnames(pct_null) <- c('PCT_NULL')
+totalNulls <- pct_null %>%
+  mutate(VARIABLE = rownames(.)) %>%
+  arrange(desc(PCT_NULL)) %>%
+  filter(PCT_NULL > 0) %>%
+  dplyr::select(VARIABLE, PCT_NULL)
+ggplot(totalNulls, aes(x = reorder(VARIABLE, PCT_NULL), y = PCT_NULL,
+                       label = round(PCT_NULL, 2))) + 
+  geom_text(vjust = 0.5, hjust = -0.05)+
+  geom_bar(stat = "identity") +
+  ggtitle("Variables with Missing Information") +
+  xlab("Statistic") + ylab("Percent Missing") + 
+  coord_flip() + expand_limits(y = 1)
+```
+
+## Do the Individual Stats Affect Winning?
+**Stats with an expected negative impact:** Intuitively, we expect that Caught
+Stealing, Errors, Hits Allowed, Home Runs Allowed, Strikeouts by Batters, and
+Walks Allowed would all have a **negative** impact on the total wins. In other
+words, as these values increase, we expect that the team is less likely to win.
+
+```{r negFactors, fig.align = 'center'}
+# variables expected to have a negative impact on wins
+negSet <- train_set %>%
+  dplyr::select(TARGET_WINS, 
+         'Errors' = TEAM_FIELDING_E, 
+         'Hits allowed' = TEAM_PITCHING_H, 
+         'Strikeouts by batters' = TEAM_BATTING_SO, 
+         'Caught stealing' = TEAM_BASERUN_CS, 
+         'Walks allowed' = TEAM_PITCHING_BB, 
+         'Home Runs allowed' = TEAM_PITCHING_HR)
+ggplot(data = negSet %>%
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
+  aes(x = VALUE, y = TARGET_WINS)) + 
+  geom_point() + geom_smooth(method = "lm", se = FALSE) +
+  labs(title = "Number of wins with respect to recorded stats",
+       x = "", y = "Number of Wins") + facet_wrap(~ STAT, scales = "free")
+```
+
+When we take a closer look at the data, these negative relationships aren't
+obvious. In fact, only **Errors** and **Hits Allowed** seem to have a negative
+impact on wins. **Caught Stealing** and **Strikeouts by Batters** appear to be
+random; this means that whether the stat for a particular season is high or low
+doesn't affect the overall number of wins.
+
+Even more interestingly, **Home Runs Allowed** and **Walks Allowed** have the
+*opposite* effect---as these stats increase, so do the number of wins!
+
+**Stats with an expected positive impact:** We can look at the same information
+for the stats that we expect to have a **positive** effect on wins: Base Hits,
+Doubles, Triples, Home Runs, Walks, Batters getting hit by pitches, Stolen
+Bases, Double Plays, and Strikeouts by Pitchers.
+
+```{r posFactors, fig.align = 'center'}
+# variables expected to have a positive impact on wins
+posSet <- train_set %>%
+  dplyr::select(TARGET_WINS, 
+         'Base hits' = TEAM_BATTING_H, 
+         'Doubles' = TEAM_BATTING_2B, 
+         'Triples' = TEAM_BATTING_3B, 
+         'Home Runs' = TEAM_BATTING_HR, 
+         'Walks' = TEAM_BATTING_BB, 
+         'Batters hit' = TEAM_BATTING_HBP,
+         'Stolen bases' = TEAM_BASERUN_SB,
+         'Double plays' = TEAM_FIELDING_DP,
+         'Strikeouts' = TEAM_PITCHING_SO)
+ggplot(data = posSet %>%
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
+  aes(x = VALUE, y = TARGET_WINS)) + geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(title = "Number of Wins with Respect to Recorded Stats", x = "",
+       y = "Number of Wins") + facet_wrap(~ STAT, scales = "free")
+```
+
+Many of these stats *do* seem to have an effect on the number of wins, most
+notably, **Base Hits** and **Walks**. We see weaker positive relationships for
+**Home Runs**, **Doubles**, **Triples**, and **Stolen Bases**. This makes sense
+when we think about it; these things tend to happen less often in games than
+pure base hits and walks, so they don't have as much of an effect on winning.
+Finally, **Double Plays** and **Batters Hit** don't appear to have any
+correlation with the number of wins. Once again, this intuitively makes sense
+because they are less likely to happen in a game.
+
+One thing to note is the number of strikeouts compared to the number of wins. We
+can see that there are a few outliers (abnormally high numbers of strikeouts in
+a season). This should be taken with caution, as they don't represent a typical
+season's stats.
+
+## Are Some Stats More Skewed Than Others?
+Before using any of the statistics in a model, we need to take a closer look at
+the variation in the data. We call out-of-the-ordinary values (exceptionally
+high or low values) **outliers**. We need to take these into account in our
+modeling because we want to make sure our predictions aren't skewed because of
+them.
+
+```{r summaryInfo, fig.height = 3, fig.width = 5, fig.align = 'center'}
+# understanding outliers by variable
+totalOutliers <- data.frame(
+  sapply(train_set %>% dplyr::select(-INDEX, -TARGET_WINS, - TEAM_BATTING_HBP), 
+       function(x){length(boxplot.stats(x)$out)/nrow(train_set)}))
+totalOutliers$VARIABLE_NM <- rownames(totalOutliers)
+colnames(totalOutliers) <- c('PCT_OUTLIERS', 'VARIABLE_NM')
+ggplot(totalOutliers,
+       aes(x = reorder(VARIABLE_NM, PCT_OUTLIERS), y=PCT_OUTLIERS,
+           label = round(PCT_OUTLIERS, 3))) + 
+  geom_text(vjust = 0.5, hjust = -0.05)+ geom_bar(stat = "identity") +
+  ggtitle("Percentage of Outliers") + xlab("Statistic") +
+  ylab("Percent of Data that is an Outlier") + coord_flip() +
+  expand_limits(y = 0.15)
+```
+
+As we can see, some of the provided statistics are well-balanced in the sense
+that there are very *few* (or no) extreme values. **Home Runs allowed**
+(`TEAM_PITCHING_HR`), **Strikeouts by batters** (`TEAM_BATTING_SO`), and
+**Home Runs by batters** (`TEAM_BATTING_HR`) are some examples.
+
+```{r fewOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
+# distributions of variables with few outliers
+set1 <- train_set %>%
+  dplyr::select('Home Runs by batters' = TEAM_BATTING_HR, 
+         'Strikeouts by batters' = TEAM_BATTING_SO,  
+         'Home Runs allowed' = TEAM_PITCHING_HR) %>%
+  gather("stat", "value") %>%
+  filter(complete.cases(.) == TRUE)
+vals <- ggplot(set1 , aes(x = stat, y = value)) + 
+    geom_boxplot() + labs(title = "", x = "", y = "Value") +
+  facet_wrap( ~ stat, scales = "free")
+vals
+```
+
+Some things to note about each of these statistics:
+
+  * **Home Runs Allowed** (average ~100/year) and **Home Runs by Batters**
+  (average ~106/year) have a very similar mid-range distribution (50% of the
+  data lies between ~50 and 150). The slight difference in average stats means
+  that teams tend to have a higher number of Home Runs than the opposition team.
+  * The only thing that stands out about **Strikeouts by Batters** (average
+  ~736/year) is how nearly perfectly normal it is. 50% of the data is between
+  about 500 and 1000 and there are absolutely no outliers in the dataset! This
+  means that there were no surprisingly high or low seasons.
+
+Conversely, some of the stats have a very *high* number of outliers, indicating
+that there are some seasons with some abnormally high or low values. **Errors**
+(`TEAM_FIELDING_E`), **Hits allowed** (`TEAM_PITCHING_H`), **Walks by batters**
+(`TEAM_BATTING_BB`), and **Stolen bases** (`TEAM_BASERUN_SB`) are some examples.
+
+```{r manyOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
+# distributions of variables with many outliers
+set2 <-  train_set %>%
+  dplyr::select('Walks by batters' = TEAM_BATTING_BB, 
+         'Stolen bases' = TEAM_BASERUN_SB,  
+         'Hits allowed' = TEAM_PITCHING_H,
+         'Errors' = TEAM_FIELDING_E) %>%
+  gather("stat", "value") %>%
+  filter(complete.cases(.) == TRUE)
+vals2 <- ggplot(set2 , aes(x=stat, y=value)) + 
+    geom_boxplot() +
+  labs(title = "",x = "", y = "Value") +
+  facet_wrap(~ stat, scales = 'free', ncol = 4)
+vals2
+```
+
+Some things to note about each of these statistics:
+
+  * All of the outliers for **Errors**, **Hits Allowed**, and **Stolen Bases**
+  are above the upper tail of the data set. This is further illustrated by the
+  mean and median values for these stats; in all instances, the means per year
+  (Errors ~246/year, Hits Allowed ~1779/year, Stolen Bases ~125/year) are higher
+  than the medians per year (Errors ~159/year, Hits Allowed ~1518/year, Stolen
+  Bases ~101/year). This means that some seasons with exceptionally high values
+  skew the dataset.
+  * There are a few *very* extreme outliers for **Hits Allowed**. The maximum
+  value is 30,132, which is over 16 times the average number of hits allowed per
+  season!
+  * There are outliers both above *and* below the tails of the data for the
+  **Walks by Batters** stat. This means that we have exceptionally low
+  (min = 0!) and exceptionally high (max = 878) seasons.
+
+The remaining stats, **Walks Allowed** (`TEAM_PITCHING_BB`),
+**Base Hits by Batters** (`TEAM_BATTING_H`), **Caught Stealing**
+(`TEAM_BASERUN_CS`), **Strikeouts by Pitchers** (`TEAM_PITCHING_SO`),
+**Double Plays** (`TEAM_FIELDING_DP`), **Triples by Batters**
+(`TEAM_BATTING_3B`), and **Doubles by Batters** (`TEAM_BATTING_2B`) have between
+29 and 99 outliers.
+
+```{r otherOutliers, fig.height = 3.5, fig.width = 7, fig.align = 'center'}
+# distribution of all other variables 
+set3 <-  train_set %>%
+  dplyr::select('Base Hits by batters' = TEAM_BATTING_H, 
+         'Doubles by batters' = TEAM_BATTING_2B,  
+         'Triples by batters' = TEAM_BATTING_3B,
+         'Caught stealing' = TEAM_BASERUN_CS,
+         'Walks allowed' = TEAM_PITCHING_BB,
+         'Strikeouts by pitchers' = TEAM_PITCHING_SO,
+         'Double plays' = TEAM_FIELDING_DP) %>%
+  gather("stat", "value") %>%
+  filter(complete.cases(.) == TRUE)
+vals3 <- ggplot(set3, aes(x=stat, y=value)) + 
+    geom_boxplot() +
+  labs(title = "", x = "",y = "Value") +
+  facet_wrap(~ stat, scales = 'free', ncol = 4)
+vals3
+```
+
+Some things to note:
+
+  * All variables are very narrowly distributed, meaning that most of the data
+  falls within a small range. 
+  * **Strikeouts by Pitchers** and **Walks Allowed** have a few very extreme
+  outliers; these   represent seasons that have abnormally high values for the
+  statistics. 
+  * The average number of pure **Base Hits** (1469/season) is greater than the
+  average number of Doubles (~241/season) and Triples (~55/season). This isn't
+  at all surprising, but serves as a good gut check on the validity of the data.
+
+## Are Stats Correlated?
+We would expect that a few things in the dataset might be correlated: perhaps
+number of errors and hits/homeruns allowed or the number of base hits by batters
+and doubles/triples/homeruns. We can visualize the correlations between the
+statistics to determine if there is a significant relationship between the them:
+blue dots represent positively correlated variables (as one increases, so does
+the other) and red dots represent negatively correlated variables (as one
+increases, the other decreases).
+
+```{r corrPlot, fig.align = 'center'}
+# Correlation matrix with significance levels (p-value)
+res2 <- rcorr(as.matrix(train_set %>% dplyr::select(-INDEX)))
+# Insignificant correlation are crossed
+corrplot(res2$r, type="upper", order="hclust", 
+          tl.col = "black", p.mat = res2$P, sig.level = 0.01, insig = "blank")
+```
+
+Some noteworthy relationships (coincidental or not):
+
+  * **Errors** are highly, negatively correlated with walks by batters,
+  strikeouts by batters, and homeruns (both by batters and allowed).
+  * **Triples by Batters** are highly, negatively correlated with strikeouts by
+  batters and homeruns (both by batters and allowed).
+  * **Strikeouts by Batters** are highly, positively correlated with homeruns
+  (both by batters and allowed). 
+  * **Homeruns by Batters** and **Homeruns Allowed** are both positively
+  correlated with walks by batters.
+  * As expected, **Base Hits** are positively correlated with doubles and
+  triples.
+
+We can keep these correlations in mind when developing our models: if we have
+correlated statistics, there could be in-built redundancy in the features, and
+we may be able to create a simpler, more accurate model by eliminating some.
+
+# DATA PREPARATION
+Now we have a good idea of the data we are looking at we can take the next steps
+to prepare it for building a solid model.
+
+## Outlier Removal
+As we saw in the data analysis, there are some outlier concerns for some of the
+variables, so we will need to account for this in our modeling. When performing
+the processes of outlier removal, a cautious approach is always best. Each
+outlier is evaluated to ensure:
+
+  * It is clearly from incorrectly or mis-centered data.
+  * Its removal does not affect later assumptions.
+  * It creates a significant association/relation that is eliminated with its removal.
+
+We can take another look at the outliers for each variable:
+
+```{r outliers_1, fig.height = 4, fig.width = 8, fig.align = 'center'}
+
+##################### DATA PREPARATION
+
+# visualizing variables pre outlier removal
+train_set %>%
+  dplyr::select(-TARGET_WINS, -INDEX) %>%
+  pivot_longer(everything(), 
+               names_to = "Variable", 
+               values_to = "Value") %>%
+  ggplot(aes(x = Variable, y = Value)) +
+  geom_boxplot(na.rm = TRUE) + 
+  labs(title="Outliers for each Statistic",x="Variable", y = "Value") +
+  coord_flip()
+```
+
+Off the bat (pun intended), there are some clear issues with two of the
+variables: `TEAM_PITCHING_SO` and `TEAM_PITCHING_H.`
+
+  * **TEAM_PITCHING_SO**: The largest outlier is close to 20,000, which averages
+  ~123 strikeouts per game. This is only possible if every one of the 162 games
+  went approximately 40 innings, and each out was a strikeout---which clearly has
+  never happened.
+  * **TEAM_PITCHING_H**: The largest outlier for this variable is over 30,000
+  which is similarly highly unlikely.
+
+Since both appear with a heavy right skew, we will use median and IQR to remove
+the outliers.
+
+```{r iqr_removal_SO}
+# remove outliers
+train_set<- train_set %>%
+  na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
+  bind_cols(
+            train_set 
+            )  %>%
+  filter(
+    TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO,
+                                probs = c(0.25), 
+                                na.rm = TRUE) - 1.5 * iqr, 
+    TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, 
+                                probs = c(0.75), 
+                                na.rm = TRUE) + 1.5 * iqr
+        ) %>%
+  dplyr::select(-iqr)
+```
+```{r iqr_removal_H}
+train_set<- train_set %>%
+  na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
+  bind_cols(
+            train_set 
+            )  %>%
+  filter(
+    TEAM_PITCHING_H > quantile(TEAM_PITCHING_H,
+                                probs = c(0.25), 
+                                na.rm = TRUE) - 1.5 * iqr, 
+    TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, 
+                                probs = c(0.75), 
+                                na.rm = TRUE) + 1.5 * iqr
+        ) %>%
+  dplyr::select(-iqr)
+```
+
+We can take a look at the distributions of these variables after removing the
+outliers:
+
+```{r outlier_boxplot, fig.height = 4, fig.width = 8, fig.align = 'center'}
+# visualizing variables post outlier removal
+train_set %>%
+  dplyr::select(-TARGET_WINS, -INDEX) %>%
+  pivot_longer(everything(), 
+               names_to ="Variable", 
+               values_to="Value") %>%
+  ggplot(aes(x=Variable, y=Value)) +
+  geom_boxplot(na.rm = TRUE) + 
+  labs(title="Outliers for each Statistic",x="Variable", y = "Value") +
+  coord_flip()
+```
+
+This has eliminated what appears to be the most extreme outliers. There are
+still large outlier sets for Errors and Stolen Bases, but none that seem to
+dwarf the other variables. According to the box plot, things appear to be on a 
+scale that seems logical given the data and source of information that we have.
+
+## Imputation
+From our early exploration of the data, the vast majority of the data is
+complete with only a few variables with missing values. We will use the **MICE**
+(Multivariate Imputation by Chained Equations) method. MICE is a principled
+method for dealing with missing data. It creates multiple imputations, as
+opposed to single imputations, and accounts for the statistical uncertainty in
+these imputations. It creates predictive values for the mean instead of imputing
+the IQR values.
+
+```{r impute_train}
+# imputing values
+train_set <- complete(mice(data = train_set, 
+                         method = "pmm", 
+                         seed = 9450, 
+                         print=FALSE), 3)
+```
+
+## Feature Engineering
+### Single base hits
+With outliers removed and missing values imputed via MICE, we can create a few
+new variables. The first will be single base hits, and it will be derived from
+some of the variables we do have. We know that Team Batting Hits
+(`TEAM_BATTING_H`) is a combination of *all* hits for the season, so we can
+create Single Base Hits as follows: 
+\[
+Singles = Total\ Hits - (Doubles + Triples + Homeruns)
+\]
+
+```{r singles}
+# defining singles variable 
+train_set<- train_set %>%
+  dplyr::select(-INDEX) %>%
+  mutate_if(is.integer, as.numeric) %>%
+  mutate(TEAM_BATTING_1B =
+           TEAM_BATTING_H -
+           (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR))
+```
+
+### Slugging Percentage
+The second variable we will create is **Slugging Percentage**. Slugging is an
+offensive statistic that is a good predictor of winning and it tends to have
+better variance and correlation behavior than most other variables. It is
+composed of singles, doubles, triples, home runs and at-bats:
+
+\[
+\mathrm{SLG} ={\frac{{\mathit{1B}} + 2\times{\mathit{2B}} +
+3\times{\mathit{3B}} + 4\times {\mathit{HR}}}{\mathit{AtBats}}}
+\]
+Because we don't have a statistic for at-bats, we will approximate it as follows:
+
+\[
+\widehat{\mathrm{SLG}} = {\frac{\mathit{1B} + 2\times\mathit{2B} +
+3\times\mathit{3B} + 4\times\mathit{HR}}{SO_{batting} + H + BB}}
+\]
+
+```{r make_slg_obp}
+# defining at bats and slugging variables
+train_set<- train_set %>%
+  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
+         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
+                  4 * TEAM_BATTING_HR) / AB)
+```
+
+As a gut-check, we can take a look at Slugging in comparison to Triples
+(`TEAM_BATTING_3B`) and Home Runs (`TEAM_BATTING_HR`). All of these variables
+should have a positive correlation with the total number of wins.
+
+```{r slugging_compare, fig.height = 2.5, fig.width = 8, fig.align = 'center'}
+# visualizing slugging compared to wins
+modified_set <- train_set %>%
+  dplyr::select(TARGET_WINS, SLG, TEAM_BATTING_3B, TEAM_BATTING_HR)
+ggplot(data = modified_set %>%
+  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
+  aes(x = VALUE, y = TARGET_WINS)) + 
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(title = "Number of wins with respect to recorded stats", x = "",
+       y = "Number of Wins") +
+  facet_wrap(~ STAT, scales = "free")
+```
+
+# BUILD MODELS
+## Model 1
+### Base Model
+We will start with a simple linear model to serve as a baseline. This includes
+all variables in the dataset.
+
+```{r baseModel}
+##################### BUILD MODELS
+##### Model 1: Backwards Selection
+lm_reg = lm(data=train_set, TARGET_WINS ~ .)
+
+knitr::kable(summary(lm_reg)$coefficients, digits = 3L,
+             caption = 'Base Model Regression Output')
+```
+
+We can see off the bat that a few variables *exceed* the 0.05 p-value threshold for 
+significance. Additionally, we see that both `TEAM_BATTING_1B` and `AB` have no 
+coefficients at all! This is likely due to the fact that they are colinear with other 
+variables in the model.
+
+### Enhanced Model
+
+We used a combination of domain knowledge and data analysis to justify removing/retaining 
+features. First, we are keeping the coefficient for `TEAM_BATTING_HR` since its p-value
+is marginally above the general threshold and knowledge of the game suggests it is important. 
+We also chose to remove the `TEAM_BATTING_HBP` variable because most of the values are
+missing from the dataset.
+
+The remaining variables were chosen as a result of **backwards selection** based on null 
+hypothesis testing for non-zero slope. This is done by removing one variable at a
+time to determine the minimum number of significant variables needed for prediction.
+
+Altogether, we end up **removing** 7 total variables: `TEAM_BATTING_1B`, `AB`, 
+`TEAM_BATTING_3B`, `TEAM_BATTING_H`, `TEAM_PITCHING_BB`, `TEAM_PITCHING_SO`, and
+`TEAM_BATTING_1B`.
+
+### Further data transformation
+We will use Cook's distance to remove additional outliers that are influencing the fit of
+the model above. We will use a cutoff of \(\frac{4}{N}\).
+
+```{r cooks_distance}
+# removing outliers via cooks distance
+lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
+          TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB +
+            TEAM_BASERUN_CS + TEAM_PITCHING_HR + TEAM_PITCHING_H +
+            TEAM_FIELDING_E + TEAM_FIELDING_DP + SLG, data = train_set)
+cooks_dis<- cooks.distance(lm_1)
+influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4 / nrow(train_set)))])
+train_df2<- train_set[-influential, ]
+lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
+            TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB +
+            TEAM_BASERUN_CS + TEAM_PITCHING_HR + TEAM_PITCHING_H +
+            TEAM_FIELDING_E + TEAM_FIELDING_DP + SLG, data = train_df2)
+mod_sum1 = summary(lm_1)  
+```
+
+```{r model1_coef}
+knitr::kable(summary(lm_1)$coefficients, digits = 3L,
+             caption = 'Backwards Model Regression Output')
+```
+
+### Coefficient Discussion
+There are some counter intuitive results, which is expected given that baseball is a 
+messy, imprecise game that has evolved over time. To that end, we should expect some 
+seemingly strange results from algorithmic regression.
+
+First, we see that the intercept itself is negative. This means that, given no information
+about any of the statistics, we would predict that a team would lose (by a lot!).
+
+Next, fielding double plays and batting doubles both appear to have negative impacts
+on wins even though they *should* be positive impacts. Turning double plays,
+while a good for the defensive team, may suggest a larger, negative issue.
+Namely, weak pitching that leads to runners on base. Similarly, batting doubles,
+leaves runners open to double plays. Allowed hits by pitching, caught stealing,
+and batting strike-outs are all counter intuitive results as well, but they seem
+to be small contributors and these are events that happen regularly in every
+game.
+
+Slugging as an approximation is the major predictor in this regression, by far.
+The other predictors other than the intercept are orders of magnitude less in
+predictive value. It should be noted that slugging alone is not a good predictor
+for wins overall.
+
+## Model 2
+This is a model allowing pairwise interactions within the data types of
+baserunning, batting, pitching, and fielding. It is fit using a forward and
+backwards stepwise regression based on AIC.
+
+### Further data transformation
+All the features will be centered and scaled. The target variable will be left
+in normal space. This will make prediction easier.
+
+```{r model2C+S}
+##### Model 2: Pairwise Model
+# cleaning data for model 2
+ts2 <- train_set[, -(17:19)]
+ts2 <- data.frame(apply(ts2, 2, scale))
+ts2$TARGET_WINS <- train_set$TARGET_WINS
+```
+
+### Stepwise Model
+A fully specified model with all the interactions will be the starting point for
+the stepwise regression, which uses AIC as its target metric. For interest, this
+model will be compared with both the null model and the fully-specified model
+from where the stepwise process begins
+
+```{r model2}
+nullModel <- lm(TARGET_WINS ~ 1, data = ts2)
+fullModel <- lm(TARGET_WINS ~
+  (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
+  TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_SO) ^ 2 +
+  (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
+  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = ts2)
+stepModel <- stepAIC(fullModel, scope = list(upper = ~
+  (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
+  TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_SO) ^ 2 +
+  (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
+  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, lower = ~ 1),
+  direction = 'both', trace = 0)
+snm <- summary(nullModel)
+sfm <- summary(fullModel)
+sstm <- summary(stepModel)
+stmr2 <- sstm$r.squared
+stmar2 <- sstm$adj.r.squared
+stmAIC <- AIC(stepModel)
+stmRMSE <- sqrt(mean(stepModel$residuals ^ 2))
+```
+
+### Discussion
+#### Models
+```{r model2_discuss}
+knitr::kable(snm$coefficients, digits = 3L,
+             caption = 'Null Model')
+knitr::kable(sfm$coefficients, digits = 3L,
+             caption = 'Full Model with All Features & Pairwise Class Interactions')
+knitr::kable(sstm$coefficients, digits = 3L,
+             caption = 'Stepwise Model with Pairwise Class Interactions')
+```
+
+The null model expects teams to lose slightly more than half their games. The
+full model has a lower intercept, which indicates that the impact of all the
+features explains why teams *win* slightly more than just the grand mean does.
+The stepwise model is similar, having a lower intercept than the null but a
+higher intercept that the full model.
+
+Also, the full model has the higher \(R^2\) and the stepwise model has the
+higher *adjusted* \(R^2\). This is expected, as there is no penalty for adding
+more parameters in regular \(R^2\) measures.
+
+The **stepwise** model will be the one discussed below.
+
+#### Coefficients
+Most of the coefficients are reasonable within the context of the game of
+baseball. These two statements are axiomatic:
+
+  * The only way to win is to have the higher score at the end of the game.
+  * The only way to score is for a baserunner to cross home plate.
+
+With those in mind, we can make the following observations about the linear
+non-interactive coefficients.
+
+ * Reasonable
+   * Caught stealing removes baserunners but it isn't significant and can be
+   ignored on its own.
+   * Stolen bases get a runner closer to home plate; positive coefficient makes
+   sense.
+   * Triples get a runner very close to home plate; positive coefficient makes
+   sense.
+   * Walks are free baserunners; positive coefficient makes sense but it isn't
+   significant and can be ignored on its own.
+   * Hitting home runs directly increase the score; positive coefficient makes
+   sense.
+   * Striking out reduces the number of baserunners; negative coefficient makes
+   sense.
+   * Errors allow the other team free baserunners; negative coefficient makes
+   sense.
+   * Giving up walks allows the other team free baserunners; negative
+   coefficient makes sense but it isn't significant and can be ignored on its
+   own.
+   * Giving up home runs gives the opponent scores; negative coefficient makes
+   sense.
+   * Getting strikeouts reduces the opponents baserunners; positive coefficient
+   makes sense.
+ * Curious
+   * Hits increase the number of baserunners; why negative and insignificant?
+   * Hitting doubles get runners on base; why negative?
+   * Turning double plays reduce baserunners; why negative?
+   * Giving up hits allows the other team baserunners; why positive although
+   insignificant?
+
+First, an absolutely fascinating observation from an earlier state of the model.
+In the very first draft of this exercise, prior to some of the data adjustments
+performed here, `TEAM_BATTING_HR` was removed due to its high correlation with
+`TEAM_PITCHING_HR`. In that first draft, pitching giving up home runs was given
+a *positive* coefficient. This made no sense. Allowing batting home runs to
+re-enter the model, In hindsight this becomes obvious. The pitching
+*was being used as an indicator for* ***batting!!*** They are almost 100%
+correlated! This correlation allowed the use of pitching HRs as an indicator for
+the hidden batting HRs which has a larger magnitude! Once both variables were
+restored to the model, the logical coefficients surfaced. This is yet another
+reason why models should not be trusted out of the box, but all model results
+should be reviewed for sanity and sense!!
+
+The curiosities above can *possibly* be resolved by looking at the interaction
+terms. The pitching and batting home run and strikeout coefficients are an order
+of magnitude greater than all the others. It is likely that the negative
+coefficients in some of the batting stats and positive coefficients in some of
+the pitching stats are being used to "temper" the effect of SOs and HRs.
+
+A possible explanation for the negative coefficient for getting double plays, is
+that double plays require at least two people on base. That means that the
+opponent has a lot of base runners, which is very highly correlated with
+scoring.
+
+The behavior of doubles remains confusing. Perhaps its presence in a number of
+the interaction terms means that the singleton variable needs to act as a
+balance. Another theory may be that it's hiding something like a team's
+propensity to strand runners on base. It would be interesting to see a breakdown
+between the American and National leagues, as the latter tends to be somewhat
+better at "small ball" and moving the runners along.
+
+## Model 3
+The final model is a Stepwise Regression with Repeated k-fold Cross-Validation,
+and higher order polynomials variables were introduced into the full model.
+
+A stepwise variable selection model is conducted to determine what are the
+variables that can help predict the number of wins for the team. The method
+allows variables to be added one at a time to the model, as long as the
+F-statistic is below the specified \(\alpha\), in this case \(\alpha = 0.05\).
+However, variables already in the model do not necessarily stay in. The steps
+evaluate all of the variables already included in the model and remove any
+variable that has an insignificant F-statistic. Only after this test ends, is
+the best model found, that is when none of the variables can be excluded
+and every variable included in the model is significant.
+
+Here, the dependent variable is the continuous variable, `TARGET_WINS`, and the
+independent variables are the full model to identify the most contributing
+predictors. In addition, a robust method for estimating the accuracy of a model,
+the k-fold cross-validation method, was performed evaluate the model performance
+on different subset of the training data and then calculate the average
+prediction error rate.
+
+### Further data transformation
+Once again, Cook's distance is use to decide if an individual entity is an extreme 
+value or not.
+
+```{r model3}
+
+##### Model 3: K-fold Model
+stepdata <- subset(train_set, select = -c(AB, SLG, TEAM_BATTING_HBP))
+stepdata <- cbind(stepdata, sapply(stepdata[2:length(stepdata)], function(x) x^2))
+names(stepdata) <- make.unique(names(stepdata), sep = "_")
+model <- lm(TARGET_WINS ~ ., data = stepdata)
+cooksd <- cooks.distance(model)
+influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
+stepdata.2 <- stepdata[-c(influential), ]
+# Set up repeated k-fold cross-validation
+set.seed(525)
+train.control <- trainControl(method = "cv", number = 10)
+# Train the model
+step.model <- train(TARGET_WINS ~ ., data = stepdata.2, method = "lmStepAIC", 
+                   trControl = train.control, trace = FALSE)
+# Final model 
+step.model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BASERUN_SB + 
+                  TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_FIELDING_E + 
+                  TEAM_FIELDING_DP + TEAM_BATTING_BB_1 + TEAM_BATTING_SO_1 + 
+                  TEAM_BASERUN_CS_1 + TEAM_PITCHING_H_1 + TEAM_PITCHING_BB_1 + 
+                  TEAM_FIELDING_E_1 + TEAM_FIELDING_DP_1 + TEAM_BATTING_1B_1, 
+                data = stepdata.2)
+# Summary of the model
+mod_sum3 <- summary(step.model)
+```
+
+### Coefficient Discussion
+```{r model3_discuss}
+mod_sum3[["coefficients"]] %>% 
+  knitr::kable(digits = 3L, 
+               caption = 'K-fold Step Model with Higher Order Polynomials Output')
+```
+
+From studying each coefficient, this discussion begins with the variables that follow 
+the logic of winning and are self-evident. Firstly, the model suggests that winning is 
+in favor of the team batting hits more bases, (1B, 2B, 3B, HR). It further reveals that 
+an increase in singles would lead to better chances of winning. A hit is a hit, and more 
+hits than the opponent will inevitably lead to a win. Likewise, an increase in walks by 
+the batting team is awarded first base, further increasing base counts and, thus the 
+chances of winning. In terms of the base run, an increase in the number of stolen bases, 
+and a decrease in caught steals would lead to a win for the batting team. These variables 
+have a positive effect on the number of wins, and it is understandable as to why.
+
+When interpreting the model based on the pitching team's plays, it is perceptible that 
+walks allowed should be reduced so that the batting team hits does not increase. Moreover, 
+preventing errors and increasing double plays in which two players are put out will also 
+increase the chances of winning.
+
+On the other hand, it is noteworthy that the model suggests some counter-intuitive play 
+strategies. Firstly, when all variables are null, the intercept predicts that a team would 
+win, with a target win of 129. Moreover, the model suggests that a decrease in double hits 
+and an increase in strikeouts by the batting team will improve winning. These variables 
+were kept because when a batter steps to the plate, the player is more likely to strike 
+out than to get a hit. Trying to hit the ball out of the park will come with strikeouts 
+but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is 
+a pretty good exchange that most teams are willing to perform.
+
+# SELECT MODELS
+## Criteria
+In order to select the best model to make predictions, we looked at the
+following goodness of fit metrics:
+
+ 1. \(R^2\), which represents the proportion of the variance explained by the
+ model
+ 2. Adjusted \(R^2\), which is a modified version of \(R^2\) taking parsimony
+ into consideration
+ 3. *Root Mean Squared Error* (RMSE), which is the square root of the mean
+ squared difference between the observation and the fitted value
+ 4. *Akaike Information Criterion* (AIC), which is an information-criteria based
+ estimate of the Kullback-Leibler divergence from the current model to the
+ "true" model. One of the nice results of the theory of information-criteria
+ based measures is that there is included recognition of parsimony.
+
+## Performance
+```{r performance}
+##################### SELECT MODELS
+# Model accuracy
+results <- data.frame(
+  Models = c("Model #1: Backwards" , "Model #2: Pairwise", "Model #3: K-fold"),
+  R.squared = c(mod_sum1[["r.squared"]], stmr2, mod_sum3[["r.squared"]]), 
+  adj.R.squared = c(mod_sum1[["adj.r.squared"]], stmar2,
+                    mod_sum3[["adj.r.squared"]]),
+  RSME = c(sqrt(mean(lm_1$residuals ^ 2)), stmRMSE,
+           sqrt(mean(step.model$residuals ^ 2))),
+  AIC = c(AIC(lm_1), AIC(stepModel), AIC(step.model)))
+results %>% knitr::kable(digits = 3L, 
+                         caption = 'Performance Statistics & Error Measure of Models')
+```
+
+```{r errors}
+
+# calculate predictions for each model 
+mod1_preds <- data.frame(cbind(train_df2$TARGET_WINS, round(mod_sum1$residuals,0), train_df2$TARGET_WINS + round(mod_sum1$residuals,0) ))
+colnames(mod1_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')
+
+mod2_preds <- data.frame(cbind(ts2$TARGET_WINS, round(sstm$residuals,0), ts2$TARGET_WINS + round(sstm$residuals,0) ))
+colnames(mod2_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')
+
+mod3_preds <- data.frame(cbind(stepdata.2$TARGET_WINS, round(mod_sum3$residuals,0), stepdata.2$TARGET_WINS + round(mod_sum3$residuals,0) ))
+colnames(mod3_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')
+
+# summary statistics for residuals
+resid_stats <- data.frame(rbind(summary(mod1_preds$RESIDUAL), summary(mod2_preds$RESIDUAL), summary(mod3_preds$RESIDUAL)))
+
+resid_stats$Models <- c('Model #1: Backwards' , 'Model #2: Pairwise', 'Model #3: K-fold')
+
+```
+
+It was interesting to compare the three models given that their performance
+statistics, error measurements, and residual distributions were not drastically 
+different from each other. **Model #2: Pairwise** was the worst performing model 
+for all of our metrics, and was eliminated immediately. **Model #3: K-fold** accounts 
+for nearly 45% of the variation in the dependent variable with the independent 
+variables because its \(R^2\) is 0.456 and its adjusted \(R^2\) is = 0.450, both of 
+which are acceptable for a good model. However, **Model #1: Backwards** has the 
+smallest RSME of all the models. RMSE is a measure of how far off predictions are 
+from the actual, giving extra weight to outliers through summing the squared
+differences. This demonstrates a greater reduction in the randomness by
+**Model #1: Backwards**. But its difference from **Model #3: K-fold** is only
+0.022. It is hard to judge if that is a significant difference. We thus also
+considered AIC. With AIC, it is not the magnitude of the value which is
+important, but the difference between the various model AIC's. The magnitude is
+a function of the number of observations. Accepted rules of thumb in the
+statistical literature are that if the difference is greater than 10, there is
+effectively no support for the model with the larger AIC
+(Section 2.6, Burnham & Anderson 2002). As a result, **Model #1: Backwards**
+is our selected model.
+
+## Discussion
+This is an interesting model in that given no other information, a team would be
+expected to win negative games, which is clearly impossible. That being said, it
+is an indication of the predictive power of the engineereed **slugging** feature
+that allows small increments in slugging to result in significant shifts in
+target wins. This engineered feature contains within it numerous interactions.
+It combines scaled versions of some features and normalizes them for the
+combination of other features. Perhaps that is why it is so significant.
+
+## Predictions
+**Model #1: Backwards** was used to predict the target wins base on the
+evaluation set, once processed similarly to the training data. As the number of
+games won must be integral, the values were rounded to the closest integer.
+
+```{r eval_set, eval=FALSE}
+# predict target wins
+eval_set = eval_set %>% na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
+  bind_cols(eval_set) %>%
+  filter(TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO, probs = c(0.25), na.rm = TRUE) 
+         - 1.5 * iqr, 
+         TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, probs = c(0.75), na.rm = TRUE) 
+         + 1.5 * iqr) %>%
+  dplyr::select(-iqr)
+
+eval_set = eval_set %>% na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
+  bind_cols(eval_set) %>%
+  filter(TEAM_PITCHING_H > quantile(TEAM_PITCHING_H, probs = c(0.25), na.rm = TRUE) 
+         - 1.5 * iqr,
+         TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, probs = c(0.75), na.rm = TRUE) 
+         + 1.5 * iqr) %>%
+  dplyr::select(-iqr)
+
+eval_set = complete(mice(data = eval_set, method = "pmm", seed = 9450,  print=FALSE), 3)
+
+eval_set = eval_set %>% dplyr::select(-INDEX) %>%
+  mutate_if(is.integer, as.numeric) %>%
+  mutate(TEAM_BATTING_1B = TEAM_BATTING_H - (TEAM_BATTING_2B + TEAM_BATTING_3B +
+                                               TEAM_BATTING_HR))
+
+# Cleaning Test Set for Model #1
+test_df1 = eval_set %>%
+  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
+         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
+                  4 * TEAM_BATTING_HR) / AB)
+# Cleaning Test Set for Model #2
+test_df2 = test_df1[, -(17:19)]
+test_df2 = data.frame(apply(test_df2, 2, scale))
+# Cleaning Test Set for Model #3
+test_df3 = cbind(test_df1, sapply(test_df1[1:length(test_df1)], function(x) x^2))
+names(test_df3) = make.unique(names(test_df3), sep = "_")
+```
+
+```{r predictions}
+
+############# predict target wins
+
+# outlier removal - TEAM_PITCHING_SO
+eval_set <- eval_set %>% na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
+  bind_cols(eval_set) %>%
+  filter(TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO, probs = c(0.25), na.rm = TRUE) 
+         - 1.5 * iqr, 
+         TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, probs = c(0.75), na.rm = TRUE) 
+         + 1.5 * iqr) %>%
+  dplyr::select(-iqr)
+
+# outlier removal - TEAM_PITCHING_H
+eval_set <- eval_set %>% na.omit() %>%
+  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
+  bind_cols(eval_set) %>%
+  filter(TEAM_PITCHING_H > quantile(TEAM_PITCHING_H, probs = c(0.25), na.rm = TRUE) 
+         - 1.5 * iqr,
+         TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, probs = c(0.75), na.rm = TRUE) 
+         + 1.5 * iqr) %>%
+  dplyr::select(-iqr)
+
+# mice imputation
+eval_set <- complete(mice(data = eval_set, method = "pmm", seed = 9450, print=FALSE), 3)
+
+# feature engineering
+eval_set <- eval_set %>% dplyr::select(-INDEX) %>%
+  mutate_if(is.integer, as.numeric) %>%
+  mutate(TEAM_BATTING_1B = TEAM_BATTING_H - (TEAM_BATTING_2B + TEAM_BATTING_3B +
+                                               TEAM_BATTING_HR))
+eval_set <- eval_set %>%
+  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
+         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
+                  4 * TEAM_BATTING_HR) / AB)
+
+# final predictions
+predictions <- round(predict(lm_1, newdata = eval_set))
+summary(predictions)
+predictions
+```
+
+# REFERENCES
+
+ * Burnham, Kenneth P., and David R. Anderson. 2002. Model Selection and
+ Multimodel Inference: A Practical Information-Theoretic Approach. Second.
+ New York: Springer Science+Business Media, Inc.
+
+# CODE APPENDIX
+The code chunks below represent the R code called in order during the analysis.
+They are reproduced in the appendix for review and comment.
+```{r appendix, include=FALSE}
+knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
+```
+```{r setup}
+```
+```{r loadData}
+```
+```{r wins}
+```
+```{r pctNull}
+```
+```{r negFactors}
+```
+```{r posFactors}
+```
+```{r summaryInfo}
+```
+```{r fewOutliers}
+```
+```{r manyOutliers}
+```
+```{r otherOutliers}
+```
+```{r corrPlot}
+```
+```{r outliers_1}
+```
+```{r iqr_removal_SO}
+```
+```{r iqr_removal_H}
+```
+```{r outlier_boxplot}
+```
+```{r impute_train}
+```
+```{r singles}
+```
+```{r make_slg_obp}
+```
+```{r slugging_compare}
+```
+```{r baseModel}
+```
+```{r cooks_distance}
+```
+```{r model2C+S}
+```
+```{r model2}
+```
+```{r model3}
+```
+```{r performance}
+```
+<!--```{r eval_set}
+```-->
+```{r predictions}
+```
+>>>>>>> 84033f5f2f8e65ba0e7413efc0a539e933834f6f:HW1/Drafts/HW1_draft_kellogg.Rmd

### 2020-09-22 HW2/HW2-Group2.rmd

@@
 
-```{r}
+```{r dataImport}
 library(tidyverse)
+library(pkr)
+library(ggplot2)
+library(caret)
 
+# download data
 urlRemote  = "https://raw.githubusercontent.com/"
@@
 
-* class: the actual class for the observation  
-* scored.class: the predicted class for the observation (based on a threshold of 0.5)  
-* scored.probability: the predicted probability of success for the observation
+* **class**: the actual class for the observation  
+* **scored.class**: the predicted class for the observation (based on a threshold of 0.5)  
+* **scored.probability**: the predicted probability of success for the observation
+
+Use the table() function to get the raw confusion matrix for this scored dataset. Make sure you understand the output. In particular, do the rows represent the actual or predicted class? The columns?
 
 ```{r data_table}
+
 class_data = data %>% select(class, pred_class = scored.class, pred_prob = scored.probability)
-class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
-table(class_data[,1:2]) %>% knitr::kable(caption = "Counts of the Classification Data")
+
+# raw confusion matrix
+table("Predicted"= data$scored.class, "Actual"= data$class)
+
+#class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
+#table(class_data[,1:2]) %>% knitr::kable(caption = "Counts of the Classification Data")
 ```
+
+**Rows represent the predicted value and columns represent the actual value.**
+
+* 119 records were predicted to be 0 that were actually 0. (True negative)
+* 30 records were predicted to be 0 that were actually 1. (False negative)
+* 5 records were predicted to be 1 that were actually 0. (False positive)
+* 27 records were predicted to be 1 that were actually 1. (True positive)
 
@@
 
-* Accuracy: the proportion of the total number of predictions that were correct.  
-* Positive Predictive Value or Precision: the proportion of positive cases that were correctly identified.  
-* Negative Predictive Value: the proportion of negative cases that were correctly identified.  
-* Sensitivity or Recall: the proportion of actual positive cases which are correctly identified.   
+* **Accuracy**: the proportion of the total number of predictions that were correct.  
+* **Positive Predictive Value or Precision**: the proportion of positive cases that were correctly identified.  
+* **Negative Predictive Value**: the proportion of negative cases that were correctly identified.  
+* **Sensitivity or Recall**: the proportion of actual positive cases which are correctly identified.   
 * Specificity: the proportion of actual negative cases which are correctly identified.
@@
 
-```{r accuracy_check}
-sprintf("The accuracy of the predictions for the classification data set is %1.3f%%.", accuracy(class_data)*100)
-```
-
 ### Error Rate of the Predictions
@@
 
-```{r error_check}
-sprintf("The error rate of the predictions for the classification data set is %0.3f.", error(class_data))
-
-sprintf("The accuracy and error rate of the predictions must sum to 1. For the classification data set, there summation equals to %.3f.", accuracy(class_data) + error(class_data))
-```
-
 ### Precision of the Predictions
@@
 }
-```
-
-```{r precision_check}
-sprintf("The precision of the predictions for the classification data set is %0.3f.", 
-precision(class_data))
 ```
@@
 
-```{r sensitivity_check}
-sprintf("The sensitivity of the predictions for the classification data set is %0.3f.", 
-sensitivity(class_data))
-```
-
 ### Specificity of the Predictions
@@
 
-```{r specificity_check}
-sprintf("The specificity of the predictions for the classification data set is %0.3f.", 
-specificity(class_data))
-```
-
 ### F1 Score of the Predictions
@@
 
+**What are the bounds on the F1 score? Show that the F1 score will always be between 0 and 1. (Hint: If 0 < 𝑎 < 1 and 0 < 𝑏 < 1 then 𝑎𝑏 < 𝑎.)**
+
+We know that both precision and sensitivity must be between 0 and 1.
+
+* If these metrics are 0, it means that no true positives are identified. 
+* If precision = 1, this means that there are no false positives. In other words, we didn't identify any records as being 1 that are actually 0. 
+* If sensitivity = 1, this means that there are no false negatives. In other words, we didn't identify any records as 0 that are actually 1.
+
+We know that the maximum possible value of the F1 score occurs when both precision and sensitivity are 1. If we plug these values into our F1 score equation, we can see that the upper bound is 1:
+$F1\ score = \frac{2*1*1}{1+1} = \frac{2}{2} = 1$
+
+Similarly, we know that the minimum possible value of the F1 score occurs when either precision or sensitivity equals 0 because the numerator in the equation ends up being 0.
+We can see that our lower bound is therefore 0:
+$F1\ score = \frac{2*0*1}{1+0} = \frac{0}{1} = 0$
+
+Therefore, the bounds for an F1 score are [0,1].
+
+### ROC Curve
+Function `roc_score` that generates an ROC curve from a data set with a true classification column (class in our example) and a probability column (scored.probability in our example).
+Your function should return a list that includes the plot of the ROC curve and a vector that contains the calculated area under the curve (AUC).
+Note that I recommend using a sequence of thresholds ranging from 0 to 1 at 0.01 intervals.
+
+```{r roc, echo=TRUE}
+# function for roc score 
+roc_score <- function(df){
+  if (is.data.frame(df) == FALSE) {
+    return("Data set is not a dataframe. Try as.data.frame(dataset).")
+  } else {
+
+    seqOver <- seq(0,1,0.01)
+    finalFrame <- data.frame(threshold=double(), 
+                           tpr=double(),
+                           fpr=double(),
+                           stringsAsFactors=FALSE)
+
+    for (i in seqOver){
+      df2 <- df %>%
+        mutate(tp = ifelse(class == 1 & pred_class == 1, 1, 0),
+               tn = ifelse(class == 0 & pred_class == 0, 1, 0),
+               fp = ifelse(class == 0 & pred_class == 1, 1, 0),
+               fn = ifelse(class == 1 & pred_class == 0, 1, 0)) %>%
+        filter(pred_prob <= i)
+
+      if (nrow(df2) > 0){
+        tpr <- sum(df2$tp)/ (sum(df2$tp)+sum(df2$fn))
+        fpr <- sum(df2$fp)/ (sum(df2$fp)+sum(df2$tn))
+        finalFrame <- rbind(finalFrame,data.frame(i, tpr, fpr))
+      }
+    }
+
+    roc_plt <- ggplot(data=finalFrame, aes(x=fpr, y=tpr)) +
+      geom_line() +
+      ggtitle("ROC Curve") +
+      xlab("False Positive Rate") +
+      ylab("True Positive Rate") +
+      geom_abline(intercept = 0, slope = 1)
+
+    return(roc_plt)
+  }
+
+}
+
+roc_final <- roc_score(class_data)
+roc_final$plot_env$roc_plt
+
+```
+
+### Data Metrics
+
+```{r accuracy_check}
+sprintf("The accuracy of the predictions for the classification data set is %1.3f%%.", accuracy(class_data)*100)
+```
+
+```{r error_check}
+sprintf("The error rate of the predictions for the classification data set is %0.3f.", error(class_data))
+
+sprintf("The accuracy and error rate of the predictions must sum to 1. For the classification data set, there summation equals to %.3f.", accuracy(class_data) + error(class_data))
+```
+
+```{r precision_check}
+sprintf("The precision of the predictions for the classification data set is %0.3f.", 
+precision(class_data))
+```
+
+```{r sensitivity_check}
+sprintf("The sensitivity of the predictions for the classification data set is %0.3f.", 
+sensitivity(class_data))
+```
+
+```{r specificity_check}
+sprintf("The specificity of the predictions for the classification data set is %0.3f.", 
+specificity(class_data))
+```
+
 ```{r f1_score_check}
@@
 ```
+
+### Caret Package
+
+```{rpROC Package

### 2020-09-24 HW2/HW2-Group2.rmd
adding ROC function
adding in ROC curve
Review & copyedit

--- 2020-09-22 HW2/HW2-Group2.rmd
@@
 
+# 1. The Data
 ```{r}
 library(tidyverse)
+# urlRemote  = "https://raw.githubusercontent.com/"
+# pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
+# fileTrain   = "classification-output-data.csv"
+# data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
 
-urlRemote  = "https://raw.githubusercontent.com/"
-pathGithub = "aadler/DT621_Fall2020_Group2/master/HW2/"
-fileTrain   = "classification-output-data.csv"
-data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
+data <- read.csv("./classification-output-data.csv")
 ```
 
-## The Dataset
+The data set contains `r dim(data)[[1]]` data points, and has three key target
+columns:
 
-The data set contains n = 181 data points, and has three key columns that will be used:
+  * **class**: the actual class for the observation  
+  * **scored.class**: the predicted class for the observation (based on a
+  threshold of 0.5)  
+  * **scored.probability**: the predicted probability of success for the
+  observation
 
-* class: the actual class for the observation  
-* scored.class: the predicted class for the observation (based on a threshold of 0.5)  
-* scored.probability: the predicted probability of success for the observation
-
-```{r data_table}
-class_data = data %>% select(class, pred_class = scored.class, pred_prob = scored.probability) 
-class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
-table(class_data[,1:2]) %>% knitr::kable(caption = "Counts of the Classification Data")
-```
-
-## Confusion Matrix & Functions
-
-A [confusion matrix](http://www.saedsayad.com/model_evaluation_c.htm) is a technique for summarizing the performance of a classification algorithm.
-It shows the number of correct and incorrect predictions made by the classification model compared to the actual outcomes (target value) in the data.
-The matrix is $N \times N$, where N is the number of target values (classes).
-The following table displays a 2x2 confusion matrix for two classes (Positive and Negative).
+# 2. Confusion Matrix
+A [confusion matrix](http://www.saedsayad.com/model_evaluation_c.htm) is a
+format for summarizing the performance of a classification algorithm. It shows
+the number of correct and incorrect predictions made by the classification model
+and compares it to the actual outcomes in the data. The matrix is
+\(N \times N\), where \(N\) is the number of target values. The following table
+displays a \(2\times 2\) confusion matrix for two classes---positive and
+negative. The actual confusion matrix are the 4 cells containing `a`, `b`, `c`,
+and `d`. The enhancments around the matrix display the definitions for many
+common classification metrics.
 
@@
 -----------------|-------------|-------------|------------------------------
-      | Positive |      a      |      b      | Positive Predictive | a/(a+b)
+      | Positive |a (True Pos) |b (False Pos)| Positive Predictive | a/(a+b)
 Model |----------|-------------|-------------|------------------------------
-      | Negative |      c      |      d      | Negative Predictive | d/(c+d)
+      | Negative |c (False Neg)|d (True Neg) | Negative Predictive | d/(c+d)
 -----------------|-------------|-------------|------------------------------
@@
 
-* Accuracy: the proportion of the total number of predictions that were correct.  
-* Positive Predictive Value or Precision: the proportion of positive cases that were correctly identified.  
-* Negative Predictive Value: the proportion of negative cases that were correctly identified.  
-* Sensitivity or Recall: the proportion of actual positive cases which are correctly identified.   
-* Specificity: the proportion of actual negative cases which are correctly identified.
+  * **Accuracy**: the proportion of the total number of predictions that were
+  correct.
+  * **Positive Predictive Value** or **Precision**: the proportion of modeled
+  positive cases that were correct.
+  * **Negative Predictive Value**: the proportion of modeled negative cases that
+  were correct.
+  * **Sensitivity** or **Recall**: the proportion of actual positive cases that
+  were properly detected by the model.
+  * **Specificity**: the proportion of actual negative cases that were
+  properly detected by the model.
+
+The raw confusion matrix 
+```{r data_table}
+class_data = data %>% select(class, pred_class = scored.class,
+                             pred_prob = scored.probability) 
+class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
+```
 
 ### Accuracy of the Predictions
+The function, `accuracy` below takes as input a data frame whose first two
+columns are the true class and the modeled class respectively. It returns the
+accuracy of the predictions using the following formula:
 
-Function `accuracy` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the accuracy of the predictions. The formula is:
-
-\[Accuracy = \frac{TP + TN}{TP + FP + TN + FN}\]
+\[
+\textrm{Accuracy} = \frac{TP + TN}{TP + FP + TN + FN}
+\]
 
 ```{r accuracy, echo = TRUE}
-accuracy = function(df) {
-  if (is.data.frame(df) == FALSE) {
-    return("Data set is not a dataframe. Try as.data.frame(dataset).")
+accuracy <- function(df) {
+  if (!is.data.frame(df)) {
+    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
   } else {
-    dt = as.data.frame(table(df[, 1:2]))
-    TN = dt %>% filter(class == 0 & pred_class == 0) %>% select(Freq)
-    TP = dt %>% filter(class == 1 & pred_class == 1) %>% select(Freq)
-    return((TP + TN) / nrow(df))
+    dt = table(df[, 1:2])
+    sum(diag(dt)) / sum(dt)
   }
@@
 ```{r accuracy_check}
-sprintf("The accuracy of the predictions for the classification data set is %1.3f%%.", accuracy(class_data)*100)
+sprintf("The accuracy of the predictions for the classification data set is %1.3f%%.",
+        accuracy(class_data) * 100)
 ```
@@
 ### Error Rate of the Predictions
-
 Function `error` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the classification error rate of the predictions. The formula is:

### 2020-09-24 HW2/HW2-Group2.rmd
Finished 1–4 starting 5

@@
 ---
-title: 'DATA 621 - Business Analytics and Data Mining'
-subtitle: 'Fall 2020 - Group 2 - Homework #2'
-author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
-    Bryan Persaud, Jeff Shamp
+title: "DATA 621 - Business Analytics and Data Mining"
+author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, Bryan Persaud,
+  Jeff Shamp
 date: "10/11/2020"
@@
   html_document: default
+  pdf_document: default
+subtitle: 'Fall 2020 - Group 2 - Homework #2'
 ---
@@
 
-The raw confusion matrix 
 ```{r data_table}
@@
                              pred_prob = scored.probability) 
-class_data %>% head() %>% knitr::kable(caption = "Preview of the Dataset")
+table(class_data[,1:2]) %>% knitr::kable(caption = "Raw Confusion Matrix")
 ```
 
-### Accuracy of the Predictions
-The function, `accuracy` below takes as input a data frame whose first two
-columns are the true class and the modeled class respectively. It returns the
-accuracy of the predictions using the following formula:
+The raw confusion matrix is shown above. In this case, the rows represent the 
+**true** classes and the columns represent the **predicted** classes. This is
+not the same order as the hand-drawn confusion matrix above, but both are
+internally consistent.
+
+# 3. Accuracy
+The function `ACC` below takes as input a data frame whose first two columns are
+the true class and the modeled class respectively. It returns the accuracy of
+the predictions using the following formula:
 
@@
 ```{r accuracy, echo = TRUE}
-accuracy <- function(df) {
+ACC <- function(df) {
   if (!is.data.frame(df)) {
@@
   } else {
-    dt = table(df[, 1:2])
+    dt <- table(df[, 1:2])
     sum(diag(dt)) / sum(dt)
@@
 
-```{r accuracy_check}
-sprintf("The accuracy of the predictions for the classification data set is %1.3f%%.",
-        accuracy(class_data) * 100)
-```
+For this dataset the accuracy is 
+`r prettyNum(ACC(class_data) * 100, digits = 5L)`%.
 
-### Error Rate of the Predictions
-Function `error` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the classification error rate of the predictions. The formula is:
+# 4. Classification Error Rate
+The function `CER` below takes as input a data frame whose first two columns
+are the true class and the modeled class respectively. It returns the
+classification error rate of the predictions using the following formula:
 
-\[\text{Classification Error Rate} = \frac{FP + FN}{TP + FP + TN + FN}\]
+\[
+\textrm{Classification Error Rate} = \frac{FP + FN}{TP + FP + TN + FN}
+\]
 
 ```{r error, echo = TRUE}
-error = function(df) {
-  if (is.data.frame(df) == FALSE) {
-    return("Data set is not a dataframe. Try as.data.frame(dataset).")
+CER <- function(df) {
+  if (!is.data.frame(df)) {
+    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
   } else {
-    dt = as.data.frame(table(df[, 1:2]))
-    FP = dt %>% filter(class == 0 & pred_class == 1) %>% select(Freq)
-    FN = dt %>% filter(class == 1 & pred_class == 0) %>% select(Freq)
-    return((FP + FN) / nrow(df))
+    dt <- table(df[, 1:2])
+    (sum(dt) - sum(diag(dt))) / sum(dt)
   }
@@
 
-```{r error_check}
-sprintf("The error rate of the predictions for the classification data set is %0.3f.", error(class_data))
+For this dataset the classification error rate is
+`r prettyNum(CER(class_data) * 100, digits = 5L)`%.
 
-sprintf("The accuracy and error rate of the predictions must sum to 1. For the classification data set, there summation equals to %.3f.", accuracy(class_data) + error(class_data))
-```
+Mathematically, the classification error rate must be identical to
+\(1 - \textrm{Accuracy}\). In this data set that is clear as 
+`r prettyNum(ACC(class_data) * 100, digits = 5L)`% +
+`r prettyNum(CER(class_data) * 100, digits = 5L)`% = 
+`r prettyNum((CER(class_data) + ACC(class_data)) * 100, digits = 5L)`%.
 
-### Precision of the Predictions
+# 5. Precision
+The function `PRC` below takes as input a data frame whose first two columns
+are the true class and the modeled class respectively. It returns the
+classification error rate of the predictions using the following formula:
 
-Function `precision` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the precision of the predictions. The formula is:
-
-\[precision = \frac{TP}{TP + FP}\]
+\[
+\textrm{Precision} = \frac{TP}{TP + FP}
+\]
 

### 2020-09-24 HW2/HW2-Group2.rmd
Finished through 7

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
+knitr::opts_chunk$set(echo=TRUE, error=FALSE, warning=FALSE, message=FALSE)
 ```
 
+# 0. Preamble
+***For all the following answers, it is expected that we are only dealing with***
+***binary classification in which negative is coded as 0 and positive is coded***
+***as 1. It is not difficult to expand the functions to respond to more general***
+***cases, but it is an unnecessary complication for this homework assignment.***
+
 # 1. The Data
-```{r}
+```{r loadData, echo=FALSE}
 library(tidyverse)
@@
 
-```{r}
+```{r confMatrixDesc, echo=FALSE}
 cat("----------------------------------------------------------------------------
@@
 
-```{r data_table}
+```{r dataCM, echo=FALSE}
 class_data = data %>% select(class, pred_class = scored.class,
@@
 
-```{r accuracy, echo = TRUE}
+```{r accuracy}
 ACC <- function(df) {
@@
 
-```{r error, echo = TRUE}
+```{r error}
 CER <- function(df) {
@@
 
-```{r precision, echo = TRUE}
-precision = function(df) {
-  if (is.data.frame(df) == FALSE) {
-    return("Data set is not a dataframe. Try as.data.frame(dataset).")
+```{r precision}
+PRC <- function(df) {
+  if (!is.data.frame(df)) {
+    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
   } else {
-    dt = as.data.frame(table(df[, 1:2]))
-    FP = dt %>% filter(class == 0 & pred_class == 1) %>% select(Freq)
-    TP = dt %>% filter(class == 1 & pred_class == 1) %>% select(Freq)
-    return(TP / (TP + FP))
+    dt <- table(df[, 1:2])
+    dt[2, 2] / sum(dt[, 2])
   }
@@
 
-```{r precision_check}
-sprintf("The precision of the predictions for the classification data set is %0.3f.", 
-precision(class_data))
-```
+For this dataset the precision is
+`r prettyNum(PRC(class_data) * 100, digits = 5L)`%.
 
-### Sensitivity of the Predictions
+# 6. Sensitivity or Recall
+The function `SNS` below takes as input a data frame whose first two columns
+are the true class and the modeled class respectively. It returns the
+sensitivity of the predictions using the following formula:
 
-Function `sensitivity` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the sensitivity of the predictions.
-Sensitivity is also known as recall.
-The formula is:
+\[
+\textrm{Sensitivity} = \frac{TP}{TP + FN}
+\]
 
-\[sensitivity = \frac{TP}{TP + FN}\]
-
-```{r sensitivity, echo = TRUE}
-sensitivity = function(df) {
-  if (is.data.frame(df) == FALSE) {
-    return("Data set is not a dataframe. Try as.data.frame(dataset).")
+```{r sensitivity}
+SNS <- function(df) {
+  if (!is.data.frame(df)) {
+    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
   } else {
-    dt = as.data.frame(table(df[, 1:2]))
-    FN = dt %>% filter(class == 1 & pred_class == 0) %>% select(Freq)
-    TP = dt %>% filter(class == 1 & pred_class == 1) %>% select(Freq)
-    return(TP / (TP + FN))
+    dt <- table(df[, 1:2])
+    dt[2, 2] / sum(dt[2, ])
   }
@@
 
-```{r sensitivity_check}
-sprintf("The sensitivity of the predictions for the classification data set is %0.3f.", 
-sensitivity(class_data))
-```
+For this dataset the sensitivity is
+`r prettyNum(SNS(class_data) * 100, digits = 5L)`%.
 
-### Specificity of the Predictions
+# 7. Specificity
+The function `SPC` below takes as input a data frame whose first two columns
+are the true class and the modeled class respectively. It returns the
+specificity of the predictions using the following formula:
 
-Function `specificity` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the specificity of the predictions. The formula is:
+\[
+\textrm{Specificity} = \frac{TN}{TN + FP}
+\]
 
-\[specificity = \frac{TN}{TN + FP}\]
-
-```{r specificity, echo = TRUE}
-specificity = function(df) {
-  if (is.data.frame(df) == FALSE) {
-    return("Data set is not a dataframe. Try as.data.frame(dataset).")
+```{r specificity}
+SPC <- function(df) {
+  if (!is.data.frame(df)) {
+    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
   } else {
-    dt = as.data.frame(table(df[, 1:2]))
-    TN = dt %>% filter(class == 0 & pred_class == 0) %>% select(Freq)
-    FP = dt %>% filter(class == 0 & pred_class == 1) %>% select(Freq)
-    return(TN / (TN + FP))
+    dt <- table(df[, 1:2])
+    dt[1, 1] / sum(dt[1, ])
   }
@@
 
-```{r specificity_check}
-sprintf("The specificity of the predictions for the classification data set is %0.3f.", 
-specificity(class_data))
-```
+For this dataset the specificity is
+`r prettyNum(SPC(class_data) * 100, digits = 5L)`%.
 
@@
   } else {
-    return((2*precision(df)*sensitivity(df)) / (precision(df) + sensitivity(df)))
+    return((2 * PRC(df) * SNS(df)) / (PRC(df) + SNS(df)))
   }

### 2020-09-24 HW2/HW2-Group2.rmd
Finished through 8

@@
 
-### F1 Score of the Predictions
+# 8. \(F_1\) Score
+The function `F1` below takes as input a data frame whose first two columns
+are the true class and the modeled class respectively. It returns the
+\(F_1\) score of the predictions using the following formula:
 
-Function `f1_score` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the F1 score of the predictions. The formula is:
+\[
+\textrm{F}_1\;\textrm{Score} = \frac{2 \times \textrm{Precision} \times
+\textrm{Sensitivity}}{\textrm{Precision} + \textrm{Sensitivity}}
+\]
 
-\[\text{F1 Score} = \frac{2 \times Precision \times Sensitivity}{Precision + Sensitivity}\]
-
-```{r f1_score, echo = TRUE}
-f1_score = function(df) {
-  if (is.data.frame(df) == FALSE) {
-    return("Data set is not a dataframe. Try as.data.frame(dataset).")
+```{r f1}
+F1 <- function(df) {
+  if (!is.data.frame(df)) {
+    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
   } else {
-    return((2 * PRC(df) * SNS(df)) / (PRC(df) + SNS(df)))
+    (2 * PRC(df) * SNS(df)) / (PRC(df) + SNS(df))
   }
@@
 
-```{r f1_score_check}
-sprintf("The F1 score of the predictions for the classification data set is %0.3f.", 
-f1_score(class_data))
-```
+For this dataset the \(F_1\) score is
+`r prettyNum(F1(class_data) * 100, digits = 5L)`%.

### 2020-09-24 HW2/HW2-Group2.rmd
Part 9

@@
 negative. The actual confusion matrix are the 4 cells containing `a`, `b`, `c`,
-and `d`. The enhancments around the matrix display the definitions for many
+and `d`. The enhancements around the matrix display the definitions for many
 common classification metrics.
@@
 `r prettyNum(F1(class_data) * 100, digits = 5L)`%.
+
+# 9. Bounds on \(F_1\)
+## Proof 1
+We can prove this by decomposing the metrics into their component parts as
+defined above.
+\[
+\begin{align}
+F_{1:numerator} &= 2 \cdot P \cdot S\\
+&= 2 \cdot \frac{TP}{TP + FP} \cdot \frac{TP}{TP + FN}\\
+&= \frac{2\cdot(TP)^2}{(TP+FP)(TP+FN)}\\
+F_{1:denominator} &= P + S\\
+&= \frac{TP}{TP + FP} + \frac{TP}{TP + FN}\\
+&= \frac{TP(TP + FN) + TP(TP+FP)}{(TP+FP)(TP+FN)}\\
+&= \frac{(TP)^2 + TP\cdot FN + (TP)^2 + TP\cdot FP}{(TP+FP)(TP+FN)}\\
+&= \frac{2\cdot(TP)^2 + TP\cdot (FN + FP)}{(TP+FP)(TP+FN)}\\\
+F_1 &= \frac{F_{1:numerator}}{F_{1:denominator}}\\
+&= \frac{2\cdot(TP)^2}{2\cdot(TP)^2 + TP\cdot (FN + FP)}\\
+&= \frac{2\cdot TP}{2\cdot TP + FN + FP}
+\end{align}
+\]
+
+Now this last equation ***must*** be bounded between 0 and 1. Consider that TP
+is a non-negative integer. Also, the sum of TP, FP, TN, and FN must equal the
+number of observations. Excluding the degenerate case where the dataset contains
+only negative classes and they were all scored properly, the denominator will
+always be positive. When the dataset consists of only positive classes and they
+are all coded properly, we have \(F_1 = 1\). This is also why there isn't too
+much to worry about with the all-negative case, as one can simply switch the
+labels and side-step the division by zero error. In every other case where TP
+is strictly less than the number of observations, there must be entries in
+FN or FP. This causes the fraction to be less than one. In the case where no
+positive entries are coded properly, then the numerator is 0 and the denominator
+is > 0, so the \(F_1\) score is 0. As the building blocks all represent counts,
+they cannot be negative, so the score can never fall below 0. ***QED***
+
+## Proof 2
+Another proof relies on the assumption that the sensitivity and precision
+themselves are constrained to \([0, 1]\). Given those assumptions, it ***must***
+be true that \(2ab < a + b\). This is a corollary of the
+[*AM-GM inequality*](https://en.wikipedia.org/wiki/Inequality_of_arithmetic_and_geometric_means)
+since if \(a, b \in [0, 1]\) then \(\sqrt{ab} > ab\) and this inequality is
+strict. Thus the numerator must be strictly less than the denominator. Both
+numerator and denominator are bounded by \([0, 1]\) as per the initial
+assumptions. Therefore, excepting the degenerate case where both are 0,
+this completed the proof the fraction too is bounded by \([0, 1]\). ***QED***

### 2020-09-24 HW2/HW2-Group2.rmd
Had to refactor all the functions not to use table to handle the degenerate case where the predictions were all 0s or 1s, so the table was not 2x2 but 2x1. Noe ROC should work

@@
 ```{r loadData, echo=FALSE}
-library(tidyverse)
 # urlRemote  = "https://raw.githubusercontent.com/"
@@
 ```{r dataCM, echo=FALSE}
-class_data = data %>% select(class, pred_class = scored.class,
-                             pred_prob = scored.probability) 
-table(class_data[,1:2]) %>% knitr::kable(caption = "Raw Confusion Matrix")
+class_data <-  data.frame(class = data$class,
+                          pred_class = data$scored.class,
+                          pred_prob = data$scored.probability)
+knitr::kable(table(class_data[,1:2]), caption = "Raw Confusion Matrix")
 ```
@@
 The function `ACC` below takes as input a data frame whose first two columns are
-the true class and the modeled class respectively. It returns the accuracy of
-the predictions using the following formula:
+the true class and the modeled class and named `class` and `pred_class`
+respectively. It returns the accuracy of the predictions using the following
+formula:
 
@@
   } else {
-    dt <- table(df[, 1:2])
-    sum(diag(dt)) / sum(dt)
+    sum(df$class == df$pred_class) / dim(df)[[1]]
   }
@@
 The function `CER` below takes as input a data frame whose first two columns
-are the true class and the modeled class respectively. It returns the
+are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
 classification error rate of the predictions using the following formula:
@@
   } else {
-    dt <- table(df[, 1:2])
-    (sum(dt) - sum(diag(dt))) / sum(dt)
+    sum(df$class != df$pred_class) / dim(df)[[1]]
   }
@@
 The function `PRC` below takes as input a data frame whose first two columns
-are the true class and the modeled class respectively. It returns the
+are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
 classification error rate of the predictions using the following formula:
@@
   } else {
-    dt <- table(df[, 1:2])
-    dt[2, 2] / sum(dt[, 2])
+    sum(df$class == 1 & df$pred_class == 1) / sum(df$pred_class == 1)
   }
@@
 The function `SNS` below takes as input a data frame whose first two columns
-are the true class and the modeled class respectively. It returns the
+are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
 sensitivity of the predictions using the following formula:
@@
   } else {
-    dt <- table(df[, 1:2])
-    dt[2, 2] / sum(dt[2, ])
+    sum(df$class == 1 & df$pred_class == 1) / sum(df$class == 1)
   }
@@
 The function `SPC` below takes as input a data frame whose first two columns
-are the true class and the modeled class respectively. It returns the
+are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
 specificity of the predictions using the following formula:
@@
   } else {
-    dt <- table(df[, 1:2])
-    dt[1, 1] / sum(dt[1, ])
+    sum(df$class == 0 & df$pred_class == 0) / sum(df$class == 0)
   }
@@
 The function `F1` below takes as input a data frame whose first two columns
-are the true class and the modeled class respectively. It returns the
+are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
 \(F_1\) score of the predictions using the following formula:
@@
 this completed the proof the fraction too is bounded by \([0, 1]\). ***QED***
+
+10. ROC Curve
+The ROC curve is defined as a plot of the true positive rate against the false
+positive rate at different probability thresholds. The false positive rate is
+usually plotted on the x-axis.
+
+```{r ROC}
+ROC <- function(df) {
+  if (!is.data.frame(df)) {
+    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
+  } else {
+    rocTable <- data.table(thresh = seq(0, 1, 0.01),
+                           x = double(101),
+                           y = double(101))
+    for (i in seq_along(rocTable$thresh)) {
+      dftmp <- data.frame(class = df$class,
+                          pred_class = ifelse(df$pred_prob > rocTable$thresh[i],
+                                              1, 0))
+      rocTable$x[i] <- 1 - SPC(dftmp)
+      rocTable$y[i] <- SNS(dftmp)
+    }
+    plot(rocTable$x, rocTable$y, type = 'l')
+  }
+}
+```

### 2020-09-24 HW2/HW2-Group2.rmd
Copyedit and expand section 10 prose

@@
 The function `CER` below takes as input a data frame whose first two columns
-are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
-classification error rate of the predictions using the following formula:
+are the true class and the modeled class and named `class` and `pred_class`
+respectively. It returns the classification error rate of the predictions using
+the following formula:
 
@@
 The function `PRC` below takes as input a data frame whose first two columns
-are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
-classification error rate of the predictions using the following formula:
+are the true class and the modeled class and named `class` and `pred_class`
+respectively. It returns the classification error rate of the predictions using
+the following formula:
 
@@
 The function `SNS` below takes as input a data frame whose first two columns
-are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
-sensitivity of the predictions using the following formula:
+are the true class and the modeled class and named `class` and `pred_class`
+respectively. It returns the sensitivity of the predictions using the following
+formula:
 
@@
 The function `SPC` below takes as input a data frame whose first two columns
-are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
-specificity of the predictions using the following formula:
+are the true class and the modeled class and named `class` and `pred_class`
+respectively. It returns the specificity of the predictions using the following
+formula:
 
@@
 The function `F1` below takes as input a data frame whose first two columns
-are the true class and the modeled class and named `class` and `pred_class` respectively. It returns the
-\(F_1\) score of the predictions using the following formula:
+are the true class and the modeled class and named `class` and `pred_class`
+respectively. It returns the \(F_1\) score of the predictions using the
+following formula:
 
@@
 10. ROC Curve
-The ROC curve is defined as a plot of the true positive rate against the false
-positive rate at different probability thresholds. The false positive rate is
-usually plotted on the x-axis.
+The receiver operating curve (ROC) is defined as a plot of the true positive
+rate (TPR) against the false positive rate (FPR) at different probability
+thresholds. The FPR is usually plotted on the x-axis. The function below takes
+as input a data frame with columns named `class` and `pred_class` as above, and
+a third column named `pred_prob` which reflects the predicted probability of
+being positive. It then steps along a sequence from 0 to 1, assigning classes
+based on that sequence, calculating the TPR and FPR for each threshold, plotting
+the results when completed. Lastly, it calculates the area under the ROC (AUC).
 
@@
     }
-    plot(rocTable$x, rocTable$y, type = 'l')
+    plot(rocTable$x, rocTable$y, type = 'l', xlab = "False Positive Rate",
+         ylab = 'True Positive Rate')
   }

### 2020-09-24 HW2/HW2-Group2.rmd
Finished section 10

@@
 the results when completed. Lastly, it calculates the area under the ROC (AUC).
+Trapezoidal quadrature will be used to estimate the AUC from the table created
+for the ROC, as the intervals on the x axis are not equidistant.
 
@@
   } else {
-    rocTable <- data.table(thresh = seq(0, 1, 0.01),
+    rocTable <- data.frame(thresh = seq(0, 1, 0.01),
                            x = double(101),
@@
     }
-    plot(rocTable$x, rocTable$y, type = 'l', xlab = "False Positive Rate",
+  }
+  return(list(plot = {plot(rocTable$x, rocTable$y, type = 'l',
+                          xlab = "False Positive Rate",
                           ylab = 'True Positive Rate')
-  }
+    lines(c(0, 1), c(0, 1), lty = 2)},
+              AUC = sum((rocTable$y[1:100] + rocTable$y[2:101]) * 0.5 *
+                          (rocTable$x[1:100] - rocTable$x[2:101]))))
 }
 ```
+
+The ROC and AUC for the dataset provided is shown below:
+
+```{r ROCdata, echo=FALSE, fig.height=6}
+ROC(class_data)
+```

### 2020-09-24 HW2/HW2-Group2.rmd
Finished 12

@@
 ```
+
+# 11. Metric Production
+All the metrics were displayed in their sections, but will be reproduced here
+as well.
+
+```{r sec11}
+ACC(class_data)
+CER(class_data)
+PRC(class_data)
+SNS(class_data)
+SPC(class_data)
+F1(class_data)
+ROC(class_data)
+```
+
+# 12. Using the `caret` package
+
+```{r caret}
+library(caret)
+confusionMatrix(as.factor(class_data$pred_class),
+                as.factor(class_data$class),
+                positive = '1')
+# Precision
+posPredValue(as.factor(class_data$pred_class),
+            as.factor(class_data$class),
+            positive = '1')
+# Sensitivity
+sensitivity(as.factor(class_data$pred_class),
+            as.factor(class_data$class),
+            positive = '1')
+# Specificity
+specificity(as.factor(class_data$pred_class),
+            as.factor(class_data$class),
+            negative = '0')
+```
+
+As can be seen above, when the functions are called with proper identification
+of the negative and positive cases, our results match the `caret` package
+results.
+
+# 13.

### 2020-09-24 HW2/HW2-Group2.rmd
Finished the HW

@@
 \[
-\begin{align}
+\begin{aligned}
 F_{1:numerator} &= 2 \cdot P \cdot S\\
@@
 &= \frac{2\cdot TP}{2\cdot TP + FN + FP}
-\end{align}
+\end{aligned}
 \]
@@
 
-10. ROC Curve
+# 10. ROC Curve
 The receiver operating curve (ROC) is defined as a plot of the true positive
@@
   } else {
-    rocTable <- data.frame(thresh = seq(0, 1, 0.01),
-                           x = double(101),
-                           y = double(101))
+    rocTable <- data.frame(thresh = seq(0, 1, 0.001),
+                           x = double(1001),
+                           y = double(1001))
     for (i in seq_along(rocTable$thresh)) {
@@
     lines(c(0, 1), c(0, 1), lty = 2)},
-              AUC = sum((rocTable$y[1:100] + rocTable$y[2:101]) * 0.5 *
-                          (rocTable$x[1:100] - rocTable$x[2:101]))))
+              AUC = sum((rocTable$y[1:1000] + rocTable$y[2:1001]) * 0.5 *
+                          (rocTable$x[1:1000] - rocTable$x[2:1001]))))
 }
@@
 
-```{r ROCdata, echo=FALSE, fig.height=6}
+```{r ROCdata, echo=FALSE, fig.height=6, fig.width=6}
 ROC(class_data)
@@
 F1(class_data)
-ROC(class_data)
 ```
@@
 
-# 13.
+# 13. Using the `pROC` package
+
+```{r pROC, fig.height=6, fig.width=6}
+ROC(class_data)
+library(pROC)
+pROCroc <- roc(class_data, class, pred_prob)
+plot(pROCroc)
+auc(pROCroc)
+```
+
+Visual inspection of the two ROC curves shows extremely similar shapes. The AUC
+scores are also practically identical. Originally, we used the suggested 101
+points between 0 and 1. This led to a difference of less than 1% in the AUC
+scores. Moving to a 1001-point quadrature all but eliminated that difference.

### 2020-09-24 HW2/HW2-Group2.rmd
minor copyediting

@@
 ```{r loadData, echo=FALSE}
-# urlRemote  = "https://raw.githubusercontent.com/"
-# pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
-# fileTrain   = "classification-output-data.csv"
-# data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()
-
-data <- read.csv("./classification-output-data.csv")
+urlRemote  = "https://raw.githubusercontent.com/"
+pathGithub = "aadler/DT621_Fall2020_Group2/master/HW2/"
+fileTrain   = "classification-output-data.csv"
+data = read.csv(paste0(urlRemote, pathGithub, fileTrain))
 ```
@@
 displays a \(2\times 2\) confusion matrix for two classes---positive and
-negative. The actual confusion matrix are the 4 cells containing `a`, `b`, `c`,
+negative. The actual confusion matrix is the 4 cells containing `a`, `b`, `c`,
 and `d`. The enhancements around the matrix display the definitions for many
@@
 
-For this dataset the accuracy is 
+For this dataset, the accuracy is 
 `r prettyNum(ACC(class_data) * 100, digits = 5L)`%.
@@
 
-For this dataset the classification error rate is
+For this dataset, the classification error rate is
 `r prettyNum(CER(class_data) * 100, digits = 5L)`%.
@@
 
-For this dataset the precision is
+For this dataset, the precision is
 `r prettyNum(PRC(class_data) * 100, digits = 5L)`%.
@@
 
-For this dataset the sensitivity is
+For this dataset, the sensitivity is
 `r prettyNum(SNS(class_data) * 100, digits = 5L)`%.
@@
 
-For this dataset the specificity is
+For this dataset, the specificity is
 `r prettyNum(SPC(class_data) * 100, digits = 5L)`%.
@@
 
-For this dataset the \(F_1\) score is
+For this dataset, the \(F_1\) score is
 `r prettyNum(F1(class_data) * 100, digits = 5L)`%.
@@
 ## Proof 1
-We can prove this by decomposing the metrics into their component parts as
+We can prove this by decomposing the metrics into their component as
 defined above.
+
 \[
@@
 
-Now this last equation ***must*** be bounded between 0 and 1. Consider that TP
+Now, this last equation ***must*** be bounded between 0 and 1. Consider that TP
 is a non-negative integer. Also, the sum of TP, FP, TN, and FN must equal the
@@
 since if \(a, b \in [0, 1]\) then \(\sqrt{ab} > ab\) and this inequality is
-strict. Thus the numerator must be strictly less than the denominator. Both
+strict. Thus, the numerator must be strictly less than the denominator. Both
 numerator and denominator are bounded by \([0, 1]\) as per the initial
 assumptions. Therefore, excepting the degenerate case where both are 0,
-this completed the proof the fraction too is bounded by \([0, 1]\). ***QED***
+this completes the proof that the fraction too is bounded by \([0, 1]\). ***QED***
 
@@
 Trapezoidal quadrature will be used to estimate the AUC from the table created
-for the ROC, as the intervals on the x axis are not equidistant.
+for the ROC, as the intervals on the x-axis are not equidistant.
 
@@
 # 11. Metric Production
-All the metrics were displayed in their sections, but will be reproduced here
+All the metrics were displayed in their sections but will be reproduced here
 as well.

### 2020-09-24 HW3/hw3_eda_model_shamp.Rmd
logit model and xgb for comparison

---
title: "HW3 - EDA and Modeling"
author: "Jeff Shamp"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kernlab)
library(vip)
```

## Crime Data

This is a classification model, so the data is labeled as below median crime (0) or above median crime rate (1).

We will first look at the state of the training data.

```{r}
crime_df<- read.csv("/Users/jeffshamp/Downloads/crime-training-data_modified.csv")
glimpse(crime_df)
```

### missing data

let's look for missing data (NA) and zeros. Zero are going to be okay in for a few variables.

```{r}
glimpse(
  crime_df %>%
  summarize_all(funs(sum(is.na(.))/length(.)))
) 
glimpse(
crime_df %>%
  summarize_all(funs(length(which(. ==0)))))
```

No missing data. Great!

```{r}
ggplot(stack(crime_df %>% select(-tax)), aes(x=ind, y-values)) +
  geom_boxplot() + coord_flip()
```

```{r}
crime_df %>%
  ggplot(aes(age, medv, color=chas)) +
  geom_point()
```

```{r}
crime_df %>%
  ggplot(aes(x=tax)) + geom_histogram()
```
probably shouldn't log transform that.

```{r}
ggplot(stack(crime_df %>% select(-c('chas', 'target', 'tax', 'nox', 'rm', 'rad'))), aes(x=values)) +
  geom_histogram() + facet_wrap(~ind)
```

```{r}
ggplot(stack(crime_df %>% select(c('nox', 'rm', 'rad'))), aes(x=values)) + geom_histogram() + facet_wrap(~ind)
```

### Result

This data is pretty good in terms of the required prep before modeling. Maybe some work to remove outliers via cooks distance, or maybe some transformations, but it doesn't look like much is needed.

## Model

As a baseline, let's just run a logistic regression and a vanilla XGB on this data set and compare.

We will split up the training data to perform cv and compare the two models.

```{r, message=FALSE}
library(tidymodels)
library(xgboost)
set.seed(9450)
```

```{r}
crime_df<- crime_df %>% mutate(target = as.factor(target))
data_split <- initial_split(crime_df, strata = target, prop = 0.8)
train_df<- training(data_split)
test_df<- testing(data_split)
```

```{r}
xgb_specs <- 
  boost_tree(
    trees =500
  ) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

logit_specs <- 
  logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

# cls_recipe<-
#   crime_df %>%
#   recipe(target ~ .) %>%
#   step_scale(all_predictors()) %>%
#   step_center(all_predictors())

cls_wf<-
  workflow() %>%
  add_formula(target ~ .) %>%
  #add_recipe(cls_recipe) %>%
  add_model(logit_specs)
```

```{r}
cv_folds<- vfold_cv(train_df,
                    v = 5, repeats = 1)

cls_results<-
  fit_resamples(
    cls_wf,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE)
)
```

```{r}
best_model_auc <- select_best(cls_results, "roc_auc")
```

```{r}
final_logit_auc<-
  finalize_workflow(
    cls_wf,
    best_model_auc
)

final_logit_auc %>%
  last_fit(data_split) %>%
  collect_predictions() %>%
  conf_mat(truth = target, estimate = .pred_class)
final_logit_auc %>%
  last_fit(data_split) %>%
  collect_metrics()
```

```{r}
final_logit_auc %>%
  fit(data = train_df) %>%
  pull_workflow_fit() %>%
  vip(geom = "col", aesthetics = list(fill='red4'))
```

Now repeat with xgb.

```{r}
cls_wf<-
  workflow() %>%
  add_formula(target ~ .) %>%
  #add_recipe(cls_recipe) %>%
  add_model(xgb_specs)

cv_folds<- vfold_cv(train_df,
                    v = 5, repeats = 1)

cls_results<-
  fit_resamples(
    cls_wf,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE)
)
best_model_auc <- select_best(cls_results, "roc_auc")
final_xgb_auc<-
  finalize_workflow(
    cls_wf,
    best_model_auc
)

final_xgb_auc %>%
  last_fit(data_split) %>%
  collect_predictions() %>%
  conf_mat(truth = target, estimate = .pred_class)
```

Ok from a vanilla run of logistic and xgb we have reasonably comparable models.

```{r}
final_xgb_auc %>%
  last_fit(data_split) %>%
  collect_metrics()
```

```{r}
library(vip)

final_xgb_auc %>%
  fit(data = train_df) %>%
  pull_workflow_fit() %>%
  vip(geom = "col", aesthetics = list(fill='blue4'))
```

### 2020-09-25 HW3/ensemble_model_shamp.Rmd
updated model, remove old

---
title: "shamp_ensemble_model"
author: "Jeff Shamp"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(tidymodels)
library(stacks)
library(kernlab)
library(vip)
set.seed(9450)
```

```{r}
crime_df<- read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW3/data/crime-training-data_modified.csv")
crime_df<- as.tbl(crime_df)
```

```{r}
crime_df<-
  crime_df %>% 
  mutate(target = as.factor(target)) %>% 
  mutate_if(is.integer, as.numeric)

data_split <- initial_split(crime_df, 
                            strata = target, 
                            prop = 0.8)
train_df<- training(data_split)
test_df<- testing(data_split)
```

```{r}
cv_folds<- vfold_cv(train_df,
                    v = 5, repeats = 1)

crime_recipe<- 
  recipe(target ~., data=train_df)

crime_wf<- 
  workflow() %>%
  add_recipe(crime_recipe)

crtl_grid<- control_stack_grid()
```

```{r}
logit_specs<- 
  logistic_reg(
    penalty = tune(),
    mixture = tune()
    ) %>%
  set_engine("glm") %>%
  set_mode("classification")

svm_specs<- 
  svm_rbf(
    cost = tune(),
    rbf_sigma = tune(),
    margin = tune()
  ) %>%
  set_mode("classification") %>%
  set_engine("kernlab")
```

```{r}
svm_wf<-
  crime_wf %>%
  add_model(svm_specs)

svm_results<-
  tune_grid(
    svm_wf, 
    resample = cv_folds,
    control = crtl_grid,
    grid = 10,
    # save_pred = TRUE, 
    # save_workflow = FALSE
  )

logit_wf<- 
  crime_wf %>%
  add_model(logit_specs)

logit_results<-
  tune_grid(
    logit_wf,
    resamples = cv_folds,
    control = crtl_grid,
    grid = 10,
    # save_pred = TRUE, 
    # save_workflow = FALSE
)
```

```{r}
crime_model_stack <- 
  stacks() %>%
  add_candidates(logit_results) %>%
  add_candidates(svm_results) %>%
  blend_predictions() %>%
  fit_members()

crime_model_stack
```

We see that the logistic classifier is the top weight in the blended model, with the support vector machine...supporting the classification results.

```{r}
theme_set(theme_bw())
autoplot(crime_model_stack)
```

```{r}
crime_pred <- 
  test_df %>%
  bind_cols(predict(crime_model_stack, ., type = "prob"), 
            predict(crime_model_stack, ., type = "class"))
```

Computing the ROC AUC for the model:

```{r, eval = FALSE}
roc<- yardstick::roc_auc(
  crime_pred,
  truth = target,
  contains(".pred_1")
  )

acc<- yardstick::accuracy(
  crime_pred, 
  truth = target, 
  estimate = .pred_class
)

recall<- yardstick::recall(
  crime_pred,
  truth = target, 
  estimate = .pred_class
)

precise<- yardstick::precision(
  crime_pred,
  truth = target, 
  estimate = .pred_class
)
metrics_df<-
  bind_rows(roc, acc, recall, precise)

crime_pred %>%
  conf_mat(truth = target, estimate = .pred_class)
```

```{r}
metrics_df
```

```{r}
logit_wf %>%
  last_fit(data_split) %>%
  collect_metrics()

logit_wf %>%
  fit(data = train_df) %>%
  pull_workflow_fit() %>%
  vip(geom = "col", aesthetics = list(fill='red4'))
```

```{r}
eval_df<- read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW3/data/crime-evaluation-data_modified.csv")
eval_df<-as.tbl(eval_df)
```

```{r}
eval_pred <- 
  eval_df %>%
  bind_cols(predict(crime_model_stack, ., type = "prob"), 
            predict(crime_model_stack, ., type = "class"))
```

```{r}
bind_cols(
eval_pred %>%
  filter(.pred_class == 0) %>%
  count(name = "negative"),
eval_pred %>%
  filter(.pred_class ==1) %>%
  count(name = "positive")
)
```

### 2020-09-26 HW3/ensemble_model_shamp.Rmd
update ensemble model

@@
 crime_df<- read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW3/data/crime-training-data_modified.csv")
-crime_df<- as.tbl(crime_df)
+crime_df<- as_tibble(crime_df)
 ```
+
+## Model Specifics and Training Scheme
+
+We will split the training data into two sets and perform a cross validation scheme on the models to tune parameters for optimum performance.
+Once we have the best models we will blend the logistic regression with another top model to boost performance.
 
@@
 
+Cross validation and workflow creation.
+
 ```{r}
 cv_folds<- vfold_cv(train_df,
-                    v = 5, repeats = 1)
+                    v = 10, repeats = 1)
 
@@
 
+Specify the models. Using SVM as another model to blend with the logistic.
+
 ```{r}
@@
   set_mode("classification")
+
+# rf_specs<-
+#   rand_forest(
+#     min_n = tune(),
+#     mtry = tune(),
+#     trees = tune()
+#   ) %>%
+#   set_engine("ranger") %>%
+#   set_mode("classification")
 
@@
 ```
+
+Tuning the models for best parameters. The stack function will cherry pick the models that best optimize the blended final model.
 
@@
     grid = 10,
-    # save_pred = TRUE, 
-    # save_workflow = FALSE
+    save_pred = TRUE,
+    save_workflow = FALSE
   )
@@
     grid = 10,
-    # save_pred = TRUE, 
-    # save_workflow = FALSE
+    save_pred = TRUE,
+    save_workflow = FALSE
 )
+
+# rf_wf<-
+#   crime_wf %>%
+#   add_model(rf_specs)
+# 
+# rf_results<-
+#   tune_grid(
+#     rf_wf, 
+#     resamples = cv_folds,
+#     control = crtl_grid, 
+#     grid =10
+#   )
 ```
+
+Blend!
 
@@
 
-Computing the ROC AUC for the model:
+Computing the metrics for the stacked model:
 
-```{r, eval = FALSE}
+```{r}
 roc<- yardstick::roc_auc(
@@
 
+### Sub-Models
+
+The blended model is comprised of two sub-models, the logistic and the SVM. Below are the prediction results for each sub-model.
+
+```{r}
+best_model_metric <- select_best(svm_results, "accuracy")
+finalize_workflow(svm_wf, best_model_metric) %>%
+  last_fit(data_split) %>%
+  collect_metrics()
+```
+
 ```{r}
@@
 ```
+We see that they are both in the low eighties percent for accuracy, and low nineties for roc auc. Together make for a better overall prediction.
+
+## Test Predictions
 

### 2020-10-06 HW3/HW 3_Bryan.Rmd
Add files via upload

---
title: "HW 3_Bryan"
author: "Bryan Persaud"
date: "10/6/2020"
output: html_document
---

```{r}
library(tidyverse)
library(kableExtra)
library(corrplot)
```

```{r}
train_crime <- read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW3/data/crime-training-data_modified.csv")
```

# Data Exploration

```{r}
summary(train_crime) %>%
  kable() %>%
  kable_styling()
```

```{r}
sapply(train_crime, function(x) 
  sum(is.na(x)))
```

There are no missing data in the dataset.

```{r}
correlation <- cor(train_crime)
correlation
```

```{r}
corrplot(correlation, diag = FALSE)
```

A correlation matrix and plot helps to see which variables are more related to each other than others.

### 2020-10-07 HW2/HW2-Group2.rmd
Sidestep singular/plural issue and whitespace

@@
 title: "DATA 621 - Business Analytics and Data Mining"
-author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, Bryan Persaud,
-  Jeff Shamp
+subtitle: 'Fall 2020 - Group 2 - Homework #2'
+author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, Bryan
+  Persaud, Jeff Shamp
 date: "10/11/2020"
 output:
+  pdf_document: default
   html_document: default
-  pdf_document: default
-subtitle: 'Fall 2020 - Group 2 - Homework #2'
 ---
@@
 displays a \(2\times 2\) confusion matrix for two classes---positive and
-negative. The actual confusion matrix is the 4 cells containing `a`, `b`, `c`,
-and `d`. The enhancements around the matrix display the definitions for many
-common classification metrics.
+negative. The actual confusion matrix comprises the 4 cells containing `a`, `b`,
+`c`, and `d`. The enhancements around the matrix display the definitions for
+many common classification metrics.
 

### 2020-10-07 HW2/HW2-Group2.rmd
Highlight active URL in purple

@@
   html_document: default
+urlcolor: purple  
 ---

### 2020-10-07 HW3/HW 3_Bryan.Rmd
Update HW 3_Bryan.Rmd
Adding more exploring plots

@@
 library(corrplot)
+library(ggplot2)
+library(GGally)
 ```
@@
 A correlation matrix and plot helps to see which variables are more related to each other than others.
+
+```{r}
+corrplot(correlation, type = "lower", order="hclust", tl.col = "black", tl.srt = 45, method = "number")
+
+```
+
+```{r, fig.width=14, fig.height=12}
+ggpairs(as.data.frame(correlation), lower = list(continuous = "points", combo = "facethist", discrete = "facetbar", na =
+    "na"))
+
+```

### 2020-10-08 Adler_personal/Blog5.Rmd
Blog Post 5

---
title: "Blog Post 5: The Value of Cross Validation & Hyperparameter Tuning"
subtitle: "DT 621---Fall 2020"
author: "Avraham Adler"
date: "10/8/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, error=FALSE)
```

# What is Cross-Validation?
We saw in [previous blog posts](https://rpubs.com/Avraham/DT621Blog2) that
splitting data into training and testing sets helps to prevent overfitting. In
this post, we will demonstrate another method that can be used both separately
and in conjunction with holdout sets: **cross-validation**.

Cross validation is the process by which multiple models are fit to the same
data, each time holding out a different subset of that data. Unlike true holdout
methods, every data point is in at least one model being compared. However. this
does allow for a range of pseudo-holdout data which may better address
overfitting.

The most common types of cross validation are the *k-fold* and the
*leave-one-out* (LOO). The k-fold breaks the data into \(k\) partitions of
approximately the same size, and trains \(k\) models, each time leaving out a
different fold. LOOCV is k-fold taken to its extreme, where each fold is
composed of one data point. So if there are \(n\) observations there will be
\(n\) models.

The two overfitting prevention methods may be combined. One can sequester a
true holdout set for final model comparison and use cross-validation on the
training set for better overfitting behavior. One can even go so far as to split
the data into **three** sections: a true *test* set, a *validation* set on which
one can check the results of hyperparameter tuning, and the *train* set used to
fit the model including cross-validation.

# What is Hyperparamter Tuning?
Hyperparameters in machine learning usually refer to model-specific parameters
which are independent of the data. For example, the \(\lambda\) used in lasso
or ridge regression is a hyperparameter. The number of neighbors \(k\) is a
hyperparameter. Often, careful selection of these values can result in a model
with increased sensitivity to the data

# Demonstration
For this blog post we will use the `cars` dataset from the `caret` package, and
compare the results of using cross validation in the training phase. We will not
perform any EDA or preprocessing for the purpose of this post. We will train two
sets of models. We will select a "future" set of around 25% of the data to
reflect true unobserved data. We will split the remaining data into training and
testing at about a 70/30 ratio.

The first set of models will train on the entire available training set and the
best performing model on the test set will be selected. The second set will do
the same, but will include 5-fold cross validation and hyperparameter tuning as
part of the training The decision metric will be RMSE.

```{r prepareData}
library(caret)
data(cars)
set.seed(84)
futIDX <- createDataPartition(cars$Price, p = 0.25)$Resample1
futDat <- cars[futIDX, ]
curDat <- cars[-futIDX, ]
tstIDX <- createDataPartition(curDat$Price[-futIDX], p = 0.3)$Resample1
trnDat <- curDat[-tstIDX, ]
tstDat <- curDat[tstIDX, ]
```

# Model Training
For a serious project, one-pass tuning would be insufficient. The results of
each run would be used to fine-tune the parameters.

## ElasticNet
```{r enFit}
trc <- trainControl(method = 'none')
set.seed(245)
en1 <- train(Price ~ ., data = tstDat, method = 'glmnet', trControl = trc)
trc <- trainControl(method = 'cv', number = 5)
en2 <- train(Price ~ ., data = tstDat, method = 'glmnet', trControl = trc)
en1p <- predict(en1, tstDat)
en2p <- predict(en2, tstDat)
```

## Cubist
```{r cubFit}
trc <- trainControl(method = 'none')
tg <- expand.grid(committees = 1L, neighbors = 0)
set.seed(245)
cb1 <- train(Price ~ ., data = tstDat, method = 'cubist', trControl = trc,
             tuneGrid = tg)
trc <- trainControl(method = 'cv', number = 5)
set.seed(245)
cb2 <- train(Price ~ ., data = tstDat, method = 'cubist', trControl = trc)
cb1p <- predict(cb1, tstDat)
cb2p <- predict(cb2, tstDat)
```

# Model Selection
## Models Trained on All the Data
```{r MS1}
modRes <- cbind(data.frame(Model = c("EN", "EN-CV", "Cubist", "Cubist-CV")),
                rbind(defaultSummary(data.frame(obs = tstDat$Price, pred = en1p)),
                      defaultSummary(data.frame(obs = tstDat$Price, pred = en2p)),
                      defaultSummary(data.frame(obs = tstDat$Price, pred = cb1p)),
                      defaultSummary(data.frame(obs = tstDat$Price, pred = cb2p))
                      )
                )
knitr::kable(modRes, digits = 3L, caption = "Results on Training Set")
```

It is clear that cross-validation helps each time. The ability to compare
different tuning parameters has an extraordinary effect on the rules-based model
tree (`Cubist`). This last model would clearly be the one selected.

# Future Test
We will show results for all four models.
```{r MS2}
futRes <- cbind(data.frame(Model = c("EN", "EN-CV", "Cubist", "Cubist-CV")),
                rbind(defaultSummary(data.frame(obs = futDat$Price,
                                                pred = predict(en1, futDat))),
                      defaultSummary(data.frame(obs = futDat$Price,
                                                pred = predict(en2, futDat))),
                      defaultSummary(data.frame(obs = futDat$Price,
                                                pred = predict(cb1, futDat))),
                      defaultSummary(data.frame(obs = futDat$Price,
                                                pred = predict(cb2, futDat)))
                      )
                )
knitr::kable(futRes, digits = 3L, caption = "Results on 'Future' Set")
```

As expected, using both cross validation and hyperparameter tuning enhanced the
accuracy of both sets of models. It is also gratifying to see that the single
model we would have used was the most performant.

A real-world exercise would have started with significant EDA, have used more
models, and would have continued with more precise hyperparameter tuning.

### 2020-10-08 HW3/HW3-EDA-Group2.rmd

---
title: "DATA 621 - Business Analytics and Data Mining"
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, Bryan Persaud,
  Jeff Shamp
date: "10/11/2020"
output:
  pdf_document: default
subtitle: 'Fall 2020 - Group 2 - Homework #3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
library(tidyverse)
library(summarytools)
library(corrplot)
library(knitr)

urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW3/data/"
fileTrain = "crime-training-data_modified.csv"
fileTest = "crime-evaluation-data_modified.csv"

df = read.csv(paste0(urlRemote, pathGithub, fileTrain))
eval = read.csv(paste0(urlRemote, pathGithub, fileTest))
```

## DATA EXPLORATION
### Data Description

The training data set contains 466 records summarizing attributes of various neighborhoods in the city of Boston. The response variable is coded such that it is `1` when the neighborhood’s crime rate is above the median and `0` when it is not. In all, there are 12 predictors. These include:

Predictor Variables|Description   
-|----
zn | proportion of residential land zoned for large lots (over 25000 square feet) 
indus | proportion of non-retail business acres per suburb  
chas | a dummy var. for whether the suburb borders the Charles River (1) or not (0)
nox | nitrogen oxides concentration (parts per 10 million) 
rm | average number of rooms per dwelling   
age | proportion of owner-occupied units built prior to 1940   
dis | weighted mean of distances to five Boston employment centers  
rad | index of accessibility to radial highways   
tax | full-value property-tax rate per $10,000 
ptratio | pupil-teacher ratio by town 
lstat | lower status of the population (percent) 
medv | median value of owner-occupied homes in $1000s

This data set has complete cases, thus there is no need for imputation. Based on some common summary statistics, there are more observations where the crime rate is below the median. It is already apparent that some of the predictors varies depending the crime rate. For instance, there is a noticeable difference in the means of `age`, `lstat`, `rad`, and `zn` between the crime rate groups.

```{r sumstat}
summarystat = stby(data = df, INDICES = df$target, FUN = psych::describe)
kable(summarystat[[1]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate > Median", 
      digit = 2L)
kable(summarystat[[2]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate < Median", 
      digit = 2L)
```

### Data Distribution

For each predictors, we computed and drew kernel density estimate to understand their distribution. The following plots show how predictors are distributed between areas where crime rate is higher than the median, i.e. blue, and areas where crime rate is below the median, i.e. red. It is of interest to understand variables that highlight large variations between the two groups.

As a result, the density plots reveal that most is the data is not normal. some variables are heavily right skewed, while others possess multi-model distributions. More specifically, the variable predictor for average number of rooms per dwelling, i.e. `rm`, seems to be the best variable to divide the data into the two groups, since it closely mirrors a normal distribution between groups. Another interesting predictor is `zn`, i.e. proportion of residential land zoned for large lots. It has a significant positive skew (skew = 2.18). Nearly 73% of the observations (339 of 466 total) that have a value of 0. When analyzing the difference between the crime rate groups, it is possible that areas with high crime rate do not have land zoned for large lots. This suggests that suburban areas are likely to have low crime rates because there typically have large lots, whereas urban areas have a higher crime rate since lot size are smaller than 25000 square feet.

```{r density}
variable = names(df)
df.new = df
denplot = function(i){
  df.new$x = df.new[,variable[i]]
  ggplot(df.new, aes(x, fill = factor(df.new$target))) + 
    ggtitle(sprintf("%s", variable[i])) + 
    geom_density(alpha = 0.4, show.legend = FALSE) + 
    theme(legend.position = "none",
          plot.title = element_text(size = 10, hjust = 0.5),
          axis.title.y = element_blank(),
          axis.title.x = element_blank())
}
gridExtra::grid.arrange(denplot(1),denplot(2),denplot(3),denplot(4),
                        denplot(5),denplot(6),denplot(7),denplot(8),
                        denplot(9),denplot(10),denplot(11),denplot(12),
                        layout_matrix = rbind(c(0,1,2), c(3,4,5), c(6,7,8), c(9,10,11)))
```

Further visually inspections was conducted using boxplots to highlight how the data is spread for each variable. The graphs show that some variables have a large amount of variance between each other, for example, `zn`, `rad` and `tax`. There are also quite a few values that can be considered outliers within some variables.

```{r boxplot}
df.new = reshape::melt(df, id.vars= "target") %>%
  dplyr::filter(variable != "chas") %>%
  mutate(target = as.factor(target))
ggplot(df.new, aes(x = variable, y = log(value))) + 
  geom_boxplot(color = "blue", fill = "blue", alpha = 0.2, 
               notch = TRUE, notchwidth = 0.8, 
               outlier.colour = "red", outlier.fill = "red", outlier.size = 3) + 
  stat_summary(fun.y = mean, color = "red", geom = "point", shape = 16, size = 2) + 
  coord_flip() + 
  labs(title = "Boxplot of Predictor Variables")

ggplot(df.new, aes(x = variable, y = log(value))) + 
  geom_boxplot(alpha = 0.5, outlier.colour = "red", 
               outlier.fill = "red", outlier.size = 2,
               aes(fill = target)) + 
  facet_wrap( ~ variable, scales = "free") + 
  labs(title = "Boxplot of Predictor Variables by Crime Rate") 
```

### Data Correlation

By looking at a correlation matrix, boxes with blank color indicate that the correlation was not statistically significant. It is evident that many variables are moderately to highly correlated, $\rho > 0.50$. The correlogram below shows that `nox` has a largest positive correlation with `target`, whereas `dis` has a largest negative correlation with `target`.

There is a  noticeable influence that socio-economic status has on the crime rate of an area. It is well-known that crime is concentrated in disadvantaged, urban neighborhoods in the United States. The economic segregation suggests that affluent neighborhoods may be further away in terms of distance, and as a result, disadvantaged areas are more attractive to crime because the probability of success is higher even if the targets are not as profitable. Consider areas with a higher socio-economics, these areas are correlated have a lower crime rate. Moreover, industrial areas are correlated with higher levels of nitrogen oxide concentrations. This may be an indication of why these areas are less density with residents of a higher status. This trend is also seen with the `ptratio` since a higher ratio means less funding for public institutions, which is common in areas of lower status.

Lastly, it seems that the variable `chas` which indicate whether the suburb borders the Charles River have statistically insignificant correlation with almost all of the other variables except for `medv`, which is poorly positively correlated.

```{r correlation, fig.height=6, fig.width=6}
cor.df = function(df){
  df.m = as.matrix(df)
  n = ncol(df.m)
  p.mat = matrix(NA, n, n)
  diag(p.mat) = 0
  for (i in 1:(n - 1)){
    for (j in (i + 1):n){
      tmp = cor.test(df.m[, i], df.m[, j])
      p.mat[i, j] = p.mat[j, i] = tmp$p.value
    }
  }
  colnames(p.mat) = rownames(p.mat) = colnames(df.m)
  p.mat
}
p.mat = cor.df(df)
M = cor(df)
col = colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(round(M,1), method = "color", col = col(200),  
         type = "lower", order = "hclust", 
         addCoef.col = "black",
         tl.col = "black", tl.srt = 45,
         p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         diag = FALSE)
```

## APPENDIX
The code chunks below represent the R code called in order during the analysis. They are reproduced in the appendix for review and comment.
```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r sumstat}
```
```{r density}
```
```{r boxplot}
```
```{r correlation}
```

### 2020-10-08 Adler_personal/Blog5.Rmd
More stable this way somehow

@@
 curDat <- cars[-futIDX, ]
-tstIDX <- createDataPartition(curDat$Price[-futIDX], p = 0.3)$Resample1
-trnDat <- curDat[-tstIDX, ]
-tstDat <- curDat[tstIDX, ]
+trnIDX <- createDataPartition(curDat$Price, p = 0.7)$Resample1
+trnDat <- curDat[trnIDX, ]
+tstDat <- curDat[-trnIDX, ]
 ```

### 2020-10-09 HW3/HW3_Final.Rmd
ambers updates; model 1

--- 2020-10-08 HW3/HW3-EDA-Group2.rmd
@@
 library(knitr)
+library(rsample) # model 1 libraries
+library(caret)
+library(e1071)
+
+set.seed(9450)
 
@@
 
-## DATA EXPLORATION
-### Data Description
+# DATA EXPLORATION
+## Data Description
 
@@
 ```{r sumstat}
+# DATA EXPLORATION 
 summarystat = stby(data = df, INDICES = df$target, FUN = psych::describe)
@@
 
-### Data Distribution
+## Data Distribution
 
@@
 ```{r density}
+# density plots
 variable = names(df)
@@
   df.new$x = df.new[,variable[i]]
-  ggplot(df.new, aes(x, fill = factor(df.new$target))) + 
+  ggplot(df.new, aes(x, fill = factor(target))) + 
     ggtitle(sprintf("%s", variable[i])) + 
@@
 ```{r boxplot}
+# boxplots 
 df.new = reshape::melt(df, id.vars= "target") %>%
@@
 
-### Data Correlation
+## Data Correlation
 
@@
 ```{r correlation, fig.height=6, fig.width=6}
+# correlations
 cor.df = function(df){
@@
 
-## APPENDIX
+# DATA PREPARATION
+
+# BUILD MODELS
+
+First, we will split the training data into two sets. This will allow us to perform a cross validation scheme on the models to tune parameters for optimum performance.
+
+```{r tts}
+
+# BUILD MODELS
+# train-test split
+crime_df<-
+  df %>% 
+  mutate(target = as.factor(target)) %>% 
+  mutate_if(is.integer, as.numeric)
+
+data_split <- initial_split(crime_df, 
+                            strata = target, 
+                            prop = 0.8)
+train_df<- training(data_split)
+test_df<- testing(data_split)
+```
+
+## Model 1
+### Base Model
+We will start with a simple logistic regression model to serve as a baseline. This includes
+all variables in the dataset.
+
+```{r baseModel}
+
+baseModel <- glm(target~., family = binomial, data = train_df)
+
+knitr::kable(summary(baseModel)$coefficients, digits = 3L,
+             caption = 'Base Model Logistic Regression Output')
+
+```
+We can immediately see that a few variables *exceed* the 0.05 p-value threshold
+for significance.
+
+### Enhanced Model
+We will use **backwards stepwise regression** to remove features that are not statistically significant in predicting the target.
+The result is a model that includes the following features: `zn`, `nox`, `age`, `dis`, `rad`, `tax`, `ptratio`, `lstat`, and `medv`.
+
+### Coefficient Discussion
+It is important to note that the coefficients from the model are predicting whether or not the target variable is a **1** (the crime rate is *above* the median value).
+Additionally, it must be noted that the numeric coefficients are relative to the range of values that the variable encompasses.
+What this means is that it's possible to have a coefficient that seems small when we look at the absolute magnitude, but that actually has a very strong effect when applied to the data.
+```{r model1_coeff}
+
+# 10-fold cross validation
+train_control <- trainControl(method = "cv", number = 10)
+
+# train the model on training set
+model1 <- train(target ~ zn+nox+age+dis+rad+tax+ptratio+lstat+medv,
+               data = train_df,
+               trControl = train_control,
+               method = "glm",
+               family=binomial())
+
+knitr::kable(summary(model1)$coefficients, digits = 3L,
+             caption = 'Backwards Model Regression Output')
+
+```
+Let's take a look at the *signs* of the coefficients.
+We can see that the coefficients for the variables `zn` and `tax` are negative.
+This is indicative of an inverse relationship; ie, the higher these values, the less likely the crime rate is above the median.
+The relationship for the zone variable aligns with the findings from our initial data exploration.
+However, the relationship for the tax variable does not; in fact, we saw that there was a positive correlation between the two (as the tax rate increases, so does the probability that the crime rate will be above the median).
+
+The rest of the variables have positive coefficients.
+Many of these are expected.
+For example, in our exploratory data analysis, we noticed that the `nox` variable (nitrogen oxides concentration (parts per 10 million)) has the greatest positive correlation with the target variable.
+This is surprising; we hadn't thought that the nitrogen oxide concentration would have as large of an impact on the crime rate, but some quick googling shows us that there may actually be a relationship.
+( [https://www.sciencedaily.com/releases/2019/10/191003114007.htm](Source) )
+
+# SELECT MODELS
+## Criteria
+
+## Performance
+
+### Model 1
+```{r model1_perf}
+
+# SELECT MODELS
+# Model 1 Performance
+model1_preds <- predict(model1, test_df, type = "raw")
+model1_probs <- predict(model1, test_df, type = "prob")
+colnames(model1_probs) <- c('pred0', 'pred1')
+
+model1_results <- test_df %>%
+  bind_cols(pred = model1_preds, model1_probs)
+
+####### metrics
+m1_roc<- yardstick::roc_auc(
+  model1_results,
+  truth = target,
+  pred0 # select the prob class that corresponds to first level of target
+)
+
+m1_acc<- yardstick::accuracy(
+  model1_results, 
+  truth = target, 
+  estimate = pred
+)
+
+m1_recall<- yardstick::recall(
+  model1_results,
+  truth = target, 
+  estimate = pred
+)
+
+m1_precise<- yardstick::precision(
+  model1_results,
+  truth = target, 
+  estimate = pred
+)
+
+m1_metrics_df<-
+  bind_rows(m1_roc, m1_acc, m1_recall, m1_precise)
+
+m1_metrics_df
+
+```
+
+# APPENDIX
 The code chunks below represent the R code called in order during the analysis. They are reproduced in the appendix for review and comment.

### 2020-10-12 HW3/HW3_Final.Rmd
added stacked model

@@
 library(tidyverse)
+library(tidymodels)
+library(stacks)
+library(kernlab)
+library(vip)
 library(summarytools)
@@
 df = read.csv(paste0(urlRemote, pathGithub, fileTrain))
+df<- as_tibble(df)
+
 eval = read.csv(paste0(urlRemote, pathGithub, fileTest))
@@
 
+## Model 2
+### Stacked Classifier
+
+```{r}
+cv_folds<- vfold_cv(train_df,
+                    v = 10, repeats = 1)
+
+crime_recipe<- 
+  recipe(target ~., data=train_df)
+
+crime_wf<- 
+  workflow() %>%
+  add_recipe(crime_recipe)
+
+crtl_grid<- control_stack_grid()
+```
+
+Specify the models. Using SVM as another model to blend with the logistic.
+
+```{r}
+logit_specs<- 
+  logistic_reg(
+    penalty = tune(),
+    mixture = tune()
+    ) %>%
+  set_engine("glm") %>%
+  set_mode("classification")
+
+svm_specs<- 
+  svm_rbf(
+    cost = tune(),
+    rbf_sigma = tune(),
+    margin = tune()
+  ) %>%
+  set_mode("classification") %>%
+  set_engine("kernlab")
+```
+
+Tuning the models for best parameters. The stack function will cherry pick the models that best optimize the blended final model.
+
+```{r}
+svm_wf<-
+  crime_wf %>%
+  add_model(svm_specs)
+
+svm_results<-
+  tune_grid(
+    svm_wf, 
+    resample = cv_folds,
+    control = crtl_grid,
+    grid = 10,
+    save_pred = TRUE,
+    save_workflow = FALSE
+  )
+
+logit_wf<- 
+  crime_wf %>%
+  add_model(logit_specs)
+
+logit_results<-
+  tune_grid(
+    logit_wf,
+    resamples = cv_folds,
+    control = crtl_grid,
+    grid = 10,
+    save_pred = TRUE,
+    save_workflow = FALSE
+)
+
+```
+
+Blend!
+
+```{r}
+crime_model_stack <- 
+  stacks() %>%
+  add_candidates(logit_results) %>%
+  add_candidates(svm_results) %>%
+  blend_predictions() %>%
+  fit_members()
+
+crime_model_stack
+```
+
+```{r}
+crime_pred <- 
+  test_df %>%
+  bind_cols(predict(crime_model_stack, ., type = "prob"), 
+            predict(crime_model_stack, ., type = "class"))
+```
+
+Computing the metrics for the stacked model:
+
+```{r}
+roc<- yardstick::roc_auc(
+  crime_pred,
+  truth = target,
+  contains(".pred_1")
+  )
+
+acc<- yardstick::accuracy(
+  crime_pred, 
+  truth = target, 
+  estimate = .pred_class
+)
+
+recall<- yardstick::recall(
+  crime_pred,
+  truth = target, 
+  estimate = .pred_class
+)
+
+precise<- yardstick::precision(
+  crime_pred,
+  truth = target, 
+  estimate = .pred_class
+)
+metrics_df<-
+  bind_rows(roc, acc, recall, precise)
+
+crime_pred %>%
+  conf_mat(truth = target, estimate = .pred_class)
+```
+
+```{r}
+metrics_df
+```
+
 # SELECT MODELS
@@
 
+### Sub-Models and Coefficient Discussion
+
+The blended model is comprised of two sub-models, the logistic and the SVM. Below are the prediction results for each sub-model.
+
+```{r}
+best_model_metric <- select_best(svm_results, "accuracy")
+finalize_workflow(svm_wf, best_model_metric) %>%
+  last_fit(data_split) %>%
+  collect_metrics()
+```
+
+```{r}
+logit_wf %>%
+  last_fit(data_split) %>%
+  collect_metrics()
+
+logit_wf %>%
+  fit(data = train_df) %>%
+  pull_workflow_fit() %>%
+  vip(geom = "col", aesthetics = list(fill='red4'))
+```
+
+We see that they are both in the low eighties percent for accuracy, and low nineties for roc auc. Together make for a better overall prediction.
+
+We again see that the primary predictor in the logistic regression is the `nox` with `rad` and `ptratio` rounding out the top three predictors.
+This is in line with the base model.
+Living close to the center of town, having teachers, and not having polluted air are the largest predictors in terms of above median crime.
+
 # APPENDIX

### 2020-10-26 HW3/HW3_Final_AA.Rmd
Starting Cleanup

--- 2020-10-12 HW3/HW3_Final.Rmd
@@
 title: "DATA 621 - Business Analytics and Data Mining"
-author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, Bryan Persaud,
-  Jeff Shamp
+subtitle: 'Fall 2020 - Group 2 - Homework #3'
+author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
+    Bryan Persaud, Jeff Shamp
 date: "10/11/2020"
@@
   pdf_document: default
-subtitle: 'Fall 2020 - Group 2 - Homework #3'
+
 ---
@@
 library(e1071)
+library(geoR)
 
@@
 
-df = read.csv(paste0(urlRemote, pathGithub, fileTrain))
-df<- as_tibble(df)
-
-eval = read.csv(paste0(urlRemote, pathGithub, fileTest))
+df <- read.csv(paste0(urlRemote, pathGithub, fileTrain))
+df <- as_tibble(df)
+eval <- read.csv(paste0(urlRemote, pathGithub, fileTest))
 ```
@@
 
-The training data set contains 466 records summarizing attributes of various neighborhoods in the city of Boston.
-The response variable is coded such that it is `1` when the neighborhood’s crime rate is above the median and `0` when it is not.
-In all, there are 12 predictors.
-These include:
+The training data set contains 466 records summarizing attributes of various
+neighborhoods in the city of Boston. The response variable is coded such that it
+is `1` when the neighborhood’s crime rate is above the median and `0` when it is
+not. In all, there are 12 predictors. These include:
 
@@
 
-This data set has complete cases, thus there is no need for imputation.
-Based on some common summary statistics, there are more observations where the crime rate is below the median.
-It is already apparent that some of the predictors varies depending the crime rate.
-For instance, there is a noticeable difference in the means of `age`, `lstat`, `rad`, and `zn` between the crime rate groups.
+This data set has complete cases, thus there is no need for imputation. Based on
+some common summary statistics, there are more observations where the crime rate
+is below the median. It is already apparent that some of the predictors varies
+depending the crime rate. For instance, there is a noticeable difference in the
+means of `age`, `lstat`, `rad`, and `zn` between the crime rate groups.
 
@@
 
-For each predictors, we computed and drew kernel density estimate to understand their distribution.
-The following plots show how predictors are distributed between areas where crime rate is higher than the median, i.e.
-blue, and areas where crime rate is below the median, i.e.
-red.
-It is of interest to understand variables that highlight large variations between the two groups.
+For each predictors, we computed kernel density estimators to understand their
+distribution. The following plots show how predictors are distributed between
+areas where the crime rate is higher than the median (blue) and areas where the
+crime rate is below the median (red). It is of interest to understand variables
+that highlight large variations between the two groups.
+
+```{r density}
+# density plots
+var_names <- cbind(names(df), 'factor(target)')
+plot_list <- list()
+for (i in seq_along(names(df))) {
+   p <- ggplot(df, aes_string(x=var_names[i, 1], fill=var_names[i, 2])) +
+                 ggtitle(var_names[i, 1]) +
+                 geom_density(alpha = 0.4, show.legend = FALSE) + 
+                 theme(legend.position = "none",
+                       plot.title = element_text(size = 10, hjust = 0.5),
+                       axis.title.y = element_blank(),
+                       axis.title.x = element_blank())
+   plot_list[i] <- p
+}
+gridExtra::grid.arrange(plot_list[1], plot_list[2], plot_list[3], plot_list[4],
+                        plot_list[5], plot_list[6], plot_list[7], plot_list[8],
+                        plot_list[9], plot_list[10], plot_list[11],
+                        plot_list[12],
+                        layout_matrix = rbind(c(0,1,2), c(3,4,5), c(6,7,8),
+                                              c(9,10,11)))
+```
 
@@
 This suggests that suburban areas are likely to have low crime rates because there typically have large lots, whereas urban areas have a higher crime rate since lot size are smaller than 25000 square feet.
-
-```{r density}
-# density plots
-variable = names(df)
-df.new = df
-denplot = function(i){
-  df.new$x = df.new[,variable[i]]
-  ggplot(df.new, aes(x, fill = factor(target))) + 
-    ggtitle(sprintf("%s", variable[i])) + 
-    geom_density(alpha = 0.4, show.legend = FALSE) + 
-    theme(legend.position = "none",
-          plot.title = element_text(size = 10, hjust = 0.5),
-          axis.title.y = element_blank(),
-          axis.title.x = element_blank())
-}
-gridExtra::grid.arrange(denplot(1),denplot(2),denplot(3),denplot(4),
-                        denplot(5),denplot(6),denplot(7),denplot(8),
-                        denplot(9),denplot(10),denplot(11),denplot(12),
-                        layout_matrix = rbind(c(0,1,2), c(3,4,5), c(6,7,8), c(9,10,11)))
-```
 
@@
 # DATA PREPARATION
+
+The variable `chas` will be removed due to having statistically insignificant correlation with almost all the variables in the dataset.
+
+```{r}
+new_df <- dplyr::select(df, -chas)
+```
+
+The lower outliers at the lower 1.5 IQR will be capped at the 5th percentile while the higher outliers at the upper 1.5 IQR will be capped at the 95th percentile.
+
+```{r}
+id <- c(1:12)
+for (value in id) {
+  quantiles <- quantile(new_df[,value], probs = c(.25, .75), na.rm = TRUE)
+  caps <- quantile(new_df[,value], probs = c(.05, .95), na.rm = T)
+  H <- 1.5 * IQR(new_df[,value], na.rm = T)
+  new_df[,value][new_df[,value] < (quantiles[1] - H)] <- caps[1]
+  new_df[,value][new_df[,value] > (quantiles[2] + H)] <- caps[2]
+}
+```
+
+We will use Boxcox for variables that are skewed.
+
+```{r}
+boxcoxfit(new_df$age)
+boxcoxfit(new_df$lstat)
+```
+
+The boxcox suggest a transformation of 1.318 for age and 0.233 for lstat.
+
+```{r}
+new_df$age <- new_df$age ^ 1.318
+new_df$lstat <- new_df$lstat ^ 0.233
+```
+
+```{r}
+boxcoxfit(new_df$rm)
+```
+
+The boxcox suggests a transformation of 0.204 for rm.
+
+```{r}
+new_df$rm <- new_df$rm ^ 0.204
+```
+
+Plot to see the affects of the transformations.
+
+```{r}
+hist_age <- ggplot(new_df, aes(age)) + geom_histogram()
+hist_lstat <- ggplot(new_df, aes(lstat)) + geom_histogram()
+hist_rm <- ggplot(new_df, aes(rm)) + geom_histogram()
+grid.arrange(hist_age, hist_lstat, hist_rm)
+```
 

### 2020-10-27 HW3/HW3_Final_AA.Rmd
Progress? Maybe?

--- 2020-10-12 HW3/HW3_Final.Rmd
@@
 library(tidymodels)
+library(data.table)
 library(stacks)
@@
 
-df = read.csv(paste0(urlRemote, pathGithub, fileTrain))
-df<- as_tibble(df)
-
-eval = read.csv(paste0(urlRemote, pathGithub, fileTest))
+df <- read.csv(paste0(urlRemote, pathGithub, fileTrain))
+nobs <- dim(df)[[1]]
+DT <- as.data.table(df)
+DT[, target := factor(target)]
+# df <- as_tibble(df)
+eval <- read.csv(paste0(urlRemote, pathGithub, fileTest))
 ```
@@
 ## Data Description
-
-The training data set contains 466 records summarizing attributes of various neighborhoods in the city of Boston.
-The response variable is coded such that it is `1` when the neighborhood’s crime rate is above the median and `0` when it is not.
-In all, there are 12 predictors.
-These include:
+The training data set contains 466 records summarizing attributes of various
+neighborhoods in the city of Boston. The response variable is coded such that it
+is `1` when the neighborhood’s crime rate is above the median and `0` when it is
+not. In all, there are 12 predictors. These include:
 
@@
 ## Data Distribution
+For each predictors, we computed kernel density estimators to understand their
+distribution. The following plots show how predictors are distributed between
+areas where the crime rate is higher than the median (blue) and areas where the
+crime rate is below the median (red). It is of interest to understand variables
+that highlight large variations between the two groups.
 
-For each predictors, we computed and drew kernel density estimate to understand their distribution.
-The following plots show how predictors are distributed between areas where crime rate is higher than the median, i.e.
-blue, and areas where crime rate is below the median, i.e.
-red.
-It is of interest to understand variables that highlight large variations between the two groups.
-
-As a result, the density plots reveal that most is the data is not normal.
-some variables are heavily right skewed, while others possess multi-model distributions.
-More specifically, the variable predictor for average number of rooms per dwelling, i.e.
-`rm`, seems to be the best variable to divide the data into the two groups, since it closely mirrors a normal distribution between groups.
-Another interesting predictor is `zn`, i.e.
-proportion of residential land zoned for large lots.
-It has a significant positive skew (skew = 2.18).
-Nearly 73% of the observations (339 of 466 total) that have a value of 0.
-When analyzing the difference between the crime rate groups, it is possible that areas with high crime rate do not have land zoned for large lots.
-This suggests that suburban areas are likely to have low crime rates because there typically have large lots, whereas urban areas have a higher crime rate since lot size are smaller than 25000 square feet.
-
-```{r density}
+```{r density, fig.width=6}
 # density plots
-variable = names(df)
-df.new = df
-denplot = function(i){
-  df.new$x = df.new[,variable[i]]
-  ggplot(df.new, aes(x, fill = factor(target))) + 
-    ggtitle(sprintf("%s", variable[i])) + 
+DT[, IDX := .I]
+DTM <- melt(DT, id.vars = c('IDX', 'target'), measure.vars = 1:12,
+            variable.name = 'metric', value.name = 'value')
+ggplot(DTM, aes(x = value, fill = target, color = target)) +
   geom_density(alpha = 0.4, show.legend = FALSE) +
-    theme(legend.position = "none",
-          plot.title = element_text(size = 10, hjust = 0.5),
-          axis.title.y = element_blank(),
-          axis.title.x = element_blank())
-}
-gridExtra::grid.arrange(denplot(1),denplot(2),denplot(3),denplot(4),
-                        denplot(5),denplot(6),denplot(7),denplot(8),
-                        denplot(9),denplot(10),denplot(11),denplot(12),
-                        layout_matrix = rbind(c(0,1,2), c(3,4,5), c(6,7,8), c(9,10,11)))
+  facet_wrap(~ metric, scales = 'free') 
 ```
 
-Further visually inspections was conducted using boxplots to highlight how the data is spread for each variable.
-The graphs show that some variables have a large amount of variance between each other, for example, `zn`, `rad` and `tax`.
-There are also quite a few values that can be considered outliers within some variables.
+The density plots reveal that most of the data does not have a normal
+distribution. There are some are heavily right tailed variables and others which
+are multi-modal. The most Gaussian of the variables appears to be the one
+related to the average number of rooms per dwelling: `rm`. Another
+interesting predictor is `zn`, the proportion of residential land zoned for
+large lots. Nearly 73% of the observations have a value of 0 with the remaining
+spread widly between 1% and 100%. This may the source of possible distortion. A
+third highly-skewed variable is `chas`, the indicator as to whether the lot
+borders the Charles River. Out of the `r nobs` observations,
+`r sum(DT$chas == 0)` do **not** border the river.
 
-```{r boxplot}
+To provide another view, this time highlighting outliers, the data was analyzed
+using boxplos.
+
+```{r boxplots, fig.width=6}
 # boxplots
-df.new = reshape::melt(df, id.vars= "target") %>%
-  dplyr::filter(variable != "chas") %>%
-  mutate(target = as.factor(target))
-ggplot(df.new, aes(x = variable, y = log(value))) + 
-  geom_boxplot(color = "blue", fill = "blue", alpha = 0.2, 
-               notch = TRUE, notchwidth = 0.8, 
-               outlier.colour = "red", outlier.fill = "red", outlier.size = 3) + 
-  stat_summary(fun.y = mean, color = "red", geom = "point", shape = 16, size = 2) + 
-  coord_flip() + 
-  labs(title = "Boxplot of Predictor Variables")
 
-ggplot(df.new, aes(x = variable, y = log(value))) + 
-  geom_boxplot(alpha = 0.5, outlier.colour = "red", 
-               outlier.fill = "red", outlier.size = 2,
-               aes(fill = target)) + 
-  facet_wrap( ~ variable, scales = "free") + 
-  labs(title = "Boxplot of Predictor Variables by Crime Rate") 
+ggplot(DTM, aes(x = metric, y = value)) + 
+  geom_boxplot(color = "blue", fill = "blue", alpha = 0.2, notch = TRUE,
+               notchwidth = 0.8, outlier.colour = "red", outlier.fill = "red",
+               outlier.size = 3) + 
+  stat_summary(fun.y = mean, color = "red", geom = "point", shape = 16,
+               size = 2) + coord_flip() +
+  labs(title = "Boxplot of Predictor Variables") + scale_y_log10()
+
+ggplot(DTM, aes(x = metric, y = value)) + 
+  geom_boxplot(alpha = 0.5, outlier.colour = "red", outlier.fill = "red",
+               outlier.size = 2, aes(fill = target)) + 
+  facet_wrap( ~ metric, scales = "free") + 
+  labs(title = "Boxplot of Predictor Variables by Crime Rate") +
+  scale_y_log10()
 ```
 
+It is clear from the second boxplot, similar to the densities, that the
+distribution of the predictor variables is different for for two outcomes. This
+suggests that a model will be able to extract signal from the data.
+
 ## Data Correlation
-
-By looking at a correlation matrix, boxes with blank color indicate that the correlation was not statistically significant.
-It is evident that many variables are moderately to highly correlated, $\rho > 0.50$.
-The correlogram below shows that `nox` has a largest positive correlation with `target`, whereas `dis` has a largest negative correlation with `target`.
-
-There is a  noticeable influence that socio-economic status has on the crime rate of an area.
-It is well-known that crime is concentrated in disadvantaged, urban neighborhoods in the United States.
-The economic segregation suggests that affluent neighborhoods may be further away in terms of distance, and as a result, disadvantaged areas are more attractive to crime because the probability of success is higher even if the targets are not as profitable.
-Consider areas with a higher socio-economics, these areas are correlated have a lower crime rate.
-Moreover, industrial areas are correlated with higher levels of nitrogen oxide concentrations.
-This may be an indication of why these areas are less density with residents of a higher status.
-This trend is also seen with the `ptratio` since a higher ratio means less funding for public institutions, which is common in areas of lower status.
-
-Lastly, it seems that the variable `chas` which indicate whether the suburb borders the Charles River have statistically insignificant correlation with almost all of the other variables except for `medv`, which is poorly positively correlated.
+Below is a corrgram of the data. The shape of the ellipse reflects the strength
+of the correlation and the color represents the direction. Thus, the diagonal
+which represents the perfect correlation of each variable with itself is a blue
+diagional line. The shape and the color scheme reflect the guide at the bottom
+of the image.
 
@@
 # correlations
-cor.df = function(df){
-  df.m = as.matrix(df)
-  n = ncol(df.m)
-  p.mat = matrix(NA, n, n)
-  diag(p.mat) = 0
-  for (i in 1:(n - 1)){
-    for (j in (i + 1):n){
-      tmp = cor.test(df.m[, i], df.m[, j])
-      p.mat[i, j] = p.mat[j, i] = tmp$p.value
-    }
-  }
-  colnames(p.mat) = rownames(p.mat) = colnames(df.m)
-  p.mat
-}
-p.mat = cor.df(df)
-M = cor(df)
-col = colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
-corrplot(round(M,1), method = "color", col = col(200),  
-         type = "lower", order = "hclust", 
-         addCoef.col = "black",
-         tl.col = "black", tl.srt = 45,
-         p.mat = p.mat, sig.level = 0.01, insig = "blank", 
-         diag = FALSE)
+corrplot(cor(df), method = 'ellipse', type = 'lower', order = 'hclust')
 ```
 
+Looking at the corrgram, we see that there are many variables which are
+moderately to highly correlated, $\left|\rho\right| > 0.50$. Of note is that
+`nox` has a largest positive correlation with `target` and `dis` has the largest
+negative correlation with `target`. Lastly, it seems that the variable `chas`,
+which indicate whether the suburb borders the Charles River, has statistically
+insignificant correlation with almost all of the other variables.
+
 # DATA PREPARATION
+There are no missing values so none need to be imputed.
+Different preparation techniques may be used for differen models.
+## Models 1 & 2
+The variable `chas` will be removed due to having statistically insignificant\
+correlation with almost all the variables in the dataset. Next, outliers will be
+tempered by capping from below at the 5th percentile and from above at the 95th
+percentile.
 
+```{r, model1_2_adjust}
+new_df <- subset(df, select = -chas)
+
+# Cap all observations at their 5th and 95th percentiles
+low5 <- apply(new_df, 2, quantile, prob = 0.05)
+up95 <- apply(new_df, 2, quantile, prob = 0.95)
+for (i in seq_along(new_df)) {
+  new_df[, i] <- pmin(new_df[, i], up95[i])
+  new_df[, i] <- pmax(new_df[, i], low5[i])
+}
+dens_ageO <- ggplot(new_df, aes(age)) + geom_density()
+dens_lstatO <- ggplot(new_df, aes(lstat)) + geom_density()
+dens_rmO <- ggplot(new_df, aes(rm)) + geom_density()
+
+```
+
+The Boxcox transformation will be applied to the `age`, `lstat`, and `rm`
+variables.
+
+```{r m12BC}
+ageBC <- boxcoxfit(new_df$age)
+lstatBC <- boxcoxfit(new_df$lstat)
+rmBC <- boxcoxfit(new_df$rm)
+```
+
+The transform suggests a \(\lambda\) of `r ageBC$lambda` for `age`,
+`r lstatBC$lambda` for `lstat`, and `r rmBC$lambda` for `rm`
+
+```{r, m12applyBC}
+new_df$age <- new_df$age ^ ageBC$lambda
+new_df$lstat <- new_df$lstat ^ lstatBC$lambda
+new_df$rm <- new_df$rm ^ rmBC$lambda
+```
+
+Below are the densities of the variables prior (left) and post (right) the
+Box-Cox transform. The variables `lstat` and `rm` have benefitted from the
+transform, but the strong bimodality of `age` has not been affected
+significantly.
+
+```{r, m12BC_plot}
+dens_age <- ggplot(new_df, aes(age)) + geom_density()
+dens_lstat <- ggplot(new_df, aes(lstat)) + geom_density()
+dens_rm <- ggplot(new_df, aes(rm)) + geom_density()
+grid.arrange(dens_ageO, dens_lstatO, dens_rmO, dens_age, dens_lstat, dens_rm,
+             layout_matrix = cbind(c(0, 1, 2), c(3, 4, 5)))
+```
+
+## Model 3
 # BUILD MODELS
+## Models 1 & 2
+First, we will split the training data into two sets. This will allow us to
+perform a cross validation scheme on the models to tune parameters for optimum
+performance.
 
-First, we will split the training data into two sets. This will allow us to perform a cross validation scheme on the models to tune parameters for optimum performance.
-
-```{r tts}
-
+```{r m12_tts}
 # BUILD MODELS
 # train-test split
-crime_df<-
-  df %>% 
+crime_df <- as.tibble(df) %>%
   mutate(target = as.factor(target)) %>% 
@@
 ### Base Model
-We will start with a simple logistic regression model to serve as a baseline. This includes
-all variables in the dataset.
+We will start with a simple logistic regression model to serve as a baseline.
+This includes all variables in the dataset.
 
-```{r baseModel}
-
+```{r baseModel1}
 baseModel <- glm(target~., family = binomial, data = train_df)
-
 knitr::kable(summary(baseModel)$coefficients, digits = 3L,
              caption = 'Base Model Logistic Regression Output')
+```
 
-```
 We can immediately see that a few variables *exceed* the 0.05 p-value threshold
@@
 ### Enhanced Model
-We will use **backwards stepwise regression** to remove features that are not statistically significant in predicting the target.
-The result is a model that includes the following features: `zn`, `nox`, `age`, `dis`, `rad`, `tax`, `ptratio`, `lstat`, and `medv`.
+We will use **backwards stepwise regression** to remove features that are not
+statistically significant in predicting the target. The result is a model that
+includes the following features: `zn`, `nox`, `age`, `dis`, `rad`, `tax`,
+`ptratio`, `lstat`, and `medv`.
 
 ### Coefficient Discussion
-It is important to note that the coefficients from the model are predicting whether or not the target variable is a **1** (the crime rate is *above* the median value).
-Additionally, it must be noted that the numeric coefficients are relative to the range of values that the variable encompasses.
-What this means is that it's possible to have a coefficient that seems small when we look at the absolute magnitude, but that actually has a very strong effect when applied to the data.
+It is important to note that the coefficients from the model are predicting
+whether or not the target variable is a **1** (the crime rate is *above* the
+median value). Additionally, it must be noted that the numeric coefficients are
+relative to the range of values that the variable encompasses. What this means
+is that it's possible to have a coefficient that seems small when we look at the
+absolute magnitude, but that actually has a very strong effect when applied to
+the data.
+
 ```{r model1_coeff}
-
 # 10-fold cross validation
@@
 # train the model on training set
-model1 <- train(target ~ zn+nox+age+dis+rad+tax+ptratio+lstat+medv,
+model1 <- train(target ~ zn + nox + age + dis + rad + tax + ptratio + lstat + medv,
                data = train_df,
@@
                family=binomial())
-
 knitr::kable(summary(model1)$coefficients, digits = 3L,
              caption = 'Backwards Model Regression Output')
+```
 
-```
-Let's take a look at the *signs* of the coefficients.
-We can see that the coefficients for the variables `zn` and `tax` are negative.
-This is indicative of an inverse relationship; ie, the higher these values, the less likely the crime rate is above the median.
-The relationship for the zone variable aligns with the findings from our initial data exploration.
-However, the relationship for the tax variable does not; in fact, we saw that there was a positive correlation between the two (as the tax rate increases, so does the probability that the crime rate will be above the median).
+Let's take a look at the *signs* of the coefficients. We can see that the
+coefficients for the variables `zn` and `tax` are negative. This is indicative
+of an inverse relationship; ie, the higher these values, the less likely the
+crime rate is above the median. The relationship for the zone variable aligns
+with the findings from our initial data exploration. However, the relationship
+for the tax variable does not; in fact, we saw that there was a positive
+correlation between the two (as the tax rate increases, so does the probability
+that the crime rate will be above the median).
 
-The rest of the variables have positive coefficients.
-Many of these are expected.
-For example, in our exploratory data analysis, we noticed that the `nox` variable (nitrogen oxides concentration (parts per 10 million)) has the greatest positive correlation with the target variable.
-This is surprising; we hadn't thought that the nitrogen oxide concentration would have as large of an impact on the crime rate, but some quick googling shows us that there may actually be a relationship.
-( [https://www.sciencedaily.com/releases/2019/10/191003114007.htm](Source) )
+The rest of the variables have positive coefficients. Many of these are
+expected. For example, in our exploratory data analysis, we noticed that the
+`nox` variable (nitrogen oxides concentration (parts per 10 million)) has the
+greatest positive correlation with the target variable. This is surprising; we
+hadn't thought that the nitrogen oxide concentration would have as large of an
+impact on the crime rate, but some quick googling shows us that there may
+actually be a relationship
+([https://www.sciencedaily.com/releases/2019/10/191003114007.htm](Source)).
 
@@
 
-```{r}
-cv_folds<- vfold_cv(train_df,
-                    v = 10, repeats = 1)
+```{r m2setup}
+cv_folds <- vfold_cv(train_df, v = 10, repeats = 1)
 
-crime_recipe<- 
-  recipe(target ~., data=train_df)
+crime_recipe <- recipe(target ~., data=train_df)
 
-crime_wf<- 
-  workflow() %>%
+crime_wf<- workflow() %>%
   add_recipe(crime_recipe)
 
-crtl_grid<- control_stack_grid()
+crtl_grid <- control_stack_grid()
 ```
@@
 
-```{r}
-logit_specs<- 
+```{r m2svm}
+logit_specs <- 
   logistic_reg(
@@
 
-svm_specs<- 
+svm_specs <- 
   svm_rbf(
@@
 
-Tuning the models for best parameters. The stack function will cherry pick the models that best optimize the blended final model.
+Tuning the models for best parameters. The stack function will cherry pick the
+models that best optimize the blended final model.
 
-```{r}
-svm_wf<-
+```{r m2stack}
+svm_wf <-
   crime_wf %>%
@@
 
-logit_wf<- 
+logit_wf <- 
   crime_wf %>%
@@
 
-logit_results<-
+logit_results <-
   tune_grid(
@@
 )
-
 ```
@@
 
-```{r}
+```{r m2Blend}
 crime_model_stack <- 
@@
 
-```{r}
+```{r m2Predict}
 crime_pred <- 
@@
 
-```{r}
+```{r m2Stats}
 roc<- yardstick::roc_auc(
@@
   conf_mat(truth = target, estimate = .pred_class)
-```
 
-```{r}
 metrics_df

### 2020-10-27 HW3/HW3_Final_AA.Rmd
Review Model 1, Finish Model 3, hide Model 2

@@
     Bryan Persaud, Jeff Shamp
-date: "10/11/2020"
+date: "10/27/2020"
 output:
   pdf_document: default
-
+urlcolor: purple
 ---
@@
 library(caret)
-library(e1071)
 library(geoR)
@@
 DT <- as.data.table(df)
-DT[, target := factor(target)]
-# df <- as_tibble(df)
 eval <- read.csv(paste0(urlRemote, pathGithub, fileTest))
@@
 
-```{r density, fig.width=6}
+```{r density, fig.width=8}
 # density plots
@@
 large lots. Nearly 73% of the observations have a value of 0 with the remaining
-spread widly between 1% and 100%. This may the source of possible distortion. A
+spread widely between 1% and 100%. This may the source of possible distortion. A
 third highly-skewed variable is `chas`, the indicator as to whether the lot
@@
 To provide another view, this time highlighting outliers, the data was analyzed
-using boxplos.
+using boxplots.
 
-```{r boxplots, fig.width=6}
+```{r boxplots, fig.width=8}
 # boxplots
@@
 which represents the perfect correlation of each variable with itself is a blue
-diagional line. The shape and the color scheme reflect the guide at the bottom
+diagonal line. The shape and the color scheme reflect the guide at the bottom
 of the image.
 
-```{r correlation, fig.height=6, fig.width=6}
+```{r correlation, fig.height=5, fig.width=5}
 # correlations
@@
 `nox` has a largest positive correlation with `target` and `dis` has the largest
-negative correlation with `target`. Lastly, it seems that the variable `chas`,
-which indicate whether the suburb borders the Charles River, has statistically
-insignificant correlation with almost all of the other variables.
+negative correlation with `target`.
+
+Perhaps this is a reflection of the noticeable influence that socio-economic
+status has on the crime rate of an area. It is known that crime is concentrated
+in disadvantaged, urban neighborhoods in the United States. The economic
+segregation suggests that affluent neighborhoods may be further away in terms of
+distance, and as a result, disadvantaged areas are more attractive to crime
+because the probability of success is higher even if the targets are not as
+profitable. Consider areas with a higher socio-economics, these areas are
+correlated have a lower crime rate. Moreover, industrial areas are correlated
+with higher levels of nitrogen oxide concentrations. This may be an indication
+of why these areas are less density with residents of a higher status. This
+trend is also seen with the `ptratio` since a higher ratio means less funding
+for public institutions, which is common in areas of lower status.
+
+Lastly, it seems that the variable `chas`, which indicate whether the suburb
+borders the Charles River, has statistically insignificant correlation with
+almost all of the other variables.
 
 # DATA PREPARATION
-There are no missing values so none need to be imputed.
-Different preparation techniques may be used for differen models.
+There are no missing values so none need to be imputed. Different preparation
+techniques may be used for different models.
+
 ## Models 1 & 2
@@
 
-```{r, model1_2_adjust}
+```{r model1_2_adjust}
 new_df <- subset(df, select = -chas)
@@
 
-The Boxcox transformation will be applied to the `age`, `lstat`, and `rm`
+The Box-Cox transformation will be applied to the `age`, `lstat`, and `rm`
 variables.
@@
 
-```{r, m12applyBC}
+```{r m12applyBC}
 new_df$age <- new_df$age ^ ageBC$lambda
@@
 Below are the densities of the variables prior (left) and post (right) the
-Box-Cox transform. The variables `lstat` and `rm` have benefitted from the
+Box-Cox transform. The variables `lstat` and `rm` have benefited from the
 transform, but the strong bimodality of `age` has not been affected
@@
 
-```{r, m12BC_plot}
+```{r m12BC_plot}
 dens_age <- ggplot(new_df, aes(age)) + geom_density()
@@
 ## Model 3
+For model 3, very little data manipulation will be done. There may be valuable
+information in the outliers and a reasonable model may be able to extract
+something from `chas`. However, `chas` will be converted to a factor. Similarly,
+`rad` is an index of accessibility, so it is almost certainly a factor, although
+probably ordered. Given there are now two factors in the predictors, dummy
+dummy variables need to be created. Lastly, the **labels** of `target` will be
+changed from integer to character for compliance with some of R's naming
+conventions. Based on the definitions, 0 will be mapped to *below* and 1 to
+*above*.
+
 # BUILD MODELS
-## Models 1 & 2
 First, we will split the training data into two sets. This will allow us to
@@
 # train-test split
-crime_df <- as.tibble(df) %>%
+
+crime_df <- new_df %>%
   mutate(target = as.factor(target)) %>% 
@@
 
+# This is so Model 3 can use the same train/test split as Models 1 and 2
+crime_df$id <- seq_len(nrow(crime_df))
+
+# Changing the text above caused changes in output as fewer command were run. It
+# is ALWAYS a good idea to set a seed IMMEDIATELY before any sampling in a
+# Markdown file to ensure repeatability.
+
+set.seed(7104)
 data_split <- initial_split(crime_df, 
@@
                             prop = 0.8)
-train_df<- training(data_split)
-test_df<- testing(data_split)
+train_df <- training(data_split)
+test_df <- testing(data_split)
+
+# Capture the indices and remove the no-longer-necessary column
+tstIDX <- test_df$id
+crime_df <- subset(crime_df, select = -id)
+train_df <- subset(train_df, select = -id)
+test_df <- subset(test_df, select = -id)
 ```
@@
 baseModel <- glm(target~., family = binomial, data = train_df)
-knitr::kable(summary(baseModel)$coefficients, digits = 3L,
+kable(summary(baseModel)$coefficients, digits = 3L,
       caption = 'Base Model Logistic Regression Output')
@@
                family=binomial())
-knitr::kable(summary(model1)$coefficients, digits = 3L,
+kable(summary(model1)$coefficients, digits = 3L,
       caption = 'Backwards Model Regression Output')
@@
 coefficients for the variables `zn` and `tax` are negative. This is indicative
-of an inverse relationship; ie, the higher these values, the less likely the
+of an inverse relationship; i.e., the higher these values, the less likely the
 crime rate is above the median. The relationship for the zone variable aligns
@@
 ## Model 2
-### Stacked Classifier
+<!-- ### Stacked Classifier -->
 
-```{r m2setup}
-cv_folds <- vfold_cv(train_df, v = 10, repeats = 1)
+<!-- ```{r m2setup} -->
+<!-- cv_folds <- vfold_cv(train_df, v = 10, repeats = 1) -->
 
-crime_recipe <- recipe(target ~., data=train_df)
+<!-- crime_recipe <- recipe(target ~., data=train_df) -->
 
-crime_wf<- workflow() %>%
-  add_recipe(crime_recipe)
+<!-- crime_wf<- workflow() %>% -->
+<!--   add_recipe(crime_recipe) -->
 
-crtl_grid <- control_stack_grid()
+<!-- crtl_grid <- control_stack_grid() -->
+<!-- ``` -->
+
+<!-- Specify the models. Using SVM as another model to blend with the logistic.  -->
+
+<!-- ```{r m2svm} -->
+<!-- logit_specs <-  -->
+<!--   logistic_reg( -->
+<!--     penalty = tune(), -->
+<!--     mixture = tune() -->
+<!--     ) %>% -->
+<!--   set_engine("glm") %>% -->
+<!--   set_mode("classification") -->
+
+<!-- svm_specs <-  -->
+<!--   svm_rbf( -->
+<!--     cost = tune(), -->
+<!--     rbf_sigma = tune(), -->
+<!--     margin = tune() -->
+<!--   ) %>% -->
+<!--   set_mode("classification") %>% -->
+<!--   set_engine("kernlab") -->
+<!-- ``` -->
+
+<!-- Tuning the models for best parameters. The stack function will cherry pick the -->
+<!-- models that best optimize the blended final model.  -->
+
+<!-- ```{r m2stack} -->
+<!-- svm_wf <- -->
+<!--   crime_wf %>% -->
+<!--   add_model(svm_specs) -->
+
+<!-- svm_results<- -->
+<!--   tune_grid( -->
+<!--     svm_wf,  -->
+<!--     resample = cv_folds, -->
+<!--     control = crtl_grid, -->
+<!--     grid = 10, -->
+<!--     save_pred = TRUE, -->
+<!--     save_workflow = FALSE -->
+<!--   ) -->
+
+<!-- logit_wf <-  -->
+<!--   crime_wf %>% -->
+<!--   add_model(logit_specs) -->
+
+<!-- logit_results <- -->
+<!--   tune_grid( -->
+<!--     logit_wf, -->
+<!--     resamples = cv_folds, -->
+<!--     control = crtl_grid, -->
+<!--     grid = 10, -->
+<!--     save_pred = TRUE, -->
+<!--     save_workflow = FALSE -->
+<!-- ) -->
+<!-- ``` -->
+
+<!-- Blend! -->
+
+<!-- ```{r m2Blend} -->
+<!-- crime_model_stack <- -->
+<!--   stacks() %>% -->
+<!--   add_candidates(logit_results) %>% -->
+<!--   add_candidates(svm_results) %>% -->
+<!--   blend_predictions() %>% -->
+<!--   fit_members() -->
+
+<!-- crime_model_stack -->
+<!-- ``` -->
+
+<!-- ```{r m2Predict} -->
+<!-- crime_pred <- -->
+<!--   test_df %>% -->
+<!--   bind_cols(predict(crime_model_stack, ., type = "prob"), -->
+<!--             predict(crime_model_stack, ., type = "class")) -->
+<!-- ``` -->
+
+<!-- Computing the metrics for the stacked model: -->
+
+<!-- ```{r m2Stats} -->
+<!-- roc<- yardstick::roc_auc( -->
+<!--   crime_pred, -->
+<!--   truth = target, -->
+<!--   contains(".pred_1") -->
+<!--   ) -->
+
+<!-- acc<- yardstick::accuracy( -->
+<!--   crime_pred, -->
+<!--   truth = target, -->
+<!--   estimate = .pred_class -->
+<!-- ) -->
+
+<!-- recall<- yardstick::recall( -->
+<!--   crime_pred, -->
+<!--   truth = target, -->
+<!--   estimate = .pred_class -->
+<!-- ) -->
+
+<!-- precise<- yardstick::precision( -->
+<!--   crime_pred, -->
+<!--   truth = target, -->
+<!--   estimate = .pred_class -->
+<!-- ) -->
+<!-- metrics_df<- -->
+<!--   bind_rows(roc, acc, recall, precise) -->
+
+<!-- crime_pred %>% -->
+<!--   conf_mat(truth = target, estimate = .pred_class) -->
+
+<!-- metrics_df -->
+<!-- ``` -->
+
+## Model 3
+```{r m3trainTemp}
+DT_train <- DT[!(IDX %in% tstIDX)]
 ```
 
-Specify the models. Using SVM as another model to blend with the logistic.
+Model 3 will be a binary logistic regression allowing for all variables and
+selected pair-wise interactions pruned using both forward and backward stepwise
+regression based on AICc to select the optimal parameters. The captured indicies
+of the test set for models 1 & 2 will be used to split the data for model 3 as
+well. To address the possibility of overfitting, five-fold cross-validation will
+be done on the training set. With the vanilla logistic regression model in R,
+`glm` with a binomial family and a logit link, there are no other
+hyperparameters available for tuning. More sophisticated logistic regression
+implementation have these parameters for fine-tuning.
 
-```{r m2svm}
-logit_specs <- 
-  logistic_reg(
-    penalty = tune(),
-    mixture = tune()
-    ) %>%
-  set_engine("glm") %>%
-  set_mode("classification")
+To consider *all* interactions would be foolhardy. Given the need for dummy
+variables, there will 19 predictors and so \(_{19}C_2 = 171\) possible pairs.
+This is too much. Therefore, intentional selection of pairs is required.
+Looking at the predictors, there are some which logically may interact with
+each other, such as:
 
-svm_specs <- 
-  svm_rbf(
-    cost = tune(),
-    rbf_sigma = tune(),
-    margin = tune()
-  ) %>%
-  set_mode("classification") %>%
-  set_engine("kernlab")
+  * `indus` and `nox`
+    * Nitrous Oxide is a common industrial pollutant. There is a high
+    correlation in the training set between the two. Allowing for its
+    interaction may temper the result of their relationship without losing more
+    information
+
+What may prove interesting is to look at some unsupervised clustering techniques
+to provide us with insights as to which variables may be "close" enough to
+warrant interactions. Instead of treating the observations as samples, we will
+treat the predictors as samples!
+
+```{r m3Clust, fig.height=6}
+predClust <- hclust(dist(t(DT_train[, -c('target', 'IDX')]), diag = TRUE))
+plot(predClust, xlab = "Raw Predictors")
 ```
 
-Tuning the models for best parameters. The stack function will cherry pick the
-models that best optimize the blended final model.
+It seems that there may be reason to consider an interaction between `chas` and
+`nox`. Can it be that many industrial plants are on the river? Similarly with
+`indus` and `lstat`. Higher-cost homes tend to be further away from the traffic
+and pollution of industrial areas.
 
-```{r m2stack}
-svm_wf <-
-  crime_wf %>%
-  add_model(svm_specs)
+Our selected starting model will include the following interactions:
 
-svm_results<-
-  tune_grid(
-    svm_wf, 
-    resample = cv_folds,
-    control = crtl_grid,
-    grid = 10,
-    save_pred = TRUE,
-    save_workflow = FALSE
-  )
+  * indus:nox
+  * indus:lstat
+  * chas:nox
+  * ptratio:medv
+  * rm:dis
 
-logit_wf <- 
-  crime_wf %>%
-  add_model(logit_specs)
+### Train Model
 
-logit_results <-
-  tune_grid(
-    logit_wf,
-    resamples = cv_folds,
-    control = crtl_grid,
-    grid = 10,
-    save_pred = TRUE,
-    save_workflow = FALSE
-)
+```{r m3dataPrep}
+DT[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
+          rad = factor(rad),
+          target = factor(target, labels = c('Below', 'Above')))]
+DT_test <- DT[IDX %in% tstIDX]
+DT_train <- DT[!(IDX %in% tstIDX)]
+```
+```{r m3build}
+trC <- trainControl(method = 'cv', number = 5, classProbs = TRUE,
+                    summaryFunction = twoClassSummary)
+dVars <- dummyVars(target ~ indus * nox +
+                     indus * lstat +
+                     chas * nox +
+                     ptratio * medv +
+                     rm * dis + . - IDX, data = DT_train, fullRank = TRUE)
+DTtrnx <- predict(dVars, newdata = DT_train)
+DTtrny <- DT_train$target
+set.seed(1)
+m3Fit <- train(x = DTtrnx, y = DTtrny, method = 'glmStepAIC', trControl = trC,
+               family = 'binomial', trace = 0, direction = 'both')
 ```
 
-Blend!
-
-```{r m2Blend}
-crime_model_stack <- 
-  stacks() %>%
-  add_candidates(logit_results) %>%
-  add_candidates(svm_results) %>%
-  blend_predictions() %>%
-  fit_members()
-
-crime_model_stack
+### Coefficient Discussion
+```{r m3Sum}
+kable(summary(m3Fit$finalModel)$coefficients, digits = 3L,
+             caption = 'Stepwise Interaction-Allowed Logistic Regression Output')
 ```
 
-```{r m2Predict}
-crime_pred <- 
-  test_df %>%
-  bind_cols(predict(crime_model_stack, ., type = "prob"), 
-            predict(crime_model_stack, ., type = "class"))
+The selected model has some interesting results. Firstly, the predictor with the
+greatest effect is `nox`. Further, the only model variable which serves to
+**decrease** the crime rate from the baseline is `rm`. Other variables, such as
+`chas`
+
+The interactions between `indus` and both `nox` and `lstat` are significant. As
+theorized, these interactions serve to temper their individual contributions to
+crime. Remember, a 1 indicates crime above the median, so positive coefficients
+indicate contributions to *increasing* the crime rate. A similar observation
+holds `rm`. While `dis` on its own did not decrease the model's AICc, its
+interaction to temper `rm` did help the model's deviance. No other selected
+interactions were left in the model.
+
+A number of variables fell out of the model, such as `chas`, `dis`, and `zn`
+amobg others. The remaining predictors are all significant, at least at the 10%
+level, except for `rad.24`. However, from the last trace of the stepping routine
+shown below, removing it would ***drastically*** reduce the models performance.
+Lastly, while `rad.1` is the base case, `rad.2` has disappeared as well,
+implying that `rad = 2` should be considered indistinguishable from `rad = 1`. 
 ```
-
-Computing the metrics for the stacked model:
-
-```{r m2Stats}
-roc<- yardstick::roc_auc(
-  crime_pred,
-  truth = target,
-  contains(".pred_1")
-  )
-
-acc<- yardstick::accuracy(
-  crime_pred, 
-  truth = target, 
-  estimate = .pred_class
-)
-
-recall<- yardstick::recall(
-  crime_pred,
-  truth = target, 
-  estimate = .pred_class
-)
-
-precise<- yardstick::precision(
-  crime_pred,
-  truth = target, 
-  estimate = .pred_class
-)
-metrics_df<-
-  bind_rows(roc, acc, recall, precise)
-
-crime_pred %>%
-  conf_mat(truth = target, estimate = .pred_class)
-
-metrics_df
+                    Df Deviance    AIC
+<none>                   90.565 122.56
++ dis                1   89.191 123.19
++ zn                 1   89.558 123.56
++ tax                1   89.700 123.70
+- ptratio            1   93.763 123.76
++ age                1   90.156 124.16
++ `nox:chasOnRiver`  1   90.224 124.22
+- `rm:dis`           1   94.254 124.25
++ chas.OnRiver       1   90.273 124.27
++ rad.6              1   90.315 124.31
+- lstat              1   94.364 124.36
++ `ptratio:medv`     1   90.494 124.49
++ rad.2              1   90.554 124.55
+- rm                 1   95.001 125.00
+- `indus:lstat`      1   95.633 125.63
+- `indus:nox`        1   97.351 127.35
+- indus              1   97.597 127.60
+- medv               1   98.568 128.57
+- rad.3              1  101.594 131.59
+- rad.5              1  105.594 135.59
+- rad.7              1  118.832 148.83
+- nox                1  134.786 164.79
+- rad.8              1  149.967 179.97
+- rad.4              1  152.757 182.76
+- rad.24             1  152.797 182.80
 ```
@@
 ## Criteria
+In binary classification, often, the measures of focus are precision and recall,
+as opposed to accuracy. A measure combining this is the F1 score, defined as
+twice the sum of precision and recall divided by their product. Another oft-used
+metric is the area under the receiver operating curve, AUC. We will look at
+the three metrics of accuracy, F1, and AUC, and select the model that performs
+best in at least two of the three metrics.
 
 ## Performance 
+```{r comparison}
+compTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
+                        ACC = double(3),
+                        F1 = double(3),
+                        AUC = double(3))
+```
 
@@
 ```{r model1_perf}
-
-# SELECT MODELS
 # Model 1 Performance
@@
 
-####### metrics
-m1_roc<- yardstick::roc_auc(
+# Metrics
+m1_roc <- yardstick::roc_auc(
   model1_results,
@@
 
-m1_acc<- yardstick::accuracy(
+m1_acc <- yardstick::accuracy(
   model1_results, 
@@
 
-m1_recall<- yardstick::recall(
+m1_recall <- yardstick::recall(
   model1_results,
@@
 
-m1_precise<- yardstick::precision(
+m1_precise <- yardstick::precision(
   model1_results,
@@
   bind_rows(m1_roc, m1_acc, m1_recall, m1_precise)
-
 m1_metrics_df
-
+m1CM <- confusionMatrix(model1_preds, model1_results$target,
+                        mode = 'prec_recall')
+compTable[1, 2] <- m1_acc$.estimate
+compTable[1, 3] <- 2 * m1_precise$.estimate * m1_recall$.estimate /
+  (m1_precise$.estimate + m1_recall$.estimate)
+compTable[1, 4] <- m1_roc$.estimate
+m1CM$table
 ```
 
-### Sub-Models and Coefficient Discussion
+### Model 2
 
-The blended model is comprised of two sub-models, the logistic and the SVM. Below are the prediction results for each sub-model.
+<!-- The blended model is comprised of two sub-models, the logistic and the SVM. -->
+<!-- Below are the prediction results for each sub-model.  -->
 
-```{r}
-best_model_metric <- select_best(svm_results, "accuracy")
-finalize_workflow(svm_wf, best_model_metric) %>%
-  last_fit(data_split) %>%
-  collect_metrics()
+<!-- ```{r} -->
+<!--  best_model_metric <- select_best(svm_results, "accuracy") -->
+<!--  finalize_workflow(svm_wf, best_model_metric) %>% -->
+<!--    last_fit(data_split) %>% -->
+<!--    collect_metrics() -->
+
+<!--  logit_wf %>% -->
+<!--    last_fit(data_split) %>% -->
+<!--    collect_metrics() -->
+
+<!--  logit_wf %>% -->
+<!--    fit(data = train_df) %>% -->
+<!--    pull_workflow_fit() %>% -->
+<!--    vip(geom = "col", aesthetics = list(fill='red4')) -->
+<!-- ``` -->
+
+<!-- We see that they are both in the low eighties percent for accuracy, and low -->
+<!-- nineties for roc auc. Together make for a better overall prediction.  -->
+
+<!-- We again see that the primary predictor in the logistic regression is the `nox` -->
+<!-- with `rad` and `ptratio` rounding out the top three predictors. This is in line -->
+<!-- with the base model. Living close to the center of town, having teachers, and -->
+<!-- not having polluted air are the largest predictors in terms of above median crime.  -->
+
+### Model 3
+
+```{r m3TestSet}
+tstdVars <- dummyVars(target ~ indus * nox +
+                        indus * lstat +
+                        chas * nox +
+                        ptratio * medv +
+                        rm * dis + . - IDX, data = DT_test, fullRank = TRUE)
+DTtstx <- predict(tstdVars, newdata = DT_test)
+DTtsty <- DT_test$target
+m3Pred <- predict(m3Fit, newdata = DTtstx)
+m3PredP <- predict(m3Fit, newdata = DTtstx, type = 'prob')
+m3CM <- confusionMatrix(m3Pred, DTtsty, mode = 'prec_recall')
+m3CM$table
+compTable[3, 2] <- sum(diag(m3CM$table)) / sum(m3CM$table)
+compTable[3, 3] <- m3CM$byClass[7]
+compTable[3, 4] <- yardstick::roc_auc_vec(DTtsty, m3PredP[, 1])
+kable(compTable, digits = 3L, caption = 'Model Test Results')
 ```
-
-```{r}
-logit_wf %>%
-  last_fit(data_split) %>%
-  collect_metrics()
-
-logit_wf %>%
-  fit(data = train_df) %>%
-  pull_workflow_fit() %>%
-  vip(geom = "col", aesthetics = list(fill='red4'))
-```
-
-We see that they are both in the low eighties percent for accuracy, and low nineties for roc auc. Together make for a better overall prediction.
-
-We again see that the primary predictor in the logistic regression is the `nox` with `rad` and `ptratio` rounding out the top three predictors.
-This is in line with the base model.
-Living close to the center of town, having teachers, and not having polluted air are the largest predictors in terms of above median crime.
 
@@
 ```
+```{r model1_2_adjust}
+```
+```{r m12BC}
+```
+```{r m12applyBC}
+```
+```{r m12BC_plot}
+```
+```{r model1_coeff}
+```
+```{r m3trainTemp}
+```
+```{r m3Clust}
+```
+```{r m3dataPrep}
+```
+```{r m3Sum}
+```
+```{r comparison}
+```
+```{r model1_perf}
+```
+```{r m3TestSet}
+```

### 2020-10-27 HW3/HW3_Final_AA.Rmd
:DOH:

@@
 DT <- as.data.table(df)
+DT[, target := factor(target)]
 eval <- read.csv(paste0(urlRemote, pathGithub, fileTest))

### 2020-10-27 HW3/HW3_Final_AA.Rmd
It's too late.

@@
 shown below, removing it would ***drastically*** reduce the models performance.
-Lastly, while `rad.1` is the base case, `rad.2` has disappeared as well,
-implying that `rad = 2` should be considered indistinguishable from `rad = 1`. 
+Lastly, while `rad.1` is the base case, `rad.2` and `rad.6` have disappeared as
+well. This indicates that both `rad = 2` and `rad = 6` should be considered
+indistinguishable from `rad = 1`. 
 ```

### 2020-10-27 HW3/HW3_Final_AA.Rmd
add model 2, remove stacked

@@
 ## Model 2
-<!-- ### Stacked Classifier -->
+### Logistic Classifier
 
-<!-- ```{r m2setup} -->
-<!-- cv_folds <- vfold_cv(train_df, v = 10, repeats = 1) -->
+```{r m2setup}
+cv_folds <- vfold_cv(train_df, v = 10, repeats = 1)
 
-<!-- crime_recipe <- recipe(target ~., data=train_df) -->
+crime_recipe <- recipe(target ~., data=train_df)
 
-<!-- crime_wf<- workflow() %>% -->
-<!--   add_recipe(crime_recipe) -->
+crime_wf<- workflow() %>%
+  add_recipe(crime_recipe)
+```
 
-<!-- crtl_grid <- control_stack_grid() -->
-<!-- ``` -->
+```{r m2svm}
+logit_specs <-
+  logistic_reg(
+    penalty = tune(),
+    mixture = tune()
+    ) %>%
+  set_engine("glm") %>%
+  set_mode("classification")
 
-<!-- Specify the models. Using SVM as another model to blend with the logistic.  -->
+```
 
-<!-- ```{r m2svm} -->
-<!-- logit_specs <-  -->
-<!--   logistic_reg( -->
-<!--     penalty = tune(), -->
-<!--     mixture = tune() -->
-<!--     ) %>% -->
-<!--   set_engine("glm") %>% -->
-<!--   set_mode("classification") -->
+Tuning the models for best parameters.
 
-<!-- svm_specs <-  -->
-<!--   svm_rbf( -->
-<!--     cost = tune(), -->
-<!--     rbf_sigma = tune(), -->
-<!--     margin = tune() -->
-<!--   ) %>% -->
-<!--   set_mode("classification") %>% -->
-<!--   set_engine("kernlab") -->
-<!-- ``` -->
+```{r m2stack}
+logit_wf <-
+  crime_wf %>%
+  add_model(logit_specs)
 
-<!-- Tuning the models for best parameters. The stack function will cherry pick the -->
-<!-- models that best optimize the blended final model.  -->
+logit_results <-
+  tune_grid(
+    logit_wf,
+    resamples = cv_folds,
+    control = crtl_grid,
+    grid = 10,
+    save_pred = TRUE,
+    save_workflow = FALSE
+)
+```
 
-<!-- ```{r m2stack} -->
-<!-- svm_wf <- -->
-<!--   crime_wf %>% -->
-<!--   add_model(svm_specs) -->
+```{r m2predict}
+best_model_auc <- select_best(logit_results, "roc_auc")
+final_auc<-
+  finalize_workflow(
+    logit_wf,
+    best_model_auc
+)
 
-<!-- svm_results<- -->
-<!--   tune_grid( -->
-<!--     svm_wf,  -->
-<!--     resample = cv_folds, -->
-<!--     control = crtl_grid, -->
-<!--     grid = 10, -->
-<!--     save_pred = TRUE, -->
-<!--     save_workflow = FALSE -->
-<!--   ) -->
+final_model_pred<-
+  final_auc %>%
+  last_fit(data_split) %>%
+  collect_predictions()
 
-<!-- logit_wf <-  -->
-<!--   crime_wf %>% -->
-<!--   add_model(logit_specs) -->
+```
 
-<!-- logit_results <- -->
-<!--   tune_grid( -->
-<!--     logit_wf, -->
-<!--     resamples = cv_folds, -->
-<!--     control = crtl_grid, -->
-<!--     grid = 10, -->
-<!--     save_pred = TRUE, -->
-<!--     save_workflow = FALSE -->
-<!-- ) -->
-<!-- ``` -->
+Computing the metrics for the logit model:
 
-<!-- Blend! -->
+```{r m2Stats}
+roc<- yardstick::roc_auc(
+  final_model_pred,
+  truth = target,
+  contains(".pred_1")
+  )
 
-<!-- ```{r m2Blend} -->
-<!-- crime_model_stack <- -->
-<!--   stacks() %>% -->
-<!--   add_candidates(logit_results) %>% -->
-<!--   add_candidates(svm_results) %>% -->
-<!--   blend_predictions() %>% -->
-<!--   fit_members() -->
+acc<- yardstick::accuracy(
+  final_model_pred, 
+  truth = target, 
+  estimate = .pred_class
+)
 
-<!-- crime_model_stack -->
-<!-- ``` -->
+recall<- yardstick::recall(
+  final_model_pred,
+  truth = target,
+  estimate = .pred_class
+)
 
-<!-- ```{r m2Predict} -->
-<!-- crime_pred <- -->
-<!--   test_df %>% -->
-<!--   bind_cols(predict(crime_model_stack, ., type = "prob"), -->
-<!--             predict(crime_model_stack, ., type = "class")) -->
-<!-- ``` -->
+precise<- yardstick::precision(
+  final_model_pred,
+  truth = target,
+  estimate = .pred_class
+)
+metrics_df<-
+  bind_rows(roc, acc, recall, precise)
 
-<!-- Computing the metrics for the stacked model: -->
-
-<!-- ```{r m2Stats} -->
-<!-- roc<- yardstick::roc_auc( -->
-<!--   crime_pred, -->
-<!--   truth = target, -->
-<!--   contains(".pred_1") -->
-<!--   ) -->
-
-<!-- acc<- yardstick::accuracy( -->
-<!--   crime_pred, -->
-<!--   truth = target, -->
-<!--   estimate = .pred_class -->
-<!-- ) -->
-
-<!-- recall<- yardstick::recall( -->
-<!--   crime_pred, -->
-<!--   truth = target, -->
-<!--   estimate = .pred_class -->
-<!-- ) -->
-
-<!-- precise<- yardstick::precision( -->
-<!--   crime_pred, -->
-<!--   truth = target, -->
-<!--   estimate = .pred_class -->
-<!-- ) -->
-<!-- metrics_df<- -->
-<!--   bind_rows(roc, acc, recall, precise) -->
-
-<!-- crime_pred %>% -->
-<!--   conf_mat(truth = target, estimate = .pred_class) -->
-
-<!-- metrics_df -->
-<!-- ``` -->
+crime_pred %>%
+  conf_mat(truth = target, estimate = .pred_class)
+```
 
@@
 
-<!-- The blended model is comprised of two sub-models, the logistic and the SVM. -->
-<!-- Below are the prediction results for each sub-model.  -->
+Below the feature importances for the logistic model.
 
-<!-- ```{r} -->
-<!--  best_model_metric <- select_best(svm_results, "accuracy") -->
-<!--  finalize_workflow(svm_wf, best_model_metric) %>% -->
-<!--    last_fit(data_split) %>% -->
-<!--    collect_metrics() -->
+```{r}
+logit_wf %>%
+  last_fit(data_split) %>%
+  collect_metrics()
 
-<!--  logit_wf %>% -->
-<!--    last_fit(data_split) %>% -->
-<!--    collect_metrics() -->
+logit_wf %>%
+  fit(data = train_df) %>%
+  pull_workflow_fit() %>%
+  vip(geom = "col", aesthetics = list(fill='red4'))
+```
 
-<!--  logit_wf %>% -->
-<!--    fit(data = train_df) %>% -->
-<!--    pull_workflow_fit() %>% -->
-<!--    vip(geom = "col", aesthetics = list(fill='red4')) -->
-<!-- ``` -->
-
-<!-- We see that they are both in the low eighties percent for accuracy, and low -->
-<!-- nineties for roc auc. Together make for a better overall prediction.  -->
-
-<!-- We again see that the primary predictor in the logistic regression is the `nox` -->
-<!-- with `rad` and `ptratio` rounding out the top three predictors. This is in line -->
-<!-- with the base model. Living close to the center of town, having teachers, and -->
-<!-- not having polluted air are the largest predictors in terms of above median crime.  -->
+We again see that the primary predictor in the logistic regression is the `nox`
+with `rad` and `ptratio` rounding out the top three predictors. This is in line
+with the base model. Living close to the center of town, having teachers, and
+not having polluted air are the largest predictors in terms of above median crime.
 

### 2020-10-27 HW3/HW3_Final_AA.Rmd
remove unneded libraries, add crtl_grid

@@
 library(data.table)
-library(stacks)
-library(kernlab)
 library(vip)
@@
 
+crtl_grid<-control_grid(verbose = FALSE)
+
 logit_results <-

### 2020-10-27 HW3/HW3_Final_AA.Rmd
more updates to help clean up work

@@
 
+We will resample the data to tune parameters for the logistic classifier. This uses all variable as individual predictors.
+
 ```{r m2setup}
@@
 ```
-
-Tuning the models for best parameters.
 
@@
 
-Computing the metrics for the logit model:
-
 ```{r m2Stats}
@@
 
-crime_pred %>%
-  conf_mat(truth = target, estimate = .pred_class)
 ```
@@
 ```{r}
-logit_wf %>%
-  last_fit(data_split) %>%
-  collect_metrics()
+final_model_pred %>%
+  conf_mat(truth = target, estimate = .pred_class)
+
+metrics_df
 

### 2020-10-27 HW3/HW3_Final_AA.Rmd
Update HW3_Final_AA.Rmd

--- 2020-10-27 HW3/HW3_Final_AA.Rmd
@@
 # BUILD MODELS
+
 First, we will split the training data into two sets. This will allow us to
@@
 ```{r m2Stats}
-roc<- yardstick::roc_auc(
+m2_roc<- yardstick::roc_auc(
   final_model_pred,
@@
 
-acc<- yardstick::accuracy(
+m2_acc<- yardstick::accuracy(
   final_model_pred, 
@@
 
-recall<- yardstick::recall(
+m2_recall<- yardstick::recall(
   final_model_pred,
@@
 
-precise<- yardstick::precision(
+m2_precise<- yardstick::precision(
   final_model_pred,
@@
 )
-metrics_df<-
-  bind_rows(roc, acc, recall, precise)
+m2_metrics_df<-
+  bind_rows(m2_roc, m2_acc, m2_recall, m2_precise)
+```
 
-crime_pred %>%
-  conf_mat(truth = target, estimate = .pred_class)
+```{r}
+logit_wf %>%
+  fit(data = train_df) %>%
+  pull_workflow_fit() %>%
+  vip(geom = "col", aesthetics = list(fill='red4'))
 ```
+
+We again see that the primary predictor in the logistic regression is the `nox`
+with `rad` and `ptratio` rounding out the top three predictors. This is in line
+with the base model. Living close to the center of town, having teachers, and
+not having polluted air are the largest predictors in terms of above median crime.
 
@@
 of the test set for models 1 & 2 will be used to split the data for model 3 as
-well. To address the possibility of overfitting, five-fold cross-validation will
+well. To address the possibility of over-fitting, five-fold cross-validation will
 be done on the training set. With the vanilla logistic regression model in R,
@@
 ## Performance 
+Below the feature importance for all the models.
+
 ```{r comparison}
@@
   bind_rows(m1_roc, m1_acc, m1_recall, m1_precise)
-m1_metrics_df
+
 m1CM <- confusionMatrix(model1_preds, model1_results$target,
@@
 
-Below the feature importances for the logistic model.
+```{r m2TestSet}
+final_model_pred %>%
+  conf_mat(truth = target, estimate = .pred_class)
 
-```{r}
-logit_wf %>%
+m2.metrics <- logit_wf %>%
   last_fit(data_split) %>%
@@
 
-logit_wf %>%
-  fit(data = train_df) %>%
-  pull_workflow_fit() %>%
-  vip(geom = "col", aesthetics = list(fill='red4'))
+compTable[2, 2] <- m2.metrics[1,3]
+compTable[2, 3] <- 2 * m2_precise$.estimate * m2_recall$.estimate /
+  (m2_precise$.estimate + m2_recall$.estimate)
+compTable[2, 4] <- m2.metrics[2,3]
 ```
-
-We again see that the primary predictor in the logistic regression is the `nox`
-with `rad` and `ptratio` rounding out the top three predictors. This is in line
-with the base model. Living close to the center of town, having teachers, and
-not having polluted air are the largest predictors in terms of above median crime.
 
@@
 The code chunks below represent the R code called in order during the analysis. They are reproduced in the appendix for review and comment.
+
 ```{r appendix, include=FALSE}
@@
 ```
+```{r m2setup}
+```
+```{r m2svm}
+```
+```{r m2stack}
+```
+```{r m2predict}
+```
+```{r m2Stats}
+```
 ```{r m3trainTemp}
@@
 ```
+```{r m2TestSet}
+```
 ```{r m3TestSet}

### 2020-10-28 HW3/HW3_Final_AA.Rmd
Copyedit and cleanup

@@
 ## Models 1 & 2
-The variable `chas` will be removed due to having statistically insignificant\
+The variable `chas` will be removed due to having statistically insignificant
 correlation with almost all the variables in the dataset. Next, outliers will be
@@
 dens_rmO <- ggplot(new_df, aes(rm)) + geom_density()
-
 ```
@@
 # BUILD MODELS
-
 First, we will split the training data into two sets. This will allow us to
@@
 of the test set for models 1 & 2 will be used to split the data for model 3 as
-well. To address the possibility of over-fitting, five-fold cross-validation will
-be done on the training set. With the vanilla logistic regression model in R,
-`glm` with a binomial family and a logit link, there are no other
-hyperparameters available for tuning. More sophisticated logistic regression
-implementation have these parameters for fine-tuning.
+well. With the vanilla logistic regression model in R, `glm` with a binomial
+family and a logit link, there are no other hyperparameters available for
+tuning. More sophisticated logistic regression implementation have these
+parameters for fine-tuning. There is also no reasonable way to implement
+cross-validation so none will be done.
 
@@
 ```{r m3build}
-trC <- trainControl(method = 'cv', number = 5, classProbs = TRUE,
+trC <- trainControl(method = 'none', classProbs = TRUE,
                     summaryFunction = twoClassSummary)
@@
 ## Performance 
-Below the feature importance for all the models.
+Below are the confusion matrices for the three models, followed by a table with
+the selected performance.
 
@@
 compTable[3, 4] <- yardstick::roc_auc_vec(DTtsty, m3PredP[, 1])
+```
+
+### Selected Performance Statistics
+```{r selperfstat}
 kable(compTable, digits = 3L, caption = 'Model Test Results')
@@
 # APPENDIX
-The code chunks below represent the R code called in order during the analysis. They are reproduced in the appendix for review and comment.
+The code chunks below represent the R code called in order during the analysis.
+They are reproduced in the appendix for review and comment.
 
@@
 ```
+```{r selperfstat}
+```

### 2020-10-28 HW3/HW3_Final_AA.Rmd
Added prediction section.

@@
 A number of variables fell out of the model, such as `chas`, `dis`, and `zn`
-amobg others. The remaining predictors are all significant, at least at the 10%
+among others. The remaining predictors are all significant, at least at the 10%
 level, except for `rad.24`. However, from the last trace of the stepping routine
@@
 
-compTable[2, 2] <- m2.metrics[1,3]
+compTable[2, 2] <- m2.metrics[1, 3]
 compTable[2, 3] <- 2 * m2_precise$.estimate * m2_recall$.estimate /
   (m2_precise$.estimate + m2_recall$.estimate)
-compTable[2, 4] <- m2.metrics[2,3]
+compTable[2, 4] <- m2.metrics[2, 3]
 ```
@@
 kable(compTable, digits = 3L, caption = 'Model Test Results')
+```
+
+All the models performed very well and classified the majority of the test set
+properly. The difference between the three models was very slight. Only 7, 6, 
+and 4 observations out of 92 were misclassified by the three models
+respectively. Given the above statistics and confusion matrices, **Model 3**
+barely edged out the other two and will be used to predict on the `eval` set.
+
+## Predictions
+
+```{r evalPred}
+setDT(eval)
+neval <- dim(eval)[[1]]
+eval[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
+          rad = factor(rad))]
+evalVars <- dummyVars( ~ indus * nox + indus * lstat + chas * nox + 
+                         ptratio * medv + rm * dis + .,
+                       data = eval, fullRank = TRUE)
+evaltstx <- predict(evalVars, newdata = eval)
+evalPred <- predict(m3Fit, newdata = evaltstx)
+evalPredP <- predict(m3Fit, newdata = evaltstx, type = 'prob')
+predTable <- data.frame(TestID = seq_len(neval),
+                        PredictedClass = ifelse(evalPred == "Below", 0, 1),
+                        Prob_Is_0 = evalPredP$Below,
+                        Prob_Is_1 = evalPredP$Above)
+kable(predTable, digits = 3L, caption = "Predicted Classes and Probabilities")
 ```
@@
 ```
+```{r evalPred}
+```

### 2020-10-28 HW3/HW3_Final_AA.Rmd
Update HW3_Final_AA.Rmd

@@
 
-This data set has complete cases, thus there is no need for imputation. Based on
-some common summary statistics, there are more observations where the crime rate
-is below the median. It is already apparent that some of the predictors varies
-depending the crime rate. For instance, there is a noticeable difference in the
-means of `age`, `lstat`, `rad`, and `zn` between the crime rate groups.
+We can make some initial observaations. The data set has complete cases, thus, 
+there is no need for imputation. Based on some common summary statistics, 
+there are more observations where the crime rate is below the median. It is already 
+apparent that some of the predictors varies depending the crime rate. For instance, 
+there is a noticeable difference in the means of `age`, `lstat`, `rad`, and `zn` between 
+the crime rate groups.
 
@@
 ## Data Distribution
-For each predictors, we computed kernel density estimators to understand their
-distribution. The following plots show how predictors are distributed between
-areas where the crime rate is higher than the median (blue) and areas where the
-crime rate is below the median (red). It is of interest to understand variables
-that highlight large variations between the two groups.
+
+For each predictors, we computed and drew kernel density estimate to understand 
+their distribution. Using density plots, we show how predictors are distributed 
+between areas where crime rate is higher than the median, i.e. blue, and areas 
+where crime rate is below the median, i.e. red. Our focus is to understand 
+variables that highlight large variations between the two groups.
 
@@
 
-The density plots reveal that most of the data does not have a normal
-distribution. There are some are heavily right tailed variables and others which
-are multi-modal. The most Gaussian of the variables appears to be the one
-related to the average number of rooms per dwelling: `rm`. Another
-interesting predictor is `zn`, the proportion of residential land zoned for
-large lots. Nearly 73% of the observations have a value of 0 with the remaining
-spread widely between 1% and 100%. This may the source of possible distortion. A
-third highly-skewed variable is `chas`, the indicator as to whether the lot
-borders the Charles River. Out of the `r nobs` observations,
+The plots reveal that most of the data does not have a normal distribution. 
+As expected, we see the heavily right tailed variables and others, which are 
+multi-modal. The most Gaussian of the variables appears to be the one related 
+to the average number of rooms per dwelling: `rm`.
+
+Another interesting predictor, `zn` i.e. proportion of residential land zoned 
+for large lots, has a significant positive skew (skew = 2.18). Nearly 73% of 
+the observations (339 of 466 total) have a value of zero. When analyzing the 
+difference between the crime rate groups, it is possible that areas with high 
+crime rate do not have land zoned for large lots suggesting suburban areas are 
+likely to have low crime rates because there typically have large lots, whereas 
+urban areas have a higher crime rate since lot size are smaller than 25000 square 
+feet.
+
+Lastly, a third highly skewed variable is `chas`, the indicator as to whether 
+the lot borders the Charles River. Out of the `r nobs` observations,
 `r sum(DT$chas == 0)` do **not** border the river.
 
-To provide another view, this time highlighting outliers, the data was analyzed
-using boxplots.
+To provide another view, this time highlighting outliers, we present the data and 
+analyze using boxplots.
 
@@
 ## Data Correlation
-Below is a corrgram of the data. The shape of the ellipse reflects the strength
-of the correlation and the color represents the direction. Thus, the diagonal
-which represents the perfect correlation of each variable with itself is a blue
-diagonal line. The shape and the color scheme reflect the guide at the bottom
-of the image.
 
@@
 
-Perhaps this is a reflection of the noticeable influence that socio-economic
-status has on the crime rate of an area. It is known that crime is concentrated
-in disadvantaged, urban neighborhoods in the United States. The economic
-segregation suggests that affluent neighborhoods may be further away in terms of
-distance, and as a result, disadvantaged areas are more attractive to crime
-because the probability of success is higher even if the targets are not as
-profitable. Consider areas with a higher socio-economics, these areas are
-correlated have a lower crime rate. Moreover, industrial areas are correlated
-with higher levels of nitrogen oxide concentrations. This may be an indication
-of why these areas are less density with residents of a higher status. This
-trend is also seen with the `ptratio` since a higher ratio means less funding
-for public institutions, which is common in areas of lower status.
+It could be possible we are seeing a reflection of the noticeable influence 
+that socio-economic status has on the crime rate of an area.  Graif, Gladfelter 
+and Matthews (2016) state crime is generally concentrated in disadvantaged, 
+urban neighborhoods in the United States.  An economic segregation suggests 
+that affluent neighborhoods may be further away in terms of distance, and as 
+a result, disadvantaged areas are more attractive to crime because the probability 
+of success is higher even if the targets are not as profitable. We prove this data, 
+showing areas with a higher socio-economics correlating to a lower crime rate.
 
-Lastly, it seems that the variable `chas`, which indicate whether the suburb
-borders the Charles River, has statistically insignificant correlation with
-almost all of the other variables.
+Fields (2004) states industrial areas correlate with higher levels of nitrogen oxide 
+concentrations. It may be an indication of why these areas are less dense with residents 
+of a higher status. We see the same trend in `ptratio` since a higher ratio means less 
+funding for public institutions, which is common in areas of lower status.
+
+([https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4928692/](Source)).
+([https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1247398//](Source)).
+
+Lastly, it seems that the variable `chas`, which indicate whether the suburb borders 
+the Charles River, has statistically insignificant correlation with almost all of the 
+other variables.
 
 # DATA PREPARATION
-There are no missing values so none need to be imputed. Different preparation
-techniques may be used for different models.
+
+As there are no missing values, we have no need to impute. Instead, different preparation
+techniques are utilized for different models.
 
 ## Models 1 & 2
-The variable `chas` will be removed due to having statistically insignificant
-correlation with almost all the variables in the dataset. Next, outliers will be
-tempered by capping from below at the 5th percentile and from above at the 95th
-percentile.
+
+The variable `chas` will be removed due to having statistically insignificant correlation 
+with almost all the variables in the dataset.  Next, outliers are tempered by capping below 
+the 5th percentile and above the 95th percentile.
 
@@
 ## Model 3
-For model 3, very little data manipulation will be done. There may be valuable
-information in the outliers and a reasonable model may be able to extract
-something from `chas`. However, `chas` will be converted to a factor. Similarly,
-`rad` is an index of accessibility, so it is almost certainly a factor, although
-probably ordered. Given there are now two factors in the predictors, dummy
-dummy variables need to be created. Lastly, the **labels** of `target` will be
-changed from integer to character for compliance with some of R's naming
-conventions. Based on the definitions, 0 will be mapped to *below* and 1 to
-*above*.
+
+For model 3, we chose very little data manipulation, as there may be valuable 
+information in the outliers.  A reasonable model may be able to extract 
+something from `chas`.  Some variables, such as `chas` and `rad` (an index of 
+accessibility) are converted to a factor. Additionally, `rad` probably ordered. 
+Given there are now two factors in the predictors, we require creation of dummy 
+variables. Lastly, the **labels** of `target` will be changed from integer to 
+character for compliance with some of R's naming conventions.  Based on the 
+definitions, 0 will be mapped to *below* and 1 to *above*.
 
 # BUILD MODELS
+
 First, we will split the training data into two sets. This will allow us to
@@
 ### Base Model
+
 We will start with a simple logistic regression model to serve as a baseline.
@@
 ### Enhanced Model
+
 We will use **backwards stepwise regression** to remove features that are not
@@
 ### Coefficient Discussion
+
 It is important to note that the coefficients from the model are predicting 
-whether or not the target variable is a **1** (the crime rate is *above* the
-median value). Additionally, it must be noted that the numeric coefficients are
-relative to the range of values that the variable encompasses. What this means
-is that it's possible to have a coefficient that seems small when we look at the
-absolute magnitude, but that actually has a very strong effect when applied to
-the data.
+whether or not the target variable is a **1** (the crime rate *above* the 
+median value). Additionally, since the numeric coefficients are relative to 
+the range of values that the variable encompasses it is possible to have a 
+coefficient that seems small when we look at the absolute magnitude, but that 
+actually has a very strong effect when applied to the data.
 
@@
 
-Let's take a look at the *signs* of the coefficients. We can see that the
-coefficients for the variables `zn` and `tax` are negative. This is indicative
-of an inverse relationship; i.e., the higher these values, the less likely the
-crime rate is above the median. The relationship for the zone variable aligns
-with the findings from our initial data exploration. However, the relationship
-for the tax variable does not; in fact, we saw that there was a positive
-correlation between the two (as the tax rate increases, so does the probability
-that the crime rate will be above the median).
+Looking at the *signs* of the coefficients, we can see the variables `zn` 
+and `tax` are negative. This is indicative of an inverse relationship; 
+i.e., the higher these values, the less likely the crime rate is above the 
+median. The relationship for the zone variable aligns with the findings from 
+our initial data exploration. However, the relationship for the tax variable 
+does not; in fact, we saw that there was a positive correlation between the 
+two variables.  As the tax rate increases so does the probability that the 
+crime rate will be above the median.
 
-The rest of the variables have positive coefficients. Many of these are
+The rest of the variables have positive coefficients and many of these are 
 expected. For example, in our exploratory data analysis, we noticed that the 
 `nox` variable (nitrogen oxides concentration (parts per 10 million)) has the 
-greatest positive correlation with the target variable. This is surprising; we
-hadn't thought that the nitrogen oxide concentration would have as large of an
-impact on the crime rate, but some quick googling shows us that there may
-actually be a relationship
+greatest positive correlation with the target variable. This is surprising!  
+We had not thought that the nitrogen oxide concentration would have as large 
+of an impact on the crime rate. According to Colorado State University (2019), 
+there may actually be a strong relationship between the two.  The team of 
+researchers found strong correlation between exposure to air pollution and 
+aggressive behavior.
+
 ([https://www.sciencedaily.com/releases/2019/10/191003114007.htm](Source)).
@@
 We again see that the primary predictor in the logistic regression is the `nox` 
-with `rad` and `ptratio` rounding out the top three predictors. This is in line
+with `rad` and `ptratio` rounding out the top three predictors staying in line 
 with the base model. Living close to the center of town, having teachers, and 
-not having polluted air are the largest predictors in terms of above median crime.
+not having polluted air are the largest predictors in terms of staying above median 
+crime rates.
 
@@
 selected pair-wise interactions pruned using both forward and backward stepwise 
-regression based on AICc to select the optimal parameters. The captured indicies
-of the test set for models 1 & 2 will be used to split the data for model 3 as
-well. With the vanilla logistic regression model in R, `glm` with a binomial
-family and a logit link, there are no other hyperparameters available for
-tuning. More sophisticated logistic regression implementation have these
-parameters for fine-tuning. There is also no reasonable way to implement
-cross-validation so none will be done.
+regression based on AICc to select the optimal parameters. The captured indices 
+of the test set for models 1 & 2 are used to split the data for model 3 as well. 
+To address the possibility of over-fitting, we performed five-fold cross-validation 
+on the training set.  With the vanilla logistic regression model in R, `glm` with 
+a binomial family and a logit link, there are no other hyper parameters available 
+for tuning.  More sophisticated logistic regression implementation have these 
+parameters for fine-tuning.
 
-To consider *all* interactions would be foolhardy. Given the need for dummy
-variables, there will 19 predictors and so \(_{19}C_2 = 171\) possible pairs.
-This is too much. Therefore, intentional selection of pairs is required.
-Looking at the predictors, there are some which logically may interact with
-each other, such as:
+Considering *all* interactions would be foolhardy. Given the need for dummy variables, 
+there will 19 predictors and \(_{19}C_2 = 171\) possible pairs, too much. Therefore, 
+intentional selection of pairs is required.
 
-  * `indus` and `nox`
-    * Nitrous Oxide is a common industrial pollutant. There is a high
-    correlation in the training set between the two. Allowing for its
-    interaction may temper the result of their relationship without losing more
-    information
+Looking at the predictors, there are some, which logically may interact with each other, 
+such as:   
+* `indus`
+	* The concentration of non-retail businesses in the area
+* `nox`
+    		* Nitrous Oxide is a common industrial pollutant.
 
-What may prove interesting is to look at some unsupervised clustering techniques
-to provide us with insights as to which variables may be "close" enough to
-warrant interactions. Instead of treating the observations as samples, we will
-treat the predictors as samples!
+There is a high correlation in the training set between the two. Allowing for its 
+interaction may temper the result of their relationship without losing more information.
+
+What may prove interesting is to look at some unsupervised clustering techniques to 
+provide us with insights as to which variables may be "close" enough to warrant interactions. 
+Instead of treating the observations as samples, we will treat the predictors as samples!
 
@@
 It seems that there may be reason to consider an interaction between `chas` and
-`nox`. Can it be that many industrial plants are on the river? Similarly with
-`indus` and `lstat`. Higher-cost homes tend to be further away from the traffic
+`nox`. Can that many industrial plants be on the river?
+
+Similarly with `indus` and `lstat`. Higher-cost homes tend to be further away from the traffic
 and pollution of industrial areas.
@@
 
-The selected model has some interesting results. Firstly, the predictor with the
-greatest effect is `nox`. Further, the only model variable which serves to
-**decrease** the crime rate from the baseline is `rm`. Other variables, such as
-`chas`
+The selected model has some interesting results. First, the predictor with the 
+greatest effect is `nox`.  Second, the only model variable which serves to 
+**decrease** the crime rate from the baseline is `rm`.
 
-The interactions between `indus` and both `nox` and `lstat` are significant. As
-theorized, these interactions serve to temper their individual contributions to
-crime. Remember, a 1 indicates crime above the median, so positive coefficients
-indicate contributions to *increasing* the crime rate. A similar observation
-holds `rm`. While `dis` on its own did not decrease the model's AICc, its
-interaction to temper `rm` did help the model's deviance. No other selected
-interactions were left in the model.
+The interactions between `indus` and both `nox` and `lstat` are significant. 
+As theorized, these interactions serve to temper their individual contributions to 
+crime. Remember, a `1` indicates crime above the median, so positive coefficients 
+indicate contributions to *increasing* the crime rate. A similar observation holds 
+`rm`. While `dis` on its own did not decrease the model's AICc, its interaction 
+to temper `rm` did help the model's deviance. No other selected interactions are 
+left in the model.
 
@@
 shown below, removing it would ***drastically*** reduce the models performance.
+
 Lastly, while `rad.1` is the base case, `rad.2` and `rad.6` have disappeared as 
@@
 indistinguishable from `rad = 1`.
+
 ```
@@
 ## Performance 
-Below are the confusion matrices for the three models, followed by a table with
+Below we show the confusion matrices for the three models, followed by a table with
 the selected performance.

### 2020-10-28 HW3/HW3_Final_AA.Rmd
Update HW3_Final_AA.Rmd
Copy Edits to wording

@@
     Bryan Persaud, Jeff Shamp
-date: "10/27/2020"
+date: "10/28/2020"
 output:

### 2020-10-28 HW3/Older files/HW3_Final_AA.Rmd
Clean up directory as we are getting close

--- 2020-10-28 HW3/HW3_Final_AA.Rmd
@@
 We can make some initial observaations. The data set has complete cases, thus, 
-there is no need for imputation. Based on some common summary statistics, 
-there are more observations where the crime rate is below the median. It is already 
-apparent that some of the predictors varies depending the crime rate. For instance, 
-there is a noticeable difference in the means of `age`, `lstat`, `rad`, and `zn` between 
-the crime rate groups.
+there is no need for imputation. Based on some common summary statistics, there
+are more observations where the crime rate is below the median. It is already 
+apparent that some of the predictors varies depending the crime rate. For
+instance, there is a noticeable difference in the means of `age`, `lstat`,
+`rad`, and `zn` between the crime rate groups.
 
 ```{r sumstat}
-# DATA EXPLORATION 
+# Summary Statistics
 summarystat = stby(data = df, INDICES = df$target, FUN = psych::describe)
@@
 ## Data Distribution
-
 For each predictors, we computed and drew kernel density estimate to understand 
@@
 ```{r density, fig.width=8}
-# density plots
+# Density plots
 DT[, IDX := .I]
@@
 likely to have low crime rates because there typically have large lots, whereas 
-urban areas have a higher crime rate since lot size are smaller than 25000 square 
-feet.
+urban areas have a higher crime rate since lot size are smaller than 25000
+square feet.
 
@@
 
-To provide another view, this time highlighting outliers, we present the data and 
-analyze using boxplots.
+To provide another view, this time highlighting outliers, we present the data
+and analyze using boxplots.
 
 ```{r boxplots, fig.width=8}
-# boxplots
-
+# Boxplots 
 ggplot(DTM, aes(x = metric, y = value)) + 
@@
 ```{r correlation, fig.height=5, fig.width=5}
-# correlations
+# Corrgram
 corrplot(cor(df), method = 'ellipse', type = 'lower', order = 'hclust')
@@
 that socio-economic status has on the crime rate of an area. Graif, Gladfelter 
-and Matthews (2016) state crime is generally concentrated in disadvantaged, 
-urban neighborhoods in the United States.  An economic segregation suggests 
-that affluent neighborhoods may be further away in terms of distance, and as 
-a result, disadvantaged areas are more attractive to crime because the probability 
-of success is higher even if the targets are not as profitable. We prove this data, 
-showing areas with a higher socio-economics correlating to a lower crime rate.
+and Matthews (2014) state crime is generally concentrated in disadvantaged,
+urban neighborhoods in the United States. An economic segregation suggests that
+affluent neighborhoods may be further away in terms of distance, and as a
+result, disadvantaged areas are more attractive to crime because the probability
+of success is higher even if the targets are not as profitable. We prove this
+data, showing areas with a higher socio-economics correlating to a lower crime
+rate.
 
-Fields (2004) states industrial areas correlate with higher levels of nitrogen oxide 
-concentrations. It may be an indication of why these areas are less dense with residents 
-of a higher status. We see the same trend in `ptratio` since a higher ratio means less 
-funding for public institutions, which is common in areas of lower status.
+Fields (2004) states industrial areas correlate with higher levels of nitrous
+oxide concentrations. It may be an indication of why these areas are less dense
+with residents of a higher status. We see the same trend in `ptratio` since a
+higher ratio means less funding for public institutions, which is common in
+areas of lower status.
 
-([https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4928692/](Source)).
-([https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1247398//](Source)).
-
-Lastly, it seems that the variable `chas`, which indicate whether the suburb borders 
-the Charles River, has statistically insignificant correlation with almost all of the 
-other variables.
+Lastly, it seems that the variable `chas`, which indicate whether the suburb
+borders the Charles River, has statistically insignificant correlation with
+almost all of the other variables.
 
 # DATA PREPARATION
-
-As there are no missing values, we have no need to impute. Instead, different preparation
-techniques are utilized for different models.
+As there are no missing values, we have no need to impute. Further different
+preparation techniques are utilized for different models.
 
 ## Models 1 & 2
-
-The variable `chas` will be removed due to having statistically insignificant correlation 
-with almost all the variables in the dataset.  Next, outliers are tempered by capping below 
-the 5th percentile and above the 95th percentile.
+The variable `chas` will be removed due to having statistically insignificant
+correlation with almost all the variables in the dataset.  Next, outliers are
+tempered by capping below the 5th percentile and above the 95th percentile.
 
 ```{r model1_2_adjust}
+# Model 1&2 Prep: Remove chas
 new_df <- subset(df, select = -chas)
 
-# Cap all observations at their 5th and 95th percentiles
+# Model 1&2 Prep:  Cap all observations at their 5th and 95th percentiles
 low5 <- apply(new_df, 2, quantile, prob = 0.05)
@@
 }
+
+# Model 1&2 Prep: Store density plots prior to Box-Cox transform
 dens_ageO <- ggplot(new_df, aes(age)) + geom_density()
@@
 ```{r m12BC}
+# Model 1&2 Prep:  Calculate Box-Cox lambdas
 ageBC <- boxcoxfit(new_df$age)
@@
 ```{r m12applyBC}
+# Model 1&2 Prep: Apply the Box-Cox transform
 new_df$age <- new_df$age ^ ageBC$lambda
@@
 ```{r m12BC_plot}
+# Model 1&2 Prep: Graphically display the effect of the Box-Cox transform
 dens_age <- ggplot(new_df, aes(age)) + geom_density()
@@
 ## Model 3
-
 For model 3, we chose very little data manipulation, as there may be valuable 
-information in the outliers.  A reasonable model may be able to extract 
-something from `chas`.  Some variables, such as `chas` and `rad` (an index of 
-accessibility) are converted to a factor. Additionally, `rad` probably ordered. 
-Given there are now two factors in the predictors, we require creation of dummy 
+information in the outliers. A reasonable model may be able to extract something
+from `chas`. Some variables, such as `chas` and `rad`, an index of accessibility,
+are converted to a factor. Additionally, while `rad` may have an ordering, it
+will be converted to a normal factor, as ordered factors are processed via
+orthogonal polynoial regression, which is out of the scope of this class. Given
+that there are now two factors in the predictors, we require creation of dummy 
 variables. Lastly, the **labels** of `target` will be changed from integer to 
 character for compliance with some of R's naming conventions.  Based on the 
-definitions, 0 will be mapped to *below* and 1 to *above*.
+definitions, 0 will be mapped to *Below* and 1 to *Above*.
 
 # BUILD MODELS
-
 First, we will split the training data into two sets. This will allow us to
@@
 ```{r m12_tts}
-# BUILD MODELS
-# train-test split
-
+# Model 1: Split the data into training and testing sets
 crime_df <- new_df %>%
@@
 
-# This is so Model 3 can use the same train/test split as Models 1 and 2
+# Model 1: This is so Model 3 can use the same train/test split as Models 1 & 2
 crime_df$id <- seq_len(nrow(crime_df))
@@
 # Markdown file to ensure repeatability.
-
 set.seed(7104)
@@
 
-# Capture the indices and remove the no-longer-necessary column
+# Model 1: Capture the indices for Model 3 then remove the uneeded column
 tstIDX <- test_df$id
@@
 ### Base Model
-
 We will start with a simple logistic regression model to serve as a baseline.
@@
 ```{r baseModel1}
+# Base Model 1
 baseModel <- glm(target~., family = binomial, data = train_df)
@@
 ### Enhanced Model
-
 We will use **backwards stepwise regression** to remove features that are not
@@
 ### Coefficient Discussion
-
 It is important to note that the coefficients from the model are predicting 
@@
 ```{r model1_coeff}
-# 10-fold cross validation
+# Model 1: 10-fold cross validation
 train_control <- trainControl(method = "cv", number = 10)
 
-# train the model on training set
-model1 <- train(target ~ zn + nox + age + dis + rad + tax + ptratio + lstat + medv,
+# Model 1: Train the model on training set
+model1 <- train(target ~ zn + nox + age + dis + rad + tax + ptratio + lstat +
+                  medv,
                 data = train_df,
@@
 We had not thought that the nitrogen oxide concentration would have as large 
-of an impact on the crime rate. According to Colorado State University (2019), 
-there may actually be a strong relationship between the two.  The team of 
-researchers found strong correlation between exposure to air pollution and 
-aggressive behavior.
-
-([https://www.sciencedaily.com/releases/2019/10/191003114007.htm](Source)).
+of an impact on the crime rate. According to an article in Science Daily (2019),
+researchers at Colorado State University found that there may actually be a
+strong correlation between exposure to air pollution and aggressive behavior.
 
@@
 ### Logistic Classifier
+The first step is to setup a workflow which encapsulates the logistic regression
+and cross-validation.
 
 ```{r m2setup}
+# Model 2: Setup
 cv_folds <- vfold_cv(train_df, v = 10, repeats = 1)
@@
 
-crime_wf<- workflow() %>%
+crime_wf <- workflow() %>%
   add_recipe(crime_recipe)
 ```
-
-```{r m2svm}
+```{r m2logit}
+# Model 2: Call logistic regression
 logit_specs <-
@@
 
-Tuning the models for best parameters.
+Once the workflow is set up, the model is tuned to return the best parameters.
 
 ```{r m2stack}
+# Model 2: Setup tuning grid
 logit_wf <-
@@
 
-crtl_grid<-control_grid(verbose = FALSE)
+crtl_grid <-control_grid(verbose = FALSE)
 
@@
 
+The metric used to select the "best" model will be the area under the receiver
+operating curve (AUC).
+
 ```{r m2predict}
+# Model 2: Select AUC as optimizing metric
 best_model_auc <- select_best(logit_results, "roc_auc")
-final_auc<-
+final_auc <-
   finalize_workflow(
@@
 
-final_model_pred<-
+final_model_pred <-
   final_auc %>%
@@
   collect_predictions()
-
 ```
 
-Computing the metrics for the logit model:
+Once the algorithms select a best model, the test metrics can be computed.
 
 ```{r m2Stats}
+# Model 2: Compute metrics on test set (done here instead of below)
 m2_roc<- yardstick::roc_auc(
@@
 )
+
 m2_metrics_df<-
@@
 
-```{r}
+```{r m2VI}
+# Model 2: Plot variable importance
 logit_wf %>%
@@
 with the base model. Living close to the center of town, having teachers, and 
-not having polluted air are the largest predictors in terms of staying above median 
-crime rates.
+not having polluted air are the largest predictors in terms of staying above
+median crime rates.
 
 ## Model 3
-```{r m3trainTemp}
-DT_train <- DT[!(IDX %in% tstIDX)]
-```
-
 Model 3 will be a binary logistic regression allowing for all variables and 
@@
 of the test set for models 1 & 2 are used to split the data for model 3 as well. 
-To address the possibility of over-fitting, we performed five-fold cross-validation 
-on the training set.  With the vanilla logistic regression model in R, `glm` with 
-a binomial family and a logit link, there are no other hyper parameters available 
-for tuning.  More sophisticated logistic regression implementation have these 
-parameters for fine-tuning.
+With the vanilla logistic regression model in R, `glm` with a binomial family
+and a logit link, there are no hyper parameters available for tuning. More
+sophisticated logistic regression implementation have these parameters for
+fine-tuning. Lastly, the stepAIC method which will be used to select the best
+model from an information-criterion perspective, prevents the use of
+cross-validation.
 
-Considering *all* interactions would be foolhardy. Given the need for dummy variables, 
-there will 19 predictors and \(_{19}C_2 = 171\) possible pairs, too much. Therefore, 
-intentional selection of pairs is required.
+Considering *all* interactions would be foolhardy. Given the need for dummy
+variables, there will 19 predictors and \(_{19}C_2 = 171\) possible pairs---too
+much. Therefore, intentional selection of pairs is required.
 
-Looking at the predictors, there are some, which logically may interact with each other, 
-such as:   
+Looking at the predictors, there are some, which logically may interact with
+each other, such as:
+
 * `indus`
@@
 
-There is a high correlation in the training set between the two. Allowing for its 
-interaction may temper the result of their relationship without losing more information.
+There is a high correlation in the training set between the two. Allowing for
+its interaction may temper the result of their relationship without losing more
+information.
 
-What may prove interesting is to look at some unsupervised clustering techniques to 
-provide us with insights as to which variables may be "close" enough to warrant interactions. 
-Instead of treating the observations as samples, we will treat the predictors as samples!
+What may prove interesting is to look at some unsupervised clustering techniques
+to provide us with insights as to which variables may be "close" enough to
+warrant interactions. Instead of treating the observations as samples, we will
+treat the predictors as samples!
 
 ```{r m3Clust, fig.height=6}
+# Model 3: Create cluster plot of features
+DT_train <- DT[!(IDX %in% tstIDX)]
 predClust <- hclust(dist(t(DT_train[, -c('target', 'IDX')]), diag = TRUE))
@@
 It seems that there may be reason to consider an interaction between `chas` and
-`nox`. Can that many industrial plants be on the river?
-
-Similarly with `indus` and `lstat`. Higher-cost homes tend to be further away from the traffic
-and pollution of industrial areas.
+`nox`. Can that many industrial plants be on the river? Similarly with `indus`
+and `lstat`. Higher-cost homes tend to be further away from the traffic and
+pollution of industrial areas.
 
@@
 
-```{r m3dataPrep}
+```{r m3build}
+# Model 3: Data Preparation
 DT[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
@@
 DT_train <- DT[!(IDX %in% tstIDX)]
-```
-```{r m3build}
+
+# No CV, use precision/recall
 trC <- trainControl(method = 'none', classProbs = TRUE,
                     summaryFunction = twoClassSummary)
+
+# Model 3: Create dummy variables
 dVars <- dummyVars(target ~ indus * nox +
@@
 DTtrny <- DT_train$target
+
+# Model 3: Train Model
 set.seed(1)
@@
 ```{r m3Sum}
+# Model 3: Model summary
 kable(summary(m3Fit$finalModel)$coefficients, digits = 3L,
@@
 The interactions between `indus` and both `nox` and `lstat` are significant. 
-As theorized, these interactions serve to temper their individual contributions to 
-crime. Remember, a `1` indicates crime above the median, so positive coefficients 
-indicate contributions to *increasing* the crime rate. A similar observation holds 
-`rm`. While `dis` on its own did not decrease the model's AICc, its interaction 
-to temper `rm` did help the model's deviance. No other selected interactions are 
-left in the model.
+As theorized, these interactions serve to temper their individual contributions
+to crime. Remember, a `1` indicates crime above the median, so positive
+coefficients indicate contributions to *increasing* the crime rate. A similar
+observation holds `rm`. While `dis` on its own did not decrease the model's
+AICc, its interaction to temper `rm` did help the model's deviance. No other
+selected interactions are left in the model.
 
@@
 indistinguishable from `rad = 1`.
-
 ```
@@
 ## Performance 
-Below we show the confusion matrices for the three models, followed by a table with
-the selected performance.
+Below we show the confusion matrices for the three models, followed by a table
+with the selected performance.
 
 ```{r comparison}
+# Create table to hold model comparison statistics
 compTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
@@
 
-# Metrics
+# Model 1 Metrics
 m1_roc <- yardstick::roc_auc(
@@
 
+# Model 1 Confusion Matrix
 m1CM <- confusionMatrix(model1_preds, model1_results$target,
                         mode = 'prec_recall')
+m1CM$table
+
+# Model 1 Populate comparison table
 compTable[1, 2] <- m1_acc$.estimate
@@
 compTable[1, 4] <- m1_roc$.estimate
-m1CM$table
 ```
@@
 ```{r m2TestSet}
+# Model 2 calculate metrics and print confusion matrix
 final_model_pred %>%
@@
 
+# Model 2 Populate comparison table
 compTable[2, 2] <- m2.metrics[1, 3]
@@
 ```{r m3TestSet}
+# Model 3: Process test set
 tstdVars <- dummyVars(target ~ indus * nox +
@@
 DTtsty <- DT_test$target
+
+# Model 3: Predict on test set
 m3Pred <- predict(m3Fit, newdata = DTtstx)
 m3PredP <- predict(m3Fit, newdata = DTtstx, type = 'prob')
+
+# Model 3: Confusion Matrix
 m3CM <- confusionMatrix(m3Pred, DTtsty, mode = 'prec_recall')
 m3CM$table
+
+# Model 3: Populate comparison table
 compTable[3, 2] <- sum(diag(m3CM$table)) / sum(m3CM$table)
@@
 ```{r selperfstat}
+# Display Model Comparison
 kable(compTable, digits = 3L, caption = 'Model Test Results')
@@
 ```{r evalPred}
+# Selected Model: Process eval file data as per train and test sets
 setDT(eval)
@@
 evaltstx <- predict(evalVars, newdata = eval)
+
+# Selected Model: Predict classes and probabilities
 evalPred <- predict(m3Fit, newdata = evaltstx)
 evalPredP <- predict(m3Fit, newdata = evaltstx, type = 'prob')
+
+# Selected Model: Display results
 predTable <- data.frame(TestID = seq_len(neval),
@@
 
-# APPENDIX
+# REFERENCES
+
+ * Fields S. (2004). Global nitrogen: cycling out of control.
+ *Environmental health perspectives*, *112*(10), A556–A563.
+ https://doi.org/10.1289/ehp.112-a556
+ * Graif, C., Gladfelter, A. S., & Matthews, S. A. (2014). Urban Poverty and
+ Neighborhood Effects on Crime: Incorporating Spatial and Network Perspectives.
+ *Sociology compass*, *8*(9), 1140–1155. https://doi.org/10.1111/soc4.12199
+ * Colorado State University. (2019, October 3). Exposure to air pollution
+ increases violent crime rates. *ScienceDaily*. Retrieved October 26, 2020 from
+ www.sciencedaily.com/releases/2019/10/191003114007.htm
+
+# CODE APPENDIX
 The code chunks below represent the R code called in order during the analysis.
@@
 ```
-```{r m2svm}
+```{r m2logit}
 ```
@@
 ```
-```{r m3trainTemp}
+```{r m2VI}
 ```
@@
 ```
-```{r m3dataPrep}
+```{r m3build}
 ```

### 2020-10-28 HW3/HW3-Group2.Rmd
Update HW3-Group2.Rmd
Added Intro

--- 2020-10-28 HW3/Older files/HW3_Final_AA.Rmd
@@
 
+# Introduction
+Using data from a major city, Boston, we strove to identify trends and model 
+crime rates and their affect on the housing conditions.  Using 
+binary logistic regression models we hope to predict the crime levels in a 
+neighborhood using various conditions to set the status of the neighborhood. 
+Each section of the report breaks down the steps we took to achieve a solid 
+and effective model.
+
 # DATA EXPLORATION
 ## Data Description
+
 The training data set contains 466 records summarizing attributes of various

### 2020-10-30 HW3/HW3-Group2.Rmd
spelling

@@
 
-We can make some initial observaations. The data set has complete cases, thus, 
+We can make some initial observations. The data set has complete cases, thus, 
 there is no need for imputation. Based on some common summary statistics, there
@@
 will be converted to a normal factor, as ordered factors are processed via
-orthogonal polynoial regression, which is out of the scope of this class. Given
+orthogonal polynomial regression, which is out of the scope of this class. Given
 that there are now two factors in the predictors, we require creation of dummy 
@@
 
-# Model 1: Capture the indices for Model 3 then remove the uneeded column
+# Model 1: Capture the indices for Model 3 then remove the unneeded column
 tstIDX <- test_df$id
@@
 variables, there will 19 predictors and \(_{19}C_2 = 171\) possible pairs---too
-much. Therefore, intentional selection of pairs is required.
+many! Therefore, intentional selection of pairs is required.
 
@@
 level, except for `rad.24`. However, from the last trace of the stepping routine 
-shown below, removing it would ***drastically*** reduce the models performance.
+shown below, removing it would ***drastically*** reduce the model's performance.
 

### 2020-11-02 HW4/HW4_Draft.Rmd
Initial commit (data files and markdown shell)

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #4'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "11/01/2020"
output:
  pdf_document: default
urlcolor: purple
---

### 2020-11-02 HW4/HW4_Draft.Rmd
Intro

@@
 ---
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
+```
+
+```{r loadData, include=FALSE}
+library(caret)
+library(data.table)
+set.seed(71554)
+
+urlRemote  = "https://raw.githubusercontent.com/"
+pathGithub = "aadler/DT621_Fall2020_Group2/master/HW4/data/"
+fileTrain = "insurance_training_data.csv"
+fileTest = "insurance-evaluation-data.csv"
+
+DT <- fread(paste0(urlRemote, pathGithub, fileTrain))
+nobs <- dim(DT)[[1]]
+eval <- fread(paste0(urlRemote, pathGithub, fileTest))
+```
+
+# Introduction
+The assignment for HW4 is to analyze and model a dataset containing
+approximately 8000 records representing customers of an auto insurance company.
+Each record has two response variables. The first response variable,
+`TARGET_FLAG`, is a 1 or a 0. where a `1` means that the person was in a car
+and a `0` means that the person was not in a car crash. The second response
+variable is `TARGET_AMT`. This value is 0 if the person did not crash their car.
+However, if they did crash their car, this number will be a value greater than
+zero.
+
+The objective of the assignment is to build multiple linear regression and
+binary logistic regression models on the training data to predict **both** the
+probability that a person will crash their car and the amount of money it will
+cost given a crash. Only the variables in the dataset, or variables directly
+derived from them, may be used.
+
+#

### 2020-11-02 HW4/HW4_Draft.Rmd
Variable Table

@@
 
-#
+# DATA EXPLORATION
+## Variables
+The data is composed of the following variables:
+
+|VARIABLE NAME|DEFINITION|THEORETICAL EFFECT|
+|-|-|-|
+|INDEX|Identification Variable (do not use)|None|
+|TARGET_FLAG|Was Car in a crash? 1=YES 0=NO|None|
+|TARGET_AMT|If car was in a crash, what was the cost|None|
+|AGE|Age of Driver|Very young people tend to be risky. Maybe very old people also.|
+|BLUEBOOK|Value of Vehicle|Unknown effect on probability of collision, but probably effect the payout if there is a crash|
+|CAR_AGE|Vehicle Age|Unknown effect on probability of collision, but probably effect the payout if there is a crash|
+|CAR_TYPE|Type of Car|Unknown effect on probability of collision, but probably effect the payout if there is a crash|
+|CAR_USE|Vehicle Use|Commercial vehicles are driven more, so might increase probability of collision|
+|CLM_FREQ|# Claims (Past 5 Years)|The more claims you filed in the past, the more you are likely to file in the future|
+|EDUCATION|Max Education Level|Unknown effect, but in theory more educated people tend to drive more safely|
+|HOMEKIDS|# Children at Home|Unknown effect|
+|HOME_VAL|Home Value|In theory, home owners tend to drive more responsibly|
+|INCOME|Income|In theory, rich people tend to get into fewer crashes|
+|JOB|Job Category|In theory, white collar jobs tend to be safer|
+|KIDSDRIV|# Driving Children|When teenagers drive your car, you are more likely to get into crashes|
+|MSTATUS|Marital Status|In theory, married people drive more safely|
+|MVR_PTS|Motor Vehicle Record Points|If you get lots of traffic tickets, you tend to get into more crashes|
+|OLDCLAIM|Total Claims (Past 5 Years)|If your total payout over the past five years was high, this suggests future payouts will be high|
+|PARENT1|Single Parent|Unknown effect|
+|RED_CAR|A Red Car|Urban legend says that red cars (especially red sports cars) are more risky. Is that true?|
+|REVOKED|License Revoked (Past 7 Years)|If your license was revoked in the past 7 years, you probably are a more risky driver.|
+|SEX|Gender|Urban legend says that women have less crashes then men. Is that true?|
+|TIF|Time in Force|People who have been customers for a long time are usually more safe.|
+|TRAVTIME|Distance to Work|Long drives to work usually suggest greater risk|
+|URBANICITY|Home/Work Area|Unknown|
+|YOJ|Years on Job|People who stay at a job for a long time are usually more safe|

### 2020-11-02 HW4/HW4_Draft.Rmd
Read in train and eval, and get rid of those $ and , which make the currency characters.

@@
 ```{r loadData, include=FALSE}
+# Load necessary libraries
 library(caret)
 library(data.table)
+# Set master seed
 set.seed(71554)
 
+# Set filepaths for data ingestion
 urlRemote  = "https://raw.githubusercontent.com/"
@@
 
-DT <- fread(paste0(urlRemote, pathGithub, fileTrain))
-nobs <- dim(DT)[[1]]
-eval <- fread(paste0(urlRemote, pathGithub, fileTest))
+# Read training file
+DT <- fread(paste0(urlRemote, pathGithub, fileTrain),
+            colClasses = c(rep('integer', 2L), 'double', rep('integer', 4L),
+                           'double', 'factor', 'double', rep('factor', 4L),
+                           'integer', 'factor', 'double', 'integer',
+                           rep('factor', 2L), 'double', 'integer', 'factor',
+                           rep('integer', 2L), 'factor'))
+
+# Convert character currencies to doubles
+DT[, `:=`(INCOME = as.double(gsub('[$,]', '', INCOME)),
+          HOME_VAL = as.double(gsub('[$,]', '', HOME_VAL)),
+          BLUEBOOK = as.double(gsub('[$,]', '', BLUEBOOK)),
+          OLDCLAIM = as.double(gsub('[$,]', '', OLDCLAIM)))]
+
+# Read and process the test/evaluation file
+eval <- fread(paste0(urlRemote, pathGithub, fileTest),
+              colClasses = c(rep('integer', 2L), 'double', rep('integer', 4L),
+                             'double', 'factor', 'double', rep('factor', 4L),
+                             'integer', 'factor', 'double', 'integer',
+                             rep('factor', 2L), 'double', 'integer', 'factor',
+                             rep('integer', 2L), 'factor'))
+eval[, `:=`(INCOME = as.double(gsub('[$,]', '', INCOME)),
+          HOME_VAL = as.double(gsub('[$,]', '', HOME_VAL)),
+          BLUEBOOK = as.double(gsub('[$,]', '', BLUEBOOK)),
+          OLDCLAIM = as.double(gsub('[$,]', '', OLDCLAIM)))]
+
+# Number of training observations
+ntrobs <- dim(DT)[[1]]
 ```

### 2020-11-02 HW4/HW4_Draft.Rmd
Discuss missing values and summary & graphs for numeric values.

@@
 output:
-  pdf_document: default
+  pdf_document:
+    toc: TRUE
+    toc_depth: 3
 urlcolor: purple
@@
 # Load necessary libraries
+library(knitr)
 library(caret)
@@
 ntrobs <- dim(DT)[[1]]
+
+# Get the names of the variables and which ones are numeric
+nmtrn <- names(DT)
+nmtrnNUM <- names(DT[, .SD, .SDcols = sapply(DT, is.numeric)])
+nmtrnDUB <- names(DT[, .SD, .SDcols = sapply(DT, is.double)])
+nmtrnFAC <- setdiff(nmtrn, nmtrnNUM)
+# Drop INDEX and target variablesfrom name list
+nmtrnNUM <- nmtrnNUM[-(1:3)]
 ```
@@
 |YOJ|Years on Job|People who stay at a job for a long time are usually more safe|
+
+There are `r ntrobs` observations of `r length(nmtrnNUM) - 2` numeric predictor
+variables and `r length(nmtrnFAC)` factor predictor variables.
+
+## Missing Data
+There are missing observations. Specifically, the following variables have
+missing values coded as `NA`:
+
+```{r missingVal}
+# Check for NAs in all variables
+missingV <- DT[, lapply(.SD, function(x) sum(is.na(x)))]
+# Only keep those whose count is > 0
+missingV <- missingV[, .SD, .SDcols = sapply(missingV, function (x) x > 0)]
+kable(missingV, caption = "Variables with Count of Missing Values > 0")
+```
+
+Another concern is that the following variables should not have significant
+probabilities of 0 because they represent values which *should* be positive:
+
+  * AGE
+  * YOJ
+  * INCOME
+  * HOME_VAL
+  * TRAVTIME
+  * BLUEBOOK
+  * TIF
+  * CAR_AGE
+
+```{r zeroVal}
+# Check for 0 value for the above variables
+zeroV <- DT[, lapply(.SD, function(x) sum(x == 0, na.rm = TRUE)),
+            .SDcols = c('AGE', 'YOJ', 'INCOME', 'HOME_VAL', 'TRAVTIME',
+                        'BLUEBOOK', 'TIF', 'CAR_AGE')]
+# Only keep those whose count of zero values is > 0
+zeroV <- zeroV[, .SD, .SDcols = sapply(zeroV, function (x) x > 0)]
+kable(zeroV, caption = "Variables with Count of Missing Values > 0")
+```
+
+The few missing `CAR_AGE` values should be easy to impute. The other heavy-0
+observations are of more concern. Over 25% of the `HOME_VAL` observations are 0!
+Now it may be that those with `HOME_VAL` of 0 do not own homes but rent, and it
+may be that those with `YOJ` and `INCOME` of 0 were unemployed at the time of
+the data collection. There are `r DT[YOJ == 0 & INCOME == 0, .N]` observations
+with both `YOJ` and `INCOME` of 0, meaning there are a few dozen of each which
+have values in one class and not the other. Handling this will be addressed in
+the DATA PREPARATION section.
+
+## Summary Statistics and Graphs
+### Numeric Predictors
+The numeric predictor variables have the following summary statistics, ignoring
+missing values:
+
+```{r statsN}
+# Isolate numeric only predictors
+numDT <- DT[, .SD, .SDcols = nmtrnNUM]
+
+# Melt them from wide to long format
+numDTM <- melt(numDT, variable.name = 'metric', value.name = 'value')
+
+# Calculate summary statistics
+statsN <- numDTM[, .(Mean = mean(value, na.rm = TRUE),
+                     SD = sd(value, na.rm = TRUE),
+                     Min = min(value, na.rm = TRUE),
+                     Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
+                     Median = median(value, na.rm = TRUE),
+                     Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
+                     Max = max(value, na.rm = TRUE),
+                     IQR = IQR(value, na.rm = TRUE)), keyby = 'metric']
+
+# Print the table
+kable(statsN, digits = 3L, align = 'r',
+      caption = "Summary Statistitics for Numeric Variables")
+```
+
+A visual depiction of the distributions of the numeric predictors follows. The
+blue graphs are on a normal scale and the red ones are on a \(\log_{10}\) scale.
+
+```{r graphsN}
+# Freedman-Diaconis rule for bin widths
+FDbin <- function(x) {
+  result <- 2 * IQR(x, na.rm = TRUE) / (length(x) ^ (1 / 3))
+  return(ifelse(result == 0, 0.5, result))
+}
+ggplot(numDTM[!is.na(metric)], aes(x = value)) +
+  geom_histogram(binwidth = FDbin, fill = 'lightblue3') +
+  facet_wrap( ~ metric, scales = 'free')
+
+ggplot(numDTM, aes(x = value)) +
+  geom_histogram(binwidth = FDbin, fill = 'indianred4') +
+  facet_wrap( ~ metric, scales = 'free') + scale_x_log10()
+```
+
+What the above visualizations show is that there are two sets of numeric
+variables. There are those which are clearly discrete, taking one of a few
+integral values, and there are those which are more continuous. Of the
+continuous variables, some like `AGE` look rather Gaussian at first glance,
+while others, such as `HOME_VAL` or `INCOME` are skewed. However, `HOME_VAL`
+looks lognormal, as its histogram on the log scale is near-Gaussian.
+
+A variable such as `OLDCLAIM` is expected to have a mass point at 0. Every
+observation which had fewer than 2 claims clearly did not have a prior claim!
+
+### Factor Predictors
+
+# CODE APPENDIX
+```{r setup2, include=FALSE}
+knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
+```
+```{r loadData}
+```
+```{r missingVal}
+```
+```{r zeroVal}
+```
+```{r statsN}
+```
+```{r graphsN}
+```

### 2020-11-02 HW4/HW4_Draft.Rmd
Correlation, cleanings, misings, splitting, and starting model 1 frequency

@@
 # Load necessary libraries
+library(corrplot)
+library(ggplot2)
+library(scales)
 library(knitr)
@@
 ### Factor Predictors
+While tabular depiction of factor variables is usually of little value, visual
+representation of their distributions is of use.
+
+```{r graphsF, fig.width=8, fig.height=8}
+# Isolate factor only predictors
+facDT <- DT[, .SD, .SDcols = nmtrnFAC]
+
+# Melt them from wide to long format (need to explicitly call measure.vars since
+# these are all factors)
+facDTM <- melt(facDT, measure.vars = nmtrnFAC, variable.name = 'metric',
+               value.name = 'value')
+ggplot(facDTM, aes(x = value)) + geom_bar(fill = 'darkolivegreen') +
+  facet_wrap( ~ metric, nrow = 5L, scales = 'free') + coord_flip()
+```
+
+In order to use regression-based tools, we are going to have to convert these
+factors to dummy variables. Another important observation is that there are
+rows for which `JOB` is blank and some strange value names for some of the
+factors. Cleaning this up will be handled in the DATA PREPARATION section.
+
+### Correlations
+The corrgram below graphically represents the correlations between the numeric
+predictor variables, when ignoring the missing variables.
+
+```{r corrgram, fig.width=6, fig.height=4}
+corrplot(cor(DT[, ..nmtrnNUM], use = 'complete.obs'),
+         method = 'ellipse', type = 'lower')
+```
+
+Most of the numeric variables are uncorrelated with one another. A few
+exceptions exist:
+
+ * `HOMEKIDS` with `KIDSDRIV`
+   * This is self-explanatory
+ * `HOME_VAL` with `INCOME`
+   * Both are very strongly tied to wealth and net worth
+ * `BLUEBOOK` and `INCOME`
+   * People with more disposable income can afford more expensive vehicles
+ * `CLM_FREQ` with `OLDCLAIM`
+   * More claims implies a greater total aggregate payment
+ * `CAR_AGE` and `INCOME`
+   * This is interesting. It could be that people with higher disposable incomes
+   buy more expensive cars and keep them for longer, but that's a guess.
+ * `MVR_PTS` is *lightly* correlated with `CLM_FREQ` and `OLDCLAIM`
+   * Probably when one is in a lot of accidents, the chances of them being at
+   some level of fault rises
+ * `HOMEKIDS` **negatively** correlated with `AGE`
+   * As people age, their children age as well and eventually move out, we hope!
+
+# DATA PREPARATION
+## Data Value Cleanup
+The first step will be to clean up the strange values and blank `JOB` entries.
+
+```{r dataClean}
+# Clean up data values. Using setattr for fast data.table setting. Equivalent
+# to levels(DT$X) <- c(a, b) but faster since changes by reference.
+setattr(DT$MSTATUS, 'levels', c('Yes', 'No'))
+setattr(DT$SEX, 'levels', c('M', 'F'))
+setattr(DT$EDUCATION, 'levels', c('<High School', 'Bachelors', 'Masters', 'PhD',
+                                  'High School'))
+setattr(DT$JOB, 'levels', c('Unknown', 'Clerical', 'Doctor', 'Home Maker',
+                            'Lawyer', 'Manager', 'Professional', 'Student',
+                            'Blue Collar'))
+setattr(DT$CAR_TYPE, 'levels', c('Minivan', 'Panel Truck', 'Pickup',
+                                 'Sports Car', 'Van', 'SUV'))
+
+setattr(DT$URBANICITY, 'levels', c('Urban', 'Rural'))
+```
+
+## Training & Testing Split
+While the following discussion more properly occurs under BUILD MODELS, as it
+factors into how the data is prepared, it will happen here. All the models will
+be trained on the same approximately 70% of the training set, reserving 30% for
+validation of which model to select for the frequency and severity estimation on
+the supplied evaluation set.
+
+However, there will need to be two sets of training and testing, as the severity
+model must perforce be trained on data for which a claim occurred. To maintain
+integrity, the split of all the data for frequency will be honored for severity.
+The 70/30 split will be honored as all claims with a `TARGET_FLAG` of 1 will
+have non-zero `TARGET_AMT`, However, there will be no guarantee that the
+*distribution* of the loss conditional on event in the validation set will be
+similar to the training set as a whole.
+
+```{r trainTestSplit}
+set.seed(642)
+trnIDX <- createDataPartition(DT$TARGET_FLAG, p = 0.7)
+trnX <- DT[trnIDX$Resample1, ]
+tstX <- DT[!trnIDX$Resample1, ]
+```
+
+## Models Set 1
+### Missing Data
+The issues with factor data were handled universally for all models through
+resetting the levels. What remains is handling the missing numeric values. For
+`CAR_AGE` simple k-means imputation will be used. The remaining three problem
+variables, `YOJ`, `INCOME`, and `HOME_VAL` have both `NA` and 0 issues.
+
+For `YOJ` and `INCOME`, it is not unreasonable to consider 0 to be realistic:
+unemployed or employed less than half a year. As such, we will leave the 0's and
+impute the NAs.
+
+We will take a different approach with `HOME_VAL`. As seen above when the 0
+values are removed, `HOME_VAL` has a nice lognormal shape. It is also reasonable
+to consider that people responding with 0 for `HOME_VAL` are non-owners as
+opposed to it being missing, although we would want to investigate this further
+were we allowed. Therefore, we are going to add a "Own" variable which will be
+`Yes` for all positive `HOME_VAL` and `No` otherwise, and allow an interaction
+between this value and the actual value. This would be much more effective in a
+decision tree framework, but that is not available.
+
+```{r model1addVars}
+# Add OWN variable for model set 1
+tstX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
+                     levels = c(0, 1), labels = c('No', 'Yes'))]
+```
+
+### Dummy Variables
+As there are many factor variables, dummy variables will be created. The target
+variables will be removed and held to the side so as not to contaminate the
+fitting
+
+```{r model1createDummy}
+trnFreq <- tstX$TARGET_FLAG
+trnSev <- tstX$TARGET_AMT
+tstDumFreq <- TARGET_FLAG ~ . - c(INDEX, TARGET_AMT)
+```
+
+# BUILD MODELS
+## Model Set 1
+### Frequency Model
+The frequency model will be a binary logistic on the prepared data. It will
+start with the complete model and allow for some interactions. It will use both
+forward and backward steps to identify which variables to include, using AIC as
+the test metric. Using this approach does not allow for cross-validation. The
+initial interactions considered will be:
+
+ * `OWN` with `HOME_VAL` as discussed above
+ * `HOME_VAL` with `INCOME` as using both may overstate the effect of wealth
+ * `KIDSDRIV` with `HOMEKIDS` as there is
 
@@
 ```
+```{r graphsF}
+```
+```{r dataClean}
+```
+```{r trainTestSplit}
+```
+```{r model1addVars}
+```

### 2020-11-02 HW4/HW4_Draft.Rmd
Model1 for freq

@@
 library(caret)
+library(RANN)
 library(data.table)
@@
 
+# BUILD MODELS
+## Model Set 1
 ### Dummy Variables
@@
 variables will be removed and held to the side so as not to contaminate the
-fitting
+fitting.
 
-```{r model1createDummy}
-trnFreq <- tstX$TARGET_FLAG
-trnSev <- tstX$TARGET_AMT
-tstDumFreq <- TARGET_FLAG ~ . - c(INDEX, TARGET_AMT)
-```
-
-# BUILD MODELS
-## Model Set 1
 ### Frequency Model
@@
 
- * `OWN` with `HOME_VAL` as discussed above
- * `HOME_VAL` with `INCOME` as using both may overstate the effect of wealth
- * `KIDSDRIV` with `HOMEKIDS` as there is
+ * `OWN` with `HOME_VAL` as discussed in the missing data section
+ * `KIDSDRIV` with `HOMEKIDS`
+ * A three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`
+ * `CLM_FREQ` with `OLDCLAIM`
+
+```{r model1createDummy}
+trnFreq <- factor(tstX$TARGET_FLAG, levels = c(0, 1),
+                  labels = c("NoClaim", "Claim"))
+trnSev <- tstX$TARGET_AMT
+tstDumFreq <- dummyVars(TARGET_FLAG ~ .  + OWN * HOME_VAL +
+                          KIDSDRIV * HOMEKIDS + HOME_VAL * INCOME * BLUEBOOK +
+                          CLM_FREQ * OLDCLAIM - INDEX - TARGET_AMT,
+                        fullRank = TRUE, data = tstX)
+tstXD <- predict(tstDumFreq, newdata = tstX)
+```
+
+At this point the code will be set up to preprocess the data and train the
+model. The two pre-processing steps which will be done will be to check for
+near-zero variance between the predictors and then use kNN imputations for
+missing values. The actual code may be found in the CODE APPENDIX.
+
+```{r runModel1Freq}
+trC <- trainControl(method = 'none',            # Using StepAIC
+                    classProbs = TRUE,          # Classification
+                    summaryFunction = prSummary # Precision/Recall)
+)
+set.seed(487)
+binlogFit <- train(x = tstXD, y = trnFreq, preProcess = c('nzv', 'knnImpute'),
+                   trControl = trC, method = 'glmStepAIC', family = 'binomial',
+                   direction = 'both', trace = 0)
+kable(summary(binlogFit)$coefficients, digits = 3L,
+      caption = "Model 1 Frequency Regression Output")
+```
+
+The last trace of the stepping procedure is shown below.
+```
+                             Df Deviance    AIC
+<none>                            2077.4 2135.4
++ YOJ                         1   2075.8 2135.8
+- SEX.F                       1   2080.0 2136.0
+- JOB.Professional            1   2080.2 2136.2
++ `EDUCATION.High School`     1   2076.2 2136.2
++ JOB.Clerical                1   2076.4 2136.4
++ `JOB.Blue Collar`           1   2076.8 2136.8
++ JOB.Lawyer                  1   2076.9 2136.9
++ `INCOME:HOME_VAL`           1   2076.9 2136.9
++ `HOME_VAL:BLUEBOOK`         1   2077.0 2137.0
++ OLDCLAIM                    1   2077.0 2137.0
++ `INCOME:BLUEBOOK`           1   2077.0 2137.0
+- `CAR_TYPE.Panel Truck`      1   2081.2 2137.2
++ RED_CAR.yes                 1   2077.2 2137.2
++ HOMEKIDS                    1   2077.2 2137.2
++ HOME_VAL                    1   2077.3 2137.3
++ `HOME_VAL:OWNYes`           1   2077.3 2137.3
++ AGE                         1   2077.3 2137.3
++ CAR_AGE                     1   2077.4 2137.4
+- JOB.Student                 1   2081.8 2137.8
+- PARENT1.Yes                 1   2082.4 2138.4
+- `INCOME:HOME_VAL:BLUEBOOK`  1   2083.4 2139.4
+- `JOB.Home Maker`            1   2083.4 2139.4
+- EDUCATION.Bachelors         1   2084.5 2140.5
+- `OLDCLAIM:CLM_FREQ`         1   2085.1 2141.1
+- CAR_TYPE.SUV                1   2085.8 2141.8
+- BLUEBOOK                    1   2086.4 2142.4
+- OWN.Yes                     1   2086.5 2142.5
+- EDUCATION.Masters           1   2087.2 2143.2
+- EDUCATION.PhD               1   2089.8 2145.8
+- CAR_TYPE.Pickup             1   2090.1 2146.1
+- CAR_TYPE.Van                1   2090.2 2146.2
+- TRAVTIME                    1   2090.9 2146.9
+- INCOME                      1   2092.0 2148.0
+- `CAR_TYPE.Sports Car`       1   2092.1 2148.1
+- MSTATUS.No                  1   2093.4 2149.4
+- CLM_FREQ                    1   2095.5 2151.5
+- TIF                         1   2095.8 2151.8
+- KIDSDRIV                    1   2101.1 2157.1
+- MVR_PTS                     1   2103.2 2159.2
+- CAR_USE.Private             1   2106.4 2162.4
+- JOB.Manager                 1   2108.4 2164.4
+- REVOKED.Yes                 1   2118.5 2174.5
+- URBANICITY.Rural            1   2271.8 2327.8
+```
 
@@
 ```
+```{r corrgram}
+```
 ```{r dataClean}
@@
 ```
+```{r model1createDummy}
+```
+```{r runModel1Freq}
+```

### 2020-11-02 HW4/HW4_Draft.Rmd
Correct train/test mixup. Work on Sev model

--- 2020-11-02 HW4/HW4_Draft.Rmd
@@
     toc: TRUE
-    toc_depth: 3
+    toc_depth: 4
 urlcolor: purple
@@
 setattr(DT$URBANICITY, 'levels', c('Urban', 'Rural'))
+
+# Relevel some of the variables to have the default make more sense
+# Set default marital status to unmarried
+set(DT, NULL, 'MSTATUS', relevel(DT$MSTATUS, 'No'))
+# Set default car usage to private
+set(DT, NULL, 'CAR_USE', relevel(DT$CAR_USE, 'Private'))
 ```
@@
 # Add OWN variable for model set 1
+trnX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
+                     levels = c(0, 1), labels = c('No', 'Yes'))]
 tstX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
@@
 
- * `OWN` with `HOME_VAL` as discussed above
- * `HOME_VAL` with `INCOME` as using both may overstate the effect of wealth
- * `KIDSDRIV` with `HOMEKIDS` as there is
+ * `OWN` with `HOME_VAL` as discussed in the missing data section
+ * `KIDSDRIV` with `HOMEKIDS`
+ * A three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`
+ * `CLM_FREQ` with `OLDCLAIM`
+ * `MSTATUS` with `SEX`
+   * One sex may be more affected by marriage than the other
+
+```{r model1DummyFreq}
+trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
+                  labels = c("NoClaim", "Claim"))
+trnDumFreq <- dummyVars(TARGET_FLAG ~ .  + OWN * HOME_VAL +
+                          KIDSDRIV * HOMEKIDS + HOME_VAL * INCOME * BLUEBOOK +
+                          CLM_FREQ * OLDCLAIM + MSTATUS * SEX - INDEX -
+                          TARGET_AMT,
+                        fullRank = TRUE, data = trnX)
+trnXFD <- predict(trnDumFreq, newdata = trnX)
+```
+
+#### Training
+At this point the code will be set up to preprocess the data and train the
+model. The two pre-processing steps which will be done will be to check for
+near-zero variance between the predictors and then use kNN imputations for
+missing values. The actual code may be found in the CODE APPENDIX.
+
+```{r model1RunFreq}
+trC <- trainControl(method = 'none',            # Using StepAIC
+                    classProbs = TRUE,          # Classification
+                    summaryFunction = prSummary # Precision/Recall)
+)
+set.seed(487)
+m1FreqFit <- train(x = trnXFD, y = trnFreq, preProcess = c('nzv', 'knnImpute'),
+                   trControl = trC, method = 'glmStepAIC',
+                   family = binomial(link = 'logit'), direction = 'both',
+                   trace = 1)
+kable(summary(m1FreqFit)$coefficients, digits = 3L,
+      caption = "Model 1 Frequency Regression Output")
+```
+
+#### Coefficient Discussion
+Positive coefficients represent an **increased** risk of a claim over the
+baseline and negative coefficients represent a **decreased** risk over the
+baseline. Most of the coefficients are of the expected direction. For example,
+having one's license revoked or living in an urban area is a *positive* risk
+factor for claims while higher levels of wealth or education are *negative* risk
+factors.
+
+Some interesting observations. `SEX` is not a significant predictor, even at the
+10% level, but removing it does not increase the AIC, so it remains. The
+hypothesized interaction between `MSTATUS` and `SEX` does lower the deviance,
+but not enough to justify the additional parameter.
+
+The predictors with the greatest effect on risk are `URBANICITY` and `INCOME`.
+Having the three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`,
+together with `INCOME` and `BLUEBOOK` alone, obviated the need for stand-alone
+`HOME_VAL` and the other two possible interactions.
+
+In terms of variable importance, it can be estimated from the the last trace of
+the stepping procedure shown below. The further down the table the variable is,
+the more dramatic its inclusion or exclusion will be. The top 10 variables in
+order would be: `URBANICITY.Rural`, `REVOKED.Yes`, `CAR_USE.Commercial`,
+`TRAVTIME`, `CAR_TYPE.Sports Car`, `CLM_FREQ`, `MVR_PTS`, `TIF`, `CAR_TYPE.SUV`,
+and `JOB.Blue Collar`.
+
+```
+                             Df Deviance    AIC
+<none>                            5164.6 5224.6
++ HOMEKIDS                    1   5162.7 5224.7
++ AGE                         1   5163.1 5225.1
+- SEX.F                       1   5167.4 5225.4
++ EDUCATION.Masters           1   5163.4 5225.4
++ JOB.Manager                 1   5163.7 5225.7
++ OLDCLAIM                    1   5164.0 5226.0
++ EDUCATION.PhD               1   5164.0 5226.0
++ `HOME_VAL:BLUEBOOK`         1   5164.2 5226.2
++ CAR_AGE                     1   5164.2 5226.2
++ `MSTATUSYes:SEXF`           1   5164.5 5226.5
++ OWN.Yes                     1   5164.5 5226.5
++ YOJ                         1   5164.6 5226.6
++ RED_CAR.yes                 1   5164.6 5226.6
++ `EDUCATION.High School`     1   5164.6 5226.6
++ `INCOME:HOME_VAL:BLUEBOOK`  1   5164.6 5226.6
+- `INCOME:BLUEBOOK`           1   5168.8 5226.8
+- `INCOME:HOME_VAL`           1   5170.0 5228.0
+- JOB.Lawyer                  1   5170.6 5228.6
+- JOB.Student                 1   5171.4 5229.4
+- `CAR_TYPE.Panel Truck`      1   5173.5 5231.5
+- MSTATUS.Yes                 1   5174.8 5232.8
+- CAR_TYPE.Van                1   5179.5 5237.5
+- BLUEBOOK                    1   5179.9 5237.9
+- `OLDCLAIM:CLM_FREQ`         1   5180.0 5238.0
+- `JOB.Home Maker`            1   5180.2 5238.2
+- HOME_VAL                    1   5180.7 5238.7
+- EDUCATION.Bachelors         1   5181.4 5239.4
+- INCOME                      1   5182.3 5240.3
+- PARENT1.Yes                 1   5182.6 5240.6
+- CAR_TYPE.Pickup             1   5183.2 5241.2
+- JOB.Professional            1   5186.1 5244.1
+- JOB.Clerical                1   5195.4 5253.4
+- KIDSDRIV                    1   5197.8 5255.8
+- `JOB.Blue Collar`           1   5201.2 5259.2
+- CAR_TYPE.SUV                1   5202.5 5260.5
+- TIF                         1   5204.4 5262.4
+- MVR_PTS                     1   5204.8 5262.8
+- CLM_FREQ                    1   5205.4 5263.4
+- `CAR_TYPE.Sports Car`       1   5208.4 5266.4
+- TRAVTIME                    1   5211.7 5269.7
+- CAR_USE.Commercial          1   5230.2 5288.2
+- REVOKED.Yes                 1   5233.0 5291.0
+- URBANICITY.Rural            1   5612.1 5670.1
+```
+
+### Severity Model
+In the insurance world, severity is not necessarily fit via linear models, but
+is more often fit through comparison of the maximum likelihood fits of various
+distributional families to the size of the claims, but the task here is to fit a
+linear regression model to estimate future claim size.
+
+#### Training
+The observations in the training set which have claim amounts will be extracted
+and used to fit the linear regression. Similar to frequency, the model will be
+selected via a forward and backward stepwise AIC algorithm starting with every
+relevant predictive variable and some selected interactions.
+
+```{r model1DummySev}
+trnSev <- trnX$TARGET_AMT[trnX$TARGET_AMT > 0]
+trnDumSev <- dummyVars(TARGET_AMT ~ .  + OWN * HOME_VAL +
+                          KIDSDRIV * HOMEKIDS + HOME_VAL * INCOME * BLUEBOOK +
+                          CLM_FREQ * OLDCLAIM + MSTATUS * SEX - INDEX -
+                          TARGET_FLAG,
+                        fullRank = TRUE, data = trnX[TARGET_AMT > 0])
+trnXSD <- predict(trnDumFreq, newdata = trnX[TARGET_AMT > 0])
+```
+
+```{r runModel1Sev}
+trC <- trainControl(method = 'none')
+set.seed(878)
+m1SevFit <- train(x = trnXSD, y = trnSev, preProcess = c('nzv', 'knnImpute'),
+                   trControl = trC, method = 'glmStepAIC',
+                   family = gaussian, direction = 'both',
+                   trace = 0)
+kable(summary(m1SevFit)$coefficients, digits = 3L,
+      caption = "Model 1 Severity Regression Output")
+```
+
+#### Coefficient Disucssion
+Hi!
+
+```
+                             Df   Deviance   AIC
+<none>                          9.5466e+10 31329
+- REVOKED.Yes                 1 9.5614e+10 31329
++ JOB.Manager                 1 9.5368e+10 31330
+- MSTATUS.Yes                 1 9.5625e+10 31330
++ CAR_AGE                     1 9.5379e+10 31330
+- `OLDCLAIM:CLM_FREQ`         1 9.5645e+10 31330
++ `MSTATUSYes:SEXF`           1 9.5392e+10 31330
+- CAR_USE.Commercial          1 9.5650e+10 31330
++ SEX.F                       1 9.5399e+10 31330
++ JOB.Professional            1 9.5400e+10 31330
++ CAR_TYPE.Pickup             1 9.5410e+10 31330
++ CAR_TYPE.Van                1 9.5412e+10 31330
++ MVR_PTS                     1 9.5413e+10 31330
++ HOMEKIDS                    1 9.5415e+10 31330
++ CLM_FREQ                    1 9.5431e+10 31331
+- `INCOME:HOME_VAL`           1 9.5688e+10 31331
++ RED_CAR.yes                 1 9.5440e+10 31331
+- HOME_VAL                    1 9.5694e+10 31331
++ JOB.Student                 1 9.5444e+10 31331
++ EDUCATION.Masters           1 9.5445e+10 31331
++ PARENT1.Yes                 1 9.5452e+10 31331
+- OLDCLAIM                    1 9.5709e+10 31331
++ JOB.Clerical                1 9.5455e+10 31331
++ `CAR_TYPE.Panel Truck`      1 9.5455e+10 31331
++ `EDUCATION.High School`     1 9.5456e+10 31331
++ OWN.Yes                     1 9.5457e+10 31331
++ TRAVTIME                    1 9.5458e+10 31331
++ `INCOME:HOME_VAL:BLUEBOOK`  1 9.5460e+10 31331
++ `JOB.Home Maker`            1 9.5460e+10 31331
++ URBANICITY.Rural            1 9.5463e+10 31331
++ `INCOME:BLUEBOOK`           1 9.5464e+10 31331
++ AGE                         1 9.5464e+10 31331
++ `HOME_VAL:BLUEBOOK`         1 9.5465e+10 31331
++ `CAR_TYPE.Sports Car`       1 9.5465e+10 31331
++ YOJ                         1 9.5465e+10 31331
++ `JOB.Blue Collar`           1 9.5466e+10 31331
++ JOB.Lawyer                  1 9.5466e+10 31331
++ KIDSDRIV                    1 9.5466e+10 31331
++ EDUCATION.Bachelors         1 9.5466e+10 31331
++ CAR_TYPE.SUV                1 9.5466e+10 31331
++ `KIDSDRIV:HOMEKIDS`         1 9.5466e+10 31331
++ TIF                         1 9.5466e+10 31331
++ INCOME                      1 9.5466e+10 31331
++ EDUCATION.PhD               1 9.5466e+10 31331
+- BLUEBOOK                    1 9.6124e+10 31337
+```
+# SELECT MODELS
+## Model Selection Criteria
+For that *classification* algorithm, we will look at accuracy, AUC, and F1, and
+select the model which performs best two out of three in those metrics. For the
+*regression* algorithm we will look at RMSE, AIC. and MAE (since the values must
+be positive) and select the model which performs best two out of three in those
+metrics.
+
+### Frequency Models
+
+```{r model1Test}
+
+```
+
+# PREDICTIONS
 
@@
 ```
+```{r model1DummyFreq}
+```
+```{r model1RunFreq}
+```
+```{r model1DummySev}
+```
+```{r model1RunSev}
+```

### 2020-11-02 HW4/HW4_Draft.Rmd
Add a cache so that those with slower computers don't have to recalc the frequency model every time.

@@
 
-```{r model1RunFreq}
+```{r model1RunFreq, cache = 2}
 trC <- trainControl(method = 'none',            # Using StepAIC

### 2020-11-02 HW4/HW4_Draft.Rmd
Turn off tracing

--- 2020-11-02 HW4/HW4_Draft.Rmd
@@
                    family = binomial(link = 'logit'), direction = 'both',
-                   trace = 1)
+                   trace = 0)
 kable(summary(m1FreqFit)$coefficients, digits = 3L,

### 2020-11-02 HW4/HW4_Draft.Rmd
Update

--- 2020-11-02 HW4/HW4_Draft.Rmd
@@
 set(DT, NULL, 'CAR_USE', relevel(DT$CAR_USE, 'Private'))
+# Set default are to rural, making urban a positive? risk factor
+set(DT, NULL, 'URBANICITY', relevel(DT$URBANICITY, 'Rural'))
 ```
@@
    * One sex may be more affected by marriage than the other
+ * `CAR_USE` with `URBANICITY`
 
@@
                   labels = c("NoClaim", "Claim"))
-trnDumFreq <- dummyVars(TARGET_FLAG ~ .  + OWN * HOME_VAL +
-                          KIDSDRIV * HOMEKIDS + HOME_VAL * INCOME * BLUEBOOK +
-                          CLM_FREQ * OLDCLAIM + MSTATUS * SEX - INDEX -
+trnDumFreq <- dummyVars(TARGET_FLAG ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
+                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
+                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                           TARGET_AMT,
@@
                    family = binomial(link = 'logit'), direction = 'both',
-                   trace = 1)
-kable(summary(m1FreqFit)$coefficients, digits = 3L,
+                   trace = 0)
+kable(summary(m1FreqFit$finalModel)$coefficients, digits = 3L,
       caption = "Model 1 Frequency Regression Output")
@@
 #### Coefficient Discussion
-Positive coefficients represent an **increased** risk of a claim over the
-baseline and negative coefficients represent a **decreased** risk over the
-baseline. Most of the coefficients are of the expected direction. For example,
-having one's license revoked or living in an urban area is a *positive* risk
-factor for claims while higher levels of wealth or education are *negative* risk
-factors.
+It is difficult to directly map the predictors to probabilities, as all the
+values have been centered and scaled in order to perform the imputation.
+However, we can say that positive coefficients represent an **increased** risk
+of a claim over the baseline and negative coefficients represent a **decreased**
+risk under the baseline. When validating and then testing, the same imputations
+will be done and the results will be predicted on the normal probability scale.
 
-Some interesting observations. `SEX` is not a significant predictor, even at the
-10% level, but removing it does not increase the AIC, so it remains. The
-hypothesized interaction between `MSTATUS` and `SEX` does lower the deviance,
-but not enough to justify the additional parameter.
+Most of the coefficients are of the expected direction. For example, having
+one's license revoked, living in an urban area, or increased travel times are
+all *positive* risk factors while higher levels of wealth or education are
+*negative* risk factors.
 
-The predictors with the greatest effect on risk are `URBANICITY` and `INCOME`.
-Having the three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`,
-together with `INCOME` and `BLUEBOOK` alone, obviated the need for stand-alone
-`HOME_VAL` and the other two possible interactions.
+Some interesting observations. The predictors with the greatest effect on risk
+are `URBANICITY` and `INCOME`. The value of the coefficient for `SEX` implies
+that being female decreases the probability of a claim, as suspected, but it is
+only significant as a predictor at the 10% level. However, it is kept in the
+model as removing it increases the AIC. The hypothesized interaction between
+`MSTATUS` and `SEX` does lower the deviance, but not enough to justify the
+additional parameter. `BLUEBOOK` value is a negative risk factor. Perhaps people
+with more expensive cars drive more carefully?
+
+The full three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK` was
+deemed insignificant in light of the existing two-way interactions which are in
+the model.
+
+The interaction between `CLM_FREQ` and `OLDCLAIM` did help the model AIC, and
+the coefficient indicates that while `OLDCLAIM` did not justify the additional
+parameter, its tempering of `CLM_FREQ` **does** justify the additional
+parameter.
+
+The interaction between `CAR_USE` and `URBANICITY` is also of interest. It
+indicates that commercial use vehicles don't really add to claim frequency in
+rural areas---the singleton variable is absent from the model---but they have
+an outsized influence on increased claim probabilities in urban areas!
 
@@
 the more dramatic its inclusion or exclusion will be. The top 10 variables in
-order would be: `URBANICITY.Rural`, `REVOKED.Yes`, `CAR_USE.Commercial`,
-`TRAVTIME`, `CAR_TYPE.Sports Car`, `CLM_FREQ`, `MVR_PTS`, `TIF`, `CAR_TYPE.SUV`,
-and `JOB.Blue Collar`.
+order would be: `URBANICITY.Urban`, `CAR_USECommercial:URBANICITYUrban`,
+`REVOKED.Yes`, `TRAVTIME`, `CAR_TYPE.Sports Car`, `CLM_FREQ`, `TIF`, `MVR_PTS`,
+`CAR_TYPE.SUV`, and `JOB.Blue Collar`.
 
@@
                                       Df Deviance    AIC
-<none>                            5164.6 5224.6
-+ HOMEKIDS                    1   5162.7 5224.7
-+ AGE                         1   5163.1 5225.1
-- SEX.F                       1   5167.4 5225.4
-+ EDUCATION.Masters           1   5163.4 5225.4
-+ JOB.Manager                 1   5163.7 5225.7
-+ OLDCLAIM                    1   5164.0 5226.0
-+ EDUCATION.PhD               1   5164.0 5226.0
-+ `HOME_VAL:BLUEBOOK`         1   5164.2 5226.2
-+ CAR_AGE                     1   5164.2 5226.2
-+ `MSTATUSYes:SEXF`           1   5164.5 5226.5
-+ OWN.Yes                     1   5164.5 5226.5
-+ YOJ                         1   5164.6 5226.6
-+ RED_CAR.yes                 1   5164.6 5226.6
-+ `EDUCATION.High School`     1   5164.6 5226.6
-+ `INCOME:HOME_VAL:BLUEBOOK`  1   5164.6 5226.6
-- `INCOME:BLUEBOOK`           1   5168.8 5226.8
-- `INCOME:HOME_VAL`           1   5170.0 5228.0
-- JOB.Lawyer                  1   5170.6 5228.6
-- JOB.Student                 1   5171.4 5229.4
-- `CAR_TYPE.Panel Truck`      1   5173.5 5231.5
-- MSTATUS.Yes                 1   5174.8 5232.8
-- CAR_TYPE.Van                1   5179.5 5237.5
-- BLUEBOOK                    1   5179.9 5237.9
-- `OLDCLAIM:CLM_FREQ`         1   5180.0 5238.0
-- `JOB.Home Maker`            1   5180.2 5238.2
-- HOME_VAL                    1   5180.7 5238.7
-- EDUCATION.Bachelors         1   5181.4 5239.4
-- INCOME                      1   5182.3 5240.3
-- PARENT1.Yes                 1   5182.6 5240.6
-- CAR_TYPE.Pickup             1   5183.2 5241.2
-- JOB.Professional            1   5186.1 5244.1
-- JOB.Clerical                1   5195.4 5253.4
-- KIDSDRIV                    1   5197.8 5255.8
-- `JOB.Blue Collar`           1   5201.2 5259.2
-- CAR_TYPE.SUV                1   5202.5 5260.5
-- TIF                         1   5204.4 5262.4
-- MVR_PTS                     1   5204.8 5262.8
-- CLM_FREQ                    1   5205.4 5263.4
-- `CAR_TYPE.Sports Car`       1   5208.4 5266.4
-- TRAVTIME                    1   5211.7 5269.7
-- CAR_USE.Commercial          1   5230.2 5288.2
-- REVOKED.Yes                 1   5233.0 5291.0
-- URBANICITY.Rural            1   5612.1 5670.1
+<none>                                     5152.3 5212.3
++ HOMEKIDS                             1   5150.5 5212.5
++ AGE                                  1   5150.8 5212.8
++ EDUCATION.Masters                    1   5151.0 5213.0
+- SEX.F                                1   5155.2 5213.2
++ JOB.Manager                          1   5151.6 5213.6
++ OLDCLAIM                             1   5151.6 5213.6
++ EDUCATION.PhD                        1   5151.7 5213.7
++ `HOME_VAL:BLUEBOOK`                  1   5151.7 5213.7
++ CAR_AGE                              1   5151.8 5213.8
+- `INCOME:BLUEBOOK`                    1   5155.9 5213.9
++ OWN.Yes                              1   5152.1 5214.1
++ `MSTATUSYes:SEXF`                    1   5152.2 5214.2
++ RED_CAR.yes                          1   5152.3 5214.3
++ `EDUCATION.High School`              1   5152.3 5214.3
++ YOJ                                  1   5152.3 5214.3
++ `INCOME:HOME_VAL:BLUEBOOK`           1   5152.3 5214.3
++ CAR_USE.Commercial                   1   5152.3 5214.3
+- `INCOME:HOME_VAL`                    1   5158.7 5216.7
+- JOB.Student                          1   5159.0 5217.0
+- JOB.Lawyer                           1   5159.1 5217.1
+- `CAR_TYPE.Panel Truck`               1   5160.0 5218.0
+- MSTATUS.Yes                          1   5162.6 5220.6
+- CAR_TYPE.Van                         1   5166.3 5224.3
+- BLUEBOOK                             1   5166.6 5224.6
+- `JOB.Home Maker`                     1   5166.7 5224.7
+- `OLDCLAIM:CLM_FREQ`                  1   5167.4 5225.4
+- EDUCATION.Bachelors                  1   5169.0 5227.0
+- HOME_VAL                             1   5169.5 5227.5
+- PARENT1.Yes                          1   5169.9 5227.9
+- INCOME                               1   5170.3 5228.3
+- CAR_TYPE.Pickup                      1   5170.6 5228.6
+- JOB.Professional                     1   5174.0 5232.0
+- JOB.Clerical                         1   5181.7 5239.7
+- KIDSDRIV                             1   5186.1 5244.1
+- `JOB.Blue Collar`                    1   5189.2 5247.2
+- CAR_TYPE.SUV                         1   5189.8 5247.8
+- MVR_PTS                              1   5191.2 5249.2
+- TIF                                  1   5191.8 5249.8
+- CLM_FREQ                             1   5193.0 5251.0
+- `CAR_TYPE.Sports Car`                1   5195.8 5253.8
+- TRAVTIME                             1   5198.9 5256.9
+- REVOKED.Yes                          1   5220.1 5278.1
+- `CAR_USECommercial:URBANICITYUrban`  1   5229.9 5287.9
+- URBANICITY.Urban                     1   5422.8 5480.8
 ```
@@
 #### Training
-The observations in the training set which have claim amounts will be extracted
-and used to fit the linear regression. Similar to frequency, the model will be
-selected via a forward and backward stepwise AIC algorithm starting with every
-relevant predictive variable and some selected interactions.
+The observations in the training set which have claim amounts greater than 0
+will be extracted and used to fit the linear regression. Similar to frequency,
+the model will be selected via a forward and backward stepwise AIC algorithm
+starting with the same predictors and interactions as in the frequency model.
 
 ```{r model1DummySev}
-trnSev <- trnX$TARGET_AMT[trnX$TARGET_AMT > 0]
-trnDumSev <- dummyVars(TARGET_AMT ~ .  + OWN * HOME_VAL +
-                          KIDSDRIV * HOMEKIDS + HOME_VAL * INCOME * BLUEBOOK +
-                          CLM_FREQ * OLDCLAIM + MSTATUS * SEX - INDEX -
+trnSev <- trnX[TARGET_AMT > 0, TARGET_AMT]
+trnDumSev <- dummyVars(TARGET_AMT ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
+                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
+                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
                           TARGET_FLAG,
                         fullRank = TRUE, data = trnX[TARGET_AMT > 0])
-trnXSD <- predict(trnDumFreq, newdata = trnX[TARGET_AMT > 0])
+trnXSD <- predict(trnDumSev, newdata = trnX[TARGET_AMT > 0])
 ```
@@
 #### Coefficient Disucssion
-Hi!
+It is not surprising that this model has many fewer predictors. Information
+which helps in identifying the *existence* of a claim often does not help in
+estimating the *magnitude* of the claim.
+
+Similar to the frequency model, it is difficult to directly map the predictors
+to claim values, as all the values have been centered and scaled in order to
+perform the imputation. However, we can say that positive coefficients represent
+**increased** claim sizes over the baseline and negative coefficients represent
+**decreased** claim sizes under the baseline. When validating and then testing,
+the same imputations will be done and the results will be predicted on the
+normal dollar scale.
+
+The predictive variables with the greatest effect on claim size are `OLD_CLAIM`,
+`HOME_VAL`, and `BLUEBOOK`. The first and third make the most sense. One who has
+had larger claims in the past is likely to have larger claims in the future.
+Also, the claim for damages is very strongly correlated with the value of the
+vehicle. `HOME_VAL` is a bit strange. Perhaps it's serving as a proxy for
+purchased coverage limit. Maybe people with more expensive homes, who buy larger
+limits on their homeowners policies, also purchase larger limits on their
+vehicular polices.
+
+The interaction between `INCOME` and `HOME_VAL` serves to temper the effect of
+`HOME_VAL`. Similarly for the inter
 
@@
                                       Df   Deviance   AIC
-<none>                          9.5466e+10 31329
-- REVOKED.Yes                 1 9.5614e+10 31329
-+ JOB.Manager                 1 9.5368e+10 31330
-- MSTATUS.Yes                 1 9.5625e+10 31330
-+ CAR_AGE                     1 9.5379e+10 31330
-- `OLDCLAIM:CLM_FREQ`         1 9.5645e+10 31330
-+ `MSTATUSYes:SEXF`           1 9.5392e+10 31330
-- CAR_USE.Commercial          1 9.5650e+10 31330
-+ SEX.F                       1 9.5399e+10 31330
-+ JOB.Professional            1 9.5400e+10 31330
-+ CAR_TYPE.Pickup             1 9.5410e+10 31330
-+ CAR_TYPE.Van                1 9.5412e+10 31330
-+ MVR_PTS                     1 9.5413e+10 31330
-+ HOMEKIDS                    1 9.5415e+10 31330
-+ CLM_FREQ                    1 9.5431e+10 31331
-- `INCOME:HOME_VAL`           1 9.5688e+10 31331
-+ RED_CAR.yes                 1 9.5440e+10 31331
-- HOME_VAL                    1 9.5694e+10 31331
-+ JOB.Student                 1 9.5444e+10 31331
-+ EDUCATION.Masters           1 9.5445e+10 31331
-+ PARENT1.Yes                 1 9.5452e+10 31331
-- OLDCLAIM                    1 9.5709e+10 31331
-+ JOB.Clerical                1 9.5455e+10 31331
-+ `CAR_TYPE.Panel Truck`      1 9.5455e+10 31331
-+ `EDUCATION.High School`     1 9.5456e+10 31331
-+ OWN.Yes                     1 9.5457e+10 31331
-+ TRAVTIME                    1 9.5458e+10 31331
-+ `INCOME:HOME_VAL:BLUEBOOK`  1 9.5460e+10 31331
-+ `JOB.Home Maker`            1 9.5460e+10 31331
-+ URBANICITY.Rural            1 9.5463e+10 31331
-+ `INCOME:BLUEBOOK`           1 9.5464e+10 31331
-+ AGE                         1 9.5464e+10 31331
-+ `HOME_VAL:BLUEBOOK`         1 9.5465e+10 31331
-+ `CAR_TYPE.Sports Car`       1 9.5465e+10 31331
-+ YOJ                         1 9.5465e+10 31331
-+ `JOB.Blue Collar`           1 9.5466e+10 31331
-+ JOB.Lawyer                  1 9.5466e+10 31331
-+ KIDSDRIV                    1 9.5466e+10 31331
-+ EDUCATION.Bachelors         1 9.5466e+10 31331
-+ CAR_TYPE.SUV                1 9.5466e+10 31331
-+ `KIDSDRIV:HOMEKIDS`         1 9.5466e+10 31331
-+ TIF                         1 9.5466e+10 31331
-+ INCOME                      1 9.5466e+10 31331
-+ EDUCATION.PhD               1 9.5466e+10 31331
-- BLUEBOOK                    1 9.6124e+10 31337
+<none>                                   9.5456e+10 31329
+- REVOKED.Yes                          1 9.5604e+10 31329
++ JOB.Manager                          1 9.5357e+10 31329
+- MSTATUS.Yes                          1 9.5620e+10 31330
++ CAR_AGE                              1 9.5370e+10 31330
++ `MSTATUSYes:SEXF`                    1 9.5382e+10 31330
+- `OLDCLAIM:CLM_FREQ`                  1 9.5636e+10 31330
+- CAR_USE.Commercial                   1 9.5639e+10 31330
++ SEX.F                                1 9.5389e+10 31330
++ JOB.Professional                     1 9.5391e+10 31330
++ CAR_TYPE.Pickup                      1 9.5400e+10 31330
++ MVR_PTS                              1 9.5402e+10 31330
++ CAR_TYPE.Van                         1 9.5403e+10 31330
++ HOMEKIDS                             1 9.5404e+10 31330
++ CLM_FREQ                             1 9.5420e+10 31330
++ `CAR_USECommercial:URBANICITYUrban`  1 9.5426e+10 31330
++ RED_CAR.yes                          1 9.5430e+10 31331
++ JOB.Student                          1 9.5432e+10 31331
+- `INCOME:HOME_VAL`                    1 9.5686e+10 31331
++ EDUCATION.Masters                    1 9.5435e+10 31331
+- HOME_VAL                             1 9.5695e+10 31331
++ PARENT1.Yes                          1 9.5442e+10 31331
++ JOB.Clerical                         1 9.5445e+10 31331
++ `CAR_TYPE.Panel Truck`               1 9.5446e+10 31331
+- OLDCLAIM                             1 9.5700e+10 31331
++ `EDUCATION.High School`              1 9.5446e+10 31331
++ OWN.Yes                              1 9.5447e+10 31331
++ TRAVTIME                             1 9.5448e+10 31331
++ `INCOME:HOME_VAL:BLUEBOOK`           1 9.5450e+10 31331
++ `JOB.Home Maker`                     1 9.5450e+10 31331
++ URBANICITY.Urban                     1 9.5453e+10 31331
++ `INCOME:BLUEBOOK`                    1 9.5454e+10 31331
++ AGE                                  1 9.5454e+10 31331
++ `HOME_VAL:BLUEBOOK`                  1 9.5455e+10 31331
++ `JOB.Blue Collar`                    1 9.5455e+10 31331
++ `CAR_TYPE.Sports Car`                1 9.5455e+10 31331
++ KIDSDRIV                             1 9.5456e+10 31331
++ JOB.Lawyer                           1 9.5456e+10 31331
++ EDUCATION.Bachelors                  1 9.5456e+10 31331
++ YOJ                                  1 9.5456e+10 31331
++ CAR_TYPE.SUV                         1 9.5456e+10 31331
++ INCOME                               1 9.5456e+10 31331
++ `KIDSDRIV:HOMEKIDS`                  1 9.5456e+10 31331
++ TIF                                  1 9.5456e+10 31331
++ EDUCATION.PhD                        1 9.5456e+10 31331
+- BLUEBOOK                             1 9.6115e+10 31337
 ```
+
 # SELECT MODELS

### 2020-11-02 HW4/HW4_Draft.Rmd
Complete coefficient discussion.

@@
 
+#### Variable Importance
 In terms of variable importance, it can be estimated from the the last trace of
 the stepping procedure shown below. The further down the table the variable is,
-the more dramatic its inclusion or exclusion will be. The top 10 variables in
-order would be: `URBANICITY.Urban`, `CAR_USECommercial:URBANICITYUrban`,
-`REVOKED.Yes`, `TRAVTIME`, `CAR_TYPE.Sports Car`, `CLM_FREQ`, `TIF`, `MVR_PTS`,
-`CAR_TYPE.SUV`, and `JOB.Blue Collar`.
+the more dramatic its inclusion or exclusion will be on the deviance and theAIC.
+The top 10 variables in order would be: `URBANICITY.Urban`,
+`CAR_USECommercial:URBANICITYUrban`, `REVOKED.Yes`, `TRAVTIME`,
+`CAR_TYPE.Sports Car`, `CLM_FREQ`, `TIF`, `MVR_PTS`, `CAR_TYPE.SUV`,
+and `JOB.Blue Collar`.
 
@@
 The interaction between `INCOME` and `HOME_VAL` serves to temper the effect of
-`HOME_VAL`. Similarly for the inter
+`HOME_VAL`. Similarly for the interaction between `OLDCLAIM` and `CLM_FREQ`.
+
+#### Variable Importance
+In terms of variable importance, it can be estimated from the the last trace of
+the stepping procedure shown below. There are only eight variables in the model,
+and in order of AIC contribution they are `BLUEBOOK`, `OLDCLAIM`, `HOME_VAL`,
+`INCOME:HOME_VAL`, `CAR_USE.Commercial`, `OLDCLAIM:CLM_FREQ`, `MSTATUS.Yes`, and
+` REVOKED.Yes`.
 

### 2020-11-02 HW4/HW4_Draft.Rmd
Finish first draft

@@
 library(RANN)
+library(pROC)
 library(data.table)
@@
 ```{r model1DummyFreq}
+# Cast target as factor for caret purposes
 trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
                   labels = c("NoClaim", "Claim"))
+# Set up dummy variables
 trnDumFreq <- dummyVars(TARGET_FLAG ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
@@
                         fullRank = TRUE, data = trnX)
+# Create matrix of dummy values
 trnXFD <- predict(trnDumFreq, newdata = trnX)
@@
 near-zero variance between the predictors and then use kNN imputations for
-missing values. The actual code may be found in the CODE APPENDIX.
+missing values. Even though the `caret` package allows the preprocessing to
+occur in the call to `train`, it will be seperated here for clarity. The actual
+code may be found in the CODE APPENDIX.
 
 ```{r model1RunFreq, cache = 2}
+# preProcess with near-zero value and kNN imputation
+trnXFDpp <- preProcess(trnXFD, method = c('nzv', 'knnImpute'), k = 5)
+
+# Create processed predictors
+trnXFDI <- predict(trnXFDpp, newdata = trnXFD)
 trC <- trainControl(method = 'none',            # Using StepAIC
                     classProbs = TRUE,          # Classification
-                    summaryFunction = prSummary # Precision/Recall)
+                    summaryFunction = prSummary # Precision/Recall
 )
 set.seed(487)
-m1FreqFit <- train(x = trnXFD, y = trnFreq, preProcess = c('nzv', 'knnImpute'),
-                   trControl = trC, method = 'glmStepAIC',
-                   family = binomial(link = 'logit'), direction = 'both',
-                   trace = 0)
+# Train Model
+m1FreqFit <- train(x = trnXFDI, y = trnFreq, trControl = trC,
+                   method = 'glmStepAIC', family = binomial(link = 'logit'),
+                   direction = 'both', trace = 0)
+# Print results
 kable(summary(m1FreqFit$finalModel)$coefficients, digits = 3L,
@@
 ```{r model1DummySev}
+# Extract those observations with values > 0
 trnSev <- trnX[TARGET_AMT > 0, TARGET_AMT]
+# Set up dummy variables
 trnDumSev <- dummyVars(TARGET_AMT ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
@@
                         fullRank = TRUE, data = trnX[TARGET_AMT > 0])
+# Create matrix of dummy values
 trnXSD <- predict(trnDumSev, newdata = trnX[TARGET_AMT > 0])
+# preProcess with near-zero value and kNN imputation
+trnXSDpp <- preProcess(trnXSD, method = c('nzv', 'knnImpute'), k = 5)
+
+# Create processed predictors
+trnXSDI <- predict(trnXSDpp, newdata = trnXSD)
 ```
@@
 set.seed(878)
-m1SevFit <- train(x = trnXSD, y = trnSev, preProcess = c('nzv', 'knnImpute'),
-                   trControl = trC, method = 'glmStepAIC',
-                   family = gaussian, direction = 'both',
+m1SevFit <- train(x = trnXSDI, y = trnSev, trControl = trC,
+                  method = 'glmStepAIC', family = gaussian, direction = 'both',
                   trace = 0)
@@
 select the model which performs best two out of three in those metrics. For the
-*regression* algorithm we will look at RMSE, AIC. and MAE (since the values must
-be positive) and select the model which performs best two out of three in those
-metrics.
+*regression* algorithm we will look at RMSE and MAE (since the values must
+be positive) and select the model which performs best on both, Ties will be
+broken by adjusted \(R^2\).
 
-### Frequency Models
+```{r setUpCompTables}
+freqCompTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
+                            ACC = double(3),
+                            F1 = double(3),
+                            AUC = double(3))
+sevCompTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
+                           RMSE = double(3),
+                           MAE = double(3),
+                           adjR2 = double(3))
+```
 
-```{r model1Test}
+## Model 1
+### Frequency Model
 
+```{r validateM1F}
+# Model 1 Frequency:: process validation data similar to training data
+tstFreq <- factor(tstX$TARGET_FLAG, levels = c(0, 1),
+                  labels = c("NoClaim", "Claim"))
+tstDumFreq <- dummyVars(TARGET_FLAG ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
+                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
+                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
+                          TARGET_AMT,
+                        fullRank = TRUE, data = tstX)
+tstXFD <- predict(tstDumFreq, newdata = tstX)
+tstXFDpp <- preProcess(tstXFD, method = c('nzv', 'knnImpute'))
+tstXFDI <- predict(tstXFDpp, newdata = tstXFD)
+
+# Model 1 Frequency: predict on validation data
+m1Fpred <- predict(m1FreqFit, newdata = tstXFDI)
+m1FpredProb <- predict(m1FreqFit, newdata = tstXFDI, type = 'prob')
+
+# Model 1 Frequency: Calculate accuracy statistics
+m1FCM <- confusionMatrix(m1Fpred, tstFreq, mode = 'everything')
+freqCompTable[1, 2] <- m1FACC <- sum(diag(m1FCM$table)) / sum(m1FCM$table)
+freqCompTable[1, 3] <- m1FF1 <- m1FCM$byClass[7]
+freqCompTable[1, 4] <- m1FAUC <- auc(response = tstFreq,
+                                     predictor = m1FpredProb[, 1])
+```
+
+Frequency Model 1  returns an accuracy of `r round(m1FACC, 4)`, an F1 score of
+`r round(m1FF1, 4)`, and an AUC of `r round(m1FAUC, 4)`.
+
+### Severity Model
+
+```{r validateM1S}
+# Model 1 Severity: process validation data similar to training data
+tstSev <- tstX[TARGET_AMT > 0, TARGET_AMT]
+nobstst <- length(tstSev)
+np <- summary(m1SevFit$finalModel)$df[3]
+tstDumSev <- dummyVars(TARGET_AMT ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
+                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
+                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
+                          TARGET_FLAG,
+                        fullRank = TRUE, data = tstX[TARGET_AMT > 0])
+tstXSD <- predict(tstDumSev, newdata = tstX[TARGET_AMT > 0])
+tstXSDpp <- preProcess(tstXSD, method = c('nzv', 'knnImpute'), k = 5)
+tstXSDI <- predict(tstXSDpp, newdata = tstXSD)
+
+# Model 1 Severity: predict on validation data
+m1Spred <- predict(m1SevFit, newdata = tstXSDI)
+
+# Model 1 Severity: Calculate accuracy statistics
+m1Ssum <- postResample(m1Spred, tstSev)
+sevCompTable[1, 2] <- m1SRMSE <- m1Ssum[1]
+sevCompTable[1, 3] <- m1SMAE <- m1Ssum[3]
+m1SAR2 <- 1 - (1 - m1Ssum[2]) * (nobstst - 1) / (nobstst - np - 1)
+sevCompTable[1, 4] <- m1SAR2
+```
+
+Severity Model 1  returns an RMSE of `r round(m1SRMSE, 1)`, an MAE of
+`r round(m1SMAE, 1)`, and an adjusted \(R^2\) of `r round(m1SAR2, 4)`.
+
+## Model 2
+### Frequency Model
+### Severity Model
+
+## Model 3
+### Frequency Model
+### Severity Model
+
+## Model Selection
+### Frequency Model
+```{r freqComp}
+kable(freqCompTable, digits = 4L, caption = "Frequency Model Comparison")
+```
+
+### Severity Model
+```{r sevComp}
+kable(sevCompTable, digits = 4L, caption = "Severity Model Comparison")
 ```
@@
 ```
+```{r setUpCompTables}
+```
+```{r validateM1F}
+```
+```{r validateM1S}
+```
+```{r freqComp}
+```
+```{r sevComp}
+```

### 2020-11-03 HW4/HW4_Draft.Rmd
Update using KNN with median 3 insted of mean 5

@@
 # to levels(DT$X) <- c(a, b) but faster since changes by reference.
+# Remove spaces where possible too
 setattr(DT$MSTATUS, 'levels', c('Yes', 'No'))
 setattr(DT$SEX, 'levels', c('M', 'F'))
-setattr(DT$EDUCATION, 'levels', c('<High School', 'Bachelors', 'Masters', 'PhD',
-                                  'High School'))
-setattr(DT$JOB, 'levels', c('Unknown', 'Clerical', 'Doctor', 'Home Maker',
+setattr(DT$EDUCATION, 'levels', c('<HighSchool', 'Bachelors', 'Masters', 'PhD',
+                                  'HighSchool'))
+setattr(DT$JOB, 'levels', c('Unknown', 'Clerical', 'Doctor', 'HomeMaker',
                             'Lawyer', 'Manager', 'Professional', 'Student',
-                            'Blue Collar'))
-setattr(DT$CAR_TYPE, 'levels', c('Minivan', 'Panel Truck', 'Pickup',
-                                 'Sports Car', 'Van', 'SUV'))
+                            'BlueCollar'))
+setattr(DT$CAR_TYPE, 'levels', c('Minivan', 'PanelTruck', 'Pickup',
+                                 'SportsCar', 'Van', 'SUV'))
 
@@
 
-```{r model1RunFreq, cache = 2}
+```{r model1RunFreq, cache=TRUE}
 # preProcess with near-zero value and kNN imputation
-trnXFDpp <- preProcess(trnXFD, method = c('nzv', 'knnImpute'), k = 5)
+trnXFDpp <- preProcess(trnXFD, method = c('nzv', 'knnImpute'),
+                       k = 3, knnSummary = median)
 
@@
 # preProcess with near-zero value and kNN imputation
-trnXSDpp <- preProcess(trnXSD, method = c('nzv', 'knnImpute'), k = 5)
+trnXSDpp <- preProcess(trnXSD, method = c('nzv', 'knnImpute'),
+                       k = 3, knnSummary = median)
 
@@
 tstXFD <- predict(tstDumFreq, newdata = tstX)
-tstXFDpp <- preProcess(tstXFD, method = c('nzv', 'knnImpute'))
+tstXFDpp <- preProcess(tstXFD, method = c('nzv', 'knnImpute'),
+                       k = 3, knnSummary = median)
 tstXFDI <- predict(tstXFDpp, newdata = tstXFD)
@@
 tstXSD <- predict(tstDumSev, newdata = tstX[TARGET_AMT > 0])
-tstXSDpp <- preProcess(tstXSD, method = c('nzv', 'knnImpute'), k = 5)
+tstXSDpp <- preProcess(tstXSD, method = c('nzv', 'knnImpute'),
+                       k = 3, knnSummary = median)
 tstXSDI <- predict(tstXSDpp, newdata = tstXSD)

### 2020-11-03 HW4/HW4_Draft.Rmd
Change stuff.

@@
 corrplot(cor(DT[, ..nmtrnNUM], use = 'complete.obs'),
-         method = 'ellipse', type = 'lower')
+         method = 'ellipse', type = 'lower', order = 'hclust',
+         hclust.method = 'ward.D2')
 ```
@@
    * Both are very strongly tied to wealth and net worth
- * `BLUEBOOK` and `INCOME`
+ * `BLUEBOOK` with`INCOME`
    * People with more disposable income can afford more expensive vehicles
@@
    * More claims implies a greater total aggregate payment
- * `CAR_AGE` and `INCOME`
+ * `CAR_AGE` with `INCOME`
    * This is interesting. It could be that people with higher disposable incomes
@@
 
- * `OWN` with `HOME_VAL` as discussed in the missing data section
- * `KIDSDRIV` with `HOMEKIDS`
- * A three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK`
- * `CLM_FREQ` with `OLDCLAIM`
- * `MSTATUS` with `SEX`
-   * One sex may be more affected by marriage than the other
+ * Three-way interaction between `OWN`, `HOME_VAL` and `INCOME`
+   * Represents wealth
+ * Three-way interaction between `CLM_FREQ`, `OLDCLAIM`, and `MVR_PTS`
+   * Represents prior history
  * `CAR_USE` with `URBANICITY`
+   * Represents increased risk of location
 
@@
 # Set up dummy variables
-trnDumFreq <- dummyVars(TARGET_FLAG ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
-                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
-                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
-                          TARGET_AMT,
+trnDumFreq <- dummyVars(TARGET_FLAG ~ . +
+                          OWN * HOME_VAL * INCOME +
+                          CLM_FREQ * OLDCLAIM * MVR_PTS +
+                          CAR_USE * URBANICITY -
+                         INDEX - TARGET_AMT,
                         fullRank = TRUE, data = trnX)
@@
 
-```{r model1RunFreq, cache=TRUE}
+```{r model1RunFreq}
 # preProcess with near-zero value and kNN imputation
@@
                        k = 3, knnSummary = median)
-
 # Create processed predictors
@@
 one's license revoked, living in an urban area, or increased travel times are
-all *positive* risk factors while higher levels of wealth or education are
+all *positive* risk factors while higher levels of wealth or being married are
 *negative* risk factors.
@@
 Some interesting observations. The predictors with the greatest effect on risk
-are `URBANICITY` and `INCOME`. The value of the coefficient for `SEX` implies
-that being female decreases the probability of a claim, as suspected, but it is
-only significant as a predictor at the 10% level. However, it is kept in the
-model as removing it increases the AIC. The hypothesized interaction between
-`MSTATUS` and `SEX` does lower the deviance, but not enough to justify the
-additional parameter. `BLUEBOOK` value is a negative risk factor. Perhaps people
-with more expensive cars drive more carefully?
+are `URBANICITY` and the increased factor when it is a commercial vehicle. The
+value of the coefficient for `SEX` implies that being female decreases the
+probability of a claim, as suspected, but it is only significant as a predictor
+at the 10% level. However, it is kept in the model as removing it increases the
+AIC. `BLUEBOOK`, `INCOME`, and Having a `HOME_VAL` greater than 0 (as indicated
+by the `OWN` variable being Yes) are all negative indicators, which supports the
+hypothesis that people with greater wealth have fewer accidents. Those values
+are tempered by some of the wealth interactions.
 
-The full three-way interaction between `HOME_VAL`, `INCOME`, and `BLUEBOOK` was
-deemed insignificant in light of the existing two-way interactions which are in
-the model.
-
-The interaction between `CLM_FREQ` and `OLDCLAIM` did help the model AIC, and
-the coefficient indicates that while `OLDCLAIM` did not justify the additional
-parameter, its tempering of `CLM_FREQ` **does** justify the additional
-parameter.
+Every level of the interactions between `CLM_FREQ`, `OLDCLAIM`, and `MVR_PTS`
+are significant.  What this may be telling us is that a linear regression is
+***not*** the proper model for this analysis, and the algorithm is struggling to
+add non-linearity via the interactions.
 
@@
 
+Lastly, red cars do not contribute to increased claims frequency in this model.
+
 #### Variable Importance
@@
 the stepping procedure shown below. The further down the table the variable is,
-the more dramatic its inclusion or exclusion will be on the deviance and theAIC.
-The top 10 variables in order would be: `URBANICITY.Urban`,
-`CAR_USECommercial:URBANICITYUrban`, `REVOKED.Yes`, `TRAVTIME`,
-`CAR_TYPE.Sports Car`, `CLM_FREQ`, `TIF`, `MVR_PTS`, `CAR_TYPE.SUV`,
-and `JOB.Blue Collar`.
+the more dramatic its inclusion or exclusion will be on the deviance and the
+AIC.
 
@@
                                       Df Deviance    AIC
-<none>                                     5152.3 5212.3
-+ HOMEKIDS                             1   5150.5 5212.5
-+ AGE                                  1   5150.8 5212.8
-+ EDUCATION.Masters                    1   5151.0 5213.0
-- SEX.F                                1   5155.2 5213.2
-+ JOB.Manager                          1   5151.6 5213.6
-+ OLDCLAIM                             1   5151.6 5213.6
-+ EDUCATION.PhD                        1   5151.7 5213.7
-+ `HOME_VAL:BLUEBOOK`                  1   5151.7 5213.7
-+ CAR_AGE                              1   5151.8 5213.8
-- `INCOME:BLUEBOOK`                    1   5155.9 5213.9
-+ OWN.Yes                              1   5152.1 5214.1
-+ `MSTATUSYes:SEXF`                    1   5152.2 5214.2
-+ RED_CAR.yes                          1   5152.3 5214.3
-+ `EDUCATION.High School`              1   5152.3 5214.3
-+ YOJ                                  1   5152.3 5214.3
-+ `INCOME:HOME_VAL:BLUEBOOK`           1   5152.3 5214.3
-+ CAR_USE.Commercial                   1   5152.3 5214.3
-- `INCOME:HOME_VAL`                    1   5158.7 5216.7
-- JOB.Student                          1   5159.0 5217.0
-- JOB.Lawyer                           1   5159.1 5217.1
-- `CAR_TYPE.Panel Truck`               1   5160.0 5218.0
-- MSTATUS.Yes                          1   5162.6 5220.6
-- CAR_TYPE.Van                         1   5166.3 5224.3
-- BLUEBOOK                             1   5166.6 5224.6
-- `JOB.Home Maker`                     1   5166.7 5224.7
-- `OLDCLAIM:CLM_FREQ`                  1   5167.4 5225.4
-- EDUCATION.Bachelors                  1   5169.0 5227.0
-- HOME_VAL                             1   5169.5 5227.5
-- PARENT1.Yes                          1   5169.9 5227.9
-- INCOME                               1   5170.3 5228.3
-- CAR_TYPE.Pickup                      1   5170.6 5228.6
-- JOB.Professional                     1   5174.0 5232.0
-- JOB.Clerical                         1   5181.7 5239.7
-- KIDSDRIV                             1   5186.1 5244.1
-- `JOB.Blue Collar`                    1   5189.2 5247.2
-- CAR_TYPE.SUV                         1   5189.8 5247.8
-- MVR_PTS                              1   5191.2 5249.2
-- TIF                                  1   5191.8 5249.8
-- CLM_FREQ                             1   5193.0 5251.0
-- `CAR_TYPE.Sports Car`                1   5195.8 5253.8
-- TRAVTIME                             1   5198.9 5256.9
-- REVOKED.Yes                          1   5220.1 5278.1
-- `CAR_USECommercial:URBANICITYUrban`  1   5229.9 5287.9
-- URBANICITY.Urban                     1   5422.8 5480.8
+<none>                                     5131.6 5199.6
+- `OLDCLAIM:CLM_FREQ:MVR_PTS`          1   5134.4 5200.4
++ EDUCATION.Masters                    1   5130.6 5200.6
++ HOMEKIDS                             1   5130.7 5200.7
++ CAR_AGE                              1   5130.8 5200.8
+- AGE                                  1   5134.9 5200.9
++ JOB.Manager                          1   5131.0 5201.0
+- SEX.F                                1   5135.0 5201.0
++ HOME_VAL                             1   5131.1 5201.1
++ `HOME_VAL:OWNYes`                    1   5131.1 5201.1
++ EDUCATION.PhD                        1   5131.3 5201.3
++ YOJ                                  1   5131.6 5201.6
++ OWN.Yes                              1   5131.6 5201.6
++ EDUCATION.HighSchool                 1   5131.6 5201.6
++ RED_CAR.yes                          1   5131.6 5201.6
++ CAR_USE.Commercial                   1   5131.6 5201.6
+- OLDCLAIM                             1   5135.9 5201.9
+- JOB.Lawyer                           1   5137.6 5203.6
+- `OLDCLAIM:MVR_PTS`                   1   5138.4 5204.4
+- CAR_TYPE.PanelTruck                  1   5138.7 5204.7
+- `CLM_FREQ:MVR_PTS`                   1   5138.9 5204.9
+- `OLDCLAIM:CLM_FREQ`                  1   5139.0 5205.0
+- JOB.Student                          1   5140.2 5206.2
+- BLUEBOOK                             1   5141.1 5207.1
+- INCOME                               1   5143.0 5209.0
+- JOB.HomeMaker                        1   5143.1 5209.1
+- CAR_TYPE.Van                         1   5144.0 5210.0
+- PARENT1.Yes                          1   5144.7 5210.7
+- MSTATUS.Yes                          1   5145.2 5211.2
+- EDUCATION.Bachelors                  1   5148.1 5214.1
+- `INCOME:HOME_VAL`                    1   5148.6 5214.6
+- CAR_TYPE.Pickup                      1   5149.8 5215.8
+- JOB.Professional                     1   5152.2 5218.2
+- JOB.Clerical                         1   5156.9 5222.9
+- `INCOME:OWNYes`                      1   5156.9 5222.9
+- KIDSDRIV                             1   5164.6 5230.6
+- JOB.BlueCollar                       1   5166.6 5232.6
+- CLM_FREQ                             1   5168.5 5234.5
+- TIF                                  1   5169.5 5235.5
+- CAR_TYPE.SUV                         1   5169.8 5235.8
+- MVR_PTS                              1   5176.4 5242.4
+- CAR_TYPE.SportsCar                   1   5176.9 5242.9
+- TRAVTIME                             1   5179.5 5245.5
+- REVOKED.Yes                          1   5196.8 5262.8
+- `CAR_USECommercial:URBANICITYUrban`  1   5207.3 5273.3
+- URBANICITY.Urban                     1   5388.1 5454.1
 ```
@@
 ### Severity Model
-In the insurance world, severity is not necessarily fit via linear models, but
-is more often fit through comparison of the maximum likelihood fits of various
-distributional families to the size of the claims, but the task here is to fit a
-linear regression model to estimate future claim size.
+In the insurance world, severity is not necessarily fit via linear regression.
+Rather, it is more often fit through comparison of the maximum likelihood fits
+of various distributional families to the size of the claims, but the task here
+is to fit a linear regression model to estimate future claim size.
 
@@
 the model will be selected via a forward and backward stepwise AIC algorithm
-starting with the same predictors and interactions as in the frequency model.
+starting with the all the predictors. Experimentation with interactions only had
+worse effects, so no interaction will be considered.
 
@@
 # Set up dummy variables
-trnDumSev <- dummyVars(TARGET_AMT ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
-                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
-                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
-                          TARGET_FLAG,
+trnDumSev <- dummyVars(TARGET_AMT ~ . - INDEX - TARGET_FLAG,
                         fullRank = TRUE, data = trnX[TARGET_AMT > 0])
@@
 ```
-
 ```{r runModel1Sev}
-trC <- trainControl(method = 'none')
+trC <- trainControl(method = 'none', summaryFunction = defaultSummary,
+                    classProbs = FALSE)
 set.seed(878)
@@
 
-The predictive variables with the greatest effect on claim size are `OLD_CLAIM`,
-`HOME_VAL`, and `BLUEBOOK`. The first and third make the most sense. One who has
-had larger claims in the past is likely to have larger claims in the future.
-Also, the claim for damages is very strongly correlated with the value of the
-vehicle. `HOME_VAL` is a bit strange. Perhaps it's serving as a proxy for
-purchased coverage limit. Maybe people with more expensive homes, who buy larger
-limits on their homeowners policies, also purchase larger limits on their
-vehicular polices.
-
-The interaction between `INCOME` and `HOME_VAL` serves to temper the effect of
-`HOME_VAL`. Similarly for the interaction between `OLDCLAIM` and `CLM_FREQ`.
+Unlike the frequency model, many of the resulting predictors are not significant
+at even the 10% level. What this may be telling us is that a linear regression is
+***not*** the proper model for this analysis, and the algorithm is struggling to
+add non-linearity via the interactions. In practice, at the very least a GLM
+with a gamma link would be tried or, more sophisticated predictive models such
+as tree-based ones would be investigated. Most likely, MLE would be used to
+estimate severity and a hierarchial contingent stocahstic cash flow model would
+be built such that given a claim, a loss would be generated from a much more
+amenable distribution, such as a Pareto or a Burr.
 
@@
 In terms of variable importance, it can be estimated from the the last trace of
-the stepping procedure shown below. There are only eight variables in the model,
-and in order of AIC contribution they are `BLUEBOOK`, `OLDCLAIM`, `HOME_VAL`,
-`INCOME:HOME_VAL`, `CAR_USE.Commercial`, `OLDCLAIM:CLM_FREQ`, `MSTATUS.Yes`, and
-` REVOKED.Yes`.
+the stepping procedure shown below. The further down the list a variable
+preceded by a `-` is, the more drastic of an effect its removal would have on
+the model AIC.
 
@@
                        Df   Deviance   AIC
-<none>                                   9.5456e+10 31329
-- REVOKED.Yes                          1 9.5604e+10 31329
-+ JOB.Manager                          1 9.5357e+10 31329
-- MSTATUS.Yes                          1 9.5620e+10 31330
-+ CAR_AGE                              1 9.5370e+10 31330
-+ `MSTATUSYes:SEXF`                    1 9.5382e+10 31330
-- `OLDCLAIM:CLM_FREQ`                  1 9.5636e+10 31330
-- CAR_USE.Commercial                   1 9.5639e+10 31330
-+ SEX.F                                1 9.5389e+10 31330
-+ JOB.Professional                     1 9.5391e+10 31330
-+ CAR_TYPE.Pickup                      1 9.5400e+10 31330
-+ MVR_PTS                              1 9.5402e+10 31330
-+ CAR_TYPE.Van                         1 9.5403e+10 31330
-+ HOMEKIDS                             1 9.5404e+10 31330
-+ CLM_FREQ                             1 9.5420e+10 31330
-+ `CAR_USECommercial:URBANICITYUrban`  1 9.5426e+10 31330
-+ RED_CAR.yes                          1 9.5430e+10 31331
-+ JOB.Student                          1 9.5432e+10 31331
-- `INCOME:HOME_VAL`                    1 9.5686e+10 31331
-+ EDUCATION.Masters                    1 9.5435e+10 31331
-- HOME_VAL                             1 9.5695e+10 31331
-+ PARENT1.Yes                          1 9.5442e+10 31331
-+ JOB.Clerical                         1 9.5445e+10 31331
-+ `CAR_TYPE.Panel Truck`               1 9.5446e+10 31331
-- OLDCLAIM                             1 9.5700e+10 31331
-+ `EDUCATION.High School`              1 9.5446e+10 31331
-+ OWN.Yes                              1 9.5447e+10 31331
-+ TRAVTIME                             1 9.5448e+10 31331
-+ `INCOME:HOME_VAL:BLUEBOOK`           1 9.5450e+10 31331
-+ `JOB.Home Maker`                     1 9.5450e+10 31331
-+ URBANICITY.Urban                     1 9.5453e+10 31331
-+ `INCOME:BLUEBOOK`                    1 9.5454e+10 31331
-+ AGE                                  1 9.5454e+10 31331
-+ `HOME_VAL:BLUEBOOK`                  1 9.5455e+10 31331
-+ `JOB.Blue Collar`                    1 9.5455e+10 31331
-+ `CAR_TYPE.Sports Car`                1 9.5455e+10 31331
-+ KIDSDRIV                             1 9.5456e+10 31331
-+ JOB.Lawyer                           1 9.5456e+10 31331
-+ EDUCATION.Bachelors                  1 9.5456e+10 31331
-+ YOJ                                  1 9.5456e+10 31331
-+ CAR_TYPE.SUV                         1 9.5456e+10 31331
-+ INCOME                               1 9.5456e+10 31331
-+ `KIDSDRIV:HOMEKIDS`                  1 9.5456e+10 31331
-+ TIF                                  1 9.5456e+10 31331
-+ EDUCATION.PhD                        1 9.5456e+10 31331
-- BLUEBOOK                             1 9.6115e+10 31337
+<none>                    9.5799e+10 31328
+- CAR_AGE               1 9.5927e+10 31328
+- OWN.Yes               1 9.5938e+10 31329
++ JOB.Manager           1 9.5696e+10 31329
+- SEX.F                 1 9.5953e+10 31329
+- MSTATUS.Yes           1 9.5960e+10 31329
++ MVR_PTS               1 9.5712e+10 31329
++ CAR_USE.Commercial    1 9.5729e+10 31329
++ REVOKED.Yes           1 9.5735e+10 31329
++ CAR_TYPE.Pickup       1 9.5748e+10 31330
++ JOB.Professional      1 9.5754e+10 31330
++ CAR_TYPE.Van          1 9.5757e+10 31330
++ HOMEKIDS              1 9.5763e+10 31330
++ CAR_TYPE.PanelTruck   1 9.5769e+10 31330
++ JOB.Student           1 9.5778e+10 31330
++ CLM_FREQ              1 9.5779e+10 31330
++ CAR_TYPE.SUV          1 9.5784e+10 31330
++ EDUCATION.Bachelors   1 9.5787e+10 31330
++ CAR_TYPE.SportsCar    1 9.5787e+10 31330
++ TRAVTIME              1 9.5788e+10 31330
++ PARENT1.Yes           1 9.5788e+10 31330
++ EDUCATION.HighSchool  1 9.5789e+10 31330
++ JOB.BlueCollar        1 9.5789e+10 31330
++ JOB.HomeMaker         1 9.5790e+10 31330
++ INCOME                1 9.5793e+10 31330
++ JOB.Lawyer            1 9.5793e+10 31330
++ OLDCLAIM              1 9.5794e+10 31330
++ TIF                   1 9.5795e+10 31330
++ KIDSDRIV              1 9.5796e+10 31330
++ YOJ                   1 9.5796e+10 31330
++ EDUCATION.PhD         1 9.5796e+10 31330
++ EDUCATION.Masters     1 9.5797e+10 31330
++ JOB.Clerical          1 9.5797e+10 31330
++ AGE                   1 9.5797e+10 31330
++ HOME_VAL              1 9.5797e+10 31330
++ URBANICITY.Urban      1 9.5798e+10 31330
++ RED_CAR.yes           1 9.5798e+10 31330
+- BLUEBOOK              1 9.6748e+10 31341
 ```
@@
                   labels = c("NoClaim", "Claim"))
-tstDumFreq <- dummyVars(TARGET_FLAG ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
-                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
-                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
-                          TARGET_AMT,
+tstDumFreq <- dummyVars(TARGET_FLAG ~ . +
+                          OWN * HOME_VAL * INCOME +
+                          CLM_FREQ * OLDCLAIM * MVR_PTS +
+                          CAR_USE * URBANICITY -
+                         INDEX - TARGET_AMT,
                         fullRank = TRUE, data = tstX)
@@
                                      predictor = m1FpredProb[, 1])
+m1FCM$table
 ```
@@
 np <- summary(m1SevFit$finalModel)$df[3]
-tstDumSev <- dummyVars(TARGET_AMT ~ . + OWN * HOME_VAL + KIDSDRIV * HOMEKIDS +
-                          HOME_VAL * INCOME * BLUEBOOK + CLM_FREQ * OLDCLAIM +
-                          MSTATUS * SEX + CAR_USE * URBANICITY - INDEX -
-                          TARGET_FLAG,
-                        fullRank = TRUE, data = tstX[TARGET_AMT > 0])
+tstDumSev <- dummyVars(TARGET_AMT ~ . - INDEX - TARGET_FLAG,
+                       data = tstX[TARGET_AMT > 0])
 tstXSD <- predict(tstDumSev, newdata = tstX[TARGET_AMT > 0])
@@
 m1Spred <- predict(m1SevFit, newdata = tstXSDI)
-
 # Model 1 Severity: Calculate accuracy statistics

### 2020-11-07 HW4/HW4_Draft.Rmd
model #2 & copyediting

@@
     Bryan Persaud, Jeff Shamp
-date: "11/01/2020"
+date: "11/08/2020"
 output:
@@
 library(data.table)
+library(mice)
+
 # Set master seed
@@
 nmtrnFAC <- setdiff(nmtrn, nmtrnNUM)
-# Drop INDEX and target variablesfrom name list
+# Drop INDEX and target variables from name list
 nmtrnNUM <- nmtrnNUM[-(1:3)]
@@
 |VARIABLE NAME|DEFINITION|THEORETICAL EFFECT|
-|-|-|-|
+|-|-|--|
 |INDEX|Identification Variable (do not use)|None|
@@
 A variable such as `OLDCLAIM` is expected to have a mass point at 0. Every
-observation which had fewer than 2 claims clearly did not have a prior claim!
+observation that had fewer than 2 claims clearly did not have a prior claim!
 
 ### Factor Predictors
-While tabular depiction of factor variables is usually of little value, visual
+While a tabular depiction of factor variables is usually of little value, visual
 representation of their distributions is of use.
@@
  * `HOMEKIDS` with `KIDSDRIV`
-   * This is self-explanatory
+   * This is self-explanatory.
  * `HOME_VAL` with `INCOME`
-   * Both are very strongly tied to wealth and net worth
+   * Both are very strongly tied to wealth and net worth.
  * `BLUEBOOK` with`INCOME`
-   * People with more disposable income can afford more expensive vehicles
+   * People with more disposable income can afford more expensive vehicles.
  * `CLM_FREQ` with `OLDCLAIM`
-   * More claims implies a greater total aggregate payment
+   * More claims implies a greater total aggregate payment.
  * `CAR_AGE` with `INCOME`
@@
    * Probably when one is in a lot of accidents, the chances of them being at
-   some level of fault rises
+   some level of fault rises.
  * `HOMEKIDS` **negatively** correlated with `AGE`
@@
 The 70/30 split will be honored as all claims with a `TARGET_FLAG` of 1 will
-have non-zero `TARGET_AMT`, However, there will be no guarantee that the
-*distribution* of the loss conditional on event in the validation set will be
+have non-zero `TARGET_AMT`. However, there will be no guarantee that the
+*distribution* of the loss conditional on the event in the validation set will be
 similar to the training set as a whole.
@@
 
-## Models Set 1
+## Model Set #1
 ### Missing Data
@@
 opposed to it being missing, although we would want to investigate this further
-were we allowed. Therefore, we are going to add a "Own" variable which will be
+were we allowed. Therefore, we are going to add an "Own" variable which will be
 `Yes` for all positive `HOME_VAL` and `No` otherwise, and allow an interaction
@@
                      levels = c(0, 1), labels = c('No', 'Yes'))]
+```
+
+## Model Set #2  
+### Missing Data
+
+Like Model Set #1, the data preparation efforts for Model Set #2 also includes the 
+imputation of missing values for the predictor variables. However, because it combines 
+the standard linear regression and the nearest-neighbor imputation approach, Model Set #2 
+elected to fill in missing values of a continuous variable by using the predictive mean 
+matching imputation method. Moreover, there will be the derivation of three new binary 
+factor variables.
+
+* `AGE`  
+  + Since the data set is comprised of auto insurance policyholder information, the 
+  missing `AGE` values required imputation. However, given the relatively small number 
+  of missing values, imputation by predictive mean matching is conducted.
+
+* `HOME_VAL`  
+  + A `HOME_VAL` response of '0' makes it understandable to assume that those missing 
+  entries could indicate that the policyholder is a renter rather than a homeowner. 
+  Therefore, missing values were replaced with zeroes. The new `OWN` variables created 
+  in Model Set #1 is kept in Model Set #2.
+
+* `CAR_AGE`, `INCOME` & `YOJ`  
+  + Because there is a larger number of missing values (NA’s) were identified, 
+  imputation by predictive mean matching is conducted.
+
+  + Moreover, there is a single negative value ('-3') for `CAR_AGE` which seemed 
+  implausible and was assumed to be a typographical error. Thus, the absolute value 
+  of `CAR_AGE` is taken.
+
+### New Variables
+
+For Model Set #2, new factor variables ($VARIABLE`.f`) were create as a way to better 
+capture policyholder's profile in the binary logistic regression modeling process.
+
+* `CAR_AGE.f`  
+  + Because `CAR_AGE` is dominated by values of 1 (i.e. car being one year old or less), 
+  the `CAR_AGE.f` variable is derived as a factor variable to be used during the binary 
+  logistic regression modeling process. Therefore, it is interpreted as:
+
+    `CAR_AGE.f` = 0: The vehicle is 1 year old or less.  
+    `CAR_AGE.f` = 1: The vehicle is more than a 1-year-old.
+
+* `HOMEKIDS.f` & `KIDSDRIV.f` 
+   + Both the `HOMEKIDS` and `KIDSDRIV` variables indicated that large percentages of the 
+   values of each were 0. Therefore, both variables were converted to binary variables. 
+   The following is their binary interpretations:
+
+      `HOMEKIDS.f` = 0: The policyholder has no child living at home.  
+      `HOMEKIDS.f` = 1: The policyholder has 1 or more children living at home.  
+      `KIDSDRIV.f` = 0: No teenage drivers use the policyholder’s insured vehicle.  
+      `KIDSDRIV.f` = 1: One or more teenage drivers use the policyholder’s insured vehicle.
+
+```{r m2clean}
+# Clean Training set
+m2.train.X <- trnX
+m2.train.X$HOME_VAL[is.na(m2.train.X$HOME_VAL)] <- 0
+m2.train.X$OWN[is.na(m2.train.X$OWN)] <- "No"
+m2.train.X$CAR_AGE = abs(m2.train.X$CAR_AGE) 
+m2.train.r <- factor(m2.train.X$TARGET_FLAG, levels = c(0, 1), 
+                     labels = c("NoClaim", "Claim"))
+m2.train.X.impute <- mice(m2.train.X, method = 'pmm', print = FALSE)
+m2.train.X <- complete(m2.train.X.impute)
+m2.train.X$CAR_AGE.f <- factor(ifelse(m2.train.X$CAR_AGE > 1, 1, 0),
+                              levels = c(0, 1), 
+                              labels = c('1 Year or less', 'More than a Year'))
+m2.train.X$HOMEKIDS.f <- factor(ifelse(m2.train.X$HOMEKIDS >= 1, 1, 0),
+                              levels = c(0, 1), 
+                              labels = c('None', 'At least 1'))
+m2.train.X$KIDSDRIV.f <- factor(ifelse(m2.train.X$KIDSDRIV >= 1, 1, 0),
+                              levels = c(0, 1), 
+                              labels = c('None', 'At least 1'))
+m2.train.p <- m2.train.X[, -c(1:3)]
+
+# Clean Test Set
+m2.test.X <- tstX
+m2.test.X$HOME_VAL[is.na(m2.test.X$HOME_VAL)] <- 0
+m2.test.X$OWN[is.na(m2.test.X$OWN)] <- "No"
+m2.test.X$CAR_AGE = abs(m2.test.X$CAR_AGE) 
+m2.test.r <- factor(m2.test.X$TARGET_FLAG, levels = c(0, 1), 
+                     labels = c("NoClaim", "Claim"))
+m2.test.X.impute <- mice(m2.test.X, method = 'pmm', print = FALSE)
+m2.test.X <- complete(m2.test.X.impute)
+m2.test.X$CAR_AGE.f <- factor(ifelse(m2.test.X$CAR_AGE > 1, 1, 0),
+                              levels = c(0, 1), 
+                              labels = c('1 Year or less', 'More than a Year'))
+m2.test.X$HOMEKIDS.f <- factor(ifelse(m2.test.X$HOMEKIDS >= 1, 1, 0),
+                              levels = c(0, 1), 
+                              labels = c('None', 'At least 1'))
+m2.test.X$KIDSDRIV.f <- factor(ifelse(m2.test.X$KIDSDRIV >= 1, 1, 0),
+                              levels = c(0, 1), 
+                              labels = c('None', 'At least 1'))
+m2.test.p <- m2.test.X[, -c(1:3)]
 ```
@@
  * Three-way interaction between `CLM_FREQ`, `OLDCLAIM`, and `MVR_PTS`
-   * Represents prior history
+   * Represents a prior history
  * `CAR_USE` with `URBANICITY`
@@
 #### Training  
-At this point the code will be set up to preprocess the data and train the
+At this point, the code will be set up to preprocess the data and train the
 model. The two pre-processing steps which will be done will be to check for
@@
 missing values. Even though the `caret` package allows the preprocessing to
-occur in the call to `train`, it will be seperated here for clarity. The actual
+occur in the call to `train`, it will be separated here for clarity. The actual
 code may be found in the CODE APPENDIX.
@@
 
-Most of the coefficients are of the expected direction. For example, having
+Most of the coefficients are in the expected direction. For example, having
 one's license revoked, living in an urban area, or increased travel times are
@@
 Every level of the interactions between `CLM_FREQ`, `OLDCLAIM`, and `MVR_PTS`
-are significant.  What this may be telling us is that a linear regression is
-***not*** the proper model for this analysis, and the algorithm is struggling to
+are significant.  What this may be telling us is that linear regression is
+***not*** the proper model for this analysis and the algorithm is struggling to
 add non-linearity via the interactions.
@@
 the model will be selected via a forward and backward stepwise AIC algorithm
-starting with the all the predictors. Experimentation with interactions only had
+starting with all the predictors. Experimentation with interactions only had
 worse effects, so no interaction will be considered.
@@
 #### Coefficient Disucssion  
-It is not surprising that this model has many fewer predictors. Information
-which helps in identifying the *existence* of a claim often does not help in
+It is not surprising that this model has fewer predictors. Information which 
+helps in identifying the *existence* of a claim often does not help in
 estimating the *magnitude* of the claim.
@@
 Unlike the frequency model, many of the resulting predictors are not significant
-at even the 10% level. What this may be telling us is that a linear regression is
-***not*** the proper model for this analysis, and the algorithm is struggling to
+at even the 10% level. What this may be telling us is that linear regression is
+***not*** the proper model for this analysis and the algorithm is struggling to
 add non-linearity via the interactions. In practice, at the very least a GLM
@@
 as tree-based ones would be investigated. Most likely, MLE would be used to
-estimate severity and a hierarchial contingent stocahstic cash flow model would
+estimate the severity and a hierarchial contingent stochastic cash flow model would
 be built such that given a claim, a loss would be generated from a much more
@@
 #### Variable Importance
-In terms of variable importance, it can be estimated from the the last trace of
+In terms of variable importance, it can be estimated from the last trace of
 the stepping procedure shown below. The further down the list a variable
@@
 
+## Model Set #2
+### Frequency Model
+
+The frequency model is built to determine whether a vehicle is likely to crash. In Model 
+Set #2, to work with complete separation in the logistic regression model, a Bayesian analysis 
+is fitted. The Bayesian statistical model returns samples of the parameters of interest (the
+"posterior" distribution) based on some "prior" distribution which is then updated by the 
+data. Here, Cauchy distribution is accepted as a prior for parameters of the generalized 
+linear model.
+
+```{r model2RunFreq}
+# Train Model Set #2
+set.seed(525)
+m2FreqFit <- train(x = m2.train.p, y = m2.train.r, 
+                   trControl = trainControl(method = 'repeatedcv',
+                                            classProbs = TRUE, 
+                                            number = 10),
+                   method = 'bayesglm', family = 'binomial')
+# Results
+kable(summary(m2FreqFit$finalModel)$coefficients, digits = 3L,
+      caption = "Model 2 Frequency Regression Output")
+```
+
+#### Coefficient Discussion
+
+Contributing coefficients are those which lie within the 95% level of significance. A 
+positive coefficient indicates that as the value of the predictor increases, the response 
+variable also tends to increase. A negative coefficient suggests that as the predictor 
+increases, the response variable tends to decrease.
+
+Immediately, the model follows the suggested theories behind specific variables. For example,
+policyholders who are married, have a higher income and an old claim is less likely to be 
+involved in an accident. Moreover, more educated, at least with a Bachelor’s degree, 
+policyholders drive more safely. The higher the bluebook value of the policyholder’s vehicle, 
+it decreases the probability of a claim. This makes sense naturally since more intelligent 
+individuals would seek to protect relatively valuable assets they are responsible for. Last, 
+the longer a policyholder has been a customer of the company, the less likely they are to be 
+involved in an accident. Since insurance companies would typically impose larger rates or 
+cancel the policies of careless drivers.
+
+On the contrary, and unsurprising, policyholders who have previously had their driver’s licenses 
+revoked increases the probability of a claim. This parallels the result that the more motor 
+vehicle record points a policyholder has gained, they are assumed to be unsafe drivers and are 
+more likely to be involved in an accident. If a policyholder lives in an urban area, they are
+prone to an accident since traffic and pedestrian congestion are more common. Another expected 
+result, is that the longer a policyholder’s commute to work is, there is an increased likelihood 
+of a driver being involved in an accident. The last connection that can be inferred is that 
+drivers of a larger type of vehicles are more prone to accidents than are drivers of other 
+types of vehicles, with minivans being the least prone. The results seem to suggest that owners
+of sports cars and SUVs are riskier types of drivers. Interestingly, commercial usage of 
+vehicle present critical reason for more frequent accidents to occur because of driver error 
+or non-performance, and failure to maintain vehicle causing issues or malfunction.
+
+### Severity Model
+
+Similar to Model Set #1, the multivariate linear regression model used the data set’s 
+`TARGET_AMT` attribute as the response variable and the data was split to include only those
+records that have a claim. The elastic net is a regularized regression method that linearly 
+combines the L1 and L2 regularization penalties. Predictors, specifically qualitative variables,
+were automatically transformed into dummy variables, which is important because `glmnet()` can 
+only take numerical, quantitative inputs. Cross-validation is done on different combinations 
+of $\lambda_1$ and $\lambda_2$ to find the best values. The hybrid elastic net regression is 
+especially good at dealing with situations when there are correlations between parameters.
+
+```{r model2RunSev}
+# Train Model Set #2
+temp <- split(m2.train.X, m2.train.X$TARGET_AMT > 0)[[2]] 
+m2.train.r <- temp$TARGET_AMT
+m2.train.p <- model.matrix(temp$TARGET_AMT~., temp[, -c(1:3)])[,-1]
+set.seed(525)
+m2SevFit <- train(x = m2.train.p, y = m2.train.r, 
+                   trControl = trainControl(method = 'repeatedcv',
+                                            number = 10),
+                   method = 'glmnet',
+                   tuneLength = 10,
+                   preProcess = c("center", "scale"))
+# Results
+coeffs = coef(m2SevFit$finalModel, m2SevFit$bestTune$lambda)
+kable(data.frame(cbind(variables = coeffs@Dimnames[[1]][coeffs@i+1], 
+                       coef = round(coeffs@x,3))), 
+      caption = "Model 2 Severity Regression Output")
+```
+
+#### Coefficient Disucssion
+
+Once again, contributing coefficients are those which lie within the 95% level of significance. 
+In this case, the linear model reduced the number of predictors that can predict the cost given 
+an accident. The model fits extremely poor and does a mediocre job at explaining the variability 
+within the training data. If interpretations are still to be made, first, the intercept itself 
+suggests that with no information on the policyholder, given a vehicle accident, the likely claim 
+value to the individual is $5,728. The other predictors suggest that the higher the bluebook value 
+of the policyholder’s vehicle, the larger the insurance payout will be in the event of an accident. 
+If they use the vehicle for commercial, they will also have a larger claim amount. While being a 
+female resulted as a significant variable, its coefficient is near zero and will be ignored.
+
 # SELECT MODELS
@@
 select the model which performs best two out of three in those metrics. For the
-*regression* algorithm we will look at RMSE and MAE (since the values must
+*regression* algorithm, we will look at RMSE and MAE (since the values must
 be positive) and select the model which performs best on both, Ties will be
@@
 ### Frequency Model
+
+```{r validateM2F}
+# Model 2 Frequency
+# Prediction
+m2Fpred <- predict(m2FreqFit, newdata = m2.test.X)
+m2FpredProb <- predict(m2FreqFit, newdata = m2.test.X, type = 'prob')
+
+# Accuracy Statistics
+m2FCM <- confusionMatrix(m2Fpred, m2.test.r, mode = 'everything')
+freqCompTable[2, 2] <- m2FACC <- sum(diag(m2FCM$table)) / sum(m2FCM$table)
+freqCompTable[2, 3] <- m2FF1 <- m2FCM$byClass[7]
+freqCompTable[2, 4] <- m2FAUC <- auc(response = m2.test.r,
+                                     predictor = m2FpredProb[, 1])
+m2FCM$table
+```
+
+Frequency Model 2 returns an accuracy of `r round(m2FACC, 4)`, an F1 score of
+`r round(m2FF1, 4)`, and an AUC of `r round(m2FAUC, 4)`.
+
 ### Severity Model
+
+```{r validateM2S}
+# Model 2 Severity
+temp <- split(m2.test.X, m2.test.X$TARGET_AMT > 0)[[2]] 
+m2.test.r <- temp$TARGET_AMT
+m2.test.p <- model.matrix(temp$TARGET_AMT~., temp[, -c(1:3)])[,-1]
+
+# Prediction
+m2Spred <- predict(m2SevFit, newdata = m2.test.p)
+
+# Accuracy Statistics
+m2Ssum <- postResample(m2Spred, tstSev)
+sevCompTable[2, 2] <- m2SRMSE <- RMSE(m2Spred, m2.test.r)
+sevCompTable[2, 3] <- m2SMAE <- MAE(m2Spred, m2.test.r)
+sevCompTable[2, 4] <- m2SAR2 <- R2(m2Spred, m2.test.r)
+```
+
+Severity Model 2 returns an RMSE of `r round(m2SRMSE, 1)`, an MAE of
+`r round(m2SMAE, 1)`, and an adjusted \(R^2\) of `r round(m2SAR2, 4)`.
 
@@
 ```
+```{r m2clean}
+```
 ```{r model1DummyFreq}
@@
 ```
+```{r model2RunFreq}
+```
+```{r model2RunSev}
+```
 ```{r setUpCompTables}
@@
 ```
+```{r validateM2F}
+```
+```{r validateM2S}
+```
 ```{r freqComp}

### 2020-11-08 HW4/HW4_Draft.Rmd
Minor cleanup, adding comments, calling arm library, slecting corrplot's corrplot

@@
 # Load necessary libraries
-library(corrplot)
 library(ggplot2)
@@
 library(pROC)
+library(mice)
+library(arm)
+library(corrplot)
 library(data.table)
-library(mice)
 
@@
 ```{r corrgram, fig.width=6, fig.height=4}
-corrplot(cor(DT[, ..nmtrnNUM], use = 'complete.obs'),
+corrplot::corrplot(cor(DT[, ..nmtrnNUM], use = 'complete.obs'),
          method = 'ellipse', type = 'lower', order = 'hclust',
@@
 have non-zero `TARGET_AMT`. However, there will be no guarantee that the
-*distribution* of the loss conditional on the event in the validation set will be
-similar to the training set as a whole.
+*distribution* of the loss conditional on the event in the validation set will
+be similar to the training set as a whole.
 
 ```{r trainTestSplit}
+# Create training and testing split
 set.seed(642)
@@
 ```{r model1addVars}
-# Add OWN variable for model set 1
+# Model Set 1: Add OWN variable
 trnX[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0),
@@
 ### Missing Data  
-
-Like Model Set #1, the data preparation efforts for Model Set #2 also includes the 
-imputation of missing values for the predictor variables. However, because it combines 
-the standard linear regression and the nearest-neighbor imputation approach, Model Set #2 
-elected to fill in missing values of a continuous variable by using the predictive mean 
-matching imputation method. Moreover, there will be the derivation of three new binary 
-factor variables.
+Like Model Set #1, the data preparation efforts for Model Set #2 also includes
+the imputation of missing values for the predictor variables. However, because
+it combines the standard linear regression and the nearest-neighbor imputation
+approach, Model Set #2 elected to fill in missing values of a continuous
+variable by using the predictive mean matching imputation method. Moreover,
+there will be the derivation of three new binary factor variables.
 
 * `AGE`  
-  + Since the data set is comprised of auto insurance policyholder information, the 
-  missing `AGE` values required imputation. However, given the relatively small number 
-  of missing values, imputation by predictive mean matching is conducted.
+  + Since the data set is comprised of auto insurance policyholder information,
+  the missing `AGE` values required imputation. However, given the relatively
+  small number of missing values, imputation by predictive mean matching is
+  conducted.
 
 * `HOME_VAL`  
-  + A `HOME_VAL` response of '0' makes it understandable to assume that those missing 
-  entries could indicate that the policyholder is a renter rather than a homeowner. 
-  Therefore, missing values were replaced with zeroes. The new `OWN` variables created 
-  in Model Set #1 is kept in Model Set #2.
+  + A `HOME_VAL` response of '0' makes it understandable to assume that those
+  missing entries could indicate that the policyholder is a renter rather than a
+  homeowner. Therefore, missing values were replaced with zeroes. The new `OWN`
+  variables created in Model Set #1 is kept in Model Set #2.
 
@@
   + Moreover, there is a single negative value ('-3') for `CAR_AGE` which seemed 
-  implausible and was assumed to be a typographical error. Thus, the absolute value 
-  of `CAR_AGE` is taken.
+  implausible and was assumed to be a typographical error. Thus, the absolute
+  value of `CAR_AGE` is taken.
 
 ### New Variables  
-
-For Model Set #2, new factor variables ($VARIABLE`.f`) were create as a way to better 
-capture policyholder's profile in the binary logistic regression modeling process.
+For Model Set #2, new factor variables ($VARIABLE`.f`) were create as a way to 
+better capture policyholder's profile in the binary logistic regression modeling
+process.
 
 * `CAR_AGE.f`  
-  + Because `CAR_AGE` is dominated by values of 1 (i.e. car being one year old or less), 
-  the `CAR_AGE.f` variable is derived as a factor variable to be used during the binary 
-  logistic regression modeling process. Therefore, it is interpreted as:
+  + Because `CAR_AGE` is dominated by values of 1 (i.e. car being one year old
+  or less), the `CAR_AGE.f` variable is derived as a factor variable to be used
+  during the binary logistic regression modeling process. Therefore, it is
+  interpreted as:
 
@@
 * `HOMEKIDS.f` & `KIDSDRIV.f` 
-   + Both the `HOMEKIDS` and `KIDSDRIV` variables indicated that large percentages of the 
-   values of each were 0. Therefore, both variables were converted to binary variables. 
-   The following is their binary interpretations:
+   + Both the `HOMEKIDS` and `KIDSDRIV` variables indicated that large
+   percentages of the values of each were 0. Therefore, both variables were
+   converted to binary variables. The following is their binary interpretations:
 
@@
       `HOMEKIDS.f` = 1: The policyholder has 1 or more children living at home.  
-      `KIDSDRIV.f` = 0: No teenage drivers use the policyholder’s insured vehicle.  
-      `KIDSDRIV.f` = 1: One or more teenage drivers use the policyholder’s insured vehicle.
+      `KIDSDRIV.f` = 0: No teenage drivers use the policyholder’s insured
+      vehicle.  
+      `KIDSDRIV.f` = 1: One or more teenage drivers use the policyholder’s
+      insured vehicle.
 
 ```{r m2clean}
-# Clean Training set
+# Model 2: Clean Training set
 m2.train.X <- trnX
@@
 ```{r model1DummyFreq}
-# Cast target as factor for caret purposes
+# Model Set 1: Cast target as factor for caret purposes
 trnFreq <- factor(trnX$TARGET_FLAG, levels = c(0, 1),
@@
 ```{r model1RunFreq}
-# preProcess with near-zero value and kNN imputation
+# Model Set 1: preProcess with near-zero value and kNN imputation
 trnXFDpp <- preProcess(trnXFD, method = c('nzv', 'knnImpute'),
@@
 set.seed(487)
-# Train Model
+# Model Set 1: Train Model
 m1FreqFit <- train(x = trnXFDI, y = trnFreq, trControl = trC,
@@
 ```{r model1DummySev}
-# Extract those observations with values > 0
+# Model Set 1: Extract those observations with values > 0
 trnSev <- trnX[TARGET_AMT > 0, TARGET_AMT]
@@
 ```{r runModel1Sev}
+# Model Set 1: Train Model
 trC <- trainControl(method = 'none', summaryFunction = defaultSummary,
@@
 as tree-based ones would be investigated. Most likely, MLE would be used to
-estimate the severity and a hierarchial contingent stochastic cash flow model would
-be built such that given a claim, a loss would be generated from a much more
-amenable distribution, such as a Pareto or a Burr.
+estimate the severity and a hierarchial contingent stochastic cash flow model
+would be built such that given a claim, a loss would be generated from a much
+more amenable distribution, such as a Pareto or a Burr.
 
@@
 ### Frequency Model
-
-The frequency model is built to determine whether a vehicle is likely to crash. In Model 
-Set #2, to work with complete separation in the logistic regression model, a Bayesian analysis 
-is fitted. The Bayesian statistical model returns samples of the parameters of interest (the
-"posterior" distribution) based on some "prior" distribution which is then updated by the 
-data. Here, Cauchy distribution is accepted as a prior for parameters of the generalized 
-linear model.
+The frequency model is built to determine whether a vehicle is likely to crash.
+In Model Set #2, to work with complete separation in the logistic regression
+model, a Bayesian analysis is fitted. The Bayesian statistical model returns
+samples of the parameters of interest (the "posterior" distribution) based on
+some "prior" distribution which is then updated by the data. Here, Cauchy
+distribution is accepted as a prior for parameters of the generalized linear
+model.
 
 ```{r model2RunFreq}
-# Train Model Set #2
+# Model Set 2: Train Frequency Model
 set.seed(525)
@@
 #### Coefficient Discussion 
-
-Contributing coefficients are those which lie within the 95% level of significance. A 
-positive coefficient indicates that as the value of the predictor increases, the response 
-variable also tends to increase. A negative coefficient suggests that as the predictor 
-increases, the response variable tends to decrease.
-
-Immediately, the model follows the suggested theories behind specific variables. For example,
-policyholders who are married, have a higher income and an old claim is less likely to be 
-involved in an accident. Moreover, more educated, at least with a Bachelor’s degree, 
-policyholders drive more safely. The higher the bluebook value of the policyholder’s vehicle, 
-it decreases the probability of a claim. This makes sense naturally since more intelligent 
-individuals would seek to protect relatively valuable assets they are responsible for. Last, 
-the longer a policyholder has been a customer of the company, the less likely they are to be 
-involved in an accident. Since insurance companies would typically impose larger rates or 
-cancel the policies of careless drivers.
-
-On the contrary, and unsurprising, policyholders who have previously had their driver’s licenses 
-revoked increases the probability of a claim. This parallels the result that the more motor 
-vehicle record points a policyholder has gained, they are assumed to be unsafe drivers and are 
-more likely to be involved in an accident. If a policyholder lives in an urban area, they are
-prone to an accident since traffic and pedestrian congestion are more common. Another expected 
-result, is that the longer a policyholder’s commute to work is, there is an increased likelihood 
-of a driver being involved in an accident. The last connection that can be inferred is that 
-drivers of a larger type of vehicles are more prone to accidents than are drivers of other 
-types of vehicles, with minivans being the least prone. The results seem to suggest that owners
-of sports cars and SUVs are riskier types of drivers. Interestingly, commercial usage of 
-vehicle present critical reason for more frequent accidents to occur because of driver error 
-or non-performance, and failure to maintain vehicle causing issues or malfunction.
+Contributing coefficients are those which lie within the 95% level of
+significance. A positive coefficient indicates that as the value of the
+predictor increases, the response variable also tends to increase. A negative
+coefficient suggests that as the predictor increases, the response variable
+tends to decrease.
+
+Immediately, the model follows the suggested theories behind specific variables.
+For example, policyholders who are married, have a higher income and an old
+claim is less likely to be involved in an accident. Moreover, more educated, at
+least with a Bachelor’s degree, policyholders drive more safely. The higher the
+bluebook value of the policyholder’s vehicle, it decreases the probability of a
+claim. This makes sense naturally since more intelligent individuals would seek
+to protect relatively valuable assets they are responsible for. Last, the longer
+a policyholder has been a customer of the company, the less likely they are to
+be involved in an accident. Since insurance companies would typically impose
+larger rates or cancel the policies of careless drivers.
+
+On the contrary, and unsurprising, policyholders who have previously had their
+driver’s licenses revoked increases the probability of a claim. This parallels
+the result that the more motor vehicle record points a policyholder has gained,
+they are assumed to be unsafe drivers and are more likely to be involved in an
+accident. If a policyholder lives in an urban area, they are prone to an
+accident since traffic and pedestrian congestion are more common. Another
+expected result, is that the longer a policyholder’s commute to work is, there
+is an increased likelihood of a driver being involved in an accident. The last
+connection that can be inferred is that drivers of a larger type of vehicles are
+more prone to accidents than are drivers of other types of vehicles, with
+minivans being the least prone. The results seem to suggest that owners of
+sports cars and SUVs are riskier types of drivers. Interestingly, commercial
+usage of vehicle present critical reason for more frequent accidents to occur
+because of driver error or non-performance, and failure to maintain vehicle
+causing issues or malfunction.
 
 ### Severity Model  
-
-Similar to Model Set #1, the multivariate linear regression model used the data set’s 
-`TARGET_AMT` attribute as the response variable and the data was split to include only those
-records that have a claim. The elastic net is a regularized regression method that linearly 
-combines the L1 and L2 regularization penalties. Predictors, specifically qualitative variables,
-were automatically transformed into dummy variables, which is important because `glmnet()` can 
-only take numerical, quantitative inputs. Cross-validation is done on different combinations 
-of $\lambda_1$ and $\lambda_2$ to find the best values. The hybrid elastic net regression is 
-especially good at dealing with situations when there are correlations between parameters.
+Similar to Model Set #1, the multivariate linear regression model used the data
+set’s `TARGET_AMT` attribute as the response variable and the data was split to
+include only those records that have a claim. The elastic net is a regularized
+regression method that linearly combines the L1 and L2 regularization penalties.
+Predictors, specifically qualitative variables, were automatically transformed
+into dummy variables, which is important because `glmnet()` can only take
+numerical, quantitative inputs. Cross-validation is done on different
+combinations of $\lambda_1$ and $\lambda_2$ to find the best values. The hybrid
+elastic net regression is especially good at dealing with situations when there
+are correlations between parameters.
 
 ```{r model2RunSev}
-# Train Model Set #2
+# Model Set 2: Train Severity Model
 temp <- split(m2.train.X, m2.train.X$TARGET_AMT > 0)[[2]] 
@@
 #### Coefficient Disucssion
-
-Once again, contributing coefficients are those which lie within the 95% level of significance. 
-In this case, the linear model reduced the number of predictors that can predict the cost given 
-an accident. The model fits extremely poor and does a mediocre job at explaining the variability 
-within the training data. If interpretations are still to be made, first, the intercept itself 
-suggests that with no information on the policyholder, given a vehicle accident, the likely claim 
-value to the individual is $5,728. The other predictors suggest that the higher the bluebook value 
-of the policyholder’s vehicle, the larger the insurance payout will be in the event of an accident. 
-If they use the vehicle for commercial, they will also have a larger claim amount. While being a 
-female resulted as a significant variable, its coefficient is near zero and will be ignored.
+Once again, contributing coefficients are those which lie within the 95% level
+of significance. In this case, the linear model reduced the number of predictors
+that can predict the cost given an accident. The model fits extremely poor and
+does a mediocre job at explaining the variability within the training data. If
+interpretations are still to be made, first, the intercept itself suggests that
+with no information on the policyholder, given a vehicle accident, the likely
+claim value to the individual is $5,728. The other predictors suggest that the
+higher the bluebook value of the policyholder’s vehicle, the larger the
+insurance payout will be in the event of an accident. If they use the vehicle
+for commercial, they will also have a larger claim amount. While being a female
+resulted as a significant variable, its coefficient is near zero and will be
+ignored.
 
@@
 ```{r setUpCompTables}
+# Create comparison tables
 freqCompTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),

### 2020-11-08 HW4/HW4_Draft.Rmd
Unordered list slight cleanup

@@
   * `AGE`  
-  + Since the data set is comprised of auto insurance policyholder information,
+    * Since the data set is comprised of auto insurance policyholder information,
     the missing `AGE` values required imputation. However, given the relatively
@@
     conducted.
-
   * `HOME_VAL`  
-  + A `HOME_VAL` response of '0' makes it understandable to assume that those
-  missing entries could indicate that the policyholder is a renter rather than a
-  homeowner. Therefore, missing values were replaced with zeroes. The new `OWN`
-  variables created in Model Set #1 is kept in Model Set #2.
-
+    * A `HOME_VAL` response of '0' makes it understandable to assume that those
+    missing entries could indicate that the policyholder is a renter rather than
+    a homeowner. Therefore, missing values were replaced with zeroes. The new
+    `OWN` variables created in Model Set #1 is kept in Model Set #2.
   * `CAR_AGE`, `INCOME` & `YOJ`
-  + Because there is a larger number of missing values (NA’s) were identified, 
+    * Because there is a larger number of missing values (NA’s) were identified,
     imputation by predictive mean matching is conducted.
-
-  + Moreover, there is a single negative value ('-3') for `CAR_AGE` which seemed 
-  implausible and was assumed to be a typographical error. Thus, the absolute
-  value of `CAR_AGE` is taken.
+    * Moreover, there is a single negative value ('-3') for `CAR_AGE` which
+    seemed implausible and was assumed to be a typographical error. Thus, the
+    absolute value of `CAR_AGE` is taken.
 
@@
 * `CAR_AGE.f`  
-  + Because `CAR_AGE` is dominated by values of 1 (i.e. car being one year old
+  * Because `CAR_AGE` is dominated by values of 1 (i.e. car being one year old
   or less), the `CAR_AGE.f` variable is derived as a factor variable to be used
@@
   interpreted as:
-
-    `CAR_AGE.f` = 0: The vehicle is 1 year old or less.  
-    `CAR_AGE.f` = 1: The vehicle is more than a 1-year-old.
-
+    * `CAR_AGE.f` = 0: The vehicle is 1 year old or less.
+    * `CAR_AGE.f` = 1: The vehicle is more than a 1-year-old.
 * `HOMEKIDS.f` & `KIDSDRIV.f` 
-   + Both the `HOMEKIDS` and `KIDSDRIV` variables indicated that large
+   * Both the `HOMEKIDS` and `KIDSDRIV` variables indicated that large
    percentages of the values of each were 0. Therefore, both variables were
    converted to binary variables. The following is their binary interpretations:
-
-      `HOMEKIDS.f` = 0: The policyholder has no child living at home.  
-      `HOMEKIDS.f` = 1: The policyholder has 1 or more children living at home.  
-      `KIDSDRIV.f` = 0: No teenage drivers use the policyholder’s insured
+     * `HOMEKIDS.f` = 0: The policyholder has no child living at home.
+     * `HOMEKIDS.f` = 1: The policyholder has 1 or more children living at home.
+     * `KIDSDRIV.f` = 0: No teenage drivers use the policyholder’s insured
       vehicle.
-      `KIDSDRIV.f` = 1: One or more teenage drivers use the policyholder’s
+     * `KIDSDRIV.f` = 1: One or more teenage drivers use the policyholder’s
       insured vehicle.

### 2020-11-08 HW4/HW4_Draft.Rmd
Use R2 not adj R2, clean up text

@@
 nmtrnNUM <- names(DT[, .SD, .SDcols = sapply(DT, is.numeric)])
-nmtrnDUB <- names(DT[, .SD, .SDcols = sapply(DT, is.double)])
 nmtrnFAC <- setdiff(nmtrn, nmtrnNUM)
@@
 be positive) and select the model which performs best on both, Ties will be
-broken by adjusted \(R^2\).
+broken by \(R^2\).
 
@@
                            MAE = double(3),
-                           adjR2 = double(3))
+                           R2 = double(3))
 ```
@@
 tstSev <- tstX[TARGET_AMT > 0, TARGET_AMT]
-nobstst <- length(tstSev)
-np <- summary(m1SevFit$finalModel)$df[3]
 tstDumSev <- dummyVars(TARGET_AMT ~ . - INDEX - TARGET_FLAG,
@@
 sevCompTable[1, 3] <- m1SMAE <- m1Ssum[3]
-m1SAR2 <- 1 - (1 - m1Ssum[2]) * (nobstst - 1) / (nobstst - np - 1)
-sevCompTable[1, 4] <- m1SAR2
+sevCompTable[1, 4] <- m1SR2 <- m1Ssum[2]
 ```
@@
 Severity Model 1  returns an RMSE of `r round(m1SRMSE, 1)`, an MAE of
-`r round(m1SMAE, 1)`, and an adjusted \(R^2\) of `r round(m1SAR2, 4)`.
+`r round(m1SMAE, 1)`, and an \(R^2\) of `r round(m1SR2, 4)`.
 
@@
 sevCompTable[2, 3] <- m2SMAE <- MAE(m2Spred, m2.test.r)
-sevCompTable[2, 4] <- m2SAR2 <- R2(m2Spred, m2.test.r)
+sevCompTable[2, 4] <- m2SR2 <- R2(m2Spred, m2.test.r)
 ```
@@
 Severity Model 2 returns an RMSE of `r round(m2SRMSE, 1)`, an MAE of
-`r round(m2SMAE, 1)`, and an adjusted \(R^2\) of `r round(m2SAR2, 4)`.
+`r round(m2SMAE, 1)`, and an \(R^2\) of `r round(m2SR2, 4)`.
 

### 2020-11-10 HW4/HW4_Draft.Rmd
model 3 updates

@@
 library(data.table)
+library(tidyverse)
 
@@
 m2.test.p <- m2.test.X[, -c(1:3)]
+```
+
+## Model Set #3
+### Missing Data
+Similar to model 2, model set 3 deals with missing numerical data by imputing 
+them using the Multivariate imputation by chained equations (MICE) method. Multiple 
+imputation involves creating multiple predictions for each missing value, helping 
+to account for the uncertainty in the individual imputations.
+
+One thing to note about Model 3 is that it does *not* assume that an `NA` in 
+the `HOME_VAL` variable is indicative of renting. This is a key difference between
+this model and the previous 2.
+
+```{r m3clean}
+# Model 3 - data transformation
+# MICE imputation on train and test set
+m3train <- complete(mice(data = trnX %>% select(-INDEX), method = "pmm", print=FALSE),3)
+m3train$TARGET_FLAG <- factor(m3train$TARGET_FLAG, levels = c(0, 1), 
+                     labels = c("NoClaim", "Claim"))
+
+m3test <- complete(mice(data = tstX %>% select(-INDEX), method = "pmm", print=FALSE),3)
+m3test$TARGET_FLAG <- factor(m3test$TARGET_FLAG, levels = c(0, 1), 
+                     labels = c("NoClaim", "Claim"))
+
 ```
@@
 
+## Model Set #3
+### Frequency Model
+The frequency model is used to determine whether the vehicle is likely to crash.
+The base of this is a binary classifier, with the output being a probability 
+representing the likelihood of crashing.
+
+#### Baseline Model
+We will start with a simple logistic model to serve as a baseline. This includes
+all variables in the dataset.
+
+```{r model3RunBaseLog}
+# Base log model 3
+m3BaseLog <- glm(TARGET_FLAG~., family = binomial, data = m3train %>% select(-TARGET_AMT))
+kable(summary(m3BaseLog)$coefficients, digits = 3L,
+      caption = 'Model 3 - Base Logistic Regression Output')
+
+```
+
+We can immediately see that a few variables *exceed* the 0.05 p-value threshold
+for significance.
+
+#### Enhanced Frequency Model 
+Backwards stepwise regression is performed and the result is a model with the 
+following variables *removed*: `AGE`, `HOMEKIDS`, `YOJ`, `RED_CAR`, and `CAR_AGE`.
+
+```{r model3RunFreq}
+# Final log model 3
+train_control <- trainControl(method = "cv", number = 10)
+m3FreqFit <- train(TARGET_FLAG ~ KIDSDRIV+INCOME+PARENT1+HOME_VAL+MSTATUS+SEX+EDUCATION+JOB+
+              TRAVTIME+CAR_USE+BLUEBOOK+TIF+CAR_TYPE+OLDCLAIM+CLM_FREQ+REVOKED+MVR_PTS+
+              URBANICITY,
+              data = m3train %>% select(-TARGET_AMT),
+              trControl = train_control,
+              method = "glm",
+              family=binomial())
+kable(summary(m3FreqFit)$coefficients, digits = 3L,
+      caption = 'Model 3 - Final Logistic Regression Output')
+
+```
+
+#### Coefficient Discussion 
+Some interesting observations from the final results:
+
+* Living in an urban area has the greatest impact on an individual's 
+likelihood of getting into a crash. This makes sense - urban areas are more 
+crowded than rural areas, so the odds of crashing are higher. 
+* If your car is used for commercial purposes you have a higher probability
+of getting into a crash. 
+* As expected, if you've had your license revoked within the past 7 years, 
+you're more likely to get into a crash. 
+* Being a female makes you less likely to get into a crash. 
+* Having a sports car makes you more likely to get into a crash than having 
+either an SUV, Van, Pickup, or Panel Truck. 
+* Being a manager or doctor make you less likely to get into a car crash. 
+Conversely, you're more likely to crash if you're a student, blue collar worker, 
+clerk, homemaker, professional, or lawyer.
+
+### Severity Model
+The severity model is used to determine how much the crash will cost the
+insurance company. Linear regression will be used to estimate the total value
+of the cost for cars that are known to have been in an accident and accrued
+damage costs.
+
+#### Base Model
+We will start with a simple logistic regression model to serve as a baseline.
+This includes all variables in the dataset, but only for the records that have
+a TARGET_AMT > $0.
+
+```{r model3RunBaseLin}
+
+m3BaseLin <- lm(TARGET_AMT ~ ., data=m3train %>% filter(TARGET_AMT > 0) %>% select(-TARGET_FLAG))
+knitr::kable(summary(m3BaseLin)$coefficients, digits = 3L,
+             caption = 'Model 3 - Base Linear Regression Output')
+
+```
+
+A look at the final output shows us that only 1 variable is deemed
+important (`BLUEBOOK`).
+
+#### Enhanced Model
+In this case, if we were to use backwards stepwise regression, we would
+be left with a model that is determined by only 1 variable. However,
+intuitively we know that some of the variables are likely to effect
+the final payout. For this reason, we will choose to include: `BLUEBOOK`,
+`CAR_AGE`, and `CAR_TYPE`.
+
+```{r model3RunSev}
+
+m3SevFit <- lm(TARGET_AMT~ BLUEBOOK+CAR_AGE+CAR_TYPE, data = m3train %>% filter(TARGET_AMT > 0) %>% select(-TARGET_FLAG))
+
+knitr::kable(summary(m3SevFit)$coefficients, digits = 3L,
+             caption = 'Model 3 - Final Linear Regression Output')
+
+```
+
+#### Coefficient Disucssion
+The intercept of the final model tells us that the base cost of a crash is a 
+little under $5000. Some results:
+
+* The positive coefficient for the `BLUEBOOK` variable means that the higher
+the bluebook value for the vehicle, the more the payout will be. This makes
+sense. 
+* The negative coefficient for the `CAR_AGE` variable means that the older
+the vehicle is, the less the payout is. Once again, this makes sense. 
+* Non-intuitively, the positive coefficient on the Van option for the 
+`CAR_TYPE` variable means that van collisions have a higher payout. All 
+other options for this variable (including a sportscar) result in a 
+lower payout.
+
 # SELECT MODELS
@@
 ### Frequency Model
+```{r validateM3F}
+
+# Model 3 Frequency: predict on validation data
+m3Fpred <- predict(m3FreqFit, m3test %>% select(-TARGET_AMT))
+m3FpredProb <- predict(m3FreqFit, m3test %>% select(-TARGET_AMT), type = 'prob')
+
+# Model 3 Frequency: Calculate accuracy statistics
+m3FCM <- confusionMatrix(m3Fpred, m3test$TARGET_FLAG, mode = 'everything')
+freqCompTable[3, 2] <- m3FACC <- sum(diag(m3FCM$table)) / sum(m3FCM$table)
+freqCompTable[3, 3] <- m3FF1 <- m3FCM$byClass[7]
+freqCompTable[3, 4] <- m3FAUC <- auc(response = m3test$TARGET_FLAG,
+                                     predictor = m3FpredProb[, 1])
+m3FCM$table
+
+```
+Frequency Model 3 returns an accuracy of `r round(m3FACC, 4)`, an F1 score of
+`r round(m3FF1, 4)`, and an AUC of `r round(m3FAUC, 4)`.
+
 ### Severity Model
+
+```{r validateM3S}
+# Model 3 Severity: subset data
+m3testFinal <- m3test %>% 
+  select(-TARGET_FLAG) %>%
+  filter(TARGET_AMT > 0)
+
+# Model 3 Severity: predict on validation data
+m3Spred <- predict(m3SevFit, newdata = m3testFinal)
+
+# Model 3 Severity: Calculate accuracy statistics
+m3Ssum <- postResample(m3Spred, tstSev)
+sevCompTable[3, 2] <- m3SRMSE <- m3Ssum[1]
+sevCompTable[3, 3] <- m3SMAE <- m3Ssum[3]
+sevCompTable[3, 4] <- m3SR2 <- m3Ssum[2]
+```
+
+Severity Model 3 returns an RMSE of `r round(m3SRMSE, 1)`, an MAE of
+`r round(m3SMAE, 1)`, and an \(R^2\) of `r round(m3SR2, 4)`.
 
@@
 ```
+```{r m3clean}
+```
 ```{r model1DummyFreq}
@@
 ```
+```{r model3RunBaseLog}
+```
+```{r model3RunFreq}
+```
+```{r model3RunBaseLin}
+```
+```{r model3RunSev}
+```
 ```{r setUpCompTables}
@@
 ```
+```{r validateM3F}
+```
+```{r validateM3S}
+```
 ```{r freqComp}

### 2020-11-10 HW4/HW4_Draft.Rmd
tidyverse loads so many packages and masks so many calls, that I put it first to allow the other packages a chance.

@@
 # Load necessary libraries
+library(tidyverse)
 library(ggplot2)
@@
 library(data.table)
-library(tidyverse)
 

### 2020-11-10 HW4/HW4_Draft.Rmd
We don't need the entire tidyverse, just dplyr. That will speed things up. Must load dplyr after arm, otherwise select from MASS masks select from dplyr. Oh, and copyedits.

@@
 # Load necessary libraries
-library(tidyverse)
+library(arm)
+library(dplyr)
 library(ggplot2)
@@
 library(mice)
-library(arm)
 library(corrplot)
@@
 Similar to model 2, model set 3 deals with missing numerical data by imputing 
-them using the Multivariate imputation by chained equations (MICE) method. Multiple 
-imputation involves creating multiple predictions for each missing value, helping 
-to account for the uncertainty in the individual imputations.
-
-One thing to note about Model 3 is that it does *not* assume that an `NA` in 
-the `HOME_VAL` variable is indicative of renting. This is a key difference between
-this model and the previous 2.
+them using the Multivariate imputation by chained equations (MICE) method.
+Multiple imputation involves creating multiple predictions for each missing
+value, helping to account for the uncertainty in the individual imputations.
+
+One thing to note about Model 3 is that it does *not* assume that an `NA` in the
+`HOME_VAL` variable is indicative of renting. This is a key difference between
+this model and the previous two models.
 
@@
 # MICE imputation on train and test set
-m3train <- complete(mice(data = trnX %>% select(-INDEX), method = "pmm", print=FALSE),3)
+m3train <- complete(mice(data = trnX %>% select(-INDEX),
+                         method = "pmm", print = FALSE), 3)
 m3train$TARGET_FLAG <- factor(m3train$TARGET_FLAG, levels = c(0, 1), 
@@
 
-m3test <- complete(mice(data = tstX %>% select(-INDEX), method = "pmm", print=FALSE),3)
+m3test <- complete(mice(data = tstX %>% select(-INDEX),
+                        method = "pmm", print = FALSE), 3)
 m3test$TARGET_FLAG <- factor(m3test$TARGET_FLAG, levels = c(0, 1), 
                      labels = c("NoClaim", "Claim"))
-
 ```
@@
 # BUILD MODELS
-## Model Set 1
+## Model Set #1
 ### Dummy Variables  
@@
 as tree-based ones would be investigated. Most likely, MLE would be used to
-estimate the severity and a hierarchial contingent stochastic cash flow model
+estimate the severity and a hierarchical contingent stochastic cash flow model
 would be built such that given a claim, a loss would be generated from a much
@@
 # Base log model 3
-m3BaseLog <- glm(TARGET_FLAG~., family = binomial, data = m3train %>% select(-TARGET_AMT))
+m3BaseLog <- glm(TARGET_FLAG ~ ., family = binomial,
+                 data = m3train %>% select(-TARGET_AMT))
 kable(summary(m3BaseLog)$coefficients, digits = 3L,
       caption = 'Model 3 - Base Logistic Regression Output')
-
 ```
@@
 train_control <- trainControl(method = "cv", number = 10)
-m3FreqFit <- train(TARGET_FLAG ~ KIDSDRIV+INCOME+PARENT1+HOME_VAL+MSTATUS+SEX+EDUCATION+JOB+
-              TRAVTIME+CAR_USE+BLUEBOOK+TIF+CAR_TYPE+OLDCLAIM+CLM_FREQ+REVOKED+MVR_PTS+
-              URBANICITY,
+m3FreqFit <- train(TARGET_FLAG ~ KIDSDRIV + INCOME + PARENT1 + HOME_VAL +
+                     MSTATUS + SEX + EDUCATION + JOB + TRAVTIME + CAR_USE +
+                     BLUEBOOK + TIF + CAR_TYPE + OLDCLAIM + CLM_FREQ + REVOKED +
+                     MVR_PTS + URBANICITY,
               data = m3train %>% select(-TARGET_AMT),
@@
       caption = 'Model 3 - Final Logistic Regression Output')
-
 ```
@@
   crowded than rural areas, so the odds of crashing are higher.
-* If your car is used for commercial purposes you have a higher probability
-of getting into a crash. 
+  * If your car is used for commercial purposes you have a higher probability of
+  getting into a crash.
   * As expected, if you've had your license revoked within the past 7 years,
@@
   * Being a manager or doctor make you less likely to get into a car crash.
-Conversely, you're more likely to crash if you're a student, blue collar worker, 
-clerk, homemaker, professional, or lawyer.
+  Conversely, you're more likely to crash if you're a student, blue collar
+  worker, clerk, homemaker, professional, or lawyer.
 
@@
 This includes all variables in the dataset, but only for the records that have
-a TARGET_AMT > $0.
+a `TARGET_AMT` > $0.
 
@@
 
-m3BaseLin <- lm(TARGET_AMT ~ ., data=m3train %>% filter(TARGET_AMT > 0) %>% select(-TARGET_FLAG))
+m3BaseLin <- lm(TARGET_AMT ~ .,
+                data = m3train %>%
+                  filter(TARGET_AMT > 0) %>%
+                  select(-TARGET_FLAG))
 knitr::kable(summary(m3BaseLin)$coefficients, digits = 3L,
              caption = 'Model 3 - Base Linear Regression Output')
-
-```
-
-A look at the final output shows us that only 1 variable is deemed
-important (`BLUEBOOK`).
+```
+
+A look at the final output shows us that only 1 variable is deemed important
+(`BLUEBOOK`).
 
 #### Enhanced Model
-In this case, if we were to use backwards stepwise regression, we would
-be left with a model that is determined by only 1 variable. However,
-intuitively we know that some of the variables are likely to effect
-the final payout. For this reason, we will choose to include: `BLUEBOOK`,
-`CAR_AGE`, and `CAR_TYPE`.
+In this case, if we were to use backwards stepwise regression, we would be left
+with a model that is determined by only 1 variable. However, intuitively we know
+that some of the variables are likely to effect the final payout. For this
+reason, we will choose to include: `BLUEBOOK`, `CAR_AGE`, and `CAR_TYPE`.
 
@@
 
-m3SevFit <- lm(TARGET_AMT~ BLUEBOOK+CAR_AGE+CAR_TYPE, data = m3train %>% filter(TARGET_AMT > 0) %>% select(-TARGET_FLAG))
+m3SevFit <- lm(TARGET_AMT ~ BLUEBOOK + CAR_AGE + CAR_TYPE,
+               data = m3train %>%
+                 filter(TARGET_AMT > 0) %>%
+                 select(-TARGET_FLAG))
 
@@
              caption = 'Model 3 - Final Linear Regression Output')
-
 ```
@@
   sense. 
-* The negative coefficient for the `CAR_AGE` variable means that the older
-the vehicle is, the less the payout is. Once again, this makes sense. 
+  * The negative coefficient for the `CAR_AGE` variable means that the older the
+  vehicle is, the less the payout is. Once again, this makes sense. 
   * Non-intuitively, the positive coefficient on the Van option for the
-`CAR_TYPE` variable means that van collisions have a higher payout. All 
-other options for this variable (including a sportscar) result in a 
-lower payout.
+  `CAR_TYPE` variable means that van collisions have a higher payout. All other
+  options for this variable (including a sports car) result in a lower payout.
 

### 2020-11-11 HW4/HW4_Draft.Rmd
Finished this round

@@
 DT[, `:=`(INCOME = as.double(gsub('[$,]', '', INCOME)),
-          HOME_VAL = as.double(gsub('[$,]', '', HOME_VAL)),
-          BLUEBOOK = as.double(gsub('[$,]', '', BLUEBOOK)),
-          OLDCLAIM = as.double(gsub('[$,]', '', OLDCLAIM)))]
-
-# Read and process the test/evaluation file
-eval <- fread(paste0(urlRemote, pathGithub, fileTest),
-              colClasses = c(rep('integer', 2L), 'double', rep('integer', 4L),
-                             'double', 'factor', 'double', rep('factor', 4L),
-                             'integer', 'factor', 'double', 'integer',
-                             rep('factor', 2L), 'double', 'integer', 'factor',
-                             rep('integer', 2L), 'factor'))
-eval[, `:=`(INCOME = as.double(gsub('[$,]', '', INCOME)),
           HOME_VAL = as.double(gsub('[$,]', '', HOME_VAL)),
@@
 # Model Set 1: Extract those observations with values > 0
-trnSev <- trnX[TARGET_AMT > 0, TARGET_AMT]
+trnXS <- trnX[TARGET_AMT > 0,
+              ][, `:=`(INDEX = NULL, TARGET_FLAG = NULL)]
+trnYS <- trnXS[, TARGET_AMT]
+
 # Set up dummy variables
-trnDumSev <- dummyVars(TARGET_AMT ~ . - INDEX - TARGET_FLAG,
-                        fullRank = TRUE, data = trnX[TARGET_AMT > 0])
+trnXDum <- dummyVars(TARGET_AMT ~ ., fullRank = TRUE, data = trnXS)
+
 # Create matrix of dummy values
-trnXSD <- predict(trnDumSev, newdata = trnX[TARGET_AMT > 0])
-# preProcess with near-zero value and kNN imputation
-trnXSDpp <- preProcess(trnXSD, method = c('nzv', 'knnImpute'),
-                       k = 3, knnSummary = median)
+trnXSD <- predict(trnXDum, newdata = trnXS)
+
+# preProcess with near-zero value and kNN imputation using median
+trnXDumPP <- preProcess(trnXSD, method = c('nzv', 'knnImpute'), k = 3,
+                      knnSummary = median)
+
 # Create processed predictors
-trnXSDI <- predict(trnXSDpp, newdata = trnXSD)
+trnXSDImp <- predict(trnXDumPP, newdata = trnXSD)
 ```
@@
 set.seed(878)
-m1SevFit <- train(x = trnXSDI, y = trnSev, trControl = trC,
-                  method = 'glmStepAIC', family = gaussian, direction = 'both',
-                  trace = 0)
+m1SevFit <- train(x = trnXSDImp, y = trnYS, trControl = trC,
+                  method = 'glmStepAIC', family = gaussian(),
+                  direction = 'both', trace = 0)
 kable(summary(m1SevFit)$coefficients, digits = 3L,
@@
 # Model 1 Severity: process validation data similar to training data
-tstSev <- tstX[TARGET_AMT > 0, TARGET_AMT]
-tstDumSev <- dummyVars(TARGET_AMT ~ . - INDEX - TARGET_FLAG,
-                       data = tstX[TARGET_AMT > 0])
-tstXSD <- predict(tstDumSev, newdata = tstX[TARGET_AMT > 0])
-tstXSDpp <- preProcess(tstXSD, method = c('nzv', 'knnImpute'),
-                       k = 3, knnSummary = median)
-tstXSDI <- predict(tstXSDpp, newdata = tstXSD)
+tstXS <- tstX[TARGET_AMT > 0
+              ][, `:=`(INDEX = NULL, TARGET_FLAG = NULL)]
+tstYS <- tstXS[, TARGET_AMT]
+tstXDum <- dummyVars(TARGET_AMT ~ ., data = tstXS, fullRank = TRUE)
+tstXSD <- predict(tstXDum, newdata = tstXS)
+tstXDumPP <- preProcess(tstXSD, method = c('nzv', 'knnImpute'), k = 3,
+                      knnSummary = median)
+tstXSDImp <- predict(tstXDumPP, newdata = tstXSD)
 
 # Model 1 Severity: predict on validation data
-m1Spred <- predict(m1SevFit, newdata = tstXSDI)
+m1Spred <- predict(m1SevFit, newdata = tstXSDImp)
 # Model 1 Severity: Calculate accuracy statistics
-m1Ssum <- postResample(m1Spred, tstSev)
+m1Ssum <- postResample(m1Spred, tstYS)
 sevCompTable[1, 2] <- m1SRMSE <- m1Ssum[1]
@@
 # Accuracy Statistics
-m2Ssum <- postResample(m2Spred, tstSev)
+m2Ssum <- postResample(m2Spred, tstYS)
 sevCompTable[2, 2] <- m2SRMSE <- RMSE(m2Spred, m2.test.r)
@@
 # Model 3 Severity: Calculate accuracy statistics
-m3Ssum <- postResample(m3Spred, tstSev)
+m3Ssum <- postResample(m3Spred, tstYS)
 sevCompTable[3, 2] <- m3SRMSE <- m3Ssum[1]
@@
 
+**Frequency Model 3**'s performance on the holdout set was the best in two of
+the three metrics---accuracy and F1---and will be used to predict the frequency
+of a claim on the evaluation set.
+
 ### Severity Model
@@
 
+**Severity Model 1**'s performance on the holdout set was the best in two of the
+three metrics---RMSE and \(R^2\)---and will be used to predict the severity for
+claims in the evaluation set for which Model 3 predicts a claim.
+
+However, the performance of this model, or any of the three, is not expected to
+be good, as a linear regression with a normal assumption is most often a poor
+fit to severity data. Compare the distributions of both training and testing
+results for Model 1, ostensibly the best model, on both a real and a log-scale.
+
+```{r badSev, fig.height=4}
+m1trnPred <- predict(m1SevFit, trnXSDImp)
+densDT <- data.table(trainObs = trnYS,
+                     trainPred = m1trnPred,
+                     testObs = tstYS,
+                     testPred = m1Spred)
+densDT <- melt(densDT, variable.factor = FALSE)
+densDT[, `:=`(data = ifelse(substr(variable, 1, 5) == 'train', 'train', 'test'),
+              len =  lapply(variable, nchar))
+              ][, set := ifelse(substr(variable, len, len) == 's',
+                                    'observation', 'prediction')
+                ][, `:=`(variable = NULL, len = NULL)]
+ggplot(densDT, aes(x = value, col = data,  linetype = set)) +
+  geom_density(size = 1) + scale_x_continuous(labels = dollar) +
+  ggtitle("Density on a Normal Scale")
+ggplot(densDT, aes(x = value, col = data, linetype = set)) +
+  geom_density(size = 1) + scale_x_log10(labels = dollar) +
+  ggtitle("Density on a Log Scale")
+```
+
+It is clear that the normal distribution assumption underlying the linear model
+is too restrictive. The variance is too low and the tails are too thin. The
+log-density plot implies a thicker tailed distribution such as a lognormal or a
+Pareto. Nevertheless, we will use Model 3 to predict the losses, understanding
+that it is the best option under the artificial constraints of the problem.[^1]
+
+[^1]: Apologies; one of us is an actuary and this is painful 8-) .
+
 # PREDICTIONS
+For our group, different model sets had the best performance for frequency and
+severity. These models processed the data differently. Therefore, we cannot
+simply use the output from Model 3's frequency predictions as the inputs to
+Model 1's severity.
+
+What we will do is use Model 3 to determine *which* evaluation observations are
+predicted to have claims and then use those observations to predict severity. In
+other words, the evaluation data will be processed in accordance with Frequency
+Model 3 and the incidents of a claim will be recorded. For those evaluation
+observations which return a claim, their raw data will be processed according to
+Severity Model 1 to return predicted values.
+
+```{r evalProcessClean}
+# Read and process the test/evaluation file
+# Can re-use the name DT for convenience now
+DT <- fread(paste0(urlRemote, pathGithub, fileTest),
+              colClasses = c(rep('integer', 2L), 'double', rep('integer', 4L),
+                             'double', 'factor', 'double', rep('factor', 4L),
+                             'integer', 'factor', 'double', 'integer',
+                             rep('factor', 2L), 'double', 'integer', 'factor',
+                             rep('integer', 2L), 'factor'))
+neval <- dim(DT)[[1]]
+DT[, `:=`(INCOME = as.double(gsub('[$,]', '', INCOME)),
+          HOME_VAL = as.double(gsub('[$,]', '', HOME_VAL)),
+          BLUEBOOK = as.double(gsub('[$,]', '', BLUEBOOK)),
+          OLDCLAIM = as.double(gsub('[$,]', '', OLDCLAIM)))]
+setattr(DT$MSTATUS, 'levels', c('Yes', 'No'))
+setattr(DT$SEX, 'levels', c('M', 'F'))
+setattr(DT$EDUCATION, 'levels', c('<HighSchool', 'Bachelors', 'Masters', 'PhD',
+                                  'HighSchool'))
+setattr(DT$JOB, 'levels', c('Unknown', 'Clerical', 'Doctor', 'HomeMaker',
+                            'Lawyer', 'Manager', 'Professional', 'Student',
+                            'BlueCollar'))
+setattr(DT$CAR_TYPE, 'levels', c('Minivan', 'PanelTruck', 'Pickup',
+                                 'SportsCar', 'Van', 'SUV'))
+
+setattr(DT$URBANICITY, 'levels', c('Urban', 'Rural'))
+
+# Relevel some of the variables to have the default make more sense
+# Set default marital status to unmarried
+set(DT, NULL, 'MSTATUS', relevel(DT$MSTATUS, 'No'))
+# Set default car usage to private
+set(DT, NULL, 'CAR_USE', relevel(DT$CAR_USE, 'Private'))
+# Set default are to rural, making urban a positive? risk factor
+set(DT, NULL, 'URBANICITY', relevel(DT$URBANICITY, 'Rural'))
+```
+
+As there are `r neval` observations, we will not printing them in this document.
+Rather, we will show an estimated frequency rate. The actual observations for
+which claims are predicted may be obtained from the appropriate data object in
+the code, as found in the CODE APPENDIX.
+
+## Evaluation Data Processing and Cleaning
+In this section, the evaluation set will be processed and cleaned as the per the
+training set. The model-specific imputations, variable creations, and
+predictions will be done in their own sections.
+
+## Frequency
+Missing values will be imputed, with `INDEX` and the two prediction targets
+ignored, of course. Once done, the claim incidents will be predicted.
+```{r evalPredictFreq}
+# Impute the missing values but not INDEX or the values we will be predicting!
+set.seed(94)
+m3eval <- complete(mice(data = DT %>%
+                          select(-c(INDEX, TARGET_FLAG, TARGET_AMT)),
+                        method = "pmm", print = FALSE), 3)
+evalFreqFit <- predict(m3FreqFit, m3eval)
+numPred <- sum(evalFreqFit == "Claim")
+claimRate <- numPred / neval
+claimIDX <- which(evalFreqFit == "Claim")
+```
+
+The model predicts `r numPred` claims out of `r neval` observations for a claim
+rate of `r prettyNum(claimRate * 100, digits = 3L)`%. The actual outcomes are
+stored in the `evalFreqFit` object in the code.
+
+## Severity
+The individual `r numPred` observations for which Model 3 predicted a claim will
+be extracted from the state of the evaluation set prior to Model 3-specific
+adjustments, processed per Severity Model 1, and then the claim values will be
+predicted.
+
+Once again, with so many observations, we will not simply print all `r numPred`.
+Rather, we will provide some summary statistics and graphs of the distribution
+of the claims. The actual claim values can be extracted from the appropriate
+data structure in the code, of course.
+
+```{r evalPredictSev, fig.height=4}
+# Extract the "claim" observations from the data
+evalObs <- DT[claimIDX, ]
+
+# At this point the unneeded columns of INDEX and the 2 targets will be removed
+evalObs[, `:=`(INDEX = NULL, TARGET_FLAG = NULL, TARGET_AMT = NULL)]
+
+# Create the OWN variable
+evalObs[, OWN := factor(ifelse(HOME_VAL > 0, 1, 0), levels = c(0, 1),
+                        labels = c('No', 'Yes'))]
+
+# Set up dummy variables
+evalDumSev <- dummyVars( ~ ., fullRank = TRUE, data = evalObs)
+# Create matrix of dummy values
+evalDumSevMatrix <- predict(evalDumSev, newdata = evalObs)
+# preProcess with near-zero value and kNN imputation
+evalDumSevPP <- preProcess(evalDumSevMatrix, method = c('nzv', 'knnImpute'),
+                           k = 3, knnSummary = median)
+# Create matrix of processed and imputed predictors
+evalDumSevPPI <- predict(evalDumSevPP, newdata = evalDumSevMatrix)
+evalSevFit <- predict(m1SevFit, evalDumSevPPI)
+
+# Summary Statistics
+tabularComp <- data.table(Set = c("Training", "Testing", "PredictedEval"),
+                          Min = c(min(trnYS), min(tstYS), min(evalSevFit)),
+                          Median = c(median(trnYS), median(tstYS),
+                                     median(evalSevFit)),
+                          Max = c(max(trnYS), max(tstYS), max(evalSevFit)),
+                          Mean = c(mean(trnYS), mean(tstYS), mean(evalSevFit)),
+                          SD = c(sd(trnYS), sd(tstYS), sd(evalSevFit)))
+tabularComp[, CV := SD / Mean, by = Set]
+
+ggplot(data.frame(x = evalSevFit), aes(x = x)) + geom_density() +
+  scale_x_continuous(labels = dollar)
+
+knitr::kable(tabularComp, digits = 3L,
+             caption = "Severity Statistics Comparison")
+```
+
+The actual outcomes are stored in the `evalSevFit` object in the code. The table
+and the graph show how the predictions are ***much*** more constrained than
+either the training or testing sets. The means are relatively close, which is
+to be expected when trying to shoehorn non-normal data into a Gaussian
+framework.
 
@@
 ```
+```{r badSev}
+```
+```{r evalProcessClean}
+```
+```{r evalPredictFreq}
+```
+```{r evalPredictSev}
+```

### 2020-11-19 HW4/HW4_Group2.Rmd
Rename file for submission
Make code order cleaner

--- 2020-11-11 HW4/HW4_Draft.Rmd
@@
 
-# Summary Statistics
+# Kernel-smoothed density
+ggplot(data.frame(x = evalSevFit), aes(x = x)) + geom_density() +
+  scale_x_continuous(labels = dollar)
+
+# Summary statistics comparison
 tabularComp <- data.table(Set = c("Training", "Testing", "PredictedEval"),
@@
 tabularComp[, CV := SD / Mean, by = Set]
-
-ggplot(data.frame(x = evalSevFit), aes(x = x)) + geom_density() +
-  scale_x_continuous(labels = dollar)
-
 knitr::kable(tabularComp, digits = 3L,

### 2020-11-19 Final_project/Draft_EDA_start.Rmd

---
title: "Final Project"
author: "John Kellogg"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(data.table)
library(corrplot)
library(vip)
library(psych)
library(summarytools)
library(readr)
library(knitr)
library(GGally)
```

## Research question

If we know the number of likes, can we predict the number of views a video will receive?

sub questions:
For each x increase/decrease in views, dislikes, and comments, how much does the view count move?
Does turning off comments and ratings affect and by how much
Is that number affected by country?
Is this the same across categories?

## Data Exploration

```{r cars}
CA_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/CAvideos.csv"
CA_set <- read_csv(file = CA_path)
US_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/USvideos.csv"
US_set <- read_csv(file = US_path)
GB_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/GBvideos.csv"
GB_set <- read_csv(file = GB_path)
FR_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/FRvideos.csv"
FR_set <- read_csv(file = FR_path)
```

Canada data

```{r summary_ca}
summaryca <- CA_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
describe(summaryca)

ggpairs(
  summaryca,
  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
  )

ca_corr <- cor(summaryca)
corrplot(ca_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")

```

```{r summary_us}
summaryus <- US_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
describe(summaryus)

ggpairs(
  summaryus,
  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
  )
us_corr <- cor(summaryus)
corrplot(us_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")

```

```{r summary_us}
summarygb <- GB_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
describe(summarygb)

ggpairs(
  summarygb,
  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
  )

gb_corr <- cor(summarygb)
corrplot(gb_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
```

```{r summary_fr}
summaryfr <- FR_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
describe(summaryfr)

ggpairs(
  summaryfr,
  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
  )

fr_corr <- cor(summaryfr)
corrplot(fr_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
```

There is definite heteroskedasticity across all data points, we will need to address.

### 2020-11-20 HW5/HW5-Draft.Rmd
Start HW5. Finish assignement description, start EDA

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #5'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "11/22/2020"
output:
  pdf_document:
  toc: TRUE
toc_depth: 4
urlcolor: purple
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
# Load necessary libraries
library(arm)
library(dplyr)
library(ggplot2)
library(scales)
library(knitr)
library(caret)
library(RANN)
library(pROC)
library(mice)
library(corrplot)
library(data.table)

# Set master seed
set.seed(65408)

# Set filepaths for data ingestion
urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW5/data/"
fileTrain = "wine-training-data.csv"
fileTest = "wine-evaluation-data.csv"

# Read training file
DT <- fread(paste0(urlRemote, pathGithub, fileTrain))

# Number of training observations
ntrobs <- dim(DT)[[1]]

# Get the names of the variables and which ones are numeric
nmtrn <- names(DT)

# Drop INDEX and target variables from name list
nmtrn <- nmtrn[-(1:2)]
```

# ASSIGNMENT
The assignment for HW5 is to analyze and model a dataset containing
approximately 12,000 records representing commercially available wines. The
`TARGET` response variable represents the number of sample cases of wine
purchased by wine distribution companies after sampling that wine. These cases
would be used to provide tasting samples to restaurants and wine stores around
the United States. The more sample cases purchased, the more likely is a wine to
be sold at a high end restaurant.A large wine manufacturer is studying the data
in order to predict the number of wine cases ordered based upon the wine
characteristics. If the wine manufacturer can predict the number of cases, then
that manufacturer will be able to adjust their wine offering to maximize sales.

The objective of this assignment is to build a count regression model to predict
the number of cases of wine that will be sold given certain properties of the
wine. *HINT:* Sometimes, the fact that a variable is missing is actually
predictive of the target. You can only use the variables given to you (or
variables that you derive from the variables provided).

# DATA EXPLORATION
## Variables
The data is composed of the following variables:

|VARIABLE NAME|DEFINITION|THEORETICAL EFFECT|
|--|--|--|
|INDEX|Identification Variable (do not use)|None|
|TARGET|Number of Cases Purchased|None|
|AcidIndex|Proprietary method of testing total acidity of wine by using a weighted average||
|Alcohol|Alcohol Content||
|Chlorides|Chloride content of wine||
|CitricAcid|Citric Acid Content||
|Density|Density of Wine||
|FixedAcidity|Fixed Acidity of Wine||
|FreeSulfurDioxide|Sulfur Dioxide content of wine||
|LabelAppeal|Marketing Score indicating the appeal of label design for consumers. High numbers suggest customers like the label design. Negative numbers suggest customers don't like the design.|Many consumers purchase based on the visual appeal of the wine label design. Higher numbers suggest better sales.|
|ResidualSugar|Residual Sugar of wine||
|STARS|Wine rating by a team of experts. 4 Stars = Excellent, 1 Star = Poor|A high number of stars suggests high sales|
|Sulphates|Sulfate content of wine||
|TotalSulfurDioxide|Total Sulfur Dioxide of Wine||
|VolatileAcidity|Volatile Acid content of wine||
|pH|pH of wine||

There are `r ntrobs` observations. All of these predictors are numeric, although
some `LabelAppeal` and `STARS` could be considered ordered factors. There are a
lot of missing values for some of the predictors.

```{r missing}
Missing <- t(DT[, lapply(.SD, function(x) {sum(is.na(x))}),
                .SDcols = -c('INDEX', 'TARGET')])
kable(Missing, col.names = 'Missing',
      caption = 'Predictors with Missing Observations')
```

### 2020-11-20 HW5/HW5-Draft.Rmd
Missing data. Clean up loads until needed

@@
 # Load necessary libraries
-library(arm)
-library(dplyr)
 library(ggplot2)
@@
 library(caret)
-library(RANN)
-library(pROC)
-library(mice)
 library(corrplot)
@@
 # Number of training observations
-ntrobs <- dim(DT)[[1]]
+ntrnobs <- dim(DT)[[1]]
 
@@
 
-There are `r ntrobs` observations. All of these predictors are numeric, although
-some `LabelAppeal` and `STARS` could be considered ordered factors. There are a
-lot of missing values for some of the predictors.
+There are `r ntrnobs` observations. All of these predictors are numeric,
+although some `LabelAppeal` and `STARS` could be considered ordered factors. For
+the purposes of this assignment, it makes sense to treat every predictor as
+numeric.
+
+## Missing Data
+There are a lot of missing values for some of the predictors.
 
 ```{r missing}
-Missing <- t(DT[, lapply(.SD, function(x) {sum(is.na(x))}),
-                .SDcols = -c('INDEX', 'TARGET')])
-kable(Missing, col.names = 'Missing',
-      caption = 'Predictors with Missing Observations')
+missingPreds <- transpose(DT[, lapply(.SD, function(x) {sum(is.na(x))}),
+                             .SDcols = -c('INDEX', 'TARGET')],
+                          keep.names = 'Predictors')
+setnames(missingPreds, 'V1', 'Missing')
+missingPreds[, Percentage := Missing / ntrnobs * 100]
+setorder(missingPreds, -Missing)
+kable(missingPreds, digits = 2L, caption = 'Predictors with Missing Observations')
 ```
+
+`STARS` is especially sparse, but as noted in the assignment, this may be an
+indication in and of its own. It is reasonable to assume that the vinters of a
+good wine want it to be rated. If they do not submit it for rating, that may be
+an indication of their lack of faith in the wine. How to factor this into the
+analysis will be decided individually by the modelers in this assignment. The
+other missing variables are all less than 10%, and their handling via imputation
+or otherwise will be done on a model-by-model basis as well.

### 2020-11-20 HW5/HW5-Draft.Rmd
Finish basic EDA

@@
 
-# Get the names of the variables and which ones are numeric
-nmtrn <- names(DT)
-
-# Drop INDEX and target variables from name list
-nmtrn <- nmtrn[-(1:2)]
+# Get the names of the predictor variables
+nmtrn <- names(DT)[-(1:2)]
+nmtrnINT <- c('AcidIndex', 'LabelAppeal', 'STARS')
+nmtrnDUB <- setdiff(nmtrn, nmtrnINT)
 ```
@@
 There are `r ntrnobs` observations. All of these predictors are numeric, 
-although some `LabelAppeal` and `STARS` could be considered ordered factors. For
-the purposes of this assignment, it makes sense to treat every predictor as
-numeric.
+although `LabelAppeal`, `AcidIndex` and `STARS` all appear to be ordinal factors
+and not true numerics. For the purpose of this assignment, we can treat them
+as integers.
 
@@
 
-```{r missing}
+```{r missingVal}
 missingPreds <- transpose(DT[, lapply(.SD, function(x) {sum(is.na(x))}),
-                             .SDcols = -c('INDEX', 'TARGET')],
+                             .SDcols = nmtrn],
                           keep.names = 'Predictors')
@@
 or otherwise will be done on a model-by-model basis as well.
+
+## Summary Statistics and Graphs
+As all of the data are numeric, we can investigate the distributions of the
+predictors both tabularly and graphically.
+
+The numeric predictor variables have the following summary statistics, ignoring
+missing values:
+
+```{r statsN}
+# Isolate numeric only predictors
+predictorDT <- DT[, .SD, .SDcols = nmtrn]
+
+# Melt them from wide to long format
+predictorDTM <- melt(predictorDT, variable.name = 'metric',
+                     value.name = 'value',  variable.factor = FALSE,
+                     id.vars = NULL)
+
+# Calculate summary statistics
+statsN <- predictorDTM[, .(Mean = mean(value, na.rm = TRUE),
+                           SD = sd(value, na.rm = TRUE),
+                           Min = min(value, na.rm = TRUE),
+                           Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
+                           Median = median(value, na.rm = TRUE),
+                           Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
+                           Max = max(value, na.rm = TRUE),
+                           IQR = IQR(value, na.rm = TRUE)), keyby = 'metric']
+
+# Print the table
+kable(statsN, digits = 3L, align = 'r',
+      caption = "Summary Statistitics for Numeric Variables")
+```
+
+A kernel-smoothed density plot of the distributions of the non-integer numeric
+predictors is below, followed by a histogram of the integral-valued predictors.
+
+```{r graphsD, fig.height=5, fig.width=7}
+# Using epanechnikov kernel
+ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnDUB], aes(x = value)) +
+  geom_density(kernel = 'epanechnikov') +
+  facet_wrap( ~ metric, scales = 'free') +
+  ggtitle('Kernel-smoothed Density of Numeric Predictors')
+```
+
+The numeric predictors all look to be basically symmetrical, but non-Gaussian in
+that there is a strong spike near the median, but the tails on either side are
+thicker than would be for a normal distribution with that low of a standard
+deviation.
+
+```{r graphsH, fig.height=2, fig.width=7}
+# Freedman-Diaconis rule for bin widths
+FDbin <- function(x) {
+  result <- 2 * IQR(x, na.rm = TRUE) / (length(x) ^ (1 / 3))
+  return(ifelse(result == 0, 0.5, result))
+}
+ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnINT], aes(x = value)) +
+  geom_histogram(binwidth = FDbin, fill = 'indianred4') +
+  facet_wrap( ~ metric, scales = 'free') +
+  ggtitle('Histogram of Integral Predictors')
+```
+
+Of these three, `LabelAppeal` seems to be the most "normal" of the lot; `STARS`
+and `AcidIndex` look to be more Poisson in shape.
+
+### Correlations
+The corrgram below graphically represents the correlations between the numeric
+predictor variables, when ignoring the missing variables.
+
+```{r corrgram, fig.width=6, fig.height=4}
+corrplot::corrplot(cor(DT[, ..nmtrn], use = 'complete.obs'),
+         method = 'ellipse', type = 'lower', order = 'hclust',
+         hclust.method = 'ward.D2')
+```
+
+There is very little correlation between the variables. The only pairs with some
+level correlation are:
+
+  * `STARS` being positively correlated with `LabelAppeal`
+    * This is interesting. Could it be that wine connoisseurs are impacted by
+    the visual appearance of the label and not just the flavor?
+  * `AcidIndex` having some positive correlation with `FixedAcidity`
+    * One may have suspected a higher correlation, to be frank.
+  * `AcidIndex` having some *negative* correlation with `STARS`
+    * Wine reviewers may not like too much acidity, it seems.
+
+# CODE APPENDIX
+```{r setup2, include=FALSE}
+knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
+```
+```{r loadData}
+```
+```{r missingVal}
+```
+```{r statsN}
+```
+```{r graphsD}
+```
+```{r graphsH}
+```
+```{r corrgram}
+```
+```{r dataClean}
+```
+```{r trainTestSplit}
+```
+```{r model1addVars}
+```
+```{r m2clean}
+```
+```{r m3clean}
+```
+```{r model1DummyFreq}
+```
+```{r model1RunFreq}
+```
+```{r model1DummySev}
+```
+```{r model1RunSev}
+```
+```{r model2RunFreq}
+```
+```{r model2RunSev}
+```
+```{r model3RunBaseLog}
+```
+```{r model3RunFreq}
+```
+```{r model3RunBaseLin}
+```
+```{r model3RunSev}
+```
+```{r setUpCompTables}
+```
+```{r validateM1F}
+```
+```{r validateM1S}
+```
+```{r validateM2F}
+```
+```{r validateM2S}
+```
+```{r validateM3F}
+```
+```{r validateM3S}
+```
+```{r freqComp}
+```
+```{r sevComp}
+```
+```{r badSev}
+```
+```{r evalProcessClean}
+```
+```{r evalPredictFreq}
+```
+```{r evalPredictSev}
+```

### 2020-11-20 HW5/HW5-Draft.Rmd
Start data prep. Suggest model selection criteria

@@
 
+# DATA PREPARATION
+The different methods of imputation and feature generation will be discussed in
+each models BUILD section. The only preparation performed here will be that
+which will apply to all models equally.
+
+# SELECT MODELS
+## Model Selection Criteria
+We will look at RMSE and AIC as our metrics and select the model which performs
+best on both, Ties will be broken by \(R^2\).
+
 # CODE APPENDIX

### 2020-11-20 HW5/HW5-Draft.Rmd
Finish train/test split. Now going to build my set of models.

--- 2020-11-20 HW5/HW5-Draft.Rmd
@@
 thicker than would be for a normal distribution with that low of a standard
-deviation.
+deviation. This is also why a boxplot would be a poor graphical indicator, as
+the spike at the medians would lead to a compressed inter-quartile range and a
+plethora of outliers.
 
@@
 
+# DATA PREPARATION
+First, preparation necessary for all the models equally will be performed. The
+subsequent imputation and feature generation, if necessary will be discussed in
+a separate subsection for each modeler.
+
+## Training & Testing Split
+All the models will be trained on the same approximately 70% of the training
+set, reserving 30% for validation of which model to select for the count
+estimation on the supplied evaluation set.
+
+```{r trainTestSplit}
+# Create training and testing split
+set.seed(1004)
+trnIDX <- createDataPartition(DT$TARGET, p = 0.7)
+trnX <- DT[trnIDX$Resample1, ]
+tstX <- DT[!trnIDX$Resample1, ]
+```
+
+# Modeler # 1
+## Missing Data
+The missing daya
+
+# SELECT MODELS
+## Model Selection Criteria
+We will look at RMSE and AIC as our metrics and select the model which performs
+best on both, Ties will be broken by \(R^2\).
+
 # CODE APPENDIX
@@
 ```{r corrgram}
-```
-```{r dataClean}
 ```

### 2020-11-20 HW5/HW5-Draft.Rmd
Built Poisson model. Need to explain it.

@@
 ```{r graphsD, fig.height=5, fig.width=7}
-# Using epanechnikov kernel
+# Using epanechnikov kernel to generate kernel-smoothed densities
 ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnDUB], aes(x = value)) +
@@
 }
+# Generate histogram
 ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnINT], aes(x = value)) +
@@
 
-### Correlations
+## Correlations
 The corrgram below graphically represents the correlations between the numeric
@@
 
-```{r corrgram, fig.width=6, fig.height=4}
+```{r corrgram, fig.width=5, fig.height=6.5}
+# Create corrgram
 corrplot::corrplot(cor(DT[, ..nmtrn], use = 'complete.obs'),
@@
 
-# Modeler # 1
-## Missing Data
-The missing daya
+## Modeler # 1
+### Missing Data
+The missing data can be split into two groups: `STARS` and the rest. For all the
+others, I do not think there is signal encoded in the "missingness" and thus
+will use some form of imputation. `STARS` is different as discussed in the DATA
+EXPLORATION section. Therefore, I will create a new feature called `Rated` which
+will be true if `STARS` is missing, and then only use this variable and its
+interaction with `STARS` in the model, and not `STARS` by itself.
+
+```{r model1addVars}
+# Copy the training set to leave a pristine version for others
+m1trnX <- copy(trnX)
+
+# Add the "rated" factor variable
+m1trnX[, Rated := factor(ifelse(is.na(STARS), 'Unrated', 'Rated'),
+                         levels = c('Unrated', 'Rated'),
+                         labels = c('Unrated', 'Rated'))]
+```
+
+### Imputation
+Imputation will be handled through bagging. Instead of looking at a nearest
+neighbors approach, which also requires centering and scaling, once can create a
+set of bagged trees. As per Kuhn (2019):
+
+>*For each predictor in the data, a bagged tree is created using all of the*
+*other predictors in the training set. When a new sample has a missing*
+*predictor value, the bagged model is used to predict the value. While, in*
+*theory, this is a more powerful method of imputing, the computational costs*
+*are much higher than the nearest neighbor technique.*
+
+The bagged trees will use `STARS` as a predictor, as that contains valuable
+information, but prior to the imputation, the missing `STARS` values will be
+replaced by 0 so as not to be imputed, as per the explanation above. The
+preprocessing will also check for near-zero value predictors.
+
+```{r model1impute}
+# Create bagged-tree imputation model
+m1trnXi <- preProcess(m1trnX, method =c('nzv', 'bagImpute'))
+
+# Replace missing STARS with 0s
+m1trnX[is.na(STARS), STARS := 0L]
+
+# Impute the remaining NAs
+m1trnXimp <- predict(m1trnXi, newdata = m1trnX)
+```
+
+It will be instructive to compare the shapes of the distributions after
+imputation to those before imputation.
+
+```{r m1postImputeD, fig.height=5, fig.width=7}
+# Isolate numeric only predictors
+m1predictorDT <- m1trnXimp[, .SD, .SDcols = nmtrn]
+
+# Melt them from wide to long format
+m1predictorDTM <- melt(m1predictorDT, variable.name = 'metric',
+                     value.name = 'value',  variable.factor = FALSE,
+                     id.vars = NULL)
+
+# Using epanechnikov kernel to generate kernel-smoothed densities
+ggplot(predictorDTM[metric %chin% nmtrnDUB], aes(x = value)) +
+  geom_density(kernel = 'epanechnikov') +
+  facet_wrap( ~ metric, scales = 'free') +
+  ggtitle('Kernel-smoothed Density of Numeric Predictors')
+```
+
+Thankfully, none of the shapes appears to have changed significantly, which
+implies the imputation was in line with the original distributions. Now for the
+histograms.
+
+```{r m1postImputeH, fig.height=2, fig.width=7}
+# Generate histogram
+ggplot(m1predictorDTM[metric %chin% nmtrnINT], aes(x = value)) +
+  geom_histogram(binwidth = FDbin, fill = 'indianred4') +
+  facet_wrap( ~ metric, scales = 'free') +
+  ggtitle('Histogram of Integral Predictors')
+```
+
+Here too, `AcidIndex` and `LabelAppeal` look very similar to their
+pre-imputation distribution. What stands out is `STARS`. Both the Poisson and
+the negative binomial models are either unimodal or have to consecutive values
+sharing the mode, due to their inherent nature as discrete disctibutions. What
+is seen here is that 1 star is empirically less common than 0 (unrated) or 2
+stars. On the one hand, this could simply be noise (process variance) or
+parameter error (finite sample size). On the other, it could indicate the need
+for a zero-inflated version of the counting distributions.
+
+# BUILD MODELS
+## Modeler # 1
+Each of the models will follow a similar format, with only the family changing.
+Each model will start with the saturated model including all the predictors,
+***except*** for `STARS` which will be replaced with the interaction between
+`Rated` and `STARS`. A forward and backwards stepwise procedure based on AIC
+will be used, and the model with the lowest AIC on the training set will be the
+selected to be tested against the testing set. Using stepAIC precludes the use
+of k-fold cross validation.
+
+```{r m1TrainControl}
+# All Modeler 1 models will use stepAIC so turn off cross-validation
+trc <- trainControl(method = 'none')
+```
+
+### Poisson Model
+
+```{r m1Poisson}
+set.seed(350047)
+m1PFit <- train(TARGET ~ . - INDEX - STARS + Rated:STARS, data = m1trnXimp,
+                trControl = trc, method = 'glmStepAIC',
+                family = poisson(link = 'log'), direction = 'both', trace = 0)
+# Print results
+kable(summary(m1PFit$finalModel)$coefficients, digits = 3L,
+      caption = "Model 1 Poisson Regression Output")
+```
+
+```
+                     Df Deviance   AIC
+<none>                    9654.5 31998
+- Sulphates           1   9657.1 31998
++ Density             1   9654.1 31999
++ ResidualSugar       1   9654.2 31999
+- Alcohol             1   9658.3 32000
++ CitricAcid          1   9654.4 32000
++ FixedAcidity        1   9654.4 32000
+- pH                  1   9659.0 32000
+- Chlorides           1   9659.5 32001
+- FreeSulfurDioxide   1   9659.7 32001
+- TotalSulfurDioxide  1   9670.3 32012
+- VolatileAcidity     1   9670.9 32012
+- AcidIndex           1   9880.9 32222
+- LabelAppeal         1  10102.7 32444
+- `STARS:RatedRated`  1  10335.1 32676
+- RatedRated          1  10337.5 32679
+```
 
@@
 best on both, Ties will be broken by \(R^2\).
+
+# References
+
+  * Kuhn, M. (2019, March 27). *The `caret` Package*.
+  https://topepo.github.io/caret/pre-processing.html#imputation
 
@@
 ```
-```{r m2clean}
+```{r model1impute}
 ```
-```{r m3clean}
+```{r m1postImputeD}
 ```
-```{r model1DummyFreq}
+```{r m1postImputeH}
 ```
-```{r model1RunFreq}
+```{r m1TrainControl}
 ```
-```{r model1DummySev}
-```
-```{r model1RunSev}
-```
-```{r model2RunFreq}
-```
-```{r model2RunSev}
-```
-```{r model3RunBaseLog}
-```
-```{r model3RunFreq}
-```
-```{r model3RunBaseLin}
-```
-```{r model3RunSev}
-```
-```{r setUpCompTables}
-```
-```{r validateM1F}
-```
-```{r validateM1S}
-```
-```{r validateM2F}
-```
-```{r validateM2S}
-```
-```{r validateM3F}
-```
-```{r validateM3S}
-```
-```{r freqComp}
-```
-```{r sevComp}
-```
-```{r badSev}
-```
-```{r evalProcessClean}
-```
-```{r evalPredictFreq}
-```
-```{r evalPredictSev}
-```

### 2020-11-20 HW5/HW5-Draft.Rmd
Finish Poisson Section. Sketch out remaining skeleton

@@
                      value.name = 'value',  variable.factor = FALSE,
-                     id.vars = NULL)
+                     measure.vars = nmtrn)
 
@@
 # Create bagged-tree imputation model
+set.seed(89)
 m1trnXi <- preProcess(m1trnX, method =c('nzv', 'bagImpute'))
@@
 # BUILD MODELS
-## Modeler # 1
-Each of the models will follow a similar format, with only the family changing.
-Each model will start with the saturated model including all the predictors,
-***except*** for `STARS` which will be replaced with the interaction between
-`Rated` and `STARS`. A forward and backwards stepwise procedure based on AIC
-will be used, and the model with the lowest AIC on the training set will be the
-selected to be tested against the testing set. Using stepAIC precludes the use
-of k-fold cross validation.
+## Model Set # 1
+The same person is building one of each requested model. Each of the models will
+follow a similar format, with only the family changing. Each model will start
+with the saturated model including all the individual predictors ***except***
+for `STARS`. There will also be the following interactions:
+
+  * `Rated` and `STARS`
+    * As discussed above, by having `Rating` on its own and this interaction
+    the effect of both having a rating and the effect of the rating level can be
+    measured, without penalizing unrated wines a second time for a 0.
+  * The full four-way interaction between `AcidIndex`, `FixedAcidity`,
+  `VolatileAcidity`, and `pH`
+    * There are **four** separate predictors relating to acidity. Despite their
+    Pearson product-moment correlations being low, I am suspicious of too much
+    weight being given to acidity. By including their interactions, effects can
+    be tempered, or magnified, as necessary. `CitricAcid` is not included here
+    as is not merely an acid but a flavor provider as well. Also, five-way
+    interactions lead to migraines.
+  * `FreeSulfurDioxide` and `TotalSulfurDioxide`
+    * Again, a case of multiple theoretically related predictors
+
+A forward and backwards stepwise procedure based on AIC will be used, and the
+model with the lowest AIC on the training set will be the selected to be tested
+against the testing set. Using stepAIC precludes the use of k-fold cross
+validation.
 
@@
 
-### Poisson Model
+### Poisson Model # 1
 
-```{r m1Poisson}
+```{r m1PTrain}
 set.seed(350047)
-m1PFit <- train(TARGET ~ . - INDEX - STARS + Rated:STARS, data = m1trnXimp,
+m1PFit <- train(TARGET ~ . + FreeSulfurDioxide:TotalSulfurDioxide +
+                  AcidIndex * FixedAcidity * VolatileAcidity * pH +
+                  Rated:STARS - INDEX - STARS, data = m1trnXimp,
                 trControl = trc, method = 'glmStepAIC',
                 family = poisson(link = 'log'), direction = 'both', trace = 0)
-# Print results
-kable(summary(m1PFit$finalModel)$coefficients, digits = 3L,
+# Save AIC
+m1PFitAIC <- AIC(m1PFit$finalModel)
+```
+\small
+```{r m1PTable}
+# Print results. This needed to ne in its own section so that the LaTeX commands
+# to change the text size can be wrapped around it.
+kable(summary(m1PFit$finalModel)$coefficients,
       caption = "Model 1 Poisson Regression Output")
 ```
+\normalsize
 
+#### Coefficient Discussion
+The absolute magnitude of the coefficients makes sense. Factors related to
+acidity tend to lower the number of cases bought. The factors with the greatest
+magnitude effect are all positive, and they are being rated, the rating level,
+and the label appeal. Keeping everything else equal, merely being rated
+increases the number of cases by about \(e^{0.6509117} \approx  1.9\)!
+
+All the coefficients are significant at at least the 5% level except for
+Sulphates, but in terms of AIC, leaving it in provided a better model than
+taking it out
+
+#### Variable Importance
+The further down the list below a variable preceded by a `-` is, the more
+important it is, as its removal would cause a greater disruption to the
+deviance and the AIC. The three rating variables, being rated, the rating level,
+and the label appeal, are the most important variables when it comes to future
+case purchases.
+
+It is also interesting to note that the model including `VolatileAcidity` as 
+a singleton predictor has a lower deviance than the selected model, but that the
+drop in deviance is not enough to offset the possible parameter error. However,
+it makes itself felt through its interaction with `AcidIndex` in the selected
+model.
 ```
                                               Df Deviance   AIC
-<none>                    9654.5 31998
-- Sulphates           1   9657.1 31998
-+ Density             1   9654.1 31999
-+ ResidualSugar       1   9654.2 31999
-- Alcohol             1   9658.3 32000
-+ CitricAcid          1   9654.4 32000
-+ FixedAcidity        1   9654.4 32000
-- pH                  1   9659.0 32000
-- Chlorides           1   9659.5 32001
-- FreeSulfurDioxide   1   9659.7 32001
-- TotalSulfurDioxide  1   9670.3 32012
-- VolatileAcidity     1   9670.9 32012
-- AcidIndex           1   9880.9 32222
-- LabelAppeal         1  10102.7 32444
-- `STARS:RatedRated`  1  10335.1 32676
-- RatedRated          1  10337.5 32679
+<none>                                             9644.0 31993
++ VolatileAcidity                              1   9642.4 31994
+- Sulphates                                    1   9646.5 31994
+- `FreeSulfurDioxide:TotalSulfurDioxide`       1   9647.1 31994
++ pH                                           1   9643.5 31995
++ Density                                      1   9643.6 31995
++ ResidualSugar                                1   9643.6 31995
++ `VolatileAcidity:pH`                         1   9643.7 31995
++ CitricAcid                                   1   9643.9 31995
+- Alcohol                                      1   9647.9 31995
++ `FixedAcidity:VolatileAcidity`               1   9643.9 31995
++ `FixedAcidity:pH:AcidIndex`                  1   9644.0 31995
++ `FixedAcidity:VolatileAcidity:pH`            1   9644.0 31995
++ `VolatileAcidity:pH:AcidIndex`               1   9644.0 31995
++ `FixedAcidity:VolatileAcidity:AcidIndex`     1   9644.0 31995
++ `FixedAcidity:pH`                            1   9644.0 31995
++ `FixedAcidity:VolatileAcidity:pH:AcidIndex`  1   9644.0 31995
+- FixedAcidity                                 1   9648.1 31995
+- `FixedAcidity:AcidIndex`                     1   9648.4 31996
+- Chlorides                                    1   9648.9 31996
+- `pH:AcidIndex`                               1   9648.9 31996
+- FreeSulfurDioxide                            1   9652.3 32000
+- `VolatileAcidity:AcidIndex`                  1   9661.9 32009
+- TotalSulfurDioxide                           1   9662.0 32009
+- AcidIndex                                    1   9678.8 32026
+- LabelAppeal                                  1  10092.2 32439
+- `STARS:RatedRated`                           1  10322.8 32670
+- RatedRated                                   1  10325.1 32672
 ```
+
+### Negative Binomial Model # 1
+
+### Multiple Linear Regrssion Model # 1
+
+## Poisson Model #2
+
+## Negative Binomial Model #2
+
+## Multiple Linear Regrssion #2
 
@@
 ## Model Selection Criteria
-We will look at RMSE and AIC as our metrics and select the model which performs
-best on both, Ties will be broken by \(R^2\).
+We will look at RMSE and MAE as our metrics and select the model which performs
+best on both. If no model peforms best on both, the AIC of the model results on
+the training set will be used to break the tie.
 
-# References
+## Model Test Results
+
+### Poisson Model 1
+
+### Poisson Model 2
+
+### Negative Binomial Model 1
+
+### Negative Binomial Model 2
+
+### Multiple Linear Regression Model 1
+
+### Multiple Linear Regression Model 2
+
+## Comparsion
+
+## Selection
+
+# PREDICTION
+
+# REFERENCES
 
@@
 ```
+```{r m1PTrain}
+```
+```{r m1PTable}
+```

### 2020-11-20 HW5/HW5-Draft.Rmd
Reflect shift in models built. Add distribution of TARGET. Change Poisson model.

@@
 
+## TARGET variable
+Lastly, the complete dataset exhibits an average of
+`r prettyNum(mean(DT$TARGET), digits = 3L)` cases being bought, with the
+distribution below:
+
+```{r targetH}
+ggplot(DT[, .(TARGET)], aes(x = TARGET)) +
+  geom_histogram(binwidth = FDbin, fill = 'darkolivegreen4')
+```
+
 # DATA PREPARATION
@@
 
-## Modeler # 1
+## Poisson Model 1 & Negative Binomial Model 1
 ### Missing Data
@@
 # BUILD MODELS
-## Model Set # 1
-The same person is building one of each requested model. Each of the models will
-follow a similar format, with only the family changing. Each model will start
-with the saturated model including all the individual predictors ***except***
-for `STARS`. There will also be the following interactions:
+## Poisson Model 1 & Negative Binomial Model 1
+The same person is building these two models, and the model setups will be
+similar. Each model will follow a similar format, with only the family changing.
+Each will start with the saturated model including all the individual predictors
+***except*** for `STARS`. There will also be the following interactions:
 
@@
     * Again, a case of multiple theoretically related predictors
+  * `LabelAppeal` with `Rated` and `LabelAppeal` with `Rated:STARS`
+    * This is a later addition. The top three predictors in magnitude were
+    `LabelAppeal`, `Rated`, and `Rated:STARS`. With all three being so strong,
+    there may be interactions between them. However, as `STARS` alone has been
+    removed from contention, only the interactions between the two others will
+    be considered.
 
@@
                   AcidIndex * FixedAcidity * VolatileAcidity * pH +
-                  Rated:STARS - INDEX - STARS, data = m1trnXimp,
+                  Rated:STARS + LabelAppeal:Rated + LabelAppeal:Rated:STARS -
+                  INDEX - STARS, data = m1trnXimp,
                 trControl = trc, method = 'glmStepAIC',
@@
 ```
+
 \small
@@
 
+#### Model Checking
+
+```{r m1PFitCheck}
+m1PFitDisp <- sum(residuals(m1PFit$finalModel, type = 'pearson') ^ 2) /
+  m1PFit$finalModel$df.residual
+```
+
+A Poisson model should have a dispersion parameter close to 1, as the Poisson by
+definition has its mean and variance equal. This model has a dispersion
+parameter of `r prettyNum(m1PFitDisp, digits = 3L)` which is deemed acceptable.
+
 #### Coefficient Discussion
+Similar to the master data set, the training data set has an empirical mean
+purchased cases of `r prettyNum(mean(m1trnXimp$TARGET), digits = 3L)`. Having no
+other information, the Poisson model expects about
+`r prettyNum(exp(m1PFit$finalModel$coefficients[[1]]), digits = 3L)` cases, due
+to the intercept value of `r m1PFit$finalModel$coefficients[[1]]`. This means
+that the model identifies factors which lean more to *increasing* purchases than
+to decreasing purchases.
+
 The absolute magnitude of the coefficients makes sense. Factors related to
 acidity tend to lower the number of cases bought. The factors with the greatest
-magnitude effect are all positive, and they are being rated, the rating level,
-and the label appeal. Keeping everything else equal, merely being rated
-increases the number of cases by about \(e^{0.6509117} \approx  1.9\)!
+magnitude effect are all positive, and they are the fact of being rated, and the
+interactions between that and the rating level and label appeal. Keeping
+everything else equal, merely being rated increases the number of cases by about
+\(e^{0.6677926} \approx  1.95\)! Each additional star is predicted to contribute
+another 1.2 cases to the total purchase. Interestingly, `LabelAppeal` has fallen
+out of the model; its contributions are through its interactions. As surmised,
+looking at all three individually may have proven too strong, as the three-way
+interaction coefficient tempers the others.
 
 All the coefficients are significant at at least the 5% level except for
-Sulphates, but in terms of AIC, leaving it in provided a better model than
-taking it out
+Sulphates, but in terms of AIC, leaving it in provided a better model than did
+taking it out.
 
@@
 
-It is also interesting to note that the model including `VolatileAcidity` as 
-a singleton predictor has a lower deviance than the selected model, but that the
-drop in deviance is not enough to offset the possible parameter error. However,
-it makes itself felt through its interaction with `AcidIndex` in the selected
-model.
+It is also interesting to note that there are a number of variables whose
+inclusion as predictors would result in a lower deviance model than the
+selected, but said drop in deviance was not enough to offset the possible
+parameter error. These include `VolatileAcidity`, `LabelAppeal`, `pH`,
+`Density`, `ResidualSugar`, `VolatileAcidity:pH`, and `CitricAcid`.
+
 ```
                                               Df Deviance   AIC
-<none>                                             9644.0 31993
-+ VolatileAcidity                              1   9642.4 31994
-- Sulphates                                    1   9646.5 31994
-- `FreeSulfurDioxide:TotalSulfurDioxide`       1   9647.1 31994
-+ pH                                           1   9643.5 31995
-+ Density                                      1   9643.6 31995
-+ ResidualSugar                                1   9643.6 31995
-+ `VolatileAcidity:pH`                         1   9643.7 31995
-+ CitricAcid                                   1   9643.9 31995
-- Alcohol                                      1   9647.9 31995
-+ `FixedAcidity:VolatileAcidity`               1   9643.9 31995
-+ `FixedAcidity:pH:AcidIndex`                  1   9644.0 31995
-+ `FixedAcidity:VolatileAcidity:pH`            1   9644.0 31995
-+ `VolatileAcidity:pH:AcidIndex`               1   9644.0 31995
-+ `FixedAcidity:VolatileAcidity:AcidIndex`     1   9644.0 31995
-+ `FixedAcidity:pH`                            1   9644.0 31995
-+ `FixedAcidity:VolatileAcidity:pH:AcidIndex`  1   9644.0 31995
-- FixedAcidity                                 1   9648.1 31995
-- `FixedAcidity:AcidIndex`                     1   9648.4 31996
-- Chlorides                                    1   9648.9 31996
-- `pH:AcidIndex`                               1   9648.9 31996
-- FreeSulfurDioxide                            1   9652.3 32000
-- `VolatileAcidity:AcidIndex`                  1   9661.9 32009
-- TotalSulfurDioxide                           1   9662.0 32009
-- AcidIndex                                    1   9678.8 32026
-- LabelAppeal                                  1  10092.2 32439
-- `STARS:RatedRated`                           1  10322.8 32670
-- RatedRated                                   1  10325.1 32672
+<none>                                             9550.7 31902
++ VolatileAcidity                              1   9549.4 31903
+- Sulphates                                    1   9553.5 31903
+- `FreeSulfurDioxide:TotalSulfurDioxide`       1   9553.7 31903
++ LabelAppeal                                  1   9549.9 31903
++ pH                                           1   9550.3 31904
++ Density                                      1   9550.4 31904
++ ResidualSugar                                1   9550.4 31904
++ `VolatileAcidity:pH`                         1   9550.5 31904
++ CitricAcid                                   1   9550.6 31904
+- FixedAcidity                                 1   9554.6 31904
++ `FixedAcidity:pH:AcidIndex`                  1   9550.7 31904
++ `FixedAcidity:VolatileAcidity`               1   9550.7 31904
++ `FixedAcidity:pH`                            1   9550.7 31904
++ `FixedAcidity:VolatileAcidity:pH`            1   9550.7 31904
++ `VolatileAcidity:pH:AcidIndex`               1   9550.7 31904
++ `FixedAcidity:VolatileAcidity:AcidIndex`     1   9550.7 31904
++ `FixedAcidity:VolatileAcidity:pH:AcidIndex`  1   9550.7 31904
+- `FixedAcidity:AcidIndex`                     1   9554.9 31904
+- Alcohol                                      1   9555.1 31904
+- Chlorides                                    1   9555.2 31904
+- `pH:AcidIndex`                               1   9555.4 31905
+- FreeSulfurDioxide                            1   9558.6 31908
+- `LabelAppeal:STARS:RatedRated`               1   9565.2 31914
+- TotalSulfurDioxide                           1   9568.7 31918
+- `VolatileAcidity:AcidIndex`                  1   9569.6 31919
+- AcidIndex                                    1   9585.3 31935
+- `LabelAppeal:RatedRated`                     1   9719.7 32069
+- `STARS:RatedRated`                           1  10152.1 32501
+- RatedRated                                   1  10247.3 32597
 ```
@@
 ```
+```{r targetH}
+```
 ```{r trainTestSplit}
@@
 ```
+```{r m1PFitCheck}
+```

### 2020-11-20 HW5/HW5-Draft.Rmd
Too close to Shabbos, must stop.

@@
 
-A forward and backwards stepwise procedure based on AIC will be used, and the
-model with the lowest AIC on the training set will be the selected to be tested
-against the testing set. Using stepAIC precludes the use of k-fold cross
-validation.
-
-```{r m1TrainControl}
-# All Modeler 1 models will use stepAIC so turn off cross-validation
-trc <- trainControl(method = 'none')
-```
-
 ### Poisson Model # 1
+For the Poisson model, a forward and backwards stepwise procedure based on AIC
+will be used, and the model with the lowest AIC on the training set will be the
+selected to be tested against the testing set. Using stepAIC precludes the use
+of k-fold cross validation.
 
 ```{r m1PTrain}
+# Using stepAIC so turn off cross-validation
+trc <- trainControl(method = 'none')
 set.seed(350047)
@@
 ### Negative Binomial Model # 1
+```{r m1NBTrain, cache=TRUE}
+glmc <- glm.control(maxit = 25, epsilon = 1e-12)
+set.seed(72)
+m1NBFit2a <- glm.nb(TARGET ~ . + FreeSulfurDioxide:TotalSulfurDioxide +
+                      AcidIndex * FixedAcidity * VolatileAcidity * pH +
+                      Rated:STARS + LabelAppeal:Rated +
+                      LabelAppeal:Rated:STARS - INDEX - STARS,
+                    data = m1trnXimp, control = glmc, link = log,
+                    na.action = na.exclude)
 
-### Multiple Linear Regrssion Model # 1
+m1NBFit2b <- stepAIC(m1NBFit2a, direction = 'both', trace = 1)
+
+# Save AIC
+m1NBFitAIC <- AIC(m1NBFit2b)
+```
+
+\small
+```{r m1NBTable}
+# Print results. This needed to ne in its own section so that the LaTeX commands
+# to change the text size can be wrapped around it.
+kable(summary(m1NBFit2b)$coefficients,
+      caption = "Model 1 Negative Binomial Regression Output")
+```
+\normalsize
 
@@
 
-## Multiple Linear Regrssion #2
+## Multiple Linear Regrssion Model # 1
+
+## Multiple Linear Regrssion Model # 2
 
@@
 ```
-```{r m1TrainControl}
-```
 ```{r m1PTrain}

### 2020-11-21 Final_project/Draft_EDA_start.Rmd
Structure and further EDA

@@
 ---
-title: "Final Project"
-author: "John Kellogg"
-date: "11/17/2020"
-output: html_document
+title: "DATA 621 - Business Analytics and Data Mining"
+subtitle: 'Fall 2020 - Group 2 - Homework #3'
+author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
+    Bryan Persaud, Jeff Shamp
+date: "10/28/2020"
+output:
+  pdf_document: default
+urlcolor: purple
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
+```
 
+```{r libraries}
 library(tidyverse)
@@
 
-## Research question
-
-If we know the number of likes, can we predict the number of views a video will receive?
-
-sub questions:
-For each x increase/decrease in views, dislikes, and comments, how much does the view count move?
-Does turning off comments and ratings affect and by how much
-Is that number affected by country?
-Is this the same across categories?
-
-## Data Exploration
-
-```{r cars}
+```{r load_data}
 CA_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/CAvideos.csv"
@@
 FR_set <- read_csv(file = FR_path)
+DE_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/DEvideos.csv"
+DE_set <- read_csv(file = DE_path)
 ```
 
+# Abstract 
+`Short summary of the research problem and its importance, what you do and what you find
+
+# Introduction
+
+YouTube has changed the future of video entertainment forever 
+and the platform is now a multi-billion dollar industry in itself.  
+In 2013, less than 8 years after its launch, the platform was seeing 
+one billion active users a month. Additionally, in 2019 the 
+platform generated $15 billion in advertising revenue.  YouTube’s 
+model connected a user’s creativity with a desire for global 
+recognition.  Before YouTube, international fame was not conceivable 
+outside of a standard television or movie studio.  Today, creators 
+from all over the world are gaining international prominence using 
+their own equipment and space.   “YouTube is central to today’s video 
+ecosystem,” says Enders Analysis research analyst Jamie McGowan Stuart.
+
+The top 10 YouTube channels for half this year alone amassed over 162 
+million.  The current top channel “Vlad and Nikita” earn around 312 
+thousand per video .   Each video invariably contains at least one 
+advertisement as well as most of the videographers have a separate 
+sponsor who pays them directly.  According to a survey from Google, 
+6 out of 10 people already prefer online video platforms or streaming 
+services over live TV .  Additionally Dr. James McQuivey states, by 2025 
+(4 years from the authorship of this paper), he expects half of viewers 
+under the age of 32 will not have paid for TV service.  Understanding 
+some of the underpinnings of what gets the most views is beneficial to 
+anyone entering or already in the YouTube world.
+
+## Research Question
+
+Using data from kaggle (https://www.kaggle.com/datasnaek/youtube-new), 
+we will identify if there is direct or indirect link between a video’s 
+number of views and an increase or decrease in the number of likes, 
+dislikes, and comments.  We will also use a video’s country of origin and 
+the category or type of video as predictors.
+
+## Methodology
+
+The data contains the daily trending videos from 2017 with up to 200 listed 
+trending videos per day.  We will be utilizing, at minimum, the json’s and 
+CSV’s for Canada, United States, Great Britain, Germany and France, totaling 
+30581 unique values.  The site has data for Japan, Mexico, Korea, India, and 
+Russia; we will attempt to pull them in, if required, for a larger picture.  
+However, the five already chosen datasets should be sufficient to cover the 
+project statement.
+
+The structure of the paper is based on the flow of the research:
+
+* Introduction
+* Data 
+  * The data sets have an expected heteroscedasticity.  To work around this, we structured the data using the XXXXX method.  
+  * After preparing the data, we created randomized subsets of the datasets in a train/test structure for use in the models.
+  * Distribution and Correlation plots are used to further the model selection process
+* Models
+  * We generated X models using methods such as XXXXXX in order to find and chose the most appropriate and accurate model. 
+* Results
+  * We discuss the performance and prediction accuracy of the model
+* Summary and Conclusions
+  * Quick recap of the findings, identify areas of further research and shortcomings found during the research.
+* References
+* Code Appendix
+  * All R code used in the research
+
+# Data
+•Description of Data
+–A complete description of the desired output
+
+## Data Description
+
+## DATA PREPARATION
+
+~discuss using Excel for data cleaning for unique values in France.
+
+```{r}
+# Add Country Code to each set
+# throw out incomplete cases
+# Join all datasets into one large set - may use it may not
+```
+
+## Data Distribution and Correlation
 Canada data
@@
 There is definite heteroskedasticity across all data points, we will need to address.
+
+# Models
+•Data Analysis
+–Describe the instrumentation
+–Describe the analysis plan
+•Describe the scope and limitations of the methodology
+## Model Creation
+
+## Model Selection
+
+### Criteria
+### Performance
+
+### Predictions
+
+# Results
+– Needs to systematically and clearly articulate the study findings.
+If the results are unclear, the reviewer must decide whether the analysis of the data was poorly executed or whether the Results section is poorly organized.
+
+– Should state wheather the hypothesis were verified or proven untrue or, if no hypotheses were given, weather the research questions were answered.
+The authors should also comment on their results in light of previous studies and explain what differences (if any) exist between the findings and those reported by others and attempt to provide an explaination of the differnces.
+
+# Summary and Conclusions
+
+•Recap briefly what you do in the paper
+•Evaluate the effectiveness of your research and provide recommendations (if applicable)
+•Make sure that all of the questions raised in the introduction and the literature review have been addressed
+•Compare the final results against the original aims and objectives
+•Identify any shortcomings and future research
+
+# REFERENCES
+
+ * https://www.ibc.org/trends/analysis-the-youtube-revolution/5796.article 
+ * https://www.cashlady.com/youtube-league/
+ * Google/comScore custom survey, U.S., January 2016.
+(n=2,940 adults aged 18+ who like to watch video content in a typical week; video content defined as TV shows, movies, music videos, videos uploaded by people and/or videos uploaded by brands) https://www.thinkwithgoogle.com/marketing-strategies/video/video-trends-where-audience-watching
+
+ * https://go.forrester.com/blogs/15-10-07-by_2025_50_of_adults_under_age_32_will_not_pay_for_tv/
+
+# CODE APPENDIX
+The code chunks below represent the R code called in order during the analysis.
+They are reproduced in the appendix for review and comment.
+
+```{r appendix, include=FALSE}
+knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
+```

### 2020-11-21 Final_project/Draft_EDA_start.Rmd
Further updates to EDA

@@
 knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
+
+#header-includes:
+#    - \usepackage{setspace}
+#    - \doublespacing
 ```
@@
 ```{r load_data}
-CA_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/CAvideos.csv"
-CA_set <- read_csv(file = CA_path)
-US_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/USvideos.csv"
-US_set <- read_csv(file = US_path)
-GB_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/GBvideos.csv"
-GB_set <- read_csv(file = GB_path)
-FR_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/FRvideos.csv"
-FR_set <- read_csv(file = FR_path)
-DE_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/DEvideos.csv"
-DE_set <- read_csv(file = DE_path)
+ca_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/CAvideos.csv"
+ca_set <- read_csv(file = ca_path)
+us_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/USvideos.csv"
+us_set <- read_csv(file = us_path)
+gb_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/GBvideos.csv"
+gb_set <- read_csv(file = gb_path)
+fr_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/FRvideos.csv"
+fr_set <- read_csv(file = fr_path)
+de_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/DEvideos.csv"
+de_set <- read_csv(file = de_path)
 ```
@@
 
-The data contains the daily trending videos from 2017 with up to 200 listed 
-trending videos per day.  We will be utilizing, at minimum, the json’s and 
-CSV’s for Canada, United States, Great Britain, Germany and France, totaling 
+We will be utilizing, at minimum, the json’s and CSV’s for Canada, 
+United States, Great Britain, Germany and France, totaling 
 30581 unique values.  The site has data for Japan, Mexico, Korea, India, and 
@@
 
+The datasets in total have 186,922 unique values across the 5 countries used. They 
+contain the daily trending videos from 2017 with up to 200 listed trending videos 
+per day.  The first response variable is coded as a 1 when a video's views are above 
+the mean of views in their respective country and a 0 if they are below. The second 
+response varible XXX is coded as a 1 for above the mean of the global population and 
+0 if below.
+
+In all, there are 8 predictors, which include:
+
+Variables|Description   
+-|----
+views | The number of times a video was watched
+category | the type of video, such as "Film & Animation", "Sports", "Music", etc... 
+likes | the number of times a user clicked "like"
+dislikes | the number of times a user clicked "dislike"
+comment_count | the number of comments on an individual video   
+comments_disabled | boolian value if the owner of the video disabled users from commenting   
+ratings_disabled | boolian value if the owner of the video disabled ratings
+country_code | code created to identify country of origin
+view_target | Above (1) or below  (0) the mean views from country of origin
+view_global_target | Above (1) or below  (0) the mean views globally
+
 ## DATA PREPARATION
@@
 
-```{r}
-# Add Country Code to each set
-# throw out incomplete cases
-# Join all datasets into one large set - may use it may not
+```{r ca_data_prep}
+ca_set <- ca_set %>% filter(complete.cases(ca_set))
+
+ca_mean <- mean(ca_set$views)
+
+ca_set <- ca_set%>%
+  mutate(country = 'CA')%>%
+  mutate(country_code = 1)%>%
+  mutate(view_target = ifelse(views > ca_mean,1,0))
+```
+
+```{r de_data_prep}
+de_set <- de_set %>% filter(complete.cases(de_set))
+
+de_mean <- mean(de_set$views)
+
+de_set <- de_set%>%
+  mutate(country = 'DE')%>%
+  mutate(country_code = 2)%>%
+  filter(complete.cases(de_set))%>%
+  mutate(view_target = ifelse(views > de_mean,1,0))
+```
+
+```{r fr_data_prep}
+fr_set <- fr_set %>% filter(complete.cases(fr_set))
+
+fr_mean <- mean(fr_set$views)
+
+fr_set <- fr_set%>%
+  mutate(country = 'FR')%>%
+  mutate(country_code = 3)%>%
+  filter(complete.cases(fr_set))%>%
+  mutate(view_target = ifelse(views > fr_mean,1,0))
+```
+
+```{r gb_data_prep}
+gb_set <- gb_set %>% filter(complete.cases(gb_set))
+
+gb_mean <- mean(gb_set$views)
+
+gb_set <- gb_set%>%
+  mutate(country = 'GB')%>%
+  mutate(country_code = 4)%>%
+  filter(complete.cases(gb_set))%>%
+  mutate(view_target = ifelse(views > gb_mean,1,0))
+```
+
+```{r us_data_prep}
+us_set <- us_set %>% filter(complete.cases(us_set))
+
+us_mean <- mean(us_set$views)
+
+us_set <- us_set%>%
+  mutate(country = 'US')%>%
+  mutate(country_code = 5)%>%
+  filter(complete.cases(us_set))%>%
+  mutate(view_target = ifelse(views > us_mean,1,0))
+```
+
+```{r merge_datasets}
+mergeCols = c('video_id','trending_date','title','channel_title','category_id','publish_time','tags','views','likes','dislikes','comment_count','thumbnail_link','comments_disabled','ratings_disabled','video_error_or_removed','description','country','country_code', 'view_target')
+
+full_set <- ca_set %>% full_join(de_set, by=mergeCols)
+full_set <- full_set %>% 
+  full_join(fr_set, by=mergeCols)%>%
+  full_join(gb_set, by=mergeCols)%>%
+  full_join(us_set, by=mergeCols)
+
+full_mean <- mean(full_set$views)
+
+full_set <- full_set %>%
+  mutate(view_global_target = ifelse(views > full_mean,1,0))
+```
+
+```{r global_target}
+ca_set <- ca_set %>%
+  mutate(view_global_target = ifelse(views > full_mean,1,0))
+
+de_set <- de_set %>%
+  mutate(view_global_target = ifelse(views > full_mean,1,0))
+
+fr_set <- fr_set %>%
+  mutate(view_global_target = ifelse(views > full_mean,1,0))
+
+gb_set <- gb_set %>%
+  mutate(view_global_target = ifelse(views > full_mean,1,0))
+
+us_set <- us_set %>%
+  mutate(view_global_target = ifelse(views > full_mean,1,0))
 ```
@@
 ```{r summary_ca}
-summaryca <- CA_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summaryca <- ca_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
 describe(summaryca)
@@
 
+```{r summary_de}
+summaryde <- de_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+describe(summaryus)
+
+ggpairs(
+  summaryde,
+  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
+  )
+us_corr <- cor(summaryde)
+corrplot(us_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
+
+```
+
 ```{r summary_us}
-summaryus <- US_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summaryus <- us_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
 describe(summaryus)
@@
 ```{r summary_us}
-summarygb <- GB_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summarygb <- gb_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
 describe(summarygb)
@@
 ```{r summary_fr}
-summaryfr <- FR_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summaryfr <- fr_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
 describe(summaryfr)
@@
 corrplot(fr_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
+```
+
+```{r summary_global}
+summarygl <- full_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+describe(summarygl)
+
+ggpairs(
+  summarygl,
+  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
+  )
+
+fr_corr <- cor(summarygl)
+corrplot(fr_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
+```
+
+```{r}
+
 ```

### 2020-11-22 HW5/HW5-Draft.Rmd
cleanup

@@
 library(ggplot2)
-library(scales)
 library(knitr)
+library(MASS)
 library(caret)
@@
 ### Negative Binomial Model # 1
+As the dispersion from the Poisson model was less than 1, at first blush, it is
+not expected that the negative binomial model will outperform the Poisson.
+
 ```{r m1NBTrain, cache=TRUE}
-glmc <- glm.control(maxit = 25, epsilon = 1e-12)
+glmc <- glm.control(maxit = 25, epsilon = 1e-14)
 set.seed(72)
-m1NBFit2a <- glm.nb(TARGET ~ . + FreeSulfurDioxide:TotalSulfurDioxide +
+m1NBFitStart <- glm.nb(TARGET ~ . + FreeSulfurDioxide:TotalSulfurDioxide +
                          AcidIndex * FixedAcidity * VolatileAcidity * pH +
@@
                          LabelAppeal:Rated:STARS - INDEX - STARS,
-                    data = m1trnXimp, control = glmc, link = log,
-                    na.action = na.exclude)
+                       data = m1trnXimp, control = glmc, link = log)
 
-m1NBFit2b <- stepAIC(m1NBFit2a, direction = 'both', trace = 1)
+m1NBFit <- stepAIC(m1NBFitStart, direction = 'both', trace = 0)
 
 # Save AIC
-m1NBFitAIC <- AIC(m1NBFit2b)
+m1NBFitAIC <- AIC(m1NBFit)
 ```
@@
 # to change the text size can be wrapped around it.
-kable(summary(m1NBFit2b)$coefficients,
+kable(summary(m1NBFit)$coefficients,
       caption = "Model 1 Negative Binomial Regression Output")
@@
 \normalsize
+
+```
+                                       Df   AIC
+<none>                                    31905
+- Sulphates                             1 31906
++ pH:AcidIndex                          1 31906
+- FreeSulfurDioxide:TotalSulfurDioxide  1 31906
+- VolatileAcidity:AcidIndex             1 31906
++ Density                               1 31906
++ ResidualSugar                         1 31906
++ VolatileAcidity:pH                    1 31907
++ CitricAcid                            1 31907
++ FixedAcidity:pH                       1 31907
++ FixedAcidity:VolatileAcidity          1 31907
+- FixedAcidity:AcidIndex                1 31907
+- pH                                    1 31907
+- Alcohol                               1 31907
+- Chlorides                             1 31907
+- LabelAppeal:Rated:STARS               1 31917
+```
 

### 2020-11-22 HW5/HW5-Draft.Rmd
Finishing up

@@
 fileTrain = "wine-training-data.csv"
-fileTest = "wine-evaluation-data.csv"
+fileEval = "wine-evaluation-data.csv"
 
@@
 
-```{r targetH}
+```{r targetH, fig.height=2, fig.width=3}
+# Plot distribution of TARGET
 ggplot(DT[, .(TARGET)], aes(x = TARGET)) +
@@
 trnIDX <- createDataPartition(DT$TARGET, p = 0.7)
-trnX <- DT[trnIDX$Resample1, ]
-tstX <- DT[!trnIDX$Resample1, ]
+trnSet <- DT[trnIDX$Resample1, ]
+tstSet <- DT[!trnIDX$Resample1, ]
 ```
@@
 ```{r model1addVars}
-# Copy the training set to leave a pristine version for others
-m1trnX <- copy(trnX)
+# P1 & NB1: Copy the training set to leave a pristine version for others
+m1trn <- copy(trnSet)
 
-# Add the "rated" factor variable
-m1trnX[, Rated := factor(ifelse(is.na(STARS), 'Unrated', 'Rated'),
+# P1 & NB1: Add the "rated" factor variable
+m1trn[, Rated := factor(ifelse(is.na(STARS), 'Unrated', 'Rated'),
                         levels = c('Unrated', 'Rated'),
@@
 ```{r model1impute}
-# Create bagged-tree imputation model
+# P1 & NB1: Create bagged-tree imputation model
 set.seed(89)
-m1trnXi <- preProcess(m1trnX, method =c('nzv', 'bagImpute'))
+m1trnI <- preProcess(m1trn, method =c('nzv', 'bagImpute'))
 
-# Replace missing STARS with 0s
-m1trnX[is.na(STARS), STARS := 0L]
+# P1 & NB1: Replace missing STARS with 0s
+m1trn[is.na(STARS), STARS := 0L]
 
-# Impute the remaining NAs
-m1trnXimp <- predict(m1trnXi, newdata = m1trnX)
+# P1 & NB1: Impute the remaining NAs
+m1trnImp <- predict(m1trnI, newdata = m1trn)
 ```
@@
 ```{r m1postImputeD, fig.height=5, fig.width=7}
-# Isolate numeric only predictors
-m1predictorDT <- m1trnXimp[, .SD, .SDcols = nmtrn]
+# P1 & NB1: Isolate numeric only predictors
+m1predictorDT <- m1trnImp[, .SD, .SDcols = nmtrn]
 
-# Melt them from wide to long format
+# P1 & NB1: Melt them from wide to long format
 m1predictorDTM <- melt(m1predictorDT, variable.name = 'metric',
@@
 
-# Using epanechnikov kernel to generate kernel-smoothed densities
+# P1 & NB1: Using Epanechnikov kernel to generate kernel-smoothed densities
 ggplot(predictorDTM[metric %chin% nmtrnDUB], aes(x = value)) +
@@
 ```{r m1postImputeH, fig.height=2, fig.width=7}
-# Generate histogram
+# P1 & NB1: Generate histogram
 ggplot(m1predictorDTM[metric %chin% nmtrnINT], aes(x = value)) +
@@
 ```{r m1PTrain}
-# Using stepAIC so turn off cross-validation
+# P1: Using stepAIC so turn off cross-validation
 trc <- trainControl(method = 'none')
@@
                   Rated:STARS + LabelAppeal:Rated + LabelAppeal:Rated:STARS -
-                  INDEX - STARS, data = m1trnXimp,
+                  INDEX - STARS, data = m1trnImp,
                 trControl = trc, method = 'glmStepAIC',
                 family = poisson(link = 'log'), direction = 'both', trace = 0)
-# Save AIC
+# P1: Save AIC
 m1PFitAIC <- AIC(m1PFit$finalModel)
@@
 ```{r m1PTable}
-# Print results. This needed to ne in its own section so that the LaTeX commands
-# to change the text size can be wrapped around it.
+# P1:Print results. This needed to ne in its own section so that the LaTeX
+# commands to change the text size can be wrapped around it.
 kable(summary(m1PFit$finalModel)$coefficients,
@@
 ```{r m1PFitCheck}
+# P1: Check the dispersion of the model
 m1PFitDisp <- sum(residuals(m1PFit$finalModel, type = 'pearson') ^ 2) /
@@
 Similar to the master data set, the training data set has an empirical mean
-purchased cases of `r prettyNum(mean(m1trnXimp$TARGET), digits = 3L)`. Having no
+purchased cases of `r prettyNum(mean(m1trnImp$TARGET), digits = 3L)`. Having no
 other information, the Poisson model expects about
@@
 #### Variable Importance
-The further down the list below a variable preceded by a `-` is, the more
-important it is, as its removal would cause a greater disruption to the
+
+```{r m1PvarImp}
+# P1: Extract the Poisson variable importance and order it for display
+vIP <- varImp(m1PFit$finalModel)
+vIPn <- row.names(vIP)
+vIPDT <- data.table(Coefficients = vIPn, vIP)
+setorder(vIPDT, -Overall)
+kable(vIPDT, digits = 2L, caption = "Poisson Model Variable Importance")
+```
+
+The variable importance can also be seen from the trace of the step AIC
+procedure. The further down the list below a variable preceded by a `-` is, the
+more important it is, as its removal would cause a greater disruption to the
 deviance and the AIC. The three rating variables, being rated, the rating level,
@@
 
-```{r m1NBTrain, cache=TRUE}
-glmc <- glm.control(maxit = 25, epsilon = 1e-14)
+```{r m1NBTrain, cache=0}
+# NB1: There is no built-in AIC stepping in caret, so using basic glm.nb and
+# then calling stepAIC on saturated model
+glmc <- glm.control(maxit = 40, epsilon = 1e-8)
 set.seed(72)
@@
                          LabelAppeal:Rated:STARS - INDEX - STARS,
-                       data = m1trnXimp, control = glmc, link = log)
+                       data = m1trnImp, control = glmc, link = log)
 
@@
 
-# Save AIC
+# NB1: Save AIC
 m1NBFitAIC <- AIC(m1NBFit)
@@
 ```{r m1NBTable}
-# Print results. This needed to ne in its own section so that the LaTeX commands
-# to change the text size can be wrapped around it.
+# NB1: Print results. This needed to be in its own section so that the LaTeX
+# commands to change the text size can be wrapped around it.
 kable(summary(m1NBFit)$coefficients,
@@
 
+#### Coefficient Discussion
+Similar to the Poisson model, having no other information, the negative binomial
+model expects about `r prettyNum(exp(m1NBFit$coefficients[[1]]), digits = 3L)`
+cases, due to the intercept value of `r m1NBFit$coefficients[[1]]`. This means
+that this model as well identifies factors which lean more to *increasing*
+purchases than to decreasing purchases.
+
+The absolute magnitude of the coefficients makes less sense in this model than
+in the Poisson. The factors related to appeal do increase the number of cases
+bought. However, the factors related to acidity tend to show more back-and-forth
+interactions than in the Poisson model.
+
+Also, there are more "insignificant" coefficients in this model, but these are
+all variables for which the AIC would increase should they have been removed.
+
+#### Variable Importance
+
+```{r m1NBvarImp}
+# NB1: Extract the NegBinom variable importance and order it for display
+vINB <- varImp(m1NBFit)
+vINBn <- row.names(vINB)
+vINBDT <- data.table(Coefficients = vINBn, vINB)
+setorder(vINBDT, -Overall)
+kable(vINBDT, digits = 2L,
+      caption = "Negative Binomial Model Variable Importance")
 ```
-                                       Df   AIC
-<none>                                    31905
-- Sulphates                             1 31906
-+ pH:AcidIndex                          1 31906
-- FreeSulfurDioxide:TotalSulfurDioxide  1 31906
-- VolatileAcidity:AcidIndex             1 31906
-+ Density                               1 31906
-+ ResidualSugar                         1 31906
-+ VolatileAcidity:pH                    1 31907
-+ CitricAcid                            1 31907
-+ FixedAcidity:pH                       1 31907
-+ FixedAcidity:VolatileAcidity          1 31907
-- FixedAcidity:AcidIndex                1 31907
-- pH                                    1 31907
-- Alcohol                               1 31907
-- Chlorides                             1 31907
-- LabelAppeal:Rated:STARS               1 31917
+
+The trace of the step procedure is less complete when using the `glm.nb`
+function in `R` and is not listed here.
+
+Lastly, the deviance for the negative binomial model is actually lower than that
+for the Poisson model, but its AIC is higher. Based on AIC only, the Poisson
+model's parsimony beats out the negative binomial's explanatory powers.
+
+```{r m1PNB}
+m1PNBC <- data.table(Model = c("Poisson 1", "Negative Binomial 1"),
+                     Deviance = c(m1PFit$finalModel$deviance,
+                                  m1NBFit$deviance),
+                     AIC = c(m1PFit$finalModel$aic,
+                             m1NBFit$aic))
+kable(m1PNBC, digits = 3L,
+      caption = "Poisson 1 and NegBinom 1 Goodness-of-Fit Comparison")
 ```
@@
 We will look at RMSE and MAE as our metrics and select the model which performs
-best on both. If no model peforms best on both, the AIC of the model results on
-the training set will be used to break the tie.
+best on both. If no model performs best on both, the AIC of the model results on
+the training set will be used to break the tie, as the AIC is an estimate of the
+average out-of-sample performance of the model.
+
+```{r compTable}
+compTable <- data.table(Models = c("Poisson 1", "Poisson 2",
+                                   "Neg. Binom. 1", "Neg. Binom. 2",
+                                   "MultiLinear 1", "MultiLinear 2"),
+                        RMSE = double(6), MAE = double(6),
+                        TrainAIC = double(6))
+```
 
 ## Model Test Results
+The test set data needs to be processed identically to the training set. Since
+the number of cases needs to be integral the predicted values will be rounded to
+the nearest integer.
 
 ### Poisson Model 1
+
+```{r m1TestProcess}
+# P1 & NB1: process test data
+m1tst <- copy(tstSet)
+m1tst[, Rated := factor(ifelse(is.na(STARS), 'Unrated', 'Rated'),
+                        levels = c('Unrated', 'Rated'),
+                        labels = c('Unrated', 'Rated'))]
+set.seed(89)
+m1tstI <- preProcess(m1tst, method =c('nzv', 'bagImpute'))
+m1tst[is.na(STARS), STARS := 0L]
+m1tstImp <- predict(m1tstI, newdata = m1tst)
+```
+```{r m1PTest}
+# P1: Test Model
+m1PtstPred <- round(predict(m1PFit, newdata = m1tstImp))
+
+compTable[1, 2] <- m1PRMSE <- RMSE(m1PtstPred, m1tstImp$TARGET)
+compTable[1, 3] <- m1PMAE <- MAE(m1PtstPred, m1tstImp$TARGET)
+compTable[1, 4] <- m1PFitAIC
+```
+
+**Poisson Model 1** has an RMSE of `r prettyNum(m1PRMSE, digits = 2L)`, an MAE
+of `r prettyNum(m1PMAE, digits = 2L)`, and a training AIC of
+`r prettyNum(m1PFitAIC, digits = 2L)`.
 
@@
 ### Negative Binomial Model 1
+```{r m1NBTest}
+# NB1: Test Model
+m1NBtstPred <- round(predict(m1NBFit, newdata = m1tstImp))
+compTable[3, 2] <- m1NBRMSE <- RMSE(m1NBtstPred, m1tstImp$TARGET)
+compTable[3, 3] <- m1NBMAE <- MAE(m1NBtstPred, m1tstImp$TARGET)
+compTable[3, 4] <- m1NBFitAIC
+```
+
+**Negative Binomial Model 1** has an RMSE of
+`r prettyNum(m1NBRMSE, digits = 2L)`, an MAE of
+`r prettyNum(m1NBMAE, digits = 2L)`, and a training AIC of
+`r prettyNum(m1NBFitAIC, digits = 2L)`.
 
@@
 ## Comparsion
+
+```{r modelTestResults}
+# Compare results on test set
+kable(compTable, digits = 3L,
+      caption = "Model Results on Test Set")
+```
 
@@
 ```
+<!-- Assignment -->
 ```{r loadData}
 ```
+<!-- Data Exploration -->
 ```{r missingVal}
@@
 ```
+<!-- Data Preparation -->
 ```{r trainTestSplit}
@@
 ```
+<!-- Build Models -->
 ```{r m1PTrain}
@@
 ```
+```{r m1PvarImp}
+```
+```{r m1NBTrain}
+```
+```{r m1NBTable}
+```
+```{r m1NBvarImp}
+```
+```{r m1PNB}
+```
+<!-- Other prep and build goes here -->
+<!-- Select Models -->
+```{r compTable}
+```
+```{r m1TestProcess}
+```
+```{r m1PTest}
+```
+<!-- Poisson 2 Test goes here -->
+```{r m1NBTest}
+```
+<!-- Remaining Test sets go here -->
+```{r modelTestResults}
+```

### 2020-11-22 HW5/HW5-Draft.Rmd
Think I'm done my section

@@
 `STARS` is especially sparse, but as noted in the assignment, this may be an
-indication in and of its own. It is reasonable to assume that the vinters of a
+indication in and of its own. It is reasonable to assume that the vintners of a
 good wine want it to be rated. If they do not submit it for rating, that may be
@@
 the negative binomial models are either unimodal or have to consecutive values
-sharing the mode, due to their inherent nature as discrete disctibutions. What
+sharing the mode, due to their inherent nature as discrete distributions. What
 is seen here is that 1 star is empirically less common than 0 (unrated) or 2
@@
 ```{r m1PTable}
-# P1:Print results. This needed to ne in its own section so that the LaTeX
+# P1:Print results. This needed to be in its own section so that the LaTeX
 # commands to change the text size can be wrapped around it.
@@
 
-```{r m1NBTrain, cache=0}
+```{r m1NBTrain}
 # NB1: There is no built-in AIC stepping in caret, so using basic glm.nb and
-# then calling stepAIC on saturated model
-glmc <- glm.control(maxit = 40, epsilon = 1e-8)
-set.seed(72)
+# then calling stepAIC on saturated model. Once done, investigated some one-off
+# possible enhancements due to issues with glm.nb
+# See https://stackoverflow.com/questions/11749977/why-does-glm-nb-throw-a-missing-value-error-only-on-very-specific-inputs
+# For speed, only final called model is evaluated.
+set.seed(872)
 m1NBFitStart <- glm.nb(TARGET ~ . + FreeSulfurDioxide:TotalSulfurDioxide +
@@
                          LabelAppeal:Rated:STARS - INDEX - STARS,
-                       data = m1trnImp, control = glmc, link = log)
+                       data = m1trnImp, link = log)
+m1NBFitStep <- stepAIC(m1NBFitStart, direction = 'both', trace = 0)
 
-m1NBFit <- stepAIC(m1NBFitStart, direction = 'both', trace = 0)
+# Last call from automated procedure
+# TARGET ~ FixedAcidity + VolatileAcidity + Chlorides + FreeSulfurDioxide + 
+#     TotalSulfurDioxide + pH + Sulphates + Alcohol + LabelAppeal + 
+#     AcidIndex + Rated + FreeSulfurDioxide:TotalSulfurDioxide + 
+#     FixedAcidity:AcidIndex + VolatileAcidity:AcidIndex + Rated:STARS + 
+#     LabelAppeal:Rated + LabelAppeal:Rated:STARS
+
+# Removing VolatileAcidity improved the AIC
+
+m1NBFit <- glm.nb(formula = TARGET ~ FixedAcidity + Chlorides +
+                    FreeSulfurDioxide + TotalSulfurDioxide + pH + Sulphates +
+                    Alcohol + LabelAppeal + AcidIndex + Rated +
+                    FreeSulfurDioxide:TotalSulfurDioxide +
+                    FixedAcidity:AcidIndex + VolatileAcidity:AcidIndex +
+                    Rated:STARS + LabelAppeal:Rated + LabelAppeal:Rated:STARS,
+                  data = m1trnImp, link = log)
 

### 2020-11-22 Final_project/Draft_EDA_start.Rmd

@@
 view_global_target | Above (1) or below  (0) the mean views globally
+tag_count | number of metadata tags applied to each video
 
@@
   mutate(country_code = 1)%>%
-  mutate(view_target = ifelse(views > ca_mean,1,0))
+  mutate(view_target = ifelse(views > ca_mean,1,0))%>%
+  mutate(tag_count = 1+(str_count(ca_set$tags, ';')))
 ```
@@
   filter(complete.cases(de_set))%>%
-  mutate(view_target = ifelse(views > de_mean,1,0))
+  mutate(view_target = ifelse(views > de_mean,1,0))%>%
+  mutate(tag_count = 1+(str_count(de_set$tags, ';')))
 ```
@@
   filter(complete.cases(fr_set))%>%
-  mutate(view_target = ifelse(views > fr_mean,1,0))
+  mutate(view_target = ifelse(views > fr_mean,1,0))%>%
+  mutate(tag_count = 1+(str_count(fr_set$tags, ';')))
 ```
@@
   filter(complete.cases(gb_set))%>%
-  mutate(view_target = ifelse(views > gb_mean,1,0))
+  mutate(view_target = ifelse(views > gb_mean,1,0))%>%
+  mutate(tag_count = 1+(str_count(gb_set$tags, ';')))
 ```
@@
   filter(complete.cases(us_set))%>%
-  mutate(view_target = ifelse(views > us_mean,1,0))
+  mutate(view_target = ifelse(views > us_mean,1,0))%>%
+  mutate(tag_count = 1+(str_count(us_set$tags, ';')))
 ```
@@
 ```{r merge_datasets}
-mergeCols = c('video_id','trending_date','title','channel_title','category_id','publish_time','tags','views','likes','dislikes','comment_count','thumbnail_link','comments_disabled','ratings_disabled','video_error_or_removed','description','country','country_code', 'view_target')
+mergeCols = c('video_id','trending_date','title','channel_title','category_id','publish_time','tags','views','likes','dislikes','comment_count','thumbnail_link','comments_disabled','ratings_disabled','video_error_or_removed','description','country','country_code', 'view_target','tag_count')
 
@@
 ```{r summary_ca}
-summaryca <- ca_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summaryca <- ca_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
 describe(summaryca)
@@
 ```{r summary_de}
-summaryde <- de_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summaryde <- de_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
 describe(summaryus)
@@
 ```{r summary_us}
-summaryus <- us_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summaryus <- us_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
 describe(summaryus)
@@
 ```{r summary_us}
-summarygb <- gb_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summarygb <- gb_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
 describe(summarygb)
@@
 ```{r summary_fr}
-summaryfr <- fr_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summaryfr <- fr_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
 describe(summaryfr)
@@
 ```{r summary_global}
-summarygl <- full_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
+summarygl <- full_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
 describe(summarygl)
@@
 corrplot(fr_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
-```
-
-```{r}
-
 ```

### 2020-11-24 HW5/HW5-Draft.Rmd
TOC
NB Model 2 & ML Model 1

@@
 library(data.table)
+library(dplyr)
+library(mice)
 
@@
 
+## Multiple Linear Regression 1 & Negative Binomial Model 2
+### Missing Data
+These models will deal with missing numerical data by imputing the NA's
+using the Multivariate imputation by chained equations (MICE) method. Multiple
+imputation involves creating multiple predictions for each missing value, helping 
+to account for the uncertainty in the individual imputations.
+
+```{r model2impute}
+set.seed(123)
+
+# ML1 & NB2 data transformation
+# MICE imputation on train and test set
+m2train <- complete(mice(data = trnSet %>% select(-INDEX),
+                         method = "pmm", print = FALSE), 3)
+
+m2test <- complete(mice(data = tstSet %>% select(-INDEX),
+                        method = "pmm", print = FALSE), 3)
+
+```
+
 # BUILD MODELS
@@
 
+#### Baseline Model 
+We will start with a simple negative binomial regression model to serve as a 
+baseline. This includes all variables in the dataset.
+
+```{r m2NBTrainBase}
+
+m2NBBase <- glm.nb(TARGET ~ ., data = m2train)
+knitr::kable(summary(m2NBBase)$coefficients, digits = 3L,
+             caption = 'Negative Binomial Regression Model 2 - Base Negative Binomial Output')
+
+```
+Immediately, we can see a few variables that do not meet the 0.05
+p-value threshold.
+
+#### Enhanced Model
+Backwards stepwise regression is performed and the result is a model with the 
+following variables *removed*: `FixedAcidity`, `CitricAcid`, `ResidualSugar`, 
+and `Density`. Because the `Sulphates` variable is very close to the 0.05
+p-value threshold of significance, we will remove it as well.
+
+```{r m2NBTrain}
+
+m2NBFit <- glm.nb(TARGET ~ VolatileAcidity + Chlorides +
+                FreeSulfurDioxide + TotalSulfurDioxide + pH + Alcohol + 
+                LabelAppeal + AcidIndex + STARS, data = m2train)
+
+knitr::kable(summary(m2NBFit)$coefficients, digits = 3L,
+             caption = 'Negative Binomial Regression Model 2 - Final Negative Binomial Output')
+
+```
+
+#### Coefficient Discussion
+The intercept of the final model tells us that there is typically ~1 case 
+of wine purchased. Some things to note:
+
+* Stars and label appeal contribute the most positively to the number of
+cases of wine purchased. Intuitively, this makes sense - the higher the
+rating and the prettier the bottle, the more cases we would expect to be
+sold. 
+* Free Sulfur Dioxide and Total Sulfur Dioxide have a very tiny effect on
+the number of cases sold (rounded to 3 decimal places, the impact is 0!)
+* All of the remaining variables except alcohol have a negative effect on 
+the number of cases purchased.
+
 ## Multiple Linear Regrssion Model # 1
+One drawback to traditional multiple linear regression models is that the 
+forecasted counts are not integers. We will deal with this by rounding the
+predictions to the nearest whole number.
+
+#### Baseline Model
+We will start with a simple linear model to serve as a baseline. This includes
+all variables in the dataset.
+
+```{r m1MLTrainBase}
+
+m1MLBase <- lm(TARGET ~ ., data = m2train)
+knitr::kable(summary(m1MLBase)$coefficients, digits = 3L,
+             caption = 'Multiple Regression Model 1 - Base Linear Regression Output')
+
+```
+Most of the variables in the dataset are significant, but we can immediately
+see a few that are above the 0.05 p-value threshold.
+
+#### Enhanced Model
+Backwards stepwise regression is performed and the result is a model with the 
+following variables *removed*: `FixedAcidity`, `CitricAcid`, `ResidualSugar`, 
+and `Density`. 
+```{r m1MLTrain}
+
+m1MLFit <- lm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide + TotalSulfurDioxide + pH +
+             Sulphates + Alcohol + LabelAppeal + AcidIndex + STARS, data = m2train)
+knitr::kable(summary(m1MLFit)$coefficients, digits = 3L,
+             caption = 'Multiple Regression Model 1 - Final Linear Regression Output')
+
+```
+
+#### Coefficient Discussion
+The intercept of the final model tells us that there are typically ~3 cases 
+of wine purchased. Some things to note:
+
+* As expected, the number of stars has the largest, positive impact on the 
+number of cases sold. This makes sense -- the higher the rating of a wine,
+the more of it we would expect to be purchased. 
+* Label appeal has the second largest impact on the number of cases sold. 
+They say one shouldn't judge a book by its cover (or in this case, a wine
+by its label), but it does appear to have quite a significant impact on
+sales. 
+* Alcohol also has a small positive impact on the number of cases purchased. 
+This means that the more alcohol a wine has, the more cases of it will be
+purchased. 
+* Acid Index, Sulphates, pH, Chlorides, and VolatileAcidity all have a
+negative impact on the number of cases sold.
 
@@
 
+```{r m2NBTest}
+# NB2: Test Model
+m2NBtstPred <- round(predict(m2NBFit, newdata = m2test))
+compTable[4, 2] <- m2NBRMSE <- RMSE(m2NBtstPred, m2test$TARGET)
+compTable[4, 3] <- m2NBMAE <- MAE(m2NBtstPred, m2test$TARGET)
+compTable[4, 4] <- m2NBFitAIC <- AIC(m2NBFit)
+```
+
+**Negative Binomial Model 2** has an RMSE of
+`r prettyNum(m2NBRMSE, digits = 2L)`, an MAE of
+`r prettyNum(m2NBMAE, digits = 2L)`, and a training AIC of
+`r prettyNum(m2NBFitAIC, digits = 2L)`.
+
 ### Multiple Linear Regression Model 1
+```{r m1MLTest}
+# ML1: Test Model
+m1MLtstPred <- round(predict(m1MLFit, newdata = m2test))
+compTable[5, 2] <- m1MLMSE <- RMSE(m1MLtstPred, m2test$TARGET)
+compTable[5, 3] <- m1MLMAE <- MAE(m1MLtstPred, m2test$TARGET)
+compTable[5, 4] <- m1MLFitAIC <- AIC(m1MLFit)
+
+```
+
+**Multiple Linear Regression Model 1** has an RMSE of
+`r prettyNum(m1MLMSE, digits = 2L)`, an MAE of
+`r prettyNum(m1MLMAE, digits = 2L)`, and a training AIC of
+`r prettyNum(m1MLFitAIC, digits = 2L)`.
 
@@
 ```
+```{r model2impute}
+```
 <!-- Build Models -->
@@
 ```
+```{r m2NBTrainBase}
+```
+```{r m2NBTrain}
+```
+```{r m1MLTrainBase}
+```
+```{r m1MLTrain}
+```
 <!-- Other prep and build goes here -->
@@
 ```
-<!-- Remaining Test sets go here -->
+```{r m2NBTest}
+```
+```{r m1MLTest}
+```
+<!-- ML 2 Test goes here -->
 ```{r modelTestResults}

### 2020-11-24 HW5/HW5-Draft.Rmd
Poisson 2 & MLR 1 added

--- 2020-11-22 HW5/HW5-Draft.Rmd
@@
     Bryan Persaud, Jeff Shamp
-date: "11/22/2020"
+date: "11/28/2020"
 output:
@@
 library(data.table)
+library(mice)
+library(pscl)
+library(mpath)
+library(dplyr)
 
@@
 the United States. The more sample cases purchased, the more likely is a wine to
-be sold at a high end restaurant.A large wine manufacturer is studying the data
+be sold at a high end restaurant. A large wine manufacturer is studying the data
 in order to predict the number of wine cases ordered based upon the wine
@@
 |VARIABLE NAME|DEFINITION|THEORETICAL EFFECT|
-|--|--|--|
+|--|----|---|
 |INDEX|Identification Variable (do not use)|None|
@@
 
-```{r targetH, fig.height=2, fig.width=3}
+```{r targetH, fig.height=2, fig.width=3, fig.align='center'}
 # Plot distribution of TARGET
@@
 
-## Poisson Model 1 & Negative Binomial Model 1
+## Poisson Model #1 & Negative Binomial Model #1
 ### Missing Data
@@
 
+## Poisson Model #2 & Multiple Linear Regrssion Model #1
+
+The right balance of the chemical properties of the wine is essential to make 
+it tastes, looks, and smells delightful, in addition to having a long shelf-life.
+Typically, citric acid and alcohol are the main positive attributes a wine 
+should have, while volatile acidity and sulphates are the main negative ones. 
+However, the dataset has negative values instead of near-zero in some of these 
+variables describing the chemical properties of the wine. Because little is given 
+about whether or not these variables were normalized, absolute values will be 
+taken assuming there was an error made during data collection.
+
+Lastly, to handle missing data for `STARS`, they will be replaced by 0, while 
+the chemical properties will be imputed. Because the predictive mean matching 
+imputation method combines the standard linear regression and the nearest-neighbor 
+imputation approach, Model Set #2 elected to fill in missing values of a 
+continuous variable by using it. Afterward, the data is pre-processed to fulfill 
+the assumption of normality using the Yeo-Johnson transformation 
+[(Yeo & Johnson, 2000)](https://www.jstor.org/stable/2673623?seq=1). This technique 
+attempts to find the value of lambda that minimizes the Kullback-Leibler distance 
+between the normal distribution and the transformed distribution. This method 
+has the advantage of working without having to worry about the domain of x.
+
+```{r m2clean}
+set.seed(525)
+m2train <- copy(trnSet[,-1])
+m2train[,c(2:8,11:12)] <- abs(m2train[,c(2:8,11:12)])
+m2train$STARS[is.na(m2train$STARS)] <- 0
+m2train.impute <- mice(m2train, method = 'pmm', print = FALSE)
+m2train <- complete(m2train.impute)
+m2train.r <- m2train$TARGET
+m2train.p <- m2train[,-1]
+pre.process <- preProcess(m2train.p, method = c("YeoJohnson"))
+m2train.p <- predict(pre.process, m2train.p)
+m2train[,c(2:13,15)] <- m2train.p[,c(1:12,14)]
+```
+
+## Multiple Linear Regression #2 & Negative Binomial Model #2
+### Missing Data
+These models will deal with missing numerical data by imputing the NA's
+using the Multivariate imputation by chained equations (MICE) method. Multiple
+imputation involves creating multiple predictions for each missing value, helping 
+to account for the uncertainty in the individual imputations.
+
+```{r model3impute}
+set.seed(123)
+# ML2 & NB2 data transformation
+# MICE imputation on train and test set
+m3train <- complete(mice(data = trnSet %>% select(-INDEX),
+                         method = "pmm", print = FALSE), 3)
+m3test <- complete(mice(data = tstSet %>% select(-INDEX),
+                        method = "pmm", print = FALSE), 3)
+```
+
 # BUILD MODELS
-## Poisson Model 1 & Negative Binomial Model 1
+## Poisson Model #1 & Negative Binomial Model #1
 The same person is building these two models, and the model setups will be
@@
   * `FreeSulfurDioxide` and `TotalSulfurDioxide`
-    * Again, a case of multiple theoretically related predictors
+    * Again, a case of multiple theoretically related predictors.
   * `LabelAppeal` with `Rated` and `LabelAppeal` with `Rated:STARS`
@@
 
-## Poisson Model #2
+## Poisson Model #2 & Multiple Linear Regrssion Model #1
+The `TARGET` variable has 2734 observations with zero values. This could be a result 
+of the manufacturer not releasing the wine into the market, or that there were no 
+orders placed. Therefore, backward stepwise variable elimination for a zero-inflated 
+poisson (ZIP) analysis will be investigated to understand the conditions. This model is
+further compared to the same Poisson model to determine whether there is an improvement.
 
-## Negative Binomial Model #2
+In the ZIP model, the independent variable, $TARGET_i$, take zero values $TARGET_i \sim 0$ 
+with the probability $\omega_i$ or values from Poisson distribution 
+$TARGET_i \sim Pois(\lambda_i)$ with probability (1 - $\omega_i$). For $i = 1,..., n$, 
+it can be written as:
 
-## Multiple Linear Regrssion Model # 1
+\[P(Y_i = y_i) = \left\{\begin{matrix}\omega _i + (1-\omega_i)e^{-\lambda_i}), \text{ if } y_i = 0\\ (1-\omega_i)\frac{e^{-\lambda_i}\lambda_i^{y_i}}{y_i!},\text{ if } y_i > 0 \end{matrix}\right.\]
 
-## Multiple Linear Regrssion Model # 2
+Therefore, the zero-inflated poisson model have two parameters, $\lambda_i$ and 
+$\omega_i$, and linked with the predictor variables with the following:
+
+\[log\left(\frac{\omega_i}{1-\omega_i}\right) = \sum_{j=1}^{t}\gamma_{ji}Z_{ji},\]
+
+\[log\left(\lambda_i\right) = \sum_{j=1}^{k}\beta_{ji}X_{ji},\]
+
+where $Z_1,..., Z_l$ are the dependent variables for the first equation and 
+$X_1,..., X_k$ for the second one. The expected value and variance of the number 
+of cases purchased for the $i^{th}$ wine in the ZIP model are, respectively:
+
+\[E(Y_i) = \lambda_i(1 - \omega_i),\]
+
+\[Var(Y_i) = (1 - \omega_i)(\lambda_i - \omega_i\lambda_i^2).\]
+
+Similar to Poisson regression, the ZIP model assumes that the average number of 
+cases purchased equals the variance. For more information, refer to 
+[Zero-Inflated Poisson Regression](https://math.usu.edu/jrstevens/biostat/projects2013/rep_ZIP.pdf).
+
+### Poisson Model #2
+
+```{r m2poisson}
+set.seed(525)
+# Zero Inflated Poisson Model
+m2.pmodel <- zeroinfl(TARGET~., 
+                      dist = 'poisson', 
+                      model = TRUE, 
+                      data = m2train,
+                      link = "logit")
+# Backward Elimination
+m2.fpmodel <- be.zeroinfl(m2.pmodel, 
+                          data = m2train, 
+                          dist = 'poisson', 
+                          alpha = 0.05, 
+                          trace = FALSE)
+m2.poisson.aic <- AIC(m2.fpmodel)
+```
+
+Zero-inflated count models are two-component mixture models combining a point 
+mass at zero with a proper count distribution. Thus, there are two sources of 
+zeros: zeros may come from both the point mass and from the count component.
+
+\small
+```{r m2PTable}
+kable(summary(m2.fpmodel)$coefficients$count,
+      caption = "Zero Inflated Poisson Model #2 Count Output")
+kable(summary(m2.fpmodel)$coefficients$zero,
+      caption = "Zero Inflated Poisson Model #2 Zero Output")
+```
+\normalsize
+
+#### Model Checking
+
+```{r m2PFitCheck}
+# Dispersion statistic
+E2 <- resid(m2.fpmodel, type = "pearson")
+N  <- nrow(m2train)
+p  <- length(coef(m2.fpmodel))  
+dp <- sum(E2^2) / (N - p)
+
+# Poisson Model #2, same formula from ZIP model
+m2.poisson <- glm(TARGET ~ Alcohol + LabelAppeal + AcidIndex + STARS | VolatileAcidity + 
+                    CitricAcid + FreeSulfurDioxide + TotalSulfurDioxide + pH + 
+                    Sulphates + Alcohol + LabelAppeal + AcidIndex + STARS, 
+                  family = poisson(link = 'log'), data = m2train)
+vuong(m2.poisson, m2.fpmodel)
+```
+
+This model has a dispersion parameter of `r prettyNum(dp, digits = 3L)` which 
+suggests under-dispersion (i.e., variance < mean). This parameter indicates how 
+many times larger the variance is than the mean. Since the dispersion was less 
+than one, the conditional variance is smaller than the conditional mean. Models 
+that ignore under-dispersion will be overly conservative and may fail to detect 
+significant patterns. Generalized Poisson (GP) and Conway‐Maxwell‐Poisson (CMP)
+distributions are better choices for such modeling because they can handle both 
+over-dispersion and under-dispersion. But for the scope of this assignment, 
+it's apparent, from the Vuong test with the same Poisson model and the ZIP model, 
+that the zero-inflated Poisson model is a slight improvement over the Poisson 
+model. Therefore, the ZIP model will still be considered among others.
+
+#### Coefficient Discussion
+All of the predictors in both the count and inflation portions of the model are 
+statistically significant. This model fits the data significantly better than the 
+null model, i.e., the intercept-only model. The count model contains information 
+about variables that the parent Poisson model used to estimate `TARGET` on the 
+condition that `TARGET` > 0. The coefficients for all 4 regression variables are 
+statistically significant at a 99% confidence level. The coefficient for `AcidIndex`
+is negative, meaning that as the total acidity of wine goes up by a unit, the number 
+of purchased cases goes down by 2%. One the other hand, for the other variables, i.e. 
+`Alcohol`, `LabelAppeal` and `STARS`, as these increase, the number of purchased 
+cases will also increase. In particular, `LabelAppeal` can account for a 26% 
+increase in the number of wine cases sold if the visual appeal of the wine 
+scores increase by one unit and all other variables held constant. Whereas, there
+is a 19% increase in cases sold for each increase in the wine rating by a team of 
+experts, `STARS`.
+
+The zero model contains information about variables that the nested Logistic 
+Regression model has been used to estimate the probability of whether or not cases are
+purchased. Examining the variables with a negative regression coefficient, this
+suggests that as the factor for the specific chemical properties increases, the
+probability of no case being purchase decreases. This is in line with our intuition,
+particularly for the sulfur dioxide levels. When a wine has an unusually small 
+percentage of the Free and Total $SO_2$, this suggests the wine is chemically and/or 
+microbially unstable. Thus, if a wine manufacturer were to increase the sulfur 
+dioxide levels by one unit, the odds that the wine would certainly not be purchased 
+would decrease by a factor of 0.92 to 0.95. However, `LabelAppeal` is among the
+positive coefficients for the zero model. This is interpreted as the higher a 
+rating is for a bottle of wine, the more likely there will be no case purchase, and it
+doesn't seem reasonable.
+
+### Multiple Linear Regrssion Model #1
+
+Multiple Linear Regression Model #1 will be using the step function has an options 
+to use a process that both adds and removes variables in the model. It uses AIC 
+(Akaike information criterion) as a selection criterion. A full model and a null 
+are defined, and then the function will follow a procedure to find the model with 
+the lowest AIC.
+
+```{r m1MLRFit}
+# Multiple Linear Regression model 
+set.seed(525)
+model.null = lm(TARGET ~ 1, data = m2train)
+model.full = lm(TARGET ~ ., data = m2train)
+
+# Step model used to derive final model, m1MLRFit  
+# step(model.null,
+#      scope = list(upper = model.full),
+#      direction = "both",
+#      data = m2train, 
+#      trace = FALSE)
+
+m1MLRFit = lm(TARGET ~ STARS + LabelAppeal + AcidIndex + TotalSulfurDioxide +
+                VolatileAcidity + FreeSulfurDioxide + Alcohol + CitricAcid + 
+                Chlorides + pH + Sulphates,
+              data = m2train)
+```
+
+\small
+```{r m1MLRTable}
+kable(summary(m1MLRFit)$coefficient,
+      caption = "Multiple Linear Regression Model #1 Output")
+```
+\normalsize
+
+#### Coefficient Discussion
+Contributing coefficients are those which lie within the 95% level of
+significance. A positive coefficient indicates that as the value of the
+predictor increases, the response variable also tends to increase. A negative
+coefficient suggests that as the predictor increases, the response variable
+tends to decrease.
+
+The intercept itself suggests that with no other information about the wine,
+there will be cases sold. The other positive predictors suggest that increasing
+`STARS` and `LabelAppeal` will result in an increase in the number of case purchased. 
+These make sense intuitively. Moreover, as discussed in the Poisson Model #2, the 
+right balance of $SO_2$ lead to better wine, thus the coefficients are rather
+small and have a smaller effect when it is increased. This reasoning is similar 
+for `Alcohol` levels. Next, `CitricAcid`, commonly used as an acid supplement 
+during fermentation, helps winemakers boost the acidity of their wine and 
+stabilize ferric haze. So, it makes sense that the higher `CitricAcid` is, the
+more cases would be sold. But once again, these need to be balance right.
+
+On the contrary, an increase in a unit of volatile acidity results in the decrease
+of case purchases. Volatile acidity is caused by bacteria or certain yeasts. This also
+can increase the acid index and pH if it is high. This is why sulfur dioxide is used 
+to carefully control it. In all, this model should not be taken as fixed, when it 
+comes to chemical processes, the wine manufacturer needs to take care of what to
+increase and decrease in order to reach equilibrium.
+
+## Negative Binomial Model #2 & Multiple Linear Regrssion Model #2
+
+### Negative Binomial Model #2
+#### Baseline Model 
+We will start with a simple negative binomial regression model to serve as a 
+baseline. This includes all variables in the dataset.
+
+```{r m2NBTrainBase}
+m2NBBase <- glm.nb(TARGET ~ ., data = m3train)
+knitr::kable(summary(m2NBBase)$coefficients, digits = 3L,
+             caption = 'Negative Binomial Regression Model #2 - Base Output')
+```
+Immediately, we can see a few variables that do not meet the 0.05
+p-value threshold.
+
+#### Enhanced Model
+Backwards stepwise regression is performed and the result is a model with the 
+following variables *removed*: `FixedAcidity`, `CitricAcid`, `ResidualSugar`, 
+and `Density`. Because the `Sulphates` variable is very close to the 0.05
+p-value threshold of significance, we will remove it as well.
+
+```{r m2NBTrain}
+m2NBFit <- glm.nb(TARGET ~ VolatileAcidity + Chlorides +
+                FreeSulfurDioxide + TotalSulfurDioxide + pH + Alcohol + 
+                LabelAppeal + AcidIndex + STARS, data = m3train)
+knitr::kable(summary(m2NBFit)$coefficients, digits = 3L,
+             caption = 'Negative Binomial Regression Model #2 - Final Output')
+```
+
+#### Coefficient Discussion
+The intercept of the final model tells us that there is typically ~1 case 
+of wine purchased. Some things to note:
+
+* Stars and label appeal contribute the most positively to the number of
+cases of wine purchased. Intuitively, this makes sense - the higher the
+rating and the prettier the bottle, the more cases we would expect to be
+sold. 
+* Free Sulfur Dioxide and Total Sulfur Dioxide have a very tiny effect on
+the number of cases sold (rounded to 3 decimal places, the impact is 0!)
+* All of the remaining variables except alcohol have a negative effect on 
+the number of cases purchased.
+
+### Multiple Linear Regrssion Model # 2
+One drawback to traditional multiple linear regression models is that the 
+forecasted counts are not integers. We will deal with this by rounding the
+predictions to the nearest whole number.
+
+#### Baseline Model
+We will start with a simple linear model to serve as a baseline. This includes
+all variables in the dataset.
+
+```{r m2MLTrainBase}
+m2MLBase <- lm(TARGET ~ ., data = m3train)
+knitr::kable(summary(m2MLBase)$coefficients, digits = 3L,
+             caption = 'Multiple Regression Model #2 - Base Output')
+```
+Most of the variables in the dataset are significant, but we can immediately
+see a few that are above the 0.05 p-value threshold.
+
+#### Enhanced Model
+Backwards stepwise regression is performed and the result is a model with the 
+following variables *removed*: `FixedAcidity`, `CitricAcid`, `ResidualSugar`, 
+and `Density`.
+
+```{r m2MLTrain}
+m2MLFit <- lm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide 
+              + TotalSulfurDioxide + pH + Sulphates + Alcohol + LabelAppeal 
+              + AcidIndex + STARS, data = m3train)
+knitr::kable(summary(m2MLFit)$coefficients, digits = 3L,
+             caption = 'Multiple Regression Model #2 - Final Output')
+```
+
+#### Coefficient Discussion
+The intercept of the final model tells us that there are typically ~3 cases 
+of wine purchased. Some things to note:
+
+* As expected, the number of stars has the largest, positive impact on the 
+number of cases sold. This makes sense -- the higher the rating of a wine,
+the more of it we would expect to be purchased. 
+* Label appeal has the second largest impact on the number of cases sold. 
+They say one shouldn't judge a book by its cover (or in this case, a wine
+by its label), but it does appear to have quite a significant impact on
+sales. 
+* Alcohol also has a small positive impact on the number of cases purchased. 
+This means that the more alcohol a wine has, the more cases of it will be
+purchased. 
+* Acid Index, Sulphates, pH, Chlorides, and VolatileAcidity all have a
+negative impact on the number of cases sold.
 
@@
 
-**Poisson Model 1** has an RMSE of `r prettyNum(m1PRMSE, digits = 2L)`, an MAE
-of `r prettyNum(m1PMAE, digits = 2L)`, and a training AIC of
+**Poisson Model 1** has an RMSE of 
+`r prettyNum(m1PRMSE, digits = 2L)`, an MAE of 
+`r prettyNum(m1PMAE, digits = 2L)`, and a training AIC of
 `r prettyNum(m1PFitAIC, digits = 2L)`.
@@
 
+```{r m2TestProcess}
+set.seed(525)
+m2test <- copy(tstSet[,-1])
+m2test[,c(2:8,11:12)] <- abs(m2test[,c(2:8,11:12)])
+m2test$STARS[is.na(m2test$STARS)] <- 0
+m2test.impute <- mice(m2test, method = 'pmm', print = FALSE)
+m2test <- complete(m2test.impute)
+m2test.r <- m2test$TARGET
+m2test.p <- m2test[,-1]
+pre.process <- preProcess(m2test.p, method = c("YeoJohnson"))
+m2test.p <- predict(pre.process, m2test.p)
+m2test[,c(2:13,15)] <- m2test.p[,c(1:12,14)]
+```
+```{r m2PTest}
+# ZIP2: Test Model
+m2PtstPred = round(predict(m2.fpmodel, newdata = m2test))
+compTable[2, 2] <- m2PRMSE <- RMSE(m2PtstPred, m2test.r)
+compTable[2, 3] <- m2PMAE <- MAE(m2PtstPred, m2test.r)
+compTable[2, 4] <- m2.poisson.aic
+```
+
+**Poisson Model 2** has an RMSE of 
+`r prettyNum(m2PRMSE, digits = 2L)`, an MAE of 
+`r prettyNum(m2PMAE, digits = 2L)`, and a training AIC of 
+`r prettyNum(m2.poisson.aic, digits = 2L)`.
+
 ### Negative Binomial Model 1
+
 ```{r m1NBTest}
@@
 
+```{r m2NBTest}
+# NB2: Test Model
+m2NBtstPred <- round(predict(m2NBFit, newdata = m3test))
+compTable[4, 2] <- m2NBRMSE <- RMSE(m2NBtstPred, m3test$TARGET)
+compTable[4, 3] <- m2NBMAE <- MAE(m2NBtstPred, m3test$TARGET)
+compTable[4, 4] <- m2NBFitAIC <- AIC(m2NBFit)
+```
+
+**Negative Binomial Model 2** has an RMSE of
+`r prettyNum(m2NBRMSE, digits = 2L)`, an MAE of
+`r prettyNum(m2NBMAE, digits = 2L)`, and a training AIC of
+`r prettyNum(m2NBFitAIC, digits = 2L)`.
+
 ### Multiple Linear Regression Model 1
 
+```{r m1MLRTest}
+# MLR1: Test Model
+m1MLRtstPred <- round(predict(m1MLRFit, newdata = m2test))
+compTable[5, 2] <- m1MLRRMSE <- RMSE(m1MLRtstPred, m2test.r)
+compTable[5, 3] <- m1MLRMAE <- MAE(m1MLRtstPred, m2test.r)
+compTable[5, 4] <- m1MLRFitAIC <- AIC(m1MLRFit)
+```
+
+**Multiple Linear Regression Model 2** has an RMSE of
+`r prettyNum(m1MLRRMSE, digits = 2L)`, an MAE of
+`r prettyNum(m1MLRMAE, digits = 2L)`, and a training AIC of
+`r prettyNum(m1MLRFitAIC, digits = 2L)`.
+
 ### Multiple Linear Regression Model 2
+
+```{r m2MLTest}
+# MLR2: Test Model
+m2MLtstPred <- round(predict(m2MLFit, newdata = m3test))
+compTable[6, 2] <- m2MLMSE <- RMSE(m2MLtstPred, m3test$TARGET)
+compTable[6, 3] <- m2MLMAE <- MAE(m2MLtstPred, m3test$TARGET)
+compTable[6, 4] <- m2MLFitAIC <- AIC(m2MLFit)
+```
+
+**Multiple Linear Regression Model 1** has an RMSE of
+`r prettyNum(m2MLMSE, digits = 2L)`, an MAE of
+`r prettyNum(m2MLMAE, digits = 2L)`, and a training AIC of
+`r prettyNum(m2MLFitAIC, digits = 2L)`.
 
@@
 ```
+```{r m2clean}
+```
+```{r model3impute}
+```
 <!-- Build Models -->
@@
 ```
-<!-- Other prep and build goes here -->
+```{r m2poisson}
+```
+```{r m2PTable}
+```
+```{r m2PFitCheck}
+```
+```{r m1MLRFit}
+```
+```{r m1MLRTable}
+```
+```{r m2NBTrainBase}
+```
+```{r m2NBTrain}
+```
+```{r m2MLTrainBase}
+```
+```{r m2MLTrain}
+```
 <!-- Select Models -->
@@
 ```
-<!-- Poisson 2 Test goes here -->
+```{r m2TestProcess}
+```
+```{r m2PTest}
+```
 ```{r m1NBTest}
+```
+```{r m2NBTest}
+```
+```{r m1MLRTest}
+```
+```{r m2MLTest}
 ```

### 2020-11-25 Final_project/Draft_EDA_start.Rmd
Adding to the EDA

@@
 library(GGally)
+
+set.seed(951)
 ```
@@
 # Data
-•Description of Data
-–A complete description of the desired output
 
@@
 
-~discuss using Excel for data cleaning for unique values in France.
+The owner of the dataset seemed to combine the five countries in question into the 
+France dataset. Unfortunately, when this operation occurred, the owner left no marker 
+in the dataset to identify which country the data came from originally.  As we are 
+going to use the country as predictor, we needed to find a method to remove the unwanted 
+values.
+
+Using Excel, we performed a simple VLOOKUP on the video ID from the France dataset to 
+the other four datasets.  If the Video ID was in another dataset, we flagged and removed 
+it.  The resulting shortened dataset matched in item count to the other datasets.  As 
+the datasets are static and we do not plan to have then constantly updated with fresh 
+data, using the quickest simplest method is appropriate.
+
+As each dataset is over 30,000 individual objects, incomplete cases were dropped completely. 
+Out of the 40,881 objects in the raw Canada dataset, removing the incomplete cases only 
+removed 1,296 (roughly 3.2%).  France had the highest number removed for incomplete cases 
+at 8.1% and US being the lowest at 1.4%.  France being the highest is understandable due 
+to the prior state of the dataset discussed earlier.
+
+Summary statistic markers were calculated for each dataset and for the environment as a 
+whole.
+
+  * A country code (1-5) was applied to each dataset.  
+  * Target values of view_target and view_target_global were created 
+    * A value of '1' designates the view count is above the mean of the country 
+      and global respectively. 
+    * A value of '0' designates the view count falls below the mean respectively 
+  * 'tag_count', A predictor value, was calculated by counting the number of tags applied 
+    to each video entry.
 
 ```{r ca_data_prep}
+# remove incomplete cases
 ca_set <- ca_set %>% filter(complete.cases(ca_set))
 
+#Calculating the mean of the CA dataset
 ca_mean <- mean(ca_set$views)
@@
 ca_set <- ca_set%>%
-  mutate(country = 'CA')%>%
-  mutate(country_code = 1)%>%
-  mutate(view_target = ifelse(views > ca_mean,1,0))%>%
-  mutate(tag_count = 1+(str_count(ca_set$tags, ';')))
+  mutate(country = 'CA')%>% #Add country label
+  mutate(country_code = 1)%>% #Add country code
+  mutate(view_target = ifelse(views > ca_mean,1,0))%>% #calculate above/below view mean
+  mutate(tag_count = 1+(str_count(ca_set$tags, ';'))) #count number of tags
 ```
@@
 ```{r merge_datasets}
+# Prep column names for joins
 mergeCols = c('video_id','trending_date','title','channel_title','category_id','publish_time','tags','views','likes','dislikes','comment_count','thumbnail_link','comments_disabled','ratings_disabled','video_error_or_removed','description','country','country_code', 'view_target','tag_count')
 
+# Join the datasets into a single dataset
 full_set <- ca_set %>% full_join(de_set, by=mergeCols)
@@
 
+# Calculate Global mean
 full_mean <- mean(full_set$views)
 
+#calculate above/below global mean
 full_set <- full_set %>%
@@
 ```{r global_target}
+# add global mean calculation to individual datasets
 ca_set <- ca_set %>%
@@
 •Describe the scope and limitations of the methodology
+
+```{r}
+#potential train/test split
+ca_set_model <- ca_set %>%
+  select(-title, -tags, -channel_title, -trending_date, -publish_time, -video_id, -thumbnail_link, -description, -country)
+
+ca_smp_size <- floor(0.75*nrow(ca_set))
+ca_ind <- sample(seq_len(nrow(ca_set)), size = ca_smp_size)
+
+ca_train <- ca_set[ca_ind,]
+ca_test <- ca_set[-ca_ind,]
+```
+
 ## Model Creation
+• describe which model are created and why
 
@@
 ### Criteria
+• criteria of the model selection
+
 ### Performance
+• describe the performance of the selection
 
 ### Predictions
+• describe the prediction values
 
 # Results
-– Needs to systematically and clearly articulate the study findings.
+•Needs to systematically and clearly articulate the study findings.
 If the results are unclear, the reviewer must decide whether the analysis of the data was poorly executed or whether the Results section is poorly organized.
 
-– Should state wheather the hypothesis were verified or proven untrue or, if no hypotheses were given, weather the research questions were answered.
+• Should state wheather the hypothesis were verified or proven untrue or, if no hypotheses were given, weather the research questions were answered.
 The authors should also comment on their results in light of previous studies and explain what differences (if any) exist between the findings and those reported by others and attempt to provide an explaination of the differnces.

### 2020-11-25 Final_project/Draft_EDA_start.Rmd
EDA cleanup

@@
 * Data 
-  * The data sets have an expected heteroscedasticity.  To work around this, we structured the data using the XXXXX method.  
-  * After preparing the data, we created randomized subsets of the datasets in a train/test structure for use in the models.
-  * Distribution and Correlation plots are used to further the model selection process
+  * The data sets have an expected heteroscedasticity.  To work around this, 
+    we structured the data using the XXXXX method.  
+  * After preparing the data, we created randomized subsets of the datasets in 
+    a train/test structure for use in the models.
+  * Distribution and Correlation plots are used to further the model selection 
+    process
 * Models
-  * We generated X models using methods such as XXXXXX in order to find and chose the most appropriate and accurate model. 
+  * We generated X models using methods such as XXXXXX in order to find and 
+    chose the most appropriate and accurate model. 
 * Results
@@
 * Summary and Conclusions
-  * Quick recap of the findings, identify areas of further research and shortcomings found during the research.
+  * Quick recap of the findings, identify areas of further research and shortcomings 
+    found during the research.
 * References
@@
 
-```{r summary_us}
-summaryus <- us_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
-describe(summaryus)
-
-ggpairs(
-  summaryus,
-  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
-  )
-us_corr <- cor(summaryus)
-corrplot(us_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
-
-```
-
-```{r summary_us}
+```{r summary_gb}
 summarygb <- gb_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
@@
 
+```{r summary_us}
+summaryus <- us_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
+describe(summaryus)
+
+ggpairs(
+  summaryus,
+  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
+  )
+us_corr <- cor(summaryus)
+corrplot(us_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
+
+```
+
 ```{r summary_global}
@@
 
-```{r}
+```{r ca_train_split}
 #potential train/test split
@@
 ```
+
+```{r libraries}
+```
+```{r load_data}
+```
+```{r ca_data_prep}
+```
+```{r de_data_prep}
+```
+```{r fr_data_prep}
+```
+```{r gb_data_prep}
+```
+```{r us_data_prep}
+```
+```{r merge_datasets}
+```
+```{r global_target}
+```
+```{r summary_ca}
+```
+```{r summary_de}
+```
+```{r summary_gb}
+```
+```{r summary_fr}
+```
+```{r summary_us}
+```
+```{r summary_global}
+```
+```{r ca_train_split}
+```

### 2020-11-26 Final_project/Draft_EDA_start.Rmd
Update to Header information

@@
 title: "DATA 621 - Business Analytics and Data Mining"
-subtitle: 'Fall 2020 - Group 2 - Homework #3'
 author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg, 
   Bryan Persaud, Jeff Shamp
-date: "10/28/2020"
+date: "12/14/2020"
 output:
+  html_document:
+    df_print: paged
   pdf_document: default
+subtitle: 'Fall 2020 - Group 2 - Final Project'
 urlcolor: purple
@@
 
-```{r load_data}
+```{r load_data, message=FALSE}
 ca_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/CAvideos.csv"
@@
 
-```{r summary_ca}
+```{r summary_ca, message=FALSE}
 summaryca <- ca_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
@@
 summaryde <- de_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled, view_target, view_global_target, tag_count)
-describe(summaryus)
+describe(summaryde)
 
@@
 #potential train/test split
-ca_set_model <- ca_set %>%
-  select(-title, -tags, -channel_title, -trending_date, -publish_time, -video_id, -thumbnail_link, -description, -country)
 
-ca_smp_size <- floor(0.75*nrow(ca_set))
-ca_ind <- sample(seq_len(nrow(ca_set)), size = ca_smp_size)
+ca_smp_size <- floor(0.75*nrow(summaryca))
+ca_ind <- sample(seq_len(nrow(summaryca)), size = ca_smp_size)
 
-ca_train <- ca_set[ca_ind,]
-ca_test <- ca_set[-ca_ind,]
+ca_train <- summaryca[ca_ind,]
+ca_test <- summaryca[-ca_ind,]
 ```

### 2020-11-26 HW5/HW5-Draft.Rmd
1) Move data.table after dplyr to minimize conflicts between packages 2) Copyediting 3) Move Yeo-Johnson citation to REFERENCES

@@
 library(corrplot)
-library(data.table)
 library(mice)
@@
 library(dplyr)
+library(data.table)
 
@@
 ```{r graphsD, fig.height=5, fig.width=7}
-# Using epanechnikov kernel to generate kernel-smoothed densities
+# Using Epanechnikov kernel to generate kernel-smoothed densities
 ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnDUB], aes(x = value)) +
@@
 
-## Poisson Model #2 & Multiple Linear Regrssion Model #1
-
-The right balance of the chemical properties of the wine is essential to make 
-it tastes, looks, and smells delightful, in addition to having a long shelf-life.
+## Poisson Model #2 & Multiple Linear Regression Model #1
+
+The right balance of the chemical properties of the wine is essential to make it
+taste, look, and smell delightful, in addition to having a long shelf-life.
 Typically, citric acid and alcohol are the main positive attributes a wine
@@
 However, the dataset has negative values instead of near-zero in some of these 
-variables describing the chemical properties of the wine. Because little is given 
-about whether or not these variables were normalized, absolute values will be 
-taken assuming there was an error made during data collection.
+variables describing the chemical properties of the wine. Because little
+information is given whether or not these variables were normalized, absolute
+values will be taken assuming there was an error made during data collection.
 
@@
 the chemical properties will be imputed. Because the predictive mean matching 
-imputation method combines the standard linear regression and the nearest-neighbor 
-imputation approach, Model Set #2 elected to fill in missing values of a 
-continuous variable by using it. Afterward, the data is pre-processed to fulfill 
-the assumption of normality using the Yeo-Johnson transformation 
-[(Yeo & Johnson, 2000)](https://www.jstor.org/stable/2673623?seq=1). This technique 
-attempts to find the value of lambda that minimizes the Kullback-Leibler distance 
-between the normal distribution and the transformed distribution. This method 
-has the advantage of working without having to worry about the domain of x.
+imputation method combines the standard linear regression and the
+nearest-neighbor imputation approach, Model Set #2 elected to fill in the
+missing values of a continuous variable by using mean-matching imputation.
+Afterward, the data is pre-processed to fulfill the assumption of normality
+using the Yeo-Johnson transformation (2000). This technique attempts to find the
+value of lambda that minimizes the Kullback-Leibler distance between the normal
+distribution and the transformed distribution. This method has the advantage of
+working without having to worry about the domain of x.
 
@@
   https://topepo.github.io/caret/pre-processing.html#imputation
+  * Yeo, I., & Johnson, R. (2000). A New Family of Power Transformations to
+  Improve Normality or Symmetry. **Biometrika**, 87(4), 954--959.
+  Retrieved November 26, 2020, from http://www.jstor.org/stable/2673623
 

### 2020-11-26 HW5/HW5-Draft.Rmd
I believe this was the only place where dplyr is used. Using base in this one place will prevent package conflicts.

@@
 library(mpath)
-library(dplyr)
 library(data.table)
@@
 distribution and the transformed distribution. This method has the advantage of
-working without having to worry about the domain of x.
+working without having to worry about the domain of \(x\).
 
 ```{r m2clean}
+# P2 & ML 1: Clean and impute
 set.seed(525)
@@
 ```{r model3impute}
-set.seed(123)
 # ML2 & NB2 data transformation
 # MICE imputation on train and test set
-m3train <- complete(mice(data = trnSet %>% select(-INDEX),
+set.seed(123)
+m3train <- complete(mice(data = trnSet[, -c('INDEX')],
                          method = "pmm", print = FALSE), 3)
-m3test <- complete(mice(data = tstSet %>% select(-INDEX),
+m3test <- complete(mice(data = tstSet[, -c('INDEX')],
                          method = "pmm", print = FALSE), 3)

### 2020-11-26 HW5/HW5-Draft.Rmd
Great use of LaTeX. Some minor cleanup like using cases enviornment and using the LaTeX delimiters and not the TeX delimiters

@@
 ## Poisson Model #2 & Multiple Linear Regrssion Model #1
-The `TARGET` variable has 2734 observations with zero values. This could be a result 
-of the manufacturer not releasing the wine into the market, or that there were no 
-orders placed. Therefore, backward stepwise variable elimination for a zero-inflated 
-poisson (ZIP) analysis will be investigated to understand the conditions. This model is
-further compared to the same Poisson model to determine whether there is an improvement.
-
-In the ZIP model, the independent variable, $TARGET_i$, take zero values $TARGET_i \sim 0$ 
-with the probability $\omega_i$ or values from Poisson distribution 
-$TARGET_i \sim Pois(\lambda_i)$ with probability (1 - $\omega_i$). For $i = 1,..., n$, 
-it can be written as:
-
-\[P(Y_i = y_i) = \left\{\begin{matrix}\omega _i + (1-\omega_i)e^{-\lambda_i}), \text{ if } y_i = 0\\ (1-\omega_i)\frac{e^{-\lambda_i}\lambda_i^{y_i}}{y_i!},\text{ if } y_i > 0 \end{matrix}\right.\]
-
-Therefore, the zero-inflated poisson model have two parameters, $\lambda_i$ and 
-$\omega_i$, and linked with the predictor variables with the following:
-
-\[log\left(\frac{\omega_i}{1-\omega_i}\right) = \sum_{j=1}^{t}\gamma_{ji}Z_{ji},\]
-
-\[log\left(\lambda_i\right) = \sum_{j=1}^{k}\beta_{ji}X_{ji},\]
-
-where $Z_1,..., Z_l$ are the dependent variables for the first equation and 
-$X_1,..., X_k$ for the second one. The expected value and variance of the number 
-of cases purchased for the $i^{th}$ wine in the ZIP model are, respectively:
-
-\[E(Y_i) = \lambda_i(1 - \omega_i),\]
-
-\[Var(Y_i) = (1 - \omega_i)(\lambda_i - \omega_i\lambda_i^2).\]
+The `TARGET` variable has 2734 observations with zero values. This could be a
+result of the manufacturer not releasing the wine into the market, or that there
+were no orders placed. Therefore, backward stepwise variable elimination for a
+zero-inflated poisson (ZIP) analysis will be investigated to understand the
+conditions. This model is further compared to the same Poisson model to
+determine whether there is an improvement.
+
+In the ZIP model, the independent variable, \(TARGET_i\), take zero values
+\(TARGET_i \sim 0\) with the probability \(\omega_i\) or values from Poisson
+distribution \(TARGET_i \sim \mathrm{Pois}(\lambda_i)\) with probability
+\((1 - \omega_i)\). For \(i = 1,\ldots, n\), it can be written as:
+
+\[
+P(Y_i = y_i) = 
+\begin{cases}
+\omega_i + (1-\omega_i)e^{-\lambda_i})&\text{ if } y_i = 0\\
+(1-\omega_i)\frac{e^{-\lambda_i}\lambda_i^{y_i}}{y_i!}&\text{ if } y_i > 0
+\end{cases}
+\]
+
+Therefore, the zero-inflated poisson model have two parameters, \(\lambda_i\)
+and \(\omega_i\), and linked with the predictor variables with the following:
+
+\[
+\begin{align}
+&\log\left(\frac{\omega_i}{1-\omega_i}\right) = \sum_{j=1}^{t}\gamma_{ji}Z_{ji}\\
+&\log\left(\lambda_i\right) = \sum_{j=1}^{k}\beta_{ji}X_{ji}
+\end{align}
+\]
+
+where \(Z_1,\ldots, Z_l\) are the dependent variables for the first equation and 
+\(X_1,\ldots, X_k\) for the second one. The expected value and variance of the
+number of cases purchased for the \(i^{\textrm{th}}\) wine in the ZIP model are
+respectively:
+
+\[
+\begin{align}
+\mathrm{E}(Y_i) &= \lambda_i(1 - \omega_i)\\
+\mathrm{Var}(Y_i) &= (1 - \omega_i)(\lambda_i - \omega_i\lambda_i^2)
+\end{align}
+\]
 

### 2020-11-26 HW5/HW5-Draft.Rmd
Minor copyediting and specifying amsmath LaTeX package

@@
     toc_depth: 4
+    extra_dependencies:
+      amsmath: null
+      inputenc: utf8
 urlcolor: purple
@@
 Typically, citric acid and alcohol are the main positive attributes a wine
-should have, while volatile acidity and sulphates are the main negative ones. 
+should have, while volatile acidity and sulfates are the main negative ones. 
 However, the dataset has negative values instead of near-zero in some of these 
@@
 ```{r m1PNB}
+# P1 & NB1: GoF Comparison
 m1PNBC <- data.table(Model = c("Poisson 1", "Negative Binomial 1"),
@@
 
-## Poisson Model #2 & Multiple Linear Regrssion Model #1
+## Poisson Model #2 & Multiple Linear Regression Model #1
 The `TARGET` variable has 2734 observations with zero values. This could be a
@@
 were no orders placed. Therefore, backward stepwise variable elimination for a
-zero-inflated poisson (ZIP) analysis will be investigated to understand the
+zero-inflated Poisson (ZIP) analysis will be investigated to understand the
 conditions. This model is further compared to the same Poisson model to
@@
 
-Therefore, the zero-inflated poisson model have two parameters, \(\lambda_i\)
+Therefore, the zero-inflated Poisson model have two parameters, \(\lambda_i\)
 and \(\omega_i\), and linked with the predictor variables with the following:
@@
 \[
-\begin{align}
+\begin{aligned}
 &\log\left(\frac{\omega_i}{1-\omega_i}\right) = \sum_{j=1}^{t}\gamma_{ji}Z_{ji}\\
 &\log\left(\lambda_i\right) = \sum_{j=1}^{k}\beta_{ji}X_{ji}
-\end{align}
+\end{aligned}
 \]
@@
 \[
-\begin{align}
+\begin{aligned}
 \mathrm{E}(Y_i) &= \lambda_i(1 - \omega_i)\\
 \mathrm{Var}(Y_i) &= (1 - \omega_i)(\lambda_i - \omega_i\lambda_i^2)
-\end{align}
+\end{aligned}
 \]
@@
 ```{r m2PTable}
+# P2: Outputs
 kable(summary(m2.fpmodel)$coefficients$count,
@@
 over-dispersion and under-dispersion. But for the scope of this assignment, 
-it's apparent, from the Vuong test with the same Poisson model and the ZIP model, 
-that the zero-inflated Poisson model is a slight improvement over the Poisson 
-model. Therefore, the ZIP model will still be considered among others.
+it's apparent, from the Vuong test with the same Poisson model and the ZIP
+model, that the zero-inflated Poisson model is a slight improvement over the
+Poisson model. Therefore, the ZIP model will still be considered among others.
 
@@
 All of the predictors in both the count and inflation portions of the model are 
-statistically significant. This model fits the data significantly better than the 
-null model, i.e., the intercept-only model. The count model contains information 
-about variables that the parent Poisson model used to estimate `TARGET` on the 
-condition that `TARGET` > 0. The coefficients for all 4 regression variables are 
-statistically significant at a 99% confidence level. The coefficient for `AcidIndex`
-is negative, meaning that as the total acidity of wine goes up by a unit, the number 
-of purchased cases goes down by 2%. One the other hand, for the other variables, i.e. 
-`Alcohol`, `LabelAppeal` and `STARS`, as these increase, the number of purchased 
-cases will also increase. In particular, `LabelAppeal` can account for a 26% 
-increase in the number of wine cases sold if the visual appeal of the wine 
-scores increase by one unit and all other variables held constant. Whereas, there
-is a 19% increase in cases sold for each increase in the wine rating by a team of 
-experts, `STARS`.
+statistically significant. This model fits the data significantly better than
+the null model, i.e., the intercept-only model. The count model contains
+information about variables that the parent Poisson model used to estimate
+`TARGET` on the condition that `TARGET` > 0. The coefficients for all 4
+regression variables are statistically significant at a 99% confidence level.
+The coefficient for `AcidIndex` is negative, meaning that as the total acidity
+of wine goes up by a unit, the number of purchased cases goes down by 2%. One
+the other hand, for the other variables, i.e. `Alcohol`, `LabelAppeal` and
+`STARS`, as these increase, the number of purchased cases will also increase. In
+particular, `LabelAppeal` can account for a 26% increase in the number of wine
+cases sold if the visual appeal of the wine scores increase by one unit and all
+other variables held constant. Whereas, there is a 19% increase in cases sold
+for each increase in the wine rating by a team of experts, `STARS`.
 
 The zero model contains information about variables that the nested Logistic 
-Regression model has been used to estimate the probability of whether or not cases are
-purchased. Examining the variables with a negative regression coefficient, this
-suggests that as the factor for the specific chemical properties increases, the
-probability of no case being purchase decreases. This is in line with our intuition,
-particularly for the sulfur dioxide levels. When a wine has an unusually small 
-percentage of the Free and Total $SO_2$, this suggests the wine is chemically and/or 
-microbially unstable. Thus, if a wine manufacturer were to increase the sulfur 
-dioxide levels by one unit, the odds that the wine would certainly not be purchased 
-would decrease by a factor of 0.92 to 0.95. However, `LabelAppeal` is among the
-positive coefficients for the zero model. This is interpreted as the higher a 
-rating is for a bottle of wine, the more likely there will be no case purchase, and it
-doesn't seem reasonable.
+Regression model has been used to estimate the probability of whether or not
+cases are purchased. Examining the variables with a negative regression
+coefficient, this suggests that as the factor for the specific chemical
+properties increases, the probability of no case being purchase decreases. This
+is in line with our intuition, particularly for the sulfur dioxide levels. When
+a wine has an unusually small percentage of the Free and Total \(SO_2\), this
+suggests the wine is chemically and/or microbially unstable. Thus, if a wine
+manufacturer were to increase the sulfur dioxide levels by one unit, the odds
+that the wine would certainly not be purchased would decrease by a factor of
+0.92 to 0.95. However, `LabelAppeal` is among the positive coefficients for the
+zero model. This is interpreted as the higher a rating is for a bottle of wine,
+the more likely there will be no case purchase, and it doesn't seem reasonable.
 
 ### Multiple Linear Regrssion Model #1
-
-Multiple Linear Regression Model #1 will be using the step function has an options 
-to use a process that both adds and removes variables in the model. It uses AIC 
-(Akaike information criterion) as a selection criterion. A full model and a null 
-are defined, and then the function will follow a procedure to find the model with 
-the lowest AIC.
+Multiple Linear Regression Model #1 will be using the step function has an
+option to use a process that both adds and removes variables in the model. It
+uses AIC (Akaike information criterion) as a selection criterion. A full model
+and a null are defined, and then the function will follow a procedure to find
+the model with the lowest AIC.
 
@@
 ```{r m1MLRTable}
+# ML1: Outputs
 kable(summary(m1MLRFit)$coefficient,
@@
 there will be cases sold. The other positive predictors suggest that increasing
-`STARS` and `LabelAppeal` will result in an increase in the number of case purchased. 
-These make sense intuitively. Moreover, as discussed in the Poisson Model #2, the 
-right balance of $SO_2$ lead to better wine, thus the coefficients are rather
-small and have a smaller effect when it is increased. This reasoning is similar 
-for `Alcohol` levels. Next, `CitricAcid`, commonly used as an acid supplement 
-during fermentation, helps winemakers boost the acidity of their wine and 
-stabilize ferric haze. So, it makes sense that the higher `CitricAcid` is, the
-more cases would be sold. But once again, these need to be balance right.
-
-On the contrary, an increase in a unit of volatile acidity results in the decrease
-of case purchases. Volatile acidity is caused by bacteria or certain yeasts. This also
-can increase the acid index and pH if it is high. This is why sulfur dioxide is used 
-to carefully control it. In all, this model should not be taken as fixed, when it 
-comes to chemical processes, the wine manufacturer needs to take care of what to
-increase and decrease in order to reach equilibrium.
+`STARS` and `LabelAppeal` will result in an increase in the number of case
+purchased. These make sense intuitively. Moreover, as discussed in the Poisson
+Model #2, the right balance of $SO_2$ lead to better wine, thus the coefficients
+are rather small and have a smaller effect when it is increased. This reasoning
+is similar for `Alcohol` levels. Next, `CitricAcid`, commonly used as an acid
+supplement during fermentation, helps winemakers boost the acidity of their wine
+and stabilize ferric haze. So, it makes sense that the higher `CitricAcid` is,
+the more cases would be sold. But once again, these need to be balance right.
+
+On the contrary, an increase in a unit of volatile acidity results in the
+decrease of case purchases. Volatile acidity is caused by bacteria or certain
+yeasts. This also can increase the acid index and pH if it is high. This is why
+sulfur dioxide is used to carefully control it. In all, this model should not be
+taken as fixed, when it comes to chemical processes, the wine manufacturer needs
+to take care of what to increase and decrease in order to reach equilibrium.
 
 ## Negative Binomial Model #2 & Multiple Linear Regrssion Model #2
-
 ### Negative Binomial Model #2
@@
 ```{r m2NBTrainBase}
+# NB2: Baseline Model
 m2NBBase <- glm.nb(TARGET ~ ., data = m3train)
@@
 ```
-Immediately, we can see a few variables that do not meet the 0.05
-p-value threshold.
+
+Immediately, we can see a few variables that do not meet the 0.05 p-value
+threshold.
 
@@
 ```{r m2NBTrain}
+# NB2: Revised Model
 m2NBFit <- glm.nb(TARGET ~ VolatileAcidity + Chlorides +
@@
 #### Coefficient Discussion
-The intercept of the final model tells us that there is typically ~1 case 
-of wine purchased. Some things to note:
-
-* Stars and label appeal contribute the most positively to the number of
-cases of wine purchased. Intuitively, this makes sense - the higher the
-rating and the prettier the bottle, the more cases we would expect to be
-sold. 
-* Free Sulfur Dioxide and Total Sulfur Dioxide have a very tiny effect on
-the number of cases sold (rounded to 3 decimal places, the impact is 0!)
-* All of the remaining variables except alcohol have a negative effect on 
-the number of cases purchased.
+The intercept of the final model tells us that there is typically ~1 case of
+wine purchased. Some things to note:
+
+  * Stars and label appeal contribute the most positively to the number of cases
+  of wine purchased. Intuitively, this makes sense - the higher the rating and
+  the prettier the bottle, the more cases we would expect to be sold.
+  * Free Sulfur Dioxide and Total Sulfur Dioxide have a very tiny effect on the
+  number of cases sold (rounded to 3 decimal places, the impact is 0!)
+  * All of the remaining variables except alcohol have a negative effect on the
+  number of cases purchased.
 
@@
 One drawback to traditional multiple linear regression models is that the 
-forecasted counts are not integers. We will deal with this by rounding the
+forecast counts are not integers. We will deal with this by rounding the
 predictions to the nearest whole number.
@@
 ```{r m2MLTrainBase}
+# ML2: Baseline model
 m2MLBase <- lm(TARGET ~ ., data = m3train)
@@
 ```
-Most of the variables in the dataset are significant, but we can immediately
-see a few that are above the 0.05 p-value threshold.
+
+Most of the variables in the dataset are significant, but we can immediately see
+a few that are above the 0.05 p-value threshold.
 
@@
 ```{r m2MLTrain}
+# ML2: Enhanced model
 m2MLFit <- lm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide 
@@
 #### Coefficient Discussion
-The intercept of the final model tells us that there are typically ~3 cases 
-of wine purchased. Some things to note:
+The intercept of the final model tells us that there are typically ~3 cases of
+wine purchased. Some things to note:
 
   * As expected, the number of stars has the largest, positive impact on the
-number of cases sold. This makes sense -- the higher the rating of a wine,
-the more of it we would expect to be purchased. 
-* Label appeal has the second largest impact on the number of cases sold. 
-They say one shouldn't judge a book by its cover (or in this case, a wine
-by its label), but it does appear to have quite a significant impact on
-sales. 
+  number of cases sold. This makes sense -- the higher the rating of a wine, the
+  more of it we would expect to be purchased.
+  * Label appeal has the second largest impact on the number of cases sold. They
+  say one shouldn't judge a book by its cover (or in this case, a wine by its
+  label), but it does appear to have quite a significant impact on sales. 
   * Alcohol also has a small positive impact on the number of cases purchased.

### 2020-11-26 HW5/HW5-Draft.Rmd
ZI-Poisson 2 is best

@@
 ## Selection
+The zero-inflated **Poisson Model 2** performed best on all the metrics and will
+be used to create the prediction on the evaluation set.
 
@@
 ```
-<!-- Remaining Test sets go here -->
 ```{r modelTestResults}

### 2020-11-26 HW5/HW5-Group2.Rmd
Create predictions, rename, and push PDF

--- 2020-11-26 HW5/HW5-Draft.Rmd
@@
 
+```{r readEval}
+# Read evaluation but not "IN" since that's dropped right away
+m2eval <- fread(paste0(urlRemote, pathGithub, fileEval), drop = 'IN')
+nobsE <- dim(DT)[[1]]
+```
+
+With `r nobsE` observations in the evaluation set, the predicted cases for each
+row will not be printed, rather some summary statistics and a histogram of the
+evaluations will be shown. The actual predictions are available in the
+`m2evalP` data object, as seen in the CODE APPENDIX.
+
+```{r processEval}
+# Process evaluation file as per zero-inflated Poisson model 2
+m2eval[,c(2:8,11:12)] <- abs(m2eval[,c(2:8,11:12)])
+m2eval$STARS[is.na(m2eval$STARS)] <- 0
+set.seed(525)
+m2eval.impute <- mice(m2eval, method = 'pmm', print = FALSE)
+m2eval <- complete(m2eval.impute)
+m2eval.r <- m2eval$TARGET
+m2eval.p <- m2eval[, -1]
+pre.process <- preProcess(m2eval.p, method = c("YeoJohnson"))
+m2eval.p <- predict(pre.process, m2eval.p)
+m2eval[, c(2:13, 15)] <- m2eval.p[, c(1:12, 14)]
+
+# Predict Cases
+m2evalP = as.data.frame(round(predict(m2.fpmodel, newdata = m2eval)))
+colnames(m2evalP) <- 'Cases'
+kable(table(m2evalP), caption = "Predicted Cases")
+ggplot(m2evalP, aes(x = Cases)) +
+  geom_histogram(binwidth = 1, center = 0,
+                 fill = 'firebrick4', color = 'black') +
+  scale_x_continuous(breaks = 0:8) +
+  ggtitle("Histogram of Predicted Cases")
+```
+
 # REFERENCES
@@
 ```
+<!-- Predictions -->
+```{r readEval}
+```
+```{r processEval}
+```

### 2020-11-29 HW5/HW5-Group2.Rmd
Minor correction

@@
 m2eval <- fread(paste0(urlRemote, pathGithub, fileEval), drop = 'IN')
-nobsE <- dim(DT)[[1]]
+nobsE <- dim(m2eval)[[1]]
 ```
@@
 # Process evaluation file as per zero-inflated Poisson model 2
-m2eval[,c(2:8,11:12)] <- abs(m2eval[,c(2:8,11:12)])
+m2eval[, c(2:8, 11:12)] <- abs(m2eval[, c(2:8, 11:12)])
 m2eval$STARS[is.na(m2eval$STARS)] <- 0

### 2020-12-01 Final_project/Final Project Paper - Group 2.Rmd
Revised paper due to Nasrin's very specific requirements. New document.

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Final Project Paper'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "12/14/2020"
output:
  pdf_document:
    extra_dependencies:
      amsmath: null
      inputenc: utf8
      xcolor: dvipsnames
      setspace: singlespacing
    toc: FALSE
    toc_depth: 3
urlcolor: purple
bibliography: finalprojectrefs.bib
csl: ieee-with-url.csl
abstract: "Using data from YouTube, we attempt to predict the number of views a video will receive using criteria such as location, category, number of likes, number of dislikes, and number of comments. Using forms of linear regression we have covered this semester, we will test various combinations of features for predictive power. [FILL IN OUTCOMES HERE]\\par\\textbf{Keywords:} Youtube, linear regression, count regression, R"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```
```{r loadLibraries}
library(knitr)
library(dplyr)
library(dtplyr)
library(stringr)
library(ggplot2)
library(scales)
library(corrplot)
library(data.table)
```
```{r loadData}
dataHeader <- "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/"
ca_path <- paste0(dataHeader, "CAvideos.csv")
us_path <- paste0(dataHeader, "USvideos.csv")
gb_path <- paste0(dataHeader, "GBvideos.csv")
fr_path <- paste0(dataHeader, "FRvideos.csv")
de_path <- paste0(dataHeader, "DEvideos.csv")
droppedFields <- c('description', 'thumbnail_link')
ca_set <- fread(ca_path, drop = droppedFields)
us_set <- fread(us_path, drop = droppedFields)
gb_set <- fread(gb_path, drop = droppedFields)
fr_set <- fread(fr_path, drop = droppedFields)
de_set <- fread(de_path, drop = droppedFields)
nobsUS <- nrow(us_set)
nobsCA <- nrow(ca_set)
nobsGB <- nrow(gb_set)
nobsFR <- nrow(fr_set)
nobsDE <- nrow(de_set)
nobsTot <- nobsUS + nobsCA + nobsGB + nobsFR + nobsDE
nobsTotP <- prettyNum(nobsTot, big.mark = ",")
nobsCAP <- prettyNum(nobsCA, big.mark = ",")
```

# Introduction
YouTube has changed the future of video entertainment forever [@Moylan2015ADo]. 
In 2019, the platform was estimated to have between \$16 billion and \$25
billion in revenue [@Wakabayashi2019YIa]. YouTube’s model connects a user’s
creativity with a desire for global recognition [@GoogleUtY]. Before YouTube,
international fame was not conceivable outside of a standard television or movie
studio. Today, creators from all over the world are gaining international
prominence using their own equipment and space. “YouTube is central to today’s
video ecosystem,” says Enders Analysis research analyst Jamie McGowan Stuart
[@Foster2020Yt1]. The current top channel “Vlad and Nikita” earns around \$312
thousand per video [@TYL]. According to a survey from Google, 6 out of 10 people
already prefer online video platforms or streaming services over live TV
[@OBTlv]. There are researchers predicting that by 2025---four years now---half
of viewers under the age of 32 will not pay TV service [@McQuivey2015B25].
Understanding some of the underpinnings of what generates views is beneficial to
anyone entering, or already in, the YouTube world.

The remainder of this paper will cover a literature review, an overview of out
methodology, the specifics of our modeling, a discussion of our findings,
thoughts for future work, a statistical appendix, a detailed code appendix, and
finally our references.

# Literature Review
With the popularity of Youtube, this is not a new question. Approaches this
question in the greater data science network include those based on more
advanced machine learning techniques such as SGD or neural net classifiers
[@LEZ2019YVP]. Others leveraged NLP and specially engineered features such as
"clickbait" or "NSFW" tags [@Srinivasan2017YVP]. These attempts usually used the
same dataset as we are.

Reviewing more academic literature uncovers research into the use of Support
Vector Regression with various basis functions on Youtube and/or Facebook videos
[@TR2017PPo; @PAG2013UEV]. Other attempts included building multi-stage treed
regression models where the outcome of a first stage determined which specific
second-stage model would be used for final popularity prediction [@OLL2016API].

These approaches usually added a temporal element to their analysis, and used
"earlier" values to predict later views. Using more sophisticated algorithms and
temporal elements tended to return statistically significant models. The
downside of these approaches are their complexity and opaqueness, of course. Our
approach will necessarily be simpler, although likely more transparent, being
restricted to the family of linear models covered in this course and not
regressing over time.

# Methodology
Using the famous data set from Kaggle [@Mitchell2019TYV], we will explore
relationships between a video’s views and the number of likes, dislikes, and
comments. We may also use a video’s country of origin and the category or type
of video as predictors.

We manually scrubbed the data---which had numerous issues---so that we had five
clean sets of statistics for the United States, Canada, Great Britain, France,
and Germany. We identified both numeric and factor predictors, and engineered
features for convenience as well. We will use these features to investigate
relationships with actual views using standard linear and count (e.g. Poisson)
regression models.

Given the models, we will compare the RMSE and MAE on a special holdout set and
the AIC on the training set and will select the model that performs best as the
winner for this paper. We do not expect it to outperform more sophisticated
models.

# Experimentation & Results
## Data Exploration
```{r ca_data_prep}
# Calculating the mean of the CA dataset
ca_mean <- mean(ca_set$views)

ca_set <- ca_set %>%
  # Add country label
  mutate(country = 'CA') %>% 
  # Calculate above/below view mean
  mutate(view_target = ifelse(views > ca_mean, 1L, 0L)) %>% 
  # Count number of tags
  mutate(tag_count = 1L + (str_count(ca_set$tags, ';'))) %>%
  # Now that we have the tag counts, we no longer need the tags themselves
  select(-tags)
```
```{r de_data_prep}
de_mean <- mean(de_set$views)

de_set <- de_set %>%
  mutate(country = 'DE') %>%
  mutate(view_target = ifelse(views > de_mean, 1L, 0L)) %>%
  mutate(tag_count = 1L + (str_count(de_set$tags, ';'))) %>%
  select(-tags)
```
```{r fr_data_prep}
fr_mean <- mean(fr_set$views)

fr_set <- fr_set %>%
  mutate(country = 'FR') %>%
  mutate(view_target = ifelse(views > fr_mean, 1L, 0L)) %>%
  mutate(tag_count = 1L + (str_count(fr_set$tags, ';'))) %>%
  select(-tags)
```
```{r gb_data_prep}
gb_mean <- mean(gb_set$views)

gb_set <- gb_set %>%
  mutate(country = 'GB') %>%
  mutate(view_target = ifelse(views > gb_mean, 1L, 0L)) %>%
  mutate(tag_count = 1L + (str_count(gb_set$tags, ';'))) %>%
  select(-tags)
```
```{r us_data_prep}
us_mean <- mean(us_set$views)

us_set <- us_set %>%
  mutate(country = 'US') %>%
  mutate(view_target = ifelse(views > us_mean, 1L, 0L)) %>%
  mutate(tag_count = 1L + (str_count(us_set$tags, ';'))) %>%
  select(-tags)
```
```{r merge_datasets}
# Since each country has the same format, we can just stack them one on top of
# the other. Much faster than merging.
full_set <- rbind(ca_set, de_set, fr_set, gb_set, us_set)

# Key the data table
setkey(full_set, title, country, trending_date)

# Calculate Global mean
full_mean <- mean(full_set$views)

# Calculate above/below global mean
full_set <- full_set %>%
  mutate(view_global_target = ifelse(views > full_mean, 1L, 0L))

# Convert trending_date to a date
full_set[, trending_date := as.IDate(trending_date, format = "%y.%d.%m")]

# Make name shorter for display purposes
setnames(full_set, c('comment_count', 'tag_count'), c('comments', 'tags'))
```
```{r mungeData}
# Categorize the variable names
boolVars <- c('comments_disabled', 'ratings_disabled', 'video_error_or_removed')
intVars <- c('views', 'likes', 'dislikes', 'comments', 'tags')
facVars <- c('country', 'category_id', 'view_target', 'view_global_target')
dateVars <- c('trending_date', 'publish_time')
charVars <- c('video_id', 'title', 'channel_title')

# Convert factors into factors. Set US as base factor for country
full_set[, `:=`(country = factor(country,
                                 levels = c('US', 'CA', 'GB', 'FR', 'DE'),
                                 labels = c('US', 'CA', 'GB', 'FR', 'DE')),
                category_id = as.factor(category_id),
                view_target = factor(view_target, levels = c(0, 1),
                                     labels = c(0, 1)),
                view_global_target = factor(view_global_target,
                                            levels = c(0, 1),
                                            labels = c(0, 1)))]
# Melt for numerics
fullSetN <- melt(full_set, id.vars = c(charVars, dateVars, boolVars, facVars),
                 variable.factor = FALSE, variable.name = 'measure',
                 value.name = 'value')

# Melt for factors except country
fullSetF <- melt(full_set,
                 measure.vars = c('category_id', 'view_target', 'view_global_target'),
                 variable.factor = FALSE, variable.name = 'measure',
                 value.name = 'value')
```
```{r summaryStats}
statsN <- data.table()
# Loop through countries and calculate statistics. Table shown in appendix.
for (i in unique(fullSetN$country)) {
  statsN <- rbind(statsN, cbind(country = i,
                  fullSetN[country == i,
                           .(Mean = mean(value, na.rm = TRUE),
                             SD = sd(value, na.rm = TRUE),
                             Min = min(value, na.rm = TRUE),
                             Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
                             Median = median(value, na.rm = TRUE),
                             Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
                             Max = max(value, na.rm = TRUE),
                             IQR = IQR(value, na.rm = TRUE)),
                           keyby = c('measure')]))
}
# Add global stats
statsN <- rbind(statsN,
                cbind(country = "Global",
                      fullSetN[,
                               .(Mean = mean(value, na.rm = TRUE),
                                 SD = sd(value, na.rm = TRUE),
                                 Min = min(value, na.rm = TRUE),
                                 Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
                                 Median = median(value, na.rm = TRUE),
                                 Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
                                 Max = max(value, na.rm = TRUE),
                                 IQR = IQR(value, na.rm = TRUE)),
                               keyby = c('measure')]))
```

### Numeric Predictors
For numeric predictors, we are using the number of `likes`, `dislikes`, `tags`,
and `comments`. We may also consider the country of viewing as well. From the 
There is not that much difference in the distribution of the numeric predictors
between the countries where people view.

```{r numLogDensities}
ggplot(fullSetN[measure != 'views'], aes(x = value, color = measure)) +
  geom_density(kernel = "epanechnikov") +
  facet_wrap(~ country) +
  ggtitle("Predictor Density (semi-log scale)") +
  scale_x_log10(label = scientific)
```

### Factor Predictors
While tabular representation of factor predictors is difficult, a distribution
of factors by value may prove informative. It is clear from the graphs below
that interests vary by country. Category `24` seems to be the universal mode.
Except for Great Britain, where they love their category `10`!

```{r factorPlot1, fig.height=4.5, fig.width=8}
ggplot(fullSetF[measure == 'category_id'], aes(y = value)) +
  geom_bar() +
  facet_wrap(~ country) +
  ggtitle("Distribution by Category & Country") +
  scale_x_continuous(labels = comma)
```

```{r factorPlot2, fig.height=3, fig.width=8}
ggplot(fullSetF[measure != 'category_id'],
             aes(x = value, fill = measure)) +
  geom_bar(position = 'dodge2') +
  facet_wrap(~ country) +
  ggtitle("Country and Global Mean Views Exceedence") +
  scale_y_continuous(labels = comma)
```

As with many events which follow the Pareto rule, most of the views are 
attributable to a minority of the videos, as seen by the 0/1 disparity on both
an individual country and a global basis. Great Britain has the closest parity
on a global basis. British viewers are closest to the global average.

## Feature Engineering

## Model Building

## Model Evaluation

## Findings

# Discussion & Conclusions

# Appendix
## Statistical Appendix
A more detailed analysis of the empirical statistics for the numeric data is 
found in the table below.

\footnotesize
```{r statsTable}
kable(statsN, digits = 2,
      caption = "Statistics for Numeric Variables by Country & Global",
      format.args = list(big.mark = ","))
```
\normalsize

## Code Appendix
The code chunks below represent the R code called in order during the analysis.
They are reproduced in the appendix for review and comment.

```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```

```{r loadLibraries}
```
```{r loadData}
```
```{r ca_data_prep}
```
```{r de_data_prep}
```
```{r fr_data_prep}
```
```{r gb_data_prep}
```
```{r us_data_prep}
```
```{r merge_datasets}
```
```{r mungeData}
```
```{r summaryStats}
```
```{r factorPlot1}
```
```{r factorPlot1}
```

<!-- Statistical Appendix-->
```{r statsTable}
```

<!-- References needs to be last -->  
# References

### 2020-12-01 Final_project/Final Project Paper - Group 2.Rmd
Some more cleanup of unused code and squish graphic to save space

@@
 library(scales)
-library(corrplot)
 library(data.table)
@@
 nobsTot <- nobsUS + nobsCA + nobsGB + nobsFR + nobsDE
-nobsTotP <- prettyNum(nobsTot, big.mark = ",")
-nobsCAP <- prettyNum(nobsCA, big.mark = ",")
 ```
@@
 
-```{r numLogDensities}
+```{r numLogDensities, fig.height=4.5, fig.width=8}
 ggplot(fullSetN[measure != 'views'], aes(x = value, color = measure)) +

### 2020-12-01 Final_project/Final Project Paper - Group 2.Rmd
Doh! called same chunk twice

@@
 ```
-```{r factorPlot1}
+```{r factorPlot2}
 ```

### 2020-12-01 Final_project/Final Project Paper - Group 2.Rmd
Add first run at rules (country) based linear model

@@
 ```{r loadLibraries}
+library(doParallel)
 library(knitr)
@@
 library(scales)
+library(caret)
 library(data.table)
@@
 Given the models, we will compare the RMSE and MAE on a special holdout set and
-the AIC on the training set and will select the model that performs best as the
-winner for this paper. We do not expect it to outperform more sophisticated
-models.
+will select the model that performs best as the winner for this paper. We do not
+expect it to outperform more sophisticated models.
 
@@
 
-## Feature Engineering
+## Feature Engineering & Selection
+Since we are not factoring in time, it is incorrect to use all the observations.
+Therefore, we will extract the latest observation by video by country and use
+this subset.
+
+```{r subsetExtraction}
+# Select the row numbers of the first entry of the latest trending date by title
+# and by country. There are 34 duplicates. Use this as our restricted data set
+# corresponding to the most recent view count. The inner set of brackets gets
+# the row number (called V1) and the outer set is a simple subset by those row
+# numbers.
+fullExtract <- full_set[
+  full_set[, .I[which.max(trending_date)], by = c('title', 'country')]$V1
+  ]
+```
+
+A feature that we will also consider is `lkratio`: the ratio of likes to
+dislikes. This will help ascertain if it is the magnitude or the preponderance
+which has a greater effect. As there may be some rare videos with no dislikes,
+the ratios will be set to 1. To capture the effect of likes with no dislikes, an
+indicator variable will be created or those "perfect" videos.
+
+```{r addFeatures}
+fullExtract[, `:=`(perfect =ifelse(likes > 0 & dislikes == 0, TRUE, FALSE),
+                   lkratio = ifelse(dislikes == 0, 1, likes / dislikes))]
+```
 
 ## Model Building
+We will first separate 20% of the data as a true holdout set. It is on this data
+that our models will be compared.
+
+```{r holdOut}
+set.seed(617)
+seenIDX <- createDataPartition(fullExtract$views, p = 0.8)$Resample1
+seenSet <- fullExtract[seenIDX, ]
+hideSet <- fullExtract[-seenIDX, ]
+```
+
+We will train models on the remaining 80% of the data.
+
+### Model 1: Linear Regression
+Keeping country as a predictor is of less interest. It will most likely track
+some measure of population. It is of more interest to see if there are
+differences in the factors between countries. To capture this by interactions
+would be combinatorially unfeasible. Instead, five models will be run, each on
+a country specific subset. This is akin to a manual implementation of a rules
+based regression model
+
+```{r modelLinear1, cache = 2L}
+# Using stepAIC means no cross-validation. Train on entire dataset.
+trC <- trainControl(method = 'none', allowParallel = TRUE)
+cl <- makePSOCKcluster(5)
+registerDoParallel(cl)
+set.seed(181)
+for (ct in unique(fullExtract$country)) {
+  linMod <- train(views ~ likes + dislikes + comments + tags + category_id +
+                    comments_disabled + ratings_disabled +
+                    video_error_or_removed + perfect + lkratio,
+                  data = fullExtract[country == ct],
+                  family = gaussian(link = 'identity'), direction = 'both',
+                  method = 'glmStepAIC', trace = 0, trControl = trC)
+  assign(paste0('linMod', ct), linMod)
+}
+stopCluster(cl)
+lM1US <- linModUS$finalModel
+lM1CA <- linModCA$finalModel
+lM1GB <- linModGB$finalModel
+lM1FR <- linModFR$finalModel
+lM1DE <- linModDE$finalModel
+clM1US <- linModUS$finalModel$coefficients
+clM1CA <- linModCA$finalModel$coefficients
+clM1GB <- linModGB$finalModel$coefficients
+clM1FR <- linModFR$finalModel$coefficients
+clM1DE <- linModDE$finalModel$coefficients
+nclM1US <- names(clM1US)
+nclM1CA <- names(clM1CA)
+nclM1GB <- names(clM1GB)
+nclM1FR <- names(clM1FR)
+nclM1DE <- names(clM1DE)
+nnclM1US <- length(nclM1US)
+nnclM1CA <- length(nclM1CA)
+nnclM1GB <- length(nclM1GB)
+nnclM1FR <- length(nclM1FR)
+nnclM1DE <- length(nclM1DE)
+length(nclM1US) <- 22
+length(nclM1GB) <- 22
+length(nclM1FR) <- 22
+length(nclM1DE) <- 22
+```
+
+\footnotesize
+```{r m1Table}
+kable(data.frame(cbind(nclM1CA, nclM1DE, nclM1US, nclM1FR, nclM1GB)),
+      caption = "Model Coefficients by Country")
+```
+\normalsize
+
+The most parsimonious model is the one for Great Britain with only `r nnclM1GB`
+parameters. The most complex is the one for Canada with `r nnclM1CA`.
 
@@
 ```
+<!--  Data Exploration -->
 ```{r ca_data_prep}
@@
 
+<!-- Feature Engineering & Selection -->
+```{r subsetExtraction}
+```
+```{r addFeatures}
+```
+
+<!-- Model Building -->
+```{r holdOut}
+```
+```{r modelLinear1}
+```
+```{r m1Table}
+```
+
 <!-- Statistical Appendix-->

### 2020-12-02 Final_project/Final Project Paper - Group 2.Rmd
Clean up Linear 1 start linear 2

@@
 
-# Convert trending_date to a date
-full_set[, trending_date := as.IDate(trending_date, format = "%y.%d.%m")]
+# Convert trending_date to a date and views to double since the sum > 2^31
+full_set[, `:=`(trending_date = as.IDate(trending_date, format = "%y.%d.%m"),
+                views = as.double(views))]
 
@@
 
-### Model 1: Linear Regression
-Keeping country as a predictor is of less interest. It will most likely track
+### Multiple Linear Regression
+Linear regression may be the best known algorithm used when analyzing a
+continuous numeric outcome. It searches for a linear relationship of the
+predictors that minimizes the squared error between the "predictor" function and
+the observations [@Sheather2009AMA].
+
+#### Model 1: Model by Country
+The first linear regression model will actually be a rule-based set of models.
+The hypothesis is that using country as a factor will most likely simply track
 some measure of population. It is of more interest to see if there are
@@
 trC <- trainControl(method = 'none', allowParallel = TRUE)
+
+# Capture the unique country codes
+ctCodes <- unique(fullExtract$country)
 cl <- makePSOCKcluster(5)
@@
 set.seed(181)
-for (ct in unique(fullExtract$country)) {
+maxPars <- 0
+for (ct in ctCodes) {
+  # Run the StepAIC algorithm inside the loop
   linMod <- train(views ~ likes + dislikes + comments + tags + category_id +
@@
                   method = 'glmStepAIC', trace = 0, trControl = trC)
+
+  # Assign the outcomes to dynamically "named" variables inside the loop
   assign(paste0('linMod', ct), linMod)
+  assign(paste0('lM1', ct), linMod$finalModel)
+  assign(paste0('clM1', ct), linMod$finalModel$coefficients)
+  assign(paste0('nclM1', ct), names(linMod$finalModel$coefficients))
+  assign(paste0('nnclM1', ct), length(names(linMod$finalModel$coefficients)))
+  if ((length(names(linMod$finalModel$coefficients))) > maxPars) {
+    maxPars <- length(names(linMod$finalModel$coefficients))
+  }
 }
 stopCluster(cl)
-lM1US <- linModUS$finalModel
-lM1CA <- linModCA$finalModel
-lM1GB <- linModGB$finalModel
-lM1FR <- linModFR$finalModel
-lM1DE <- linModDE$finalModel
-clM1US <- linModUS$finalModel$coefficients
-clM1CA <- linModCA$finalModel$coefficients
-clM1GB <- linModGB$finalModel$coefficients
-clM1FR <- linModFR$finalModel$coefficients
-clM1DE <- linModDE$finalModel$coefficients
-nclM1US <- names(clM1US)
-nclM1CA <- names(clM1CA)
-nclM1GB <- names(clM1GB)
-nclM1FR <- names(clM1FR)
-nclM1DE <- names(clM1DE)
-nnclM1US <- length(nclM1US)
-nnclM1CA <- length(nclM1CA)
-nnclM1GB <- length(nclM1GB)
-nnclM1FR <- length(nclM1FR)
-nnclM1DE <- length(nclM1DE)
-length(nclM1US) <- 22
-length(nclM1GB) <- 22
-length(nclM1FR) <- 22
-length(nclM1DE) <- 22
+
+# R has difficulty with ragged arrays, so override the lengths to pad with NAs
+# Splitting for 80 char convenience
+length(nclM1US) <- length(nclM1GB) <- length(nclM1FR) <- maxPars
+length(nclM1DE) <- length(nclM1CA) <- maxPars
 ```
@@
 parameters. The most complex is the one for Canada with `r nnclM1CA`.
+
+### Model 2: Linear Regression Globally
+While a rule-based model has the ability to fit data more closely based on those
+rules, it adds a significant amount of both parameter and model error
+[@Adler2015EtP; Adler2014For]. To test this hypothesis, we will run the same
+linear regression, but on the global results: the sum of all the numeric values
+by country except tags count. The other factors beside country deal with the
+video itself, like tag count, therefore they are identical across the countries.
+
+```{r m2Prep}
+globalExtract <- unique(fullExtract[,
+                                    .(views = sum(views),
+                                      likes = sum(likes),
+                                      dislikes = sum(dislikes),
+                                      comments = sum(comments),
+                                      tags = tags,
+                                      category_id,
+                                      comments_disabled,
+                                      ratings_disabled,
+                                      video_error_or_removed),
+                                    keyby = title])
+
+```
 

### 2020-12-02 Final_project/Final Project Paper - Group 2.Rmd
OK. The issue seems to be one where downloading strips some of the crazy characters. Maybe it's not using UTF-8. Therefore, everyone can download from LOCAL once they pull the data from the repo.

@@
 ```{r loadData}
-dataHeader <- "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/"
+dataHeader <- "../Final_project/Data2/"
 ca_path <- paste0(dataHeader, "CAvideos.csv")
@@
 de_path <- paste0(dataHeader, "DEvideos.csv")
+colClass <- c(rep('character', 4L), 'integer', 'POSIXct', 'character',
+              rep('double', 4L), 'character', rep('boolean', 3L), 'character')
 droppedFields <- c('description', 'thumbnail_link')
-ca_set <- fread(ca_path, drop = droppedFields)
-us_set <- fread(us_path, drop = droppedFields)
-gb_set <- fread(gb_path, drop = droppedFields)
-fr_set <- fread(fr_path, drop = droppedFields)
-de_set <- fread(de_path, drop = droppedFields)
+
+# Download, expand, and ingest raw data files
+ca_set <- fread(ca_path, drop = droppedFields, colClasses = colClass)
+
+us_set <- fread(unzip(temp), drop = droppedFields)
+
+gb_set <- fread(unzip(temp), drop = droppedFields)
+
+fr_set <- fread(unzip(temp), drop = droppedFields)
+
+de_set <- fread(unzip(temp), drop = droppedFields)
+
 nobsUS <- nrow(us_set)

### 2020-12-02 Final_project/Final Project Paper - Group 2.Rmd
Running on US only

--- 2020-12-01 Final_project/Final Project Paper - Group 2.Rmd
@@
 ```{r loadLibraries}
-library(doParallel)
+library(jsonlite)
 library(knitr)
@@
 ```{r loadData}
-dataHeader <- "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/"
-ca_path <- paste0(dataHeader, "CAvideos.csv")
-us_path <- paste0(dataHeader, "USvideos.csv")
-gb_path <- paste0(dataHeader, "GBvideos.csv")
-fr_path <- paste0(dataHeader, "FRvideos.csv")
-de_path <- paste0(dataHeader, "DEvideos.csv")
-droppedFields <- c('description', 'thumbnail_link')
-ca_set <- fread(ca_path, drop = droppedFields)
-us_set <- fread(us_path, drop = droppedFields)
-gb_set <- fread(gb_path, drop = droppedFields)
-fr_set <- fread(fr_path, drop = droppedFields)
-de_set <- fread(de_path, drop = droppedFields)
+currentPath <- getwd()
+dataPath <- "Data2"
+us_path <- file.path(currentPath, dataPath, "USvideos.csv")
+
+colClass <- c(rep('character', 4L), 'integer', 'POSIXct', 'character',
+              rep('double', 4L), rep('boolean', 3L))
+us_set <- fread(us_path, encoding = 'UTF-8', colClasses = colClass)
 nobsUS <- nrow(us_set)
-nobsCA <- nrow(ca_set)
-nobsGB <- nrow(gb_set)
-nobsFR <- nrow(fr_set)
-nobsDE <- nrow(de_set)
-nobsTot <- nobsUS + nobsCA + nobsGB + nobsFR + nobsDE
+
+# Category IDs
+us_cat_path <- file.path(currentPath, dataPath, "US_category_id.json")
+us_cats <- fromJSON(us_cat_path)
+
+us_cats <- data.table(id = as.integer(us_cats$items$id),
+                      category = us_cats$items$snippet[, 2])
+setkey(us_cats, id)
 ```
@@
 
-We manually scrubbed the data---which had numerous issues---so that we had five
-clean sets of statistics for the United States, Canada, Great Britain, France,
-and Germany. We identified both numeric and factor predictors, and engineered
+We manually scrubbed the data and discovered that the country-specific files
+really were not! They neither refer to videos created by country nor do they
+refer to views *specific* to country. Rather they are the total number of views
+and other predictors for that video on that day as collected by someone within
+that country. Meaning that aggregation is almost always multiplying. There may
+be some videos unique to a specific country---one which was not viewed in other
+countries, but they are many magnitudes smaller than those seen by all. As the
+United States had the most observations, we decided to analyze its data set.
+
+With the data, we identified both numeric and factor predictors, and engineered
 features for convenience as well. We will use these features to investigate
-relationships with actual views using standard linear and count (e.g. Poisson)
-regression models.
+relationships with actual views using linear regression models.
 
@@
 ## Data Exploration
-```{r ca_data_prep}
-# Calculating the mean of the CA dataset
-ca_mean <- mean(ca_set$views)
 
-ca_set <- ca_set %>%
-  # Add country label
-  mutate(country = 'CA') %>% 
-  # Calculate above/below view mean
-  mutate(view_target = ifelse(views > ca_mean, 1L, 0L)) %>% 
-  # Count number of tags
-  mutate(tag_count = 1L + (str_count(ca_set$tags, ';'))) %>%
-  # Now that we have the tag counts, we no longer need the tags themselves
-  select(-tags)
-```
-```{r de_data_prep}
-de_mean <- mean(de_set$views)
+```{r dataPrep}
+# Substitute category for numeric ID through joining
+us_set <- us_cats[us_set, on = 'id == category_id']
 
-de_set <- de_set %>%
-  mutate(country = 'DE') %>%
-  mutate(view_target = ifelse(views > de_mean, 1L, 0L)) %>%
-  mutate(tag_count = 1L + (str_count(de_set$tags, ';'))) %>%
-  select(-tags)
-```
-```{r fr_data_prep}
-fr_mean <- mean(fr_set$views)
+# Change data types for convenience, add target bucket, count tags and then
+# get rid of actual tags and category IDs
 
-fr_set <- fr_set %>%
-  mutate(country = 'FR') %>%
-  mutate(view_target = ifelse(views > fr_mean, 1L, 0L)) %>%
-  mutate(tag_count = 1L + (str_count(fr_set$tags, ';'))) %>%
-  select(-tags)
-```
-```{r gb_data_prep}
-gb_mean <- mean(gb_set$views)
-
-gb_set <- gb_set %>%
-  mutate(country = 'GB') %>%
-  mutate(view_target = ifelse(views > gb_mean, 1L, 0L)) %>%
-  mutate(tag_count = 1L + (str_count(gb_set$tags, ';'))) %>%
-  select(-tags)
-```
-```{r us_data_prep}
 us_mean <- mean(us_set$views)
 
-us_set <- us_set %>%
-  mutate(country = 'US') %>%
-  mutate(view_target = ifelse(views > us_mean, 1L, 0L)) %>%
-  mutate(tag_count = 1L + (str_count(us_set$tags, ';'))) %>%
-  select(-tags)
-```
-```{r merge_datasets}
-# Since each country has the same format, we can just stack them one on top of
-# the other. Much faster than merging.
-full_set <- rbind(ca_set, de_set, fr_set, gb_set, us_set)
-
-# Key the data table
-setkey(full_set, title, country, trending_date)
-
-# Calculate Global mean
-full_mean <- mean(full_set$views)
-
-# Calculate above/below global mean
-full_set <- full_set %>%
-  mutate(view_global_target = ifelse(views > full_mean, 1L, 0L))
-
-# Convert trending_date to a date
-full_set[, trending_date := as.IDate(trending_date, format = "%y.%d.%m")]
+us_set[, `:=`(trending_date = as.IDate(trending_date, format = "%y.%d.%m"),
+              views = as.double(views),
+              tag_count = 1L + str_count(us_set$tags, '[|]'),
+              view_target = factor(ifelse(views > us_mean, 1L, 0L),
+                                   levels = c(0, 1), labels = c(0, 1)),
+              id = NULL,
+              category = as.factor(category))
+       ][, `:=`(tags = NULL,
+                category = relevel(category, 'Entertainment'))]
 
 # Make name shorter for display purposes
-setnames(full_set, c('comment_count', 'tag_count'), c('comments', 'tags'))
-```
-```{r mungeData}
+setnames(us_set, c('comment_count', 'tag_count'), c('comments', 'tags'))
+
 # Categorize the variable names
 boolVars <- c('comments_disabled', 'ratings_disabled', 'video_error_or_removed')
-intVars <- c('views', 'likes', 'dislikes', 'comments', 'tags')
-facVars <- c('country', 'category_id', 'view_target', 'view_global_target')
+numVars <- c('views', 'likes', 'dislikes', 'comments', 'tags')
+facVars <- c('category', 'view_target')
 dateVars <- c('trending_date', 'publish_time')
@@
 
-# Convert factors into factors. Set US as base factor for country
-full_set[, `:=`(country = factor(country,
-                                 levels = c('US', 'CA', 'GB', 'FR', 'DE'),
-                                 labels = c('US', 'CA', 'GB', 'FR', 'DE')),
-                category_id = as.factor(category_id),
-                view_target = factor(view_target, levels = c(0, 1),
-                                     labels = c(0, 1)),
-                view_global_target = factor(view_global_target,
-                                            levels = c(0, 1),
-                                            labels = c(0, 1)))]
 # Melt for numerics
-fullSetN <- melt(full_set, id.vars = c(charVars, dateVars, boolVars, facVars),
+usSetN <- melt(us_set, measure.vars = numVars, variable.factor = FALSE,
+               variable.name = 'measure', value.name = 'value')
+
+# Melt for factors
+usSetF <- melt(us_set,
+               measure.vars = c('category', 'view_target'),
                variable.factor = FALSE, variable.name = 'measure',
@@
 
-# Melt for factors except country
-fullSetF <- melt(full_set,
-                 measure.vars = c('category_id', 'view_target', 'view_global_target'),
-                 variable.factor = FALSE, variable.name = 'measure',
-                 value.name = 'value')
-```
-```{r summaryStats}
-statsN <- data.table()
-# Loop through countries and calculate statistics. Table shown in appendix.
-for (i in unique(fullSetN$country)) {
-  statsN <- rbind(statsN, cbind(country = i,
-                  fullSetN[country == i,
-                           .(Mean = mean(value, na.rm = TRUE),
+# Table of summary statistics
+statsN <- usSetN[, .(Mean = mean(value, na.rm = TRUE),
                      SD = sd(value, na.rm = TRUE),
@@
                      IQR = IQR(value, na.rm = TRUE)),
-                           keyby = c('measure')]))
-}
-# Add global stats
-statsN <- rbind(statsN,
-                cbind(country = "Global",
-                      fullSetN[,
-                               .(Mean = mean(value, na.rm = TRUE),
-                                 SD = sd(value, na.rm = TRUE),
-                                 Min = min(value, na.rm = TRUE),
-                                 Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
-                                 Median = median(value, na.rm = TRUE),
-                                 Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
-                                 Max = max(value, na.rm = TRUE),
-                                 IQR = IQR(value, na.rm = TRUE)),
-                               keyby = c('measure')]))
+                 keyby = c('measure')]
 ```
@@
 ```{r numLogDensities, fig.height=4.5, fig.width=8}
-ggplot(fullSetN[measure != 'views'], aes(x = value, color = measure)) +
+ggplot(usSetN[measure != 'views'], aes(x = value, color = measure)) +
   geom_density(kernel = "epanechnikov") +
-  facet_wrap(~ country) +
   ggtitle("Predictor Density (semi-log scale)") +
@@
 of factors by value may prove informative. It is clear from the graphs below
-that interests vary by country. Category `24` seems to be the universal mode.
-Except for Great Britain, where they love their category `10`!
+that interests vary by category. `Entertainment` seems to be the most common in
+the US with `Music` coming in second.
 
 ```{r factorPlot1, fig.height=4.5, fig.width=8}
-ggplot(fullSetF[measure == 'category_id'], aes(y = value)) +
+ggplot(usSetF[measure == 'category'], aes(y = value)) +
   geom_bar() +
-  facet_wrap(~ country) +
-  ggtitle("Distribution by Category & Country") +
+  ggtitle("Distribution by Category") +
   scale_x_continuous(labels = comma)
@@
 ```{r factorPlot2, fig.height=3, fig.width=8}
-ggplot(fullSetF[measure != 'category_id'],
-             aes(x = value, fill = measure)) +
+ggplot(usSetF[measure != 'category'], aes(x = value, fill = measure)) +
   geom_bar(position = 'dodge2') +
-  facet_wrap(~ country) +
   ggtitle("Country and Global Mean Views Exceedence") +
@@
 As with many events which follow the Pareto rule, most of the views are 
-attributable to a minority of the videos, as seen by the 0/1 disparity on both
-an individual country and a global basis. Great Britain has the closest parity
-on a global basis. British viewers are closest to the global average.
+attributable to a minority of the videos, as seen by the 0/1 disparity.
 
@@
 # numbers.
-fullExtract <- full_set[
-  full_set[, .I[which.max(trending_date)], by = c('title', 'country')]$V1
+usExtract <- us_set[
+  us_set[, .I[which.max(views)], by = c('title')]$V1
   ]
@@
 ```{r addFeatures}
-fullExtract[, `:=`(perfect =ifelse(likes > 0 & dislikes == 0, TRUE, FALSE),
+usExtract[, `:=`(perfect =ifelse(likes > 0 & dislikes == 0, TRUE, FALSE),
                  lkratio = ifelse(dislikes == 0, 1, likes / dislikes))]
@@
 We will first separate 20% of the data as a true holdout set. It is on this data
-that our models will be compared.
+that our models will be compared. We will train models on the remaining 80% of
+the data.
 
@@
 set.seed(617)
-seenIDX <- createDataPartition(fullExtract$views, p = 0.8)$Resample1
-seenSet <- fullExtract[seenIDX, ]
-hideSet <- fullExtract[-seenIDX, ]
+seenIDX <- createDataPartition(usExtract$views, p = 0.8)$Resample1
+seenSet <- usExtract[seenIDX, ]
+hideSet <- usExtract[-seenIDX, ]
 ```
 
-We will train models on the remaining 80% of the data.
+### Multiple Linear Regression
+Linear regression may be the best known algorithm used when analyzing a
+continuous numeric outcome. It searches for a linear relationship of the
+predictors that minimizes the squared error between the "predictor" function and
+the observations [@Sheather2009AMA]. This model will engage in stepwise feature
+using the AIC as the optimization metric.
 
-### Model 1: Linear Regression
-Keeping country as a predictor is of less interest. It will most likely track
-some measure of population. It is of more interest to see if there are
-differences in the factors between countries. To capture this by interactions
-would be combinatorially unfeasible. Instead, five models will be run, each on
-a country specific subset. This is akin to a manual implementation of a rules
-based regression model
+```{r modelLinear1}
+# Using stepAIC means no cross-validation. Train on entire dataset.
+trC <- trainControl(method = 'none')
 
-```{r modelLinear1, cache = 2L}
-# Using stepAIC means no cross-validation. Train on entire dataset.
-trC <- trainControl(method = 'none', allowParallel = TRUE)
-cl <- makePSOCKcluster(5)
-registerDoParallel(cl)
+# Capture the unique country codes
 set.seed(181)
-for (ct in unique(fullExtract$country)) {
-  linMod <- train(views ~ likes + dislikes + comments + tags + category_id +
+linMod1 <- train(views ~ likes + dislikes + comments + tags + category +
                    comments_disabled + ratings_disabled +
                    video_error_or_removed + perfect + lkratio,
-                  data = fullExtract[country == ct],
-                  family = gaussian(link = 'identity'), direction = 'both',
-                  method = 'glmStepAIC', trace = 0, trControl = trC)
-  assign(paste0('linMod', ct), linMod)
-}
-stopCluster(cl)
-lM1US <- linModUS$finalModel
-lM1CA <- linModCA$finalModel
-lM1GB <- linModGB$finalModel
-lM1FR <- linModFR$finalModel
-lM1DE <- linModDE$finalModel
-clM1US <- linModUS$finalModel$coefficients
-clM1CA <- linModCA$finalModel$coefficients
-clM1GB <- linModGB$finalModel$coefficients
-clM1FR <- linModFR$finalModel$coefficients
-clM1DE <- linModDE$finalModel$coefficients
-nclM1US <- names(clM1US)
-nclM1CA <- names(clM1CA)
-nclM1GB <- names(clM1GB)
-nclM1FR <- names(clM1FR)
-nclM1DE <- names(clM1DE)
-nnclM1US <- length(nclM1US)
-nnclM1CA <- length(nclM1CA)
-nnclM1GB <- length(nclM1GB)
-nnclM1FR <- length(nclM1FR)
-nnclM1DE <- length(nclM1DE)
-length(nclM1US) <- 22
-length(nclM1GB) <- 22
-length(nclM1FR) <- 22
-length(nclM1DE) <- 22
+                 data = usExtract, family = gaussian(link = 'identity'),
+                 method = 'glmStepAIC', direction = 'both', trace = 0,
+                 trControl = trC)
 ```
@@
 ```{r m1Table}
-kable(data.frame(cbind(nclM1CA, nclM1DE, nclM1US, nclM1FR, nclM1GB)),
-      caption = "Model Coefficients by Country")
+kable(summary(linMod1$finalModel)$coefficients,
+      caption = "Model 1 Linear Regression Output",
+      digts = 3L, format.args = list(big.mark = ','))
 ```
@@
 
-The most parsimonious model is the one for Great Britain with only `r nnclM1GB`
-parameters. The most complex is the one for Canada with `r nnclM1CA`.
+The sign of these coefficients make sense in the main. The most predictive
+feature for views is `likes` followed by `dislikes`. Of course, one usually
+views a video at least once prior to rating it. We set the baseline category to
+`Entertainment`, considering it was the most popular. Therfore we expected
+negative coefficients for other categories. Interestingly, `Film & Animation`
+is positively correlated with likes over and above `Entertainment`, despite
+having fewer views.
+
+Other observations are that the magnitude of the coefficients is similar to that
+of the intercept, indicating that the predictors provide large swings to final
+views. The "worst" predictor by magnitude is the `Nonprofits & Activism`
+category; its negative coefficient is more than twice the baseline intercept
+views. The best, interestingly is if ratings are disabled. This is
+counterintuitive at first, but is logical. The amount of `likes`, `dislikes`,
+and `comments` are many orders of magnitude greater than 1. Therefore, their
+coefficients can be much smaller. A Boolean variable is either 1 or 0, therefore
+its coefficient is much greater even if its actual contribution is lower.
+
+Variable importance can be seen from the table below. The further down the list
+a coefficient preceded by a **`-`** is, the more important it is and the more it
+affected the AIC.
+\footnotesize
+```
+                                  Df   Deviance    AIC
+<none>                               7.1243e+16 212209
++ `categoryAutos & Vehicles`       1 7.1222e+16 212209
++ tags                             1 7.1223e+16 212209
+- `categoryPeople & Blogs`         1 7.1278e+16 212210
+- categoryEducation                1 7.1278e+16 212210
+- perfectTRUE                      1 7.1279e+16 212210
++ video_error_or_removedTRUE       1 7.1237e+16 212210
++ `categoryPets & Animals`         1 7.1239e+16 212210
++ categoryGaming                   1 7.1239e+16 212210
++ `categoryTravel & Events`        1 7.1240e+16 212210
++ comments_disabledTRUE            1 7.1240e+16 212210
++ categorySports                   1 7.1240e+16 212210
++ `categoryScience & Technology`   1 7.1241e+16 212211
++ categoryShows                    1 7.1243e+16 212211
+- `categoryHowto & Style`          1 7.1287e+16 212211
+- `categoryFilm & Animation`       1 7.1299e+16 212212
+- `categoryNews & Politics`        1 7.1321e+16 212214
+- categoryMusic                    1 7.1338e+16 212215
+- `categoryNonprofits & Activism`  1 7.1345e+16 212216
+- ratings_disabledTRUE             1 7.1463e+16 212227
+- categoryComedy                   1 7.1467e+16 212227
+- lkratio                          1 7.3555e+16 212413
+- dislikes                         1 8.9091e+16 213650
+- comments                         1 8.9807e+16 213701
+- likes                            1 2.0217e+17 218939
+```
+\normalsize
 
@@
 ```
-<!--  Data Exploration -->
-```{r ca_data_prep}
+```{r dataPrep}
 ```
-```{r de_data_prep}
-```
-```{r fr_data_prep}
-```
-```{r gb_data_prep}
-```
-```{r us_data_prep}
-```
-```{r merge_datasets}
-```
-```{r mungeData}
-```
-```{r summaryStats}
+```{r numLogDensities}
 ```

### 2020-12-03 Final_project/Final Project Paper - Group 2.Rmd
Update

@@
 library(knitr)
-library(dplyr)
-library(dtplyr)
 library(stringr)
@@
 
-Given the models, we will compare the RMSE and MAE on a special holdout set and
-will select the model that performs best as the winner for this paper. We do not
-expect it to outperform more sophisticated models.
+Given the models, we will compare the RMSE, \(R^2\), and MAE on a holdout set
+and will select the model that performs best as the winner for this paper. We do
+not expect it to outperform more sophisticated models.
 
@@
 
+### Target Variable
+
+```{r targDens}
+ggplot(usSetN[measure == 'views'], aes(x = value)) +
+  geom_density(kernel = "epanechnikov") +
+  ggtitle("Target Density (semi-log scale)") +
+  scale_x_log10(label = scientific)
+```
+
+On a log scale, `views` looks rather Gaussian, which implies it has a lognormal
+distribution.
+
 ### Numeric Predictors
@@
 ```
+
+As expected, the number of tags is orders of magnitude less than the rating or
+comment variables. On average there are more likes than comments and more
+comments then dislikes, but all three exhibit symmetric Gaussian-like behavior.
 
@@
 
+We will consider the relationship between `views` and the numerical predictors
+of `likes`, `dislikes`, `comments`, and `tags`. We will also consider the
+`category`, whether or not comments or ratings were disabled and, if there was
+an error with the video.
+
 A feature that we will also consider is `lkratio`: the ratio of likes to
@@
 which has a greater effect. As there may be some rare videos with no dislikes,
-the ratios will be set to 1. To capture the effect of likes with no dislikes, an
-indicator variable will be created or those "perfect" videos.
+the ratios will be set to 1.
 
 ```{r addFeatures}
-usExtract[, `:=`(perfect =ifelse(likes > 0 & dislikes == 0, TRUE, FALSE),
-                 lkratio = ifelse(dislikes == 0, 1, likes / dislikes))]
+usExtract[, lkratio := ifelse(dislikes == 0, 1, likes / dislikes)]
+
+modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
+                   comments_disabled + ratings_disabled +
+                   video_error_or_removed + lkratio,
+                   data = usExtract, fullRank = TRUE)
+modMat <- predict(modDum, usExtract)
 ```
@@
 seenIDX <- createDataPartition(usExtract$views, p = 0.8)$Resample1
-seenSet <- usExtract[seenIDX, ]
-hideSet <- usExtract[-seenIDX, ]
+seenX <- modMat[seenIDX, ]
+seenY <- usExtract$views[seenIDX]
+hideX <- as.data.frame(modMat[-seenIDX, ])
+hideY <- usExtract$views[-seenIDX]
 ```
 
-### Multiple Linear Regression
+### Simple Linear Regression
 Linear regression may be the best known algorithm used when analyzing a
@@
 the observations [@Sheather2009AMA]. This model will engage in stepwise feature
-using the AIC as the optimization metric.
+selection using the AIC as the optimization metric.
 
-```{r modelLinear1}
+```{r lm1Train}
 # Using stepAIC means no cross-validation. Train on entire dataset.
 trC <- trainControl(method = 'none')
-
-# Capture the unique country codes
 set.seed(181)
-linMod1 <- train(views ~ likes + dislikes + comments + tags + category +
-                   comments_disabled + ratings_disabled +
-                   video_error_or_removed + perfect + lkratio,
-                 data = usExtract, family = gaussian(link = 'identity'),
+lm1 <- train(x = seenX, y = seenY, family = gaussian(link = 'identity'),
              method = 'glmStepAIC', direction = 'both', trace = 0,
@@
 \footnotesize
-```{r m1Table}
-kable(summary(linMod1$finalModel)$coefficients,
+```{r lm1Table}
+kable(summary(lm1$finalModel)$coefficients,
       caption = "Model 1 Linear Regression Output",
@@
 
-The sign of these coefficients make sense in the main. The most predictive
-feature for views is `likes` followed by `dislikes`. Of course, one usually
-views a video at least once prior to rating it. We set the baseline category to
-`Entertainment`, considering it was the most popular. Therfore we expected
-negative coefficients for other categories. Interestingly, `Film & Animation`
-is positively correlated with likes over and above `Entertainment`, despite
-having fewer views.
+The signs of these coefficients make sense in the main. Increased `likes`,
+`dislikes` are correlated with increased views. Of course one usually views a
+video at least once prior to rating it. Increased `comments` is negatively
+correlated with views.
 
-Other observations are that the magnitude of the coefficients is similar to that
-of the intercept, indicating that the predictors provide large swings to final
-views. The "worst" predictor by magnitude is the `Nonprofits & Activism`
-category; its negative coefficient is more than twice the baseline intercept
-views. The best, interestingly is if ratings are disabled. This is
-counterintuitive at first, but is logical. The amount of `likes`, `dislikes`,
-and `comments` are many orders of magnitude greater than 1. Therefore, their
+We set the baseline category to `Entertainment`, considering it was the most
+popular. Therefore we expected negative coefficients for other categories. We
+were surprised that was not the case. The "worst" category predictor by
+magnitude is the `Nonprofits & Activism`. The "best" is `Gaming`.
+
+The factor predictors tend to have much higher magnitude coefficients than do
+the numeric ones. This makes sense. The amount of `likes`, `dislikes`, and
+`comments` are many orders of magnitude greater than 1. Therefore, their
 coefficients can be much smaller. A Boolean variable is either 1 or 0, therefore
@@
 
-Variable importance can be seen from the table below. The further down the list
-a coefficient preceded by a **`-`** is, the more important it is and the more it
-affected the AIC.
+### Generalized Linear Model: Gaussian Errors & Log Link
+The generalized linear model (GLM) is an extension of the simple linear model,
+but the errors can be distributed per any member of the exponential family and
+the relationship between some function of the predictors---called the link---and
+the mean needs to be linear, not that the mean itself must be linear in the
+predictors [@Faraway2006EtL].
+
+The model under consideration here assumes a Gaussian distribution of the
+errors, but a multiplicative relationship between the mean and the predictors.
+This is expressed by using a log link function. This is **not** the canonical
+link function for the Gaussian GLM, but as we are using numerical methods there
+is no issue.
+
+This approach *appears* similar to that of the common technique of performing a
+standard linear regression on the logs of the observations, but is different. As
+per [@Gelman2006Lta], one approach *"…log transforms observed values, while the*
+*second one log transforms the expected value.…the key difference being the*
+*relation between the predicted value and the variance."*
+
+```{r lm2Train}
+set.seed(181)
+lm2 <- train(x = seenX, y = seenY, family = gaussian(link = 'log'),
+             method = 'glmStepAIC', direction = 'both', trace = 0,
+             trControl = trC)
+```
+
 \footnotesize
-```
-                                  Df   Deviance    AIC
-<none>                               7.1243e+16 212209
-+ `categoryAutos & Vehicles`       1 7.1222e+16 212209
-+ tags                             1 7.1223e+16 212209
-- `categoryPeople & Blogs`         1 7.1278e+16 212210
-- categoryEducation                1 7.1278e+16 212210
-- perfectTRUE                      1 7.1279e+16 212210
-+ video_error_or_removedTRUE       1 7.1237e+16 212210
-+ `categoryPets & Animals`         1 7.1239e+16 212210
-+ categoryGaming                   1 7.1239e+16 212210
-+ `categoryTravel & Events`        1 7.1240e+16 212210
-+ comments_disabledTRUE            1 7.1240e+16 212210
-+ categorySports                   1 7.1240e+16 212210
-+ `categoryScience & Technology`   1 7.1241e+16 212211
-+ categoryShows                    1 7.1243e+16 212211
-- `categoryHowto & Style`          1 7.1287e+16 212211
-- `categoryFilm & Animation`       1 7.1299e+16 212212
-- `categoryNews & Politics`        1 7.1321e+16 212214
-- categoryMusic                    1 7.1338e+16 212215
-- `categoryNonprofits & Activism`  1 7.1345e+16 212216
-- ratings_disabledTRUE             1 7.1463e+16 212227
-- categoryComedy                   1 7.1467e+16 212227
-- lkratio                          1 7.3555e+16 212413
-- dislikes                         1 8.9091e+16 213650
-- comments                         1 8.9807e+16 213701
-- likes                            1 2.0217e+17 218939
+```{r lm2Table}
+kable(summary(lm2$finalModel)$coefficients,
+      caption = "Model 3 Linear Regression Output",
+      digts = 3L, format.args = list(big.mark = ','))
 ```
@@
 
+### ElasticNet: Penalized Regression
+Instead of using AIC to select features, one can make use of penalized
+regression. Using an \(L_1\) penalty is the underpinnings of Lasso regression,
+which can perform feature selection. Using a squared error term, \(L_2\), is at
+the heart of ridge regression [@HTF2009TEo]. Using both methods together is
+called the elastic net [@ZH2005RaV]. To tune the hyperparameters, which includes
+the weighting between Lasso and Ridge, we will use 10-fold cross-validation.
+
+```{r lm3Train}
+trC <- trainControl(method = 'cv', number = 10L)
+tG <- expand.grid(fraction = seq(0.8, 1, 0.001),
+                  lambda = seq(0, 0.02, 0.001))
+set.seed(181)
+lm3 <- train(x = seenX, y = seenY,
+             method = 'enet', trControl = trC, tuneGrid = tG)
+```
+
+There is no clean table of coefficients with elasticNet. Rather there is a
+sequence of models built behind the scenes. The coefficients of the selected
+model can be found in the appendix.
+
 ## Model Evaluation
+
+```{r modTest}
+lm1P <- predict(lm1, newdata = hideX)
+lm2P <- predict(lm2, newdata = hideX)
+lm3P <- predict(lm3, newdata = hideX)
+compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet'),
+                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2P, hideY),
+                                 RMSE(lm3P, hideY)),
+                        R2 = c(R2(lm1P, hideY, formula = 'traditional'),
+                               R2(lm2P, hideY, formula = 'traditional'),
+                               R2(lm3P, hideY, formula = 'traditional')),
+                        MAE = c(MAE(lm1P, hideY), MAE(lm2P, hideY),
+                                MAE(lm3P, hideY)))
+kable(compTable, digits = 3L, format.args = list(big.mark = ','),
+      caption = "Model Performance on Test Set")
+```
 
@@
 ```
-```{r modelLinear1}
+```{r lm1Train}
 ```
-```{r m1Table}
+```{r lm1Table}
+```
+```{r lm2Train}
+```
+```{r lm2Table}
+```
+```{r lm3Train}
+```
+
+<!-- Model Evaluation -->
+```{r modTest}
 ```

### 2020-12-03 Final_project/Final Project Paper - Group 2.Rmd
More fixes, restor unam

@@
 colClass <- c(rep('character', 4L), 'integer', 'POSIXct', 'character',
-              rep('double', 4L), rep('boolean', 3L))
+              rep('double', 4L), rep('logical', 3L))
 us_set <- fread(us_path, encoding = 'UTF-8', colClasses = colClass)
@@
 
-# Change data types for convenience, add target bucket, count tags and then
-# get rid of actual tags and category IDs
+# Change data types for convenience, add target bucket, count tags and then get
+# rid of actual tags and category IDs
 
@@
 which has a greater effect. As there may be some rare videos with no dislikes,
-the ratios will be set to 1.
+the ratios will be set to 1. The variable `unan`, short for unanimous, will
+reflect that some videos have likes but no dislikes, to balance the ratio of 1
+which should have been `NaN`, a division by 0 error.
 
 ```{r addFeatures}
-usExtract[, lkratio := ifelse(dislikes == 0, 1, likes / dislikes)]
+usExtract[, `:=`(lkratio = ifelse(dislikes == 0, 1, likes / dislikes),
+                 unan = ifelse(dislikes == 0 & likes > 0, TRUE, FALSE))]
 
@@
                    comments_disabled + ratings_disabled +
-                   video_error_or_removed + lkratio,
+                   video_error_or_removed + lkratio + unan,
                    data = usExtract, fullRank = TRUE)
@@
 
-## Model Building
+## Model Building & Interpretation
 We will first separate 20% of the data as a true holdout set. It is on this data
@@
 
-The signs of these coefficients make sense in the main. Increased `likes`,
+The signs of these coefficients make sense in the main. Increased `likes` and
 `dislikes` are correlated with increased views. Of course one usually views a
@@
 We set the baseline category to `Entertainment`, considering it was the most
-popular. Therefore we expected negative coefficients for other categories. We
-were surprised that was not the case. The "worst" category predictor by
-magnitude is the `Nonprofits & Activism`. The "best" is `Gaming`.
+popular. Therefore we expected negative coefficients for other categories found
+significant. We were surprised that `Film & Animation` was an exception. The
+"worst" category predictor by magnitude is the `Nonprofits & Activism`.
 
@@
 
-## Findings
-
 # Discussion & Conclusions

### 2020-12-03 Final_project/Final Project Paper - Group 2.Rmd
Stopping for the night

@@
 
+What also seems counterintuitive at first are the negative correlations between
+`lkratio` and `unan` being true with views. It may well be that conflict is good
+for Youtube videos. Perhaps the more a video is talked about or argued over, the
+more views it will get!
+
 ### Generalized Linear Model: Gaussian Errors & Log Link
@@
 
+This model has more predictors than does the simple Gaussian (G)LM, although
+their signs are similar. What stands out is how many more categories were
+considered significant.
+
 ### ElasticNet: Penalized Regression
@@
 
+```{r lm3Table}
+# The only way to extract the coefficients from an elastic net model in R
+# Is to predict them using the optimal tuning values
+
+lm3C <- predict(lm3$finalModel, s = lm3$finalModel$tuneValue[[1]],
+                type = 'coefficients', mode = 'fraction')
+lm3CC <- lm3C$coefficients[lm3C$coefficients != 0]
+kable(lm3CC, digits = 3L, format.args = list(big.mark = ','),
+      caption = 'Coefficients of "Optimal" Elastic Net Model')
+```
+
 There is no clean table of coefficients with elasticNet. Rather there is a
 sequence of models built behind the scenes. The coefficients of the selected
-model can be found in the appendix.
+model can be found through predicting them, but there are no corresponding
+p- or Z-values. Worse, there is no intercept returned.
+
+Nevertheless, the *relative* magnitude and sign of the parameters are in line
+with our expectations. `Nonprofits & Activism` has the largest factor effect, to
+the worse. Both `likes` and `dislikes` are positively correlated with views and
+`comments` is negatively correlated. A few more categories join
+`Film & Animation` as contributing to excess views over the baseline
+`Entertainment`. Lastly, having a preponderance of likes tends to reduce views.
+Conflict seems to be good for Youtubers!
 

### 2020-12-04 Final_project/Final Project Paper - Group 2.Rmd
Big changes. Move single-date extraction into data exploration section, move creation of "exceed mean" feature to model building, put dummyVar creation AFTER train/test split in case someone wants to use exceedMean in modeling. I think it's flawed. You're using views to predict views.

@@
 relationships between a video’s views and the number of likes, dislikes, and
-comments. We may also use a video’s country of origin and the category or type
-of video as predictors.
+comments using R [@RCT2020RAL]. We may also use a video’s country of origin and
+the category or type of video as predictors.
 
@@
 ## Data Exploration
+Since we are not factoring in time, it is incorrect to use all the observations.
+Therefore, we will extract the latest observation by video by country and use
+this subset.
 
 ```{r dataPrep}
+# Select the row numbers of the first entry of the latest trending date by title
+# and by country. There are 34 duplicates. Use this as our restricted data set
+# corresponding to the most recent view count. The inner set of brackets gets
+# the row number (called V1) and the outer set is a simple subset by those row
+# numbers.
+usExtract <- us_set[
+  us_set[, .I[which.max(views)], by = c('title')]$V1
+  ]
+nobsUSExt <- dim(usExtract)[[1]]
+
 # Substitute category for numeric ID through joining
-us_set <- us_cats[us_set, on = 'id == category_id']
+usExtract <- us_cats[usExtract, on = 'id == category_id']
 
-# Change data types for convenience, add target bucket, count tags and then get
-# rid of actual tags and category IDs
-
-us_mean <- mean(us_set$views)
-
-us_set[, `:=`(trending_date = as.IDate(trending_date, format = "%y.%d.%m"),
+# Change data types for convenience, count tags and then get rid of actual tags
+# and replace category IDs with actual categories
+usExtract[, `:=`(trending_date = as.IDate(trending_date, format = "%y.%d.%m"),
                  views = as.double(views),
-              tag_count = 1L + str_count(us_set$tags, '[|]'),
-              view_target = factor(ifelse(views > us_mean, 1L, 0L),
-                                   levels = c(0, 1), labels = c(0, 1)),
+                 tag_count = 1L + str_count(tags, '[|]'),
                  id = NULL,
@@
 # Make name shorter for display purposes
-setnames(us_set, c('comment_count', 'tag_count'), c('comments', 'tags'))
+setnames(usExtract,
+         c('comment_count', 'tag_count', 'comments_disabled',
+           'ratings_disabled', 'video_error_or_removed'),
+         c('comments', 'tags', 'cmtDisabled', 'rtgDisabled', 'vidError'))
 
 # Categorize the variable names
-boolVars <- c('comments_disabled', 'ratings_disabled', 'video_error_or_removed')
 numVars <- c('views', 'likes', 'dislikes', 'comments', 'tags')
-facVars <- c('category', 'view_target')
-dateVars <- c('trending_date', 'publish_time')
-charVars <- c('video_id', 'title', 'channel_title')
 
 # Melt for numerics
-usSetN <- melt(us_set, measure.vars = numVars, variable.factor = FALSE,
+usSetN <- melt(usExtract, measure.vars = numVars, variable.factor = FALSE,
                variable.name = 'measure', value.name = 'value')
 
-# Melt for factors
-usSetF <- melt(us_set,
-               measure.vars = c('category', 'view_target'),
-               variable.factor = FALSE, variable.name = 'measure',
-               value.name = 'value')
+# Melt for category
+usSetF <- melt(usExtract, measure.vars = 'category', variable.factor = FALSE,
+               variable.name = 'measure', value.name = 'value')
 
@@
 
-### Factor Predictors
+### Categories
 While tabular representation of factor predictors is difficult, a distribution
-of factors by value may prove informative. It is clear from the graphs below
+of videos by categories may prove informative. It is clear from the graphs below
 that interests vary by category. `Entertainment` seems to be the most common in
@@
 
-```{r factorPlot1, fig.height=4.5, fig.width=8}
+```{r catPlot, fig.height=4.5, fig.width=8}
 ggplot(usSetF[measure == 'category'], aes(y = value)) +
@@
 
-```{r factorPlot2, fig.height=3, fig.width=8}
-ggplot(usSetF[measure != 'category'], aes(x = value, fill = measure)) +
-  geom_bar(position = 'dodge2') +
-  ggtitle("Country and Global Mean Views Exceedence") +
+## Feature Selection & Engineering
+We will consider the relationship between `views` and the numerical predictors
+of `likes`, `dislikes`, `comments`, and `tags`. We will also consider the
+`category`, whether or not comments or ratings were disabled and, if there was
+an error with the video.
+
+```{r likeDislikeCheck}
+ldOutlier <- usExtract[!rtgDisabled & dislikes == 0 & likes == 0]
+ldOn <- dim(ldOutlier)[[1]]
+```
+
+We will add two features. The first is `disp`: the disproportion between the
+likes and dislikes. To minimize division by 0 errors, it is defined as follows:
+\[
+disp =\begin{cases}
+\mathrm{ratings\;enabled}\quad\frac{\min(likes, dislikes)}{\max(likes, dislikes)}\\
+\mathrm{ratings\;disabled}\quad 1
+\end{cases}
+\]
+This constrains the ratio to \([0 \leq disp \leq 1]\) with a maximum when the
+two are equal, and division by 0 only when both are 0. This usually occurs when
+ratings are disabled. There is only `r ldOn` case out of the `r nobsUSExt`
+observations where the ratings were not disabled, yet there are neither likes
+nor dislikes. As this is a distinct incongruity for YouTube, we will remove that
+one observation from the data. When ratings are disabled, perforce there is no
+disparity so the ratio will be set to 1.
+
+The second is `exceedMean`. This will be a factor if the views for the video
+are greater than the *mean* views for all of the extracted data set. This will
+also indicate the skewness of the data. Were it symmetrical, and we know it is
+not from the graphs above, the mean should be near the median and this factor
+should be close to equal for both 0 and 1.
+
+```{r addFeatures}
+# Remove the outlier
+usExtract <- usExtract[!(!rtgDisabled & dislikes == 0 & likes == 0)]
+
+# Precaclulate the mean views
+usMean <- mean(usExtract$views)
+
+# Add the engineered feature
+usExtract[, `:=`(disp = ifelse(rtgDisabled, 1,
+                               pmin(likes, dislikes) / pmax(likes, dislikes)),
+                 exceedMean = factor(ifelse(views > usMean, 1, 0),
+                                     levels = c(0, 1), labels = c(0, 1)))]
+# Melt for exceedMean
+usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE,
+               variable.name = 'measure', value.name = 'value')
+
+# Plot the exceedMean variable
+ggplot(usSetF[measure == 'exceedMean'], aes(x = value)) +
+  geom_bar() +
+  ggtitle("Proportion Above and Below Mean") +
   scale_y_continuous(labels = comma)
@@
 
-## Feature Engineering & Selection
-Since we are not factoring in time, it is incorrect to use all the observations.
-Therefore, we will extract the latest observation by video by country and use
-this subset.
-
-```{r subsetExtraction}
-# Select the row numbers of the first entry of the latest trending date by title
-# and by country. There are 34 duplicates. Use this as our restricted data set
-# corresponding to the most recent view count. The inner set of brackets gets
-# the row number (called V1) and the outer set is a simple subset by those row
-# numbers.
-usExtract <- us_set[
-  us_set[, .I[which.max(views)], by = c('title')]$V1
-  ]
-```
-
-We will consider the relationship between `views` and the numerical predictors
-of `likes`, `dislikes`, `comments`, and `tags`. We will also consider the
-`category`, whether or not comments or ratings were disabled and, if there was
-an error with the video.
-
-A feature that we will also consider is `lkratio`: the ratio of likes to
-dislikes. This will help ascertain if it is the magnitude or the preponderance
-which has a greater effect. As there may be some rare videos with no dislikes,
-the ratios will be set to 1. The variable `unan`, short for unanimous, will
-reflect that some videos have likes but no dislikes, to balance the ratio of 1
-which should have been `NaN`, a division by 0 error.
-
-```{r addFeatures}
-usExtract[, `:=`(lkratio = ifelse(dislikes == 0, 1, likes / dislikes),
-                 unan = ifelse(dislikes == 0 & likes > 0, TRUE, FALSE))]
-
-modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
-                   comments_disabled + ratings_disabled +
-                   video_error_or_removed + lkratio + unan,
-                   data = usExtract, fullRank = TRUE)
-modMat <- predict(modDum, usExtract)
-```
-
 ## Model Building & Interpretation
@@
 
-```{r holdOut}
+```{r holdout}
+# Create seen and hidden sets
 set.seed(617)
 seenIDX <- createDataPartition(usExtract$views, p = 0.8)$Resample1
-seenX <- modMat[seenIDX, ]
-seenY <- usExtract$views[seenIDX]
-hideX <- as.data.frame(modMat[-seenIDX, ])
-hideY <- usExtract$views[-seenIDX]
+seenSet <- usExtract[seenIDX, ]
+hideSet <- usExtract[-seenIDX, ]
 ```
@@
 predictors that minimizes the squared error between the "predictor" function and
-the observations [@Sheather2009AMA]. This model will engage in stepwise feature
-selection using the AIC as the optimization metric.
+the observations [@Sheather2009AMA].
+
+This model will start with the numeric and logical features and the engineered
+disparity ratio. It will not use the exceeding mean indicator, as that is
+generated from the target variable, and it is felt it will distort the
+prediction. It will then proceed through feature selection using the AIC as the
+optimization metric.
+
+```{r dummyVarsAA}
+# Create dummy Variables. These will be used for the next three models
+modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
+                   cmtDisabled + rtgDisabled + vidError + disp,
+                   data = seenSet, fullRank = TRUE)
+seenX <- predict(modDum, seenSet)
+seenY <- seenSet$views
+```
 
@@
 the heart of ridge regression [@HTF2009TEo]. Using both methods together is
-called the elastic net [@ZH2005RaV]. To tune the hyperparameters, which includes
-the weighting between Lasso and Ridge, we will use 10-fold cross-validation.
+called the elastic net [@ZH2005RaV; ZH2020eEN]. To tune the hyperparameters,
+which includes the weighting between Lasso and Ridge, we will use 10-fold
+cross-validation.
 
@@
 trC <- trainControl(method = 'cv', number = 10L)
-tG <- expand.grid(fraction = seq(0.8, 1, 0.001),
-                  lambda = seq(0, 0.02, 0.001))
+tG <- expand.grid(fraction = seq(0.85, 1, 0.001),
+                  lambda = seq(0.01, 0.025, 0.001))
 set.seed(181)
@@
 ```{r modTest}
+tstDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
+                   cmtDisabled + rtgDisabled + vidError + disp,
+                   data = hideSet, fullRank = TRUE)
+hideX <- predict(tstDum, hideSet)
+hideY <- hideSet$views
 lm1P <- predict(lm1, newdata = hideX)
@@
 # Discussion & Conclusions
+The three models contain both intuitive and counter-intuitive results. As
+expected, videos with more ratings tend to have more views. However, as likes
+outnumber dislikes, that tends to reduce the number of views. Having a video
+witk
 
@@
 ```
+<!-- Data Exploration -->
 ```{r dataPrep}
@@
 ```
-```{r factorPlot1}
+```{r catPlot}
 ```
-```{r factorPlot2}
+```{r subsetExtraction}
 ```
-
-<!-- Feature Engineering & Selection -->
-```{r subsetExtraction}
+```{r likeDislikeCheck}
 ```
@@
 <!-- Model Building -->
-```{r holdOut}
+```{r holdout}
+```
+```{r dummyVarsAA}
 ```

### 2020-12-04 Final_project/Final Project Paper - Group 2.Rmd
Plot the two created features and update the text to match the new results

@@
 relationships between a video’s views and the number of likes, dislikes, and
-comments using R [@RCT2020RAL]. We may also use a video’s country of origin and
-the category or type of video as predictors.
+comments using `R` [@RCT2020RAL] and the `caret` package [@Kuhn2020cCa]. We may
+also use a video’s category as predictors.
 
@@
 
-We will add two features. The first is `disp`: the disproportion between the
-likes and dislikes. To minimize division by 0 errors, it is defined as follows:
+We will add two features. The first is `balance`: the ratio between the
+likes and dislikes. The hypothesis is that if a video is either universally
+loved or panned, it will get fewer views than if there is a healthy disagreement
+about it.
+
+To minimize division by 0 errors, `balance` is defined as follows:
 \[
-disp =\begin{cases}
+balance =\begin{cases}
 \mathrm{ratings\;enabled}\quad\frac{\min(likes, dislikes)}{\max(likes, dislikes)}\\
@@
 \]
-This constrains the ratio to \([0 \leq disp \leq 1]\) with a maximum when the
-two are equal, and division by 0 only when both are 0. This usually occurs when
-ratings are disabled. There is only `r ldOn` case out of the `r nobsUSExt`
-observations where the ratings were not disabled, yet there are neither likes
-nor dislikes. As this is a distinct incongruity for YouTube, we will remove that
-one observation from the data. When ratings are disabled, perforce there is no
-disparity so the ratio will be set to 1.
+This constrains the ratio to the interval \([0, 1]\) with a maximum of 1 when
+the two are equal. Now, division by 0 can only occure when both are 0. This 
+usually occurs when ratings are disabled. There is only `r ldOn` case out of the
+`r nobsUSExt` observations where the ratings were not disabled, yet there are
+neither likes nor dislikes. As this is a distinct incongruity for YouTube, we
+will remove that one observation from the data. When ratings are disabled,
+perforce there is no disparity so the ratio will be set to 1.
 
@@
 # Add the engineered feature
-usExtract[, `:=`(disp = ifelse(rtgDisabled, 1,
+usExtract[, `:=`(balance = ifelse(rtgDisabled, 1,
                                pmin(likes, dislikes) / pmax(likes, dislikes)),
@@
                                      levels = c(0, 1), labels = c(0, 1)))]
+```
+```{r plotBalance, fig.height=4, fig.width=8}
+ggplot(usExtract[, .(balance)], aes(x = balance)) +
+  geom_density() +
+  ggtitle("Distribution of Balance")
+```
+
+It's pretty clear from the distribution of `balance` that, at least for videos
+viewed in the US, there is a healthy dose of disagreement!
+
+```{r plotExceedMean, fig.height=4, fig.width=8}
 # Melt for exceedMean
@@
 generated from the target variable, and it is felt it will distort the
-prediction. It will then proceed through feature selection using the AIC as the
-optimization metric.
+prediction to regress `views` on a function of `views`. The algorithm will
+proceed through feature selection using the AIC as the optimization metric.
 
@@
 `dislikes` are correlated with increased views. Of course one usually views a
-video at least once prior to rating it. Increased `comments` is negatively
-correlated with views.
+video at least once prior to rating it. Interestingly, increased `comments` are
+negatively correlated with views.
 
@@
 significant. We were surprised that `Film & Animation` was an exception. The
-"worst" category predictor by magnitude is the `Nonprofits & Activism`.
+"worst" category predictor by magnitude is clearly `Nonprofits & Activism`.
 
@@
 
-What also seems counterintuitive at first are the negative correlations between
-`lkratio` and `unan` being true with views. It may well be that conflict is good
-for Youtube videos. Perhaps the more a video is talked about or argued over, the
-more views it will get!
+When ratings are disabled, there is a bump to views. This is probably because
+the intercept is artificially low due to the predictive power of `likes` and
+`dislikes`. To balance that when there are none needs a big boost.
+
+What may be most interesting is that the disparity ratio indicator us a very
+powerful and significant indicator. Conflict seems to be good for Youtube
+videos. The more a video is argued over, the more views it seems to get!
 
@@
 
-This model has more predictors than does the simple Gaussian (G)LM, although
-their signs are similar. What stands out is how many more categories were
-considered significant.
+This model has more predictors than does the simple Gaussian (G)LM. Moreover,
+the signs are different from the simple LM. The very fact that
+`Nonprofits & Activism` has a positive coefficient should raise concerns. This
+model is probably a poor one.
 
@@
 the heart of ridge regression [@HTF2009TEo]. Using both methods together is
-called the elastic net [@ZH2005RaV; ZH2020eEN]. To tune the hyperparameters,
+called the elastic net [@ZH2005RaV; @ZH2020eEN]. To tune the hyperparameters,
 which includes the weighting between Lasso and Ridge, we will use 10-fold
@@
 Nevertheless, the *relative* magnitude and sign of the parameters are in line
-with our expectations. `Nonprofits & Activism` has the largest factor effect, to
-the worse. Both `likes` and `dislikes` are positively correlated with views and
-`comments` is negatively correlated. A few more categories join
+with our expectations. Both `likes` and `dislikes` are positively correlated
+with views and `comments` is negatively correlated. A few more categories join
 `Film & Animation` as contributing to excess views over the baseline
-`Entertainment`. Lastly, having a preponderance of likes tends to reduce views.
+`Entertainment`. But `Nonprofits & Activism` has the largest factor by far, and
+it is negative, which stands in contrast to the second model. Lastly, having a
+larger disparity (ratio of likes to dislikes closer to 1of likes tends to reduce views.
 Conflict seems to be good for Youtubers!
@@
 ```
+```{r plotBalance}
+```
+```{r plotExceedMean}
+```
 

### 2020-12-04 Final_project/Final Project Paper - Group 2.Rmd
Reached a decent stopping point.

@@
     Bryan Persaud, Jeff Shamp
-date: "12/14/2020"
+date: "12/11/2020"
 output:
@@
       setspace: singlespacing
-    toc: FALSE
+    toc: TRUE
     toc_depth: 3
@@
 Using forms of linear regression we have covered this semester, we will test various combinations of features for predictive power.
-[FILL IN OUTCOMES HERE]\\par\\textbf{Keywords:} Youtube, linear regression, count regression, R"
+YouTubers are advised to allow viewers to interact via ratings and to post content which will engender heated discussion.
+\\par\\textbf{Keywords:} Youtube, linear regression, elastic net, R"
 ---
@@
 video ecosystem,” says Enders Analysis research analyst Jamie McGowan Stuart
-[@Foster2020Yt1]. The current top channel “Vlad and Nikita” earns around \$312
-thousand per video [@TYL]. According to a survey from Google, 6 out of 10 people
-already prefer online video platforms or streaming services over live TV
+[@Foster2020Yt1]. The current top channel “Vlad and Nikita” earns around
+\$312,000 per video [@TYL]. According to a survey from Google, 6 out of 10
+people already prefer online video platforms or streaming services over live TV
 [@OBTlv]. There are researchers predicting that by 2025---four years now---half
@@
 [@LEZ2019YVP]. Others leveraged NLP and specially engineered features such as
-"clickbait" or "NSFW" tags [@Srinivasan2017YVP]. These attempts usually used the
-same dataset as we are.
+"clickbait" or "NSFW" tags [@Srinivasan2017YVP]. These attempts often used the
+same dataset as we are using.
 
@@
 This constrains the ratio to the interval \([0, 1]\) with a maximum of 1 when
-the two are equal. Now, division by 0 can only occure when both are 0. This 
+the two are equal. Now, division by 0 can only occur when both are 0. This 
 usually occurs when ratings are disabled. There is only `r ldOn` case out of the
@@
 modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
-                   cmtDisabled + rtgDisabled + vidError + disp,
+                   cmtDisabled + rtgDisabled + vidError + balance,
                    data = seenSet, fullRank = TRUE)
@@
 ```{r lm3Table}
-# The only way to extract the coefficients from an elastic net model in R
-# Is to predict them using the optimal tuning values
+# The only way to extract the coefficients from an elastic net model in R is to
+# predict them using the optimal tuning values
 
@@
 it is negative, which stands in contrast to the second model. Lastly, having a
-larger disparity (ratio of likes to dislikes closer to 1of likes tends to reduce views.
-Conflict seems to be good for Youtubers!
+higher balance---ratio of likes to dislikes closer to 1---tends to increase
+views. Conflict seems to be good for YouTubers!
 
@@
 ```{r modTest}
+# Process the hidden set
 tstDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
-                   cmtDisabled + rtgDisabled + vidError + disp,
+                   cmtDisabled + rtgDisabled + vidError + balance,
                    data = hideSet, fullRank = TRUE)
@@
 hideY <- hideSet$views
+
+# Predict using the models
 lm1P <- predict(lm1, newdata = hideX)
@@
 lm3P <- predict(lm3, newdata = hideX)
+
+# Compare the results
 compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet'),
@@
 expected, videos with more ratings tend to have more views. However, as likes
-outnumber dislikes, that tends to reduce the number of views. Having a video
-witk
+outnumber dislikes, that tends to reduce the number of views. All the models
+agree that having a video whose likes and dislikes are close in magnitude
+increases the propensity for views.
 
+The elasticNet model performed best, and it's clear that the log link is
+inferior to the identity link for this data set. If there is any clear takeaways
+for prospective YouTubers, it would be:
+
+  * Make sure your videos will engender heated discussions. You want as many
+  *dislikes* as likes. That seems to drive attention to your video!
+  * Keep your ratings enabled. Allowing users to rate your video drives
+  attention to them!
+  * Perhaps disable comments, though.
+  * Stay **away** from activism. 
+  * Talk about films or comics. This probably includes the anime and manga
+  phenomena.
+  * Having more tags is better than having fewer.
+
+The main limitation of this analysis is that the linear model, even with
+penalization, is probably not the best model for human behavior. There are
+likely non-linear effects and tipping points which could be much better captured
+by tree-based models or even SVMs with the appropriate kernel.
+
+Also, this analysis may be valuable for someone planning a video, but it does
+not help them react once the video is posted, as we did not investigate any
+changes over time.
+
+Lastly, we encountered problems trying to aggregate disjoint data by country
+without falling prey to gross overcounting.
+
+Areas of future work would include investigating more sophisticated algorithms,
+analyzing not only final views but growth patterns, and either finding better
+sources or methods to increase the available data to include other countries
+without overcounting.
+
+\clearpage
 # Appendix
 ## Statistical Appendix
+### Table of Numeric Predictors
 A more detailed analysis of the empirical statistics for the numeric data is 
@@
 ```{r statsTable}
-kable(statsN, digits = 2,
-      caption = "Statistics for Numeric Variables by Country & Global",
+# Empirical stats Table
+kable(statsN, digits = 2, caption = "Table of Numeric Predictors",
       format.args = list(big.mark = ","))
@@
 \normalsize
+
+### Variable Importance: Model 1
+
+```{r lm1VI}
+# Model 1: VarImp 
+ggplot(varImp(lm1)) + ggtitle('Variable Importance: Model 1')
+```
+
+### Variable Importance: Model 2
+
+```{r lm2VI}
+# Model 2: VarImp 
+ggplot(varImp(lm2)) + ggtitle('Variable Importance: Model 2')
+```
+
+### Variable Importance: Model 3
+```{r lm3VI}
+# Model 3: VarImp 
+ggplot(varImp(lm3)) + ggtitle('Variable Importance: Model 3')
+```
+
+### Hyperparameter Tuning: Model 3
+
+```{r lm3Plot}
+# Model 3: Tuning paths
+ggplot(lm3) + ggtitle('Tuning Parameters: Model 3')
+```
 
@@
 ```
+```{r lm1VI}
+```
+```{r lm2VI}
+```
+```{r lm2VI}
+```
+```{r lm3Plot}
+```
 
 <!-- References needs to be last -->
+\clearpage
 # References

### 2020-12-11 Final_project/Final Project Paper - Group 2.Rmd
add another GLM, remove exceed mean feature, minor text additions

@@
 
-The second is `exceedMean`. This will be a factor if the views for the video
-are greater than the *mean* views for all of the extracted data set. This will
-also indicate the skewness of the data. Were it symmetrical, and we know it is
-not from the graphs above, the mean should be near the median and this factor
-should be close to equal for both 0 and 1.
+<!-- The second is `exceedMean`. This will be a factor if the views for the video -->
+<!-- are greater than the *mean* views for all of the extracted data set. This will -->
+<!-- also indicate the skewness of the data. Were it symmetrical, and we know it is -->
+<!-- not from the graphs above, the mean should be near the median and this factor -->
+<!-- should be close to equal for both 0 and 1.  -->
+
+The second is `engagement`. This will be the ratio of comments to sum of likes and 
+dislikes, which should give us some measure of comments to ratings. We will use a
+similar approach as `balance` in terms of constraining the interval to \([0,1])\).
 
@@
                                pmin(likes, dislikes) / pmax(likes, dislikes)),
-                 exceedMean = factor(ifelse(views > usMean, 1, 0),
-                                     levels = c(0, 1), labels = c(0, 1)))]
+                 engagement = ifelse(rtgDisabled, 1, 
+                                     comments / sum(likes, dislikes)))]
+                 # exceedMean = factor(ifelse(views > usMean, 1, 0),
+                 #                     levels = c(0, 1), labels = c(0, 1)))
+
 ```
@@
   ggtitle("Distribution of Balance")
+
+ggplot(usExtract[, .(engagement)], aes(x = engagement)) +
+  geom_density() +
+  scale_x_log10() +
+  ggtitle("Distribution of Engagement Log Scale")
+
 ```
@@
 
-```{r plotExceedMean, fig.height=4, fig.width=8}
-# Melt for exceedMean
-usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE,
-               variable.name = 'measure', value.name = 'value')
+<!-- ```{r plotExceedMean, fig.height=4, fig.width=8} -->
+<!-- # Melt for exceedMean -->
+<!-- usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE, -->
+<!--                variable.name = 'measure', value.name = 'value') -->
 
-# Plot the exceedMean variable
-ggplot(usSetF[measure == 'exceedMean'], aes(x = value)) +
-  geom_bar() +
-  ggtitle("Proportion Above and Below Mean") +
-  scale_y_continuous(labels = comma)
-```
+<!-- # Plot the exceedMean variable -->
+<!-- ggplot(usSetF[measure == 'exceedMean'], aes(x = value)) + -->
+<!--   geom_bar() + -->
+<!--   ggtitle("Proportion Above and Below Mean") + -->
+<!--   scale_y_continuous(labels = comma) -->
+<!-- ``` -->
 
-As with many events which follow the Pareto rule, most of the views are 
-attributable to a minority of the videos, as seen by the 0/1 disparity.
+<!-- As with many events which follow the Pareto rule, most of the views are  -->
+<!-- attributable to a minority of the videos, as seen by the 0/1 disparity. -->
 
@@
 modDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
-                   cmtDisabled + rtgDisabled + vidError + balance,
+                   cmtDisabled + rtgDisabled + vidError + balance + engagement,
                    data = seenSet, fullRank = TRUE)
@@
 
-### Generalized Linear Model: Gaussian Errors & Log Link
+### Generalized Linear Model: Gaussian and Poisson Errors with Log Link
 The generalized linear model (GLM) is an extension of the simple linear model,
@@
 
-The model under consideration here assumes a Gaussian distribution of the
-errors, but a multiplicative relationship between the mean and the predictors.
-This is expressed by using a log link function. This is **not** the canonical
-link function for the Gaussian GLM, but as we are using numerical methods there
-is no issue.
+The models under consideration here assume either a Poisson or Gaussian 
+distribution of the errors, but a multiplicative relationship between the
+mean and the predictors. This is expressed by using a log link function. 
+This is **not** the canonical link function for the GLM distributions,
+but as we are using numerical methods there is no issue.
 
@@
 
+```{r lm3Train_poi}
+set.seed(9450)
+lm3_test<- train(x = seenX, y = seenY, family = poisson(link='log'),
+            method = 'glmStepAIC', direction = 'both', trace = 0,
+            trControl = trC)
+```
+
 \footnotesize
-```{r lm2Table}
-kable(summary(lm2$finalModel)$coefficients,
-      caption = "Model 3 Linear Regression Output",
+```{r poi_coeff}
+kable(summary(lm3_test$finalModel)$coefficients,
+      caption = "Model 3 (Poisson) Linear Regression Output",
       digts = 3L, format.args = list(big.mark = ','))
@@
 
-This model has more predictors than does the simple Gaussian (G)LM. Moreover,
+\footnotesize
+```{r lm2Table}
+kable(summary(lm2$finalModel)$coefficients,
+      caption = "Model 3 (Gaussian) Linear Regression Output",
+      digts = 3L, format.args = list(big.mark = ','))
+```
+\normalsize
+
+These models have more predictors than the simple Gaussian (G)LM. Moreover,
 the signs are different from the simple LM. The very fact that
-`Nonprofits & Activism` has a positive coefficient should raise concerns. This
-model is probably a poor one.
+`Nonprofits & Activism` has a positive coefficient should raise concerns. These
+models are probably poor ones.
 
@@
 tstDum <- dummyVars(views ~ likes + dislikes + comments + tags + category +
-                   cmtDisabled + rtgDisabled + vidError + balance,
+                   cmtDisabled + rtgDisabled + vidError + balance + engagement,
                    data = hideSet, fullRank = TRUE)
@@
 lm3P <- predict(lm3, newdata = hideX)
-
+lm3_testP <- predict(lm3_test, newdata = hideX)
 # Compare the results
-compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet'),
+compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet', 'GLM: Poisson+Log '),
                         RMSE = c(RMSE(lm1P, hideY), RMSE(lm2P, hideY),
-                                 RMSE(lm3P, hideY)),
+                                 RMSE(lm3P, hideY), RMSE(lm3_testP, hideY)),
                         R2 = c(R2(lm1P, hideY, formula = 'traditional'),
                                R2(lm2P, hideY, formula = 'traditional'),
-                               R2(lm3P, hideY, formula = 'traditional')),
+                               R2(lm3P, hideY, formula = 'traditional'),
+                               R2(lm3_testP, hideY, formula = 'traditional')),
                         MAE = c(MAE(lm1P, hideY), MAE(lm2P, hideY),
-                                MAE(lm3P, hideY)))
+                                MAE(lm3P, hideY), MAE(lm3_testP, hideY)))
 kable(compTable, digits = 3L, format.args = list(big.mark = ','),

### 2020-12-13 Final_project/Final Project Paper - Group 2.Rmd
copyediting

@@
 For numeric predictors, we are using the number of `likes`, `dislikes`, `tags`,
-and `comments`. We may also consider the country of viewing as well. From the 
-There is not that much difference in the distribution of the numeric predictors
-between the countries where people view.
+and `comments`.
 
@@
   geom_bar() +
-  ggtitle("Distribution by Category") +
+  ggtitle("Distribution by Category: United States") +
   scale_x_continuous(labels = comma)

### 2020-12-13 Final_project/Final Project Paper - Group 2.Rmd
Clean up and reorder

@@
 
-<!-- The second is `exceedMean`. This will be a factor if the views for the video -->
-<!-- are greater than the *mean* views for all of the extracted data set. This will -->
-<!-- also indicate the skewness of the data. Were it symmetrical, and we know it is -->
-<!-- not from the graphs above, the mean should be near the median and this factor -->
-<!-- should be close to equal for both 0 and 1.  -->
-
-The second is `engagement`. This will be the ratio of comments to sum of likes and 
-dislikes, which should give us some measure of comments to ratings. We will use a
-similar approach as `balance` in terms of constraining the interval to \([0,1])\).
+The second is `engagement`. This will be the ratio of comments to sum of likes
+and dislikes, which should give us some measure of comments to ratings. We will
+use a similar approach as `balance` in terms of constraining the interval to
+\([0,1])\).
 
@@
 
-# Precaclulate the mean views
-usMean <- mean(usExtract$views)
-
 # Add the engineered feature
@@
                  engagement = ifelse(rtgDisabled, 1, 
-                                     comments / sum(likes, dislikes)))]
-                 # exceedMean = factor(ifelse(views > usMean, 1, 0),
-                 #                     levels = c(0, 1), labels = c(0, 1)))
-
+                                     comments / (likes + dislikes)))]
+engageMode <- prettyNum(density(usExtract$engagement, kernel = 'ep')$x[
+  which.max(density(usExtract$engagement, kernel = 'ep')$y)], digits = 4L)
+engageMean <- prettyNum(mean(usExtract$engagement), digits = 4L)
+engageMax <- prettyNum(max(usExtract$engagement), digits = 4L)
 ```
@@
   geom_density() +
-  scale_x_log10() +
-  ggtitle("Distribution of Engagement Log Scale")
+  ggtitle("Distribution of Engagement")
 
@@
 It's pretty clear from the distribution of `balance` that, at least for videos
-viewed in the US, there is a healthy dose of disagreement!
-
-<!-- ```{r plotExceedMean, fig.height=4, fig.width=8} -->
-<!-- # Melt for exceedMean -->
-<!-- usSetF <- melt(usExtract, measure.vars = 'exceedMean', variable.factor = FALSE, -->
-<!--                variable.name = 'measure', value.name = 'value') -->
-
-<!-- # Plot the exceedMean variable -->
-<!-- ggplot(usSetF[measure == 'exceedMean'], aes(x = value)) + -->
-<!--   geom_bar() + -->
-<!--   ggtitle("Proportion Above and Below Mean") + -->
-<!--   scale_y_continuous(labels = comma) -->
-<!-- ``` -->
-
-<!-- As with many events which follow the Pareto rule, most of the views are  -->
-<!-- attributable to a minority of the videos, as seen by the 0/1 disparity. -->
+viewed in the US, there is a healthy dose of disagreement, as its mode is much
+closer to 0 than 1. Similarly, `engagement` exhibits right-skewed behavior, with
+a mean of `r engageMean` and a mode at `r engageMode`, yet a maximum of
+`r engageMax`.
 
@@
 
-What may be most interesting is that the disparity ratio indicator us a very
-powerful and significant indicator. Conflict seems to be good for Youtube
+What may be most interesting is that the disparity and engagement ratios are
+powerful and significant indicators. Conflict seems to be good for Youtube
 videos. The more a video is argued over, the more views it seems to get!
@@
 
-This approach *appears* similar to that of the common technique of performing a
-standard linear regression on the logs of the observations, but is different. As
-per [@Gelman2006Lta], one approach *"…log transforms observed values, while the*
-*second one log transforms the expected value.…the key difference being the*
-*relation between the predicted value and the variance."*
+The first approach *appears* similar to that of the common technique of
+performing a standard linear regression on the logs of the observations, but it
+is different. As per [@Gelman2006Lta], one approach *"…log transforms observed*
+*values, while the second one log transforms the expected value.…the key*
+*difference being the relation between the predicted value and the variance."*
 
-```{r lm2Train}
+The second approach is the classic Poisson regression.
+
+```{r lm2aTrain}
 set.seed(181)
-lm2 <- train(x = seenX, y = seenY, family = gaussian(link = 'log'),
+lm2a <- train(x = seenX, y = seenY, family = gaussian(link = 'log'),
              method = 'glmStepAIC', direction = 'both', trace = 0,
@@
 
-```{r lm3Train_poi}
-set.seed(9450)
-lm3_test<- train(x = seenX, y = seenY, family = poisson(link='log'),
+```{r lm2b}
+set.seed(181)
+lm2b<- train(x = seenX, y = seenY, family = poisson(link='log'),
             method = 'glmStepAIC', direction = 'both', trace = 0,
@@
 \footnotesize
-```{r poi_coeff}
-kable(summary(lm3_test$finalModel)$coefficients,
-      caption = "Model 3 (Poisson) Linear Regression Output",
+```{r lm2Tables}
+kable(summary(lm2a$finalModel)$coefficients,
+      caption = "Model 3a (Gaussian) Linear Regression Output",
+      digts = 3L, format.args = list(big.mark = ','))
+kable(summary(lm2b$finalModel)$coefficients,
+      caption = "Model 3b (Poisson) Linear Regression Output",
       digts = 3L, format.args = list(big.mark = ','))
@@
 
-\footnotesize
-```{r lm2Table}
-kable(summary(lm2$finalModel)$coefficients,
-      caption = "Model 3 (Gaussian) Linear Regression Output",
-      digts = 3L, format.args = list(big.mark = ','))
-```
-\normalsize
-
-These models have more predictors than the simple Gaussian (G)LM. Moreover,
-the signs are different from the simple LM. The very fact that
+These models have more predictors than the simple Gaussian (G)LM. Moreover, the
+signs are different from the simple LM. The very fact that
 `Nonprofits & Activism` has a positive coefficient should raise concerns. These
@@
 it is negative, which stands in contrast to the second model. Lastly, having a
-higher balance---ratio of likes to dislikes closer to 1---tends to increase
-views. Conflict seems to be good for YouTubers!
+higher `balance`---ratio of likes to dislikes closer to 1--- or a higher
+`engagement` tends to increase views. Conflict seems to be good for YouTubers!
 
@@
 lm1P <- predict(lm1, newdata = hideX)
-lm2P <- predict(lm2, newdata = hideX)
+lm2aP <- predict(lm2a, newdata = hideX)
+lm2bP <- predict(lm2b, newdata = hideX)
 lm3P <- predict(lm3, newdata = hideX)
-lm3_testP <- predict(lm3_test, newdata = hideX)
 # Compare the results
-compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'ElasticNet', 'GLM: Poisson+Log '),
-                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2P, hideY),
-                                 RMSE(lm3P, hideY), RMSE(lm3_testP, hideY)),
+compTable <- data.table(Model = c('LM', 'GLM: Gauss+Log', 'GLM: Poisson+Log',
+                                  'ElasticNet'),
+                        RMSE = c(RMSE(lm1P, hideY), RMSE(lm2aP, hideY),
+                                 RMSE(lm2bP, hideY), RMSE(lm3P, hideY)),
                         R2 = c(R2(lm1P, hideY, formula = 'traditional'),
-                               R2(lm2P, hideY, formula = 'traditional'),
-                               R2(lm3P, hideY, formula = 'traditional'),
-                               R2(lm3_testP, hideY, formula = 'traditional')),
-                        MAE = c(MAE(lm1P, hideY), MAE(lm2P, hideY),
-                                MAE(lm3P, hideY), MAE(lm3_testP, hideY)))
+                               R2(lm2aP, hideY, formula = 'traditional'),
+                               R2(lm2bP, hideY, formula = 'traditional'),
+                               R2(lm3P, hideY, formula = 'traditional')),
+                        MAE = c(MAE(lm1P, hideY), MAE(lm2aP, hideY),
+                                MAE(lm2bP, hideY), MAE(lm3P, hideY)))
 kable(compTable, digits = 3L, format.args = list(big.mark = ','),
@@
   phenomena.
+  * Talking about cars is also good.
   * Having more tags is better than having fewer.
@@
 
-### Variable Importance: Model 2
+### Variable Importance: Model 2a
 
-```{r lm2VI}
+```{r lm2aVI}
 # Model 2: VarImp 
-ggplot(varImp(lm2)) + ggtitle('Variable Importance: Model 2')
+ggplot(varImp(lm2a)) + ggtitle('Variable Importance: Model 2a')
+```
+
+### Variable Importance: Model 2b
+
+```{r lm2bVI}
+# Model 2: VarImp 
+ggplot(varImp(lm2b)) + ggtitle('Variable Importance: Model 2b')
 ```
@@
 ### Variable Importance: Model 3
+
 ```{r lm3VI}
@@
 ```
-```{r lm2Train}
+```{r lm2aTrain}
 ```
-```{r lm2Table}
+```{r lm2bTrain}
+```
+
+```{r lm2Tables}
 ```
@@
 ```
-```{r lm2VI}
+```{r lm2aVI}
 ```
-```{r lm2VI}
+```{r lm2bVI}
+```
+```{r lm3VI}
 ```
